### 2011-08-18 ltx/lit_review.tex
rearranged some stuff

\documentclass{article}

	\usepackage{fullpage} % Include this if you want to cram lots of things on a page

	\usepackage{amsmath} % these are standard macro packages of the American Mathematical Society
	\usepackage{amssymb}
	%\usepackage{stmaryrd}

	\usepackage{epsfig} % if you want figures

%	\usepackage{fancyhdr} % These 4 lines are needed to set up the running  header
%	\fancyhead[LE,RO]{Test Document}
%	\fancyhead[RE,LO]{\thepage}
%	\pagestyle{fancy}

\newcommand{\matlab}[1]
{\centerline{\parbox{.9\textwidth}{\noindent\textsc{\bf MATLAB:} #1}}}

\newcommand{\code}[1]{\texttt{#1}}

\newcommand {\x}{\V{x}}
\newcommand {\y}{\V{y}}
\newcommand {\V}[1]{\mbox{\boldmath$#1$}}

\begin{document}

\title{Literature Review}
\author{Willie}
\maketitle
\mbox{}

\subsection*{Overview}
\vspace{6pt}
The phrase `multiple target tracking' (MTT) has been used to refer to two distinct but related problems: the first is the problem of detecting and tracking the positions of multiple targets of interest in a video, and the second is the problem of associating data points, where each data point is assumed to represent the centroid of some target whose position varies with time. The first is a problem in the field of machine vision and may use data association techniques from the second to accomplish MTT after targets of interest are detected and their positions represented as points.\\
\\
The goal of this project is to apply the generalized Polya urn Dirichlet process mixture model (GPUDPM) to videos containing multiple targets in a way that accomplishes data association without requiring explicit target detection. We hope to allow for a flexible MTT algorithm that can handle videos with different numbers of targets, with diverse backdrops, whose targets display a wide range of characteristics, and where there are time-varying shifts in targets' characteristics, perspectives, or orientations.\\
\\
The following are summaries of papers focusing on MTT as it relates to the association of target centroids, MTT as it relates to machine vision, and the generalized Polya urn Dirichlet process mixture model.

\subsection*{Data Association Related MTT}
\vspace{6pt}
Papers on this topic often involve algorithms that take a set of data, where each point has a time-value and is assumed to represent the centroid position of a target at the given point in time, and aim to return the most probable set of tracks formed by `associating' these data points into sequences.\\
\\
In Bar-Shalom's classic MTT survey paper (Tracking Methods in a Multitarget Environment, 1978), he discusses the common algorithms of the time, and how they attempt to overcome issues caused by false positive and true negative data points. Blackman (Multiple Hypothesis Tracking for Multiple Target Tracking, 2004) provides an overview of multiple hypothesis tracking (MHT), a classic and perhaps most-widely-used MTT algorithm, which is extended by many of the machine-vision algorithms described in the following section. Blom and Bloem (Interacting Multiple Model Joint Probabilistic Data Association avoiding track coalescence, 2002) give a good overview of another classic data association algorithm, the joint probabilistic data association filter (JPDA). Pulford (Taxonomy of multiple target tracking methods, 2005), aiming to provide a modern update of Bar-Shalom's paper, gives a thorough overview of `classic' and `modern' data association algorithms (and gives the details of 35 different MTT algorithms). Hue et al. (Sequential Monte Carlo Methods for Multiple Target Tracking and Data Fusion, 2002) introduce particle filters for MTT. Oh et al. (Markov Chain Monte Carlo Data Association for General Multiple-Target Tracking Problems, 2004) uses an MCMC approach to perform general data association. Fox et. al (Nonparametric Bayesian Methods for Large Scale Multi-Target Tracking, 2006) uses a Dirichlet Process prior to estimate the number of targets in an MTT scheme.

\subsection*{Machine Vision Related MTT}
\vspace{6pt}
Papers on this topic often involve algorithms that have a target detection phase (where the centroids of potential targets are collected, often by looking for specific target features) and a data association phase (where the centroids are sequenced in order to carry out tracking).\\
\\
Qu et al. (Distributed Bayesian Multiple-Target Tracking in Crowded Environments Using Multiple Collaborative Cameras, 2007) term machine vision related MTT `visual multiple-target tracking', and provide a good overview of the topic (the following mirrors this overview). MacCormick and Blake (A Probabilistic Exclusion Principle for Tracking Multiple Objects, 2000) focus on distinguishing between foreground and background objects and propose a probabilistic exclusion principle to help the problem where objects occlude each other (ie where one object overlaps and blocks another). Isard and MacCormick (BraMBLe: A Bayesian Multiple-Blob Tracker, 2001) present a people-tracking algorithm that deduces foreground and background objects probabilistically, without the typical strategy of selecting foreground objects prior to MTT; the authors also talk about ways in which their techniques allow for tracking variable numbers of objects and objects that enter and exit the videos. Zhao and Nevatia (Tracking Multiple Humans in Crowded Environment, 2004) make use of human shape models to better the MTT accuracy when tracking crowds of people. Tao et al. (A Sampling Algorithm for Tracking Multiple Objects) use sampling to perform Bayesian inference and focus on handling occlusion, addition, and deletion of targets. Rasmussen et al. (Probabilistic Data Association Methods for Tracking Complex Visual Objects, 2001) apply a version of the joint probabilistic data association filter to certain shape and texture features they are  able to extract from video targets. Khan et al. (An MCMC-based Particle Filter for Tracking Multiple Interacting Targets, 2004), present an MCMC-based particle filter that also introduces a Markov random field to model the interactions of targets when they are within certain ranges of other targets (applied to videos of ants, arguing that an ant's behavior is highly influenced by nearby ants). Smith et al. (Using Particles to Track Varying Numbers of Interacting People, 2005) present an MCMC particle filter to increase efficiency in the multiple hypothesis tracker, and also incorporate target features such as color into their model. McKenna et al. (Tracking Groups of People, 2000) use color and gradient information to perform background subtraction on videos of people; their model then has a heirarchical structure where groups of occluding people are first tracked, and then later separated to track each person separately, using color information. Yamamoto et al. (Realtime Multiple Object Tracking Based on Optical Flows, 1995) uses optical flows (calculated based on a so called `generalized gradient model' of an image) to carry out multiple target tracking.\\
\\

\subsection*{Generalized Polya Urn Dirichlet Process Mixture Model}
\vspace{6pt}
Caron et al. (Generalized Polya Urn for Time-varying Dirichlet Process Mixtures, 2007) introduce this model. Gasthaus et al. (Dependent Dirichlet Process Spike Sorting, 2009) apply this model to the spike-sorting problem.

\end{document}

### 2011-08-23 ltx/lit_review.tex
attempt to put in correct tex docs for datamining final paper and literature review

@@
 	\usepackage{epsfig} % if you want figures
+
+	\usepackage[margin=1.5in]{geometry}
+
+	\usepackage[numbers]{natbib}
 
@@
 
+\subsection*{Some Terminology}
+\vspace{6pt}
+1. Multiple Target Tracking (MTT) \\
+\\
+a.k.a. Multitarget Tracking, Data Association \\
+\\
+Definition: An algorithm that takes as input a collection of data points, where each point represents the center position of an object at a given time, and returns a partition of these points; each element of the partition is comprised of a sequence of data points and represents a target's position at successive points in time.
+\\
+\\
+\\
+2. Target Detection \\
+\\
+a.k.a. Target Recognition, Target Localization, Target Representation, Target Segmentation, Object Detection \\
+\\
+Definition: The act of deducing whether an image contains a specified target and, if it is present, locating the target's position. \\
+\\
+\\
+3. Visual Multiple Target Tracking (VMTT) \\
+\\
+a.k.a. Video Tracking, Video Multiple Target Tracking, 2D Tracking, Object Tracking, Multi-Object Video Tracking \\
+\\
+Definition: An algorithm that tracks the time-varying positions of multiple targets in a video; it typically involves target detection to find the center position of each target at each point in time, and MTT on the found target centers to accomplish tracking.
+\\
+\\
+
 \subsection*{Overview}
 \vspace{6pt}
-The phrase `multiple target tracking' (MTT) has been used to refer to two distinct but related problems: the first is the problem of detecting and tracking the positions of multiple targets of interest in a video, and the second is the problem of associating data points, where each data point is assumed to represent the centroid of some target whose position varies with time.
-The first is a problem in the field of machine vision and may use data association techniques from the second to accomplish MTT after targets of interest are detected and their positions represented as points.\\
+Visual multiple target tracking (VMTT) involves tracking the positions of multiple targets in a video.
+VMTT algorithms typically carry out two main tasks: the first is target detection, the task of finding the positions of targets within an image, and the second is multiple target tracking (MTT), the task of grouping the positions into sequences, where each sequence represents a target's position at successive points in time.
 \\
-The goal of this project is to apply the generalized Polya urn Dirichlet process mixture model (GPUDPM) to videos containing multiple targets in a way that accomplishes data association without requiring explicit target detection.
-We hope to allow for a flexible MTT algorithm that can handle videos with different numbers of targets, with diverse backdrops, whose targets display a wide range of characteristics, and where there are time-varying shifts in targets' characteristics, perspectives, or orientations.\\
 \\
-The following are summaries of papers focusing on MTT as it relates to the association of target centroids, MTT as it relates to machine vision, and the generalized Polya urn Dirichlet process mixture model.
+MTT algorithms are well studied, and, given the positions of targets in each frame of a video, have been shown to achieve tracking with high accuracy \cite{blackman_2004}.
+General methods for target detection have had less success.
+Many current target detection methods are limited to specific types of targets as they perform detection by catering to some unique visual or structural characteristic of the target.
+These methods are often less robust when applied to targets that change appearance, orientation, or scale, and to scenes with diverse background characteristics.
+Furthermore, even with a target-specific feature to aid detection, there are often problems with image segmentation (decisions as to where one object ends and a neighboring object begins) and with estimating the number of objects in a scene.
+\\
+\\
+The goal of this project is to apply a time dependent Dirichlet process mixture model known as the generalized Polya urn Dirichlet process mixture model (GPUDPM) to videos containing multiple targets in a way that tracks the targets' positions without requiring explicit target detection.
+We hope to allow for a 	single flexible VMTT algorithm that can be applied to scenes with different numbers of targets, with diverse backgrounds, whose targets display a wider range of visual and structural characteristics, and where there are time-varying shifts in targets' characteristics, perspectives, or orientations.\\
 
-\subsection*{Data Association Related MTT}
+\subsection*{MTT}
 \vspace{6pt}
-Papers on this topic often involve algorithms that take a set of data, where each point has a time-value and is assumed to represent the centroid position of a target at the given point in time, and aim to return the most probable set of tracks formed by `associating' these data points into sequences.\\
+At its most basic, MTT can be approached with the Global Nearest Neighbor (GNN) technique \cite{blackman_2004}, where tracks are formed by associating each observation with its nearest neighbor, under the constraint that an observation can be associated with no more than one track.
+Often, GNN is performed until a problem such as a conflicting track assignment occurs, whereupon a more sophisticated technique is used.
+As for these more-complex methods, there are two classic MTT algorithms that form a basis for many modern techniques.
+The first is called multiple hypothesis tracking (MHT) \cite{singer_1974, reid_1979, blackman_2004}, in which (at its most basic), for each observation, a set of potential neighbors are considered for association; the process is then repeated recursively for each potential neighbor, forming a tree-like structure.
+Most branches die along the way due to lack of potential neighbors, but eventually (ideally) only one branch remains, and it is considered most-likely to be the true path.
+MHT is computationally expensive, but has gained popularity in recent times due to increases in computational capabilities \cite{blackman_2004}.
+The second classic MTT algorithm is known as the Joint Probabilistic Data Association (JPDA) method \cite{barshalom_1975, kirubarajan_2004}.
+In contrast to GNN, where the nearest potential neighbor is chosen as the associated position, and MHT, where all potential neighbors are followed until only one remains, JPDA computes a probability that each potential neighbor (that pass an initial validation test) is the next position of the current track, and predicts the next position as a probability-weighted combination of all potential positions.
+In this way, JPDA acts similar to another MTT technique known as Recursive Bayesian Filtering (RBF), where the position of a target at a future step is predicted, and this prediction allows for subsequent points to be associated to a track with higher accuracy.
+Techniques involving RBF often make use of a Kalman or particle filter, and these methods are often incorporated with the other algorithms mentioned in this section.
+Pulford \cite{pulford_2005} gives a thorough overview of `classic' and `modern' data association algorithms (and gives the details of 35 different MTT algorithms).
+Fox et al.
+\cite{fox_2006} uses a Dirichlet Process prior to estimate the number of targets in an MTT scheme.
 \\
-In Bar-Shalom's classic MTT survey paper (Tracking Methods in a Multitarget Environment, 1978), he discusses the common algorithms of the time, and how they attempt to overcome issues caused by false positive and true negative data points.
-Blackman (Multiple Hypothesis Tracking for Multiple Target Tracking, 2004) provides an overview of multiple hypothesis tracking (MHT), a classic and perhaps most-widely-used MTT algorithm, which is extended by many of the machine-vision algorithms described in the following section.
-Blom and Bloem (Interacting Multiple Model Joint Probabilistic Data Association avoiding track coalescence, 2002) give a good overview of another classic data association algorithm, the joint probabilistic data association filter (JPDA).
-Pulford (Taxonomy of multiple target tracking methods, 2005), aiming to provide a modern update of Bar-Shalom's paper, gives a thorough overview of `classic' and `modern' data association algorithms (and gives the details of 35 different MTT algorithms).
-Hue et al.
-(Sequential Monte Carlo Methods for Multiple Target Tracking and Data Fusion, 2002) introduce particle filters for MTT.
-Oh et al.
-(Markov Chain Monte Carlo Data Association for General Multiple-Target Tracking Problems, 2004) uses an MCMC approach to perform general data association.
-Fox et.
-al (Nonparametric Bayesian Methods for Large Scale Multi-Target Tracking, 2006) uses a Dirichlet Process prior to estimate the number of targets in an MTT scheme.
 
-\subsection*{Machine Vision Related MTT}
+\subsection*{VMTT}
 \vspace{6pt}
-Papers on this topic often involve algorithms that have a target detection phase (where the centroids of potential targets are collected, often by looking for specific target features) and a data association phase (where the centroids are sequenced in order to carry out tracking).\\
-\\
 Qu et al.
-(Distributed Bayesian Multiple-Target Tracking in Crowded Environments Using Multiple Collaborative Cameras, 2007) term machine vision related MTT `visual multiple-target tracking', and provide a good overview of the topic (the following mirrors this overview).
-MacCormick and Blake (A Probabilistic Exclusion Principle for Tracking Multiple Objects, 2000) focus on distinguishing between foreground and background objects and propose a probabilistic exclusion principle to help the problem where objects occlude each other (ie where one object overlaps and blocks another).
-Isard and MacCormick (BraMBLe: A Bayesian Multiple-Blob Tracker, 2001) present a people-tracking algorithm that deduces foreground and background objects probabilistically, without the typical strategy of selecting foreground objects prior to MTT; the authors also talk about ways in which their techniques allow for tracking variable numbers of objects and objects that enter and exit the videos.
-Zhao and Nevatia (Tracking Multiple Humans in Crowded Environment, 2004) make use of human shape models to better the MTT accuracy when tracking crowds of people.
+\cite{qu_2007} term computer vision related MTT `visual multiple-target tracking', and give a thorough overview of the topic.
+MacCormick and Blake \cite{maccormick_1999} focus on distinguishing between foreground and background objects and use it to help perform target detection in videos where objects occlude each other (i.e.
+where one object overlaps and blocks another).
+Isard and MacCormick \cite{isard_2001} present a people-tracking algorithm that deduces foreground and background objects probabilistically, without the typical strategy of selecting foreground objects prior to MTT; the authors also talk about ways in which their techniques allow for tracking variable numbers of objects and objects that enter and exit the videos.
+Zhao and Nevatia \cite{tao_2004} make use of human shape models to better the MTT accuracy when tracking crowds of people.
 Tao et al.
-(A Sampling Algorithm for Tracking Multiple Objects) use sampling to perform Bayesian inference and focus on handling occlusion, addition, and deletion of targets.
+\cite{tao_1999} use sampling to perform Bayesian inference and focus on handling occlusion, addition, and deletion of targets.
 Rasmussen et al.
-(Probabilistic Data Association Methods for Tracking Complex Visual Objects, 2001) apply a version of the joint probabilistic data association filter to certain shape and texture features they are  able to extract from video targets.
+\cite{rasmussen_2001} apply a version of the joint probabilistic data association filter to certain shape and texture features they are able to extract from video targets (i.e.
+in the case of conflicting track assignments, successive positions are chosen based on a combination of probabilities of potential positions which are each dependent, in part, upon these shape and texture features).
 Khan et al.
-(An MCMC-based Particle Filter for Tracking Multiple Interacting Targets, 2004), present an MCMC-based particle filter that also introduces a Markov random field to model the interactions of targets when they are within certain ranges of other targets (applied to videos of ants, arguing that an ant's behavior is highly influenced by nearby ants).
+\cite{khan_2005}, present an MCMC-based particle filter that also introduces a Markov random field to model the interactions of targets when they are within certain ranges of other targets (applied to videos of ants, arguing that an ant's behavior is highly influenced by nearby ants).
 Smith et al.
-(Using Particles to Track Varying Numbers of Interacting People, 2005) present an MCMC particle filter to increase efficiency in the multiple hypothesis tracker, and also incorporate target features such as color into their model.
+\cite{smith_2005} aim to increase the efficiency in the multiple hypothesis tracker while incorporating target features such as color into their model.
 McKenna et al.
-(Tracking Groups of People, 2000) use color and gradient information to perform background subtraction on videos of people; their model then has a heirarchical structure where groups of occluding people are first tracked, and then later separated to track each person separately, using color information.
+\cite{mckenna_2000} use color and gradient information to perform background subtraction on videos of people; their model then has a heirarchical structure where groups of occluding people are first tracked, and then later separated to track each person separately, using color information.
 Yamamoto et al.
-(Realtime Multiple Object Tracking Based on Optical Flows, 1995) uses optical flows (calculated based on a so called `generalized gradient model' of an image) to carry out multiple target tracking.\\
-\\
+\cite{yamamoto_1995} uses optical flows (calculated based on a so called `generalized gradient model' of an image, which attempts to capture the speed and direction of movement over areas in the image) to carry out multiple target tracking.\\
 
@@
 \vspace{6pt}
-Caron et al.
-(Generalized Polya Urn for Time-varying Dirichlet Process Mixtures, 2007) introduce this model.
-Gasthaus et al.
-(Dependent Dirichlet Process Spike Sorting, 2009) apply this model to the spike-sorting problem.
+Caron et al. \cite{caron_2007} introduce the generalized Polya urn Dirichlet process mixture model. Gasthaus et al. \cite{gasthaus_2008} apply this model to the spike-sorting problem.
+
+\begin{small}
+\bibliographystyle{plainnat}
+\bibliography{refs_lit_review} 
+\end{small}
+
+%\subsection*{Overview 2}
+%\vspace{6pt}
+%The phrase `multiple target tracking' (MTT) has been used to refer to two distinct but related problems: the first is the problem of detecting and tracking the positions of multiple targets of interest in a video, and the second is the problem of associating data points, where each data point is assumed to represent the centroid of some target whose position varies with time.
+The first is a problem in the field of machine vision and may use data association techniques from the second to accomplish MTT after targets of interest are detected and their positions represented as points.\\
+%\\
+%The goal of this project is to apply the generalized Polya urn Dirichlet process mixture model (GPUDPM) to videos containing multiple targets in a way that accomplishes data association without requiring explicit target detection.
+We hope to allow for a single flexible MTT algorithm that can handle videos with different numbers of targets, with diverse backdrops, whose targets display a wide range of characteristics, and where there are time-varying shifts in targets' characteristics, perspectives, or orientations.\\
+%\\
+%\\
+%The following are summaries of papers focusing on MTT as it relates to the association of target centroids, MTT as it relates to machine vision, and the generalized Polya urn Dirichlet process mixture model.
+
+% Hue et al. \cite{hue, 2002} introduce particle filters for MTT.
+
+%Oh et al. (Markov Chain Monte Carlo Data Association for General Multiple-Target Tracking Problems, 2004) uses an MCMC approach to perform general data association.
+
+% VMTT
+%Papers on this topic often involve algorithms that have a target detection phase (where the centroids of potential targets are collected, often by looking for specific target features) and a data association phase (where the centroids are sequenced in order to carry out tracking).\\
 

### 2011-08-25 ltx/lit_review.tex
might have screwed up earlier commit -- this is the good point

@@
 
+\subsection*{Overall Paper Breakdown}
+\vspace{6pt}
+Papers relating to techniques for the general tracking of multiple objects are split into four broad topics: those related to the mean-shift procedure (a very popular and effective technique in recent times), those related to object tracking with the use of color, those related to foreground modeling and motion extraction, and those related to cluster-based object tracking.
+\\
+\\
+\\
+The Mean-Shift Procedure \\
+a.k.a Kernel Based Object Detection \\
+\\ 
+The mean-shift procedure attempts to provide a robust way for non-rigid objects to be tracked.
+In this case, tracking refers to the task of maintaining the location of an object after its initial position has been specified.
+The mean shift procedure (a derivation and classic overview of which can be found in \cite{fukunaga_1975}and \cite{cheng_1995}) is a simple iterative procedure that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges for each data point.
+Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors.
+It can be shown that, through this procedure, each data point becomes associated with a local point of high density (dependent upon the underlying weight distribution specified by the kernel) which naturally allows for clustering \cite{cheng_1995}.
+The mean shift algorithm has been implemented successfully to allow for a sort of kernel-based object tracking (\cite{comaniciu_2003}, \cite{comaniciu_1999}, \cite{comaniciu_2000}).
+Given the current position of an object at a given frame, the goal of tracking is often to find a nearby position in the next frame that has the most similar distribution over some common set of features.
+Usually this must be done through an exhaustive search, comparing the similarity of distributions at each nearby position with that at the current position.
+However, if a certain type of `isotropic kernel' (mean-shift kernel) known as the Bhattacharyya coefficient is chosen as a similarity metric between the feature distributions at two positions, it creates a smooth function where gradient descent techniques can be used to quickly converge upon an optimal subsequent position without using an exhaustive search.
+Many papers are concerned with applying this kernel-based object tracking scheme to different sets of features or with different kernels.
+Additionally, the scheme has been attempted with an adaptive kernel whose shape, scale, and orientation is influenced by a target being tracked \cite{wang_2004}, with a kernel adjusted by the estimated centroid of a tracked target \cite{mehmood_2009}, and with a heirarchical version of the mean-shift procedure \cite{dementhon_2002}.
+\\
+\\
+\\
+Object Tracking with the Use of Color \\
+\\
+Color has long been used to track non-rigid objects; again, in this case, tracking refers to the task of maintaining the location of an object after its initial position has been specified.
+In general, these schemes operate by modeling an object with some color-based appearance model, and using this model to find the object in subsequent frames \cite{elgammal_2001}.
+One classic approach involves modeling the color of a target using a Gaussian mixture model (GMM), and choosing a subsequent target position by searching a surrounding area for a position that yields the most similar GMM.
+Furthermore, the GMM is often allowed to adapt over time, and slowly change to model changes in lighting or other smooth time-based variations in the target's color \cite{raja_1998}, \cite{mckenna_1999}, \cite{jepson_2003}.
+Additionally, color has been successfully applied as the feature in the mean-shift procedure (kernel-based method) for tracking \cite{comaniciu_2003}, \cite{perez_2002}, \cite{nummiaro_2003}, \cite{lee_2011}.
+\\
+\\
+\\
+Foreground Detection\\
+a.k.a Frame Differencing, Image Differencing, Background Subtraction, Motion Detection\\
+\\
+Foreground detection, frame differencing, background subtraction, and motion detection all refer to slighly different things.
+Frame differencing and background subtraction involve comparing pixel values, or groups of pixel values, from two images, to detect a change surpassing some threshold.
+Often, background subtraction refers to comparing an image containing targets with an image of the background without targets, while frame or image differencing refers to comparing two subsequent images in a video.
+Foreground detection might also refer to a similar comparison technique, but also might be achieved by modeling pixel values.
+For example, sensitive foreground detection has been carried out robustly by modeling each pixel over time in order to infer a probability distribution over pixel values; if the pixel takes on a very unlikely value at a future time, it is marked as a foreground pixel \cite{stauffer_1999}, \cite{elgammal_2000}, \cite{elgammal_2002}.
+These pixel modeling techniques have been adapted to allow for time-dependent pixel value probability distributions, which help account for slighlty changing backgrounds, such as those caused from slight movements of background objects (for example, trees in outdoor scenes).
+Note that these techniques treat each pixel independently; this has a tendency to cause errors in foreground detection, resulting in fragmentation of foreground images (which is problematic as many of these methods intend to blob the foreground objects after detection in order to perform segmentation and locate objects' centroids).
+Post processing has been applied to add information regarding the edges of objects in an attempt to decrease fragmentation \cite{turdu_2007}.
+Similar to foreground detection, motion detection might also refer to a comparison technique or to other ways of extracting movement from an image, such as through techniques involving optical flow (a calculation based on a so called `generalized gradient model' of an image, which attempts to capture the speed and direction of movement over areas in the image) \cite{horn_1981}, \cite{bobick_2001}.
+In \cite{black_2000}, a motion detection and object tracking method is developed based on the relative motion of moving objects' boundaries, which are again found through analysis of optical flow.\\
+\\
+\\
+Clustering Based Object Tracking \\
+a.k.a. Most Similar Things to Our Project \\
+\\
+Clustering based object tracking and detection schemes typically attempt to either detect objects or simultaneously detect and track objects using minimal, easily extracted, or general information about a scene and the targets within it.
+These techniques often start with foreground (e.g.
+movement) or color data from each frame in a video, and attempt to automatically perform segmentation and tracking with temporal clustering (where clusters are somehow present at or shared between multiple frames).
+This differs from blob-based tracking approaches, where a pixel is assigned to the foreground or to a specific, segmented blob (each of which represents a target) such as in \cite{stauffer_1999}.
+In \cite{pece_2002}, foreground positions are found via frame differencing and modelled with a sequence of GMMs\\
+\cite{beleznai_2006}
+\cite{dickinson_2005}
+\cite{schiele_2006}
+  \\
+\\
+
 \subsection*{Some Terminology}
@@
 Definition: An algorithm that tracks the time-varying positions of multiple targets in a video; it typically involves target detection to find the center position of each target at each point in time, and MTT on the found target centers to accomplish tracking.
+\\
+\\
+4. Tracking \\
+\\
+a.k.a most terms in 3.
+\\
+Definition: Given the current position of a target, tracking algorithms find a likely subsequent position of a target (note that this is distinct from the VMTT definition above, as it does not involve target detection; however, terms describing VMTT and tracking are ofen used to refer to both).
+\\
+\\
+5. Background Subtraction\\
+\\
+a.k.a frame differencing, image differencing, motion detection \\
+\\
+Definition: The terms above all involve comparing pixel values, or groups of pixel values, from two images, to detect a change surpassing some threshold.
+Often 'background subtraction' refers to comparing an image containing targets with an image of the background without targets, while 'frame/image differencing' refers to comparing two subsequent images in a video.
+\\
 \\

### 2011-08-30 ltx/lit_review.tex
fixed small errors in refs_lit_review.bib
new literature review, almost complete and extensive

--- 2011-08-23 ltx/lit_review.tex
@@
 
-	\usepackage[numbers]{natbib}
-
-%	\usepackage{fancyhdr} % These 4 lines are needed to set up the running  header
-%	\fancyhead[LE,RO]{Test Document}
-%	\fancyhead[RE,LO]{\thepage}
-%	\pagestyle{fancy}
+	% \usepackage[numbers]{natbib}
+	\usepackage[square]{natbib}
 
@@
 
+\newcommand{\tab}{\hspace*{2em}}
+
 \begin{document}
@@
 
-\subsection*{Some Terminology}
+\subsection*{Outline and Terminology}
 \vspace{6pt}
-1. Multiple Target Tracking (MTT) \\
-\\
-a.k.a. Multitarget Tracking, Data Association \\
-\\
-Definition: An algorithm that takes as input a collection of data points, where each point represents the center position of an object at a given time, and returns a partition of these points; each element of the partition is comprised of a sequence of data points and represents a target's position at successive points in time.
-\\
-\\
-\\
-2. Target Detection \\
-\\
-a.k.a. Target Recognition, Target Localization, Target Representation, Target Segmentation, Object Detection \\
-\\
-Definition: The act of deducing whether an image contains a specified target and, if it is present, locating the target's position. \\
-\\
-\\
-3. Visual Multiple Target Tracking (VMTT) \\
-\\
-a.k.a. Video Tracking, Video Multiple Target Tracking, 2D Tracking, Object Tracking, Multi-Object Video Tracking \\
-\\
-Definition: An algorithm that tracks the time-varying positions of multiple targets in a video; it typically involves target detection to find the center position of each target at each point in time, and MTT on the found target centers to accomplish tracking.
-\\
-\\
+\begin{enumerate}
+\item Multiple Target Tracking (MTT)
+	\begin{enumerate}
+	\item a.k.a. Point MTT, Multitarget Tracking, Data Association, Point Matching, Plot Association
+	\item Definition: (will include later)
+	\item Multiple Hypothesis Tracking (MHT), Joint Probabilistic Data Association Filter (JPDAF), and their variants and spin-offs
+		\begin{enumerate}
+		\item These algorithms may be used to track targets in videos if they can be represented as points.
+		\end{enumerate}
+	\end{enumerate}
+\item Visual Object Tracking (VOT)
+	\begin{enumerate}
+	\item a.k.a. Video Tracking, Target Tracking, Visual Single Object Tracking, 2D Tracking, `Tracking of non-rigid objects'
+	\item Definition: (will include later)
+	\item Recursive Bayesian filtering
+	\item Mean-shift methods
+	\item Color tracking % (particularly, GMMs and `adaptive' stuff)
+	\item These can be extended to multiple targets if detection / initialization is given (sometimes involve more sophisticated techniques when multiple targets introduced).
+	\end{enumerate}
+\item Foreground Extraction and Object Detection
+	\begin{enumerate}
+	\item a.k.a. Foreground segmentation, background modeling, frame differencing, background subtraction, motion detection, and \\
+	a.k.a. Target localization, target representation, target segmentation 
+	\item Definition: (will include later)
+	\item Foreground / Background Modeling
+		\begin{enumerate}
+		\item GMM for background / foreground modeling (many studies)
+		\end{enumerate}
+	\item Motion detection, frame differencing, background subtraction
+	\item Segmentation
+	\end{enumerate}
+\item Multiple Visual Object Tracking and Detection
+	\begin{enumerate}
+	\item a.k.a Multiple Target Tracking and Localization
+	\item Definition: (will include later) 
+	\item blob tracking
+		\begin{enumerate}
+		\item foreground detection (motion, thresholding, subtraction, color), segmentation (pixel assignment), (sometimes) centroid placement (localization?), and (sometimes) point MTT.
+		\end{enumerate}
+	\item clustering methods for automated segmentation and tracking
+		\begin{enumerate}
+		% \item Scene modeling (often uses Gaussians perpetuated over time, though each one isn't meant to represent a single target, i.e. doesn't intend to perform segmentation).
+		\item Pece cluster tracker on motion (time-varying GMM with additions, applied to frame differencing)
+		\item Mean-shift time-dependent clustering (1 paper)
+		%  \item Other papers mentioned in Pece
+		% \item Clustering feature points
+		% \item gpudpm4mtt
+		\end{enumerate}
+	\end{enumerate}
+\end{enumerate}
 
-\subsection*{Overview}
+\subsection*{Related Work}
 \vspace{6pt}
-Visual multiple target tracking (VMTT) involves tracking the positions of multiple targets in a video.
-VMTT algorithms typically carry out two main tasks: the first is target detection, the task of finding the positions of targets within an image, and the second is multiple target tracking (MTT), the task of grouping the positions into sequences, where each sequence represents a target's position at successive points in time.
-\\
-\\
-MTT algorithms are well studied, and, given the positions of targets in each frame of a video, have been shown to achieve tracking with high accuracy \cite{blackman_2004}.
-General methods for target detection have had less success.
-Many current target detection methods are limited to specific types of targets as they perform detection by catering to some unique visual or structural characteristic of the target.
-These methods are often less robust when applied to targets that change appearance, orientation, or scale, and to scenes with diverse background characteristics.
-Furthermore, even with a target-specific feature to aid detection, there are often problems with image segmentation (decisions as to where one object ends and a neighboring object begins) and with estimating the number of objects in a scene.
-\\
-\\
-The goal of this project is to apply a time dependent Dirichlet process mixture model known as the generalized Polya urn Dirichlet process mixture model (GPUDPM) to videos containing multiple targets in a way that tracks the targets' positions without requiring explicit target detection.
-We hope to allow for a 	single flexible VMTT algorithm that can be applied to scenes with different numbers of targets, with diverse backgrounds, whose targets display a wider range of visual and structural characteristics, and where there are time-varying shifts in targets' characteristics, perspectives, or orientations.\\
+The follow sections provide details on work related to the automated detection and tracking of objects. The following numbered sections correspond to the numbers in the outline above.
 
-\subsection*{MTT}
-\vspace{6pt}
-At its most basic, MTT can be approached with the Global Nearest Neighbor (GNN) technique \cite{blackman_2004}, where tracks are formed by associating each observation with its nearest neighbor, under the constraint that an observation can be associated with no more than one track.
+\subsection*{1. Multiple Target Tracking (MTT)}
+
+\subsubsection*{(a)-(c)}
+At its most basic, MTT can be approached with the Global Nearest Neighbor (GNN) technique \citep{blackman_2004}, where tracks are formed by associating each observation with its nearest neighbor, under the constraint that an observation can be associated with no more than one track.
 Often, GNN is performed until a problem such as a conflicting track assignment occurs, whereupon a more sophisticated technique is used.
 As for these more-complex methods, there are two classic MTT algorithms that form a basis for many modern techniques.
-The first is called multiple hypothesis tracking (MHT) \cite{singer_1974, reid_1979, blackman_2004}, in which (at its most basic), for each observation, a set of potential neighbors are considered for association; the process is then repeated recursively for each potential neighbor, forming a tree-like structure.
-Most branches die along the way due to lack of potential neighbors, but eventually (ideally) only one branch remains, and it is considered most-likely to be the true path.
-MHT is computationally expensive, but has gained popularity in recent times due to increases in computational capabilities \cite{blackman_2004}.
-The second classic MTT algorithm is known as the Joint Probabilistic Data Association (JPDA) method \cite{barshalom_1975, kirubarajan_2004}.
+The first is called multiple hypothesis tracking (MHT) \citep{singer_1974, reid_1979, blackman_2004}, in which, for each observation, a set of neighbors is collected as potentials for association; the process is then repeated recursively for each potential neighbor, forming a tree-like structure of possible paths.
+Once the possible branches for a given path are enumerated, one branch is considered most-likely and is chosen as the true path (where branches may be favored if, for example, they are longer or adhere to some expected behavior of the target).
+MHT is computationally expensive, but has gained popularity due to increases in computational capabilities \citep{blackman_2004}.
+The second classic MTT algorithm is known as the Joint Probabilistic Data Association (JPDA) method \citep{barshalom_1975, kirubarajan_2004}.
 In contrast to GNN, where the nearest potential neighbor is chosen as the associated position, and MHT, where all potential neighbors are followed until only one remains, JPDA computes a probability that each potential neighbor (that pass an initial validation test) is the next position of the current track, and predicts the next position as a probability-weighted combination of all potential positions.
@@
 Techniques involving RBF often make use of a Kalman or particle filter, and these methods are often incorporated with the other algorithms mentioned in this section.
-Pulford \cite{pulford_2005} gives a thorough overview of `classic' and `modern' data association algorithms (and gives the details of 35 different MTT algorithms).
-Fox et al.
-\cite{fox_2006} uses a Dirichlet Process prior to estimate the number of targets in an MTT scheme.
-\\
+Pulford gives a thorough overview of `classic' and `modern' data association algorithms (and gives the details of 35 different MTT algorithms) \citep{pulford_2005}.
+Taking a different approach to estimating the number targets, Fox et al.
+uses a Dirichlet Process prior to estimate the number of targets in an MTT scheme \citep{fox_2006}.
 
-\subsection*{VMTT}
-\vspace{6pt}
-Qu et al.
-\cite{qu_2007} term computer vision related MTT `visual multiple-target tracking', and give a thorough overview of the topic.
-MacCormick and Blake \cite{maccormick_1999} focus on distinguishing between foreground and background objects and use it to help perform target detection in videos where objects occlude each other (i.e.
-where one object overlaps and blocks another).
-Isard and MacCormick \cite{isard_2001} present a people-tracking algorithm that deduces foreground and background objects probabilistically, without the typical strategy of selecting foreground objects prior to MTT; the authors also talk about ways in which their techniques allow for tracking variable numbers of objects and objects that enter and exit the videos.
-Zhao and Nevatia \cite{tao_2004} make use of human shape models to better the MTT accuracy when tracking crowds of people.
-Tao et al.
-\cite{tao_1999} use sampling to perform Bayesian inference and focus on handling occlusion, addition, and deletion of targets.
-Rasmussen et al.
-\cite{rasmussen_2001} apply a version of the joint probabilistic data association filter to certain shape and texture features they are able to extract from video targets (i.e.
-in the case of conflicting track assignments, successive positions are chosen based on a combination of probabilities of potential positions which are each dependent, in part, upon these shape and texture features).
-Khan et al.
-\cite{khan_2005}, present an MCMC-based particle filter that also introduces a Markov random field to model the interactions of targets when they are within certain ranges of other targets (applied to videos of ants, arguing that an ant's behavior is highly influenced by nearby ants).
-Smith et al.
-\cite{smith_2005} aim to increase the efficiency in the multiple hypothesis tracker while incorporating target features such as color into their model.
-McKenna et al.
-\cite{mckenna_2000} use color and gradient information to perform background subtraction on videos of people; their model then has a heirarchical structure where groups of occluding people are first tracked, and then later separated to track each person separately, using color information.
-Yamamoto et al.
-\cite{yamamoto_1995} uses optical flows (calculated based on a so called `generalized gradient model' of an image, which attempts to capture the speed and direction of movement over areas in the image) to carry out multiple target tracking.\\
+\subsection*{2. Visual Object Tracking (VOT)}
 
-\subsection*{Generalized Polya Urn Dirichlet Process Mixture Model}
-\vspace{6pt}
-Caron et al. \cite{caron_2007} introduce the generalized Polya urn Dirichlet process mixture model. Gasthaus et al. \cite{gasthaus_2008} apply this model to the spike-sorting problem.
+\subsubsection*{(a)-(b)}
+Given the position of a target in the current frame of a video, VOT is the task of predicting the target's position in the following frame.
+VOT is not concerned with detecting new targets, but instead with maintaining the location of an object after its initial position has been specified.
+VOT is also sometimes used to refer to the task of maintaining the shape or size of a target over time, in addition to its position.
+Most VOT schemes operate by considering the region that represents a target in a given frame of a video, and looking in the vicinity of this region in a subsequent frame for the most similar region.
+Areas of research are primarily concerned with ways to represent the appearance of targets, predict future positions of targets, accurately or quickly find similar regions in successive frames, and handle troublesome targets (for example, targets which may appear, dissapear, or have time-varying characteristics) \citep{yilmaz_2006}.
+
+\subsubsection*{(c)}
+Recursive Bayesian filtering involves the task of estimating the position of a target from noisy measurements.
+(This section is not finished.
+Need to discuss numerous methods which use particle filtering techniques for object tracking.
+Are these techniques even very similar / under the same umbrella?)
+
+\subsubsection*{(d)}
+The mean-shift procedure (also known as `kernel-based object tracking') attempts to provide a robust way for non-rigid objects to be tracked.
+This method is beneficial because it optimizes the search for the `next' position of an object (i.e.
+the search for the region in a frame which is most similar to a target region in the previous frame).
+This procedure--a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges for each data point.
+Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors.
+It can be shown that, through this procedure, each data point becomes associated with a local point of high density (dependent upon the underlying weight distribution specified by the kernel) which naturally allows for clustering \citep{cheng_1995}.
+The mean shift algorithm has been implemented successfully to allow for a sort of kernel-based object tracking \citep{comaniciu_2003, comaniciu_1999, comaniciu_2000}.
+Given the current position of an object at a given frame, the goal of tracking is often to find a nearby position in the next frame that has the most similar distribution over some common set of features.
+Usually this must be done through an exhaustive search, comparing the similarity of distributions at each nearby position with that at the current position.
+However, if a certain type of `isotropic kernel' (mean-shift kernel) known as the Bhattacharyya coefficient is chosen as a similarity metric between the feature distributions at two positions, it creates a smooth function where gradient descent techniques can be used to quickly converge upon an optimal subsequent position without using an exhaustive search.
+Many papers are concerned with applying this kernel-based object tracking scheme to different sets of features or with different kernels.
+Additionally, the scheme has been attempted with an adaptive kernel whose shape, scale, and orientation is influenced by a target being tracked \citep{wang_2004}, with a kernel adjusted by the estimated centroid of a tracked target \citep{mehmood_2009}, and with a heirarchical version of the mean-shift procedure \citep{dementhon_2002}.
+
+\subsubsection*{(e)}
+Color has long been used to track non-rigid objects; again, in this case, tracking refers to the task of maintaining the location of an object after its initial position has been specified.
+In general, these schemes operate by modeling an object with some color-based appearance model, and using this model to find the object in subsequent frames.
+One classic approach involves modeling the color of a target (in particular, a distribution over the hue-saturation space of a target region) with a Gaussian mixture model (GMM), and choosing subsequent target positions by searching surrounding areas for regions that yield similar GMMs.
+Furthermore, the GMM is often allowed to adapt over time, and slowly change to model changes in lighting or other smooth time-based variations in the target's color \citep{raja_1998, mckenna_1999, jepson_2003}.
+Others have attempted to abstract this work with a non-parametric color modeling approach based on kernel density estimation, which does not assume a specific underlying distribution (such as the Gaussian mixture in the previous case) and instead converges to reasonable distribution that depends on the data \citep{elgammal_2001}.
+Additionally, color has been successfully applied as the feature in the mean-shift procedure (kernel-based method) for tracking \citep{comaniciu_2003, perez_2002, nummiaro_2003, lee_2011}.
+
+\subsubsection*{(f)}
+VOT schemes are easily extendable to tracking multiple targets; at the most basic level, this involves initializing multiple targets and running an instance of (single) VOT for each, either simultaneously or in succession \citep{perez_2002}.
+Furthermore, if performed simultaneously, the tracking of each target can be made dependent upon characteristics of other targets (such as their proximity) to resolve errors and improve tracking (such as those caused by the incorrect merging of two targets) \citep{khan_2004, vermaak_2003} .
+
+\subsection*{3. Foreground Extraction and Object Detection}
+
+\subsubsection*{(a)-(b)}
+The following methods are not concerned with tracking, but instead with the detection of and differentiation between individual objects in a given frame (or over a sequence of frames) of a video.
+Usually, objects of interest for tracking are in the foreground of a scene; as such many of the following techniques involve attempts to separate the foreground of a scene from the background.
+Another issue is that of segmentation, or deciding which pixels are associated with which object; this is also what determines the number of objects present in a scene.
+
+\subsubsection*{(c)}
+Foreground extraction, foreground segmentation, foreground detection, and background modeling all refer to the task of deducing which pixels represent foreground objects (i.e.
+objects likely to be desired for tracking), and which represent background objects.
+Foreground extraction, though possible to carry out by following certain heuristics involving temporal changes in pixel value characteristics (see section (d)), is often achieved, instead, using model-oriented techniques.
+For example, foreground extraction has been robustly carried out by modeling each pixel over time in order to infer a probability distribution over pixel values; if the pixel takes on a very unlikely value at a future time, it is marked as a foreground pixel at that point \citep{stauffer_1999, elgammal_2000, elgammal_2002}.
+These pixel modeling techniques have been adapted to allow for time-dependent probability distributions over pixel values, which help account for slighlty changing backgrounds, such as those caused from brief movements of background objects (for example, trees in outdoor scenes).
+Note that these techniques treat each pixel independently; this has a tendency to cause errors in foreground detection, resulting in fragmentation of foreground images (which is problematic as many of these methods intend to blob the foreground objects after detection in order to perform segmentation and locate objects' centroids).
+Post processing has been applied to add information regarding the edges of objects in an attempt to decrease fragmentation \citep{turdu_2007}.
+Others have attempted to achieve better foreground modeling in cases where the foreground and background exhibit similar color distributions (the so called `color similarity problem') by, for example, incorporating models of the foreground into a typical background model (such as the one described above) \citep{gallego_2009, mchugh_2009}.
+
+\subsubsection*{(d)}
+Frame differencing and background subtraction involve comparing pixel values, or groups of pixel values, between two images, to detect a change surpassing some threshold.
+The goal of these techniques is to find locations in a scene that display motion, and to deem these moving areas the foreground of a scene.
+Often, background subtraction refers to comparing an image containing targets with an image of the background without targets or with some model of the background that is learned as the video progresses (in these cases, objects that do not move throughout a scene will be treated as the background), while frame or image differencing refers to comparing two subsequent images in a video.
+Frame differencing has been used as the sole data extraction method for object tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking schemes \citep{perez_2002}.
+In addition to the two techniques mentioned above, there exists other ways of extracting movement from an image, such as through techniques involving optical flow (a calculation based on a so called `generalized gradient model' of an image, which attempts to capture the speed and direction of movement over areas in the image) \citep{horn_1981, bobick_2001}.
+In \citep{black_2000}, a motion detection and object tracking method is developed based on the relative motion of moving objects' boundaries, which are again found through analysis of optical flow.
+
+\subsubsection*{(e)}
+Target localization, target representation, and target segmentation all relate to the task of finding the position of objects in a scene.
+In the context of the foreground extraction described previously, the process of segmenting foreground pixels into distinct foreground objects allows for object localization.
+Segmentation can be very simple: the watershed transform, mean-shift procedure, and other basic clustering algorithms have been used directly on foreground extracted pixels to accomplish segmentation \citep{yang_2006}.
+
+\subsection*{4. Multiple Visual Object Tracking and Detection}
+
+\subsubsection*{(a)-(b)}
+There are certain methods which attempt to combine object detection and VOT in order to provide a start-to-finish technique for temporal object detection (detecting a given object throughout its time in a scene).
+Some methods involve foreground extraction and segmentation as initialization for VOT, which results in the tracking of all objects present at the beginning of a scene, while others attempt more sophisticated methods in order to handle scenes where objects enter and exit throughout.
+
+\subsubsection*{(c)}
+Blob tracking often refers to the task of foreground extraction (in blob trackers this is often done by extracting pixels whose value is above some threshold, which only works for videos where foreground objects have pixel values that are very dissimilar to those of background objects), segmentation of foreground pixels, and MTT, where each `blob' created during segmentation is treated as a point particle.
+More sophisticated blob tracking may involve statistical appearance models for robust foreground extraction, such as those described in 3.(c), which allow for background subtraction in cases where the foreground and background pixels have similar colors or where the characteristics of the foreground targets show strong temporal dependence \citep{stauffer_1999, wren_1997, kjeldsen_1996}.
+Blob trackers have even incorporated the ability to estimate the correct number of targets in a scene and introduce recursive Bayesian filtering for improved VOT \citep{isard_2001}.
+
+\subsubsection*{(d)}
+Clustering based object tracking and detection schemes typically attempt to either detect objects or simultaneously detect and track objects using minimal, easily extracted, or general information about a scene and the targets within it.
+These techniques often start with foreground (e.g.
+movement) pixels, color data from each frame in a video, or so called `feature points' (described in the following section) and attempt to automatically perform segmentation and tracking with temporal clustering (where clusters are somehow present at or shared between multiple frames).
+This differs from blob-based tracking approaches, where a pixel is assigned to the foreground or to a specific, segmented blob (each of which represents a target) at each frame.
+
+\paragraph*{i}
+In \citep{pece_2002}, foreground positions are found via frame differencing and modeled with a sequence of GMMs.
+Pece provides criteria for eliminating, merging, and estimating the correct number of clusters.
+Additionally, he uses simple heuristics for the initialization of clusters.
+Tracking is carried out on a dataset consisting of varied objects, and this method is able to track the various types of objects, though fails when objects become too close in space.
+
+\paragraph*{ii}
+The mean-shift procedure was used by \cite{beleznai_2006} to perform clustering on foreground pixels extracted via frame-differencing.
+The authors promote this technique over that in \citep{pece_2002}, and say that it has a better ability to keep nearby density maxima (in the difference pixels data) separate.
+
+% \paragraph*{ii}
+
+% \cite{dickinson_2005}
+% \cite{schiele_2006}
+% Another method of detecting objects in the foreground involves extraction of objects' `feature points'. These could be any sort of characteristic feature able to be extracted
+
+% \subsection*{Generalized Polya Urn Dirichlet Process Mixture Model}
+% \vspace{6pt}
+% Caron et al. \cite{caron_2007} introduce the generalized Polya urn Dirichlet process mixture model. Gasthaus et al. \cite{gasthaus_2008} apply this model to the spike-sorting problem.
 

### 2011-09-01 ltx/model.tex
added the beginning of the write up about the model

\documentclass{article}

	\usepackage{fullpage} % Include this if you want to cram lots of things on a page

	\usepackage{amsmath} % these are standard macro packages of the American Mathematical Society
	\usepackage{amssymb}
	%\usepackage{stmaryrd}

	\usepackage{epsfig} % if you want figures

	\usepackage[margin=1.5in]{geometry}

	% \usepackage[numbers]{natbib}
	\usepackage[square]{natbib}

	\usepackage{amsmath, epsfig}
	\usepackage{amsfonts}

	\font\nipstenhv  = phvb at 8pt

\newcommand{\matlab}[1]
{\centerline{\parbox{.9\textwidth}{\noindent\textsc{\bf MATLAB:} #1}}}

\newcommand{\code}[1]{\texttt{#1}}

\newcommand {\x}{\V{x}}
\newcommand {\y}{\V{y}}
\newcommand {\V}[1]{\mbox{\boldmath$#1$}}

\newcommand{\tab}{\hspace*{2em}}

\begin{document}

\title{Model}
\author{Willie}
\maketitle
\mbox{}

\subsection*{The Data}
Data are extracted from stationary, single-camera videos that contain moving objects. A moving object produces a trail of `motion points', which are data representing groups of pixels that exhibit sufficient change between consecutive frames. The $D$-dimensional position of each motion point (where $D=2$ for a typical video) is recorded, along with a $V$-dimensional vector representing a distribution over the motion point's pixel color values. Hence, the data consists of motion point observations $x = \{ x^{pos_{1}}, \cdots, x^{pos_{D}}, x^{col_{1}}, \cdots, x^{col_{V}} \}$.

\subsection*{Likelihood}
For a given object $k$ at video frame $t$, the set of positions of the object's motion points are modeled with a product (over each position dimension) of one-dimensional Gaussians
\begin{equation}
P(x_{t_{i}}^{pos} | \mu_{t, k}, \lambda_{t, k}) = \prod_{d=1}^{D} \mathcal{N}(x_{t_{i}}^{pos_{d}} | \mu_{t, k}^{d}, \lambda_{t, k}^{d})
\end{equation}
where $x_{t_{i}}^{pos} = \{ x_{t_{i}}^{pos_{1}}, \cdots, x_{t_{i}}^{pos_{D}} \}$,  $\mu_{t, k}^{d}$ is the mean of the $k^{th}$ Gaussian at time $t$ for dimension $d$, and $\lambda_{t, k}^{d}$ is the precision of the $k^{th}$ Gaussian at time $t$ for dimension $d$.  Likewise, for a given object $k$ at video frame $t$, the set of color vectors of the object's motion points are modeled with a multinomial
\begin{equation}
P(x_{t_{i}}^{col} | m_{t, k}) = Mult(x_{t_{i}}^{col} | m_{t, k})
\end{equation}
where $x_{t_{i}}^{col} = \{ x_{t_{i}}^{col_{1}}, \cdots, x_{t_{i}}^{col_{V}} \}$, $m_{t,k} = \{ m_{t,k}^{1}, \cdots, m_{t,k}^{V} \}$, $\sum_{v=1}^{V} m_{t, k}^{v} = 1$, and $m_{t,k}^{v}>0 \hspace{8pt} \forall v \in \{ 1, \cdots, V \}$. The likelihood for a motion point observation is thus
\begin{equation}
P(x_{t_{i}} | \mu_{t, k}, \lambda_{t, k}, m_{t, k}) =  Mult(x_{t_{i}}^{col} | m_{t, k})   \prod_{d=1}^{D}  \mathcal{N}(x_{t_{i}}^{pos_{d}} | \mu_{t, k}^{d}, \lambda_{t, k}^{d}) 
\end{equation}

\subsection*{Base Distribution $\mathbb{G}_{0}$}
$\mathbb{G}_{0}$ is the base distribution of the underlying time-dependent Dirichlet process mixture; it also serves as a prior distribution for the parameters present in the likelihood. We make use of conjugate priors in the base distribution to allow for more efficient computation. In particular, for an object $k$ at time $t$, we let the prior distribution over the parameters of the motion point position distribution, $\mu_{t, k}$ and $\lambda_{t, k}$, be 
\begin{equation}
\mathbb{G}_{0}^{pos}(\mu_{t, k}, \lambda_{t, k} | \mu_{0}, n_{0}, a, b) = \prod_{d=1}^{D}[\mathcal{N}(\mu_{t, k}^{d} | \mu_{0}, n_{0} \lambda_{t, k}^{d})   Ga(\lambda_{t, k}^{d} | a, b)]
\end{equation}
where $\mu_{0}, n_{0}, a,$ and $b$ are parameters of the base distribution. Furthermore, for object $k$ at time $t$, we let the prior distribuion over the parameters of the motion point color vector distribution be
\begin{equation}
\mathbb{G}_{0}^{col}(m_{t, k} | q) = Dir(m_{t, k} | q)
\end{equation}
where $q = \{ q^{1}, \cdots, q^{V} \}$ is a parameter of the base distribution, where $q^{v}>0 \hspace{8pt} \forall v \in \{ 1, \cdots, V \}$. The base distribution is thus 
\begin{equation}
\mathbb{G}_{0}(\mu_{t, k}, \lambda_{t, k}, m_{t, k} | \mu_{0}, n_{0}, a, b, q) = Dir(m_{t, k} | q)  \prod_{d=1}^{D}[\mathcal{N}(\mu_{t, k}^{d} | \mu_{0}, n_{0} \lambda_{t, k}^{d})   Ga(\lambda_{t, k}^{d} | a, b)]
\end{equation}

\subsection*{Transition Kernels}
Let $\phi_{t,k}^{pos} = \{ \mu_{t, k}, \lambda{t, k} \}$ and $\phi_{t,k}^{col} = \{ m_{t, k} \}$.  We define two transition kernels, $P(\phi_{t, k}^{pos} | \phi_{t-1, k}^{pos})$ and $P(\phi_{t, k}^{col} | \phi_{t-1, k}^{col})$, which dictate, respectively, the random walk of the motion point position parameters and the motion point color vector parameters over time. The transition kernels must be chosen so that their invariation distributions are $\mathbb{G}_{0}$, i.e. such that the following hold:
\begin{equation}
\int \mathbb{G}_{0}^{pos}(\phi_{t-1, k}^{pos}) P(\phi_{t, k}^{pos} | \phi_{t-1, k}^{pos}) d\phi_{t-1, k}^{pos}  =  \mathbb{G}_{0}^{pos}(\phi_{t, k}^{pos})
\end{equation}
\begin{equation}
\int \mathbb{G}_{0}^{col}(\phi_{t-1, k}^{col}) P(\phi_{t, k}^{col} | \phi_{t-1, k}^{col}) d\phi_{t-1, k}^{col}  =  \mathbb{G}_{0}^{col}(\phi_{t, k}^{col})
\end{equation}

\end{document}

### 2011-09-06 ltx/paper_draft.tex
first draft of the paper. a lot of parts going well, a lot of parts still need work

\documentclass{article}

% \usepackage{fullpage} % Include this if you want to cram lots of things on a page 
% \usepackage{amsmath} % these are standard macro packages of the American Mathematical Society
% \usepackage{amssymb}
% \usepackage{epsfig} % if you want figures
% \usepackage[margin=1.5in]{geometry}
% \usepackage[numbers]{natbib}
% \usepackage[square]{natbib}
\newcommand{\matlab}[1]
% {\centerline{\parbox{.9\textwidth}{\noindent\textsc{\bf MATLAB:} #1}}}
% \newcommand{\code}[1]{\texttt{#1}}
% \newcommand {\x}{\V{x}}
% \newcommand {\y}{\V{y}}
% \newcommand {\V}[1]{\mbox{\boldmath$#1$}}
% \newcommand{\tab}{\hspace*{2em}}

\usepackage[square,numbers]{natbib}
\usepackage{amsmath, epsfig}
\usepackage{amsfonts}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{easybmat}
\usepackage{footmisc}
\usepackage{subfig}
\renewcommand\algorithmiccomment[1]{// \textit{#1}}
%
\newcommand{\ignore}[1]{}
\newcommand{\comment}[1]{}
\DeclareMathOperator*{\argmax}{arg\,max}

% need to include style sheet for journal

\begin{document}

\title{Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures}
\author{Willie Neiswanger and Frank Wood}
\maketitle
\mbox{}

\section{Introduction}

% \subsection{Topic Overview and Terminology}

This work is related to the study of algorithms that perform automated detection and tracking of arbitrary objects in videos; we define such algorithms as those with the ability to decide which spatiotemporal regions of a video constitute objects (extraction), find the positions and/or shapes of distinct objects (localization), and maintain the positions of the detected objects over time (tracking). Robust algorithms which can carry out these tasks are applicable to problems in the fields of surveillance, autonomous systems and robotics, and video summarization and compression. Approaches to these tasks have been proposed under different headings and in a variety of contexts. This section serves to introduce relevant terminology and topics of research.

The task of extraction is closely related to the topics of foregound segmentation, background modeling, motion detection, and appearance feature extraction. In general, these methods aim to extract features from regions of a video frame that are considered objects or are distinguished in some way from the background of a scene. Areas of research in this field include attempts to handle situations where the background displays motion or other dynamic characteristics, reduce noise caused by the false extraction of backgrounds, handle abberant image effects present in some videos, and operate accurately on scenes involving illumination changes.

The task of localization is closely related to the topics of object detection, position-finding, target segmentation, contour-detection, template matching, and recognition. These methods are mainly concerned with finding the positions and/or shapes of objects in a given video frame, or with discerning between objects similar in position or appearance. There has been a focus in this field on developing techniques for representing an object's appearance, both to allow for localization of objects on a per-frame basis and to allow for consistent localization of a given object over time. Research into the unsupervised detection of arbitrary objects (which we define as methods that accomplish both extraction and localization) has included work involving image segmentation, parsing, and pixel clustering. These topics usually relate to the automated partitioning of images into meaningful regions. Note that the popular phrase `object detection' is also typically used to refer to procedures that carry out extraction and localization; however, it is often reserved to denote methods that are tailored to find the position of specified objects or object types (as opposed to arbitrary object detection), or to denote supervised methods, where extraction is peformed implicitely during searches for specified objects and localization procedures.

The task of tracking is closely related to the topics of video tracking, data association, filtering, point matching, and `tracking of non-rigid objects'. In general, these tasks aim to maintain localization of the positions and/or shapes of objects over time. Areas of interest in this field involve developing metrics for judging the similarity between two object representations (for example, potential representations of the same object at adjacent frames in a video), techniques that provide a way to conduct efficient searches over a frame for regions that are similar to a given object, and methods for predicting the future position or appearance of an object. A great deal of research in this field over the past decade has focused on the development of algorithms to track multiple objects simultaneously. There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera), complex object interactions, object coalescence (where one object is tracked by multiple times/redundantly), objects with similar appearances, variable (and potentially high) numbers of objects, and objects that enter and exit a field of view at different times.

There does not yet exist a major body of research concerned primarily with methods for combining extraction, localization, and tracking to perform cohesive unsupervised detection of arbitrary objects in videos. Many papers that attempt to accomplish this goal focus on one of extraction, localization, or tracking, and attempt to integrate in methods from the other two tasks. However, a few key schemes for unsupervised detection and tracking do indeed exist. A widely applied, though unsophisticated, technique to carry out such a procedure is called blob tracking. Blob tracking typically refers to methods that perform extraction and localization at each frame independently--for example, by assuming that object pixels have a different intensity than background pixels, setting a fixed pixel intensity threshold for extraction of object pixel locations, and segmenting the extracted data into `blobs' to accomplish localization. Afterwards, tracking is often carried out by finding the centroid of each blob and applying standard multiple target tracking algorithms (described in [section]). Note that blob tracking carries out the three main tasks described above, so it satisfies the criteria for the type of unsupervised detection and tracking algorithms studied in this work; there are, however, many challenges associated with blob tracking that render it impractical in many settings (detailed in [section]).

This paper presents a new method for unsupervised multiple object detection and tracking. We introduce, and provide details on the theory behind, a technique involving dependent Dirichlet process mixture models. Specifically, this work applies the model to data gained during the extraction process; note that data of this type is refered to at different points in the paper as `extraction data', `data points', 'points', or 'observations'. The model is first presented in an abstract form, which could be adapted to a wide range of extraction procedures and object appearance representations (mainly left for future studies). We then describe inference algorithms that allow object localization and tracking to be carried out. Afterwards, we formulate a definite model for data gained during a specific extraction procedure and implement the described inference procedures to carry out multiple object detection and tracking in a number of experiments.

\subsection{Related Work}

Related work can be broken down into studies relating to extraction procedures, unsupervised localization procedures, single object tracking methods, multiple object tracking methods, supervised multiple object detection and tracking methods, and unsupervised multiple object detection and tracking methods.

Extraction procedures can usually be classified as either model-based methods or heuristic-based methods. Model-based methods emphasize the detection of shapes of regions containing objects accurately and reducing background noise. For example, foreground extraction has been robustly carried out by modeling each pixel over time in order to infer a probability distribution over pixel values; if the pixel takes on a very unlikely value at a future time, it is marked as a foreground pixel at that point \citep{stauffer_1999, elgammal_2000, elgammal_2002}. These pixel modeling techniques have been adapted to allow for time-dependent probability distributions over pixel values, which help account for slighlty changing backgrounds, such as those caused from brief movements of background objects (for example, trees in outdoor scenes). Note that these techniques treat each pixel independently; this has a tendency to cause errors in foreground detection, resulting in fragmentation of foreground images (which is problematic as many of these methods intend to blob the foreground objects after detection in order to perform segmentation and locate objects' centroids). Post processing has been applied to add information regarding the edges of objects in an attempt to decrease fragmentation \citep{turdu_2007}. Others have attempted to achieve better foreground modeling in cases where the foreground and background exhibit similar color distributions (the so called `color similarity problem') by, for example, incorporating models of the foreground into a typical background model (such as the one described above) \citep{gallego_2009, mchugh_2009}.

On the other hand, heuristic-based methods perform extracting by looking for temporal changes in pixel value characteristics. While these methods locate regions with objects less accurately than model-based methods, they are often much less computationally expensive. Frame differencing and background subtraction are two such methods, and involve comparing pixel values, or groups of pixel values, between two images, to detect a change surpassing some threshold. The goal of these techniques is to find locations in a scene that display motion, and to deem these moving areas the foreground of a scene.  Often, background subtraction refers to comparing an image containing targets with an image of the background without targets or with some model of the background that is learned as the video progresses (in these cases, objects that do not move throughout a scene will be treated as the background), while frame or image differencing refers to comparing two subsequent images in a video. Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking schemes \citep{perez_2002}. In addition to the two techniques mentioned above, there exists other ways of extracting movement from an image, such as through techniques involving optical flow (a calculation based on a so called `generalized gradient model' of an image, which attempts to capture the speed and direction of movement over areas in the image) \citep{horn_1981, bobick_2001}. In \citep{black_2000}, a motion detection and object tracking method is developed based on the relative motion of moving objects' boundaries, which are again found through analysis of optical flow.

There are an extensive collection of methods for tracking single, non-rigid, objects. For example, color has long been used for tracking. In this case, tracking refers to the task of maintaining the location of an object after its initial position has been specified. In general, these schemes operate by modeling an object with some color-based appearance model, and using this model to find the object in subsequent frames. One classic approach involves modeling the color of a target (in particular, a distribution over the hue-saturation space of a target region) with a Gaussian mixture model (GMM), and choosing subsequent target positions by searching surrounding areas for regions that yield similar GMMs. Furthermore, the GMM is often allowed to adapt over time, and slowly change to model changes in lighting or other smooth time-based variations in the target's color \citep{raja_1998, mckenna_1999, jepson_2003}. Others have attempted to abstract this work with a non-parametric color modeling approach based on kernel density estimation, which does not assume a specific underlying distribution (such as the Gaussian mixture in the previous case) and instead converges to reasonable distribution that depends on the data \citep{elgammal_2001}. Additionally, color has been successfully applied as the feature in the mean-shift procedure (kernel-based method) for tracking \citep{comaniciu_2003, perez_2002, nummiaro_2003, lee_2011}.

The mean-shift procedure (also known as `kernel-based object tracking') attempts to provide a robust way for non-rigid objects to be tracked. This method is beneficial because it optimizes the search for the `next' position of an object (i.e. the search for the region in a frame which is most similar to a target region in the previous frame). This procedure--a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges for each data point. Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors. It can be shown that, through this procedure, each data point becomes associated with a local point of high density (dependent upon the underlying weight distribution specified by the kernel) which naturally allows for clustering \citep{cheng_1995}. The mean shift algorithm has been implemented successfully to allow for a sort of kernel-based object tracking \citep{comaniciu_2003, comaniciu_1999, comaniciu_2000}. Given the current position of an object at a given frame, the goal of tracking is often to find a nearby position in the next frame that has the most similar distribution over some common set of features. Usually this must be done through an exhaustive search, comparing the similarity of distributions at each nearby position with that at the current position. However, if a certain type of `isotropic kernel' (mean-shift kernel) known as the Bhattacharyya coefficient is chosen as a similarity metric between the feature distributions at two positions, it creates a smooth function where gradient descent techniques can be used to quickly converge upon an optimal subsequent position without using an exhaustive search. Many papers are concerned with applying this kernel-based object tracking scheme to different sets of features or with different kernels. Additionally, the scheme has been attempted with an adaptive kernel whose shape, scale, and orientation is influenced by a target being tracked \citep{wang_2004}, with a kernel adjusted by the estimated centroid of a tracked target \citep{mehmood_2009}, and with a heirarchical version of the mean-shift procedure \citep{dementhon_2002}.

Some have tried to extent single object tracking schemes to tracking multiple targets; at the most basic level, this involves initializing multiple targets and running an instance of (single) VOT for each, either simultaneously or in succession \citep{perez_2002}. Furthermore, if performed simultaneously, the tracking of each target can be made dependent upon characteristics of other targets (such as their proximity) to resolve errors and improve tracking (such as those caused by the incorrect merging of two targets) \citep{khan_2004, vermaak_2003} .

% Localization-focused. Clustering methods, segmentation methods, object appearance models and representations (allows for consistent object localization over time).

% Methods for tracking single objects in videos: filtering, similarity metrics / localization (and how they play a role), mean shift procedures.

% extensions to multiple visual object tracking. problems with multiple independent single object trackers (coalesence, object interaction, object occlusion, objects entering and leaving scene [and, relatedly, variable numbers of objects]).
% Multiple object tracking attempts at overcoming various the various challenges specified above (include nonparametric varmaak/doucet for maintaining multi-modality here?---sure, and then mention again later with fox's work?)
% End with description of initialization problem.

% Methods for detection and tracking of objects. Heuristic initialization at beginning and then carrying out above methods (problems with varying numbers of obects / objects entering or leaving scene). Supervised learning of specific targets to allow for initialization and provide better localization throughout. Various research has been towards integrating unsupervised extraction and localization into multiple object tracking methods. Particle filters with integrated detection (at least two of these). Introduce concept of blob tracking and define basic algorithm (some manner of extraction, independent/per-frame localization, tracking; note that each object is represented in each frame as a low-dimensional point [or higher dimensional thing like region/outline with sophisticated blob/contour detectors like bramble and others]). The key thing characterizing a blob detector is that the position of each object (or the spatially separate region associated with each object) is assigned/fixed at each frame without the aid of tracking (ie if localization/segmentation can be carried out at each frame independently), and tracking is done with this data. Schemes often are simply segmentation of extraction data into spatially disjoint blobs at each frame (where blobs  are not modeled in any principled manner), returning the position or region associated with each object at each frame.

% When the state of objects (for example, position and/or other characteristics) can be found independently at each frame, well established methods in the field known as Multiple Target Tracking (MTT) can be carried out to perform tracking of multiple objects. Two classic algorithms for MTT are known as MHT and JPDA. In general, MTT methods are useful for procedures that can represent each object at each frame as a point or region without localization information from this object at a previous frame (which requires the use of tracking). MTT in the literature is often quite separate from fields of computer vision involving object detection and tracking research. In general, the situation arises as video data without modification is very high dimensional (discussion of difference between MTT and detection methods here?).

% Methods like blob tracking which attempt to find the state of each object independently at each frame have a hard time dealing with occlusion and other complexities (except for sophisticated blob trackers like bramble which isn't really blob tracking). Attempts have been made to perform time dependent clustering on extraction data (look at cluster tracking paper for good def), to carry out joint localization and tracking; this has been refered to as cluster tracking. Work of Pece, MTT of multiple targets work(?). The work presented in this paper most closely resembes the topics listed here. Does cluster tracking on whichever extraction data, using a time dependent Dirichlet Process mixture model (to better estimate and maintain the number of targets, as was shown to be effective in vermaak and fox), which takes the form of a sequence of clusters, where the objects' appearances are modeled in a principled manner by the cluster component distributons, and the tracking behavior is modeled by time-dependencies between adjacent elements in the cluster-sequences.

\subsection{Contributions}

This paper applies a time dependent Dirichlet process mixture to model data gained during extraction procedures, provides background as to how this Bayesian nonparametric model allows for the incorporation of prior knowledge about object appearances and motion behaviors, describes methods for unsupervised detection and tracking of arbitrary objects in terms of Bayesian inference procedures, and presents a specific implementation of the model and inference procedures to carry out detection and tracking on a number of target-types in different settings.

% List here the benefits of this idea that should be discussed/listed:
This paper is heavily model-based, which differs from many papers in the field of multiple object tracking that do not explicitely describe models which underlie various procedures that are carried out. Solutions in many papers are, instead, algorithm oriented. The model-based approach allows rigor to be kept as inference procedures are developed. Furthermore, development of inference procedures are performed separately from development of the model, and various well studied inference and approximate inference methods can be carried out to achieve algorithmic results that are (in some cases) known to be optimal.
%treated separately, which allows for multiple methods to be carried out in a unified framework and   these solutions are  but provide algorithm-based presentations.
%	This way allows rigor to be kept as inference procedures are developed.
%	Furthermore, inference is treated separately; this allows multiple detection and tracking algorithms to be developed (in a well-formed/rigorous fashion) by developing different inference techniques. Also, inference on graphical models is a well-structured field and allows for use of techniques in this field as well as better interpretations of the resulting detection/tracking algorithms.

% (related to previous?) models objects as distribution-components (of a mixture model) as opposed to point-targets or targets with noise that adheres to some distribution (is this truly a distinction and is it good?)

This idea presents a framework which allows for time-varying models of objects' appearances and motion behaviors to be incorporated, and is formulated in a Bayesian manner that allows distributions underlying the parameters of these models to be specified in a principled manner. It has been shown to work successfully in conjunction with simple (noisy) extraction methods. More robust, expensive, or in general accurate methods can be used for further accuracy. This method is able to handle the appearance and disappearance of objects from a scene, variable numbers of objects, cases where objects show time varying changes in behavior, and cases of occlusion.

Section two of this paper provides an overview of the model and describes how aspects can be modified for for modeling object appearance and motion behavior. Section three presents a specific implementation of the model and a number of inference procedures for carrying out object detection and tracking. Section four describes a number of experiments carried out with this implementation. Section five describes results of these experiments, comments on further directions of study, and provides concluding remarks.

\section{Model Overview}
%
This section provides details on the nature of the data to be modeled, Dirichlet process mixture (DPM) models, and dependent Dirichlet process mixture (DDPM) models. We end by making explicit the connection between elements of the model and object characteristics that these elements represent (for example, the appearance and motion behavior of objects are specifically captured). The provided model is formulated here in a sufficently abstract manner to allow for various representations of objects to be specified in future studies.

\subsection{Data}
%
We defined the task of extraction to be the unsupervised decision of which spatiotemporal regions of a video constitute objects; hence, the extraction data modeled in this work are D-dimensional observations that in some way provide information about the regions of the video frame that contain objects. A simple example of this information is shown by the three-dimensional extraction data $x = ( d_{1}, d_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $d_{1}$ and $d_{2}$ are the spatial dimensions of pixels that comprise regions containing objects in a given frame, and $t$ is the discrete time index of the frame. These three features are always available once regions containing objects are discerned, and should always be kept as features of the data for modeling (as position information is key for reasonable tracking and time information is necessary for our intention to model the temporal dependencies of the data).

Examples of additional features that could potentially be extracted from video frame regions containing objects include color information, pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, texture representations, transform results (such as results of hough or fourier transforms), and quantiative region descriptions derived from image segmentation procedures; the main idea is to choose features which are able to be easily extracted or computed, capture variability in the appearance of objects, are applicable to a wide variety of object types, and, when viewed in cumulation for multiple observations, are distributed in a way that allows for modeling and tractable inference (this is discussed further in section 3).

The baseline three-dimensional extraction data $x = ( d_{1}, d_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$ from pixels present in regions containing objects are distributed roughly as worm-like shapes that follow the paths of individual objects as they move, when they are plotted in three dimensions (where we let the frame number, our discrete representation of time, ascend the vertical axis and the spatial positions occupy the horizontal axes). An example of data with such a distribution can been seen in Figure 1. The data consists of the three-dimensional baseline features and was gained by extraction peformed on a 50 frame sequence showing five ants moving across the video scene.
\begin{figure}[h]
        \center{\includegraphics[width=75mm]{paths.png}}
        \caption{\label{Figure 1:} Baseline three-dimensional extraction data gained from a video of multiple moving targets. The vertical axis denotes the frame number and the horizontal axes denote $d_{1}$ and $d_{2}$ spatial coordinates, respectively, of pixels within a given frame. The data comes from a 50 frame video segment showing 5 moving ants (note that one ant enters the scene near the middle of the segment).}
\end{figure}
Inference performed on the time dependent Dirichlet process mixture model (described in the following sections) carries out clustering of the data, which allows each worm-like data cloud to be isolated and independent objects to be localized and tracked.

\subsection{Dirichlet Process Mixtures}

Dirichlet process mixtures fall under the heading of Bayesian nonparametric (BNP) models. These models have been widely used over the past decade to perform nonparametric density estimation and cluster analysis. They are, in particular, useful for estimating the number of latent classes (clusters) in mixture models. In this work, estimating the number of clusters in the data generated via extraction is shown in the following sections to be equivalent to estimating the number of distinct objects in a video. Consequentially, a model which is able to perform robust nonparametric estimation of the number of cluster is well suited to be applied to the task of unsupervised object detection and tracking.

The following sections are intended to provide a solid introduction to DPMs without a major foray into the theoretical aspects of the model that require a more rigorus technical setup. We give background on finite mixture models, Bayesian finite mixture models, and Dirichlet processes, presently, which allows us to subsequently describe DPMs in context.

\subsubsection{Finite Mixture Model}

A finite mixture model can be thought of as a probability distribution for an observation $x_{i}$ formulated as a linear combination of K mixture components (where the coefficients of the linear combination sum to one); each mixture component, in turn, is a probability distribution for $x_{i}$, given some parametric form. The finite mixture model can be written as

\begin{equation}
P(x) = \sum_{k=1}^{K} P(c_{i} = k)P(x|\theta_{k})
\end{equation}

where $c_{i} \in \{ 1, \ldots, K \}$ denotes the assignment of $x_{i}$ to a given mixture, which allows the coefficients of the linear combination to be written as $P(c_{i} = k)$ (note that these coefficients now sum to one naturally due to the nature of probability distributions). Additionally, $\theta_{k}$ denotes the parametric form of the $k^{\text{th}}$ mixture component. We also define $p_{k} = P(c_{i} = k)$ for all $k \in \{ 1, \ldots, K \} $. This model can also be written hierarchically as
\begin{align}
\begin{split}
	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}}) \\
	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
\end{split}
\end{align}
where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $c_{i}$ are the mixture component assignments associated with each observation, and the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e. the distribution to be mixed, $F(\theta_{c_{i}})$).

\subsubsection{Bayesian (Finite) Mixture Model}

The finite mixture model of the previous section can be extended to a Bayesian perspective by viewing the distribution parameters that were previously point values, $\theta_{c_{i}}$ (the mixture component parameters) and $p_{1}, \ldots, p_{K}$ (the mixture component assignment weights), as random variables and providing each with a prior distribution. In this case, the prior distribution $\mathbb{G}_{0}$ is placed on the mixture component parameters, and the prior distribution $\text{Dir}(\alpha/K, \ldots, \alpha/K)$ is placed on the K mixture component assignment weights. The resulting Bayesian mixture model can be formulated in a hierarchical representation as
\begin{align}
\begin{split}
	\label{bayesian_mixture_model}
	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}}) \\
	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
	\theta_{c_{i}}  &\sim  \mathbb{G}_{0} \\
	p_{1}, \ldots, p_{K}  &\sim  \text{Dir}(\alpha/K, \ldots, \alpha/K)
\end{split}
\end{align}
where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $c_{i}$ are the mixture component assignments associated with each observation, the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e. the distribution to be mixed, $F(\theta_{c_{i}})$), the $\theta_{c_{i}}$ are drawn from their prior distribution $\mathbb{G}_{0}$, and $p_{1}, \ldots, p_{K}$ are drawn from their prior distribution, where `Dir' refers to a standard Dirichlet distribution.

\subsubsection{Dirichlet Process}

% Define Dirichlet distribution first? Each draw from a K-dimensional Dirichlet lies on the K-simplex (i.e. is a vector of K values which sum to 1) and can therefore be thought of as a K-dimensional discrete distribution.

The Dirichlet process (DP), first introduced by \cite{ferguson_1973} in 1973, may be intuitively viewed as a probability distribution over discrete probability distributions. Accordingly, draws from a DP are probability mass functions (PMFs). A DP is parameterized by a base distribution $\mathbb{G}_{0}$, which is a probability distribution over a set $\Theta$, and a concentration parameter $\alpha \in \mathbb{R}_{+}$. We say that $G$ is a random PMF distributed according to a DP, written $G \sim \text{DP}(\alpha, \mathbb{G}_{0})$, if the following holds for all finite partitions $A_{1}, \ldots, A_{p}$ of $\Theta$:
\begin{equation}
(G(A_{1}), \ldots, G(A_{p})) \sim \text{Dir}(\alpha \mathbb{G}_{0}(A_{1}), \ldots, \alpha \mathbb{G}_{0}(A_{p}))
\end{equation}
Where `Dir' denotes a standard Dirichlet distribution. The parameters $\mathbb{G}_{0}$ and $\alpha$ may be intuitively viewed as the mean and precision of the DP. This is due to the fact that if the base distribution $\mathbb{G}_{0}$ is a distribution over $\Theta$, $A \subset \Theta$, and $G \sim \text{DP}(\alpha, \mathbb{G}_{0})$, then the following holds:
\begin{equation}
\mathbb{E}[G(A)] = \mathbb{G}_{0}(A)
\end{equation}
\begin{equation}
\text{Var}[G(A)] = \mathbb{G}_{0}(A) (1 - \mathbb{G}_{0}(A)) / (\alpha + 1)
\end{equation}
Hence, the expectation of $G(A)$ is $\mathbb{G}_{0}$, the variance of $G(A) \rightarrow 0$ as $\alpha \rightarrow \infty$, and $G$ converges pointwise to $\mathbb{G}_{0}$ when $\alpha$ is unbounded.

\subsubsection{Dirichlet Process (Infinite) Mixture Model}

A Dirichlet process mixture model, also refered to as an infinite mixture model, is an extension of the Bayesian mixture model described previously. 
When using a DP as a prior in a Bayesian mixture model, $\Theta$ can be viewed as the set of parameters of the component mixture distributions.
%
A DPM is formulated by allowing the prior mixing distribution (distribution over component weights?) in a standard mixture model to be distributed according to a DP; this allows for modeling data where the true number of latent mixture components is unknown and arbitrarily large by assuming an infinite number of components, of which only a finite amount are expressed by the data. In particular, the DPM can be defined hierarchically as
\begin{align}
\begin{split}
	x_{i}|\phi_{i} &\sim F(\phi_{i}) \\
	\phi_{i} | \mathbb{G}  &\sim  \mathbb{G} \\
	\mathbb{G} | \alpha, \mathbb{G}_{0}  &\sim  \text{DP}(\alpha, \mathbb{G}_{0})
\end{split}
\end{align}
where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $\phi_{i}$ are parameters defining the mixture component from which the $i_{th}$ observation is drawn (i.e. the distribution to be mixed, $F(\phi_{i})$), and the $\phi_{i}$ are drawn from $\mathbb{G}$, which is in turn drawn from a DP with base distribution $\mathbb{G}_{0}$ and parameter $\alpha$. See \cite{gasthaus_2008} and \cite{gasthaus_thesis} for more details on this formulation. Note that difference between the formulation and indexing of the clusters in this model compared to the previous two models. This definition can be shown to be equivalent to the Bayesian mixture model defined in \eqref{bayesian_mixture_model}, when K is taken to be unbounded. This is the origin for references to this model as the infinite mixture model.

An alternative formultion of the DPM is known as the Chinese resturant process (CRP) sampling representation. If we let $K$ be the number of distinct mixture components in the above model, we can write the distinct mixture components as $\theta_{1}, \ldots, \theta_{K}$. Let $c_{1}, \ldots, c_{N}$ (where $c_{i} \in \{1, \ldots, K \}$) be class assignment variables that indicate the cluster to which observation $x_{i}$ is assigned. This allows the DPM to be represented by the CRP, a discrete-time stochastic process that defines a partition of the set $\{ x_{1}, \ldots, x_{N} \}$ (via the elements' assignments $c_{1}, \ldots, c_{N}$). The CRP allows samples to be drawn from the conditional distribution of the indictor variables $c_{i}$, and can be expressed as
\begin{align}
\begin{split}
	P(c_{i} = c_{j} \text{  for some  } j<i) &= \frac{m_{k}}{i-1+\alpha}\\
	P(c_{i} \neq c_{j} \text{  for all  } j<i) &= \frac{\alpha}{i-1+\alpha}
\end{split}
\end{align}
where $m_{k}$ is the cardinality of the set $\{ c_{j} | (c_{j}=c_{i}=k)  \wedge  (j < i) \}$ and $\alpha$ is the parameter of the DP prior on $\mathbb{G}$.

\subsection{Dependent Dirichlet Process Mixtures}

The goal of dependent Dirichlet process mixtures (DDPMs) is to allow modeling of data that is not independent and identically distributed (i.i.d) but instead has some underlying dependencies. For example, data generated during extraction procedures from videos have some associated temporal structure, since tracked objects display time dependent characteristics (in the features that denote position, as well as those which denote other appearance characteristics).

To account for the dependent behavior of data, research has been conducted on developing models involving a sequence of DPMs, where components of the mixtures are dependent upon (or may be considered `tied to') corresponding components at adjacent positions in the sequence. For example, if the data shows temporal dependence, the goal might be to create a sequence of DPMs, one for each time-step, where the components of the mixture at each step are dependent upon corresponding components in the both the following and previous time steps. Research towards models which accomplish such goals are described in (list citations).

More rigorously, we take the definition of a DDPM to be a stochastic process defined on the space of probability distributions over a domain, which are indexed by time, space, or a selection of other covariates in such a way that the marginal distribution at any point in the domain follows a Dirichlet process (adapted from definitions found in [cite griffin and steel and gasthaus-thesis]).

\subsubsection{Generalized Polya Urn Dependent Dirichlet Process Mixture}

Data gained by performing extraction on videos containing objects is significantly time-dependent. Futhermore, the clusters of data, each of which represents an object in the video, might change in number as time progresses, since objects can enter and exit a scene. The specific DDPM used to model extraction data in this work while handling the above challenges is known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM), and was introduced by \cite{caron_2006} in 2006.

The GPUDDPM, when applied to data over T discrete time steps, can be viewed as a collection of DPMMs (one for each $t \in \{1, \ldots, T \}$), which are linked together by dependencies among their respective parameters. For example, the mixture component parameters at a given time step are dependent upon the mixture component parameters in adjacent time steps and the distribution and number of distinct cluster assignments at a given time step are dependent upon the same over a wider range of time steps, upon the distribution of observations' cluster assignments over this range, and upon a deletion procedure, described below.

To link together the parameters of mixture components in adjacent time-steps, a transition kernel $P(\phi_{k}^{t} | \phi_{k}^{t-1})$ is implemented, which provides a distribution over parameters of a mixture component in a given time-step (as a function of equivalent parameters in adjacent time-steps). One caveat is that each mixture component must be drawn independently from $\mathbb{G}_{0}$ (the base distribution of the DP, which acts as a prior distribution for the cluster parameters) which we achieve by making $\mathbb{G}_{0}$ the invariant distribution of $P(\phi_{k}^{t} | \phi_{k}^{t-1})$ (note that the transition kernel is a markov chain). Additionally, to account for varying numbers of clusters, there is a deletion procedure (described in \cite{caron_2007}) by which observations are considered "removed" from their assigned clusters at a given time-step (which modifies the value of $m_{k}$ defined above in the CRP representation). We introduce variables $d_{1}, \ldots, d_{N}$ which denote the times at which the assignments of given observations are considered to be removed. With these variables, we can now express the $m_{k}$ value at time $t$ as
\begin{equation}
m_{k,t} = \sum_{t' = 1}^{t} \mathbb{I}[(c_{t'}=k) \wedge (t < d_{t'})]
\end{equation}
where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise. Additionally, for an observation $x_{i}$ at a given time-step $t$, the deletion time $d_{i}$ can be defined to be $d_{i} = t + l_{i}$, where $l_{i}$ is considered the lifespan of an assignment, is distributed geometrically, and can be expressed as
\begin{equation}
l_{i} | \rho  \sim  \rho(1 - \rho)^{l_{i}}
\end{equation} \\
with parameter $\rho$. This process adheres to what \cite{caron_2007} calls a `uniform deletion strategy' over all the observations' assignments (since assignments to each cluster have equal probability of being deleted), though a more complex deletion strategy, dependent upon cluster size, can also be implemented.

Using the cluster size and observation deletion terms introduced above, we can define the GPUDDPM hierarchically as, for each time step $t = \{1, \ldots, T\}$
\begin{align}
\begin{split}
	x_{i}^{(t)} | c_{i}^{(t)}, \theta_{c_{i}^{(t)}}^{(t)} &\sim F(\theta_{c_{i}^{(t)}}^{(t)}) \\
	c_{i}^{(t)} | m^{(t-1)}, \alpha  &\sim  \text{CRP}(m^{(t-1)}, \alpha) \\
	m^{(t-1)} | m^{(t-2)}, c_{i}^{(t-1)}  &\sim \text{DEL}(m^{(t-2)}, c_{i}^{(t-1)}, \rho, \gamma) \\
	\theta_{k}^{(t)} | \theta_{k}^{(t-1)}   &\sim 
											\begin{cases}
											P(\theta_{k}^{(t)} | \theta_{k}^{(t-1)}) \\
											\mathbb{G}_{0}
											\end {cases}
\end{split}
\end{align}

The graphical corresponding with this formulation is shown in Figure 2.

\begin{figure}[h]
        \center{\includegraphics[width=120mm]{gpuddpm_gm_1.pdf}}
        \caption{\label{Figure 2:} Graphical Model of the Generalized Poly Urn Dependent Dirichlet Process Mixture}
\end{figure}

\subsubsection{Object Characteristics Modeled}

We would like make explicit the connection between certain elements of the GPUDDPM model and the characteristics of objects in vidoes that these elements represent.

First, the appearances of objects are moded by what we term the `likelihood and object appearance model', which we denoted by $F$. In the above modeling formulation, $F$ is the probability distribution for an observation, given that it is assigned to specified cluster (i.e. given that it is associated with a given object in a video). The likelihood and object appearance model is therefore

\begin{equation}
F(\theta_{c_{i}}) = P(x_{i}|\theta_{c_{i}})
\end{equation}

where $x_{i}$ is an observation, $c_{i}$ is its associated assignment, $\theta_{c_{i}}$ specifies the parameters of the parametric form of the $c_{i}^{th}$ cluster. The specific form of this model is dependent upon the observations $x_{i}$; recall that the observations, which as a baseline include spatial and temporal features, could include an arbitrary amount of additional features which need to be incorporated into this appearance model. A specific formulation of the appearance model can be seen in the proceeding section.

Object behavior is modeled by the dependencies between corresponding clusters at adjacent time steps. In particular, this relationship is captured by the transition kernel, $P(\theta_{k}^{t} | \theta_{k}^{t-1})$ (which captures the dependence of cluster $k$ at time $t$ on the same cluster at time $t-1$), described previously. The motion behavior of an object, i.e. the specific dependence of the parameters of cluster $k$ at a time $t$ on its parameters at time $t-1$, is therefore
\begin{equation}
Tr(\theta_{k, t} | \theta_{k, t-1}) = P(\theta_{k, t} | \theta_{k, t-1})
\end{equation}

Additionally, the base distribution $\mathbb{G}_{0}$ of the DP acts as a prior on the cluster parameters. This allows one to place a prior probability over appearances of objects. It is important that this prior is not chosen in such a specific manner that it limits this method from performing unsupervised detection of arbitrary objects.

\section{Model Specification and Inference}

A particular method for extraction from videos was chosen to generate data and a specific formulation of the model described in the previous section was constructed. Inference on this model was carried out to perform unsupervised detection and tracking of multiple arbitrary objects. Model choices are discussed and inference schemes are described in the following sections.

\subsection{Extraction}

Out of the numerous extraction procedures discussed at the beginning of this paper, we desired one that was as unsophisticated as possible, both to test the robustness of this method on potentially noisy extraction data and to ensure that the extraction procedure was applicable to a wide range of videos. We decided upon the extraction procedure of simple frame-differencing, which involves looking at the difference in pixel intensity or value in successive frames, and recording the positions of the pixels (or pixel groups) that have exhibited frame-wise change beyond a given threshold.

Frame differencing is simple, quick, able to be applied to a wide range of static, single-camera videos (note that videos used in the experiments described in the following sections were chosen to be static; moving-camera videos should be used in conjunction with applicable extraction methods that allow camera movement). We have found that, when correctly implemented, it is sufficiently general to extract the worm-like data clouds from a wide variety of videos containing multiple moving objects. A few examples of this extraction procedure on single video frames can be found in Figure 3.

\begin{figure}
\centering
{\includegraphics[width=45mm]{antpic2.png}}
%\hspace{2mm}
{\includegraphics[width=60mm]{frame_diff_ants_1.png}}
\caption{(a.) Frame of a video containing five ants. (b.) Frame differencing extraction results from the same frame (note: this is just a demo of the type of image i will have here... these aren't actually from the same frame, and also formatting is screwed up)}
\label{test}
\end{figure}

\subsection{Model Implementation in Experiments}
%

\subsubsection{Likelihood and Object Appearance Model}
For a given object $k$ at video frame $t$, the set of positions of the object's motion points are modeled with a product (over each position dimension) of one-dimensional Gaussians
\begin{equation}
P(x_{t_{i}}^{pos} | \mu_{t, k}, \lambda_{t, k}) = \prod_{d=1}^{D} \mathcal{N}(x_{t_{i}}^{pos_{d}} | \mu_{t, k}^{d}, \lambda_{t, k}^{d})
\end{equation}
where $x_{t_{i}}^{pos} = \{ x_{t_{i}}^{pos_{1}}, \cdots, x_{t_{i}}^{pos_{D}} \}$,  $\mu_{t, k}^{d}$ is the mean of the $k^{th}$ Gaussian at time $t$ for dimension $d$, and $\lambda_{t, k}^{d}$ is the precision of the $k^{th}$ Gaussian at time $t$ for dimension $d$.  Likewise, for a given object $k$ at video frame $t$, the set of color vectors of the object's motion points are modeled with a multinomial
\begin{equation}
P(x_{t_{i}}^{col} | m_{t, k}) = Mult(x_{t_{i}}^{col} | m_{t, k})
\end{equation}
where $x_{t_{i}}^{col} = \{ x_{t_{i}}^{col_{1}}, \cdots, x_{t_{i}}^{col_{V}} \}$, $m_{t,k} = \{ m_{t,k}^{1}, \cdots, m_{t,k}^{V} \}$, $\sum_{v=1}^{V} m_{t, k}^{v} = 1$, and $m_{t,k}^{v}>0 \hspace{8pt} \forall v \in \{ 1, \cdots, V \}$. The likelihood for a motion point observation is thus
\begin{equation}
P(x_{t_{i}} | \mu_{t, k}, \lambda_{t, k}, m_{t, k}) =  Mult(x_{t_{i}}^{col} | m_{t, k})   \prod_{d=1}^{D}  \mathcal{N}(x_{t_{i}}^{pos_{d}} | \mu_{t, k}^{d}, \lambda_{t, k}^{d}) 
\end{equation}

I want to justify distribution families chosen for appearance modeling in this implementation. To include: the edges of an object are detected well with frame differencing, so while extraction data densities aren't necessarily Gaussians with maxima over the targets, they are usually radially centered around the centroids of targets. Furthermore (thing frank said about gaussians and what they do---minimize least squared distance or something).

\subsubsection{Base Distribution $\mathbb{G}_{0}$ and Appearance Prior}
$\mathbb{G}_{0}$ is the base distribution of the underlying time-dependent Dirichlet process mixture; it also serves as a prior distribution for the parameters present in the likelihood. We make use of conjugate priors in the base distribution to allow for more efficient computation. In particular, for an object $k$ at time $t$, we let the prior distribution over the parameters of the motion point position distribution, $\mu_{t, k}$ and $\lambda_{t, k}$, be 
\begin{equation}
\mathbb{G}_{0}^{pos}(\mu_{t, k}, \lambda_{t, k} | \mu_{0}, n_{0}, a, b) = \prod_{d=1}^{D}[\mathcal{N}(\mu_{t, k}^{d} | \mu_{0}, n_{0} \lambda_{t, k}^{d})   Ga(\lambda_{t, k}^{d} | a, b)]
\end{equation}
where $\mu_{0}, n_{0}, a,$ and $b$ are parameters of the base distribution. Furthermore, for object $k$ at time $t$, we let the prior distribuion over the parameters of the motion point color vector distribution be
\begin{equation}
\mathbb{G}_{0}^{col}(m_{t, k} | q) = Dir(m_{t, k} | q)
\end{equation}
where $q = \{ q^{1}, \cdots, q^{V} \}$ is a parameter of the base distribution, where $q^{v}>0 \hspace{8pt} \forall v \in \{ 1, \cdots, V \}$. The base distribution is thus 
\begin{equation}
\mathbb{G}_{0}(\mu_{t, k}, \lambda_{t, k}, m_{t, k} | \mu_{0}, n_{0}, a, b, q) = Dir(m_{t, k} | q)  \prod_{d=1}^{D}[\mathcal{N}(\mu_{t, k}^{d} | \mu_{0}, n_{0} \lambda_{t, k}^{d})   Ga(\lambda_{t, k}^{d} | a, b)]
\end{equation}

\subsubsection*{Transition Kernels and Motion Behavior}
Let $\phi_{t,k}^{pos} = \{ \mu_{t, k}, \lambda{t, k} \}$ and $\phi_{t,k}^{col} = \{ m_{t, k} \}$.  We define two transition kernels, $P(\phi_{t, k}^{pos} | \phi_{t-1, k}^{pos})$ and $P(\phi_{t, k}^{col} | \phi_{t-1, k}^{col})$, which dictate, respectively, the random walk of the motion point position parameters and the motion point color vector parameters over time. The transition kernels must be chosen so that their invariation distributions are $\mathbb{G}_{0}$, i.e. such that the following hold:
\begin{equation}
\int \mathbb{G}_{0}^{pos}(\phi_{t-1, k}^{pos}) P(\phi_{t, k}^{pos} | \phi_{t-1, k}^{pos}) d\phi_{t-1, k}^{pos}  =  \mathbb{G}_{0}^{pos}(\phi_{t, k}^{pos})
\end{equation}
\begin{equation}
\int \mathbb{G}_{0}^{col}(\phi_{t-1, k}^{col}) P(\phi_{t, k}^{col} | \phi_{t-1, k}^{col}) d\phi_{t-1, k}^{col}  =  \mathbb{G}_{0}^{col}(\phi_{t, k}^{col})
\end{equation}

\subsection{Inference Implementations in Experiments}
%
Include a graphical model for each of the specified inference methods here? Or do I only need one graphical model for the model specified above?

Include algorithms for each of the inference procedures.

\section*{Experiments}

\subsection{Synthetic Videos}

\subsubsection{Occlusions}

\subsubsection{Similar Appearances}

\subsubsection{Different Numbers of Objects}

\subsubsection{Objects wih Diverse Characteristics}

\subsection{Real-life Videos}

\subsubsection{Complex Behavior}
%
Ants

\subsubsection{Occlusions}
%
Various people scenes

\subsubsection{Objects with Diverse Characteristics}
%
Traffic and Humans (PETS scenes)

\subsubsection{Standard Data for Comparison}
%
PETS and other standard datasets

\section*{Results and Discussion}

\section*{Conclusion}

\begin{small}
\bibliographystyle{plainnat}
\bibliography{ref_paper_draft} 
\end{small}

\end{document}

### 2011-09-06 ltx/paper_draft_old.tex
first draft of the paper. a lot of parts going well, a lot of parts still need work

--- 2011-08-30 ltx/lit_review.tex
@@
 \usepackage{fullpage} % Include this if you want to cram lots of things on a page 
-
 \usepackage{amsmath} % these are standard macro packages of the American Mathematical Society
 \usepackage{amssymb}
-	%\usepackage{stmaryrd}
-
 \usepackage{epsfig} % if you want figures
-
 \usepackage[margin=1.5in]{geometry}
-
 % \usepackage[numbers]{natbib}
 \usepackage[square]{natbib}
-
 \newcommand{\matlab}[1]
 {\centerline{\parbox{.9\textwidth}{\noindent\textsc{\bf MATLAB:} #1}}}
-
 \newcommand{\code}[1]{\texttt{#1}}
-
 \newcommand {\x}{\V{x}}
@@
 \newcommand {\V}[1]{\mbox{\boldmath$#1$}}
-
 \newcommand{\tab}{\hspace*{2em}}
+% need to include style sheet for journal i will submit to
 
@@
 
-\title{Literature Review}
-\author{Willie}
+\title{Object Detection and Tracking with Dependent Dirichlet Processes}
+%\title{Object Detection and Tracking with Bayesian Nonparametrics}
+\author{Willie and Frank}
 \maketitle
 \mbox{}
+
+\section*{Introduction}
+An unsupervised algorithm for the automated localization and tracking of arbitrary objects in videos is desirable; an algorithm of this sort is one with the capability to decide which parts of a video constitute objects or regions of interest (detection), find the positions of distinct objects (localization), and maintain the positions of all detected objects over time (tracking).
+\\
+\\
+The march towards such an algorithm can be seen in the approaches taken to solve a number of related problems over the past three decades.
+Multiple target tracking (MTT), also know as data association or point matching, involves providing a general localization and tracking solution for objects represented as two-dimensional points; a few algorithms have been well established since the 1980s.
+Single object tracking (SOT), also known as video tracking, 2D tracking, or 'tracking of non-rigid objects', refers to the task of finding the position of an object at a later time, given a current instantiated position.
+VOT often focuses on the
+
+At the same time, there has been much recent development in algorithms involving general methods for extracting parts of a video which may be considered objects or regions of interest often fall under the heading of foreground segmentation, background modeling, motion detection, or feature point detection.
+Schemes attempting to provide an automated method for the detection and tracking of objects or interest regions are varied, though often make use one common method falls under the general heading of blob tracking.
+\\
+One topic involves algorithms which attempt to perform combined object detection and tracking.
+\\
+Furthermore, research has focused on performing the above functions on videos with diverse object types, time varying object appearances, diverse numbers of objects, objects which may enter and exit the scene, occlusion of objects, presence of clutter, variable background characteristics, and unpredictable object behaviors.
+\\
+Details of specific methods in these topics will be covered presently. (?)
+\\
+\
+\\
+\subsection*{MTT and VOT}
+Discussion of tracking: MTT and VOT.
+Maybe I could go more into the discussion frank and I had here (instead of opening paragraph).
+More breadth, less depth on some of the VOT techniques (really read through the Survey on Object Tracking, ie include contour and template stuff).
+\\
+\\
+note: putting this here, not sure where it goes but needs to be hammered out in this intro: MTT (how i currently view it) treats each object as points, but does general tracking which leads to object detection (decides on number, structure, shape, etc.
+of paths, each of which represents an object over time).
+Current video detection and tracking schemes do this: at each frame, they try to detect an object (position, rough size/shape--need to define detection as something like this), and use these detections at each frame in object tracking (ranging from representing each detected object as a point and doing MTT methods, or a wide range of similarity-measure schemes of target characteristics/color which i call VOT methods.
+Also, object detection at further frames often aided by current detection/tracking...but even so, is still a detection at each frame, then tracking viewpoint).
+[note: this is almost a general tracking problem where each data point is a detected object].
+A slightly different paradigm when approaching this problem involves general tracking [i need to define tracking better] which leads to object detection, like MTT but where objects are not considered points but instead as having some spatial shape and other characteristics too (eg color).
+Sorta new ideas:
+The goal of this project is to demonstrate a model that provides a stable framework for general unsupervised detection and tracking of multiple objects of arbitrary nature.
+
+\subsection*{General Object Detection}
+Discussion on foreground/background modeling, feature characteristics of interest-objects, such as feature points and motion points, other general object detection methods (related to detection in video...any purely image-based object detection methods?).
+
+\subsection*{}
+Discussion of combined detection and tracking schemes.
+Example: blob tracking.
+The simplest involves detection at each time and then tracking (more particularly: feature extraction [foreground modeling], object detection [segmentation], then object tracking).
+Some have tried to do this in a different order: generalized tracking over time, which leads to object detection (more particularly: feature extraction [foreground modeling], generalized object tracking [most akin to MTT [as i view it...learn more about MTT before i say this] but for VOT], results complete object detection).
+Talk about conception benefits of this reverse order over detection then tracking.
+\\
+Discussion on this project.
+This project attempts to do the final method presented in the background (general multiple object tracking leading to object detection).
+Introduce overall goals and benefits of this idea, terms/concepts/methods/models that will be discussed in future sections.
+
+\section*{Model Overview}
+Discuss types of models necessary for this generalized object tracking goal described above (e.g.
+models that allow for nonparametric decisions as to number of targets, size of targets, time over which targets are tracked/in scene).
+\\
+\\
+Discuss Dirichlet Mixture Model and Dirichlet Process.\\
+\\
+Discuss dependent Dirichlet process concept, and time-dependent Dirichlet process mixtures, and gpudpm (gpuddp frank calls it?).
+
+\section*{Model Specification and Inference}
+
+\section*{Experiments}
+
+\section*{Results}
+
+\section*{Conclusion}
 
@@
 
-%\subsection*{Overview 2}
-%\vspace{6pt}
-%The phrase `multiple target tracking' (MTT) has been used to refer to two distinct but related problems: the first is the problem of detecting and tracking the positions of multiple targets of interest in a video, and the second is the problem of associating data points, where each data point is assumed to represent the centroid of some target whose position varies with time.
-The first is a problem in the field of machine vision and may use data association techniques from the second to accomplish MTT after targets of interest are detected and their positions represented as points.\\
-%\\
-%The goal of this project is to apply the generalized Polya urn Dirichlet process mixture model (GPUDPM) to videos containing multiple targets in a way that accomplishes data association without requiring explicit target detection.
-We hope to allow for a single flexible MTT algorithm that can handle videos with different numbers of targets, with diverse backdrops, whose targets display a wide range of characteristics, and where there are time-varying shifts in targets' characteristics, perspectives, or orientations.\\
-%\\
-%\\
-%The following are summaries of papers focusing on MTT as it relates to the association of target centroids, MTT as it relates to machine vision, and the generalized Polya urn Dirichlet process mixture model.
-
-% Hue et al. \cite{hue, 2002} introduce particle filters for MTT.
-
-%Oh et al. (Markov Chain Monte Carlo Data Association for General Multiple-Target Tracking Problems, 2004) uses an MCMC approach to perform general data association.
-
-% VMTT
-%Papers on this topic often involve algorithms that have a target detection phase (where the centroids of potential targets are collected, often by looking for specific target features) and a data association phase (where the centroids are sequenced in order to carry out tracking).\\
-
 \end{document}

### 2011-09-29 ltx/paper.tex
i updated the paper draft with the new model

\documentclass{article}

\usepackage[square,numbers]{natbib}
\usepackage{amsmath, epsfig}
\usepackage{amsfonts}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{easybmat}
\usepackage{footmisc}
\usepackage[usenames,dvipsnames]{color}
\usepackage{subfig}
\renewcommand\algorithmiccomment[1]{// \textit{#1}}
\newcommand{\ignore}[1]{}
\newcommand{\comment}[1]{}
\newcommand{\frank}[1]{\textcolor{red}{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}}}
\newcommand{\willie}[1]{\textcolor{green}{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}}}
\DeclareMathOperator*{\argmax}{arg\,max}

% need to include style sheet for journal

\begin{document}

\title{Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures}
\author{Willie Neiswanger and Frank Wood}
\maketitle
\mbox{}

\abstract{main points}

\section{Introduction}

\frank{this section should introduce our work and situate it relative to the literature}
\willie{okeydoke. this isn't ready yet. to avoid confusion I've just commented the whole section out.}

% % \subsection{Topic Overview and Terminology}

% This work is related to the study of algorithms that perform automated detection and tracking of arbitrary objects in videos; we define such algorithms as those with the ability to decide which spatiotemporal regions of a video constitute objects (extraction), find the positions and/or shapes of distinct objects (localization), and maintain the positions of the detected objects over time (tracking). Robust algorithms which can carry out these tasks are applicable to problems in the fields of surveillance, autonomous systems and robotics, and video summarization and compression. Approaches to these tasks have been proposed under different headings and in a variety of contexts. This section serves to introduce relevant terminology and topics of research.

% The task of extraction is closely related to the topics of foregound segmentation, background modeling, motion detection, and appearance feature extraction. In general, these methods aim to extract features from regions of a video frame that are considered objects or are distinguished in some way from the background of a scene. Areas of research in this field include attempts to handle situations where the background displays motion or other dynamic characteristics, reduce noise caused by the false extraction of backgrounds, handle abberant image effects present in some videos, and operate accurately on scenes involving illumination changes.

% The task of localization is closely related to the topics of object detection, position-finding, target segmentation, contour-detection, template matching, and recognition. These methods are mainly concerned with finding the positions and/or shapes of objects in a given video frame, or with discerning between objects similar in position or appearance. There has been a focus in this field on developing techniques for representing an object's appearance, both to allow for localization of objects on a per-frame basis and to allow for consistent localization of a given object over time. Research into the unsupervised detection of arbitrary objects (which we define as methods that accomplish both extraction and localization) has included work involving image segmentation, parsing, and pixel clustering. These topics usually relate to the automated partitioning of images into meaningful regions. Note that the popular phrase `object detection' is also typically used to refer to procedures that carry out extraction and localization; however, it is often reserved to denote methods that are tailored to find the position of specified objects or object types (as opposed to arbitrary object detection), or to denote supervised methods, where extraction is peformed implicitely during searches for specified objects and localization procedures.

% The task of tracking is closely related to the topics of video tracking, data association, filtering, point matching, and `tracking of non-rigid objects'. In general, these tasks aim to maintain localization of the positions and/or shapes of objects over time. Areas of interest in this field involve developing metrics for judging the similarity between two object representations (for example, potential representations of the same object at adjacent frames in a video), techniques that provide a way to conduct efficient searches over a frame for regions that are similar to a given object, and methods for predicting the future position or appearance of an object. A great deal of research in this field over the past decade has focused on the development of algorithms to track multiple objects simultaneously. There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera), complex object interactions, object coalescence (where one object is tracked by multiple times/redundantly), objects with similar appearances, variable (and potentially high) numbers of objects, and objects that enter and exit a field of view at different times.

% There does not yet exist a major body of research concerned primarily with methods for combining extraction, localization, and tracking to perform cohesive unsupervised detection of arbitrary objects in videos. Many papers that attempt to accomplish this goal focus on one of extraction, localization, or tracking, and attempt to integrate in methods from the other two tasks. However, a few key schemes for unsupervised detection and tracking do indeed exist. A widely applied, though unsophisticated, technique to carry out such a procedure is called blob tracking. Blob tracking typically refers to methods that perform extraction and localization at each frame independently--for example, by assuming that object pixels have a different intensity than background pixels, setting a fixed pixel intensity threshold for extraction of object pixel locations, and segmenting the extracted data into `blobs' to accomplish localization. Afterwards, tracking is often carried out by finding the centroid of each blob and applying standard multiple target tracking algorithms (described in [section]). Note that blob tracking carries out the three main tasks described above, so it satisfies the criteria for the type of unsupervised detection and tracking algorithms studied in this work; there are, however, many challenges associated with blob tracking that render it impractical in many settings (detailed in [section]).

% This paper presents a new method for unsupervised multiple object detection and tracking. We introduce, and provide details on the theory behind, a technique involving dependent Dirichlet process mixture models. Specifically, this work applies the model to data gained during the extraction process; note that data of this type is refered to at different points in the paper as `extraction data', `data points', 'points', or 'observations'. The model is first presented in an abstract form, which could be adapted to a wide range of extraction procedures and object appearance representations (mainly left for future studies). We then describe inference algorithms that allow object localization and tracking to be carried out. Afterwards, we formulate a definite model for data gained during a specific extraction procedure and implement the described inference procedures to carry out multiple object detection and tracking in a number of experiments.

% \subsection{Related Work}

% Related work can be broken down into studies relating to extraction procedures, unsupervised localization procedures, single object tracking methods, multiple object tracking methods, supervised multiple object detection and tracking methods, and unsupervised multiple object detection and tracking methods.

% Extraction procedures can usually be classified as either model-based methods or heuristic-based methods. Model-based methods emphasize the detection of shapes of regions containing objects accurately and reducing background noise. For example, foreground extraction has been robustly carried out by modeling each pixel over time in order to infer a probability distribution over pixel values; if the pixel takes on a very unlikely value at a future time, it is marked as a foreground pixel at that point \citep{stauffer_1999, elgammal_2000, elgammal_2002}. These pixel modeling techniques have been adapted to allow for time-dependent probability distributions over pixel values, which help account for slighlty changing backgrounds, such as those caused from brief movements of background objects (for example, trees in outdoor scenes). Note that these techniques treat each pixel independently; this has a tendency to cause errors in foreground detection, resulting in fragmentation of foreground images (which is problematic as many of these methods intend to blob the foreground objects after detection in order to perform segmentation and locate objects' centroids). Post processing has been applied to add information regarding the edges of objects in an attempt to decrease fragmentation \citep{turdu_2007}. Others have attempted to achieve better foreground modeling in cases where the foreground and background exhibit similar color distributions (the so called `color similarity problem') by, for example, incorporating models of the foreground into a typical background model (such as the one described above) \citep{gallego_2009, mchugh_2009}.

% On the other hand, heuristic-based methods perform extracting by looking for temporal changes in pixel value characteristics. While these methods locate regions with objects less accurately than model-based methods, they are often much less computationally expensive. Frame differencing and background subtraction are two such methods, and involve comparing pixel values, or groups of pixel values, between two images, to detect a change surpassing some threshold. The goal of these techniques is to find locations in a scene that display motion, and to deem these moving areas the foreground of a scene.  Often, background subtraction refers to comparing an image containing targets with an image of the background without targets or with some model of the background that is learned as the video progresses (in these cases, objects that do not move throughout a scene will be treated as the background), while frame or image differencing refers to comparing two subsequent images in a video. Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking schemes \citep{perez_2002}. In addition to the two techniques mentioned above, there exists other ways of extracting movement from an image, such as through techniques involving optical flow (a calculation based on a so called `generalized gradient model' of an image, which attempts to capture the speed and direction of movement over areas in the image) \citep{horn_1981, bobick_2001}. In \citep{black_2000}, a motion detection and object tracking method is developed based on the relative motion of moving objects' boundaries, which are again found through analysis of optical flow.

% There are an extensive collection of methods for tracking single, non-rigid, objects. For example, color has long been used for tracking. In this case, tracking refers to the task of maintaining the location of an object after its initial position has been specified. In general, these schemes operate by modeling an object with some color-based appearance model, and using this model to find the object in subsequent frames. One classic approach involves modeling the color of a target (in particular, a distribution over the hue-saturation space of a target region) with a Gaussian mixture model (GMM), and choosing subsequent target positions by searching surrounding areas for regions that yield similar GMMs. Furthermore, the GMM is often allowed to adapt over time, and slowly change to model changes in lighting or other smooth time-based variations in the target's color \citep{raja_1998, mckenna_1999, jepson_2003}. Others have attempted to abstract this work with a non-parametric color modeling approach based on kernel density estimation, which does not assume a specific underlying distribution (such as the Gaussian mixture in the previous case) and instead converges to reasonable distribution that depends on the data \citep{elgammal_2001}. Additionally, color has been successfully applied as the feature in the mean-shift procedure (kernel-based method) for tracking \citep{comaniciu_2003, perez_2002, nummiaro_2003, lee_2011}.

% The mean-shift procedure (also known as `kernel-based object tracking') attempts to provide a robust way for non-rigid objects to be tracked. This method is beneficial because it optimizes the search for the `next' position of an object (i.e. the search for the region in a frame which is most similar to a target region in the previous frame). This procedure--a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges for each data point. Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors. It can be shown that, through this procedure, each data point becomes associated with a local point of high density (dependent upon the underlying weight distribution specified by the kernel) which naturally allows for clustering \citep{cheng_1995}. The mean shift algorithm has been implemented successfully to allow for a sort of kernel-based object tracking \citep{comaniciu_2003, comaniciu_1999, comaniciu_2000}. Given the current position of an object at a given frame, the goal of tracking is often to find a nearby position in the next frame that has the most similar distribution over some common set of features. Usually this must be done through an exhaustive search, comparing the similarity of distributions at each nearby position with that at the current position. However, if a certain type of `isotropic kernel' (mean-shift kernel) known as the Bhattacharyya coefficient is chosen as a similarity metric between the feature distributions at two positions, it creates a smooth function where gradient descent techniques can be used to quickly converge upon an optimal subsequent position without using an exhaustive search. Many papers are concerned with applying this kernel-based object tracking scheme to different sets of features or with different kernels. Additionally, the scheme has been attempted with an adaptive kernel whose shape, scale, and orientation is influenced by a target being tracked \citep{wang_2004}, with a kernel adjusted by the estimated centroid of a tracked target \citep{mehmood_2009}, and with a heirarchical version of the mean-shift procedure \citep{dementhon_2002}.

% Some have tried to extent single object tracking schemes to tracking multiple targets; at the most basic level, this involves initializing multiple targets and running an instance of (single) VOT for each, either simultaneously or in succession \citep{perez_2002}. Furthermore, if performed simultaneously, the tracking of each target can be made dependent upon characteristics of other targets (such as their proximity) to resolve errors and improve tracking (such as those caused by the incorrect merging of two targets) \citep{khan_2004, vermaak_2003} .

% %  ------------------------------------------------------------------------------------
% % Localization-focused. Clustering methods, segmentation methods, object appearance models and representations (allows for consistent object localization over time).

% % Methods for tracking single objects in videos: filtering, similarity metrics / localization (and how they play a role), mean shift procedures.

% % extensions to multiple visual object tracking. problems with multiple independent single object trackers (coalesence, object interaction, object occlusion, objects entering and leaving scene [and, relatedly, variable numbers of objects]).
% % Multiple object tracking attempts at overcoming various the various challenges specified above (include nonparametric varmaak/doucet for maintaining multi-modality here?---sure, and then mention again later with fox's work?)
% % End with description of initialization problem.

% % Methods for detection and tracking of objects. Heuristic initialization at beginning and then carrying out above methods (problems with varying numbers of obects / objects entering or leaving scene). Supervised learning of specific targets to allow for initialization and provide better localization throughout. Various research has been towards integrating unsupervised extraction and localization into multiple object tracking methods. Particle filters with integrated detection (at least two of these). Introduce concept of blob tracking and define basic algorithm (some manner of extraction, independent/per-frame localization, tracking; note that each object is represented in each frame as a low-dimensional point [or higher dimensional thing like region/outline with sophisticated blob/contour detectors like bramble and others]). The key thing characterizing a blob detector is that the position of each object (or the spatially separate region associated with each object) is assigned/fixed at each frame without the aid of tracking (ie if localization/segmentation can be carried out at each frame independently), and tracking is done with this data. Schemes often are simply segmentation of extraction data into spatially disjoint blobs at each frame (where blobs  are not modeled in any principled manner), returning the position or region associated with each object at each frame.

% % When the state of objects (for example, position and/or other characteristics) can be found independently at each frame, well established methods in the field known as Multiple Target Tracking (MTT) can be carried out to perform tracking of multiple objects. Two classic algorithms for MTT are known as MHT and JPDA. In general, MTT methods are useful for procedures that can represent each object at each frame as a point or region without localization information from this object at a previous frame (which requires the use of tracking). MTT in the literature is often quite separate from fields of computer vision involving object detection and tracking research. In general, the situation arises as video data without modification is very high dimensional (discussion of difference between MTT and detection methods here?).

% % Methods like blob tracking which attempt to find the state of each object independently at each frame have a hard time dealing with occlusion and other complexities (except for sophisticated blob trackers like bramble which isn't really blob tracking). Attempts have been made to perform time dependent clustering on extraction data (look at cluster tracking paper for good def), to carry out joint localization and tracking; this has been refered to as cluster tracking. Work of Pece, MTT of multiple targets work(?). The work presented in this paper most closely resembes the topics listed here. Does cluster tracking on whichever extraction data, using a time dependent Dirichlet Process mixture model (to better estimate and maintain the number of targets, as was shown to be effective in vermaak and fox), which takes the form of a sequence of clusters, where the objects' appearances are modeled in a principled manner by the cluster component distributons, and the tracking behavior is modeled by time-dependencies between adjacent elements in the cluster-sequences.
% %  ------------------------------------------------------------------------------------

% \subsection{Contributions}

% This paper applies a time dependent Dirichlet process mixture to model data gained during extraction procedures, provides background as to how this Bayesian nonparametric model allows for the incorporation of prior knowledge about object appearances and motion behaviors, describes methods for unsupervised detection and tracking of arbitrary objects in terms of Bayesian inference procedures, and presents a specific implementation of the model and inference procedures to carry out detection and tracking on a number of target-types in different settings.

% % List here the benefits of this idea that should be discussed/listed:
% This paper is heavily model-based, which differs from many papers in the field of multiple object tracking that do not explicitely describe models which underlie various procedures that are carried out. Solutions in many papers are, instead, algorithm oriented. The model-based approach allows rigor to be kept as inference procedures are developed. Furthermore, development of inference procedures are performed separately from development of the model, and various well studied inference and approximate inference methods can be carried out to achieve algorithmic results that are (in some cases) known to be optimal.
% %treated separately, which allows for multiple methods to be carried out in a unified framework and   these solutions are  but provide algorithm-based presentations.
% %	This way allows rigor to be kept as inference procedures are developed.
% %	Furthermore, inference is treated separately; this allows multiple detection and tracking algorithms to be developed (in a well-formed/rigorous fashion) by developing different inference techniques. Also, inference on graphical models is a well-structured field and allows for use of techniques in this field as well as better interpretations of the resulting detection/tracking algorithms.

% % (related to previous?) models objects as distribution-components (of a mixture model) as opposed to point-targets or targets with noise that adheres to some distribution (is this truly a distinction and is it good?)

% This idea presents a framework which allows for time-varying models of objects' appearances and motion behaviors to be incorporated, and is formulated in a Bayesian manner that allows distributions underlying the parameters of these models to be specified in a principled manner. It has been shown to work successfully in conjunction with simple (noisy) extraction methods. More robust, expensive, or in general accurate methods can be used for further accuracy. This method is able to handle the appearance and disappearance of objects from a scene, variable numbers of objects, cases where objects show time varying changes in behavior, and cases of occlusion.

% Section two of this paper provides an overview of the model and describes how aspects can be modified for for modeling object appearance and motion behavior. Section three presents a specific implementation of the model and a number of inference procedures for carrying out object detection and tracking. Section four describes a number of experiments carried out with this implementation. Section five describes results of these experiments, comments on further directions of study, and provides concluding remarks.

\section{Model Overview}
\label{sec:modeloverview}
%
This section provides details on the nature of the data to be modeled, Dirichlet process mixture (DPM) models, and dependent Dirichlet process mixture (DDPM) models. We end by making explicit the connection between elements of the model and object characteristics that these elements represent (for example, the appearance and motion behavior of objects are specifically captured). The provided model is formulated here in a sufficently abstract manner to allow for various representations of objects to be specified in future studies.

\subsection{Data}
\label{sec:data}
%
We defined the task of extraction to be the unsupervised decision of which spatiotemporal regions of a video constitute objects; hence, the extraction data modeled in this work are D-dimensional observations that in some way provide information about the regions of the video frame that contain objects. A simple example of this information is shown by the three-dimensional extraction data $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $x_{1}$ and $x_{2}$ are the two spatial dimensions of pixels contained in regions comprising objects in a given frame, and $t$ is the discrete time index of the frame. These three features are always available once regions containing objects are discerned, and should always be kept as features of the data for modeling (as position information is key for reasonable tracking and time information is necessary for our intention to model the temporal dependencies of the data).

Examples of additional features that could potentially be extracted from video frame regions containing objects include color information, pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, texture representations, transform results (such as results of hough or fourier transforms), and quantiative region descriptions derived from image segmentation procedures; the main idea is to choose features which are able to be easily extracted or computed, capture variability in the appearance of objects, are applicable to a wide variety of object types, and, when viewed in cumulation for multiple observations, are distributed in a way that allows for modeling and tractable inference (this is discussed further in section 3).

We will use the phrase `baseline' three-dimensional extraction data to refer to the $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$ from pixels present in regions containing objects. These data are distributed roughly as worm-like shapes that follow the paths of individual objects as they move, when they are plotted in three dimensions (where we let the frame number, our discrete representation of time, ascend the vertical axis and the spatial positions occupy the horizontal axes). An example of data with such a distribution can been seen in Figure 1. The data consists of the three-dimensional baseline features and was gained by extraction peformed on a 50 frame sequence showing five ants moving across the video scene.
\begin{figure}[h]
        \center{\includegraphics[width=75mm]{paths.png}}
        \caption{\label{Figure 1:} Baseline three-dimensional extraction data gained from a video of multiple moving targets. The vertical axis denotes the frame number and the horizontal axes denote $d_{1}$ and $d_{2}$ spatial coordinates, respectively, of pixels within a given frame. The data comes from a 50 frame video segment showing 5 moving ants (note that one ant enters the scene near the middle of the segment).}
\end{figure}
Inference performed on the time dependent Dirichlet process mixture model (described in the following sections) carries out clustering of the data, which allows each worm-like data cloud to be isolated and independent objects to be localized and tracked.

\subsection{Dirichlet Process Mixtures}

Dirichlet process mixtures fall under the heading of Bayesian nonparametric (BNP) models. These models have been widely used over the past decade to perform nonparametric density estimation and cluster analysis. They are, in particular, useful for estimating the number of latent classes (clusters) in mixture models. In this work, estimating the number of clusters in the data generated via extraction is shown in the following sections to be equivalent to estimating the number of distinct objects in a video. Consequentially, a model which is able to perform robust nonparametric estimation of the number of cluster is well suited to be applied to the task of unsupervised object detection and tracking.

The following sections are intended to provide a solid introduction to DPMs without a major foray into the theoretical aspects of the model that require a more rigorus technical setup. We give background on finite mixture models, Bayesian finite mixture models, and Dirichlet processes, presently, which allows us to subsequently describe DPMs in context.

\subsubsection{Finite Mixture Model}

A finite mixture model can be thought of as a probability distribution for an observation $x_{i}$ formulated as a linear combination of K mixture components (where the coefficients of the linear combination sum to one); each mixture component, in turn, is a probability distribution for $x_{i}$, given some parametric form. The finite mixture model can be written as

\begin{equation}
P(x) = \sum_{k=1}^{K} P(c_{i} = k)P(x|\theta_{k})
\end{equation}

where $c_{i} \in \{ 1, \ldots, K \}$ denotes the assignment of $x_{i}$ to a given mixture, which allows the coefficients of the linear combination to be written as $P(c_{i} = k)$ (note that these coefficients now sum to one naturally due to the nature of probability distributions). Additionally, $\theta_{k}$ denotes the parametric form of the $k^{\text{th}}$ mixture component. We also define $p_{k} = P(c_{i} = k)$ for all $k \in \{ 1, \ldots, K \} $. This model can also be written hierarchically as
\begin{align}
\begin{split}
	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}}) \\
	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
\end{split}
\end{align}
where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $c_{i}$ are the mixture component assignments associated with each observation, and the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e. the distribution to be mixed, $F(\theta_{c_{i}})$).

\subsubsection{Bayesian (Finite) Mixture Model}

The finite mixture model of the previous section can be extended to a Bayesian perspective by viewing the distribution parameters that were previously point values, $\theta_{c_{i}}$ (the mixture component parameters) and $p_{1}, \ldots, p_{K}$ (the mixture component assignment weights), as random variables and providing each with a prior distribution. In this case, the prior distribution $\mathbb{G}_{0}$ is placed on the mixture component parameters, and the prior distribution $\text{Dir}(\alpha/K, \ldots, \alpha/K)$ is placed on the K mixture component assignment weights. The resulting Bayesian mixture model can be formulated in a hierarchical representation as
\begin{align}
\begin{split}
	\label{bayesian_mixture_model}
	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}}) \\
	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
	\theta_{c_{i}}  &\sim  \mathbb{G}_{0} \\
	p_{1}, \ldots, p_{K}  &\sim  \text{Dir}(\alpha/K, \ldots, \alpha/K)
\end{split}
\end{align}
where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $c_{i}$ are the mixture component assignments associated with each observation, the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e. the distribution to be mixed, $F(\theta_{c_{i}})$), the $\theta_{c_{i}}$ are drawn from their prior distribution $\mathbb{G}_{0}$, and $p_{1}, \ldots, p_{K}$ are drawn from their prior distribution, where `Dir' refers to a standard Dirichlet distribution.

\subsubsection{Dirichlet Process}

% Define Dirichlet distribution first? Each draw from a K-dimensional Dirichlet lies on the K-simplex (i.e. is a vector of K values which sum to 1) and can therefore be thought of as a K-dimensional discrete distribution.

The Dirichlet process (DP), first introduced by \cite{ferguson_1973} in 1973, may be intuitively viewed as a probability distribution over discrete probability distributions. Accordingly, draws from a DP are probability mass functions (PMFs). A DP is parameterized by a base distribution $\mathbb{G}_{0}$, which is a probability distribution over a set $\Theta$, and a concentration parameter $\alpha \in \mathbb{R}_{+}$. We say that $G$ is a random PMF distributed according to a DP, written $G \sim \text{DP}(\alpha, \mathbb{G}_{0})$, if the following holds for all finite partitions $A_{1}, \ldots, A_{p}$ of $\Theta$:
\begin{equation}
(G(A_{1}), \ldots, G(A_{p})) \sim \text{Dir}(\alpha \mathbb{G}_{0}(A_{1}), \ldots, \alpha \mathbb{G}_{0}(A_{p}))
\end{equation}
Where `Dir' denotes a standard Dirichlet distribution. The parameters $\mathbb{G}_{0}$ and $\alpha$ may be intuitively viewed as the mean and precision of the DP. This is due to the fact that if the base distribution $\mathbb{G}_{0}$ is a distribution over $\Theta$, $A \subset \Theta$, and $G \sim \text{DP}(\alpha, \mathbb{G}_{0})$, then the following holds:
\begin{equation}
\mathbb{E}[G(A)] = \mathbb{G}_{0}(A)
\end{equation}
\begin{equation}
\text{Var}[G(A)] = \mathbb{G}_{0}(A) (1 - \mathbb{G}_{0}(A)) / (\alpha + 1)
\end{equation}
Hence, the expectation of $G(A)$ is $\mathbb{G}_{0}$, the variance of $G(A) \rightarrow 0$ as $\alpha \rightarrow \infty$, and $G$ converges pointwise to $\mathbb{G}_{0}$ when $\alpha$ is unbounded.

\subsubsection{Dirichlet Process (Infinite) Mixture Model}

A Dirichlet process mixture model, also refered to as an infinite mixture model, is an extension of the Bayesian mixture model described previously. 
When using a DP as a prior in a Bayesian mixture model, $\Theta$ can be viewed as the set of parameters of the component mixture distributions.
%
A DPM is formulated by allowing the prior mixing distribution (distribution over component weights?) in a standard mixture model to be distributed according to a DP; this allows for modeling data where the true number of latent mixture components is unknown and arbitrarily large by assuming an infinite number of components, of which only a finite amount are expressed by the data. In particular, the DPM can be defined hierarchically as
\begin{align}
\begin{split}
	x_{i}|\phi_{i} &\sim F(\phi_{i}) \\
	\phi_{i} | \mathbb{G}  &\sim  \mathbb{G} \\
	\mathbb{G} | \alpha, \mathbb{G}_{0}  &\sim  \text{DP}(\alpha, \mathbb{G}_{0})
\end{split}
\end{align}
where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $\phi_{i}$ are parameters defining the mixture component from which the $i_{th}$ observation is drawn (i.e. the distribution to be mixed, $F(\phi_{i})$), and the $\phi_{i}$ are drawn from $\mathbb{G}$, which is in turn drawn from a DP with base distribution $\mathbb{G}_{0}$ and parameter $\alpha$. See \cite{gasthaus_2008} and \cite{gasthaus_thesis} for more details on this formulation. Note that difference between the formulation and indexing of the clusters in this model compared to the previous two models. This definition can be shown to be equivalent to the Bayesian mixture model defined in \eqref{bayesian_mixture_model}, when K is taken to be unbounded. This is the origin for references to this model as the infinite mixture model.

An alternative formultion of the DPM is known as the Chinese resturant process (CRP) sampling representation. If we let $K$ be the number of distinct mixture components in the above model, we can write the distinct mixture components as $\theta_{1}, \ldots, \theta_{K}$. Let $c_{1}, \ldots, c_{N}$ (where $c_{i} \in \{1, \ldots, K \}$) be class assignment variables that indicate the cluster to which observation $x_{i}$ is assigned. This allows the DPM to be represented by the CRP, a discrete-time stochastic process that defines a partition of the set $\{ x_{1}, \ldots, x_{N} \}$ (via the elements' assignments $c_{1}, \ldots, c_{N}$). The CRP allows samples to be drawn from the conditional distribution of the indictor variables $c_{i}$, and can be expressed as
\begin{align}
\begin{split}
	P(c_{i} = c_{j} \text{  for some  } j<i) &= \frac{m_{k}}{i-1+\alpha}\\
	P(c_{i} \neq c_{j} \text{  for all  } j<i) &= \frac{\alpha}{i-1+\alpha}
\end{split}
\end{align}
where $m_{k}$ is the cardinality of the set $\{ c_{j} | (c_{j}=c_{i}=k)  \wedge  (j < i) \}$ and $\alpha$ is the parameter of the DP prior on $\mathbb{G}$.

\subsection{Dependent Dirichlet Process Mixtures}

The goal of dependent Dirichlet process mixtures (DDPMs) is to allow modeling of data that is not independent and identically distributed (i.i.d) but instead has some underlying dependencies. For example, data generated during extraction procedures from videos have some associated temporal structure, since tracked objects display time dependent characteristics (in the features that denote position, as well as those which denote other appearance characteristics).

To account for the dependent behavior of data, research has been conducted on developing models involving a sequence of DPMs, where components of the mixtures are dependent upon (or may be considered `tied to') corresponding components at adjacent positions in the sequence. For example, if the data shows temporal dependence, the goal might be to create a sequence of DPMs, one for each time-step, where the components of the mixture at each step are dependent upon corresponding components in the both the following and previous time steps. Research towards models which accomplish such goals are described in (list citations).

More rigorously, we take the definition of a DDPM to be a stochastic process defined on the space of probability distributions over a domain, which are indexed by time, space, or a selection of other covariates in such a way that the marginal distribution at any point in the domain follows a Dirichlet process (adapted from definitions found in [cite griffin and steel and gasthaus-thesis]).

\subsubsection{Generalized Polya Urn Dependent Dirichlet Process Mixture}
\label{sec:gpudpm}

Data gained by performing extraction on videos containing objects is significantly time-dependent. Futhermore, the clusters of data, each of which represents an object in the video, might change in number as time progresses, since objects can enter and exit a scene. The specific DDPM used to model extraction data in this work while handling the above challenges is known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM), and was introduced by \cite{caron_2006} in 2006.

The GPUDDPM, when applied to data over T discrete time steps, can be viewed as a collection of DPMMs (one for each $t \in \{1, \ldots, T \}$), which are linked together by dependencies among their respective parameters. For example, the mixture component parameters at a given time step are dependent upon the mixture component parameters in adjacent time steps and the distribution and number of distinct cluster assignments at a given time step are dependent upon the same over a wider range of time steps, upon the distribution of observations' cluster assignments over this range, and upon a deletion procedure, described below.

To link together the parameters of mixture components in adjacent time-steps, a transition kernel $P(\phi_{k}^{t} | \phi_{k}^{t-1})$ is implemented, which provides a distribution over parameters of a mixture component in a given time-step (as a function of equivalent parameters in adjacent time-steps). One caveat is that each mixture component must be drawn independently from $\mathbb{G}_{0}$ (the base distribution of the DP, which acts as a prior distribution for the cluster parameters) which we achieve by making $\mathbb{G}_{0}$ the invariant distribution of $P(\phi_{k}^{t} | \phi_{k}^{t-1})$ (note that the transition kernel is a markov chain). Additionally, to account for varying numbers of clusters, there is a deletion procedure (described in \cite{caron_2007}) by which observations are considered "removed" from their assigned clusters at a given time-step (which modifies the value of $m_{k}$ defined above in the CRP representation). We introduce variables $d_{1}, \ldots, d_{N}$ which denote the times at which the assignments of given observations are considered to be removed. With these variables, we can now express the $m_{k}$ value at time $t$ as
\begin{equation}
m_{k,t} = \sum_{t' = 1}^{t} \mathbb{I}[(c_{t'}=k) \wedge (t < d_{t'})]
\end{equation}
where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise. Additionally, for an observation $x_{i}$ at a given time-step $t$, the deletion time $d_{i}$ can be defined to be $d_{i} = t + l_{i}$, where $l_{i}$ is considered the lifespan of an assignment, is distributed geometrically, and can be expressed as
\begin{equation}
l_{i} | \rho  \sim  \rho(1 - \rho)^{l_{i}}
\end{equation} \\
with parameter $\rho$. This process adheres to what \cite{caron_2007} calls a `uniform deletion strategy' over all the observations' assignments (since assignments to each cluster have equal probability of being deleted), though a more complex deletion strategy, dependent upon cluster size, can also be implemented.

Using the cluster size and observation deletion terms introduced above, we can define the GPUDDPM hierarchically as, for each time step $t = \{1, \ldots, T\}$
\begin{align}
\begin{split}
	x_{i}^{(t)} | c_{i}^{(t)}, \theta_{c_{i}^{(t)}}^{(t)} &\sim F(\theta_{c_{i}^{(t)}}^{(t)}) \\
	c_{i}^{(t)} | m^{(t-1)}, \alpha  &\sim  \text{CRP}(m^{(t-1)}, \alpha) \\
	m^{(t-1)} | m^{(t-2)}, c_{i}^{(t-1)}  &\sim \text{DEL}(m^{(t-2)}, c_{i}^{(t-1)}, \rho, \gamma) \\
	\theta_{k}^{(t)} | \theta_{k}^{(t-1)}   &\sim 
\begin{cases}
	P(\theta_{k}^{(t)} | \theta_{k}^{(t-1)}) \\
	\mathbb{G}_{0}
\end {cases}
\end{split}
\end{align}

The graphical corresponding with this formulation is shown in Figure 2.

\begin{figure}[h]
        \center{\includegraphics[width=120mm]{gpuddpm_gm_1.pdf}}
        \caption{\label{Figure 2:} Graphical Model of the Generalized Poly Urn Dependent Dirichlet Process Mixture}
\end{figure}

\subsubsection{Object Characteristics Modeled}

We would like make explicit the connection between certain elements of the GPUDDPM model and the characteristics of objects in vidoes that these elements represent.

First, the appearances of objects are moded by what we term the `likelihood and object appearance model', which we denoted by $F$. In the above modeling formulation, $F$ is the probability distribution for an observation, given that it is assigned to specified cluster (i.e. given that it is associated with a given object in a video). The likelihood and object appearance model is therefore

\begin{equation}
F(\theta_{c_{i}}) = P(x_{i}|\theta_{c_{i}})
\end{equation}

where $x_{i}$ is an observation, $c_{i}$ is its associated assignment, $\theta_{c_{i}}$ specifies the parameters of the parametric form of the $c_{i}^{th}$ cluster. The specific form of this model is dependent upon the observations $x_{i}$; recall that the observations, which as a baseline include spatial and temporal features, could include an arbitrary amount of additional features which need to be incorporated into this appearance model. A specific formulation of the appearance model can be seen in the proceeding section.

Object behavior is modeled by the dependencies between corresponding clusters at adjacent time steps. In particular, this relationship is captured by the transition kernel, $P(\theta_{k}^{t} | \theta_{k}^{t-1})$ (which captures the dependence of cluster $k$ at time $t$ on the same cluster at time $t-1$), described previously. The motion behavior of an object, i.e. the specific dependence of the parameters of cluster $k$ at a time $t$ on its parameters at time $t-1$, is therefore
\begin{equation}
Tr(\theta_{k, t} | \theta_{k, t-1}) = P(\theta_{k, t} | \theta_{k, t-1})
\end{equation}

Additionally, the base distribution $\mathbb{G}_{0}$ of the DP acts as a prior on the cluster parameters. This allows one to place a prior probability over appearances of objects. It is important that this prior is not chosen in such a specific manner that it limits this method from performing unsupervised detection of arbitrary objects.

\section{Model Specification and Inference}

In the following sections, a particular extraction method for generating data from videos and a specific formulation of the model described in Section~\ref{sec:modeloverview} are given. Inference schemes on this model, which when carried out allow for unsupervised detection and tracking of multiple arbitrary objects, are also described.

\subsection{Extraction}
\label{sec:modelspec_extraction}

We desire an extraction procedure that is as unsophisticated as possible, both to gauge the robustness of this method on potentially noisy extraction data and to ensure that the procedure is applicable to a wide range of videos. For these reasons we choose the method of frame-differencing for all extraction performed in this paper, which involves recording the positions of pixels (or pixel groups) that have exhibited differences in intensity or value in succesive frames beyond a given threshold. In particular, if we let $I_{t}$ be the image difference obtained by subtracting the value of the $(i,j)^{th}$ pixel in video frame $t$ from the value of the $(i,j)^{th}$ pixel in video frame $t+1$, we can define extraction on frame $t$ to be the process that returns the dataset
\begin{equation}
	\Omega_{t} = \{ (i,j) | I_{t}(i,j) > \phi \}
\end{equation}
where $i$ and $j$ respectively denote the two spatial positions of a pixel, and $\phi$ is a given pixel value threshold. Frame differencing on each frame $t =\{1, \ldots, T \}$ yields the dataset
\begin{equation}
	\Omega = \bigcup_{t=1}^{T} \Omega_{t} = \{ (i,j,t) | I_{t}(i,j) > \phi \}
\end{equation}
which is equivalent to data with the baseline features described in Section~\ref{sec:data}.

Frame differencing is unsophisticated, computationally inexpensive, able to be applied to a wide range of static, single-camera videos (note that videos used in the experiments described in the following sections were chosen to be static; moving-camera videos should be used in conjunction with applicable extraction methods that allow for camera movement). We have found that, when correctly implemented, frame differencing is sufficiently general to extract the desired worm-like data clouds from a variety of videos containing multiple moving objects. A few examples of this extraction procedure on single video frames can be found in Figure 3.

\begin{figure}
\centering
{\includegraphics[width=45mm]{antpic2.png}}
%\hspace{2mm}
{\includegraphics[width=60mm]{frame_diff_ants_1.png}}
\caption{(a.) Frame of a video containing five ants. (b.) Frame differencing extraction results from the same frame \willie{this is just a demo of the type of image i will have here... these aren't actually from the same frame, and also formatting is screwed up)}}
\label{test}
\end{figure}

We furthermore wish to capture the color (or grayscale) value of each pixel and incorporate this color information into our model. For each pixel $x = (i, j, t) \in \Omega$ (defined above), we specify a square, $L$ pixels in length, centered on $(i, j)$, that selects a set of pixels surrounding $\bold{x}$ in frame $t$. We also specify a scalar color value that is able to be computed for each pixel; this value could, for example, be some function of the r-g-b or h-s-v value of a pixel. The color value of each of the selected pixels surrounding $\bold{x}$ is recorded. Afterwards, the set of possible color values (ie the range of color values to which a pixel may be assigned) is partitioned into $V$ bins, and the number of pixels with a color value lying in each of the bins yields a $V$ dimensional vector of `color counts'. We will refer to this extraction technique as `color counting'.

\willie{Show example color-count vector (histogram) for different objects in an image. give more-formal definition like for frame differencing above.}

\subsection{Model Specification for Experiments}
%

\subsubsection{Data}

The frame differencing and color counting extraction outlined above yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \mathbb{Z}_{+}$, where each element $\bold{x} \in \bold{X}$ can be written
\begin{equation}
\bold{x} = ( \bold{x}^{s}, \bold{x}^{c}, t ) = ( x^{s_{1}}, x^{s_{2}}, x^{c_{1}}, \ldots, x^{c_{V}}, t )
\end{equation}
where $\bold{x}^{s} \in \mathbb{R}^{2}$ is the two dimensional vector of spatial coordinates,  $\bold{x}^{c} \in \mathbb{Z}_{+}^{V}$ is the $V$ discrete dimensional vector of color counts, and $t \in \mathbb{Z}_{+}$ is the discrete time index.

In the following experiments, the hue component of a pixel's h-s-v value is recorded from each pixel surrounding a given $\bold{x}^{s}$ in the manner described in Section~\ref{sec:modelspec_extraction} (where we choose $L=5$). Additionally, the set of possible hue values is partitioned into 10 bins, and the number of pixels with a hue value lying in each of the 10 bins is recorded to yield the vector of `color counts' $\bold{x}^{c} = x^{c_{1}}, \ldots, x^{c_{10}}$. Hue is chosen to represent object color since it has been demonstrated in previous work as a simple representation of object appearance that allows for distinct objects to be well differentiated (\willie{cite mckenna, other papers here}).

\subsubsection{Likelihood and Object Appearance Model}

At a given time $t$, we model each $\bold{x} \in \bold{X}$ as a draw from the product of a multivariate normal and multinomial distribution
\begin{equation}
P(\bold{x}|\theta) = \mathcal{N}(\bold{x}^{s} | \boldsymbol{\mu}, \Sigma)  \mathcal{M}n(\bold{x}^{c} | \bold{p})
\end{equation}
where $\theta = \{ \boldsymbol{\mu}, \Sigma, \bold{p} \}$ denote the parameters of a cluster at time $t$, with mean $\boldsymbol{\mu} \in \mathbb{R}^{2}$, covariance matrix $\Sigma \in \mathbb{R}^{2} \times \mathbb{R}^{2}$, and discrete probability vector $\bold{p} = (p_{1}, \ldots, p_{V})$ such that $\sum_{i=1}^{V}p_{i} = 1$. Also, $\mathcal{N}$ denotes the multivariate normal distribution and $\mathcal{M}n$ denotes the multinomial distribution,

The distribution families assumed to generate each $\bold{x}$ must be justified. Both the multivariate normal and multinomial distributions were chosen because they are sufficiently simple (and well studied) to allow for tractable inference and sufficiently flexible to provide a reasonable approximation to the data gained during extraction. In particular, the multivariate normal distribution over the spatial features $\bold{x}^{s}$ can be thought to represent the shape of each object as an oval; likewise, data generated by each moving object during extraction are often ovular--as noisy extraction procedures cause some smoothing of edges and corners, producing blob-like shapes even when objects are not particularly round--and centered on a given object. Furthermore, this model is justified as the maximum likelihood parameter estimate of a normal distribution corresponds to the least squares fit of data relative to the mean of the distribution. \willie{needs to be said more accurately and explained why this helps.} Modeling the color features $\bold{x}^{c}$ as draws from a multinomial distribution (equivalently, as draws from a product of discrete distributions), is justified since distinct object tend to generate pixels whose hue values are noisy but yield consistent counts in discrete hue bins.

\subsubsection{Base Distribution $\mathbb{G}_{0}$ and Appearance Prior}

$\mathbb{G}_{0}$ denotes the base distribution of the time-dependent Dirichlet process mixture; it also serves as a prior distribution for the parameters $\theta = \{ \boldsymbol{\mu}, \Sigma, \bold{p} \}$ present in the likelihood. We make use of conjugate priors in the base distribution to allow for more efficient computation. Specifically, in the experiments carried out in this paper, a normal-inverse-Wishart prior is placed on the multivariate normal parameters $\{ \boldsymbol{\mu}, \Sigma \}$ , and a Dirichlet prior is placed on the multinomial parameter $ \{  \bold{p}  \} $ (where the normal-inverse-Wishart is a conjugate prior for the multivariate normal component of the likelihood and the Dirichlet is a conjugate prior for the multinomial component). The prior can thus be written
\begin{equation}
\mathbb{G}_{0}(\theta) = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}, \Sigma | \boldsymbol{\mu}_{0}, k_{0}, v_{0}, \Lambda_{0})  \mathcal{D}ir( \bold{p} | \bold{q}_{0})
\end{equation}
where $\mathcal{N}i\mathcal{W}$ denotes the normal-inverse-Wishart distribution, $\mathcal{D}ir$ denotes the Dirichlet distribution, and the prior has the hyperparameters $\boldsymbol{\mu}_{0}, k_{0}, v_{0}, \Lambda_{0}$ and $\bold{q}_{0}$.  \willie{reference text with good background on NiW and Dir distributions}

\subsubsection{Transition Kernels and Motion Behavior}

We wish to formulate a transition kernel $P(\theta_{t} | \theta_{t-1})$ that provides a reasonable representation of how we expect tracked objects to behave. In the interest of formulating a model with the intention of using it to track arbitrary objects, we do not wish to make many assumptions about the long-term behavior of objects (though transition kernels that incorporate, for example, dynamics could potentially be implemented if one is aware beforehand of an object's behavior).

As described in Section~\ref{sec:gpudpm}, $\mathbb{G}_{0}$ must be the invariant distribution of $P(\theta_{t} | \theta_{t-1})$ in order for the the cluster parameters to remain marginally distributed according to the base distribution. In other words, the transition kernel must satisfy
\begin{equation}
\int \mathbb{G}_{0}(\theta_{t-1})P(\theta_{t} | \theta_{t-1}) d\theta_{t-1} = \mathbb{G}_{0}(\theta_{t})
\end{equation}
for a given cluster $\theta$. One way to achieve this is through the use of auxiliary variables. Auxiliary variables are a set of $M$ variables $\bold{z}_{t} = (z_{t,1}, \ldots, z_{t,M})$ associated with a each cluster $\theta$ at time $t$ that satisfy
\begin{eqnarray}
% P(\theta_{k,t} | \theta_{k,t-1}) = \int P(\theta_{k,t} | \bold{z}_{k,t}) P(\bold{z}_{k,t} | \theta_{k,t-1}) d \bold{z}_{k,t}
P(\theta_{t} | \theta_{t-1}) = \int P(\theta_{t} | \bold{z}_{t}) P(\bold{z}_{t} | \theta_{t-1}) d \bold{z}_{t}  %\\
% P(\theta', \bold{z}_{t}) = p(\bold{z}_{t}|\theta') \mathbb{G}_{0}(\theta')
\end{eqnarray}

In this way, the parameters of a cluster at a given time do not depend directly on their value at the previous time; they are instead dependent upon an intermediate sequence of auxiliary variables chosen to satisfy the above criteria, which allows the cluster parameters at each time step to be marginally distributed according to the base distribution $\mathbb{G}_{0}$.

For each cluster, we introduce $M$ auxiliary variables $z_{t, 1}, \ldots, z_{t, M}$ at time $t$ that are each drawn from the product of a multivariate normal and multinomial when conditioned on the associated cluster parameters $\theta_{t} = \{ \boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t} \}$:
\begin{equation}
z_{t, m} | \boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t}  \sim  \mathcal{N}(\boldsymbol{\mu}_{t}, \Sigma_{t}) \mathcal{M}n(\bold{p}_{t})   \hspace{15pt}   
\forall m \in \{ 1, \ldots, M \}
\end{equation}
To satisfy the above criteria for auxiliary variables, at each time $t$ we specify the dependencies of a given cluster on its associated set of auxiliary variables by
\begin{equation}
\boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t} | \bold{z}_t  \sim  \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}_{M}, k_{M}, v_{M}, \Lambda_{M})  \mathcal{D}ir(\bold{q}_{M})
\end{equation}
where $\boldsymbol{\mu}_{M}, k_{M}, v_{M}, \Lambda_{M},$ and $\bold{q}_{M}$ are parameters given by:
\willie{ran out of time. need to type up updates here.}

\subsection{Inference Implementations in Experiments}
%
Include a graphical model for each of the specified inference methods here? Or do I only need one graphical model for the model specified above?

Include algorithms for each of the inference procedures.

\section*{Experiments}

\subsection{Synthetic Videos}

\subsubsection{Occlusions}

\subsubsection{Similar Appearances}

\subsubsection{Different Numbers of Objects}

\subsubsection{Objects wih Diverse Characteristics}

\subsection{Real-life Videos}

\subsubsection{Complex Behavior}
%
Ants

\subsubsection{Occlusions}
%
Various people scenes

\subsubsection{Objects with Diverse Characteristics}
%
Traffic and Humans (PETS scenes)

\subsubsection{Standard Data for Comparison}
%
PETS and other standard datasets

\section*{Results and Discussion}

\section*{Conclusion}

% \begin{small}
% \bibliographystyle{plainnat}
% \bibliography{paper_refs} 
% \end{small}

\end{document}

### 2011-10-06 ltx/paper.tex
-The new paper update involves descriptions of performance metrics and experiments

@@
 
-Using the cluster size and observation deletion terms introduced above, we can define the GPUDDPM hierarchically as, for each time step $t = \{1, \ldots, T\}$
+Using the cluster size and observation deletion terms introduced above, we can define the GPUDDPM generatively as, for each time step $t = \{1, \ldots, T\}$
 \begin{align}
@@
 \begin{figure}[h]
-        \center{\includegraphics[width=120mm]{gpuddpm_gm_1.pdf}}
+        \center{\includegraphics[width=60mm]{gpuddpm_gm_1.pdf}}
         \caption{\label{Figure 2:} Graphical Model of the Generalized Poly Urn Dependent Dirichlet Process Mixture}
@@
 \section{Model Specification and Inference}
+\label{sec:modelspec}
 
@@
 
-\subsection{Model Specification for Experiments}
-%
+\subsection{Model Implementation in Experiments}
+
+The following sections provide details on the data and specific model formulation (determined from the general model formulation of Section~\ref{sec:modeloverview}) involved in the experiments described in Section~\ref{sec:experiments}.
 
@@
 
-The frame differencing and color counting extraction outlined above yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \mathbb{Z}_{+}$, where each element $\bold{x} \in \bold{X}$ can be written
+The frame differencing and color counting extraction outlined in Section~\ref{sec:modelspec_extraction} yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \mathbb{Z}_{+}$, where each element $\bold{x} \in \bold{X}$ can be written
 \begin{equation}
@@
 The prior can thus be written
-\begin{equation}
+\begin{equation} \label{basedistro}
 \mathbb{G}_{0}(\theta) = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}, \Sigma | \boldsymbol{\mu}_{0}, k_{0}, v_{0}, \Lambda_{0})  \mathcal{D}ir( \bold{p} | \bold{q}_{0})
@@
 For each cluster, we introduce $M$ auxiliary variables $z_{t, 1}, \ldots, z_{t, M}$ at time $t$ that are each drawn from the product of a multivariate normal and multinomial when conditioned on the associated cluster parameters $\theta_{t} = \{ \boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t} \}$:
-\begin{equation}
+\begin{equation} \label{transker1}
 z_{t, m} | \boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t}  \sim  \mathcal{N}(\boldsymbol{\mu}_{t}, \Sigma_{t}) \mathcal{M}n(\bold{p}_{t})   \hspace{15pt}   
@@
 To satisfy the above criteria for auxiliary variables, at each time $t$ we specify the dependencies of a given cluster on its associated set of auxiliary variables by
-\begin{equation}
+\begin{equation} \label{transker2}
 \boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t} | \bold{z}_t  \sim  \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}_{M}, k_{M}, v_{M}, \Lambda_{M})  \mathcal{D}ir(\bold{q}_{M})
@@
 
-\subsection{Inference Implementations in Experiments}
-%
-Include a graphical model for each of the specified inference methods here? Or do I only need one graphical model for the model specified above?
+\subsection{Inference Implementation in Experiments}
 
-Include algorithms for each of the inference procedures.
+This section provides background on the inference procedure applied to the model and extraction data, which allows for unsupervised detection and tracking of objects.
+A Markov Chain Monte Carlo (MCMC) sampler is used to perform inference, which makes use of the Metropolis-Hastings (M-H) algorithm to sample from a few distributions.
 
-\section*{Experiments}
+The generative model associated with the GPUDDPM is illustrated in Section~\ref{sec:gpudpm}.
+This representation includes cluster size $m$ (the number of observations assigned to a given cluster), which is standard in DPM constructions.
+To capture time-dependence, the GPUDDPM allows clusters to `lose assigned observations' in a deletion step, which modifies the sizes $m$ over time.
+Instead of incorporating cluster size variables $m$ directly, we can formulate an equivalent model which makes use of deletion variables $d$, which represent the time when observations are lost from a cluster .
+At each time step, cluster sizes can be recontructed from all previous assignment and deletion variables.
+The MCMC sampler described in this section operates in this alternate representation, which we can formulate as
+\begin{align}
+\begin{split}
+	 \bold{x}_{i} | c_{i}, \theta_{c_{i}} &\sim \mathcal{N}(\boldsymbol{\mu}_{c_{i}}, \Sigma_{c_{i}})  \mathcal{M}n(\bold{p}_{c_{i}}) \\
+	c_{i} | c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha  &\sim  \text{CRP}(c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha) \\
+	d_{i} | \rho  &\sim \text{Geo}(\rho) + i + 1 \\
+	\theta_{k, t} | \theta_{k, t-1}   &\sim 
+\begin{cases}
+	\text{TK} \hspace{12pt} \text{if } k \in \{ 1, \ldots, K_{t-1}  \} \\
+	\mathbb{G}_{0}  \hspace{15pt} \text{if } k = K_{t-1} + 1
+\end {cases}
+\end{split}
+\end{align}
+where TK stands for the transition kernel, given in \eqref{transker1} and \eqref{transker2}, and the base distribution $\mathbb{G}_{0}$ is given in \eqref{basedistro}.
+This formulation of the GPUDDPM is also used in \cite{gasthaus_thesis} and \cite{caron_2007}.
+The associated graphical model for this formulation is given in \willie{cite figure}.
 
-\subsection{Synthetic Videos}
+We define the state of the sampler to consist of the assignment variables for all observations, $c_{1}, \ldots, c_{N}$, the deletion variables for all observations, $d_{1}, \dots, d_{N}$, the sequence of cluster parameters for each active cluster $k$, $\theta_{k,1}, \ldots, \theta_{k, T}$, and the $M$ associated auxiliary variables for each cluster $\theta_{k,t}$, $z_{k,t,1}, \ldots, z_{k, t, M}$ (where $N$ denotes the number of observations and $T$ denotes the number of time steps).
 
-\subsubsection{Occlusions}
+The sampler moves sequentially through each of the $N$ observations, sampling the assignment $c_{i}$ and the deletion variable $d_{i}$ at each step.
+After the final observation at a given time $t$ step is sampled, the cluster parameters for all active clusters at $t$ are sampled, and a M-H step is employed to sample the $M$ auxiliary variables for each active cluster.
+How each sample is generated will be explicitely shown in the following sections.
 
-\subsubsection{Similar Appearances}
+\subsection{Modeling Background Noise}
+\willie{Do we think this is worth it? I can make extraction pretty clean on most datasets and this adds complication to an already complicated model}
 
-\subsubsection{Different Numbers of Objects}
+Noisy extraction procedures may result in erroneous data points that do not correspond to any foreground object.
+We would like to incorporate these points into our model.
+To do this, we modify our base distribution $\mathbb{G}_{0}$
 
-\subsubsection{Objects wih Diverse Characteristics}
+\section{Experiments}
+\label{sec:experiments}
 
-\subsection{Real-life Videos}
+The following sections provide details on performance evaluation metrics that have been developed for use in object detection and tracking studies (and that are used to quantify experimental results in this paper), synthetic video experiments that verify certain aspects of the technique presented in this paper, benchmark video experiments that demonstrate the performance of this technique in relation to other methods developed in recent years, and nonstandard video datasets to show the capability of this technique to perform unsupervised detection and tracking on a diverse range of object and video types.
 
-\subsubsection{Complex Behavior}
-%
-Ants
+\subsection{Performance Evaluation Metrics}
 
-\subsubsection{Occlusions}
-%
-Various people scenes
+Performance evaluation metrics, which intend to provide a standardized way of quantifying the success of a detection and tracking procedure on a given video, have started to become well studied and consistently used in the past four years.
+The metrics developed in \cite{kasturi_2008} and used in \cite{ellis_2010, taj_2007, lee_2009} have become well established metrics for evaluating the performance of object detection and tracking in videos and have been adopted by the Video Analysis and Content Extraction (VACE) program and the Classification of Events, Activities, and Relationships (CLEAR) consortium, two large projects concerned with video tracking.
+These metrics are used to quantify the experimental results in this study.
 
-\subsubsection{Objects with Diverse Characteristics}
-%
-Traffic and Humans (PETS scenes)
+The above metrics are dependent upon ground-truth data specifying the positions of each object in each frame throughout a video sequence.
+In the conducted experiments, we gained the synthetic video ground-truth during construction of the videos (described in Section~\ref{sec:syntheticvideos}), and we attempted to choose benchmark datasets with provided ground-truth data.
+However, the final benchmark dataset, in addition to all nonstandard video datasets, did not include ground-truth data; for these datasets, the Video Performance Evaluation Resource (ViPER) ground-truth tool \cite{doermann_2000} was used to gain ground-truth data.
 
-\subsubsection{Standard Data for Comparison}
-%
-PETS and other standard datasets
+\subsubsection{Mapping Ground-Truth Tracks to Output Tracks}
+
+The problem of finding a mapping between a video's ground-truth tracks and an algorithm's output tracks is nontrivial, though necesary to solve in order to compute the performance evaluation metrics used in this study.
+In short, the typical solution to this problem involves first specifying a performance metric and then choosing the mapping from ground-truth tracks to output tracks which yields the most favorable performance metric value.
+This process is described in detail by Kasturi et al.
+\cite{kasturi_2008}; we follow the method outlined in this paper to find an optimal mapping.
+Additionally, the method employed in \cite{kasturi_2008} allows erroneous and undetected tracks to be left unmapped, which is both desired and necessary in the case where there is a different number of ground-truth and output tracks.
+Note that once a mapping from a collection of ground-truth tracks to a collection of output tracks has been established, one can determine which output tracks are false positives (the output tracks to which no groundtruth track is assigned) and which input tracks are false negatives (the input tracks that are not assigned to an output track).
+The numbers of tracks displaying both of these failures are factors in the performance metrics used in this study.
+
+%, where a numerical search algorith--the Hungarian algorithm \cite{munkres_1957}--is used to find an optimal mapping from ground-truth to output tracks without requiring the performance metric value of all possible mapping combinations to be computed (reducing the time complexity from factorial to polynomial time)
+
+\subsubsection{MODA and MOTA}
+
+The two metrics used to quantify results in this study are refered to as the Multiple Object Detection Accuracy (MODA) and Multiple Object Tracking Accuracy (MOTA).
+These metrics, along with the the ground-truth tracks of all objects in the video sequence and the the mapping as defined in the previous section, can be used to evaluate detection and tracking accurancy of the video.
+For a given video frame $t$, we define MODA as
+\begin{equation}
+MODA = 1 - \frac{m^{t} + e^{t}}{N_{g}^{t}}
+\end{equation}
+where $m^{t}$ is the number of missed object detections (ie true-negative detections), $e^{t}$ is the number of erroneous object detections (ie false positive detections), and $N_{g}^{t}$ is the number of ground truth objects (all at frame $t$).
+
+For a sequence of video frames $t \in \{ 1, \ldots, T \}$ we define MOTA as
+\begin{equation}
+MOTA = 1 - \frac{\sum_{t=1}^{T} m^{t} + e^{t} + s_{t}}{\sum_{t=1}^{T} N_{g}^{t}}
+\end{equation}
+where $s$ denotes the number of instances at frame $t$ where an object's identity switches from a previous frame (ie where the object's mapping differs between frame $t$ and frame $t-1$),  and the other terms are defined the same as above.
+
+\subsection{Synthetic Video Datasets}
+\label{sec:syntheticvideos}
+
+Synthetic videos provide controlled scenarios in which aspects of the technique presented in this paper can be tested.
+Each of the following synthetic videos consists of a sequence of 200 images (each of size $500$ x $500$ pixels) containing a number of smaller colored squares of different (and potentially time-varying) sizes moving at varied speeds and trajectories over a black background.
+The synthetic videos contain instances of occlusion (where one or more objects are briefly hidden) and objects with time-varying appearances and behaviors, as these are examples of occurances that notoriously decrease the accuracy of detection and tracking.
+After each video was constructed, the extraction and inference procedures described in Section~\ref{sec:modelspec} were carried out to return a collection of centroid-position-sequences, each corresponding to the sequence of positions of a tracked object over each frame that it is present in the video.
+
+\subsubsection{Tracking by Appearance}
+
+The first synthetic video experiment aimed to test the ability of the model and inference procedure to maintain the identity of independent objects based on color information alone.
+Two videos were constructed, both containing a red square (rgb value $[255,0,0]$ and size $20$ x $20$ pixels) and a blue square (rgb value $[0,0,255]$ and size $20$ x $20$ pixels).
+In both videos, the squares begin at opposite sides of the scene at frame $f=1$ and travel towards each other, arriving at the same location at $f=100$ (where the blue square occludes the red square).
+The second half of the two videos differ in that both squares in the first video continue in the same direction and end at the other's starting positions at $f=200$, and both squares in the second video reverse directions and end at their initial starting positions at $f=200$.
+Frame difference extraction on each video yields a dataset with identical spatial features; hence, successful tracking depends fully on the reasonable incorporation of color information into the model.
+
+\subsubsection{Occlusion and Appearance Change}
+
+The second synthetic video experiment aimed to test tracking performance under occlusion, object appearance change, and motion change.
+A video was constructed, containing a red square (rgb value $[255,0,0]$), a green square (rgb value $[0,255,0]$), and a blue square (rgb value $[0,0,255]$).
+The red square was of size $20$ x $20$ pixels, the blue square was of size $15$ x $15$ pixels, and the green square began at size $50$ x $50$ pixels at frame $f=1$, linearly shrinks to $10$ x $10$ pixels at $f=100$, then linearly grows back to $50$ x $50$ pixels by the end of the video, $f=200$.
+Furthermore, the red and blue squares display the same behavior as in the second video of the first synthetic experiment (they begin at opposite sides of the scene traveling towards each other, cross at the center of the scene at $f=100$, and reverse direction, ending at their initial positions at $f=200$).
+The green square begins at a point equidistant from the other two squares, intersects with them as they overlap (causing the blue square to occlude the other two), and continues on in a direction at a 20 degree angle from its initial trajectory.
+
+% \subsubsection{Similar Appearances}
+
+% \subsubsection{Different Numbers of Objects}
+
+% \subsubsection{Objects wih Diverse Characteristics}
+
+\subsection{Benchmark Video Datasets}
+
+Benchmark video datasets for object tracking and detection have been produced to provide standard scenes on which researchers can compare detection and tracking results.
+These videos have been primarily produced for surveillance-related workshops--notably, for the International Workshop on Performance Evaluation of Tracking and Surveillance (PETS)--which provide a setting where researchers can focus on a collection of videos and computer vision goals.
+Four commonly used benchmark videos from four PETS workshops are
+
+\subsubsection{PETS2000}
+
+\subsubsection{PETS2001 or VS-PETS/Soccer}
+
+\subsubsection{PETS2004/Caviar}
+
+\subsubsection{PETS2009/2010}
+
+\subsection{Nonstandard Video Datasets}
+
+\subsubsection{Ants Videos}
+
+\subsubsection{}
 
@@
 
-% \begin{small}
-% \bibliographystyle{plainnat}
-% \bibliography{paper_refs} 
-% \end{small}
+\begin{small}
+\bibliographystyle{plainnat}
+\bibliography{paper_refs} 
+\end{small}
 

### 2011-10-13 ltx/paper.tex
added some results images to paper. left other parts of paper mostly alone.

@@
 \usepackage{amsfonts}
-\usepackage{subfigure}
 \usepackage{graphicx}
@@
 
-% \subsection{Related Work}
+\subsection{Related Work}
 
@@
 Futhermore, the clusters of data, each of which represents an object in the video, might change in number as time progresses, since objects can enter and exit a scene.
-The specific DDPM used to model extraction data in this work while handling the above challenges is known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM), and was introduced by \cite{caron_2006} in 2006.
+The specific DDPM used to model extraction data in this work while handling the above challenges is known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM), and was introduced by \cite{caron_2007} in 2006.
 
@@
 
+% \begin{figure}
+% \centering{
+% \includegraphics[width=45mm]{antpic2.png}}
+% %\hspace{2mm}
+% {\includegraphics[width=60mm]{frame_diff_ants_1.png}}
+% \caption{(a.) Frame of a video containing five ants.
+(b.) Frame differencing extraction results from the same frame \willie{this is just a demo of the type of image i will have here...
+these aren't actually from the same frame, and also formatting is screwed up)}}
+% \label{test}
+% \end{figure}
+
 \begin{figure}
   \centering               
-{\includegraphics[width=45mm]{antpic2.png}}
-%\hspace{2mm}
-{\includegraphics[width=60mm]{frame_diff_ants_1.png}}
+  \subfloat[]{\label{fig:ants_img}\includegraphics[width=0.5\textwidth]{antpic2.png}}
+  \subfloat[]{\label{fig:ants_img_framediff}\includegraphics[width=0.5\textwidth]{frame_diff_ants_1.png}}
   \caption{(a.) Frame of a video containing five ants.
@@
 these aren't actually from the same frame, and also formatting is screwed up)}}
-\label{test}
+  \label{fig:img_and_framediff}
 \end{figure}
@@
 	c_{i} | c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha  &\sim  \text{CRP}(c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha) \\
-	d_{i} | \rho  &\sim \text{Geo}(\rho) + i + 1 \\
-	\theta_{k, t} | \theta_{k, t-1}   &\sim 
+	d_{i}-i-1 | \rho  &\sim \text{Geo}(\rho)\\
+	\theta_{k, t} | \theta_{k, t-1} \willie{other-prior-params}  &\sim 
 \begin{cases}
 	\text{TK} \hspace{12pt} \text{if } k \in \{ 1, \ldots, K_{t-1}  \} \\
-	\mathbb{G}_{0}  \hspace{15pt} \text{if } k = K_{t-1} + 1
+	\mathbb{G}_{0}(\willie{prior params})  \hspace{15pt} \text{if } k = K_{t-1} + 1
 \end {cases}
@@
 \subsection{Modeling Background Noise}
-\willie{Do we think this is worth it? I can make extraction pretty clean on most datasets and this adds complication to an already complicated model}
 
@@
 We would like to incorporate these points into our model.
-To do this, we modify our base distribution $\mathbb{G}_{0}$
+To do this, we modify our base distribution $\mathbb{G}_{0}$ to be a two component mixture, where one component is the base distribution and the other is a multivariate normal distribution with a fixed `flat' covarance $\sigma = [100, 0; 0, 100]$ multiplied with an `even' multinomial distribution \willie{need more rigorus terms here} with a fixed parameter $\bold{p} = (1/K, \ldots, 1/K)$.
 
@@
 
+\begin{figure}
+  \centering               
+  \subfloat[]{\includegraphics[width=0.5\textwidth]{cross_final}}
+  \subfloat[]{\includegraphics[width=0.5\textwidth]{bounce_final}}
+  \caption{Each plot shows a sample from the posterior distribution of the state for one of the videos in synthetic experiment one, where the time axis ascends vertically, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.
+In both cases, the two objects are successfully tracked through occlusion, though the objects reverse direction in the second video.
+\willie{will show assignments by marker type in addition to color}}
+  \label{fig:synth_one_plot}
+\end{figure}
+
 \subsubsection{Occlusion and Appearance Change}
@@
 The green square begins at a point equidistant from the other two squares, intersects with them as they overlap (causing the blue square to occlude the other two), and continues on in a direction at a 20 degree angle from its initial trajectory.
+
+\begin{figure}
+  \centering               
+  \subfloat[]{\includegraphics[width=0.35\textwidth, angle=30]{synth2_img}}
+  \subfloat[]{\includegraphics[width=0.5\textwidth]{synth2_final}}
+  \caption{(a.) Video frame $f=10$ from synthetic experiment two.
+(b.) Sample from posterior of the state for synthetic experiment two, where color represents assignment and the mean and standard deviation are shown.
+The three objects of time varying size, speed, and position are tracked through simultaneous occlusion.
+\willie{will show assignments by marker type in addition to color}}
+  \label{fig:synth_two_plot}
+\end{figure}
 
@@
 
+%\subsubsection{PETS2000}
+
+\subsubsection{PETS2001}
+
+The PETS2001 dataset consists of
+
+\begin{figure}
+  \centering             
+  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2001_1}}
+  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2001_2}}
+  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2001_3}}
+  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2001_4}}
+  \caption{Four frames $f=25, 35, 45, 55$ from the PETS2001 sequence with one posterior sample mean and covarance matrix overlayed. \willie{will show path trail over prev 10 frames}}
+  \label{fig:pets2001_imgs}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\subfloat[]{\includegraphics[width=0.52\textwidth]{pets2001_results_1}}
+	\subfloat[]{\includegraphics[width=0.48\textwidth]{pets2001_results_2}}
+	\caption{Sample from posterior of the state for the PETS2001 dataset, where color represents assignment and the mean and standard deviation are shown.
+\willie{will show assignments by marker type in addition to color}}
+	\label{fig:pets2001_results}
+\end{figure}
+
+%\subsubsection{PETS2004/Caviar}
+
+\subsubsection{PETS2009/2010}
+
+The PETS2009/2010 dataset consists of
+
+\begin{figure}
+  \centering             
+  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2009_1}}
+  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2009_2}}
+  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2009_3}}
+  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2009_4}}
+  \caption{Four frames $f=10, 25, 35, 75$ from the PETS2009 sequence with one posterior sample mean and covarance matrix overlayed. \willie{will show path trail over prev 10 frames}}
+  \label{fig:pets2009_imgs}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=3in]{pets2009_results}
+	% \subfloat[]{\includegraphics[width=0.52\textwidth]{pets2001_results_1}}
+	% \subfloat[]{\includegraphics[width=0.48\textwidth]{pets2001_results_2}}
+	\caption{Sample from posterior of the state for the PETS2009 dataset, where color represents assignment and the mean and standard deviation are shown.
+\willie{will show assignments by marker type in addition to color}}
+	\label{fig:pets2009_results}
+\end{figure}
+
 \subsubsection{PETS2000}
 
-\subsubsection{PETS2001 or VS-PETS/Soccer}
+The PETS2000 dataset consists of
 
-\subsubsection{PETS2004/Caviar}
+\begin{figure}
+	\centering
+	\subfloat[]{\includegraphics[width=0.5\textwidth]{pets2000_results_1}}
+	\subfloat[]{\includegraphics[width=0.5\textwidth]{pets2000_results_2}}
+	\caption{\willie{This dataset has not done that well so far. Not sure why. Here are two typical cases.}}
+	\label{fig:pets2000_results}
+\end{figure}
 
-\subsubsection{PETS2009/2010}
+\subsubsection{PETS2004/Caviar Datasets}
+
+The Caviar datasets used in PETS2004 consist of multiple test-case-scenarios involving one or more humans.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=3in]{pets2004_results_1}
+	\caption{Sample from posterior of the mean for frames $f = 1-20$ overlayed on frame $f=20$.}
+	\label{fig:pets2009_results}
+\end{figure}
 

### 2011-11-17 ltx/paper.tex
changed structure of directory, updated paper, added more images, graphical models look good now, etc.

@@
 \section{Introduction}
-
-\frank{this section should introduce our work and situate it relative to the literature}
-\willie{okeydoke. this isn't ready yet. to avoid confusion I've just commented the whole section out.}
-
 % % \subsection{Topic Overview and Terminology}
 
-% This work is related to the study of algorithms that perform automated detection and tracking of arbitrary objects in videos; we define such algorithms as those with the ability to decide which spatiotemporal regions of a video constitute objects (extraction), find the positions and/or shapes of distinct objects (localization), and maintain the positions of the detected objects over time (tracking).
-Robust algorithms which can carry out these tasks are applicable to problems in the fields of surveillance, autonomous systems and robotics, and video summarization and compression.
-Approaches to these tasks have been proposed under different headings and in a variety of contexts.
-This section serves to introduce relevant terminology and topics of research.
+We define the task of automated detection and tracking of arbitrary objects in videos to be the fulfillment of three actions: deciding which spatiotemporal regions of a video constitute objects (`extraction'), finding the positions and/or shapes of distinct objects (`localization'), and maintaining the positions of the detected objects over time (`tracking').
+Robust algorithms capable of carrying out these tasks are applicable to problems in the fields of surveillance, autonomous systems and robotics, and video summarization and compression.
+Approaches to these tasks have been proposed under different headings and in a variety of contexts; this paper serves to introduce a new technique involving the application of a dependent Dirichlet processes mixture model, which we hope will provide the foundation for a number of robust, unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.
 
-% The task of extraction is closely related to the topics of foregound segmentation, background modeling, motion detection, and appearance feature extraction.
-In general, these methods aim to extract features from regions of a video frame that are considered objects or are distinguished in some way from the background of a scene.
-Areas of research in this field include attempts to handle situations where the background displays motion or other dynamic characteristics, reduce noise caused by the false extraction of backgrounds, handle abberant image effects present in some videos, and operate accurately on scenes involving illumination changes.
-
-% The task of localization is closely related to the topics of object detection, position-finding, target segmentation, contour-detection, template matching, and recognition.
-These methods are mainly concerned with finding the positions and/or shapes of objects in a given video frame, or with discerning between objects similar in position or appearance.
-There has been a focus in this field on developing techniques for representing an object's appearance, both to allow for localization of objects on a per-frame basis and to allow for consistent localization of a given object over time.
-Research into the unsupervised detection of arbitrary objects (which we define as methods that accomplish both extraction and localization) has included work involving image segmentation, parsing, and pixel clustering.
-These topics usually relate to the automated partitioning of images into meaningful regions.
-Note that the popular phrase `object detection' is also typically used to refer to procedures that carry out extraction and localization; however, it is often reserved to denote methods that are tailored to find the position of specified objects or object types (as opposed to arbitrary object detection), or to denote supervised methods, where extraction is peformed implicitely during searches for specified objects and localization procedures.
-
-% The task of tracking is closely related to the topics of video tracking, data association, filtering, point matching, and `tracking of non-rigid objects'.
-In general, these tasks aim to maintain localization of the positions and/or shapes of objects over time.
-Areas of interest in this field involve developing metrics for judging the similarity between two object representations (for example, potential representations of the same object at adjacent frames in a video), techniques that provide a way to conduct efficient searches over a frame for regions that are similar to a given object, and methods for predicting the future position or appearance of an object.
-A great deal of research in this field over the past decade has focused on the development of algorithms to track multiple objects simultaneously.
-There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera), complex object interactions, object coalescence (where one object is tracked by multiple times/redundantly), objects with similar appearances, variable (and potentially high) numbers of objects, and objects that enter and exit a field of view at different times.
-
-% There does not yet exist a major body of research concerned primarily with methods for combining extraction, localization, and tracking to perform cohesive unsupervised detection of arbitrary objects in videos.
+There does not yet exist a major body of research concerned primarily with methods for combining extraction, localization, and tracking to perform cohesive unsupervised detection of arbitrary objects in videos.
 Many papers that attempt to accomplish this goal focus on one of extraction, localization, or tracking, and attempt to integrate in methods from the other two tasks.
 However, a few key schemes for unsupervised detection and tracking do indeed exist.
-A widely applied, though unsophisticated, technique to carry out such a procedure is called blob tracking.
-Blob tracking typically refers to methods that perform extraction and localization at each frame independently--for example, by assuming that object pixels have a different intensity than background pixels, setting a fixed pixel intensity threshold for extraction of object pixel locations, and segmenting the extracted data into `blobs' to accomplish localization.
-Afterwards, tracking is often carried out by finding the centroid of each blob and applying standard multiple target tracking algorithms (described in [section]).
-Note that blob tracking carries out the three main tasks described above, so it satisfies the criteria for the type of unsupervised detection and tracking algorithms studied in this work; there are, however, many challenges associated with blob tracking that render it impractical in many settings (detailed in [section]).
+A widely applied, though unsophisticated, example is blob tracking.
+Blob tracking typically refers to methods that perform extraction and localization at each frame independently--for example, by assuming that object pixels have a different intensity than background pixels, setting a fixed pixel intensity threshold to extract pixel locations, and segmenting the extracted data into `blobs' to accomplish localization.
+A number of tracking techniques based on finding blobs in consecutive frames of a video with similar appearances have been attempted to maintain the identity of detected blobs over time.
+The main challenges, however, associated with blob tracking involve the difficulty of segmenting occluded objects in a given frame \cite{zhao2004tracking} and preventing multiple tracked blobs from collapsing onto and tracking a single object \cite{vermaak_2003}.
+These difficulties stem from the typical blob filter's need to perform detection via segmentation on individual frames and lack of a principled way to resist the collapse of independent tracks.
 
-% This paper presents a new method for unsupervised multiple object detection and tracking.
-We introduce, and provide details on the theory behind, a technique involving dependent Dirichlet process mixture models.
-Specifically, this work applies the model to data gained during the extraction process; note that data of this type is refered to at different points in the paper as `extraction data', `data points', 'points', or 'observations'.
-The model is first presented in an abstract form, which could be adapted to a wide range of extraction procedures and object appearance representations (mainly left for future studies).
+%Note that blob tracking carries out the three main tasks described above, so it satisfies the criteria for the type of unsupervised detection and tracking algorithms studied in this work;
+
+This paper presents the use of Dirichlet process mixture models as a technique for unsupervised multiple object detection and tracking, which we intend as a method to allow for robust object detection (particularly in cases where frame-by-frame segmentation would be difficult) and to keep the tracks of independent objects isolated.
+Specifically, this work applies a type of Bayesian nonparametric model to data gained during extraction; note that data of this type is refered to in the paper as `extraction data', `data points', or `observations'.
+The model is first presented in an abstract form, which could be adapted to a wide range of extraction procedures and object appearance representations (mainly left for exploration in future studies).
 We then describe inference algorithms that allow object localization and tracking to be carried out.
@@
 
-% Related work can be broken down into studies relating to extraction procedures, unsupervised localization procedures, single object tracking methods, multiple object tracking methods, supervised multiple object detection and tracking methods, and unsupervised multiple object detection and tracking methods.
+Related work can be broken down into studies relating to extraction procedures, localization techniques, single and multiple object tracking methods, and combination methods for joint object detection and tracking.
 
-% Extraction procedures can usually be classified as either model-based methods or heuristic-based methods.
+The task of extraction incorporates the tasks of foregound segmentation, background modeling, motion detection, and appearance feature extraction.
+In general, these methods aim to extract features from regions of a video frame that are considered objects or are distinguished in some way from the background of a scene.
+Areas of research in this field include attempts to handle situations where the background displays motion or other dynamic characteristics, reduce noise caused by the false extraction of backgrounds, handle abberant image effects present in some videos, and operate accurately on scenes involving illumination changes.
+These procedures can usually be classified as either model-based methods or heuristic-based methods.
 Model-based methods emphasize the detection of shapes of regions containing objects accurately and reducing background noise.
@@
 
-% On the other hand, heuristic-based methods perform extracting by looking for temporal changes in pixel value characteristics.
+On the other hand, heuristic-based methods perform extracting by looking for temporal changes in pixel value characteristics.
 While these methods locate regions with objects less accurately than model-based methods, they are often much less computationally expensive.
@@
 
-% There are an extensive collection of methods for tracking single, non-rigid, objects.
+The task of localization is closely related to the topics of object detection, position-finding, target segmentation, contour-detection, template matching, and recognition.
+These methods are mainly concerned with finding the positions and/or shapes of objects in a given video frame, or with discerning between objects similar in position or appearance.
+There has been a focus in this field on developing techniques for representing an object's appearance, both to allow for localization of objects on a per-frame basis and to allow for consistent localization of a given object over time.
+Research into the unsupervised detection of arbitrary objects (which we define as methods that accomplish both extraction and localization) has included work involving image segmentation, parsing, and pixel clustering.
+These topics usually relate to the automated partitioning of images into meaningful regions.
+Note that the popular phrase `object detection' is also typically used to refer to procedures that carry out extraction and localization; however, it is often reserved to denote methods that are tailored to find the position of specified objects or object types (as opposed to arbitrary object detection), or to denote supervised methods, where extraction is peformed implicitely during searches for specified objects and localization procedures.
+
+The task of tracking incorporates the topics of video tracking, data association, filtering, point matching, and `tracking of non-rigid objects'.
+In general, these tasks aim to maintain localization of the positions and/or shapes of objects over time.
+Areas of interest in this field involve developing metrics for judging the similarity between two object representations (for example, potential representations of the same object at adjacent frames in a video), techniques that provide a way to conduct efficient searches over a frame for regions that are similar to a given object, and methods for predicting the future position or appearance of an object.
+A great deal of research in this field over the past decade has focused on the development of algorithms to track multiple objects simultaneously.
+There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera), complex object interactions, object coalescence (where one object is tracked by multiple times/redundantly), objects with similar appearances, variable (and potentially high) numbers of objects, and objects that enter and exit a field of view at different times.
+
+There are an extensive collection of methods for tracking single, non-rigid, objects.
 For example, color has long been used for tracking.
@@
 Additionally, color has been successfully applied as the feature in the mean-shift procedure (kernel-based method) for tracking \citep{comaniciu_2003, perez_2002, nummiaro_2003, lee_2011}.
+
+%
+
+% Extraction
+
+%
 
@@
 This section provides details on the nature of the data to be modeled, Dirichlet process mixture (DPM) models, and dependent Dirichlet process mixture (DDPM) models.
-We end by making explicit the connection between elements of the model and object characteristics that these elements represent (for example, the appearance and motion behavior of objects are specifically captured).
-The provided model is formulated here in a sufficently abstract manner to allow for various representations of objects to be specified in future studies.
+We end by making explicit the way in which characteristics of video objects can be represented by elements of these models.
+We formulate the model in a sufficently abstract manner to allow for specific implementations tailored towards specified objects in future studies.
 
@@
 These data are distributed roughly as worm-like shapes that follow the paths of individual objects as they move, when they are plotted in three dimensions (where we let the frame number, our discrete representation of time, ascend the vertical axis and the spatial positions occupy the horizontal axes).
-An example of data with such a distribution can been seen in Figure 1.
+An example of data with such a distribution can been seen in Figure \ref{fig:baselinedata}.
 The data consists of the three-dimensional baseline features and was gained by extraction peformed on a 50 frame sequence showing five ants moving across the video scene.
 \begin{figure}[h]
-        \center{\includegraphics[width=75mm]{paths.png}}
-        \caption{\label{Figure 1:} Baseline three-dimensional extraction data gained from a video of multiple moving targets.
+        \center{\includegraphics[width=85mm]{../img/usbh_data_2.png}}
+        \caption{\label{fig:baselinedata} Baseline three-dimensional extraction data gained from a video segment containing 5 moving ants.
 The vertical axis denotes the frame number and the horizontal axes denote $d_{1}$ and $d_{2}$ spatial coordinates, respectively, of pixels within a given frame.
-The data comes from a 50 frame video segment showing 5 moving ants (note that one ant enters the scene near the middle of the segment).}
+Markers are colored by position for ease of viewing.}
 \end{figure}
@@
 Using the cluster size and observation deletion terms introduced above, we can define the GPUDDPM generatively as, for each time step $t = \{1, \ldots, T\}$
+% \begin{align}
+% \begin{split}
+% 	x_{i}^{(t)} | c_{i}^{(t)}, \theta_{c_{i}^{(t)}}^{(t)} &\sim F(\theta_{c_{i}^{(t)}}^{(t)}) \\
+% 	c_{i}^{(t)} | m^{(t-1)}, \alpha  &\sim  \text{CRP}(m^{(t-1)}, \alpha) \\
+% 	m^{(t-1)} | m^{(t-2)}, c_{i}^{(t-1)}  &\sim \text{DEL}(m^{(t-2)}, c_{i}^{(t-1)}, \rho, \gamma) \\
+% 	\theta_{k}^{(t)} | \theta_{k}^{(t-1)}   &\sim 
+% \begin{cases}
+% 	P(\theta_{k}^{(t)} | \theta_{k}^{(t-1)}) \\
+% 	\mathbb{G}_{0}
+% \end{cases}
+% \end{split}
+% \end{align}
 \begin{align}
 \begin{split}
-	x_{i}^{(t)} | c_{i}^{(t)}, \theta_{c_{i}^{(t)}}^{(t)} &\sim F(\theta_{c_{i}^{(t)}}^{(t)}) \\
-	c_{i}^{(t)} | m^{(t-1)}, \alpha  &\sim  \text{CRP}(m^{(t-1)}, \alpha) \\
-	m^{(t-1)} | m^{(t-2)}, c_{i}^{(t-1)}  &\sim \text{DEL}(m^{(t-2)}, c_{i}^{(t-1)}, \rho, \gamma) \\
-	\theta_{k}^{(t)} | \theta_{k}^{(t-1)}   &\sim 
+	\bold{x}_{t}^{i} | c_{t}^{i}, \theta_{c_{t}^{i}, t} &\sim F(\theta_{c_{t}^{i}, t}) \\
+	c_{t}^{i} | m_{c_{t}^{i}, t-1}, \alpha  &\sim  \text{CRP}(m_{c_{t}^{i}, t-1}, \alpha) \\
+	m_{c_{t-1}^{i}, t-1} | m_{c_{t-1}^{i}, t-2}, c_{t-1}^{i}  &\sim \text{DEL}(m_{c_{t-1}^{i}, t-2}, c_{t-1}^{i}, \rho, \gamma) \\
+	\theta_{k, t} | \theta_{k, t-1}   &\sim 
 \begin{cases}
-	P(\theta_{k}^{(t)} | \theta_{k}^{(t-1)}) \\
+	P(\theta_{k, t} | \theta_{k, t-1}) \\
 	\mathbb{G}_{0}
-\end {cases}
+\end{cases}
 \end{split}
@@
 
-The graphical corresponding with this formulation is shown in Figure 2.
+The graphical model corresponding with this formulation is shown in Figure \ref{fig:gpuddpm_gm_1}.
 
 \begin{figure}[h]
-        \center{\includegraphics[width=60mm]{gpuddpm_gm_1.pdf}}
-        \caption{\label{Figure 2:} Graphical Model of the Generalized Poly Urn Dependent Dirichlet Process Mixture}
+        \center{\includegraphics[width=100mm]{../img/gpuddp_gm_1.pdf}}
+        \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Poly Urn Dependent Dirichlet Process Mixture}
 \end{figure}
@@
 We have found that, when correctly implemented, frame differencing is sufficiently general to extract the desired worm-like data clouds from a variety of videos containing multiple moving objects.
-A few examples of this extraction procedure on single video frames can be found in Figure 3.
+A few examples of this extraction procedure on pairs of consecutive video frames can be found in Figure \ref{fig:img_and_framediff}.
 
@@
 % \centering{
-% \includegraphics[width=45mm]{antpic2.png}}
+% \includegraphics[width=45mm]{../img/antpic2.png}}
 % %\hspace{2mm}
-% {\includegraphics[width=60mm]{frame_diff_ants_1.png}}
+% {\includegraphics[width=60mm]{../img/frame_diff_ants_1.png}}
 % \caption{(a.) Frame of a video containing five ants.
@@
   \centering
-  \subfloat[]{\label{fig:ants_img}\includegraphics[width=0.5\textwidth]{antpic2.png}}
-  \subfloat[]{\label{fig:ants_img_framediff}\includegraphics[width=0.5\textwidth]{frame_diff_ants_1.png}}
-  \caption{(a.) Frame of a video containing five ants.
-(b.) Frame differencing extraction results from the same frame \willie{this is just a demo of the type of image i will have here...
-these aren't actually from the same frame, and also formatting is screwed up)}}
+  \subfloat[]{\label{fig:ants_img}\includegraphics[width=0.335\textwidth]{../img/usbh_frame_723.pdf}} 
+  \subfloat[]{\label{fig:ants_img}\includegraphics[width=0.335\textwidth]{../img/usbh_frame_724.pdf}}
+  \subfloat[]{\label{fig:ants_img_framediff}\includegraphics[width=0.32\textwidth]{../img/usbh_framediff.png}}
+  \caption{(a., b.) Consecutive frames of a video containing four ants. (c.) Extraction results produced by taking the pixel-wise difference between these frames}
   \label{fig:img_and_framediff}
@@
 	c_{i} | c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha  &\sim  \text{CRP}(c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha) \\
-	d_{i}-i-1 | \rho  &\sim \text{Geo}(\rho)\\
+	d_{i} | \rho  &\sim \text{Geo}(\rho) + i + 1\\
 	\theta_{k, t} | \theta_{k, t-1} \willie{other-prior-params}  &\sim 
@@
 	\mathbb{G}_{0}(\willie{prior params})  \hspace{15pt} \text{if } k = K_{t-1} + 1
-\end {cases}
+\end{cases}
 \end{split}
@@
 This formulation of the GPUDDPM is also used in \cite{gasthaus_thesis} and \cite{caron_2007}.
-The associated graphical model for this formulation is given in \willie{cite figure}.
+The associated graphical model for this formulation is given in \ref{fig:gpuddpm_gm_2}.
+\begin{figure}[h]
+        \center{\includegraphics[width=100mm]{../img/gpuddp_gm_3.pdf}}
+        \caption{\label{fig:gpuddpm_gm_2} Graphical Model of the Generalized Poly Urn Dependent Dirichlet Process Mixture}
+\end{figure}
 
@@
 
-The sampler moves sequentially through each of the $N$ observations, sampling the assignment $c_{i}$ and the deletion variable $d_{i}$ at each step.
+The sampler operates by iteratively sampling each member of the state given all other members.
+In particular, it moves sequentially through each of the $N$ observations, sampling the assignment $c_{i}$ and the deletion variable $d_{i}$ at each step.
 After the final observation at a given time $t$ step is sampled, the cluster parameters for all active clusters at $t$ are sampled, and a M-H step is employed to sample the $M$ auxiliary variables for each active cluster.
-How each sample is generated will be explicitely shown in the following sections.
+The following sections detail how each sample is generated will be explicitely shown in the following sections.
+
+\subsubsection{Sampling Assignment Variables $c_{i}$}
+The assignment of each observation, $c_{i}$, given all other members of the state,
+
+\subsubsection{Sampling Deletion Variables $d_{i}$}
+
+\subsubsection{Sampling Cluster Parameters $\theta_{k, t}$}
+
+\subsubsection{Sampling Auxiliary Variables $z_{k,t,m}$}
 
@@
 Performance evaluation metrics, which intend to provide a standardized way of quantifying the success of a detection and tracking procedure on a given video, have started to become well studied and consistently used in the past four years.
-The metrics developed in \cite{kasturi_2008} and used in \cite{ellis_2010, taj_2007, lee_2009} have become well established metrics for evaluating the performance of object detection and tracking in videos and have been adopted by the Video Analysis and Content Extraction (VACE) program and the Classification of Events, Activities, and Relationships (CLEAR) consortium, two large projects concerned with video tracking.
+The metrics developed in \cite{kasturi_2008} and used in \cite{ellis_2010, taj_2007, lee_2009} have become well established metrics for evaluating the performance of object detection and tracking in videos and have been adopted by the Video Analysis and Content Extraction (VACE) program and the Classification of Events, Activities, and Relationships (CLEAR) consortium, two large-scale efforts concerned with video tracking and interaction analysis.
 These metrics are used to quantify the experimental results in this study.
@@
 The above metrics are dependent upon ground-truth data specifying the positions of each object in each frame throughout a video sequence.
-In the conducted experiments, we gained the synthetic video ground-truth during construction of the videos (described in Section~\ref{sec:syntheticvideos}), and we attempted to choose benchmark datasets with provided ground-truth data.
-However, the final benchmark dataset, in addition to all nonstandard video datasets, did not include ground-truth data; for these datasets, the Video Performance Evaluation Resource (ViPER) ground-truth tool \cite{doermann_2000} was used to gain ground-truth data.
+In the experiments described below, we gained the synthetic video ground-truth during construction of the videos (described in Section~\ref{sec:syntheticvideos}), and we used the Video Performance Evaluation Resource (ViPER) ground-truth software \cite{doermann_2000}, an open source tool commonly used in the video tracking community, to author ground-truth data for each of the benchmark datasets.
 
-\subsubsection{Mapping Ground-Truth Tracks to Output Tracks}
+\subsubsection{Mapping Ground-Truth to Output}
+\label{sec:mappingtoground}
 
@@
 \cite{kasturi_2008}; we follow the method outlined in this paper to find an optimal mapping.
+Similiar to descriptions in this paper, we implement the Hungarian algorithm \cite{munkres_1957} as a polynomial-time ($O(n^{3})$) solution to the problem of optimally mapping two sets of tracks once the similarity between any two tracks given some specified metric is established.
 Additionally, the method employed in \cite{kasturi_2008} allows erroneous and undetected tracks to be left unmapped, which is both desired and necessary in the case where there is a different number of ground-truth and output tracks.
-Note that once a mapping from a collection of ground-truth tracks to a collection of output tracks has been established, one can determine which output tracks are false positives (the output tracks to which no groundtruth track is assigned) and which input tracks are false negatives (the input tracks that are not assigned to an output track).
+Note that once a mapping from a collection of ground-truth tracks to a collection of output tracks has been established, one can determine which output tracks are false positives (the output tracks to which no ground-truth track is assigned) and which input tracks are false negatives (the input tracks that are not assigned to an output track).
 The numbers of tracks displaying both of these failures are factors in the performance metrics used in this study.
@@
 
-\subsubsection{MODA and MOTA}
+% \subsubsection{MODA and MOTA}
 
-The two metrics used to quantify results in this study are refered to as the Multiple Object Detection Accuracy (MODA) and Multiple Object Tracking Accuracy (MOTA).
+% The two metrics used to quantify results in this study are refered to as the Multiple Object Detection Accuracy (MODA) and Multiple Object Tracking Accuracy (MOTA).
 These metrics, along with the the ground-truth tracks of all objects in the video sequence and the the mapping as defined in the previous section, can be used to evaluate detection and tracking accurancy of the video.
 For a given video frame $t$, we define MODA as
+% \begin{equation}
+% MODA = 1 - \frac{m^{t} + e^{t}}{N_{g}^{t}}
+% \end{equation}
+% where $m^{t}$ is the number of missed object detections (ie true-negative detections), $e^{t}$ is the number of erroneous object detections (ie false positive detections), and $N_{g}^{t}$ is the number of ground truth objects (all at frame $t$).
+
+% For a sequence of video frames $t \in \{ 1, \ldots, T \}$ we define MOTA as
+% \begin{equation}
+% MOTA = 1 - \frac{\sum_{t=1}^{T} m^{t} + e^{t} + s_{t}}{\sum_{t=1}^{T} N_{g}^{t}}
+% \end{equation}
+% where $s$ denotes the number of instances at frame $t$ where an object's identity switches from a previous frame (ie where the object's mapping differs between frame $t$ and frame $t-1$),  and the other terms are defined the same as above.
+
+\subsubsection{SFDA and ATA}
+
+The two metrics used to quantify performance in this study are known as the the Sequence Frame Detection Accuracy (SFDA) and the Average Tracking Accuracy (ATA).
+These metrics were developed during VACE Phase II to provide a single, comprehensive metric to describe detection, and one to describe tracking.
+
+The following terms are used in the definitions of the performance metrics:
+\begin{itemize}
+\item $G_{i}$ denotes the spatiotemporal region occupied by the $i$th ground-truth object in a video, and $G_{i}^{(t)}$ denotes the region occupied by the $i$th ground-truth object in frame $t$.
+\item $D_{i}$ denotes the spatiotemporal region occupied by the $i$th detected object in a video, and $D_{i}^{(t)}$ denotes the region occupied by the $i$th detected object in frame $t$.
+\item $N_{G}$ denotes the total number of unique ground-truth objects in a video, and $N_{G}^{(t)}$ denotes the number of unique ground-truth objects present at frame $t$.
+\item $N_{D}$ denotes the total number of unique detected objects in a video, and $N_{D}^{(t)}$ denotes the number of unique detected objects present at frame $t$.
+\item $N_{\text{frames}}$ denotes the total number of frames in a video, and $N_{\text{frames}}^{(i)}$ denotes the number of frames in which an object $i$ (which can be a ground-truth or detected object, depending on the context) is present in a video.
+\item $N_{\text{mapped}}$ denotes the number of mapped ground-truth/detect pairs in a video, and $N_{\text{mapped}}^{(t)}$ denotes the number of mapped ground-truth/detect pairs present at frame $t$.
+\end{itemize}
+
+The SFDA metric quantifies the performance of an object detection algorithm as a function of the number of correct detects, false positive detects, missed (true negative) detects, and spatial allignment of detects relative to the ground-truth.
+The SFDA is calculated by computing the Frame Detection Accuracy at frame $t$ ($\text{FDA}^{(t)}$) for each frame in a video sequence.
+The FDA provides a measure of the allignment between ground-truth and detected objects in a given frame via the overlap ratio of a ground-truth/detect pair, defined to be the ratio of the intersection of ground-truth and detect regions to the union of ground-truth and detect regions.
+Formally, we can write
 \begin{equation}
-MODA = 1 - \frac{m^{t} + e^{t}}{N_{g}^{t}}
+\text{FDA}^{(t)} = \frac{\text{Overlap Ratio}}{\left(\frac{N_{G}^{(t)} + N_{D}^{(t)}}{2}\right)}
 \end{equation}
-where $m^{t}$ is the number of missed object detections (ie true-negative detections), $e^{t}$ is the number of erroneous object detections (ie false positive detections), and $N_{g}^{t}$ is the number of ground truth objects (all at frame $t$).
+where
+\begin{equation}
+\text{Overlap Ratio} = \sum_{i = 1}^{N_{\text{mapped}}^{(t)}} \frac{\left|G_{i}^{(t)} \cap D_{i}^{(t)}\right|}{\left|G_{i}^{(t)} \cup D_{i}^{(t)}\right|}
+\end{equation}
+The term $N_{\text{mapped}}^{(t)}$ refers to an optimal mapping between ground-truth and detects at frame $t$ as specified in section \ref{sec:mappingtoground} using the $\text{FDA}^{(t)}$ as the relevant metric.
+Given the $\text{FDA}^{(t)}$ at each frame, the SFDA can be computed; this metric may be viewed as the average FDA over all frames of a video sequence.
+We define
+\begin{equation}
+\text{SFDA} = \frac{\sum_{t=1}^{N_{\text{frames}}} \text{FDA}^{(t)}}{\sum_{t=1}^{N_{\text{frames}}} \exists \left( N_{G}^{(t)} \vee N_{D}^{(t)} \right)}
+\end{equation}
+where $\exists \left( N_{G}^{(t)} \vee N_{D}^{(t)} \right)$ yields a 1 if either a detected or ground-truth object is present in frame $t$ and a 0 otherwise.
 
-For a sequence of video frames $t \in \{ 1, \ldots, T \}$ we define MOTA as
+The ATA metric quantifies the performance of an object tracking algorithm as a function of the spatial overlap of a mapped set of sequences of detected object positions to a set of sequences of groundtruth object positions.
+The ATA is calculated by first computing the Sequence Track Detection Accuracy (STDA), which can be viewed as a tracking performance measure unnormalied in terms of the number of objects.
+We can write the STDA as
 \begin{equation}
-MOTA = 1 - \frac{\sum_{t=1}^{T} m^{t} + e^{t} + s_{t}}{\sum_{t=1}^{T} N_{g}^{t}}
+\text{STDA} = \sum_{i=1}^{N_{\text{mapped}}} \frac{\sum_{t=1}^{N_{\text{frames}}} \left( \frac{\left|G_{i}^{(t)} \cap D_{i}^{(t)}\right|}{\left|G_{i}^{(t)} \cup D_{i}^{(t)}\right|}  \right) }{ N_{(G_{i} \cup D_{i} \neq \emptyset)} }
 \end{equation}
-where $s$ denotes the number of instances at frame $t$ where an object's identity switches from a previous frame (ie where the object's mapping differs between frame $t$ and frame $t-1$),  and the other terms are defined the same as above.
+where $N_{\text{mapped}}$ refers to an optimal mapping between ground-truth and detected objects as specified in section \ref{sec:mappingtoground} using the STDA as the relevant metric, and $N_{(G_{i} \cup D_{i} \neq \emptyset)}$ denotes the number of unique objects (including false positive objects, true negative objects, and correctly mapped objects) in a video sequence.
+
+Given the STDA for a video sequence, the ATA can be computed by the formula
+\begin{equation}
+\text{ATA} = \frac{\text{STDA}}{\left( \frac{N_{G} + N_{D}}{2} \right)}
+\end{equation}
 
@@
   \centering               
-  \subfloat[]{\includegraphics[width=0.5\textwidth]{cross_final}}
-  \subfloat[]{\includegraphics[width=0.5\textwidth]{bounce_final}}
+  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/cross_final}}
+  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/bounce_final}}
   \caption{Each plot shows a sample from the posterior distribution of the state for one of the videos in synthetic experiment one, where the time axis ascends vertically, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.
@@
 
-\subsubsection{Occlusion and Appearance Change}
+\subsubsection{Tracking though Occlusion and Appearance Shift}
 
@@
   \centering               
-  \subfloat[]{\includegraphics[width=0.35\textwidth, angle=30]{synth2_img}}
-  \subfloat[]{\includegraphics[width=0.5\textwidth]{synth2_final}}
+  % \subfloat[]{\includegraphics[width=0.35\textwidth, angle=30]{../img/synth2_img}}
+  \subfloat[]{\includegraphics[width=0.55\textwidth]{../img/synth2_img_b}}
+  \subfloat[]{\includegraphics[width=0.45\textwidth]{../img/synth2_final}}
   \caption{(a.) Video frame $f=10$ from synthetic experiment two.
@@
   \label{fig:synth_two_plot}
+\end{figure}
+
+\begin{figure}[h]
+        \center{\includegraphics[width=0.99\textwidth]{../img/synth2_alphaimg_hd.png}}
+        \caption{\label{fig:synth2_alphaimg} Extracted data and sample from the posterior of the model overlaid on three, time-varying images from synthetic experiment 2.}
 \end{figure}
@@
   \centering             
-  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2001_1}}
-  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2001_2}}
-  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2001_3}}
-  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2001_4}}
-  \caption{Four frames $f=25, 35, 45, 55$ from the PETS2001 sequence with one posterior sample mean and covarance matrix overlayed. \willie{will show path trail over prev 10 frames}}
+  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_1}}
+  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_2}}
+  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_3}}
+  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_4}}
+  \caption{Four frames $f=25, 35, 45, 55$ from the PETS2001 sequence with one posterior sample mean and covarance matrix overlaid. \willie{will show path trail over prev 10 frames}}
   \label{fig:pets2001_imgs}
@@
 	\centering
-	\subfloat[]{\includegraphics[width=0.52\textwidth]{pets2001_results_1}}
-	\subfloat[]{\includegraphics[width=0.48\textwidth]{pets2001_results_2}}
+	\subfloat[]{\includegraphics[width=0.52\textwidth]{../img/pets2001_results_1}}
+	\subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2001_results_2}}
 	\caption{Sample from posterior of the state for the PETS2001 dataset, where color represents assignment and the mean and standard deviation are shown.
@@
 	\label{fig:pets2001_results}
+\end{figure}
+
+\begin{figure}[h]
+        \center{\includegraphics[width=85mm]{../img/pets2001_alphaimg_hr.png}}
+        \caption{\label{fig:baselinedata} Extracted data and sample from posterior of the model overlaid on three, time-varying images from the pets2001 dataset.}
+\end{figure}
+
+\subsubsection{PETS2000}
+
+The PETS2000 dataset consists of
+
+\begin{figure}
+	\centering
+	\subfloat[]{\includegraphics[width=1\textwidth]{../img/pets2000_results_1}}
+	%\subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2000_results_2}}
+	\caption{pets2000}
+	\label{fig:pets2000_results}
 \end{figure}
@@
   \centering             
-  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2009_1}}
-  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2009_2}}
-  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2009_3}}
-  \subfloat[]{\includegraphics[width=0.25\textwidth]{pets2009_4}}
-  \caption{Four frames $f=10, 25, 35, 75$ from the PETS2009 sequence with one posterior sample mean and covarance matrix overlayed. \willie{will show path trail over prev 10 frames}}
+  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_1}}
+  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_2}}
+  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_3}}
+  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_4}}
+  \caption{Four frames $f=10, 25, 35, 75$ from the PETS2009 sequence with one posterior sample mean and covarance matrix overlaid. \willie{will show path trail over prev 10 frames}}
   \label{fig:pets2009_imgs}
@@
 	\centering
-	\includegraphics[width=3in]{pets2009_results}
-	% \subfloat[]{\includegraphics[width=0.52\textwidth]{pets2001_results_1}}
-	% \subfloat[]{\includegraphics[width=0.48\textwidth]{pets2001_results_2}}
+	\includegraphics[width=3in]{../img/pets2009_results}
+	% \subfloat[]{\includegraphics[width=0.52\textwidth]{../img/pets2001_results_1}}
+	% \subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2001_results_2}}
 	\caption{Sample from posterior of the state for the PETS2009 dataset, where color represents assignment and the mean and standard deviation are shown.
@@
 	\label{fig:pets2009_results}
-\end{figure}
-
-\subsubsection{PETS2000}
-
-The PETS2000 dataset consists of
-
-\begin{figure}
-	\centering
-	\subfloat[]{\includegraphics[width=0.5\textwidth]{pets2000_results_1}}
-	\subfloat[]{\includegraphics[width=0.5\textwidth]{pets2000_results_2}}
-	\caption{\willie{This dataset has not done that well so far. Not sure why. Here are two typical cases.}}
-	\label{fig:pets2000_results}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=3in]{pets2004_results_1}
-	\caption{Sample from posterior of the mean for frames $f = 1-20$ overlayed on frame $f=20$.}
+	\includegraphics[width=3in]{../img/pets2004_results_1}
+	\caption{Sample from posterior of the mean for frames $f = 1-20$ overlaid on frame $f=20$.}
 	\label{fig:pets2009_results}

### 2011-11-17 old/final_datamining/final.tex
changed structure of directory, updated paper, added more images, graphical models look good now, etc.

\documentclass{article}
\usepackage{nips10submit_e,times}
%\documentstyle[nips07submit_09,times]{article}
\usepackage[square,numbers]{natbib}
\usepackage{amsmath, epsfig}
\usepackage{amsfonts}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{easybmat}
\usepackage{footmisc}
\usepackage{subfig}
\renewcommand\algorithmiccomment[1]{// \textit{#1}}
%
\newcommand{\ignore}[1]{}
\newcommand{\comment}[1]{}
\DeclareMathOperator*{\argmax}{arg\,max}

\title{\large{Time Dependent Dirichlet Process Mixture Models\\ for Multiple Target Tracking}}

\author{
Willie Neiswanger \hspace{1cm} Frank Wood\\
Columbia University, New York, NY 10027, USA \\
\texttt{wdn2101@columbia.edu, fwood@stat.columbia.edu}
%\texttt{\{stud1@otherdepth.,\{stud3,fwood\}@stat.,\{stud2\}@columbia.edu}
%\texttt{pfau@neurotheory.columbia.edu} 
%\texttt{\{bartlett,fwood\}@stat.columbia.edu} 
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.
\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
\newcommand{\X}{\mathcal{X}}
\nipsfinalcopy

\begin{document}

\maketitle

\begin{abstract}
This paper proposes a Multiple Target Tracking (MTT) technique intended to reduce the need for explicit target identification and serve as a general method to track targets with arbitrary characteristics moving over diverse backgrounds. The technique makes use of a known time dependent Dirichlet process mixture model, comprised of a sequence of interdependent infinite Gaussian mixture models constructed to accommodate behavior that changes over time. This paper describes how pixel location values can be extracted from a wide variety of videos containing multiple moving targets and clustered using the aforementioned model in order to perform MTT. The technique is demonstrated on video segments showing multiple ant targets, of distinct species and with diverse physical and behavioral characteristics, moving over non-uniform backgrounds.
\end{abstract}

\section{Introduction}
\label{sec:introduction}

Multiple Target Tracking (MTT) is a machine vision problem that involves identifying multiple concurrently moving targets in a video and maintaining identification of these targets over time while tracking the frame-by-frame locations of each. Difficulties in producing a single, general-purpose algorithm capable of successfully carrying out MTT over a wide range of videos are mainly due to the potential for videos to be highly dissimilar; parameters such as the targets' physical characteristics (including size, shape, color, and behavioral patterns), backgrounds over which the targets move, and filming conditions may all differ. This paper describes the use of a time dependent Dirichlet process mixture model to carry out clustering on data that can be extracted in a general manner from videos containing multiple moving targets. This clustering is incorporated into a general MTT technique for use on a large variety of videos.

Most MTT methods expand upon an algorithmic framework consisting of a target identification phase and a data association phase (see \cite{1521260} and \cite{probmethods} for review of methods). The former phase aims to return a position value for each target in each frame of a video, and is usually accomplished by detecting some sort of distinctive target attribute or filming the targets in a structured environment where a homogeneous, computationally discernible background can be chosen that contrasts with the targets. The latter phase is often carried out probabilistically using, for example, probabilistic data association filters that allow potential location-sequences to be evaluated, and then selected or rejected as target paths. Much research has gone into data association filters that aim to help with a chief MTT hindrance, where the target identification phase returns many false positive positions (clutter) or true negative positions (detection holes).

The clustering-based MTT technique of this paper deviates from the usual two-phase method outlined above in that it does not explicitly carry out target identification. This technique was initially developed to aid ant tracking software intended to track ants of many species, having highly variable physical characteristics, filmed in numerous environments. Hence, the MTT methods in this software could not depend upon any characteristic specific to a given ant species or background property, and needed to instead use video features common to moving targets in any video. This technique begins by extracting three-dimensional location values $(x$, $y$, frame number$) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$ of certain pixels that tend to be present around moving targets. The extracted pixel location values are distributed roughly as worm-like shapes that follow the paths of individual targets when plotted in $\mathbb{R}^{2} \times \mathbb{Z}_{+}$, where we let the frame number (which can be viewed as a discrete representation of time) ascend the vertical axis (see Data section for further details). Clustering can then be performed with a time dependent Dirichlet process mixture model, which allows each worm-like data cloud to be isolated and examined separately. Regression can be carried out on each isolated data cluster to find a sequence of locations for each target, therby completing the MTT process.

In the following sections, the MTT process and its associated model are discussed in more detail; the process is then demonstrated on video segments containing multiple targets with varied characteristics moving concurrently over diverse backgrounds.

\section{Data Extraction}

Extraction of three dimensional location values $(x$, $y$, frame number$) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$ of pixels found near moving targets in videos is necessary to produce worm-like data clouds that can be clustered using the time dependent Dirichlet Process mixture model. One specific way of extracting the location values of relevant pixels is simple frame-differencing, which involves looking at the difference in pixel intensity or value in successive frames, and recording the positions of the pixels (or pixel groups) that have exhibited frame-wise change beyond a given threshold. Frame differencing was the main method used to generate the pixel location values in this project (however, note that the techniques described in the following sections can be applied to any pixel filtering technique producing data that, when plotted, form discernible worm-shaped clouds that trace the paths of targets).

Frame differencing is a rudimentary process, which is advantageous here, since it lends itself well to many MTT situations that can be encoutered. We have found that, when correctly implemented, it is sufficiently general to extract the desired worm-like data clouds from a wide variety of videos containing multiple moving targets with only a few restrictions on the filming conditions (relating to video stability) and target characteristics (which are described in detail in the Experiments section). Additionally, frame differencing can be applied to videos of arbitrary targets, and does not involve the identification of specific target characteristics, which is a desirable trait of this MTT technique.

\section{Model Overview}

The generalized Polya urn dependent Dirichlet process mixture model (GPUDPM) \cite{Caron2007}, described by Caron et al. [2007], is the specific time dependent Dirichlet process mixture model applied in this work. In order to introduce this model, we will begin by providing a brief overview of the (non-dependent) Dirichlet Process Mixture Model (DPMM).

A DPMM is formulated by allowing the mixing distribution in a standard mixture model to be distributed according to a Dirichlet Process (DP); this allows for modeling data where the true number of latent mixture components is unknown and arbitrarily large by assuming an infinite number of components, of which only a finite amount are expressed by the data. In particular, the DPMM can be defined as
\begin{align}
\begin{split}
	x_{i}|\theta_{i} &\sim F(\theta_{i}) \\
	\theta_{i} | \mathbb{G}  &\sim  \mathbb{G} \\
	\mathbb{G} | \alpha, \mathbb{G}_{0}  &\sim  DP(\alpha, \mathbb{G}_{0})
\end{split}
\end{align}
where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $\theta_{i}$ are parameters defining the mixture component (i.e. the distribution to be mixed, $F(\theta_{i})$) from which the $i_{th}$ observation is drawn, and the $\theta_{i}$ are drawn from $\mathbb{G}$, which is in turn drawn from a DP with base distribution $\mathbb{G}_{0}$ and parameter $\alpha$ (which can be intuitively viewed as the mean and precision of the DP, respectively). See \cite{Gasthaus2008} and \cite{Gasthaus_thesis} for more details on this formulation.

Let $K$ be the number of distinct mixture components in the above model. We can write the distinct mixture components as $\phi_{1}, \ldots, \phi_{K}$. Let $c_{1}, \ldots, c_{N}$ (where $c_{i} \in \{1, \ldots, K \}$) be class assignment variables that indicate the cluster to which observation $x_{i}$ is assigned. The DPMM can be represented by a Chinese Restaurant Process (CRP), a discrete-time stochastic process that defines a partition of the set $\{ x_{1}, \ldots, x_{N} \}$ (via the elements' assignments $c_{1}, \ldots, c_{N}$) at each time-step. The CRP allows samples to be drawn from the conditional distribution of the indictor variables $c_{i}$, and can be expressed as
\begin{align}
\begin{split}
	P(c_{i} = c_{j} $ for some $ j<i) &= \frac{m_{k}}{i-1+\alpha}\\
	P(c_{i} \neq c_{j} $ for all $ j<i) &= \frac{\alpha}{i-1+\alpha}
\end{split}
\end{align}
where $m_{k}$ is the cardinality of the set $\{ c_{j} | (c_{j}=c_{i}=k)  \wedge  (j < i) \}$ and $\alpha$ is the parameter of the DP prior on $\mathbb{G}$ .

Consider the situation where the distribution of the observations is not static, but evolves over time (for example, in cases where the number, location, and density of clusters is time dependent). In this case, a time dependent DPMM (see \cite{griffinsteel} for formal description of dependent DPMMs), such as the GPUDPM can be used to model the data. A GPUDPM that models data over T discrete time steps can be viewed as a collection of DPMMs (one for each $t \in \{1, \ldots, T \}$), which are linked together by dependencies among their respective parameters. For example, mixture component parameters at a given time step are dependent upon the mixture component parameters in adjacent time steps and the distribution and number of distinct cluster assignments at a given time step are dependent upon the same over a wider range of time steps (and upon a deletion procedure, described below).

To link together the parameters of mixture components in adjacent time-steps, a transition kernel $P(\phi_{k}^{t} | \phi_{k}^{t-1})$ is implemented, which provides a distribution over parameters of a mixture component in a given time-step (as a function of equivalent parameters in adjacent time-steps). One caveat is that each mixture component must be drawn independently from $\mathbb{G}_{0}$, which we achieve by making $\mathbb{G}_{0}$ the invariant distribution of $P(\phi_{k}^{t} | \phi_{k}^{t-1})$ (note that the transition kernel is a markov chain). Additionally, to account for varying numbers of clusters, there is a deletion procedure (described in Caron et al. [2007]) by which observations are considered "removed" from their assigned clusters at a given time-step (which modifies the value of $m_{k}$ defined above). We introduce variables $d_{1}, \ldots, d_{N}$ which denote the times at which the assignments of given observations are considered to be removed. With these variables, we can now express the $m_{k}$ value at time $t$ as
\begin{equation}
m_{k,t} = \sum_{t' = 1}^{t} \mathbb{I}[(c_{t'}=k) \wedge (t < d_{t'})]
\end{equation}
where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise. Additionally, for an observation $x_{i}$ at a given time-step $t$, the deletion time $d_{i}$ can be defined to be $d_{i} = t + l_{i}$, where $l_{i}$ is considered the lifespan of an assignment, is distributed geometrically, and can be expressed as
\begin{equation}
l_{i} | \rho  \sim  \rho(1 - \rho)^{l_{i}}
\end{equation} \\
with parameter $\rho$. This process adheres to what Caron et al. [2007] consider a uniform deletion strategy over all the observations' assignments (since assignments to each cluster have equal probability of being deleted), though a more complex deletion strategy, dependent upon cluster size, can also be implemented.

\section{Model Specification and Inference}

We now specify the particular modeling choices made when implementing the GPUDPM introduced above and describe the inference procedure---which makes use of a Markov Chain Monte Carlo (MCMC) sampler---to sample from the posterior distribution of the model's parameters (closely following the processes described in \cite{Gasthaus2008} and \cite{Gasthaus_thesis}) and perform clustering on the data.

The likelihood function (i.e. the distribution $F(\theta_{i})$ to be mixed in the DPMM) is chosen to be multivariate normal. In particular, the probability of observation $x_{i}$, given the Gaussian component specified by parameters $\phi_{k}$ can be defined to be
\begin{equation}
P(x_{i}|\phi_{k}) := \mathcal{N}(x_{i}|\phi_{k}) = \mathcal{N}(x_{i}|\mathbf{\mu}_{k}, \Lambda_{k}^{-1})
\end{equation}
where the Gaussian component's parameters are $\phi_{k} = \{ \mathbf{\mu}_{k}, \Lambda_{k}^{-1} \}$. However, restrictions will be placed on the covariance matrix $\Lambda_{k}^{-1}$;  in particular, the covariance matrix is restricted to be a diagonal matrix, of the form $\Lambda_{k}^{-1} = \mathbf{\lambda}_{k}\mathbf{I}_{D} = diag(\lambda_{k}^{(1)}, \ldots, \lambda_{k}^{(D)})$, which provides computational advantages when performing inference in the GPUDPM model. This makes the likelihood function a product of univariate normal distributions, one for each of the D dimensions of an observation. We can now reformulate the previous equation as
\begin{equation}
P(x_{i}|\phi_{k}) = \mathcal{N}(x_{i}|\mathbf{\mu}_{k}, \mathbf{\lambda}_{k}\mathbf{I}_{D}) = \prod_{d=1}^{D} \mathcal{N}(x_{i}^{(d)}|\mathbf{\mu}_{k}^{(d)}, \mathbf{\lambda}_{k}^{(d)})
\end{equation}
where terms with the superscript, $(\cdot)^{(d)}$, denote the $d^{th}$ scalar component of a multidimensional form. We now choose the base distribution $\mathbb{G}_{0}$, which serves as a prior for the cluster parameters, to be a product of Normal-Gamma distributions. This form is chosen so that the base distribution is the conjugate prior to the diagonal formulation of the likelihood above. Hence we can define the base distribution to be
\begin{equation}
\mathbb{G}_{0}(\mu_{k}, \lambda_{k} | \mu_{0}, n_{0}, a, b) := \prod_{d=1}^{D}[\mathcal{N}(\mu_{k}^{(d)} | \mu_{0}^{(d)}, n_{0} \lambda_{k}^{(d)})   Ga(\lambda_{k}^{(d)} | a,b)]
\end{equation}
where $\mu_{k}$ and $\lambda_{k}$ are parameters of a Gaussian component (at a given time-step), and $\mu_{0}$, $n_{0}$, $a$, and $b$ are parameters of the base distribution. As previously discussed, the transition kernel $P(\phi_{k}^{t} | \phi_{k}^{t-1})$ must be chosen such that its invariant distribution is $\mathbb{G}_{0}$, i.e. such that the following holds:
\begin{equation}
\int \mathbb{G}_{0}(\phi_{k}^{t-1}) P(\phi_{k}^{t} | \phi_{k}^{t-1}) d\phi_{k}^{t-1}  =  \mathbb{G}_{0}(\phi_{k}^{t})
\end{equation}

It is possible to construct a transition kernel with this property by introducing extra variables (suppose there are $M$ of them), which we refer to as auxiliary variables, and denote $z_{k, 1:M}^{t} = \{ z_{k, 1}^{t}, \ldots, z_{k, M}^{t} \}$, (described in detail in \cite{Gasthaus_thesis}) which satisfy
\begin{equation}
P(\phi_{k}^{t} | \phi_{k}^{t-1}) = \int P(\phi_{k}^{t} | z_{k, 1:M}^{t}) P(z_{k, 1:M}^{t} | \phi_{k}^{t-1}) dz_{k, 1:M}^{t}
\end{equation}
where, in this formulation, the terms $P(\phi_{k}^{t} | z_{k, 1:M}^{t})$ and $P(z_{k, 1:M}^{t} | \phi_{k}^{t-1})$ are conditional probabilities of the joint distribution
\begin{equation}
P(\phi_{k}^{t}, z_{k, 1:M}^{t}) = P(z_{k, 1:M}^{t} | \phi_{k}^{t}) \mathbb{G}_{0}(\phi_{k}^{t})
\vspace{4pt}
\end{equation}
where $\phi_{k}^{t}$ is a given Gaussian component at a given time-step, and $z_{k, 1:M}^{t}$ are all of the corresponding auxiliary variables. This approach is applied to the GPUDPM by Caron et al. [2007]. The above formulation of the base distribution and auxiliary varibles allows for easier sampling of mixture component parameters from the posterior distribution $P(\phi_{k}^{t} | \phi_{k}^{t-1}, x^{t})$ (where sampling mixture component parameters is further described below).

Inference is carried out using a MCMC sampler (which, specifically, makes use of the Metropolis-Hastings (M-H) algorithm, following the procedure described in \cite{Gasthaus_thesis}). The inference procedure involves an iterative process where the state of the sampler is updated with samples from its components' posterior distributions with each iteration. The state of the sampler consists of the class assignment variables $\{ c_{1}, \ldots, c_{N} \}$, the corresponding assignment deletion variables $\{ d_{1}, \ldots, d_{N} \}$, the mixture component parameters $\{ \phi_{1}, \ldots, \phi_{K} \}$ (for each component to which at least one observation was assigned at some $t \in \{1, \ldots, T \}$), and the corresponding auxiliary variables $\{ z_{1, 1:M}, \ldots, z_{K, 1:M} \}$. In each iteration, the sampler moves through the $T$ time-steps sequentially, updating the variables in the state each time. At a given time-step, the class assignment variables and mixture component parameters can be directly sampled from their respective posterior distributions, while the M-H algorithm is used to sample assignment deletion variables and auxiliary variables from their respective posterior distributions.

% Here we will describe more details of the steps in the MCMC sampler algorithm.

% The updates for parameters of the Gaussian components are given by Fink [1995].

\section{Experiments}
\hspace{18pt} Experiments were performed on data extracted from videos of ants collected by members of the Ecology and Evolutionary Biology Department at the University of Michigan. Each video shows a number of specimens of a distinct ant species. These videos were chosen to demonstrate the benefits of the MTT technique exhibited in this paper as they involve targets with diverse physical characteristics filmed moving over varied backgrounds. Additionally, each video has a different number and density of targets that display distinct behaviors.

From these videos, short segments were selected for experimentation. As a proof of concept for this MTT technique, our initial goal was to perform target tracking over a short time span---around 50 to 100 video frames. Given the frame rate of the videos used in this study, this is equivalent to approximately four to five seconds of tracking.

A consistent frame-differencing procedure is carried out on each video segment for data extraction. This procedure involves dividing the image associated with a given video frame into small blocks, comparing the blocks pairwise against those in an adjacent frame, and recording which pairs have an average pixel value difference that exceeds a set threshold. The number and size of the blocks, as well as the threshold value, can be used to specify the density and number of data points returned. With this procedure, a video containing multiple targets moving over an arbitrary background yields data representing the locations of pixel blocks whose average values undergo change beyond the specified threshold, and thus provides the desired worm-shaped clouds that trace the paths of moving targets. An example result from the data extraction process is shown in Figure 1. This data was extracted from a 50 frame sequence showing five ants moving across the video scene.\\
\begin{figure}[htb]
        \center{\includegraphics[width=75mm]{paths.png}}
        \caption{\label{Figure 1:} Data extracted from a video of multiple moving targets. The vertical axis denotes the frame number and the two horizontal axes denote $x$ and $y$ coordinate positions, respectively, within a given frame. The data comes from a 50 frame video segment showing 5 moving ants (note that one ant enters the scene near the middle of the segment). The data extraction procedure was set to collect a high density of pixel locations to allow for easier viewing.}
\end{figure}\\
\\
The main goal of the data extraction procedure is to stay simple enough to be applicable to a wide range of videos containing multiple moving targets without the need for a modified procedure when a new setting, filming condition, or collection of targets is encountered. To this ends, this procedure was used without modification for all videos in this study.

The first experiment was performed on a video segment containing two ants moving across a heterogeneous background with many visible features. A frame from this video segment is shown in Figure 2 (a.). The second experiment was performed on a video segment containing five ants, also moving across a heterogeneous background with many visible features. However, relative to the first experiment, the ant-size to frame-size ratio was larger, the ants were travelling at greater speeds, the ants were closer together (a few of the ants briefly come into contact with each other during this video segment), and the background color and texture were more similar to the color and texture of the ant targets. A frame from this video segment is shown in Figure 2 (b.).
\begin{figure}
\centering
{\includegraphics[width=65mm]{antpic1.png}}
\hspace{8mm}
{\includegraphics[width=65mm]{antpic2.png}}
\caption{(a.) A frame from the video segment used in experiment one. There are two ants in this image; one is center left and one is center slightly-right.  (b.) A frame from the video segment used in experiment two. There are five ants in this image; four are in a rough diamond shape surrounding the ant in the center.}
\label{test}
\end{figure}

All values in the state of the MCMC sampler needed to be initialized for inference to be be carried out. To start, the number of initial clusters was chosen randomly. To initialize a given cluster's parameters and auxiliary variables at each time step (both of which are used in the inference procedure), an observation was randomly chosen as a seed. The cluster's parameters were then sampled from their posterior distribution given the observation. Using these parameters, the auxiliary variables at an adjacent time step could be sampled from their posterior distribution (as before, making use of the M-H algorithm), which allowed the cluster parameters at the next time step to be sampled. This sampling chain was extended forwards and backwards from the time of the initial observation, allowing the cluster parameters and auxiliary variables to be sampled at all time steps. In this process, the means of the univariate normal distribution followed a random walk of sorts as they were successively sampled. Afterwards, for each observation, the associated assignment variable was randomly assigned to an existing cluster, and the associated deletion variable was drawn from its prior distribution (which entailed drawing a value representing the lifetime of the observation from a geometric distribution, and adding this value to the current time). To illustrate this initialization process, the extracted data from experiments one and two are displyed in Figure 3 (a.) and (b.) below; the shape of each data point represents its associated cluster assignment, and the lines running through the data points represent the paths followed by the mean-component of the cluster parameters during the initial sampling random walk.
\begin{figure}
\centering
{\includegraphics[width=110mm]{exp1_init_3.png}}
\hspace{8mm}
{\includegraphics[width=110mm]{exp2_init_5.png}}
\caption{(a.) Initialization of the state in the first experiment. (b.) Initialization of the state in the second experiment.}
\label{test}
\end{figure}

\section{Results}
Results from experiments one and two are shown in Figure 4 (a.) and (b.), respectively. In both cases, the data point assignments seem to predominantly agree, visually, with their membership to their respective worm-like data clouds. Experiment one seemed to separate the worm-like clusters better than experiment two; this is not surprising, as experiment one tracked a much less complex and better separated configuration of targets.
\begin{figure}
\centering
{\includegraphics[width=110mm]{exp1_c.png}}
\hspace{8mm}
{\includegraphics[width=110mm]{exp2_a.png}}
\caption{(a.) Clustering results from experiment one. (b.) Clustering results from experiment two.}
\label{test}
\end{figure}
%
%
%
\\\\ To complete the multiple target tracking process, linear regression was carried out on the data in each isolated cluster. Each regression line was considered the centroid path of a target, and the $\{ x, y \}$ position at which the (vertically oriented) regression line intersected with the horizontal plane of a frame number was considered the target's position at that frame. To account for a small number of false positive assignments, regression was only performed on clusters whose sizes exceeded a specified threshold. The results of regression are shown in Figure 5 (a.) and (b.); a red regression line is overlayed on the plots resulting from the inference procedure.
\begin{figure}
\centering
{\includegraphics[width=110mm]{exp1_regression.png}}
\hspace{8mm}
{\includegraphics[width=110mm]{exp2_regression.png}}
\caption{(a.) Experiment one results with regression lines. (b.) Experiment two results with regression lines.}
\label{test}
\end{figure}

\section{Conclusion}
This paper has exhibited successful MTT over short video sequences with the use of a time dependent GPUDPM model. The technique was illustrated on two multiple target tracking scenarios involving disparate targets and backgrounds. A general data extraction algorithm was outlined as a simple way to collect data from a wide variety of videos containing multiple moving targets. We hope that this research might help in development of more general MTT algorithms, with the ability to track a large variety of targets with arbitrary characteristics moving over diverse backgrounds

\begin{small}
\bibliographystyle{plainnat}
\bibliography{refs2} 
\end{small}

\end{document}

### 2011-11-17 old/newmodel/newmodel.tex
moved some excess stuff to old folder... it contains old stuff

--- 2011-10-06 ltx/paper.tex
@@
 
-% need to include style sheet for journal
-
 \begin{document}
 
-\title{Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures}
-\author{Willie Neiswanger and Frank Wood}
+\title{New Model}
+\author{Willie}
 \maketitle
-\mbox{}
-
-\abstract{main points}
-
-\section{Introduction}
-
-\frank{this section should introduce our work and situate it relative to the literature}
-\willie{okeydoke. this isn't ready yet. to avoid confusion I've just commented the whole section out.}
-
-% % \subsection{Topic Overview and Terminology}
-
-% This work is related to the study of algorithms that perform automated detection and tracking of arbitrary objects in videos; we define such algorithms as those with the ability to decide which spatiotemporal regions of a video constitute objects (extraction), find the positions and/or shapes of distinct objects (localization), and maintain the positions of the detected objects over time (tracking).
-Robust algorithms which can carry out these tasks are applicable to problems in the fields of surveillance, autonomous systems and robotics, and video summarization and compression.
-Approaches to these tasks have been proposed under different headings and in a variety of contexts.
-This section serves to introduce relevant terminology and topics of research.
-
-% The task of extraction is closely related to the topics of foregound segmentation, background modeling, motion detection, and appearance feature extraction.
-In general, these methods aim to extract features from regions of a video frame that are considered objects or are distinguished in some way from the background of a scene.
-Areas of research in this field include attempts to handle situations where the background displays motion or other dynamic characteristics, reduce noise caused by the false extraction of backgrounds, handle abberant image effects present in some videos, and operate accurately on scenes involving illumination changes.
-
-% The task of localization is closely related to the topics of object detection, position-finding, target segmentation, contour-detection, template matching, and recognition.
-These methods are mainly concerned with finding the positions and/or shapes of objects in a given video frame, or with discerning between objects similar in position or appearance.
-There has been a focus in this field on developing techniques for representing an object's appearance, both to allow for localization of objects on a per-frame basis and to allow for consistent localization of a given object over time.
-Research into the unsupervised detection of arbitrary objects (which we define as methods that accomplish both extraction and localization) has included work involving image segmentation, parsing, and pixel clustering.
-These topics usually relate to the automated partitioning of images into meaningful regions.
-Note that the popular phrase `object detection' is also typically used to refer to procedures that carry out extraction and localization; however, it is often reserved to denote methods that are tailored to find the position of specified objects or object types (as opposed to arbitrary object detection), or to denote supervised methods, where extraction is peformed implicitely during searches for specified objects and localization procedures.
-
-% The task of tracking is closely related to the topics of video tracking, data association, filtering, point matching, and `tracking of non-rigid objects'.
-In general, these tasks aim to maintain localization of the positions and/or shapes of objects over time.
-Areas of interest in this field involve developing metrics for judging the similarity between two object representations (for example, potential representations of the same object at adjacent frames in a video), techniques that provide a way to conduct efficient searches over a frame for regions that are similar to a given object, and methods for predicting the future position or appearance of an object.
-A great deal of research in this field over the past decade has focused on the development of algorithms to track multiple objects simultaneously.
-There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera), complex object interactions, object coalescence (where one object is tracked by multiple times/redundantly), objects with similar appearances, variable (and potentially high) numbers of objects, and objects that enter and exit a field of view at different times.
-
-% There does not yet exist a major body of research concerned primarily with methods for combining extraction, localization, and tracking to perform cohesive unsupervised detection of arbitrary objects in videos.
-Many papers that attempt to accomplish this goal focus on one of extraction, localization, or tracking, and attempt to integrate in methods from the other two tasks.
-However, a few key schemes for unsupervised detection and tracking do indeed exist.
-A widely applied, though unsophisticated, technique to carry out such a procedure is called blob tracking.
-Blob tracking typically refers to methods that perform extraction and localization at each frame independently--for example, by assuming that object pixels have a different intensity than background pixels, setting a fixed pixel intensity threshold for extraction of object pixel locations, and segmenting the extracted data into `blobs' to accomplish localization.
-Afterwards, tracking is often carried out by finding the centroid of each blob and applying standard multiple target tracking algorithms (described in [section]).
-Note that blob tracking carries out the three main tasks described above, so it satisfies the criteria for the type of unsupervised detection and tracking algorithms studied in this work; there are, however, many challenges associated with blob tracking that render it impractical in many settings (detailed in [section]).
-
-% This paper presents a new method for unsupervised multiple object detection and tracking.
-We introduce, and provide details on the theory behind, a technique involving dependent Dirichlet process mixture models.
-Specifically, this work applies the model to data gained during the extraction process; note that data of this type is refered to at different points in the paper as `extraction data', `data points', 'points', or 'observations'.
-The model is first presented in an abstract form, which could be adapted to a wide range of extraction procedures and object appearance representations (mainly left for future studies).
-We then describe inference algorithms that allow object localization and tracking to be carried out.
-Afterwards, we formulate a definite model for data gained during a specific extraction procedure and implement the described inference procedures to carry out multiple object detection and tracking in a number of experiments.
-
-% \subsection{Related Work}
-
-% Related work can be broken down into studies relating to extraction procedures, unsupervised localization procedures, single object tracking methods, multiple object tracking methods, supervised multiple object detection and tracking methods, and unsupervised multiple object detection and tracking methods.
-
-% Extraction procedures can usually be classified as either model-based methods or heuristic-based methods.
-Model-based methods emphasize the detection of shapes of regions containing objects accurately and reducing background noise.
-For example, foreground extraction has been robustly carried out by modeling each pixel over time in order to infer a probability distribution over pixel values; if the pixel takes on a very unlikely value at a future time, it is marked as a foreground pixel at that point \citep{stauffer_1999, elgammal_2000, elgammal_2002}.
-These pixel modeling techniques have been adapted to allow for time-dependent probability distributions over pixel values, which help account for slighlty changing backgrounds, such as those caused from brief movements of background objects (for example, trees in outdoor scenes).
-Note that these techniques treat each pixel independently; this has a tendency to cause errors in foreground detection, resulting in fragmentation of foreground images (which is problematic as many of these methods intend to blob the foreground objects after detection in order to perform segmentation and locate objects' centroids).
-Post processing has been applied to add information regarding the edges of objects in an attempt to decrease fragmentation \citep{turdu_2007}.
-Others have attempted to achieve better foreground modeling in cases where the foreground and background exhibit similar color distributions (the so called `color similarity problem') by, for example, incorporating models of the foreground into a typical background model (such as the one described above) \citep{gallego_2009, mchugh_2009}.
-
-% On the other hand, heuristic-based methods perform extracting by looking for temporal changes in pixel value characteristics.
-While these methods locate regions with objects less accurately than model-based methods, they are often much less computationally expensive.
-Frame differencing and background subtraction are two such methods, and involve comparing pixel values, or groups of pixel values, between two images, to detect a change surpassing some threshold.
-The goal of these techniques is to find locations in a scene that display motion, and to deem these moving areas the foreground of a scene.
-Often, background subtraction refers to comparing an image containing targets with an image of the background without targets or with some model of the background that is learned as the video progresses (in these cases, objects that do not move throughout a scene will be treated as the background), while frame or image differencing refers to comparing two subsequent images in a video.
-Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking schemes \citep{perez_2002}.
-In addition to the two techniques mentioned above, there exists other ways of extracting movement from an image, such as through techniques involving optical flow (a calculation based on a so called `generalized gradient model' of an image, which attempts to capture the speed and direction of movement over areas in the image) \citep{horn_1981, bobick_2001}.
-In \citep{black_2000}, a motion detection and object tracking method is developed based on the relative motion of moving objects' boundaries, which are again found through analysis of optical flow.
-
-% There are an extensive collection of methods for tracking single, non-rigid, objects.
-For example, color has long been used for tracking.
-In this case, tracking refers to the task of maintaining the location of an object after its initial position has been specified.
-In general, these schemes operate by modeling an object with some color-based appearance model, and using this model to find the object in subsequent frames.
-One classic approach involves modeling the color of a target (in particular, a distribution over the hue-saturation space of a target region) with a Gaussian mixture model (GMM), and choosing subsequent target positions by searching surrounding areas for regions that yield similar GMMs.
-Furthermore, the GMM is often allowed to adapt over time, and slowly change to model changes in lighting or other smooth time-based variations in the target's color \citep{raja_1998, mckenna_1999, jepson_2003}.
-Others have attempted to abstract this work with a non-parametric color modeling approach based on kernel density estimation, which does not assume a specific underlying distribution (such as the Gaussian mixture in the previous case) and instead converges to reasonable distribution that depends on the data \citep{elgammal_2001}.
-Additionally, color has been successfully applied as the feature in the mean-shift procedure (kernel-based method) for tracking \citep{comaniciu_2003, perez_2002, nummiaro_2003, lee_2011}.
-
-% The mean-shift procedure (also known as `kernel-based object tracking') attempts to provide a robust way for non-rigid objects to be tracked.
-This method is beneficial because it optimizes the search for the `next' position of an object (i.e.
-the search for the region in a frame which is most similar to a target region in the previous frame).
-This procedure--a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges for each data point.
-Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors.
-It can be shown that, through this procedure, each data point becomes associated with a local point of high density (dependent upon the underlying weight distribution specified by the kernel) which naturally allows for clustering \citep{cheng_1995}.
-The mean shift algorithm has been implemented successfully to allow for a sort of kernel-based object tracking \citep{comaniciu_2003, comaniciu_1999, comaniciu_2000}.
-Given the current position of an object at a given frame, the goal of tracking is often to find a nearby position in the next frame that has the most similar distribution over some common set of features.
-Usually this must be done through an exhaustive search, comparing the similarity of distributions at each nearby position with that at the current position.
-However, if a certain type of `isotropic kernel' (mean-shift kernel) known as the Bhattacharyya coefficient is chosen as a similarity metric between the feature distributions at two positions, it creates a smooth function where gradient descent techniques can be used to quickly converge upon an optimal subsequent position without using an exhaustive search.
-Many papers are concerned with applying this kernel-based object tracking scheme to different sets of features or with different kernels.
-Additionally, the scheme has been attempted with an adaptive kernel whose shape, scale, and orientation is influenced by a target being tracked \citep{wang_2004}, with a kernel adjusted by the estimated centroid of a tracked target \citep{mehmood_2009}, and with a heirarchical version of the mean-shift procedure \citep{dementhon_2002}.
-
-% Some have tried to extent single object tracking schemes to tracking multiple targets; at the most basic level, this involves initializing multiple targets and running an instance of (single) VOT for each, either simultaneously or in succession \citep{perez_2002}.
-Furthermore, if performed simultaneously, the tracking of each target can be made dependent upon characteristics of other targets (such as their proximity) to resolve errors and improve tracking (such as those caused by the incorrect merging of two targets) \citep{khan_2004, vermaak_2003} .
-
-% %  ------------------------------------------------------------------------------------
-% % Localization-focused. Clustering methods, segmentation methods, object appearance models and representations (allows for consistent object localization over time).
-
-% % Methods for tracking single objects in videos: filtering, similarity metrics / localization (and how they play a role), mean shift procedures.
-
-% % extensions to multiple visual object tracking.
-problems with multiple independent single object trackers (coalesence, object interaction, object occlusion, objects entering and leaving scene [and, relatedly, variable numbers of objects]).
-% % Multiple object tracking attempts at overcoming various the various challenges specified above (include nonparametric varmaak/doucet for maintaining multi-modality here?---sure, and then mention again later with fox's work?)
-% % End with description of initialization problem.
-
-% % Methods for detection and tracking of objects.
-Heuristic initialization at beginning and then carrying out above methods (problems with varying numbers of obects / objects entering or leaving scene).
-Supervised learning of specific targets to allow for initialization and provide better localization throughout.
-Various research has been towards integrating unsupervised extraction and localization into multiple object tracking methods.
-Particle filters with integrated detection (at least two of these).
-Introduce concept of blob tracking and define basic algorithm (some manner of extraction, independent/per-frame localization, tracking; note that each object is represented in each frame as a low-dimensional point [or higher dimensional thing like region/outline with sophisticated blob/contour detectors like bramble and others]).
-The key thing characterizing a blob detector is that the position of each object (or the spatially separate region associated with each object) is assigned/fixed at each frame without the aid of tracking (ie if localization/segmentation can be carried out at each frame independently), and tracking is done with this data.
-Schemes often are simply segmentation of extraction data into spatially disjoint blobs at each frame (where blobs  are not modeled in any principled manner), returning the position or region associated with each object at each frame.
-
-% % When the state of objects (for example, position and/or other characteristics) can be found independently at each frame, well established methods in the field known as Multiple Target Tracking (MTT) can be carried out to perform tracking of multiple objects.
-Two classic algorithms for MTT are known as MHT and JPDA.
-In general, MTT methods are useful for procedures that can represent each object at each frame as a point or region without localization information from this object at a previous frame (which requires the use of tracking).
-MTT in the literature is often quite separate from fields of computer vision involving object detection and tracking research.
-In general, the situation arises as video data without modification is very high dimensional (discussion of difference between MTT and detection methods here?).
-
-% % Methods like blob tracking which attempt to find the state of each object independently at each frame have a hard time dealing with occlusion and other complexities (except for sophisticated blob trackers like bramble which isn't really blob tracking).
-Attempts have been made to perform time dependent clustering on extraction data (look at cluster tracking paper for good def), to carry out joint localization and tracking; this has been refered to as cluster tracking.
-Work of Pece, MTT of multiple targets work(?).
-The work presented in this paper most closely resembes the topics listed here.
-Does cluster tracking on whichever extraction data, using a time dependent Dirichlet Process mixture model (to better estimate and maintain the number of targets, as was shown to be effective in vermaak and fox), which takes the form of a sequence of clusters, where the objects' appearances are modeled in a principled manner by the cluster component distributons, and the tracking behavior is modeled by time-dependencies between adjacent elements in the cluster-sequences.
-% %  ------------------------------------------------------------------------------------
-
-% \subsection{Contributions}
-
-% This paper applies a time dependent Dirichlet process mixture to model data gained during extraction procedures, provides background as to how this Bayesian nonparametric model allows for the incorporation of prior knowledge about object appearances and motion behaviors, describes methods for unsupervised detection and tracking of arbitrary objects in terms of Bayesian inference procedures, and presents a specific implementation of the model and inference procedures to carry out detection and tracking on a number of target-types in different settings.
-
-% % List here the benefits of this idea that should be discussed/listed:
-% This paper is heavily model-based, which differs from many papers in the field of multiple object tracking that do not explicitely describe models which underlie various procedures that are carried out.
-Solutions in many papers are, instead, algorithm oriented.
-The model-based approach allows rigor to be kept as inference procedures are developed.
-Furthermore, development of inference procedures are performed separately from development of the model, and various well studied inference and approximate inference methods can be carried out to achieve algorithmic results that are (in some cases) known to be optimal.
-% %treated separately, which allows for multiple methods to be carried out in a unified framework and   these solutions are  but provide algorithm-based presentations.
-% %	This way allows rigor to be kept as inference procedures are developed.
-% %	Furthermore, inference is treated separately; this allows multiple detection and tracking algorithms to be developed (in a well-formed/rigorous fashion) by developing different inference techniques.
-Also, inference on graphical models is a well-structured field and allows for use of techniques in this field as well as better interpretations of the resulting detection/tracking algorithms.
-
-% % (related to previous?) models objects as distribution-components (of a mixture model) as opposed to point-targets or targets with noise that adheres to some distribution (is this truly a distinction and is it good?)
-
-% This idea presents a framework which allows for time-varying models of objects' appearances and motion behaviors to be incorporated, and is formulated in a Bayesian manner that allows distributions underlying the parameters of these models to be specified in a principled manner.
-It has been shown to work successfully in conjunction with simple (noisy) extraction methods.
-More robust, expensive, or in general accurate methods can be used for further accuracy.
-This method is able to handle the appearance and disappearance of objects from a scene, variable numbers of objects, cases where objects show time varying changes in behavior, and cases of occlusion.
-
-% Section two of this paper provides an overview of the model and describes how aspects can be modified for for modeling object appearance and motion behavior.
-Section three presents a specific implementation of the model and a number of inference procedures for carrying out object detection and tracking.
-Section four describes a number of experiments carried out with this implementation.
-Section five describes results of these experiments, comments on further directions of study, and provides concluding remarks.
-
-\section{Model Overview}
-\label{sec:modeloverview}
-%
-This section provides details on the nature of the data to be modeled, Dirichlet process mixture (DPM) models, and dependent Dirichlet process mixture (DDPM) models.
-We end by making explicit the connection between elements of the model and object characteristics that these elements represent (for example, the appearance and motion behavior of objects are specifically captured).
-The provided model is formulated here in a sufficently abstract manner to allow for various representations of objects to be specified in future studies.
-
-\subsection{Data}
-\label{sec:data}
-%
-We defined the task of extraction to be the unsupervised decision of which spatiotemporal regions of a video constitute objects; hence, the extraction data modeled in this work are D-dimensional observations that in some way provide information about the regions of the video frame that contain objects.
-A simple example of this information is shown by the three-dimensional extraction data $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $x_{1}$ and $x_{2}$ are the two spatial dimensions of pixels contained in regions comprising objects in a given frame, and $t$ is the discrete time index of the frame.
-These three features are always available once regions containing objects are discerned, and should always be kept as features of the data for modeling (as position information is key for reasonable tracking and time information is necessary for our intention to model the temporal dependencies of the data).
-
-Examples of additional features that could potentially be extracted from video frame regions containing objects include color information, pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, texture representations, transform results (such as results of hough or fourier transforms), and quantiative region descriptions derived from image segmentation procedures; the main idea is to choose features which are able to be easily extracted or computed, capture variability in the appearance of objects, are applicable to a wide variety of object types, and, when viewed in cumulation for multiple observations, are distributed in a way that allows for modeling and tractable inference (this is discussed further in section 3).
-
-We will use the phrase `baseline' three-dimensional extraction data to refer to the $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$ from pixels present in regions containing objects.
-These data are distributed roughly as worm-like shapes that follow the paths of individual objects as they move, when they are plotted in three dimensions (where we let the frame number, our discrete representation of time, ascend the vertical axis and the spatial positions occupy the horizontal axes).
-An example of data with such a distribution can been seen in Figure 1.
-The data consists of the three-dimensional baseline features and was gained by extraction peformed on a 50 frame sequence showing five ants moving across the video scene.
-\begin{figure}[h]
-        \center{\includegraphics[width=75mm]{paths.png}}
-        \caption{\label{Figure 1:} Baseline three-dimensional extraction data gained from a video of multiple moving targets.
-The vertical axis denotes the frame number and the horizontal axes denote $d_{1}$ and $d_{2}$ spatial coordinates, respectively, of pixels within a given frame.
-The data comes from a 50 frame video segment showing 5 moving ants (note that one ant enters the scene near the middle of the segment).}
-\end{figure}
-Inference performed on the time dependent Dirichlet process mixture model (described in the following sections) carries out clustering of the data, which allows each worm-like data cloud to be isolated and independent objects to be localized and tracked.
-
-\subsection{Dirichlet Process Mixtures}
-
-Dirichlet process mixtures fall under the heading of Bayesian nonparametric (BNP) models.
-These models have been widely used over the past decade to perform nonparametric density estimation and cluster analysis.
-They are, in particular, useful for estimating the number of latent classes (clusters) in mixture models.
-In this work, estimating the number of clusters in the data generated via extraction is shown in the following sections to be equivalent to estimating the number of distinct objects in a video.
-Consequentially, a model which is able to perform robust nonparametric estimation of the number of cluster is well suited to be applied to the task of unsupervised object detection and tracking.
-
-The following sections are intended to provide a solid introduction to DPMs without a major foray into the theoretical aspects of the model that require a more rigorus technical setup.
-We give background on finite mixture models, Bayesian finite mixture models, and Dirichlet processes, presently, which allows us to subsequently describe DPMs in context.
-
-\subsubsection{Finite Mixture Model}
-
-A finite mixture model can be thought of as a probability distribution for an observation $x_{i}$ formulated as a linear combination of K mixture components (where the coefficients of the linear combination sum to one); each mixture component, in turn, is a probability distribution for $x_{i}$, given some parametric form.
-The finite mixture model can be written as
-
-\begin{equation}
-P(x) = \sum_{k=1}^{K} P(c_{i} = k)P(x|\theta_{k})
-\end{equation}
-
-where $c_{i} \in \{ 1, \ldots, K \}$ denotes the assignment of $x_{i}$ to a given mixture, which allows the coefficients of the linear combination to be written as $P(c_{i} = k)$ (note that these coefficients now sum to one naturally due to the nature of probability distributions).
-Additionally, $\theta_{k}$ denotes the parametric form of the $k^{\text{th}}$ mixture component.
-We also define $p_{k} = P(c_{i} = k)$ for all $k \in \{ 1, \ldots, K \} $.
-This model can also be written hierarchically as
-\begin{align}
-\begin{split}
-	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}}) \\
-	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
-\end{split}
-\end{align}
-where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $c_{i}$ are the mixture component assignments associated with each observation, and the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e.
-the distribution to be mixed, $F(\theta_{c_{i}})$).
-
-\subsubsection{Bayesian (Finite) Mixture Model}
-
-The finite mixture model of the previous section can be extended to a Bayesian perspective by viewing the distribution parameters that were previously point values, $\theta_{c_{i}}$ (the mixture component parameters) and $p_{1}, \ldots, p_{K}$ (the mixture component assignment weights), as random variables and providing each with a prior distribution.
-In this case, the prior distribution $\mathbb{G}_{0}$ is placed on the mixture component parameters, and the prior distribution $\text{Dir}(\alpha/K, \ldots, \alpha/K)$ is placed on the K mixture component assignment weights.
-The resulting Bayesian mixture model can be formulated in a hierarchical representation as
-\begin{align}
-\begin{split}
-	\label{bayesian_mixture_model}
-	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}}) \\
-	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
-	\theta_{c_{i}}  &\sim  \mathbb{G}_{0} \\
-	p_{1}, \ldots, p_{K}  &\sim  \text{Dir}(\alpha/K, \ldots, \alpha/K)
-\end{split}
-\end{align}
-where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $c_{i}$ are the mixture component assignments associated with each observation, the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e.
-the distribution to be mixed, $F(\theta_{c_{i}})$), the $\theta_{c_{i}}$ are drawn from their prior distribution $\mathbb{G}_{0}$, and $p_{1}, \ldots, p_{K}$ are drawn from their prior distribution, where `Dir' refers to a standard Dirichlet distribution.
-
-\subsubsection{Dirichlet Process}
-
-% Define Dirichlet distribution first?
-Each draw from a K-dimensional Dirichlet lies on the K-simplex (i.e.
-is a vector of K values which sum to 1) and can therefore be thought of as a K-dimensional discrete distribution.
-
-The Dirichlet process (DP), first introduced by \cite{ferguson_1973} in 1973, may be intuitively viewed as a probability distribution over discrete probability distributions.
-Accordingly, draws from a DP are probability mass functions (PMFs).
-A DP is parameterized by a base distribution $\mathbb{G}_{0}$, which is a probability distribution over a set $\Theta$, and a concentration parameter $\alpha \in \mathbb{R}_{+}$.
-We say that $G$ is a random PMF distributed according to a DP, written $G \sim \text{DP}(\alpha, \mathbb{G}_{0})$, if the following holds for all finite partitions $A_{1}, \ldots, A_{p}$ of $\Theta$:
-\begin{equation}
-(G(A_{1}), \ldots, G(A_{p})) \sim \text{Dir}(\alpha \mathbb{G}_{0}(A_{1}), \ldots, \alpha \mathbb{G}_{0}(A_{p}))
-\end{equation}
-Where `Dir' denotes a standard Dirichlet distribution.
-The parameters $\mathbb{G}_{0}$ and $\alpha$ may be intuitively viewed as the mean and precision of the DP.
-This is due to the fact that if the base distribution $\mathbb{G}_{0}$ is a distribution over $\Theta$, $A \subset \Theta$, and $G \sim \text{DP}(\alpha, \mathbb{G}_{0})$, then the following holds:
-\begin{equation}
-\mathbb{E}[G(A)] = \mathbb{G}_{0}(A)
-\end{equation}
-\begin{equation}
-\text{Var}[G(A)] = \mathbb{G}_{0}(A) (1 - \mathbb{G}_{0}(A)) / (\alpha + 1)
-\end{equation}
-Hence, the expectation of $G(A)$ is $\mathbb{G}_{0}$, the variance of $G(A) \rightarrow 0$ as $\alpha \rightarrow \infty$, and $G$ converges pointwise to $\mathbb{G}_{0}$ when $\alpha$ is unbounded.
-
-\subsubsection{Dirichlet Process (Infinite) Mixture Model}
-
-A Dirichlet process mixture model, also refered to as an infinite mixture model, is an extension of the Bayesian mixture model described previously. 
-When using a DP as a prior in a Bayesian mixture model, $\Theta$ can be viewed as the set of parameters of the component mixture distributions.
-%
-A DPM is formulated by allowing the prior mixing distribution (distribution over component weights?) in a standard mixture model to be distributed according to a DP; this allows for modeling data where the true number of latent mixture components is unknown and arbitrarily large by assuming an infinite number of components, of which only a finite amount are expressed by the data.
-In particular, the DPM can be defined hierarchically as
-\begin{align}
-\begin{split}
-	x_{i}|\phi_{i} &\sim F(\phi_{i}) \\
-	\phi_{i} | \mathbb{G}  &\sim  \mathbb{G} \\
-	\mathbb{G} | \alpha, \mathbb{G}_{0}  &\sim  \text{DP}(\alpha, \mathbb{G}_{0})
-\end{split}
-\end{align}
-where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $\phi_{i}$ are parameters defining the mixture component from which the $i_{th}$ observation is drawn (i.e.
-the distribution to be mixed, $F(\phi_{i})$), and the $\phi_{i}$ are drawn from $\mathbb{G}$, which is in turn drawn from a DP with base distribution $\mathbb{G}_{0}$ and parameter $\alpha$.
-See \cite{gasthaus_2008} and \cite{gasthaus_thesis} for more details on this formulation.
-Note that difference between the formulation and indexing of the clusters in this model compared to the previous two models.
-This definition can be shown to be equivalent to the Bayesian mixture model defined in \eqref{bayesian_mixture_model}, when K is taken to be unbounded.
-This is the origin for references to this model as the infinite mixture model.
-
-An alternative formultion of the DPM is known as the Chinese resturant process (CRP) sampling representation.
-If we let $K$ be the number of distinct mixture components in the above model, we can write the distinct mixture components as $\theta_{1}, \ldots, \theta_{K}$.
-Let $c_{1}, \ldots, c_{N}$ (where $c_{i} \in \{1, \ldots, K \}$) be class assignment variables that indicate the cluster to which observation $x_{i}$ is assigned.
-This allows the DPM to be represented by the CRP, a discrete-time stochastic process that defines a partition of the set $\{ x_{1}, \ldots, x_{N} \}$ (via the elements' assignments $c_{1}, \ldots, c_{N}$).
-The CRP allows samples to be drawn from the conditional distribution of the indictor variables $c_{i}$, and can be expressed as
-\begin{align}
-\begin{split}
-	P(c_{i} = c_{j} \text{  for some  } j<i) &= \frac{m_{k}}{i-1+\alpha}\\
-	P(c_{i} \neq c_{j} \text{  for all  } j<i) &= \frac{\alpha}{i-1+\alpha}
-\end{split}
-\end{align}
-where $m_{k}$ is the cardinality of the set $\{ c_{j} | (c_{j}=c_{i}=k)  \wedge  (j < i) \}$ and $\alpha$ is the parameter of the DP prior on $\mathbb{G}$.
-
-\subsection{Dependent Dirichlet Process Mixtures}
-
-The goal of dependent Dirichlet process mixtures (DDPMs) is to allow modeling of data that is not independent and identically distributed (i.i.d) but instead has some underlying dependencies.
-For example, data generated during extraction procedures from videos have some associated temporal structure, since tracked objects display time dependent characteristics (in the features that denote position, as well as those which denote other appearance characteristics).
-
-To account for the dependent behavior of data, research has been conducted on developing models involving a sequence of DPMs, where components of the mixtures are dependent upon (or may be considered `tied to') corresponding components at adjacent positions in the sequence.
-For example, if the data shows temporal dependence, the goal might be to create a sequence of DPMs, one for each time-step, where the components of the mixture at each step are dependent upon corresponding components in the both the following and previous time steps.
-Research towards models which accomplish such goals are described in (list citations).
-
-More rigorously, we take the definition of a DDPM to be a stochastic process defined on the space of probability distributions over a domain, which are indexed by time, space, or a selection of other covariates in such a way that the marginal distribution at any point in the domain follows a Dirichlet process (adapted from definitions found in [cite griffin and steel and gasthaus-thesis]).
-
-\subsubsection{Generalized Polya Urn Dependent Dirichlet Process Mixture}
-\label{sec:gpudpm}
-
-Data gained by performing extraction on videos containing objects is significantly time-dependent.
-Futhermore, the clusters of data, each of which represents an object in the video, might change in number as time progresses, since objects can enter and exit a scene.
-The specific DDPM used to model extraction data in this work while handling the above challenges is known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM), and was introduced by \cite{caron_2006} in 2006.
-
-The GPUDDPM, when applied to data over T discrete time steps, can be viewed as a collection of DPMMs (one for each $t \in \{1, \ldots, T \}$), which are linked together by dependencies among their respective parameters.
-For example, the mixture component parameters at a given time step are dependent upon the mixture component parameters in adjacent time steps and the distribution and number of distinct cluster assignments at a given time step are dependent upon the same over a wider range of time steps, upon the distribution of observations' cluster assignments over this range, and upon a deletion procedure, described below.
-
-To link together the parameters of mixture components in adjacent time-steps, a transition kernel $P(\phi_{k}^{t} | \phi_{k}^{t-1})$ is implemented, which provides a distribution over parameters of a mixture component in a given time-step (as a function of equivalent parameters in adjacent time-steps).
-One caveat is that each mixture component must be drawn independently from $\mathbb{G}_{0}$ (the base distribution of the DP, which acts as a prior distribution for the cluster parameters) which we achieve by making $\mathbb{G}_{0}$ the invariant distribution of $P(\phi_{k}^{t} | \phi_{k}^{t-1})$ (note that the transition kernel is a markov chain).
-Additionally, to account for varying numbers of clusters, there is a deletion procedure (described in \cite{caron_2007}) by which observations are considered "removed" from their assigned clusters at a given time-step (which modifies the value of $m_{k}$ defined above in the CRP representation).
-We introduce variables $d_{1}, \ldots, d_{N}$ which denote the times at which the assignments of given observations are considered to be removed.
-With these variables, we can now express the $m_{k}$ value at time $t$ as
-\begin{equation}
-m_{k,t} = \sum_{t' = 1}^{t} \mathbb{I}[(c_{t'}=k) \wedge (t < d_{t'})]
-\end{equation}
-where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise.
-Additionally, for an observation $x_{i}$ at a given time-step $t$, the deletion time $d_{i}$ can be defined to be $d_{i} = t + l_{i}$, where $l_{i}$ is considered the lifespan of an assignment, is distributed geometrically, and can be expressed as
-\begin{equation}
-l_{i} | \rho  \sim  \rho(1 - \rho)^{l_{i}}
-\end{equation} \\
-with parameter $\rho$.
-This process adheres to what \cite{caron_2007} calls a `uniform deletion strategy' over all the observations' assignments (since assignments to each cluster have equal probability of being deleted), though a more complex deletion strategy, dependent upon cluster size, can also be implemented.
-
-Using the cluster size and observation deletion terms introduced above, we can define the GPUDDPM generatively as, for each time step $t = \{1, \ldots, T\}$
-\begin{align}
-\begin{split}
-	x_{i}^{(t)} | c_{i}^{(t)}, \theta_{c_{i}^{(t)}}^{(t)} &\sim F(\theta_{c_{i}^{(t)}}^{(t)}) \\
-	c_{i}^{(t)} | m^{(t-1)}, \alpha  &\sim  \text{CRP}(m^{(t-1)}, \alpha) \\
-	m^{(t-1)} | m^{(t-2)}, c_{i}^{(t-1)}  &\sim \text{DEL}(m^{(t-2)}, c_{i}^{(t-1)}, \rho, \gamma) \\
-	\theta_{k}^{(t)} | \theta_{k}^{(t-1)}   &\sim 
-\begin{cases}
-	P(\theta_{k}^{(t)} | \theta_{k}^{(t-1)}) \\
-	\mathbb{G}_{0}
-\end {cases}
-\end{split}
-\end{align}
-
-The graphical corresponding with this formulation is shown in Figure 2.
-
-\begin{figure}[h]
-        \center{\includegraphics[width=60mm]{gpuddpm_gm_1.pdf}}
-        \caption{\label{Figure 2:} Graphical Model of the Generalized Poly Urn Dependent Dirichlet Process Mixture}
-\end{figure}
-
-\subsubsection{Object Characteristics Modeled}
-
-We would like make explicit the connection between certain elements of the GPUDDPM model and the characteristics of objects in vidoes that these elements represent.
-
-First, the appearances of objects are moded by what we term the `likelihood and object appearance model', which we denoted by $F$.
-In the above modeling formulation, $F$ is the probability distribution for an observation, given that it is assigned to specified cluster (i.e.
-given that it is associated with a given object in a video).
-The likelihood and object appearance model is therefore
-
-\begin{equation}
-F(\theta_{c_{i}}) = P(x_{i}|\theta_{c_{i}})
-\end{equation}
-
-where $x_{i}$ is an observation, $c_{i}$ is its associated assignment, $\theta_{c_{i}}$ specifies the parameters of the parametric form of the $c_{i}^{th}$ cluster.
-The specific form of this model is dependent upon the observations $x_{i}$; recall that the observations, which as a baseline include spatial and temporal features, could include an arbitrary amount of additional features which need to be incorporated into this appearance model.
-A specific formulation of the appearance model can be seen in the proceeding section.
-
-Object behavior is modeled by the dependencies between corresponding clusters at adjacent time steps.
-In particular, this relationship is captured by the transition kernel, $P(\theta_{k}^{t} | \theta_{k}^{t-1})$ (which captures the dependence of cluster $k$ at time $t$ on the same cluster at time $t-1$), described previously.
-The motion behavior of an object, i.e.
-the specific dependence of the parameters of cluster $k$ at a time $t$ on its parameters at time $t-1$, is therefore
-\begin{equation}
-Tr(\theta_{k, t} | \theta_{k, t-1}) = P(\theta_{k, t} | \theta_{k, t-1})
-\end{equation}
-
-Additionally, the base distribution $\mathbb{G}_{0}$ of the DP acts as a prior on the cluster parameters.
-This allows one to place a prior probability over appearances of objects.
-It is important that this prior is not chosen in such a specific manner that it limits this method from performing unsupervised detection of arbitrary objects.
+% \mbox{}
 
 \section{Model Specification and Inference}
-\label{sec:modelspec}
 
@@
 
-\subsection{Model Implementation in Experiments}
-
-The following sections provide details on the data and specific model formulation (determined from the general model formulation of Section~\ref{sec:modeloverview}) involved in the experiments described in Section~\ref{sec:experiments}.
+\subsection{Model Specification for Experiments}
+%
 
@@
 
-The frame differencing and color counting extraction outlined in Section~\ref{sec:modelspec_extraction} yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \mathbb{Z}_{+}$, where each element $\bold{x} \in \bold{X}$ can be written
+The frame differencing and color counting extraction outlined above yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \mathbb{Z}_{+}$, where each element $\bold{x} \in \bold{X}$ can be written
 \begin{equation}
@@
 The prior can thus be written
-\begin{equation} \label{basedistro}
+\begin{equation}
 \mathbb{G}_{0}(\theta) = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}, \Sigma | \boldsymbol{\mu}_{0}, k_{0}, v_{0}, \Lambda_{0})  \mathcal{D}ir( \bold{p} | \bold{q}_{0})
@@
 For each cluster, we introduce $M$ auxiliary variables $z_{t, 1}, \ldots, z_{t, M}$ at time $t$ that are each drawn from the product of a multivariate normal and multinomial when conditioned on the associated cluster parameters $\theta_{t} = \{ \boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t} \}$:
-\begin{equation} \label{transker1}
+\begin{equation}
 z_{t, m} | \boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t}  \sim  \mathcal{N}(\boldsymbol{\mu}_{t}, \Sigma_{t}) \mathcal{M}n(\bold{p}_{t})   \hspace{15pt}   
@@
 To satisfy the above criteria for auxiliary variables, at each time $t$ we specify the dependencies of a given cluster on its associated set of auxiliary variables by
-\begin{equation} \label{transker2}
+\begin{equation}
 \boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t} | \bold{z}_t  \sim  \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}_{M}, k_{M}, v_{M}, \Lambda_{M})  \mathcal{D}ir(\bold{q}_{M})
@@
 
-\subsection{Inference Implementation in Experiments}
+\subsubsection{Previous Model Write-up}
 
-This section provides background on the inference procedure applied to the model and extraction data, which allows for unsupervised detection and tracking of objects.
-A Markov Chain Monte Carlo (MCMC) sampler is used to perform inference, which makes use of the Metropolis-Hastings (M-H) algorithm to sample from a few distributions.
+For a given object $k$ at video frame $t$, the set of positions of the object's motion points are modeled with a product (over each position dimension) of one-dimensional Gaussians
+\begin{equation}
+P(x_{t_{i}}^{pos} | \mu_{t, k}, \lambda_{t, k}) = \prod_{d=1}^{D} \mathcal{N}(x_{t_{i}}^{pos_{d}} | \mu_{t, k}^{d}, \lambda_{t, k}^{d})
+\end{equation}
+where $x_{t_{i}}^{pos} = \{ x_{t_{i}}^{pos_{1}}, \cdots, x_{t_{i}}^{pos_{D}} \}$,  $\mu_{t, k}^{d}$ is the mean of the $k^{th}$ Gaussian at time $t$ for dimension $d$, and $\lambda_{t, k}^{d}$ is the precision of the $k^{th}$ Gaussian at time $t$ for dimension $d$.
+Likewise, for a given object $k$ at video frame $t$, the set of color vectors of the object's motion points are modeled with a multinomial
+\begin{equation}
+P(x_{t_{i}}^{col} | m_{t, k}) = Mult(x_{t_{i}}^{col} | m_{t, k})
+\end{equation}
+where $x_{t_{i}}^{col} = \{ x_{t_{i}}^{col_{1}}, \cdots, x_{t_{i}}^{col_{V}} \}$, $m_{t,k} = \{ m_{t,k}^{1}, \cdots, m_{t,k}^{V} \}$, $\sum_{v=1}^{V} m_{t, k}^{v} = 1$, and $m_{t,k}^{v}>0 \hspace{8pt} \forall v \in \{ 1, \cdots, V \}$.
+The likelihood for a motion point observation is thus
+\begin{equation}
+P(x_{t_{i}} | \mu_{t, k}, \lambda_{t, k}, m_{t, k}) =  Mult(x_{t_{i}}^{col} | m_{t, k})   \prod_{d=1}^{D}  \mathcal{N}(x_{t_{i}}^{pos_{d}} | \mu_{t, k}^{d}, \lambda_{t, k}^{d}) 
+\end{equation}
 
-The generative model associated with the GPUDDPM is illustrated in Section~\ref{sec:gpudpm}.
-This representation includes cluster size $m$ (the number of observations assigned to a given cluster), which is standard in DPM constructions.
-To capture time-dependence, the GPUDDPM allows clusters to `lose assigned observations' in a deletion step, which modifies the sizes $m$ over time.
-Instead of incorporating cluster size variables $m$ directly, we can formulate an equivalent model which makes use of deletion variables $d$, which represent the time when observations are lost from a cluster .
-At each time step, cluster sizes can be recontructed from all previous assignment and deletion variables.
-The MCMC sampler described in this section operates in this alternate representation, which we can formulate as
-\begin{align}
-\begin{split}
-	 \bold{x}_{i} | c_{i}, \theta_{c_{i}} &\sim \mathcal{N}(\boldsymbol{\mu}_{c_{i}}, \Sigma_{c_{i}})  \mathcal{M}n(\bold{p}_{c_{i}}) \\
-	c_{i} | c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha  &\sim  \text{CRP}(c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha) \\
-	d_{i} | \rho  &\sim \text{Geo}(\rho) + i + 1 \\
-	\theta_{k, t} | \theta_{k, t-1}   &\sim 
-\begin{cases}
-	\text{TK} \hspace{12pt} \text{if } k \in \{ 1, \ldots, K_{t-1}  \} \\
-	\mathbb{G}_{0}  \hspace{15pt} \text{if } k = K_{t-1} + 1
-\end {cases}
-\end{split}
-\end{align}
-where TK stands for the transition kernel, given in \eqref{transker1} and \eqref{transker2}, and the base distribution $\mathbb{G}_{0}$ is given in \eqref{basedistro}.
-This formulation of the GPUDDPM is also used in \cite{gasthaus_thesis} and \cite{caron_2007}.
-The associated graphical model for this formulation is given in \willie{cite figure}.
+I want to justify distribution families chosen for appearance modeling in this implementation.
+To include: the edges of an object are detected well with frame differencing, so while extraction data densities aren't necessarily Gaussians with maxima over the targets, they are usually radially centered around the centroids of targets.
+Furthermore (thing frank said about gaussians and what they do---minimize least squared distance or something).
 
-We define the state of the sampler to consist of the assignment variables for all observations, $c_{1}, \ldots, c_{N}$, the deletion variables for all observations, $d_{1}, \dots, d_{N}$, the sequence of cluster parameters for each active cluster $k$, $\theta_{k,1}, \ldots, \theta_{k, T}$, and the $M$ associated auxiliary variables for each cluster $\theta_{k,t}$, $z_{k,t,1}, \ldots, z_{k, t, M}$ (where $N$ denotes the number of observations and $T$ denotes the number of time steps).
+\subsubsection{Base Distribution $\mathbb{G}_{0}$ and Appearance Prior}
+$\mathbb{G}_{0}$ is the base distribution of the underlying time-dependent Dirichlet process mixture; it also serves as a prior distribution for the parameters present in the likelihood.
+We make use of conjugate priors in the base distribution to allow for more efficient computation.
+In particular, for an object $k$ at time $t$, we let the prior distribution over the parameters of the motion point position distribution, $\mu_{t, k}$ and $\lambda_{t, k}$, be
+\begin{equation}
+\mathbb{G}_{0}^{pos}(\mu_{t, k}, \lambda_{t, k} | \mu_{0}, n_{0}, a, b) = \prod_{d=1}^{D}[\mathcal{N}(\mu_{t, k}^{d} | \mu_{0}, n_{0} \lambda_{t, k}^{d})   Ga(\lambda_{t, k}^{d} | a, b)]
+\end{equation}
+where $\mu_{0}, n_{0}, a,$ and $b$ are parameters of the base distribution.
+Furthermore, for object $k$ at time $t$, we let the prior distribuion over the parameters of the motion point color vector distribution be
+\begin{equation}
+\mathbb{G}_{0}^{col}(m_{t, k} | q) = Dir(m_{t, k} | q)
+\end{equation}
+where $q = \{ q^{1}, \cdots, q^{V} \}$ is a parameter of the base distribution, where $q^{v}>0 \hspace{8pt} \forall v \in \{ 1, \cdots, V \}$. The base distribution is thus 
+\begin{equation}
+\mathbb{G}_{0}(\mu_{t, k}, \lambda_{t, k}, m_{t, k} | \mu_{0}, n_{0}, a, b, q) = Dir(m_{t, k} | q)  \prod_{d=1}^{D}[\mathcal{N}(\mu_{t, k}^{d} | \mu_{0}, n_{0} \lambda_{t, k}^{d})   Ga(\lambda_{t, k}^{d} | a, b)]
+\end{equation}
 
-The sampler moves sequentially through each of the $N$ observations, sampling the assignment $c_{i}$ and the deletion variable $d_{i}$ at each step.
-After the final observation at a given time $t$ step is sampled, the cluster parameters for all active clusters at $t$ are sampled, and a M-H step is employed to sample the $M$ auxiliary variables for each active cluster.
-How each sample is generated will be explicitely shown in the following sections.
-
-\subsection{Modeling Background Noise}
-\willie{Do we think this is worth it? I can make extraction pretty clean on most datasets and this adds complication to an already complicated model}
-
-Noisy extraction procedures may result in erroneous data points that do not correspond to any foreground object.
-We would like to incorporate these points into our model.
-To do this, we modify our base distribution $\mathbb{G}_{0}$
-
-\section{Experiments}
-\label{sec:experiments}
-
-The following sections provide details on performance evaluation metrics that have been developed for use in object detection and tracking studies (and that are used to quantify experimental results in this paper), synthetic video experiments that verify certain aspects of the technique presented in this paper, benchmark video experiments that demonstrate the performance of this technique in relation to other methods developed in recent years, and nonstandard video datasets to show the capability of this technique to perform unsupervised detection and tracking on a diverse range of object and video types.
-
-\subsection{Performance Evaluation Metrics}
-
-Performance evaluation metrics, which intend to provide a standardized way of quantifying the success of a detection and tracking procedure on a given video, have started to become well studied and consistently used in the past four years.
-The metrics developed in \cite{kasturi_2008} and used in \cite{ellis_2010, taj_2007, lee_2009} have become well established metrics for evaluating the performance of object detection and tracking in videos and have been adopted by the Video Analysis and Content Extraction (VACE) program and the Classification of Events, Activities, and Relationships (CLEAR) consortium, two large projects concerned with video tracking.
-These metrics are used to quantify the experimental results in this study.
-
-The above metrics are dependent upon ground-truth data specifying the positions of each object in each frame throughout a video sequence.
-In the conducted experiments, we gained the synthetic video ground-truth during construction of the videos (described in Section~\ref{sec:syntheticvideos}), and we attempted to choose benchmark datasets with provided ground-truth data.
-However, the final benchmark dataset, in addition to all nonstandard video datasets, did not include ground-truth data; for these datasets, the Video Performance Evaluation Resource (ViPER) ground-truth tool \cite{doermann_2000} was used to gain ground-truth data.
-
-\subsubsection{Mapping Ground-Truth Tracks to Output Tracks}
-
-The problem of finding a mapping between a video's ground-truth tracks and an algorithm's output tracks is nontrivial, though necesary to solve in order to compute the performance evaluation metrics used in this study.
-In short, the typical solution to this problem involves first specifying a performance metric and then choosing the mapping from ground-truth tracks to output tracks which yields the most favorable performance metric value.
-This process is described in detail by Kasturi et al.
-\cite{kasturi_2008}; we follow the method outlined in this paper to find an optimal mapping.
-Additionally, the method employed in \cite{kasturi_2008} allows erroneous and undetected tracks to be left unmapped, which is both desired and necessary in the case where there is a different number of ground-truth and output tracks.
-Note that once a mapping from a collection of ground-truth tracks to a collection of output tracks has been established, one can determine which output tracks are false positives (the output tracks to which no groundtruth track is assigned) and which input tracks are false negatives (the input tracks that are not assigned to an output track).
-The numbers of tracks displaying both of these failures are factors in the performance metrics used in this study.
-
-%, where a numerical search algorith--the Hungarian algorithm \cite{munkres_1957}--is used to find an optimal mapping from ground-truth to output tracks without requiring the performance metric value of all possible mapping combinations to be computed (reducing the time complexity from factorial to polynomial time)
-
-\subsubsection{MODA and MOTA}
-
-The two metrics used to quantify results in this study are refered to as the Multiple Object Detection Accuracy (MODA) and Multiple Object Tracking Accuracy (MOTA).
-These metrics, along with the the ground-truth tracks of all objects in the video sequence and the the mapping as defined in the previous section, can be used to evaluate detection and tracking accurancy of the video.
-For a given video frame $t$, we define MODA as
+\subsubsection*{Transition Kernels and Motion Behavior}
+Let $\phi_{t,k}^{pos} = \{ \mu_{t, k}, \lambda{t, k} \}$ and $\phi_{t,k}^{col} = \{ m_{t, k} \}$.
+We define two transition kernels, $P(\phi_{t, k}^{pos} | \phi_{t-1, k}^{pos})$ and $P(\phi_{t, k}^{col} | \phi_{t-1, k}^{col})$, which dictate, respectively, the random walk of the motion point position parameters and the motion point color vector parameters over time.
+The transition kernels must be chosen so that their invariation distributions are $\mathbb{G}_{0}$, i.e.
+such that the following hold:
 \begin{equation}
-MODA = 1 - \frac{m^{t} + e^{t}}{N_{g}^{t}}
+\int \mathbb{G}_{0}^{pos}(\phi_{t-1, k}^{pos}) P(\phi_{t, k}^{pos} | \phi_{t-1, k}^{pos}) d\phi_{t-1, k}^{pos}  =  \mathbb{G}_{0}^{pos}(\phi_{t, k}^{pos})
 \end{equation}
-where $m^{t}$ is the number of missed object detections (ie true-negative detections), $e^{t}$ is the number of erroneous object detections (ie false positive detections), and $N_{g}^{t}$ is the number of ground truth objects (all at frame $t$).
-
-For a sequence of video frames $t \in \{ 1, \ldots, T \}$ we define MOTA as
 \begin{equation}
-MOTA = 1 - \frac{\sum_{t=1}^{T} m^{t} + e^{t} + s_{t}}{\sum_{t=1}^{T} N_{g}^{t}}
+\int \mathbb{G}_{0}^{col}(\phi_{t-1, k}^{col}) P(\phi_{t, k}^{col} | \phi_{t-1, k}^{col}) d\phi_{t-1, k}^{col}  =  \mathbb{G}_{0}^{col}(\phi_{t, k}^{col})
 \end{equation}
-where $s$ denotes the number of instances at frame $t$ where an object's identity switches from a previous frame (ie where the object's mapping differs between frame $t$ and frame $t-1$),  and the other terms are defined the same as above.
-
-\subsection{Synthetic Video Datasets}
-\label{sec:syntheticvideos}
-
-Synthetic videos provide controlled scenarios in which aspects of the technique presented in this paper can be tested.
-Each of the following synthetic videos consists of a sequence of 200 images (each of size $500$ x $500$ pixels) containing a number of smaller colored squares of different (and potentially time-varying) sizes moving at varied speeds and trajectories over a black background.
-The synthetic videos contain instances of occlusion (where one or more objects are briefly hidden) and objects with time-varying appearances and behaviors, as these are examples of occurances that notoriously decrease the accuracy of detection and tracking.
-After each video was constructed, the extraction and inference procedures described in Section~\ref{sec:modelspec} were carried out to return a collection of centroid-position-sequences, each corresponding to the sequence of positions of a tracked object over each frame that it is present in the video.
-
-\subsubsection{Tracking by Appearance}
-
-The first synthetic video experiment aimed to test the ability of the model and inference procedure to maintain the identity of independent objects based on color information alone.
-Two videos were constructed, both containing a red square (rgb value $[255,0,0]$ and size $20$ x $20$ pixels) and a blue square (rgb value $[0,0,255]$ and size $20$ x $20$ pixels).
-In both videos, the squares begin at opposite sides of the scene at frame $f=1$ and travel towards each other, arriving at the same location at $f=100$ (where the blue square occludes the red square).
-The second half of the two videos differ in that both squares in the first video continue in the same direction and end at the other's starting positions at $f=200$, and both squares in the second video reverse directions and end at their initial starting positions at $f=200$.
-Frame difference extraction on each video yields a dataset with identical spatial features; hence, successful tracking depends fully on the reasonable incorporation of color information into the model.
-
-\subsubsection{Occlusion and Appearance Change}
-
-The second synthetic video experiment aimed to test tracking performance under occlusion, object appearance change, and motion change.
-A video was constructed, containing a red square (rgb value $[255,0,0]$), a green square (rgb value $[0,255,0]$), and a blue square (rgb value $[0,0,255]$).
-The red square was of size $20$ x $20$ pixels, the blue square was of size $15$ x $15$ pixels, and the green square began at size $50$ x $50$ pixels at frame $f=1$, linearly shrinks to $10$ x $10$ pixels at $f=100$, then linearly grows back to $50$ x $50$ pixels by the end of the video, $f=200$.
-Furthermore, the red and blue squares display the same behavior as in the second video of the first synthetic experiment (they begin at opposite sides of the scene traveling towards each other, cross at the center of the scene at $f=100$, and reverse direction, ending at their initial positions at $f=200$).
-The green square begins at a point equidistant from the other two squares, intersects with them as they overlap (causing the blue square to occlude the other two), and continues on in a direction at a 20 degree angle from its initial trajectory.
-
-% \subsubsection{Similar Appearances}
-
-% \subsubsection{Different Numbers of Objects}
-
-% \subsubsection{Objects wih Diverse Characteristics}
-
-\subsection{Benchmark Video Datasets}
-
-Benchmark video datasets for object tracking and detection have been produced to provide standard scenes on which researchers can compare detection and tracking results.
-These videos have been primarily produced for surveillance-related workshops--notably, for the International Workshop on Performance Evaluation of Tracking and Surveillance (PETS)--which provide a setting where researchers can focus on a collection of videos and computer vision goals.
-Four commonly used benchmark videos from four PETS workshops are
-
-\subsubsection{PETS2000}
-
-\subsubsection{PETS2001 or VS-PETS/Soccer}
-
-\subsubsection{PETS2004/Caviar}
-
-\subsubsection{PETS2009/2010}
-
-\subsection{Nonstandard Video Datasets}
-
-\subsubsection{Ants Videos}
-
-\subsubsection{}
-
-\section*{Results and Discussion}
-
-\section*{Conclusion}
-
-\begin{small}
-\bibliographystyle{plainnat}
-\bibliography{paper_refs} 
-\end{small}
 

### 2011-12-01 ltx/paper.tex
am writing paper

@@
 
-\abstract{main points}
+\abstract{Main points.}
 
@@
 
-We define the task of automated detection and tracking of arbitrary objects in videos to be the fulfillment of three actions: deciding which spatiotemporal regions of a video constitute objects (`extraction'), finding the positions and/or shapes of distinct objects (`localization'), and maintaining the positions of the detected objects over time (`tracking').
-Robust algorithms capable of carrying out these tasks are applicable to problems in the fields of surveillance, autonomous systems and robotics, and video summarization and compression.
-Approaches to these tasks have been proposed under different headings and in a variety of contexts; this paper serves to introduce a new technique involving the application of a dependent Dirichlet processes mixture model, which we hope will provide the foundation for a number of robust, unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.
+We can define the task of automated detection and tracking of arbitrary objects in videos to be the fulfillment of three actions: differentiating the spatiotemporal regions of a video that constitute objects from the region that constitutes the background (`extraction'), finding the positions and/or shapes of distinct objects (`localization'), and maintaining the positions of the detected objects over time (`tracking').
+An ability to carry out these tasks in an automated manner is useful for many disciplines that make use of video data, including the fields of surveillance, autonomous systems, robotics, computational biology, and video summarization.
+Algorithms of this sort that prove to be efficacious for many videos and a variety of objects are particularly desirable.
+This paper serves to introduce a new technique involving the application of a dependent Dirichlet processes mixture model, which we hope will provide the foundation for a number of robust, unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.
 
-There does not yet exist a major body of research concerned primarily with methods for combining extraction, localization, and tracking to perform cohesive unsupervised detection of arbitrary objects in videos.
-Many papers that attempt to accomplish this goal focus on one of extraction, localization, or tracking, and attempt to integrate in methods from the other two tasks.
-However, a few key schemes for unsupervised detection and tracking do indeed exist.
-A widely applied, though unsophisticated, example is blob tracking.
-Blob tracking typically refers to methods that perform extraction and localization at each frame independently--for example, by assuming that object pixels have a different intensity than background pixels, setting a fixed pixel intensity threshold to extract pixel locations, and segmenting the extracted data into `blobs' to accomplish localization.
+Research over the past ten years related to the development of methods for general detection and tracking of objects has tended to focus on one of extraction, localization, or tracking; attempts to integrate in the other two and produce a full system able to handle arbitrary objects and videos was often not a primary focus, and was often attempted afterwards, in an ad-hoc manner.
+Research aimed at developing methods to accomplish the three tasks in a cohesive manner have been pursued more recently to produce algorithms that can, in a fully automated way, accomplish the goal of deciding which spatiotemporal regions of a video constitute distinct objects.
+A widely applied, though unsophisticated, example is the technique of blob tracking.
+Blob tracking typically refers to methods that perform extraction and localization of objects at each frame independently--for example, by assuming that object pixels have a different intensity than background pixels, setting a fixed pixel intensity threshold to extract pixel locations, and segmenting the extracted data into `blobs' to accomplish localization.
 A number of tracking techniques based on finding blobs in consecutive frames of a video with similar appearances have been attempted to maintain the identity of detected blobs over time.
 The main challenges, however, associated with blob tracking involve the difficulty of segmenting occluded objects in a given frame \cite{zhao2004tracking} and preventing multiple tracked blobs from collapsing onto and tracking a single object \cite{vermaak_2003}.
-These difficulties stem from the typical blob filter's need to perform detection via segmentation on individual frames and lack of a principled way to resist the collapse of independent tracks.
+These difficulties stem from the typical blob filter's need to perform detection via segmentation on individual frames and lack of a principled way to perform tracking while resisting the collapse of independent tracks.
 
@@
 
-The PETS2009/2010 dataset consists of
+The PETS2009/2010 dataset consists of a 900 frame video.
 

### 2011-12-01 ltx/paper.tex
still messy, pushing in case you need to pull before your internet leaves
a little better formatted

@@
-\documentclass{article}
+\documentclass[smallcondensed, final]{svjour3}
 
@@
 \DeclareMathOperator*{\argmax}{arg\,max}
-
-% need to include style sheet for journal
 
@@
 The data consists of the three-dimensional baseline features and was gained by extraction peformed on a 50 frame sequence showing five ants moving across the video scene.
-\begin{figure}[h]
+\begin{figure*}[h]
         \center{\includegraphics[width=85mm]{../img/usbh_data_2.png}}
@@
 Markers are colored by position for ease of viewing.}
-\end{figure}
+\end{figure*}
 Inference performed on the time dependent Dirichlet process mixture model (described in the following sections) carries out clustering of the data, which allows each worm-like data cloud to be isolated and independent objects to be localized and tracked.

### 2011-12-01 ltx/paper.tex
some more edits

@@
 This paper serves to introduce a new technique involving the application of a dependent Dirichlet processes mixture model, which we hope will provide the foundation for a number of robust, unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.
+\willie{[i like def's in 1st paragraph, but we maybe need a better opening line hook?]}
 
-Research over the past ten years related to the development of methods for general detection and tracking of objects has tended to focus on one of extraction, localization, or tracking; attempts to integrate in the other two and produce a full system able to handle arbitrary objects and videos was often not a primary focus, and was often attempted afterwards, in an ad-hoc manner.
+Research in the past ten years involving the development of methods for general detection and tracking of objects has tended to focus on one of either extraction, localization, or tracking; attempts to integrate in the other two and produce a full system able to handle arbitrary objects and videos was often not a primary focus, and was often attempted afterwards, in an ad-hoc manner.
 Research aimed at developing methods to accomplish the three tasks in a cohesive manner have been pursued more recently to produce algorithms that can, in a fully automated way, accomplish the goal of deciding which spatiotemporal regions of a video constitute distinct objects.

### 2011-12-04 ltx/paper.tex
been working on draft of paper... better intro, edits throughout

--- 2011-12-01 ltx/paper.tex
@@
 \documentclass[smallcondensed, final]{svjour3}
+% \documentclass[twocolumn, final]{svjour3}
 
@@
 
-We can define the task of automated detection and tracking of arbitrary objects in videos to be the fulfillment of three actions: differentiating the spatiotemporal regions of a video that constitute objects from the region that constitutes the background (`extraction'), finding the positions and/or shapes of distinct objects (`localization'), and maintaining the positions of the detected objects over time (`tracking').
-An ability to carry out these tasks in an automated manner is useful for many disciplines that make use of video data, including the fields of surveillance, autonomous systems, robotics, computational biology, and video summarization.
-Algorithms of this sort that prove to be efficacious for many videos and a variety of objects are particularly desirable.
-This paper serves to introduce a new technique involving the application of a dependent Dirichlet processes mixture model, which we hope will provide the foundation for a number of robust, unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.
+We define the task of automated detection and tracking of arbitrary objects in videos to be the fulfillment of three actions: determining the spatiotemporal regions of a video that constitute objects (`extraction'), finding the positions and/or shapes of distinct objects (`localization'), and maintaining the identities of the detected objects over time (`tracking').
+An ability to carry out these tasks in an automated manner is useful for many fields that make use of video data, including robotics, video surveillance, computational biology, and video summarization.
+Algorithms proving to be efficacious for a variety of objects and video types are particularly applicable.
+This paper serves to introduce a new technique involving the use of dependent Dirichlet processes mixture models, which we hope will provide the foundation for a class of unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.
 
-Research over the past ten years related to the development of methods for general detection and tracking of objects has tended to focus on one of extraction, localization, or tracking; attempts to integrate in the other two and produce a full system able to handle arbitrary objects and videos was often not a primary focus, and was often attempted afterwards, in an ad-hoc manner.
-Research aimed at developing methods to accomplish the three tasks in a cohesive manner have been pursued more recently to produce algorithms that can, in a fully automated way, accomplish the goal of deciding which spatiotemporal regions of a video constitute distinct objects.
-A widely applied, though unsophisticated, example is the technique of blob tracking.
-Blob tracking typically refers to methods that perform extraction and localization of objects at each frame independently--for example, by assuming that object pixels have a different intensity than background pixels, setting a fixed pixel intensity threshold to extract pixel locations, and segmenting the extracted data into `blobs' to accomplish localization.
-A number of tracking techniques based on finding blobs in consecutive frames of a video with similar appearances have been attempted to maintain the identity of detected blobs over time.
-The main challenges, however, associated with blob tracking involve the difficulty of segmenting occluded objects in a given frame \cite{zhao2004tracking} and preventing multiple tracked blobs from collapsing onto and tracking a single object \cite{vermaak_2003}.
+Research in the past ten years involving the development of methods for general detection and tracking of objects has tended to focus on one of either extraction, localization, or tracking; attempts to integrate all three tasks and produce a system aimed at multiple arbitrary objects and diverse video types was not often a primary focus, or was attempted in an ad-hoc manner to test or apply a more specific technique.
+Research aimed at developing methods to accomplish the three tasks in a cohesive manner has been pursued more recently to produce algorithms that can, in a fully automated way, accomplish unsupervised detection and tracking--or in short, that can determine the spatiotemporal region of a video occupied by each distinct object.
+A widely applied, though unsophisticated, example of general detection and tracking is the technique of blob tracking.
+Blob tracking typically refers to methods that perform extraction and localization of objects at each frame independently--for example, by assuming that pixels representing an object are distinguishable from pixels representing the background via color, motion, or texture information, and segmenting the extracted data at each frame into `blobs' to accomplish localization.
+A number of tracking techniques based on finding blobs in consecutive frames of a video with similar locations and appearances have been attempted to maintain the identity of detected blobs over time.
+The main challenges, however, associated with blob tracking involve the difficulty of segmenting neighboring or occluding objects in a given frame \cite{zhao2004tracking}, preventing multiple tracked blobs from collapsing onto and tracking a single object \cite{vermaak_2003}.
 These difficulties stem from the typical blob filter's need to perform detection via segmentation on individual frames and lack of a principled way to perform tracking while resisting the collapse of independent tracks.
 
-%Note that blob tracking carries out the three main tasks described above, so it satisfies the criteria for the type of unsupervised detection and tracking algorithms studied in this work;
-
-This paper presents the use of Dirichlet process mixture models as a technique for unsupervised multiple object detection and tracking, which we intend as a method to allow for robust object detection (particularly in cases where frame-by-frame segmentation would be difficult) and to keep the tracks of independent objects isolated.
-Specifically, this work applies a type of Bayesian nonparametric model to data gained during extraction; note that data of this type is refered to in the paper as `extraction data', `data points', or `observations'.
-The model is first presented in an abstract form, which could be adapted to a wide range of extraction procedures and object appearance representations (mainly left for exploration in future studies).
-We then describe inference algorithms that allow object localization and tracking to be carried out.
-Afterwards, we formulate a definite model for data gained during a specific extraction procedure and implement the described inference procedures to carry out multiple object detection and tracking in a number of experiments.
+This paper presents the use of a Bayesian nonparametric mixture model for unsupervised multiple object detection and tracking.
+This method is intended to allow for robust detection of arbitrary objects (particularly in cases where frame-by-frame segmentation is difficult) and better isolation of independent objects during tracking.
+Specifically, this work describes how a Dirichlet process mixture model can accomplish these goals when it is applied to data that may be extracted in a general manner from videos.
+We will demonstrate how this generative model may be applied to types of noisy data that can be quickly extracted from objects in videos, and how this model allows for automated detection and tracking.
+The model is first presented in an abstract form, to allow for it to be used with arbitrary extraction procedures and object appearance representations that may be desired in future studies.
+We then formulate a specific model for data gained during a basic extraction procedure and describe inference algorithms that allow object localization and tracking to be carried out.
+We demonstrate our implementation on a number of synthetic and benchmark datasets and describe and compute performance metrics to quantify results.
 
 \subsection{Related Work}
+\label{sec:relatedwork}
 
-Related work can be broken down into studies relating to extraction procedures, localization techniques, single and multiple object tracking methods, and combination methods for joint object detection and tracking.
+Related work involves studies related to extraction procedures, localization techniques, single and multiple object tracking methods, and combination methods for joint object detection and tracking.
 
-The task of extraction incorporates the tasks of foregound segmentation, background modeling, motion detection, and appearance feature extraction.
+The task of extraction is related to the tasks of foregound segmentation, background modeling, motion detection, and appearance feature extraction.
 In general, these methods aim to extract features from regions of a video frame that are considered objects or are distinguished in some way from the background of a scene.
@@
 For example, foreground extraction has been robustly carried out by modeling each pixel over time in order to infer a probability distribution over pixel values; if the pixel takes on a very unlikely value at a future time, it is marked as a foreground pixel at that point \citep{stauffer_1999, elgammal_2000, elgammal_2002}.
-These pixel modeling techniques have been adapted to allow for time-dependent probability distributions over pixel values, which help account for slighlty changing backgrounds, such as those caused from brief movements of background objects (for example, trees in outdoor scenes).
 Note that these techniques treat each pixel independently; this has a tendency to cause errors in foreground detection, resulting in fragmentation of foreground images (which is problematic as many of these methods intend to blob the foreground objects after detection in order to perform segmentation and locate objects' centroids).
-Post processing has been applied to add information regarding the edges of objects in an attempt to decrease fragmentation \citep{turdu_2007}.
-Others have attempted to achieve better foreground modeling in cases where the foreground and background exhibit similar color distributions (the so called `color similarity problem') by, for example, incorporating models of the foreground into a typical background model (such as the one described above) \citep{gallego_2009, mchugh_2009}.
-
-On the other hand, heuristic-based methods perform extracting by looking for temporal changes in pixel value characteristics.
-While these methods locate regions with objects less accurately than model-based methods, they are often much less computationally expensive.
+Heuristic-based methods perform extracting by looking for temporal changes in pixel value characteristics.
+These methods are often less accurate than model-based methods, but also far less computationally expensive.
 Frame differencing and background subtraction are two such methods, and involve comparing pixel values, or groups of pixel values, between two images, to detect a change surpassing some threshold.
@@
 Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking schemes \citep{perez_2002}.
-In addition to the two techniques mentioned above, there exists other ways of extracting movement from an image, such as through techniques involving optical flow (a calculation based on a so called `generalized gradient model' of an image, which attempts to capture the speed and direction of movement over areas in the image) \citep{horn_1981, bobick_2001}.
+In addition to the two techniques mentioned above, there exist other ways of extracting movement from an image, such as through techniques involving optical flow (a calculation based on the `generalized gradient model' of an image, which attempts to capture the speed and direction of movement over areas in the image) \citep{horn_1981, bobick_2001}.
 In \citep{black_2000}, a motion detection and object tracking method is developed based on the relative motion of moving objects' boundaries, which are again found through analysis of optical flow.
 
-The task of localization is closely related to the topics of object detection, position-finding, target segmentation, contour-detection, template matching, and recognition.
+The task of localization is closely related to the topics of object detection, position-finding, segmentation, and target recognition.
 These methods are mainly concerned with finding the positions and/or shapes of objects in a given video frame, or with discerning between objects similar in position or appearance.
 There has been a focus in this field on developing techniques for representing an object's appearance, both to allow for localization of objects on a per-frame basis and to allow for consistent localization of a given object over time.
-Research into the unsupervised detection of arbitrary objects (which we define as methods that accomplish both extraction and localization) has included work involving image segmentation, parsing, and pixel clustering.
+Research into the unsupervised detection of arbitrary objects (which we define as methods that accomplish both extraction and localization) has included work involving image segmentation, parsing, and pixel clustering \willie{must cite here}.
 These topics usually relate to the automated partitioning of images into meaningful regions.
@@
 The task of tracking incorporates the topics of video tracking, data association, filtering, point matching, and `tracking of non-rigid objects'.
-In general, these tasks aim to maintain localization of the positions and/or shapes of objects over time.
+In general, these tasks aim to maintain the localization of the position and/or shape of an object over time.
 Areas of interest in this field involve developing metrics for judging the similarity between two object representations (for example, potential representations of the same object at adjacent frames in a video), techniques that provide a way to conduct efficient searches over a frame for regions that are similar to a given object, and methods for predicting the future position or appearance of an object.
-A great deal of research in this field over the past decade has focused on the development of algorithms to track multiple objects simultaneously.
-There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera), complex object interactions, object coalescence (where one object is tracked by multiple times/redundantly), objects with similar appearances, variable (and potentially high) numbers of objects, and objects that enter and exit a field of view at different times.
 
-There are an extensive collection of methods for tracking single, non-rigid, objects.
-For example, color has long been used for tracking.
-In this case, tracking refers to the task of maintaining the location of an object after its initial position has been specified.
-In general, these schemes operate by modeling an object with some color-based appearance model, and using this model to find the object in subsequent frames.
+Methods for tracking single, non-rigid, objects are numerous.
+Color, for example, has long been used for tracking.
+These techniques often operate by modeling an object with some color-based appearance model, and using this model to find the object in subsequent frames.
 One classic approach involves modeling the color of a target (in particular, a distribution over the hue-saturation space of a target region) with a Gaussian mixture model (GMM), and choosing subsequent target positions by searching surrounding areas for regions that yield similar GMMs.
 Furthermore, the GMM is often allowed to adapt over time, and slowly change to model changes in lighting or other smooth time-based variations in the target's color \citep{raja_1998, mckenna_1999, jepson_2003}.
-Others have attempted to abstract this work with a non-parametric color modeling approach based on kernel density estimation, which does not assume a specific underlying distribution (such as the Gaussian mixture in the previous case) and instead converges to reasonable distribution that depends on the data \citep{elgammal_2001}.
-Additionally, color has been successfully applied as the feature in the mean-shift procedure (kernel-based method) for tracking \citep{comaniciu_2003, perez_2002, nummiaro_2003, lee_2011}.
+Additionally, color has been successfully applied as the feature in a technique known a the mean-shift procedure (or the kernel-based method) for tracking \citep{comaniciu_2003, perez_2002, nummiaro_2003, lee_2011}.
+This procedure--a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--optimizes the search for the region in a frame which is most similar to a target region in the previous frame.
+It involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges for each data point.
+Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors.
+It can be shown that, through this procedure, each data point becomes associated with a local point of high density (dependent upon the underlying weight distribution specified by the kernel) which naturally allows for clustering \citep{cheng_1995}.
+The mean shift algorithm has been implemented successfully to allow for a sort of kernel-based object tracking \citep{comaniciu_2003, comaniciu_1999, comaniciu_2000}.
 
-%
+In addition to single object tracking, a great deal of research in this field over the past decade has focused on the development of algorithms to track multiple objects simultaneously.
+There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera), complex object interactions, object coalescence (where one object is tracked by multiple times/redundantly), objects with similar appearances, variable (and potentially high) numbers of objects, and objects that enter and exit a field of view at different times \willie{cite paper(s) for each of these}.
+
+In the past five years, there has been a push for methods that incorporate detection and tracking of multiple objects to allow for full systems capable of autonomously handling real-world video datasets.
+\willie{Describe tracking by detection, more on blob tracking, the other few methods i've found for automated joint detection and tracking}
+
+% Others have attempted to abstract this work with a non-parametric color modeling approach based on kernel density estimation, which does not assume a specific underlying distribution (such as the Gaussian mixture in the previous case) and instead converges to reasonable distribution that depends on the data \citep{elgammal_2001}.
 
@@
 %
-This section provides details on the nature of the data to be modeled, Dirichlet process mixture (DPM) models, and dependent Dirichlet process mixture (DDPM) models.
-We end by making explicit the way in which characteristics of video objects can be represented by elements of these models.
-We formulate the model in a sufficently abstract manner to allow for specific implementations tailored towards specified objects in future studies.
+This section provides details on the data to be modeled, the Dirichlet process mixture (DPM) model, and the dependent Dirichlet process mixture (DDPM) model.
+We make explicit the way in which characteristics of video objects can be represented by elements of these models.
+We formulate the model in a sufficently abstract manner to allow for implementations tailored for specified objects in future studies.
 
@@
 %
-We defined the task of extraction to be the unsupervised decision of which spatiotemporal regions of a video constitute objects; hence, the extraction data modeled in this work are D-dimensional observations that in some way provide information about the regions of the video frame that contain objects.
-A simple example of this information is shown by the three-dimensional extraction data $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $x_{1}$ and $x_{2}$ are the two spatial dimensions of pixels contained in regions comprising objects in a given frame, and $t$ is the discrete time index of the frame.
-These three features are always available once regions containing objects are discerned, and should always be kept as features of the data for modeling (as position information is key for reasonable tracking and time information is necessary for our intention to model the temporal dependencies of the data).
+We defined the task of extraction to be the unsupervised determination of which spatiotemporal regions of a video constitute objects, and outlined a number of extraction methods in Section~\ref{sec:relatedwork}.
+The features of the data are characteristics of pixels in the regions of the video that contain objects, and the data are gained during extraction.
+For example, at its most basic, an extraction dataset is comprised of the three-dimensional points $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $x_{1}$ and $x_{2}$ are the two spatial dimensions of a pixel contained in a region representing an object in a given frame, and $t$ is the discrete time index of the frame.
+These three features are always available once regions containing objects are discerned, and are necessary features for modeling (as position information is key for reasonable tracking and time information is necessary for maintaining the temporal dependencies in the data).
 
-Examples of additional features that could potentially be extracted from video frame regions containing objects include color information, pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, texture representations, transform results (such as results of hough or fourier transforms), and quantiative region descriptions derived from image segmentation procedures; the main idea is to choose features which are able to be easily extracted or computed, capture variability in the appearance of objects, are applicable to a wide variety of object types, and, when viewed in cumulation for multiple observations, are distributed in a way that allows for modeling and tractable inference (this is discussed further in section 3).
+Examples of additional features that could potentially be extracted from object-regions in video frames include color information, pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, texture representations, transform calculations (such as results of hough or fourier transforms), and quantiative region descriptions derived from image segmentation procedures; the main idea is to choose features which are able to be easily extracted or computed, capture variability in the appearance of objects, are applicable to a wide variety of object types, and can be represented by distributions that allow for tractable inference (discussed further in Section~\ref{sec:inference_spec}).
 
 We will use the phrase `baseline' three-dimensional extraction data to refer to the $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$ from pixels present in regions containing objects.
-These data are distributed roughly as worm-like shapes that follow the paths of individual objects as they move, when they are plotted in three dimensions (where we let the frame number, our discrete representation of time, ascend the vertical axis and the spatial positions occupy the horizontal axes).
-An example of data with such a distribution can been seen in Figure \ref{fig:baselinedata}.
+These data are distributed roughly as worm-like shapes that follow the paths of individual objects as they move, when they are plotted in three dimensions (where we let the frame number $t$, our discrete representation of time, ascend the vertical axis and the spatial positions occupy the horizontal axes).
+An example of data with such a distribution can been seen in Figure~\ref{fig:baselinedata}.
 The data consists of the three-dimensional baseline features and was gained by extraction peformed on a 50 frame sequence showing five ants moving across the video scene.
@@
         \caption{\label{fig:baselinedata} Baseline three-dimensional extraction data gained from a video segment containing 5 moving ants.
-The vertical axis denotes the frame number and the horizontal axes denote $d_{1}$ and $d_{2}$ spatial coordinates, respectively, of pixels within a given frame.
+The vertical axis denotes the frame number and the horizontal axes denote $d_{1}$ and $d_{2}$, the spatial coordinates of pixels within a given frame.
 Markers are colored by position for ease of viewing.}
 \end{figure*}
-Inference performed on the time dependent Dirichlet process mixture model (described in the following sections) carries out clustering of the data, which allows each worm-like data cloud to be isolated and independent objects to be localized and tracked.
+Inference performed on the time dependent Dirichlet process mixture model (described in Section~\ref{sec:inference_spec}) carries out clustering of the data, which isolates each worm-like data cloud and allows each object to be localized and tracked.
 
@@
 In this work, estimating the number of clusters in the data generated via extraction is shown in the following sections to be equivalent to estimating the number of distinct objects in a video.
-Consequentially, a model which is able to perform robust nonparametric estimation of the number of cluster is well suited to be applied to the task of unsupervised object detection and tracking.
+Consequentially, a model that is able to perform robust nonparametric estimation of the clusters in an extraction dataset may be applied to perform unsupervised object detection and tracking.
 
 The following sections are intended to provide a solid introduction to DPMs without a major foray into the theoretical aspects of the model that require a more rigorus technical setup.
-We give background on finite mixture models, Bayesian finite mixture models, and Dirichlet processes, presently, which allows us to subsequently describe DPMs in context.
+We give background on finite mixture models, Bayesian finite mixture models, and Dirichlet processes, which allows us to subsequently describe DPMs in context.
 
@@
 The finite mixture model can be written as
-
 \begin{equation}
@@
 \end{equation}
-
-where $c_{i} \in \{ 1, \ldots, K \}$ denotes the assignment of $x_{i}$ to a given mixture, which allows the coefficients of the linear combination to be written as $P(c_{i} = k)$ (note that these coefficients now sum to one naturally due to the nature of probability distributions).
+where $c_{i} \in \{ 1, \ldots, K \}$ denotes the assignment of $x_{i}$ to a given mixture, which allows the coefficients of the linear combination to be written as $P(c_{i} = k)$.
 Additionally, $\theta_{k}$ denotes the parametric form of the $k^{\text{th}}$ mixture component.
@@
 
-The finite mixture model of the previous section can be extended to a Bayesian perspective by viewing the distribution parameters that were previously point values, $\theta_{c_{i}}$ (the mixture component parameters) and $p_{1}, \ldots, p_{K}$ (the mixture component assignment weights), as random variables and providing each with a prior distribution.
+The finite mixture model of the previous section can be extended to a Bayesian mixture model by viewing the distribution parameters that were previously point values, $\theta_{c_{i}}$ (the mixture component parameters) and $p_{1}, \ldots, p_{K}$ (the mixture component assignment weights), as random variables and providing each with a prior distribution.
 In this case, the prior distribution $\mathbb{G}_{0}$ is placed on the mixture component parameters, and the prior distribution $\text{Dir}(\alpha/K, \ldots, \alpha/K)$ is placed on the K mixture component assignment weights.
@@
 where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $c_{i}$ are the mixture component assignments associated with each observation, the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e.
-the distribution to be mixed, $F(\theta_{c_{i}})$), the $\theta_{c_{i}}$ are drawn from their prior distribution $\mathbb{G}_{0}$, and $p_{1}, \ldots, p_{K}$ are drawn from their prior distribution, where `Dir' refers to a standard Dirichlet distribution.
+the distribution to be mixed, $F(\theta_{c_{i}})$), the $\theta_{c_{i}}$ are drawn from the prior distribution $\mathbb{G}_{0}$, and $p_{1}, \ldots, p_{K}$ are drawn from a Dirichlet prior parameterized by $\alpha/K, \ldots, \alpha/K$.
 
@@
 \end{equation}
-Where `Dir' denotes a standard Dirichlet distribution.
+Where `Dir' denotes a Dirichlet distribution.
 The parameters $\mathbb{G}_{0}$ and $\alpha$ may be intuitively viewed as the mean and precision of the DP.
@@
 %
-A DPM is formulated by allowing the prior mixing distribution (distribution over component weights?) in a standard mixture model to be distributed according to a DP; this allows for modeling data where the true number of latent mixture components is unknown and arbitrarily large by assuming an infinite number of components, of which only a finite amount are expressed by the data.
+A DPM is formulated by allowing the prior mixing distribution \willie{(the distribution over component weights?)} in a standard mixture model to be distributed according to a DP; this allows for modeling data where the true number of latent mixture components is unknown and arbitrarily large by assuming an infinite number of components, of which only a finite amount are expressed by the data.
 In particular, the DPM can be defined hierarchically as
@@
 See \cite{gasthaus_2008} and \cite{gasthaus_thesis} for more details on this formulation.
-Note that difference between the formulation and indexing of the clusters in this model compared to the previous two models.
-This definition can be shown to be equivalent to the Bayesian mixture model defined in \eqref{bayesian_mixture_model}, when K is taken to be unbounded.
+Note the difference between the indexing of the clusters in this model and the indexing in the previous two models.
+This formulation can be shown to be equivalent to the Bayesian mixture model defined in \eqref{bayesian_mixture_model}, when K is taken to be unbounded.
 This is the origin for references to this model as the infinite mixture model.
@@
 
-To account for the dependent behavior of data, research has been conducted on developing models involving a sequence of DPMs, where components of the mixtures are dependent upon (or may be considered `tied to') corresponding components at adjacent positions in the sequence.
-For example, if the data shows temporal dependence, the goal might be to create a sequence of DPMs, one for each time-step, where the components of the mixture at each step are dependent upon corresponding components in the both the following and previous time steps.
-Research towards models which accomplish such goals are described in (list citations).
+To account for the dependent behavior of data, research has been conducted on developing models involving a sequence of DPMs, where components of the mixtures are dependent upon (or may be considered `tied to') corresponding components at neighboring positions in the sequence.
+For example, if the data shows temporal dependence, the goal might be to create a sequence of DPMs, one for each time-step, where the components of the mixture at each step are dependent upon corresponding components in the both the following and previous time steps \willie{cite some of these}.
 
-More rigorously, we take the definition of a DDPM to be a stochastic process defined on the space of probability distributions over a domain, which are indexed by time, space, or a selection of other covariates in such a way that the marginal distribution at any point in the domain follows a Dirichlet process (adapted from definitions found in [cite griffin and steel and gasthaus-thesis]).
+More rigorously, we take the definition of a DDPM to be a stochastic process defined on the space of probability distributions over a domain, which are indexed by time, space, or a selection of other covariates in such a way that the marginal distribution at any point in the domain follows a Dirichlet process (adapted from definitions found in \willie{cite griffin and steel and gasthaus-thesis}).
+Hence, a time-dependent DDPM is a model which remains a Dirichlet process, marginally, at each time step, yet allows cluster parameters at a given time step to vary from (and remain dependent upon) the parameters in neighboring time steps.
 
@@
 \subsection{Inference Implementation in Experiments}
+\label{sec:inference_spec}
 
@@
 \end{equation}
-where $N_{\text{mapped}}$ refers to an optimal mapping between ground-truth and detected objects as specified in section \ref{sec:mappingtoground} using the STDA as the relevant metric, and $N_{(G_{i} \cup D_{i} \neq \emptyset)}$ denotes the number of unique objects (including false positive objects, true negative objects, and correctly mapped objects) in a video sequence.
+where $N_{\text{mapped}}$ refers to an optimal mapping between ground-truth and detected objects as specified in section \ref{sec:mappingtoground} using the STDA as the relevant metric, and $N_{(G_{i} \cup D_{i} \neq \emptyset)}$ denotes the number of frames in which a given tracked object, the ground truth object to which it is mapped, or both, are present.
 

### 2011-12-05 ltx/paper.tex
more changes, working on inference section currentely

@@
 
-A finite mixture model can be thought of as a probability distribution for an observation $x_{i}$ formulated as a linear combination of K mixture components (where the coefficients of the linear combination sum to one); each mixture component, in turn, is a probability distribution for $x_{i}$, given some parametric form.
+A finite mixture model can be thought of as a probability distribution for an observation $x_{i}$ formulated as a linear combination of K mixture components (also called `clusters'), where the coefficients of the linear combination sum to one; each mixture component, in turn, is a probability distribution for $x_{i}$, given some parametric form.
 The finite mixture model can be written as
@@
 We also define $p_{k} = P(c_{i} = k)$ for all $k \in \{ 1, \ldots, K \} $.
-This model can also be written hierarchically as
+This model can also be written generatively as
 \begin{align}
@@
 When using a DP as a prior in a Bayesian mixture model, $\Theta$ can be viewed as the set of parameters of the component mixture distributions.
-%
-A DPM is formulated by allowing the prior mixing distribution \willie{(the distribution over component weights?)} in a standard mixture model to be distributed according to a DP; this allows for modeling data where the true number of latent mixture components is unknown and arbitrarily large by assuming an infinite number of components, of which only a finite amount are expressed by the data.
-In particular, the DPM can be defined hierarchically as
+	A DPM is formulated by allowing the prior distribution over component weights in a standard mixture model to be distributed according to a DP \willie{need to make sure am saying this right}; this allows for modeling data where the true number of latent mixture components is unknown and arbitrarily large by assuming an infinite number of components, of which only a finite amount are expressed by the data.
+In particular, the DPM can be defined generatively as
 \begin{align}
@@
 \begin{split}
+\label{crp_rep}
 	P(c_{i} = c_{j} \text{  for some  } j<i) &= \frac{m_{k}}{i-1+\alpha}\\
@@
 Data gained by performing extraction on videos containing objects is significantly time-dependent.
-Futhermore, the clusters of data, each of which represents an object in the video, might change in number as time progresses, since objects can enter and exit a scene.
-The specific DDPM used to model extraction data in this work while handling the above challenges is known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM), and was introduced by \cite{caron_2007} in 2006.
+Futhermore, the clusters of data, each representing a video object, might change in number as time progresses (which is to say, clusters may get `born' and may `die' at some intermediate time step), since objects can enter and exit a scene.
+To handle these challenges, the specific DDPM chosen to model extraction data in this work is known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM), introduced by \cite{caron_2007} in 2006.
 
-The GPUDDPM, when applied to data over T discrete time steps, can be viewed as a collection of DPMMs (one for each $t \in \{1, \ldots, T \}$), which are linked together by dependencies among their respective parameters.
-For example, the mixture component parameters at a given time step are dependent upon the mixture component parameters in adjacent time steps and the distribution and number of distinct cluster assignments at a given time step are dependent upon the same over a wider range of time steps, upon the distribution of observations' cluster assignments over this range, and upon a deletion procedure, described below.
+The GPUDDPM, when applied to data over T discrete time steps, can be viewed as a sequence of DPMs (one for each $t \in \{1, \ldots, T \}$), which are linked together by dependencies between cluster parameters in neighboring time steps.
+More specifically, the parameters at a given time step are distributed as a function of the parameters in the previous time step, and the distribution and number of distinct cluster assignments at a given time step are distributed according to all previous cluster assignments and a deletion procedure, described below.
 
-To link together the parameters of mixture components in adjacent time-steps, a transition kernel $P(\phi_{k}^{t} | \phi_{k}^{t-1})$ is implemented, which provides a distribution over parameters of a mixture component in a given time-step (as a function of equivalent parameters in adjacent time-steps).
+A transition kernel $P(\phi_{k}^{t} | \phi_{k}^{t-1})$ specifies how mixture component parameters in time step $t$ are dependent upon associated mixture component parameters in time step $t-1$.
 One caveat is that each mixture component must be drawn independently from $\mathbb{G}_{0}$ (the base distribution of the DP, which acts as a prior distribution for the cluster parameters) which we achieve by making $\mathbb{G}_{0}$ the invariant distribution of $P(\phi_{k}^{t} | \phi_{k}^{t-1})$ (note that the transition kernel is a markov chain).
-Additionally, to account for varying numbers of clusters, there is a deletion procedure (described in \cite{caron_2007}) by which observations are considered "removed" from their assigned clusters at a given time-step (which modifies the value of $m_{k}$ defined above in the CRP representation).
-We introduce variables $d_{1}, \ldots, d_{N}$ which denote the times at which the assignments of given observations are considered to be removed.
+
+\willie{Do i really want to describe alternate deletion model, as below??
+shouldn't i start with usual deletion first..and introduce alternate deletion in implementation section when i do MCMC...A: no, i do not.}
+
+Recall that the distribution over the cluster assignments $c_{i}$ (one for each observation $i$) is a function of the cluster sizes (see \eqref{crp_rep}).
+Hence, to account for varying numbers of clusters--and in particular, to allow clusters to diminish the number of assigned observations and even die--there is a deletion procedure (described in \cite{caron_2007}) by which observations are considered `removed' from their assigned clusters at a given time step and value of $m_{k}$ (defined in \eqref{crp_rep} to be the size of cluster $k$) is reduced.
+
+The deletion procedure operates in the following way: at each time step, all previous assignments (that have not yet been deleted) are independently considered for deletion.
+Specifically, each (remaining) assignment is `removed' from its cluster with probability $\rho$.
+We can consider the size $m_{k,t}$ of cluster $k$ at time $t$ to be dependent upon the cluster size at time $t-1$ via a deletion step.
+In fact, it can be shown that performing deletion in this way is equivalent to, for each cluster $k$, drawing $r \sim \text{Binomial}(m_{k,t-1}, \rho)$, and reducing the cluster size $m_{k,t}$ by $r$.
+This can be written as
+\begin{equation}
+\label{del_step}
+m_{k,t} | m_{k,t-1}   \sim   m_{k,t-1}-\text{Binomial}(m_{k,t-1}, \rho)
+\end{equation}
+% We introduce variables $d_{1}, \ldots, d_{N}$ which denote the times at which the assignments of given observations are considered to be removed.
 With these variables, we can now express the $m_{k}$ value at time $t$ as
-\begin{equation}
-m_{k,t} = \sum_{t' = 1}^{t} \mathbb{I}[(c_{t'}=k) \wedge (t < d_{t'})]
-\end{equation}
-where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise.
+% \begin{equation}
+% \label{compute_clust_size}
+% m_{k,t} = \sum_{t' = 1}^{t} \mathbb{I}[(c_{t'}=k) \wedge (t < d_{t'})]
+% \end{equation}
+% where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise.
 Additionally, for an observation $x_{i}$ at a given time-step $t$, the deletion time $d_{i}$ can be defined to be $d_{i} = t + l_{i}$, where $l_{i}$ is considered the lifespan of an assignment, is distributed geometrically, and can be expressed as
-\begin{equation}
-l_{i} | \rho  \sim  \rho(1 - \rho)^{l_{i}}
-\end{equation} \\
-with parameter $\rho$.
-This process adheres to what \cite{caron_2007} calls a `uniform deletion strategy' over all the observations' assignments (since assignments to each cluster have equal probability of being deleted), though a more complex deletion strategy, dependent upon cluster size, can also be implemented.
-
-Using the cluster size and observation deletion terms introduced above, we can define the GPUDDPM generatively as, for each time step $t = \{1, \ldots, T\}$
+% \begin{equation}
+% \label{del_rho_form}
+% l_{i} | \rho  \sim  \rho(1 - \rho)^{l_{i}}
+% \end{equation} \\
+with deletion parameter $\rho$.
+This process adheres to what \cite{caron_2007} calls a `uniform deletion strategy' over all the observations' assignments (since assignments to each cluster have equal probability of being deleted), though a more complex deletion strategy, dependent upon cluster size, can also be implemented \cite{caron_2007}.
+For a given time $t$ and cluster $k$, the size $m_{k,t}$ is dependent upon the cluster size at the previous time step (via \eqref{del_step}) and on the assignments for observations at time $t$ ($c_{i}$ for $i \in \{ 1, \ldots, N_{t} \}$, where $N_{t}$ is the number of observations at time step $t$).
+We refer to this distribution over $m_{k,t} | m_{k,t-1}, c_{i}, \rho$ as DEL.
+Using the cluster size and observation deletion terms introduced above, we can define the GPUDDPM generatively as, for each time step $t \in \{1, \ldots, T\}$, and each observation $i \in \{ 1, \ldots, N_{t} \}$ at $t$
 			% \begin{align}
@@
 	c_{t}^{i} | m_{c_{t}^{i}, t-1}, \alpha  &\sim  \text{CRP}(m_{c_{t}^{i}, t-1}, \alpha) \\
-	m_{c_{t-1}^{i}, t-1} | m_{c_{t-1}^{i}, t-2}, c_{t-1}^{i}  &\sim \text{DEL}(m_{c_{t-1}^{i}, t-2}, c_{t-1}^{i}, \rho, \gamma) \\
+	m_{c_{t-1}^{i}, t-1} | m_{c_{t-1}^{i}, t-2}, c_{t-1}^{i}, \rho  &\sim \text{DEL}(m_{c_{t-1}^{i}, t-2}, c_{t-1}^{i}, \rho) \\
 	\theta_{k, t} | \theta_{k, t-1}   &\sim 
@@
 \end{align}
-
-The graphical model corresponding with this formulation is shown in Figure \ref{fig:gpuddpm_gm_1}.
+where $\text{CRP}(\cdot, \cdot)$ is given in \eqref{crp_rep}. The graphical model corresponding with this formulation is shown in Figure \ref{fig:gpuddpm_gm_1}.
 
 \begin{figure}[h]
-        \center{\includegraphics[width=100mm]{../img/gpuddp_gm_1.pdf}}
+        \center{\includegraphics[width=90mm]{../img/gpuddp_gm_1.pdf}}
         \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Poly Urn Dependent Dirichlet Process Mixture}
@@
 The likelihood and object appearance model is therefore
-
 \begin{equation}
@@
 \end{equation}
-
 where $x_{i}$ is an observation, $c_{i}$ is its associated assignment, $\theta_{c_{i}}$ specifies the parameters of the parametric form of the $c_{i}^{th}$ cluster.
@@
 \subsubsection{Sampling Assignment Variables $c_{i}$}
-The assignment of each observation, $c_{i}$, given all other members of the state,
+
+A value proportional to the posterior distribution can be computed for each possile value that $c_{i}$ may take on, i.e.
+for the posterior probability that $c_{i} = k \in \{ 1 , \ldots ,  K_{t}^{\text{possible}}, k^{*}\}$, where $K_{t}^{\text{possible}}$ is the number clusters with a non-zero size at any time between $t$ and $d_{t}^{i}$, and $k^{*}$ is a potential new cluster.
+The probably that $c_{i}$ is assigned to a cluster $k$ can be computed for each $k \in \{ 1 , \ldots ,  K_{t}^{\text{possible}}, k^{*} \}$ given all other members of the state, with
+\begin{equation}
+p(c_{t}^{i} = k | rest) \propto \prod_{t' = t}^{d_{t}^{i}} P(c_{t'}^{i} | m_{t'}^{i-1}, c_{1:t'-1}) \newline \times
+\begin{cases}
+	P(\bold{x}_{t}^{i} | \theta_{k,t})                                    \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{\text{possible}} \} \\
+	\int P(\bold{x}_{t}^{i} | \theta) * \mathbb{G}_{0}(\theta) d\theta    \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
+\end{cases}
+\end{equation}
+where the distribution under the product $P(c_{t'}^{i} | m_{t'}^{i-1}, c_{1:t'-1})$ is given in, and is taken under the situation where $c = k$, ie etc
 
@@
 \subsubsection{Sampling Cluster Parameters $\theta_{k, t}$}
+\begin{equation}
+P(x_{t}^{i} | \theta_{k,t})  P(z_{k,t+1,1:M} | \theta_{k,t})  P(\theta_{k,t} | z_{k,t,1:M}) = \newline
+\text{normal inverse wishart dirichlet with updated params...put here in addition to previous, or all only at previous and cite that?}
+\end{equation}
 
@@
 \begin{itemize}
+\renewcommand{\labelitemi}{$\bullet$}
 \item $G_{i}$ denotes the spatiotemporal region occupied by the $i$th ground-truth object in a video, and $G_{i}^{(t)}$ denotes the region occupied by the $i$th ground-truth object in frame $t$.
@@
 
-\subsubsection{Ants Videos}
+\subsubsection{Ants Video?}
 
-\subsubsection{}
+\subsubsection{Cells Video?}
 
 \section*{Results and Discussion}
+
+Discuss parameters somewhere in here.
+to add in discussion: thoughts about: deletion param vs.
+dataset size (and tracking through occlusion vs clusters taking over), mvn cluster size vs.
+tracking in occlusion vs.
+sampling correct number of objects.
+we detect movement, which isn’t exactly same as perf.
+metrics (ie our tracking stops if people stop moving).
+objects hard to track through occlusion becuase they are small, similar appearance, and our algorithm doesn’t assume dynamics to keep robust (we could, probably for better accuracy but less generality).
+background (appearance) is like a natural prior for remaining stationary
 

### 2011-12-05 ltx/paper.tex
more edits

@@
 
-\willie{Do i really want to describe alternate deletion model, as below??
-shouldn't i start with usual deletion first..and introduce alternate deletion in implementation section when i do MCMC...A: no, i do not.}
-
 Recall that the distribution over the cluster assignments $c_{i}$ (one for each observation $i$) is a function of the cluster sizes (see \eqref{crp_rep}).
-Hence, to account for varying numbers of clusters--and in particular, to allow clusters to diminish the number of assigned observations and even die--there is a deletion procedure (described in \cite{caron_2007}) by which observations are considered `removed' from their assigned clusters at a given time step and value of $m_{k}$ (defined in \eqref{crp_rep} to be the size of cluster $k$) is reduced.
+Hence, to account for varying numbers of clusters--and in particular, to allow clusters to diminish in size (i.e.
+reduce the number of assigned observations) and even die--there is a deletion procedure (described in \cite{caron_2007}) by which observations are considered `removed' from their assigned clusters at a given time step and value of $m_{k}$ (defined in \eqref{crp_rep} to be the size of cluster $k$) is reduced.
 
@@
 \end{equation}
-% We introduce variables $d_{1}, \ldots, d_{N}$ which denote the times at which the assignments of given observations are considered to be removed.
-With these variables, we can now express the $m_{k}$ value at time $t$ as
-% \begin{equation}
-% \label{compute_clust_size}
-% m_{k,t} = \sum_{t' = 1}^{t} \mathbb{I}[(c_{t'}=k) \wedge (t < d_{t'})]
-% \end{equation}
-% where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise.
-Additionally, for an observation $x_{i}$ at a given time-step $t$, the deletion time $d_{i}$ can be defined to be $d_{i} = t + l_{i}$, where $l_{i}$ is considered the lifespan of an assignment, is distributed geometrically, and can be expressed as
-% \begin{equation}
-% \label{del_rho_form}
-% l_{i} | \rho  \sim  \rho(1 - \rho)^{l_{i}}
-% \end{equation} \\
 with deletion parameter $\rho$.
@@
 We refer to this distribution over $m_{k,t} | m_{k,t-1}, c_{i}, \rho$ as DEL.
-Using the cluster size and observation deletion terms introduced above, we can define the GPUDDPM generatively as, for each time step $t \in \{1, \ldots, T\}$, and each observation $i \in \{ 1, \ldots, N_{t} \}$ at $t$
+Using the cluster size and observation deletion terms introduced above, we can define the GPUDDPM generatively as, for each time step $t \in \{1, \ldots, T\}$ and observation $i \in \{ 1, \ldots, N_{t} \}$ at $t$,
 			% \begin{align}
@@
 \begin{split}
-	\bold{x}_{t}^{i} | c_{t}^{i}, \theta_{c_{t}^{i}, t} &\sim F(\theta_{c_{t}^{i}, t}) \\
-	c_{t}^{i} | m_{c_{t}^{i}, t-1}, \alpha  &\sim  \text{CRP}(m_{c_{t}^{i}, t-1}, \alpha) \\
-	m_{c_{t-1}^{i}, t-1} | m_{c_{t-1}^{i}, t-2}, c_{t-1}^{i}, \rho  &\sim \text{DEL}(m_{c_{t-1}^{i}, t-2}, c_{t-1}^{i}, \rho) \\
-	\theta_{k, t} | \theta_{k, t-1}   &\sim 
+	\bold{x}_{i} | c_{i}, \theta_{c_{i}, t} &\sim F(\theta_{c_{i}, t}) \\
+	c_{i} | \bold{m}_{i-1}, \alpha  &\sim  \text{CRP}(\bold{m}_{i-1}, \alpha) \\
+	\bold{m}_{1} | \bold{m}_{N_{t-1}}, \rho  &\sim \text{DEL}(\bold{m}_{N_{t-1}}, \rho) \\
+	\theta_{t} | \theta_{t-1}   &\sim 
 \begin{cases}
-	P(\theta_{k, t} | \theta_{k, t-1}) \\
+	P(\theta_{t} | \theta_{t-1}) \\
 	\mathbb{G}_{0}
@@
 At each time step, cluster sizes can be recontructed from all previous assignment and deletion variables.
-The MCMC sampler described in this section operates in this alternate representation, which we can formulate as
+In particular, we introduce variables $d_{1}, \ldots, d_{N}$ which denote the times at which the assignments of given observations are considered to be removed.
+With these variables, we can now express the $m_{k}$ value at time $t$ as
+\begin{equation}
+\label{compute_clust_size}
+m_{k,t} = \sum_{t' = 1}^{t} \mathbb{I}[(c_{t'}=k) \wedge (t < d_{t'})]
+\end{equation}
+where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise.
+Additionally, for an observation $x_{i}$ at a given time-step $t$, the deletion time $d_{i}$ can be defined to be $d_{i} = t + l_{i}$, where $l_{i}$ is considered the lifespan of an assignment, is distributed geometrically, and can be expressed as
+\begin{equation}
+\label{del_rho_form}
+l_{i} | \rho  \sim  \rho(1 - \rho)^{l_{i}}
+\end{equation}
+The MCMC sampler described in this section operates in this alternate representation, which we can formulate as, for each time step $t \in \{1, \ldots, T\}$ and observation $i \in \{ 1, \ldots, N_{t} \}$ at $t$,
 \begin{align}
 \begin{split}
-	 \bold{x}_{i} | c_{i}, \theta_{c_{i}} &\sim \mathcal{N}(\boldsymbol{\mu}_{c_{i}}, \Sigma_{c_{i}})  \mathcal{M}n(\bold{p}_{c_{i}}) \\
+	 \bold{x}_{i} | c_{i}, \theta_{c_{i},t} &\sim \mathcal{N}(\boldsymbol{\mu}_{c_{i},t}, \Sigma_{c_{i},t})  \mathcal{M}n(\bold{p}_{c_{i},t}) \\
 	c_{i} | c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha  &\sim  \text{CRP}(c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha) \\
 	d_{i} | \rho  &\sim \text{Geo}(\rho) + i + 1\\
-	\theta_{k, t} | \theta_{k, t-1} \willie{other-prior-params}  &\sim 
+	\theta_{k, t} | \theta_{k, t-1}, \boldsymbol{\mu}_{0}, k_{0}, v_{0}, \Lambda_{0}  &\sim 
 \begin{cases}
 	\text{TK} \hspace{12pt} \text{if } k \in \{ 1, \ldots, K_{t-1}  \} \\
-	\mathbb{G}_{0}(\willie{prior params})  \hspace{15pt} \text{if } k = K_{t-1} + 1
+	\mathbb{G}_{0}(\boldsymbol{\mu}_{0}, k_{0}, v_{0}, \Lambda_{0})  \hspace{15pt} \text{if } k = K_{t-1} + 1
 \end{cases}
@@
 
-\subsubsection{Sampling Deletion Variables $d_{i}$}
-
 \subsubsection{Sampling Cluster Parameters $\theta_{k, t}$}
@@
 
+\subsubsection{Metropolis Hastings Algorithm}
+
 \subsubsection{Sampling Auxiliary Variables $z_{k,t,m}$}
+
+\subsubsection{Sampling Deletion Variables $d_{i}$}
 

### 2011-12-06 ltx/paper.tex
wrote more

@@
 
-Examples of additional features that could potentially be extracted from object-regions in video frames include color information, pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, texture representations, transform calculations (such as results of hough or fourier transforms), and quantiative region descriptions derived from image segmentation procedures; the main idea is to choose features which are able to be easily extracted or computed, capture variability in the appearance of objects, are applicable to a wide variety of object types, and can be represented by distributions that allow for tractable inference (discussed further in Section~\ref{sec:inference_spec}).
+Examples of additional features that could potentially be extracted from object-regions in video frames include color information, pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, texture representations, transform calculations (such as results of hough or fourier transforms), and quantiative region descriptions derived from image segmentation procedures; the main idea is to choose features which are able to be easily extracted or computed, capture variability in the appearance of objects, are applicable to a wide variety of object types, and can be represented by distributions that allow for tractable inference (discussed further in Section~\ref{sec:inference}).
 
@@
 \end{figure*}
-Inference performed on the time dependent Dirichlet process mixture model (described in Section~\ref{sec:inference_spec}) carries out clustering of the data, which isolates each worm-like data cloud and allows each object to be localized and tracked.
+Inference performed on the time dependent Dirichlet process mixture model (Section~\ref{sec:inference}) carries out clustering of the data, which isolates each worm-like data cloud and allows each object to be localized and tracked.
 
@@
 
-\section{Model Specification and Inference}
+\section{Model Specification}
 \label{sec:modelspec}
@@
 
-\subsection{Inference Implementation in Experiments}
-\label{sec:inference_spec}
+\section{Inference}
+\label{sec:inference}
+Algorithms that infer latent model parameters in a Bayesian manner given the data gained during extraction provide detection and tracking results.
+By formulating the detection and tracking task in terms of a generative model, a variety of methods previously developed for statistical inference can be readily applied; we separate presentations of model and inference so that multiple inference methods, each with its own benefits in terms of precision and speed, can be formulated and applied independently to gain detection and tracking results.
+This section provides background on the two inference procedures implemented in this study, used to perform Bayesian inference on the model described in Section~\ref{sec:modelspec}.
+The first is a type of Markov Chain Monte Carlo (MCMC) batch inference, which uses Gibbs sampling to generate samples from a posterior distribution over a model, and the second is a type of Sequential Monte Carlo (SMC) inference, also known as a particle filter, which also generates samples from a posterior information, but in this case in a sequential manner that is allowed due to the time-dependent nature of the model.
 
-This section provides background on the inference procedure applied to the model and extraction data, which allows for unsupervised detection and tracking of objects.
-A Markov Chain Monte Carlo (MCMC) sampler is used to perform inference, which makes use of the Metropolis-Hastings (M-H) algorithm to sample from a few distributions.
+\subsection{MCMC Batch Inference}
+
+This section details a Markov Chain Monte Carlo (MCMC) sampler used to perform inference, which provides another method to carry out unsupervised detection and tracking.
+It operates by iteratively sampling from conditional distributions of elements in the joint probability distribution--i.e.
+implementing Gibbs sampling--in order to approximate the posterior distribution of the cluster parameters given the data.
+used to perform inference, which makes use of the Metropolis-Hastings (M-H) algorithm to sample from a few distributions.
 
@@
 \end{equation}
-The MCMC sampler described in this section operates in this alternate representation, which we can formulate as, for each time step $t \in \{1, \ldots, T\}$ and observation $i \in \{ 1, \ldots, N_{t} \}$ at $t$,
+The MCMC sampler described in this section operates in this alternate representation, which we can formulate, for each time step $t \in \{1, \ldots, T\}$ and observation $i \in \{ 1, \ldots, N_{t} \}$ at time $t$, as
 \begin{align}
@@
 
-The sampler operates by iteratively sampling each member of the state given all other members.
+The sampler performs Gibbs sampling; it operates by iteratively sampling each member of the state given all other members.
 In particular, it moves sequentially through each of the $N$ observations, sampling the assignment $c_{i}$ and the deletion variable $d_{i}$ at each step.
@@
 
-A value proportional to the posterior distribution can be computed for each possile value that $c_{i}$ may take on, i.e.
-for the posterior probability that $c_{i} = k \in \{ 1 , \ldots ,  K_{t}^{\text{possible}}, k^{*}\}$, where $K_{t}^{\text{possible}}$ is the number clusters with a non-zero size at any time between $t$ and $d_{t}^{i}$, and $k^{*}$ is a potential new cluster.
-The probably that $c_{i}$ is assigned to a cluster $k$ can be computed for each $k \in \{ 1 , \ldots ,  K_{t}^{\text{possible}}, k^{*} \}$ given all other members of the state, with
+A value proportional to the posterior probability can be computed for each possible value that $c_{i}$ may take on; these values allow us to construct a discrete probability distribution that may be sampled from to get a posterior assignment sample.
+The possible values that the assignment may take on are $c_{i} = k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*}\}$, where $K_{t}^{'}$ is the number clusters with a non-zero size at any point between observation $i$ and its deletion point $d_{i}$, and $k^{*}$ is a potential new cluster.
+The probability that $c_{i}$ is assigned to a cluster $k$ can be computed for each $k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*} \}$, given all other members of the state, by
 \begin{equation}
-p(c_{t}^{i} = k | rest) \propto \prod_{t' = t}^{d_{t}^{i}} P(c_{t'}^{i} | m_{t'}^{i-1}, c_{1:t'-1}) \newline \times
+p(c_{i} = k | state) \propto \prod_{i' = i}^{d_{i}} P(c_{i'} | m_{i'-1}, c_{1:i'-1}) \newline \times
 \begin{cases}
-	P(\bold{x}_{t}^{i} | \theta_{k,t})                                    \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{\text{possible}} \} \\
-	\int P(\bold{x}_{t}^{i} | \theta) * \mathbb{G}_{0}(\theta) d\theta    \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
+	P(\bold{x}_{i} | \theta_{k,t})                                        \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
+	\int P(\bold{x}_{i} | \theta) * \mathbb{G}_{0}(\theta) d\theta    \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
 \end{cases}
 \end{equation}
-where the distribution under the product $P(c_{t'}^{i} | m_{t'}^{i-1}, c_{1:t'-1})$ is given in, and is taken under the situation where $c = k$, ie etc
+\willie{$d_{i}$ is actually a time; product needs to actually go up to last observation at time $d_{i}$} where the distribution under the product $P(c_{i'} | m_{i'-1}, c_{1:i'-1})$ is given in, and is taken under the situation where $c = k$, ie etc
 
 \subsubsection{Sampling Cluster Parameters $\theta_{k, t}$}
-\begin{equation}
-P(x_{t}^{i} | \theta_{k,t})  P(z_{k,t+1,1:M} | \theta_{k,t})  P(\theta_{k,t} | z_{k,t,1:M}) = \newline
-\text{normal inverse wishart dirichlet with updated params...put here in addition to previous, or all only at previous and cite that?}
-\end{equation}
+The conjugacy present in chosen distributions and auxiliary variables in the transition kernel allow for posterior samples of the cluster parameters to be easily gained, and gives us
+% \begin{eqnarray}
+% P(\theta_{k,t} | state) && = P(\bold{x}_{i} | \theta_{k,t}) P(z_{k,t+1,1:M} | \theta_{k,t})  P(\theta_{k,t} | z_{k,t,1:M}) \\
+%  && = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}, \Sigma | \boldsymbol{\mu}_{0}, k_{0}, v_{0}, \Lambda_{0})  \mathcal{D}ir( \bold{p} | \bold{q}_{0})
+% \end{eqnarray}
+\begin{align}
+\begin{split}
+P(\theta_{k,t} | state) & = P(\bold{x}_{i} | \theta_{k,t}) P(z_{k,t+1,1:M} | \theta_{k,t})  P(\theta_{k,t} | z_{k,t,1:M}) \\
+ & = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}_{k,t}, \Sigma_{k,t} | \boldsymbol{\mu}_{N}, k_{N}, v_{N}, \Lambda_{N})  \mathcal{D}ir( \bold{p}_{k,t} | \bold{q}_{N})
+\end{split}
+\end{align}
+Where the parameters in the above distribution are given when the data at time $t$, auxiliary variables $z_{k,t-1,1:M}$ at time $t-1$, and auxiliary variables $z_{k,t,1:M}$ at time $t$, are taken to be the observations in the following Bayesian updates
+\begin{eqnarray}
+Bayesian updates
+\end{eqnarray}
 
-\subsubsection{Metropolis Hastings Algorithm}
+% \subsubsection{Metropolis-Hastings Algorithm}
 
 \subsubsection{Sampling Auxiliary Variables $z_{k,t,m}$}
+The Metropolis-Hastings algorithm is used to gain samples from the posterior distribution for the auxiliary variables; it involves construction of a markov chain with the desired
 
 \subsubsection{Sampling Deletion Variables $d_{i}$}
+
+\subsection{SMC Inference}
+
+This section details a Sequential Monte Carlo (SMC) sampler--also known as a particle filter--used to perform inference, which provides another method to carry out unsupervised detection and tracking.
+\willie{general overview here: at each time step, perform sampling, and then show algorithm figure / thing}
+
+\subsubsection{Proposal Distribution for Assignments $c_{i}$}
+
+\subsubsection{Proposal Distribution $q_{1}$}
+
+\subsubsection{Proposal Distribution $q_{2}$}
 
@@
 
-The PETS2009/2010 dataset consists of a 900 frame video.
+The PETS2009/2010 dataset consists of a 900 frame video. \willie{describe which camera, sequence and view number, look at ferryman? intro paper and performance benchmark paper for good descriptions}
 
@@
 
-\subsubsection{PETS2004/Caviar Datasets}
+\subsubsection{Sensitivity Analysis}
 
-The Caviar datasets used in PETS2004 consist of multiple test-case-scenarios involving one or more humans.
+% \subsubsection{PETS2004/Caviar Datasets}
+% The Caviar datasets used in PETS2004 consist of multiple test-case-scenarios involving one or more humans.
+% \begin{figure}
+% 	\centering
+% 	\includegraphics[width=3in]{../img/pets2004_results_1}
+% 	\caption{Sample from posterior of the mean for frames $f = 1-20$ overlaid on frame $f=20$.}
+% 	\label{fig:pets2009_results}
+% \end{figure}
 
-\begin{figure}
-	\centering
-	\includegraphics[width=3in]{../img/pets2004_results_1}
-	\caption{Sample from posterior of the mean for frames $f = 1-20$ overlaid on frame $f=20$.}
-	\label{fig:pets2009_results}
-\end{figure}
+\subsection{New Video Datasets}
 
-\subsection{Nonstandard Video Datasets}
+The two video datasets in this section are video datasets first used in this paper. The datasets and groundtruth detailing the positions of each obect in each video will be published online.
 

### 2011-12-06 ltx/paper.tex
wrote more in MCMC batch inference section

@@
 where $\boldsymbol{\mu}_{M}, k_{M}, v_{M}, \Lambda_{M},$ and $\bold{q}_{M}$ are parameters given by:
-\willie{ran out of time. need to type up updates here.}
+\willie{these are given in next section, need to consider if I actually want to repeat them here}
 
@@
 
-This section details a Markov Chain Monte Carlo (MCMC) sampler used to perform inference, which provides another method to carry out unsupervised detection and tracking.
+This section details a Markov Chain Monte Carlo (MCMC) sampler used to perform inference, which provides a method to carry out unsupervised detection and tracking.
 It operates by iteratively sampling from conditional distributions of elements in the joint probability distribution--i.e.
-implementing Gibbs sampling--in order to approximate the posterior distribution of the cluster parameters given the data.
-used to perform inference, which makes use of the Metropolis-Hastings (M-H) algorithm to sample from a few distributions.
+it implements Gibbs sampling--in order to gain approximate samples from the posterior distribution of the cluster parameters given the data.
+Additionally, this sampler makes use of the Metropolis-Hastings (M-H) algorithm to sample from a few distributions.
 
@@
         \center{\includegraphics[width=100mm]{../img/gpuddp_gm_3.pdf}}
-        \caption{\label{fig:gpuddpm_gm_2} Graphical Model of the Generalized Poly Urn Dependent Dirichlet Process Mixture}
+        \caption{\label{fig:gpuddpm_gm_2} Graphical model of the deletion-variable representation of the GPUDDPM}
 \end{figure}
@@
 \subsubsection{Sampling Assignment Variables $c_{i}$}
+\label{sec:sample_assignments}
 
@@
 \end{equation}
-\willie{$d_{i}$ is actually a time; product needs to actually go up to last observation at time $d_{i}$} where the distribution under the product $P(c_{i'} | m_{i'-1}, c_{1:i'-1})$ is given in, and is taken under the situation where $c = k$, ie etc
+\willie{$d_{i}$ is actually a time; product needs to actually go up to last observation at time $d_{i}$} where the distribution under the product $P(c_{i'} | m_{i'-1}, c_{1:i'-1})$ is given in, and is taken under the situation where $c = k$.
+If a new cluster is sampled as an assignment, the cluster parameters and auxiliary variables must be initialized for all time steps.
+\willie{finish writing the few tricky technical stuff here}
 
@@
 \begin{eqnarray}
-Bayesian updates
+k_{N} &=& k_{0} + N \\
+v_{N} &=& v_{0} + N \\
+\boldsymbol{\mu}_{N} &=& \frac{k_{0}}{k_{0}+N} \boldsymbol{\mu}_{0}  +  \frac{N}{k_{0}+N} \overline{\bold{x}}^{s}\\
+\Lambda_{N} &=& \Lambda_{0} + S_{\bold{x}^{s}}\\
+\bold{q}_{N} &=& \bold{q}_{0} + \sum_{i=1}^{N} \bold{x}^{c}
+\end{eqnarray}
+where $N$ is the number of observations, $\{ \boldsymbol{\mu}_{0}, k_{0}, v_{0}, \Lambda_{N} \}$ are the $\mathcal{N}i\mathcal{W}$ prior parameters, $\bold{q}_{0}$ is the $\mathcal{D}ir$ prior parameter, $\bold{x}^{s}$ and $\bold{x}^{c}$ respectively denote the spatial and color features of the observations, and $\overline{\bold{x}}$ and $S_{\bold{x}}$ respectively denote the sample mean and sample covariance for a set of observations $\bold{x}$, defined to be
+\begin{eqnarray}
+\overline{\bold{x}}  &=&  \left( \sum_{i=1}^{N} \bold{x}_{i} \right) / N\\
+S_{\bold{x}}  &=&  \sum_{i=1}^{N} (\bold{x}_{i} - \overline{\bold{x}}) (\bold{x}_{i} - \overline{\bold{x}})^{T}
 \end{eqnarray}
@@
 \subsubsection{Sampling Auxiliary Variables $z_{k,t,m}$}
-The Metropolis-Hastings algorithm is used to gain samples from the posterior distribution for the auxiliary variables; it involves construction of a markov chain with the desired
+The posterior distribution of each of the auxiliary variables $z_{k,t,m}$ can be written as
+\begin{equation}
+P(z_{k,t,m} | state) \propto  P(z_{k,t,m} | \theta_{k,t-1}) P(\theta_{k,t} | z_{k,t,1:M})
+\end{equation}
+This distribution does not allow for samples to be easily drawn and thus requires a more sophisticated MCMC technique.
+In particular, the Metropolis-Hastings (M-H) algorithm is used to gain samples from the posterior distribution for the auxiliary variables; it involves constructing a markov chain with the sought posterior as its stationary distribution, and sampling sequentially from this constructed chain to get an approximate posterior sample.
+
+The M-H algorithm is used to sample auxiliary variables for each alive cluster.
+In this algorithm, we sample a new value for all $M$ auxiliary variables, denoted $z_{k,t,m}^{*}$, according to the following proposal distribution
+\begin{equation}
+z_{k,t,m}^{*}  \sim  P(z_{k,t,m} | \theta_{k,t}) = \mathcal{N}(z_{k,t,m}^{s} | \boldsymbol{\mu}_{k,t}, \Sigma_{k,t}) \mathcal{M}n(z_{k,t,m}^{c} | \bold{p}_{k,t})
+\end{equation}
+compute the standard M-H acceptance ratio, which in this case (due to the proposal distribution choice) simplifies to 
+\begin{equation}
+r_{accept} = \frac{P(\theta_{k,t-1} | z_{k,t,m}^{*})}{P(\theta_{k,t-1} | z_{k,t,m})}
+\end{equation}
+and keep this sample with probability $min(1, r_{accept})$.
 
 \subsubsection{Sampling Deletion Variables $d_{i}$}
+Computing the posterior distribution (or a value proportional to this distribution) for each possible value that a given deletion variable may take on (in a manner similar to how samples from the posterior distribution of the assignment variables $c_{i}$ are gained in Section~\ref{sec:sample_assignments}) may be computationally expensive due to the large number of potential deletion times.
+To remedy this, the M-H algorithm may again be used to sample from the posterior, where the prior distribution for the deletion times can be used to propose new samples $d_{i}^{*}$.
+As the `lifetime' of each assignment can be shown to be geometrically distributed, we can propose a new deletion time sample with
+\begin{align}
+\begin{split}
+l_{i}  &\sim  \text{Geo}(\rho)  \\
+d_{i}^{*}  &= l_{i} + t + 1
+\end{split}
+\end{align}
+and accept or reject this sample using the standard M-H acceptance ratio \willie{write computation of ratio}.
+
+\subsubsection{Initializing the Sampler}
+Before the samples can be iteratively drawn from each element of the state given all other elements, the state must be initialized.
+While the state of the sampler can theoretically be initialized at any value, beginning at a value closer to the maximum joint probability (over the state) allows for quicker inference.
 

### 2011-12-08 ltx/paper.tex
added more figures, paper writing continues

@@
 The data consists of the three-dimensional baseline features and was gained by extraction peformed on a 50 frame sequence showing five ants moving across the video scene.
-\begin{figure*}[h]
-        \center{\includegraphics[width=85mm]{../img/usbh_data_2.png}}
-        \caption{\label{fig:baselinedata} Baseline three-dimensional extraction data gained from a video segment containing 5 moving ants.
+% \begin{figure*}[h]
+%         \center{\includegraphics[width=85mm]{../img/usbh_data_2.png}}
+%         \caption{\label{fig:baselinedata} Baseline three-dimensional extraction data gained from a video segment containing 5 moving ants.
 The vertical axis denotes the frame number and the horizontal axes denote $d_{1}$ and $d_{2}$, the spatial coordinates of pixels within a given frame.
-Markers are colored by position for ease of viewing.}
-\end{figure*}
+Markers are colored by position for ease of viewing.
+\willie{show a number of baseline data here.
+consider whether colors help view or just add confusion.}}
+% \end{figure*}
+
+\begin{figure}[h]
+  \centering
+  \subfloat[]{\label{fig:baselinedata1}\includegraphics[width=0.33\textwidth]{../img/pets2001_data_1.png}} 
+  \subfloat[]{\label{fig:baselinedata1}\includegraphics[width=0.33\textwidth]{../img/usbh_data_3.png}} 
+  \subfloat[]{\label{fig:baselinedata2}\includegraphics[width=0.33\textwidth]{../img/traffic2_data_1.png}}\\
+  \subfloat[]{\label{fig:baselinedata1}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_1.png}} 
+  \subfloat[]{\label{fig:baselinedata1}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_2.png}} 
+  \subfloat[]{\label{fig:baselinedata2}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_3.png}}
+  \caption{Baseline extracted features from the (a.) PETS2001 Dataset, (b.) Ants Dataset, (c.) Traffic Dataset, and PETS2009/2010 benchmark dataset, frames (a.) 1-200, (b.) 201-400, (c.) 401-600.
+Time is shown on the vertical axis.}
+  \label{fig:baselinedata_both}
+\end{figure}
 Inference performed on the time dependent Dirichlet process mixture model (Section~\ref{sec:inference}) carries out clustering of the data, which isolates each worm-like data cloud and allows each object to be localized and tracked.
@@
   \subfloat[]{\label{fig:ants_img}\includegraphics[width=0.335\textwidth]{../img/usbh_frame_723.pdf}} 
-  \subfloat[]{\label{fig:ants_img}\includegraphics[width=0.335\textwidth]{../img/usbh_frame_724.pdf}}
-  \subfloat[]{\label{fig:ants_img_framediff}\includegraphics[width=0.32\textwidth]{../img/usbh_framediff.png}}
+  \subfloat[]{\label{fig:ants_img2}\includegraphics[width=0.335\textwidth]{../img/usbh_frame_724.pdf}}
+  \subfloat[]{\label{fig:ants_img_framediff}\includegraphics[width=0.323\textwidth]{../img/usbh_framediff.png}}\\
+  \subfloat[]{\label{fig:traffic2_img}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame1.png}} 
+  \subfloat[]{\label{fig:traffic2_img2}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame2.png}}
+  \subfloat[]{\label{fig:traffic2_img_framediff}\includegraphics[width=0.3273\textwidth]{../img/traffic2_framediff.png}}\\
+    \subfloat[]{\label{fig:pets2009_img}\includegraphics[width=0.333\textwidth]{../img/pets2009_frame1.png}} 
+  \subfloat[]{\label{fig:pets2009_img2}\includegraphics[width=0.333\textwidth]{../img/pets2009_frame2.png}}
+  \subfloat[]{\label{fig:pets2009_img_framediff}\includegraphics[width=0.333\textwidth]{../img/pets2009_framediff.png}}
   \caption{(a., b.) Consecutive frames of a video containing four ants. (c.) Extraction results produced by taking the pixel-wise difference between these frames}
@@
 
-A value proportional to the posterior probability can be computed for each possible value that $c_{i}$ may take on; these values allow us to construct a discrete probability distribution that may be sampled from to get a posterior assignment sample.
+A value proportional to the posterior probability can be computed for each possible value that $c_{i}$ may take on; these values allow us to construct a discrete probability distribution that may be sampled from to get a sample from the posterior distribution over assignments.
 The possible values that the assignment may take on are $c_{i} = k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*}\}$, where $K_{t}^{'}$ is the number clusters with a non-zero size at any point between observation $i$ and its deletion point $d_{i}$, and $k^{*}$ is a potential new cluster.
@@
 	P(\bold{x}_{i} | \theta_{k,t})                                        \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
-	\int P(\bold{x}_{i} | \theta) * \mathbb{G}_{0}(\theta) d\theta    \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
+	\int P(\bold{x}_{i} | \theta) \mathbb{G}_{0}(\theta) d\theta    \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
 \end{cases}
 \end{equation}
-\willie{$d_{i}$ is actually a time; product needs to actually go up to last observation at time $d_{i}$} where the distribution under the product $P(c_{i'} | m_{i'-1}, c_{1:i'-1})$ is given in, and is taken under the situation where $c = k$.
+\willie{$d_{i}$ is actually a time; product needs to actually go up to last observation at time $d_{i}$} where $t$ is the time of observation $i$, and the distribution under the product $P(c_{i'} | m_{1:K_{t}^{'}, i'-1}, c_{1:i'-1})$ is given by
+\begin{equation}
+P(c_{i} = k | m_{1:K_{t}^{'}, i-1}, c_{1:i-1}, \alpha) \propto 
+\begin{cases}
+m_{k, i-1} \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
+\alpha \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
+\end{cases}
+\end{equation}
+which is taken under the situation where $c = k$.
 If a new cluster is sampled as an assignment, the cluster parameters and auxiliary variables must be initialized for all time steps.
@@
 In particular, the Metropolis-Hastings (M-H) algorithm is used to gain samples from the posterior distribution for the auxiliary variables; it involves constructing a markov chain with the sought posterior as its stationary distribution, and sampling sequentially from this constructed chain to get an approximate posterior sample.
+The M-H algorithm is outlined in Algorithm~\ref{alg:MH}.
+\willie{am i citing wrong?}
+
+\begin{algorithm}[h]
+\label{alg:MH}
+\caption{Metropolis Hastings Algorithm}
+\begin{algorithmic}[1]
+\STATE $f(x)$ is the desired stationary PDF
+\STATE $P(x_{i} | x_{i-1})$ is the proposal pdf
+\STATE Initialize $x_{0}$
+\FOR{$i = 1 : burn\hspace{1pt}in + N$}
+\STATE Draw $x^{*} \sim P(x_{i} | x_{i-1})$
+\STATE Draw $\alpha \sim U(0,1)$
+\STATE $r_{accept} \leftarrow \frac{f(x^{*})P(x_{i-1} | x^{*})} {f(x_{i-1})P(x^{*} | x_{i-1})} $
+\IF{$\alpha < min \{ 1, r_{accept} \} $}
+\STATE $x_{i} \leftarrow x^{*}$
+\ELSE
+\STATE $x_{i} \leftarrow x_{i-1}$
+\ENDIF
+\ENDFOR
+\STATE Discard samples $x_{1}, \ldots, x_{burn\hspace{1pt}in}$
+\STATE \textbf{Output}: Remaining samples $x_{1}, \ldots, x_{N}$
+\end{algorithmic}
+%\caption{Say thing at end}
+\end{algorithm}
 
 The M-H algorithm is used to sample auxiliary variables for each alive cluster.
-In this algorithm, we sample a new value for all $M$ auxiliary variables, denoted $z_{k,t,m}^{*}$, according to the following proposal distribution
+In short, we sample a new value for all $M$ auxiliary variables, denoted $z_{k,t,m}^{*}$, according to the following proposal distribution
 \begin{equation}
@@
 This section details a Sequential Monte Carlo (SMC) sampler--also known as a particle filter--used to perform inference, which provides another method to carry out unsupervised detection and tracking.
+An overview of SMC inference is given in Algorithm~\ref{alg:SMC}.
 \willie{general overview here: at each time step, perform sampling, and then show algorithm figure / thing}
 
+\begin{algorithm}
+\caption{Sequential Monte Carlo Inference for the GPUDDPM}
+\label{alg:SMC}
+\begin{algorithmic}[1]
+\FOR{$t = 1 : T$}
+\FOR{$l = 1 : L$}
+\STATE $K_{t}^{(l)}  \leftarrow K_{t-1}^{(l)}$
+\FOR{$i = 1 : N_{t}$}
+\STATE Sample assigments from previous time sizes and cluster parameters $\theta_{k,t-1}$
+\STATE Draw $c_{i} \sim P(c_{i} | m_{1:K_{t-1}, i-1}, \theta_{1:K_{t-1}, t-1} )$
+\IF{$c_{i} = K_{t} + 1$}
+\STATE Draw a new cluster $\theta_{K_{t}+1,t} \sim \mathbb{G}_{0}(\text{priors})$   \willie{distro after algo?}
+\STATE $K_{t} \leftarrow K_{t} + 1$
+\ENDIF
+\ENDFOR
+\FOR{$s = 1 : S$}
+\FOR{$k = 1 : K$}
+\STATE Sample new cluster params via aux (given curent assignments and prev clusters)
+\ENDFOR
+\FOR{$i = 1 : N_{t}$}
+\STATE Sample new assignments (given current cluster params)
+\STATE Draw $c_{i} \sim P(c_{i} | m_{1:K_{t}, i-1}, \theta_{1:K_{t}, t} )$
+\ENDFOR
+\STATE Sample new cluster sizes.
+\ENDFOR
+\ENDFOR
+\STATE Only keep some of the $L$ samples.
+\ENDFOR
+\end{algorithmic}
+\end{algorithm}
+
 \subsubsection{Proposal Distribution for Assignments $c_{i}$}
+The probability of assignments given current cluster sizes and cluster parameters can be written as
+\begin{equation}
+P(c_{i} = k) = P(c_{i} | m_{1:K_{t}, i-1}, \theta_{1:K_{t}, t} )
+\end{equation}
 
@@
 
+\begin{equation}
+P(\theta_{k,t}) = P(c_{i} | m_{1:K_{t}, i-1}, \theta_{1:K_{t}, t} )
+\end{equation}
+
 \subsubsection{Proposal Distribution $q_{2}$}
+
+\begin{equation}
+P(\theta_{k,t}) = 	P(c_{i} | m_{1:K_{t}, i-1}, \theta_{1:K_{t}, t} )
+\end{equation}
 
@@
 
+\subsection{From Inference to Tracking Results}
+The multivariate normal-multinomial model described in this section can be used to gain object detection and tracking results.
+The nature of object detection and tracking results may vary depending on the application or performance metrics used (see Section~\ref{sec:performance_metrics}).
+In this study, we desired the centroid position and approximate spatial region for each moving object, for each frame that this object can be seen within the video.
+We chose to make our results depend completely on the inferred parameters of the multivariate normal distribution of each cluster.
+Each inferred cluster was taken to be a distinct detected object, and the sequence of means and covariance matrices for a given cluster were used, respectively, to determine the position and spatial region of a given object over a sequence of time steps.
+In particular, the mean parameter was taken to be the centroid of an object, and a 2-dimensional oval denoting a boundary around the mean that contains a specified percentage of the data (where the specification for this oval is given in Section~\ref{sec:}) was taken to be the spatial region of the object.
+Furthermore, as both provided inference methods use sampling, and each sample represents a result, we must choose a way to narrow these down to a single result.
+We have found that calculating the joint log probability for each sample and choosing the sample that yields the highest value as a result works reasonably well.
+
 \section{Experiments}
@@
 
-The following sections provide details on performance evaluation metrics that have been developed for use in object detection and tracking studies (and that are used to quantify experimental results in this paper), synthetic video experiments that verify certain aspects of the technique presented in this paper, benchmark video experiments that demonstrate the performance of this technique in relation to other methods developed in recent years, and nonstandard video datasets to show the capability of this technique to perform unsupervised detection and tracking on a diverse range of object and video types.
+This section provides details on performance evaluation metrics that have been developed to quantify results in object detection and tracking studies (and adopted in this paper), synthetic video experiments that verify certain aspects of the developed technique, benchmark video experiments that demonstrate the performance of this technique in relation to other methods developed in recent years, and new video experiments to show the capability of this technique to perform unsupervised detection and tracking on a range of object and video types.
 
 \subsection{Performance Evaluation Metrics}
+\label{sec:performance_metrics}
 
@@
 
-The problem of finding a mapping between a video's ground-truth tracks and an algorithm's output tracks is nontrivial, though necesary to solve in order to compute the performance evaluation metrics used in this study.
+The problem of finding a mapping between a video's ground-truth tracks and an algorithm's output tracks is nontrivial, though necesary to solve, in order to compute the performance evaluation metrics used in this study.
 In short, the typical solution to this problem involves first specifying a performance metric and then choosing the mapping from ground-truth tracks to output tracks which yields the most favorable performance metric value.
@@
 \cite{kasturi_2008}; we follow the method outlined in this paper to find an optimal mapping.
-Similiar to descriptions in this paper, we implement the Hungarian algorithm \cite{munkres_1957} as a polynomial-time ($O(n^{3})$) solution to the problem of optimally mapping two sets of tracks once the similarity between any two tracks given some specified metric is established.
+Similiar to descriptions in \cite{kasturi_2008}, we implement the Hungarian algorithm \cite{munkres_1957} as a polynomial-time ($O(n^{3})$) solution to the problem of optimally mapping two sets of tracks once the similarity between any two tracks given some specified metric is established.
 Additionally, the method employed in \cite{kasturi_2008} allows erroneous and undetected tracks to be left unmapped, which is both desired and necessary in the case where there is a different number of ground-truth and output tracks.
-Note that once a mapping from a collection of ground-truth tracks to a collection of output tracks has been established, one can determine which output tracks are false positives (the output tracks to which no ground-truth track is assigned) and which input tracks are false negatives (the input tracks that are not assigned to an output track).
+Note that once a mapping from a collection of ground-truth tracks to a collection of result tracks has been established, one can determine which result tracks are false positives (the result tracks to which no ground-truth track is assigned) and which ground-truth tracks are true negatives (the ground-truth tracks that are not assigned to a result track).
 The numbers of tracks displaying both of these failures are factors in the performance metrics used in this study.
@@
 The synthetic videos contain instances of occlusion (where one or more objects are briefly hidden) and objects with time-varying appearances and behaviors, as these are examples of occurances that notoriously decrease the accuracy of detection and tracking.
-After each video was constructed, the extraction and inference procedures described in Section~\ref{sec:modelspec} were carried out to return a collection of centroid-position-sequences, each corresponding to the sequence of positions of a tracked object over each frame that it is present in the video.
+After each video was constructed, the extraction and inference procedures described in Section~\ref{sec:modelspec} were carried out to return a sequence of multivariate-normal-parameters (means and covariance matrices) from which a sequence of positions and ovals approximating the shape of a tracked object over each frame that it is present in the video can be determined (as outlined in Section~\ref{sec:paramstoresults}).
 
@@
         \center{\includegraphics[width=85mm]{../img/pets2001_alphaimg_hr.png}}
-        \caption{\label{fig:baselinedata} Extracted data and sample from posterior of the model overlaid on three, time-varying images from the pets2001 dataset.}
+        \caption{\label{fig:alpha_overlay_1} Extracted data and sample from posterior of the model overlaid on three, time-varying images from the pets2001 dataset.}
 \end{figure}

### 2011-12-08 ltx/paper.tex
more writing. latex is working on my computer. most of the time

@@
   \caption{Baseline extracted features from the (a.) PETS2001 Dataset, (b.) Ants Dataset, (c.) Traffic Dataset, and PETS2009/2010 benchmark dataset, frames (a.) 1-200, (b.) 201-400, (c.) 401-600.
-Time is shown on the vertical axis.}
+Time is shown on the vertical axis.
+Marker color corresponds with spatial coordinates and is shown for ease of viewing.}
   \label{fig:baselinedata_both}
@@
 \begin{equation} \label{basedistro}
-\mathbb{G}_{0}(\theta) = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}, \Sigma | \boldsymbol{\mu}_{0}, k_{0}, v_{0}, \Lambda_{0})  \mathcal{D}ir( \bold{p} | \bold{q}_{0})
-\end{equation}
-where $\mathcal{N}i\mathcal{W}$ denotes the normal-inverse-Wishart distribution, $\mathcal{D}ir$ denotes the Dirichlet distribution, and the prior has the hyperparameters $\boldsymbol{\mu}_{0}, k_{0}, v_{0}, \Lambda_{0}$ and $\bold{q}_{0}$.
+\mathbb{G}_{0}(\theta) = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}, \Sigma | \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0})  \mathcal{D}ir( \bold{p} | \bold{q}_{0})
+\end{equation}
+where $\mathcal{N}i\mathcal{W}$ denotes the normal-inverse-Wishart distribution, $\mathcal{D}ir$ denotes the Dirichlet distribution, and the prior has the hyperparameters $\boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0}$ and $\bold{q}_{0}$.
 \willie{reference text with good background on NiW and Dir distributions}
@@
 \begin{equation} \label{transker2}
-\boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t} | \bold{z}_t  \sim  \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}_{M}, k_{M}, v_{M}, \Lambda_{M})  \mathcal{D}ir(\bold{q}_{M})
-\end{equation}
-where $\boldsymbol{\mu}_{M}, k_{M}, v_{M}, \Lambda_{M},$ and $\bold{q}_{M}$ are parameters given by:
+\boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t} | \bold{z}_t  \sim  \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}_{M}, \kappa_{M}, \nu_{M}, \Lambda_{M})  \mathcal{D}ir(\bold{q}_{M})
+\end{equation}
+where $\boldsymbol{\mu}_{M}, \kappa_{M}, \nu_{M}, \Lambda_{M},$ and $\bold{q}_{M}$ are parameters given by
 \willie{these are given in next section, need to consider if I actually want to repeat them here}
+
+\subsection{Recap of Parameters}
+\label{sec:parameter_recap}
+
+The multivariate normal-multinomial GPUDDPM implemented in this study has a number of parameters, which can be tuned to increase the efficacy of object detection and tracking.
+Object apperance parameters include\\\\
+\begin{center}
+\begin{tabular}[c]{l p{10cm}}
+$\boldsymbol{\mu}_{0}$  &  The mean prior. In experiments performed in Section~\ref{sec:experiments} the data was recentered to the origin, and this parameter was set to $(0,0)$ \\
+$\kappa_{0}$  &  scale factor of the variance of the prior on the mean\\
+$\nu_{0}$  &  scale factor of the variance of the prior on the covariance\\
+$\Lambda_{0}$  &  shape factor of the prior on the covariance\\
+$q_{0}$  &  scale factor of the prior on the multinomial counts
+\end{tabular}
+\end{center} \vspace{5mm}
+The following parameter dictates characteristics involving the movement of objects
+\begin{center}
+\begin{tabular}[c]{l p{10cm}}
+$M$  &  Number of auxiliarty variables. A large number will produce lower variation in position (better for slower objects) and a low number will produce higher variation in position.
+\end{tabular}
+\end{center} \vspace{5mm}
+Additionally, one can tune the model's tendency to detect new objects and maintain the existence of these objects (or the rate at which objects become inactive or leave the video) with the following parameters
+\begin{center}
+\begin{tabular}[c]{l p{10cm}}
+$\alpha$  &  The granularity parameter for the Dirichlet Process. A higher value will increase the tendency for new objects to be detected.\\
+$\rho$  &  The deletion parameter. A higher value will give objects an increased tendency to die off.
+\end{tabular}
+\end{center}
+\willie{Discuss choosing parameters}
 
@@
 \subsection{MCMC Batch Inference}
+\label{sec:MCMC}
 
@@
 	d_{i} | \rho  &\sim \text{Geo}(\rho) + i + 1\\
-	\theta_{k, t} | \theta_{k, t-1}, \boldsymbol{\mu}_{0}, k_{0}, v_{0}, \Lambda_{0}  &\sim 
+	\theta_{k, t} | \theta_{k, t-1}, \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0}  &\sim 
 \begin{cases}
 	\text{TK} \hspace{12pt} \text{if } k \in \{ 1, \ldots, K_{t-1}  \} \\
-	\mathbb{G}_{0}(\boldsymbol{\mu}_{0}, k_{0}, v_{0}, \Lambda_{0})  \hspace{15pt} \text{if } k = K_{t-1} + 1
+	\mathbb{G}_{0}(\boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0})  \hspace{15pt} \text{if } k = K_{t-1} + 1
 \end{cases}
@@
 \subsection{SMC Inference}
+\label{sec:SMC}
 
@@
 \subsection{From Inference to Tracking Results}
-The multivariate normal-multinomial model described in this section can be used to gain object detection and tracking results.
+\label{sec:inference_to_results}
+The multivariate normal-multinomial model on which inference is described in this section can be used to gain object detection and tracking results.
 The nature of object detection and tracking results may vary depending on the application or performance metrics used (see Section~\ref{sec:performance_metrics}).
@@
 Each inferred cluster was taken to be a distinct detected object, and the sequence of means and covariance matrices for a given cluster were used, respectively, to determine the position and spatial region of a given object over a sequence of time steps.
-In particular, the mean parameter was taken to be the centroid of an object, and a 2-dimensional oval denoting a boundary around the mean that contains a specified percentage of the data (where the specification for this oval is given in Section~\ref{sec:}) was taken to be the spatial region of the object.
+In particular, the mean parameter was taken to be the centroid of an object, and a 2-dimensional oval denoting a boundary around the mean that contains a specified percentage of the data (where the specification for this oval is given in Section~\ref{sec:pets2000_2001}) was taken to be the spatial region of the object.
 Furthermore, as both provided inference methods use sampling, and each sample represents a result, we must choose a way to narrow these down to a single result.
@@
 
+The ground-truth authored by the ViPER tool took the form of bounding boxes denoting the spatial postion of each object at each time step.
+Consequentially, to find the spatial overlap between results and ground-truth, which is intrinsic to both metrics, a rectangular bounding box was needed per object per time step from the results of the algorithm.
+We took the maximal and minimal axially aligned values of the oval produced by our algorithm (as described in Section~\ref{sec:inference_to_results}) to be the sides of a representative bounding box for a given object at a given frame.
+
 \subsection{Synthetic Video Datasets}
@@
 The synthetic videos contain instances of occlusion (where one or more objects are briefly hidden) and objects with time-varying appearances and behaviors, as these are examples of occurances that notoriously decrease the accuracy of detection and tracking.
-After each video was constructed, the extraction and inference procedures described in Section~\ref{sec:modelspec} were carried out to return a sequence of multivariate-normal-parameters (means and covariance matrices) from which a sequence of positions and ovals approximating the shape of a tracked object over each frame that it is present in the video can be determined (as outlined in Section~\ref{sec:paramstoresults}).
+After each video was constructed, the extraction and inference procedures described in Section~\ref{sec:modelspec} and \ref{sec:inference} were carried out to return a sequence of multivariate-normal-parameters (means and covariance matrices) from which a sequence of positions and ovals approximating the shape of a tracked object over each frame that it is present in the video can be determined (as outlined in Section~\ref{sec:inference_to_results}).
 
@@
 Frame difference extraction on each video yields a dataset with identical spatial features; hence, successful tracking depends fully on the reasonable incorporation of color information into the model.
+
+Parameters were set to the same values for inference on both videos, using the method outlined in Section~\ref{sec:parameter_recap} to choose parameters.
+The chosen values were $\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (5, \ldots, 5)$.
+Inference was carried out using the MCMC algorithm (Section~\ref{sec:MCMC}); the posterior sample taken as the result correctly tracked both colored squares through occlusion in both videos, and is shown in Figure~\ref{fig:synth_one_plot}.
+\willie{still need to include performance metric results for these synthetic datasets.}.
 
@@
 
+Parameter values were chosen in the same way, and set to the same values, as the first synthetic experiment.
+The MCMC inference algorithm correctly tracked all three objects through occlusion and inferred the appearance and size shifts.
+Figure~\ref{fig:synth2_alphaimg} shows a sample from the posterior distribution of the cluster parameters, where the mean and oval representation of the covariance matrix (with $0.5$ confidence value) are overlayed on the data.
+The data is plotted in three dimensional space, where time is represented on the vertical axis, and each data point is colored by its assignment to one of the three inferred clusters.
+Both the cluster parameter sample and data are overlaid on three semi-transparent planes showing the objecs in the video in three frames.
+\willie{still need to include performance metric results for this synthetic dataset}.
+
+% \begin{figure}
+%   \centering               
+%   % \subfloat[]{\includegraphics[width=0.35\textwidth, angle=30]{../img/synth2_img}}
+%   \subfloat[]{\includegraphics[width=0.55\textwidth]{../img/synth2_img_b}}
+%   \subfloat[]{\includegraphics[height=0.3\textheight]{../img/synth2_alphaimg_hd.png}}
+%   \caption{(a.) Video frame $f=10$ from synthetic experiment two.
+(b.) Sample from posterior of the state for synthetic experiment two, where color represents assignment and the mean and standard deviation are shown.
+All are overlaid on three, time-varying images from synthetic experiment.
+The three objects of time varying size, speed, and position are tracked through simultaneous occlusion.
+\willie{will show assignments by marker type in addition to color}}
+%   \label{fig:synth_two_plot}
+% \end{figure}
+
+\begin{figure}[h]
+        \center{\includegraphics[width=0.5\textwidth]{../img/synth2_alphaimg_hd.png}}
+        \caption{\label{fig:synth2_alphaimg} Sample from posterior of the state for synthetic experiment two, where color represents assignment and the mean and standard deviation are shown.
+Both the sample and data are overlaid on three images from synthetic experiment.
+The three objects of time varying size, speed, and position are tracked through simultaneous occlusion.
+\willie{will show assignments by marker type in addition to color}}
+\end{figure}
+
+% \begin{figure}[h]
+%         \center{\includegraphics[width=0.99\textwidth]{../img/synth2_alphaimg_hd.png}}
+%         \caption{\label{fig:synth2_alphaimg} Extracted data and sample from the posterior of the model overlaid on three, time-varying images from synthetic experiment 2.}
+% \end{figure}
+
+\subsection{Benchmark Video Datasets}
+
+Benchmark video datasets for object tracking and detection have been produced to provide standard scenes on which researchers can compare detection and tracking results.
+These videos have been primarily produced for surveillance-related workshops--notably, for the International Workshop on Performance Evaluation of Tracking and Surveillance (PETS)--which provide researchers with video datasets and algorithmic goals on which to focus.
+Three commonly used benchmark videos from PETS workshops--one used in PETS2000, one in PETS2001, and one used both in PETS2009 and PETS2010--were chosen to demonstrate the efficacy of the method presented in this study.
+The performance metrics and benchmark datasets allow the methods developed in this paper to be quantitatively compared against other detection and tracking algorithms.
+
+%\subsubsection{PETS2000}
+
+\subsubsection{PETS2000 and PETS2001}
+\label{sec:pets2000_2001}
+
+The PETS2000 and PETS2001 video datasets both contain a small number of humans and vehicles traveling across a parking lot, with video taken from above, emulating what might be recorded by standard outdoor surveillance equipment.
+The `Test Sequence', a set of images from a monocular, stationary camera, was used from the PETS2000 workshop, and View Two of Dataset 1, also taken via a monocular, stationary camera, was used from the PETS2001 workshop.
+The MCMC algorithm (described in Section~\ref{sec:MCMC}) was used for inference in these experiments.
+
+Due to the computation required for the MCMC batch inference method (discussed further in Section~\ref{sec:discussion}), only the final 1000 frames of the video were used.
+Parameter values were chosen in the same way, and set to the same values, as in the the synthetic experiments ($\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (5, \ldots, 5)$).
+The MCMC sampler was successful for both benchmark videos; each object was detected, tracked, and its shape estimated in manner very consistent with the ground-truth.
+The results are displayed in Figure~\ref{fig:pets2000_results}; in this figure, a sample from the posterior of the cluster parameters is overlayed the data and video frames at three time points, where the data is colored by assignment.
+\willie{also show pets 2001 result and per-frame overlays for both}.
+
+To calculate performance metrics (both SFDA and ATA), one must specify a confidence value that allows the oval representing the region occupied by an object to be computed from the inferred covariance matrix of each cluster (as discussed in Section~\ref{inference_to_results}).
+The performance metrics were found for a range of confidence intervals, and these curves, for the PETS2000 and PETS2001 video are shown in Figure~\ref{fig:pm_conf}.
+
 \begin{figure}
   \centering             
-  % \subfloat[]{\includegraphics[width=0.35\textwidth, angle=30]{../img/synth2_img}}
-  \subfloat[]{\includegraphics[width=0.55\textwidth]{../img/synth2_img_b}}
-  \subfloat[]{\includegraphics[width=0.45\textwidth]{../img/synth2_final}}
-  \caption{(a.) Video frame $f=10$ from synthetic experiment two.
-(b.) Sample from posterior of the state for synthetic experiment two, where color represents assignment and the mean and standard deviation are shown.
-The three objects of time varying size, speed, and position are tracked through simultaneous occlusion.
-\willie{will show assignments by marker type in addition to color}}
-  \label{fig:synth_two_plot}
-\end{figure}
-
-\begin{figure}[h]
-        \center{\includegraphics[width=0.99\textwidth]{../img/synth2_alphaimg_hd.png}}
-        \caption{\label{fig:synth2_alphaimg} Extracted data and sample from the posterior of the model overlaid on three, time-varying images from synthetic experiment 2.}
-\end{figure}
-
-% \subsubsection{Similar Appearances}
-
-% \subsubsection{Different Numbers of Objects}
-
-% \subsubsection{Objects wih Diverse Characteristics}
-
-\subsection{Benchmark Video Datasets}
-
-Benchmark video datasets for object tracking and detection have been produced to provide standard scenes on which researchers can compare detection and tracking results.
-These videos have been primarily produced for surveillance-related workshops--notably, for the International Workshop on Performance Evaluation of Tracking and Surveillance (PETS)--which provide a setting where researchers can focus on a collection of videos and computer vision goals.
-Four commonly used benchmark videos from four PETS workshops are
-
-%\subsubsection{PETS2000}
-
-\subsubsection{PETS2001}
-
-The PETS2001 dataset consists of
+  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2000_sfda}}
+  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2000_ata}}
+  \caption{The (a.) SFDA and (b.) ATA are shown, vs. confidence values from which an object's oval region is computed}
+  \label{fig:pm_conf}
+\end{figure}
 
@@
 
-\subsubsection{PETS2000}
-
-The PETS2000 dataset consists of
-
 \begin{figure}
@@
 
-%\subsubsection{PETS2004/Caviar}
-
 \subsubsection{PETS2009/2010}
 
-The PETS2009/2010 dataset consists of a 900 frame video. \willie{describe which camera, sequence and view number, look at ferryman? intro paper and performance benchmark paper for good descriptions}
+A video dataset used in both the PETS2009 and PETS2010 conferences was chosen for experimentation due to its prominence in a number of studies as a well tested benchmark dataset \willie{ref papers using this dataset}.
+This video consists of a 794 frame video.
+\willie{describe which camera, sequence and view number, look at ferryman?
+intro paper and performance benchmark paper for good descriptions}.
+The entire video sequence (all frames) were used in this experiment.
+
+Due to the large number of frames and objects in this video, the SMC algorithm (described in Section~\ref{sec:SMC}) was used for inference.
 
@@
 \section*{Results and Discussion}
+\label{sec:discussion}
 

### 2011-12-08 ltx/paper.tex
changed conf v performance metric graphs
wrote more. updated paper_refs.bib a little. added more figures.

@@
 The synthetic videos contain instances of occlusion (where one or more objects are briefly hidden) and objects with time-varying appearances and behaviors, as these are examples of occurances that notoriously decrease the accuracy of detection and tracking.
-After each video was constructed, the extraction and inference procedures described in Section~\ref{sec:modelspec} and \ref{sec:inference} were carried out to return a sequence of multivariate-normal-parameters (means and covariance matrices) from which a sequence of positions and ovals approximating the shape of a tracked object over each frame that it is present in the video can be determined (as outlined in Section~\ref{sec:inference_to_results}).
+After each video was constructed, the extraction procedure described in Section~\ref{sec:modelspec_extraction} (using $L=3$) and inference procedures described in \ref{sec:inference} were carried out to return a sequence of multivariate-normal-parameters (means and covariance matrices) from which a sequence of positions and ovals approximating the shape of a tracked object over each frame that it is present in the video can be determined (as outlined in Section~\ref{sec:inference_to_results}).
 
@@
 
-The PETS2000 and PETS2001 video datasets both contain a small number of humans and vehicles traveling across a parking lot, with video taken from above, emulating what might be recorded by standard outdoor surveillance equipment.
+The PETS2000 and PETS2001 video datasets both consist of a small number of humans and vehicles traveling across a parking lot, with video taken from above, emulating what might be recorded by standard outdoor surveillance equipment.
 The `Test Sequence', a set of images from a monocular, stationary camera, was used from the PETS2000 workshop, and View Two of Dataset 1, also taken via a monocular, stationary camera, was used from the PETS2001 workshop.
@@
 Due to the computation required for the MCMC batch inference method (discussed further in Section~\ref{sec:discussion}), only the final 1000 frames of the video were used.
+Extraction was performed with frame differencing as described in Section~\ref{sec:modelspec_extraction}, using $L=3$.
 Parameter values were chosen in the same way, and set to the same values, as in the the synthetic experiments ($\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (5, \ldots, 5)$).
 The MCMC sampler was successful for both benchmark videos; each object was detected, tracked, and its shape estimated in manner very consistent with the ground-truth.
-The results are displayed in Figure~\ref{fig:pets2000_results}; in this figure, a sample from the posterior of the cluster parameters is overlayed the data and video frames at three time points, where the data is colored by assignment.
+The results for the PETS2000 dataset are displayed in Figure~\ref{fig:pets2000_results} and for the PETS2001 dataset in Figure~\ref{fig:alpha_overlay_1}; in these figure, a sample from the posterior of the cluster parameters is overlayed the data and video frames at three time points, where the data is colored by assignment.
 \willie{also show pets 2001 result and per-frame overlays for both}.
 
-To calculate performance metrics (both SFDA and ATA), one must specify a confidence value that allows the oval representing the region occupied by an object to be computed from the inferred covariance matrix of each cluster (as discussed in Section~\ref{inference_to_results}).
+To calculate performance metrics (both SFDA and ATA), one must specify a confidence value that allows the oval representing the region occupied by an object to be computed from the inferred covariance matrix of each cluster (as discussed in Section~\ref{sec:	inference_to_results}).
 The performance metrics were found for a range of confidence intervals, and these curves, for the PETS2000 and PETS2001 video are shown in Figure~\ref{fig:pm_conf}.
+
+% \begin{figure}
+% 	\centering
+% 	\subfloat[]{\includegraphics[width=0.52\textwidth]{../img/pets2001_results_1}}
+% 	\subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2001_results_2}}
+% 	\caption{Sample from posterior of the state for the PETS2001 dataset, where color represents assignment and the mean and standard deviation are shown.
+\willie{will show assignments by marker type in addition to color}}
+% 	\label{fig:pets2001_results}
+% \end{figure}
+
+\begin{figure}[h]
+        \center{\includegraphics[width=85mm]{../img/pets2001_alphaimg_hr.png}}
+        \caption{\label{fig:alpha_overlay_1} Extracted data and sample from posterior of the model overlaid on three frames from the pets2001 video.}
+\end{figure}
 
@@
 	\centering
-  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2000_sfda}}
-  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2000_ata}}
-  \caption{The (a.) SFDA and (b.) ATA are shown, vs. confidence values from which an object's oval region is computed}
-  \label{fig:pm_conf}
+	\subfloat[]{\includegraphics[width=0.6\textwidth]{../img/pets2000_results_1}}
+	\subfloat[]{\includegraphics[width=0.40\textwidth]{../img/pets2000_4objects.png}}
+	\caption{Extracted data and sample (a.) from posterior of the model overlaid (b.) on a frame from the pets2000 video}
+	\label{fig:pets2000_results}
 \end{figure}
@@
   \centering             
-	\subfloat[]{\includegraphics[width=0.52\textwidth]{../img/pets2001_results_1}}
-	\subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2001_results_2}}
-	\caption{Sample from posterior of the state for the PETS2001 dataset, where color represents assignment and the mean and standard deviation are shown.
-\willie{will show assignments by marker type in addition to color}}
-	\label{fig:pets2001_results}
+  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2000_pm}}
+  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2001_pm}}
+  \caption{The SFDA (blue solid line) and ATA (red dashed line are) vs  confidence values from which an object's oval region is computed for (a.) PETS2000 and (b.) PETS2001 video datasets.}
+  \label{fig:pm_conf}
 \end{figure}
 
-\begin{figure}[h]
-        \center{\includegraphics[width=85mm]{../img/pets2001_alphaimg_hr.png}}
-        \caption{\label{fig:alpha_overlay_1} Extracted data and sample from posterior of the model overlaid on three, time-varying images from the pets2001 dataset.}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\subfloat[]{\includegraphics[width=1\textwidth]{../img/pets2000_results_1}}
-	%\subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2000_results_2}}
-	\caption{pets2000}
-	\label{fig:pets2000_results}
-\end{figure}
-
 \subsubsection{PETS2009/2010}
 
-A video dataset used in both the PETS2009 and PETS2010 conferences was chosen for experimentation due to its prominence in a number of studies as a well tested benchmark dataset \willie{ref papers using this dataset}.
-This video consists of a 794 frame video.
-\willie{describe which camera, sequence and view number, look at ferryman?
-intro paper and performance benchmark paper for good descriptions}.
-The entire video sequence (all frames) were used in this experiment.
+A video dataset used in both the PETS2009 and PETS2010 conferences, called `S2.L1 at time sequence 12.34' was chosen for experimentation due to its prominence in a number of studies \willie{ref papers using this dataset}.
+This dataset consists of a monocular, stationary camera, 794 frame video sequence.
+The entire video sequence was used in this experiment.
 
 Due to the large number of frames and objects in this video, the SMC algorithm (described in Section~\ref{sec:SMC}) was used for inference.
+This method of sequential inference was observed, on this dataset, to converge to a better sample in a shorter period of time in comparison with the MCMC algorithm.
+For this reason, we found that it was better suited for this large dataset.
+
+Extraction was performed with frame differencing as described in Section~\ref{sec:modelspec_extraction}, using $L=3$ and parameters for the model were chosen, using the method outlined in Section~\ref{sec:parameter_recap}, to be $\alpha = 0.1, \rho = 0.8, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 6, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (3, \ldots, 3)$.
+Additionlly, as with the other video datasets, ground-truth was authored using the ViPER tool.
+
+Sequential inference produced reasonably accurate results to be achieved in a short timespan.
 
@@
 
-\section*{Results and Discussion}
+\section*{Discussion}
 \label{sec:discussion}

### 2011-12-08 ltx/paper.tex
more writing. new figures.

@@
 \begin{tabular}[c]{l p{10cm}}
-$M$  &  Number of auxiliarty variables. A large number will produce lower variation in position (better for slower objects) and a low number will produce higher variation in position.
+$M$  &  Number of auxiliary variables. A large number will produce lower variation in position (better for slower objects) and a low number will produce higher variation in position.
 \end{tabular}
@@
 \begin{figure}[h]
-        \center{\includegraphics[width=0.5\textwidth]{../img/synth2_alphaimg_hd.png}}
+        \center{\includegraphics[width=65mm]{../img/synth2_alphaimg_hd.png}}
         \caption{\label{fig:synth2_alphaimg} Sample from posterior of the state for synthetic experiment two, where color represents assignment and the mean and standard deviation are shown.
@@
 The MCMC sampler was successful for both benchmark videos; each object was detected, tracked, and its shape estimated in manner very consistent with the ground-truth.
-The results for the PETS2000 dataset are displayed in Figure~\ref{fig:pets2000_results} and for the PETS2001 dataset in Figure~\ref{fig:alpha_overlay_1}; in these figure, a sample from the posterior of the cluster parameters is overlayed the data and video frames at three time points, where the data is colored by assignment.
+The results for the PETS2000 dataset are displayed in Figure~\ref{fig:pets2000_results} and for the PETS2001 dataset in Figure~\ref{fig:pets2001_alphaimg}; in these figure, a sample from the posterior of the cluster parameters is overlayed the data and video frames at three time points, where the data is colored by assignment.
 \willie{also show pets 2001 result and per-frame overlays for both}.
 
-To calculate performance metrics (both SFDA and ATA), one must specify a confidence value that allows the oval representing the region occupied by an object to be computed from the inferred covariance matrix of each cluster (as discussed in Section~\ref{sec:	inference_to_results}).
-The performance metrics were found for a range of confidence intervals, and these curves, for the PETS2000 and PETS2001 video are shown in Figure~\ref{fig:pm_conf}.
-
-% \begin{figure}
-% 	\centering
-% 	\subfloat[]{\includegraphics[width=0.52\textwidth]{../img/pets2001_results_1}}
-% 	\subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2001_results_2}}
-% 	\caption{Sample from posterior of the state for the PETS2001 dataset, where color represents assignment and the mean and standard deviation are shown.
-\willie{will show assignments by marker type in addition to color}}
-% 	\label{fig:pets2001_results}
-% \end{figure}
-
 \begin{figure}[h]
-        \center{\includegraphics[width=85mm]{../img/pets2001_alphaimg_hr.png}}
-        \caption{\label{fig:alpha_overlay_1} Extracted data and sample from posterior of the model overlaid on three frames from the pets2001 video.}
+        \center{\includegraphics[width=65mm]{../img/pets2001_alphaimg_hr.png}}
+        \caption{\label{fig:pets2001_alphaimg} Extracted data and sample from posterior of the model overlaid on three frames from the PETS2001 video.}
 \end{figure}
 
-\begin{figure}
+\begin{figure}[h]
 	\centering
-	\subfloat[]{\includegraphics[width=0.6\textwidth]{../img/pets2000_results_1}}
-	\subfloat[]{\includegraphics[width=0.40\textwidth]{../img/pets2000_4objects.png}}
-	\caption{Extracted data and sample (a.) from posterior of the model overlaid (b.) on a frame from the pets2000 video}
+	\subfloat[]{\includegraphics[width=0.58\textwidth]{../img/pets2000_results_1}}
+	\subfloat[]{\includegraphics[width=0.42\textwidth]{../img/pets2000_4objects.png}}
+	\caption{Extracted data and a sample from posterior of the model (a) overlaid on a frame from the PETS2000 video (b).}
 	\label{fig:pets2000_results}
@@
 
-\begin{figure}
+To calculate performance metrics (both SFDA and ATA), one must specify a confidence value that allows the oval representing the region occupied by an object to be computed from the inferred covariance matrix of each cluster (as discussed in Section~\ref{sec:inference_to_results}).
+The performance metrics were found for a range of confidence intervals, and these curves, for the PETS2000 and PETS2001 video are shown in Figure~\ref{fig:pm_conf}.
+
+\begin{figure}[h]
+  \centering             
+  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2000_pm}}
+  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2001_pm}}
+  \caption{The SFDA (blue solid line) and ATA (red dashed line are) vs  confidence values from which an object's oval region is computed for (a) PETS2000 and (b) PETS2001 video datasets.}
+  \label{fig:pm_conf}
+\end{figure}
+
+\begin{figure}[h]
   \centering             
@@
   \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_4}}
-  \caption{Four frames $f=25, 35, 45, 55$ from the PETS2001 sequence with one posterior sample mean and covarance matrix overlaid. \willie{will show path trail over prev 10 frames}}
+  \caption{Four frames $f=25, 35, 45, 55$ from the PETS2001 sequence with one posterior sample mean and covarance matrix overlaid.
+\willie{will show path trail over prev 10 frames} \willie{these are old images!...
+haven't replaced yet}}
   \label{fig:pets2001_imgs}
-\end{figure}
-
-\begin{figure}
-  \centering             
-  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2000_pm}}
-  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2001_pm}}
-  \caption{The SFDA (blue solid line) and ATA (red dashed line are) vs  confidence values from which an object's oval region is computed for (a.) PETS2000 and (b.) PETS2001 video datasets.}
-  \label{fig:pm_conf}
 \end{figure}
@@
 
-\begin{figure}
+\subsubsection{Sensitivity Analysis}
+
+SMC inference on the PETS2009 video dataset was carried out for a range of two key parameter values, $\alpha$ and $\rho$.
+The performance metric measures, SFDA and ATA, were computed for each combination of these two parameters.
+This sensitivity investigation focused on these parameters due to their potential to have a large effect on object detection accuracies.
+The $\alpha$ values tested included $\{ 0.01, 0.1, 1, 10, 100 \}$, and the $\rho$ values tested included $\{ 0.7, 0.75, 0.8, 0.85, 0.9 \}$.
+
+The following shows the SFDA sensitivity analysis results. The largest value is in bold.
+\begin{center}
+\begin{tabular}[c]{l l  l  l  l  l  l}
+  & & & $\hspace{8mm}\rho$ & & & \\
+  SFDA  &  	& $0.7$ & $0.75$ & $0.8$ & $0.85$ & $0.9$  \\  
+  &$0.01$ 	& $0.4121$ & $0.4262$ & $0.4219$ & $\bold{0.4331}$ & $0.4277$  \\
+  &$0.1$  	& $\bold{0.4336}$ & $0.4184$ & $0.4200$ & $0.4316$ & $0.4321$  \\ 
+$\alpha$ 	& $1$    & $0.4084$ & $0.4209$ & $0.4205$ & $0.4290$ & $0.4330$  \\  
+  &$10$   	& $0.3992$ & $0.4130$ & $0.4289$ & $0.4232$ & $0.4226$  \\  
+  &$100$  	& $0.3655$ & $0.3739$ & $0.3636$ & $0.3755$ & $0.3672$  \\
+\end{tabular}
+\end{center}
+
+The following shows the ATA sensitivity analysis results. The largest value is in bold.
+\begin{center}
+\begin{tabular}[c]{l l  l  l  l  l  l}
+		& & & $\hspace{8mm}\rho$ & & & \\
+ 	 ATA  & 	 	& $0.7$ & $0.75$ & $0.8$ & $0.85$ & $0.9$  \\  
+	  &$0.01$ 		& $0.2032$ & $0.2017$ & $0.2262$ & $0.1945$ & $0.1914$  \\
+		  &$0.1$ 	& $0.1964$ & $0.2042$ & $0.2401$ & $0.1990$ & $0.2017$  \\
+$\alpha$ & $1$  	& $0.2016$ & $0.2365$ & $0.2766$ & $0.2347$ & $0.2766$  \\   
+  		&$10$   	& $0.2277$ & $0.2284$ & $0.2613$ & $\bold{0.2781}$ & $0.2732$  \\  
+ 		 &$100$ 	& $0.2468$ & $0.2645$ & $0.2725$ & $0.2670$ & $\bold{0.2843}$  \\
+\end{tabular}
+\end{center}
+
+% By inspecting the figures, one can see that there is a point where the SFDA and ATA both achieve a maximal value.
+
+\begin{figure}[h]
   \centering             
-  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_1}}
-  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_2}}
-  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_3}}
-  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_4}}
-  \caption{Four frames $f=10, 25, 35, 75$ from the PETS2009 sequence with one posterior sample mean and covarance matrix overlaid. \willie{will show path trail over prev 10 frames}}
-  \label{fig:pets2009_imgs}
+  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2009full_sens_sfda.png}}
+  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2009full_sens_ata.png}}
+  \caption{Surface plot of the sensitivity performance metric tables}
+  \label{fig:sens_surf}
 \end{figure}
 
-\begin{figure}
-	\centering
-	\includegraphics[width=3in]{../img/pets2009_results}
-	% \subfloat[]{\includegraphics[width=0.52\textwidth]{../img/pets2001_results_1}}
-	% \subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2001_results_2}}
-	\caption{Sample from posterior of the state for the PETS2009 dataset, where color represents assignment and the mean and standard deviation are shown.
-\willie{will show assignments by marker type in addition to color}}
-	\label{fig:pets2009_results}
-\end{figure}
-
-\subsubsection{Sensitivity Analysis}
-
-% \subsubsection{PETS2004/Caviar Datasets}
-% The Caviar datasets used in PETS2004 consist of multiple test-case-scenarios involving one or more humans.
 % \begin{figure}
 %   \centering             
-% 	\includegraphics[width=3in]{../img/pets2004_results_1}
-% 	\caption{Sample from posterior of the mean for frames $f = 1-20$ overlaid on frame $f=20$.}
+%   \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2009_result1.png}}
+%   \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2009_result2.png}}
+%   \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2009_result3.png}}
+%   \caption{Surface of the sensitivity performance metric tables}
+%   \label{fig:sens_surf}
+% \end{figure}
+
+% \begin{figure}
+%   \centering             
+%   \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_1}}
+%   \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_2}}
+%   \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_3}}
+%   \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_4}}
+%   \caption{Four frames $f=10, 25, 35, 75$ from the PETS2009 sequence with one posterior sample mean and covarance matrix overlaid.
+\willie{will show path trail over prev 10 frames}\willie{these are old images!...
+haven't replaced yet}}
+%   \label{fig:pets2009_imgs}
+% \end{figure}
+
+% \begin{figure}
+% 	\centering
+% 	\includegraphics[width=3in]{../img/pets2009_results}
+% 	% \subfloat[]{\includegraphics[width=0.52\textwidth]{../img/pets2001_results_1}}
+% 	% \subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2001_results_2}}
+% 	\caption{Sample from posterior of the state for the PETS2009 dataset, where color represents assignment and the mean and standard deviation are shown.
+\willie{will show assignments by marker type in addition to color}}
 % 	\label{fig:pets2009_results}
@@
 
-\subsection{New Video Datasets}
-
-The two video datasets in this section are video datasets first used in this paper. The datasets and groundtruth detailing the positions of each obect in each video will be published online.
-
-\subsubsection{Ants Video?}
-
-\subsubsection{Cells Video?}
+% \subsection{New Video Datasets}
+
+% The two video datasets in this section are video datasets first used in this paper. The datasets and groundtruth detailing the positions of each obect in each video will be published online.
+
+% \subsubsection{Ants Video?}
+
+% \subsubsection{Cells Video?}
 

### 2011-12-18 ltx/paper.tex
more edits. removed pretty worm picture

@@
 We define the task of automated detection and tracking of arbitrary objects in videos to be the fulfillment of three actions: determining the spatiotemporal regions of a video that constitute objects (`extraction'), finding the positions and/or shapes of distinct objects (`localization'), and maintaining the identities of the detected objects over time (`tracking').
-An ability to carry out these tasks in an automated manner is useful for many fields that make use of video data, including robotics, video surveillance, computational biology, and video summarization.
-Algorithms proving to be efficacious for a variety of objects and video types are particularly applicable.
+An ability to carry out these tasks in an automated manner is useful for many fields that make use of video data, including robotics, video surveillance, time-lapse microscopy, and video summarization.
+Algorithms able to be applied to a variety of objects and video types are particularly desirable.
 This paper serves to introduce a new technique involving the use of dependent Dirichlet processes mixture models, which we hope will provide the foundation for a class of unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.
@@
 A widely applied, though unsophisticated, example of general detection and tracking is the technique of blob tracking.
-Blob tracking typically refers to methods that perform extraction and localization of objects at each frame independently--for example, by assuming that pixels representing an object are distinguishable from pixels representing the background via color, motion, or texture information, and segmenting the extracted data at each frame into `blobs' to accomplish localization.
+Blob tracking typically refers to methods that perform extraction and localization of objects at each frame independently--for example, by assuming that pixels representing an object are distinguishable from pixels representing the background via color, motion, or texture information, and segmenting the extracted data at each frame into ``blobs'' to accomplish localization.
 A number of tracking techniques based on finding blobs in consecutive frames of a video with similar locations and appearances have been attempted to maintain the identity of detected blobs over time.
-The main challenges, however, associated with blob tracking involve the difficulty of segmenting neighboring or occluding objects in a given frame \cite{zhao2004tracking}, preventing multiple tracked blobs from collapsing onto and tracking a single object \cite{vermaak_2003}.
-These difficulties stem from the typical blob filter's need to perform detection via segmentation on individual frames and lack of a principled way to perform tracking while resisting the collapse of independent tracks.
-
-This paper presents the use of a Bayesian nonparametric mixture model for unsupervised multiple object detection and tracking.
-This method is intended to allow for robust detection of arbitrary objects (particularly in cases where frame-by-frame segmentation is difficult) and better isolation of independent objects during tracking.
+The main challenges, however, associated with blob tracking involve the difficulty of segmenting neighboring or occluding objects in a given frame \cite{zhao2004tracking} and preventing multiple tracked blobs from collapsing onto and tracking a single object \cite{vermaak_2003}.
+These difficulties stem from the typical blob tracker's need to perform detection via segmentation on individual frames and lack of a principled way to perform tracking while resisting the collapse of independent tracks.
+
+This paper presents the use of a Bayesian nonparametric mixture model for general, unsupervised multiple object detection and tracking.
+This method is intended to provide a widely applicable way to detect arbitrary objects in videos--particularly in cases where frame-by-frame segmentation is difficult or video quality is low and extraction is noisy--and better maintain the isolation of independent objects during tracking.
 Specifically, this work describes how a Dirichlet process mixture model can accomplish these goals when it is applied to data that may be extracted in a general manner from videos.
-We will demonstrate how this generative model may be applied to types of noisy data that can be quickly extracted from objects in videos, and how this model allows for automated detection and tracking.
-The model is first presented in an abstract form, to allow for it to be used with arbitrary extraction procedures and object appearance representations that may be desired in future studies.
+We will demonstrate how to extract data from objects in videos and how this generative model allows for automated detection and tracking.
+The model is first presented in an abstract form, to allow it to be used with arbitrary extraction procedures and object appearance representations that may be desired in future studies.
 We then formulate a specific model for data gained during a basic extraction procedure and describe inference algorithms that allow object localization and tracking to be carried out.
@@
 
-The task of extraction is related to the tasks of foregound segmentation, background modeling, motion detection, and appearance feature extraction.
+Extraction is related to the tasks of foregound segmentation, background modeling, motion detection, and appearance feature extraction.
 In general, these methods aim to extract features from regions of a video frame that are considered objects or are distinguished in some way from the background of a scene.
@@
 
-The task of localization is closely related to the topics of object detection, position-finding, segmentation, and target recognition.
+Localization is closely related to the topics of object detection, position-finding, segmentation, and target recognition.
 These methods are mainly concerned with finding the positions and/or shapes of objects in a given video frame, or with discerning between objects similar in position or appearance.
@@
 
-The task of tracking incorporates the topics of video tracking, data association, filtering, point matching, and `tracking of non-rigid objects'.
+Tracking is related to the topics of video tracking, data association, filtering, point matching, and `tracking of non-rigid objects'.
 In general, these tasks aim to maintain the localization of the position and/or shape of an object over time.
@@
 We defined the task of extraction to be the unsupervised determination of which spatiotemporal regions of a video constitute objects, and outlined a number of extraction methods in Section~\ref{sec:relatedwork}.
-The features of the data are characteristics of pixels in the regions of the video that contain objects, and the data are gained during extraction.
-For example, at its most basic, an extraction dataset is comprised of the three-dimensional points $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $x_{1}$ and $x_{2}$ are the two spatial dimensions of a pixel contained in a region representing an object in a given frame, and $t$ is the discrete time index of the frame.
-These three features are always available once regions containing objects are discerned, and are necessary features for modeling (as position information is key for reasonable tracking and time information is necessary for maintaining the temporal dependencies in the data).
+Extraction allows us to generate a collection of data points.
+Each data point can be thought of as a hypothesis about, or noisy version of, the representation of an object; a collection of hypotheses for a given object is assumed to be distributed in some way around the true object representation.
+A video of multiple objects is therefore assumed to yield data that contains multiple groups of object hypotheses with similar characteristics (i.e.
+clusters of similar hypotheses).
+
+Determining the regions of each video frame that contain objects allows us to generate these object hypotheses; namely, we let each hypothesis be a description of the pixel data over an area centered at a point in the extracted region.
+We can, for example, let each pixel in the extracted region be a hypothesis, and let the features of the representation consist of the spatial position of the pixel.
+An extraction dataset of this form is comprised of the three-dimensional points $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $x_{1}$ and $x_{2}$ are the two spatial dimensions of a pixel contained in an extracted region in a given frame, and $t$ is the discrete time index of the frame.
+These three features are always available once regions containing objects are discerned, and are useful features for modeling (as position information is key for reasonable tracking and time information is necessary for maintaining the temporal dependencies in the data).
 
@@
 
-We will use the phrase `baseline' three-dimensional extraction data to refer to the $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$ from pixels present in regions containing objects.
+We will use the phrase ``baseline'' three-dimensional extraction data to refer to the features $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$ of pixels present in regions containing objects.
 These data are distributed roughly as worm-like shapes that follow the paths of individual objects as they move, when they are plotted in three dimensions (where we let the frame number $t$, our discrete representation of time, ascend the vertical axis and the spatial positions occupy the horizontal axes).
@@
 The data consists of the three-dimensional baseline features and was gained by extraction peformed on a 50 frame sequence showing five ants moving across the video scene.
-% \begin{figure*}[h]
-%         \center{\includegraphics[width=85mm]{../img/usbh_data_2.png}}
-%         \caption{\label{fig:baselinedata} Baseline three-dimensional extraction data gained from a video segment containing 5 moving ants.
-The vertical axis denotes the frame number and the horizontal axes denote $d_{1}$ and $d_{2}$, the spatial coordinates of pixels within a given frame.
-Markers are colored by position for ease of viewing.
-\willie{show a number of baseline data here.
-consider whether colors help view or just add confusion.}}
-% \end{figure*}
-
-\begin{figure}[h]
-  \centering
-  \subfloat[]{\label{fig:baselinedata1}\includegraphics[width=0.33\textwidth]{../img/pets2001_data_1.png}} 
-  \subfloat[]{\label{fig:baselinedata1}\includegraphics[width=0.33\textwidth]{../img/usbh_data_3.png}} 
-  \subfloat[]{\label{fig:baselinedata2}\includegraphics[width=0.33\textwidth]{../img/traffic2_data_1.png}}\\
-  \subfloat[]{\label{fig:baselinedata1}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_1.png}} 
-  \subfloat[]{\label{fig:baselinedata1}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_2.png}} 
-  \subfloat[]{\label{fig:baselinedata2}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_3.png}}
-  \caption{Baseline extracted features from the (a.) PETS2001 Dataset, (b.) Ants Dataset, (c.) Traffic Dataset, and PETS2009/2010 benchmark dataset, frames (a.) 1-200, (b.) 201-400, (c.) 401-600.
+
+% \begin{figure}[h]
+%   \centering
+%   \subfloat[]{\label{fig:baselinedata1}\includegraphics[width=0.33\textwidth]{../img/pets2001_data_1.png}} 
+%   \subfloat[]{\label{fig:baselinedata2}\includegraphics[width=0.33\textwidth]{../img/usbh_data_3.png}} 
+%   \subfloat[]{\label{fig:baselinedata3}\includegraphics[width=0.33\textwidth]{../img/traffic2_data_1.png}}\\
+%   \subfloat[]{\label{fig:baselinedata4}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_1.png}} 
+%   \subfloat[]{\label{fig:baselinedata5}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_2.png}} 
+%   \subfloat[]{\label{fig:baselinedata6}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_3.png}}
+%   \caption{Baseline extracted features from the (a.) PETS2001 Dataset, (b.) Ants Dataset, (c.) Traffic Dataset, and PETS2009/2010 benchmark dataset, frames (a.) 1-200, (b.) 201-400, (c.) 401-600.
 Time is shown on the vertical axis.
 Marker color corresponds with spatial coordinates and is shown for ease of viewing.}
-  \label{fig:baselinedata_both}
-\end{figure}
+%   \label{fig:baselinedata}
+% \end{figure}
+
 Inference performed on the time dependent Dirichlet process mixture model (Section~\ref{sec:inference}) carries out clustering of the data, which isolates each worm-like data cloud and allows each object to be localized and tracked.
@@
 \end{equation}
-where $c_{i} \in \{ 1, \ldots, K \}$ denotes the assignment of $x_{i}$ to a given mixture, which allows the coefficients of the linear combination to be written as $P(c_{i} = k)$.
-Additionally, $\theta_{k}$ denotes the parametric form of the $k^{\text{th}}$ mixture component.
-We also define $p_{k} = P(c_{i} = k)$ for all $k \in \{ 1, \ldots, K \} $.
-This model can also be written generatively as
+where $c_{i} \in \{ 1, \ldots, K \}$ denotes the assignment of $x_{i}$ to a given mixture (which allows the coefficients of the linear combination to be written as $P(c_{i} = k)$), and $\theta_{k}$ denotes the parametric form of the $k^{\text{th}}$ mixture component.
+We also define $p_{k} = P(c_{i} = k)$ for $k \in \{ 1, \ldots, K \} $.
+We can therefore write this model generatively as
 \begin{align}
 \begin{split}
-	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}}) \\
 	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
+	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}})
 \end{split}
@@
 In this case, the prior distribution $\mathbb{G}_{0}$ is placed on the mixture component parameters, and the prior distribution $\text{Dir}(\alpha/K, \ldots, \alpha/K)$ is placed on the K mixture component assignment weights.
-The resulting Bayesian mixture model can be formulated in a hierarchical representation as
+The resulting Bayesian mixture model can be formulated in a generative representation as
 \begin{align}
@@
 \label{bayesian_mixture_model}
-	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}}) \\
+	p_{1}, \ldots, p_{K}  &\sim  \text{Dir}(\alpha/K, \ldots, \alpha/K)\\
 	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
 	\theta_{c_{i}}  &\sim  \mathbb{G}_{0} \\
-	p_{1}, \ldots, p_{K}  &\sim  \text{Dir}(\alpha/K, \ldots, \alpha/K)
+	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}})
 \end{split}
@@
 \begin{split}
-	x_{i}|\phi_{i} &\sim F(\phi_{i}) \\
+	\mathbb{G} | \alpha, \mathbb{G}_{0}  &\sim  \text{DP}(\alpha, \mathbb{G}_{0}) \\
 	\phi_{i} | \mathbb{G}  &\sim  \mathbb{G} \\
-	\mathbb{G} | \alpha, \mathbb{G}_{0}  &\sim  \text{DP}(\alpha, \mathbb{G}_{0})
+	x_{i}|\phi_{i} &\sim F(\phi_{i})
 \end{split}
@@
   \caption{Each plot shows a sample from the posterior distribution of the state for one of the videos in synthetic experiment one, where the time axis ascends vertically, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.
-In both cases, the two objects are successfully tracked through occlusion, though the objects reverse direction in the second video.
+In both cases, the two objects are successfully tracked through occlusion, though both objects reverse their directions in the second video.
 \willie{will show assignments by marker type in addition to color}}
@@
 The MCMC sampler was successful for both benchmark videos; each object was detected, tracked, and its shape estimated in manner very consistent with the ground-truth.
-The results for the PETS2000 dataset are displayed in Figure~\ref{fig:pets2000_results} and for the PETS2001 dataset in Figure~\ref{fig:pets2001_alphaimg}; in these figure, a sample from the posterior of the cluster parameters is overlayed the data and video frames at three time points, where the data is colored by assignment.
+The results for the PETS2000 dataset are displayed in Figure~\ref{fig:pets2000_results} and for the PETS2001 dataset in Figure~\ref{fig:pets2001_alphaimg}; in these figure, a sample from the posterior of the cluster parameters is overlayed the data and video frames at three points in time, where the data is colored by assignment.
 \willie{also show pets 2001 result and per-frame overlays for both}.
@@
   \caption{Four frames $f=25, 35, 45, 55$ from the PETS2001 sequence with one posterior sample mean and covarance matrix overlaid.
-\willie{will show path trail over prev 10 frames} \willie{these are old images!...
+\willie{will show path trail over prev 10 frames.} \willie{These are old images!...
 haven't replaced yet}}
@@
 Due to the large number of frames and objects in this video, the SMC algorithm (described in Section~\ref{sec:SMC}) was used for inference.
-This method of sequential inference was observed, on this dataset, to converge to a better sample in a shorter period of time in comparison with the MCMC algorithm.
-For this reason, we found that it was better suited for this large dataset.
+This method of sequential inference was observed, on this dataset, to converge to a better sample in a shorter period of time in comparison with the MCMC algorithm; for this reason, we found that it was better suited for this large dataset.
 
 Extraction was performed with frame differencing as described in Section~\ref{sec:modelspec_extraction}, using $L=3$ and parameters for the model were chosen, using the method outlined in Section~\ref{sec:parameter_recap}, to be $\alpha = 0.1, \rho = 0.8, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 6, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (3, \ldots, 3)$.
-Additionlly, as with the other video datasets, ground-truth was authored using the ViPER tool.
-
-Sequential inference produced reasonably accurate results to be achieved in a short timespan.
+Additionally, as with the other video datasets, ground-truth bounding boxes around each object were authored using the ViPER tool.
+
+The SMC inference algorithm yielded a sample from the posterior distribution, from which the object detection and tracking results were obtained (as described in Section~\ref{sec:inference_to_results}).
+Figure \willie{provide figure showing PETS2009 results} shows a sample from the posterior of the cluster parameters overlayed on the data, and video frames at three points in time, where the data is colored by assignment
+\willie{show SFDA and ATA vs cov mat conf for this sample that I show figures for.}
 

### 2011-12-19 ltx/paper.tex
more edits. first draft of the tricky figure is in there

@@
 
-More rigorously, we take the definition of a DDPM to be a stochastic process defined on the space of probability distributions over a domain, which are indexed by time, space, or a selection of other covariates in such a way that the marginal distribution at any point in the domain follows a Dirichlet process (adapted from definitions found in \willie{cite griffin and steel and gasthaus-thesis}).
+More rigorously, we take the definition of a DDPM to be a stochastic process defined on the space of probability distributions over a domain, which are indexed by time, space, or a selection of other covariates in such a way that the marginal distribution at any point in the domain follows a Dirichlet process (adapted from definitions found in \cite{gasthaus_thesis} and \cite{griffin2006order}).
 Hence, a time-dependent DDPM is a model which remains a Dirichlet process, marginally, at each time step, yet allows cluster parameters at a given time step to vary from (and remain dependent upon) the parameters in neighboring time steps.
@@
 Hence, to account for varying numbers of clusters--and in particular, to allow clusters to diminish in size (i.e.
-reduce the number of assigned observations) and even die--there is a deletion procedure (described in \cite{caron_2007}) by which observations are considered `removed' from their assigned clusters at a given time step and value of $m_{k}$ (defined in \eqref{crp_rep} to be the size of cluster $k$) is reduced.
+reduce the number of assigned observations) and even die--there is a deletion procedure (described in \cite{caron_2007}) by which observations are considered ``removed'' from their assigned clusters at a given time step and value of $m_{k}$ (defined in \eqref{crp_rep} to be the size of cluster $k$) is reduced.
 
 The deletion procedure operates in the following way: at each time step, all previous assignments (that have not yet been deleted) are independently considered for deletion.
-Specifically, each (remaining) assignment is `removed' from its cluster with probability $\rho$.
+Specifically, each (remaining) assignment is removed from its cluster with probability $\rho$.
 We can consider the size $m_{k,t}$ of cluster $k$ at time $t$ to be dependent upon the cluster size at time $t-1$ via a deletion step.
-In fact, it can be shown that performing deletion in this way is equivalent to, for each cluster $k$, drawing $r \sim \text{Binomial}(m_{k,t-1}, \rho)$, and reducing the cluster size $m_{k,t}$ by $r$.
+It can be shown that performing deletion in this way is equivalent to, for each cluster $k$, drawing $r \sim \text{Binomial}(m_{k,t-1}, \rho)$, and reducing the cluster size $m_{k,t}$ by $r$.
 This can be written as
@@
 This process adheres to what \cite{caron_2007} calls a `uniform deletion strategy' over all the observations' assignments (since assignments to each cluster have equal probability of being deleted), though a more complex deletion strategy, dependent upon cluster size, can also be implemented \cite{caron_2007}.
-For a given time $t$ and cluster $k$, the size $m_{k,t}$ is dependent upon the cluster size at the previous time step (via \eqref{del_step}) and on the assignments for observations at time $t$ ($c_{i}$ for $i \in \{ 1, \ldots, N_{t} \}$, where $N_{t}$ is the number of observations at time step $t$).
+For a given time $t$ and cluster $k$, the size $m_{k,t}$ is dependent upon the cluster size at the previous time step (via \eqref{del_step}) and on the assignments $c_{i}$ for observations at time $t$.
 We refer to this distribution over $m_{k,t} | m_{k,t-1}, c_{i}, \rho$ as DEL.
-Using the cluster size and observation deletion terms introduced above, we can define the GPUDDPM generatively as, for each time step $t \in \{1, \ldots, T\}$ and observation $i \in \{ 1, \ldots, N_{t} \}$ at $t$,
-			% \begin{align}
-			% \begin{split}
-			% 	x_{i}^{(t)} | c_{i}^{(t)}, \theta_{c_{i}^{(t)}}^{(t)} &\sim F(\theta_{c_{i}^{(t)}}^{(t)}) \\
-			% 	c_{i}^{(t)} | m^{(t-1)}, \alpha  &\sim  \text{CRP}(m^{(t-1)}, \alpha) \\
-			% 	m^{(t-1)} | m^{(t-2)}, c_{i}^{(t-1)}  &\sim \text{DEL}(m^{(t-2)}, c_{i}^{(t-1)}, \rho, \gamma) \\
-			% 	\theta_{k}^{(t)} | \theta_{k}^{(t-1)}   &\sim 
-			% \begin{cases}
-			% 	P(\theta_{k}^{(t)} | \theta_{k}^{(t-1)}) \\
-			% 	\mathbb{G}_{0}
-			% \end{cases}
-			% \end{split}
-			% \end{align}
+Using the cluster size and observation deletion terms introduced above, we can define the GPUDDPM generatively as, for each time step $t \in \{1, \ldots, T\}$,  % and observation $i \in \{ 1, \ldots, N_{t} \}$ at $t$,
 \begin{align}
 \begin{split}
-	\bold{x}_{i} | c_{i}, \theta_{c_{i}, t} &\sim F(\theta_{c_{i}, t}) \\
-	c_{i} | \bold{m}_{i-1}, \alpha  &\sim  \text{CRP}(\bold{m}_{i-1}, \alpha) \\
-	\bold{m}_{1} | \bold{m}_{N_{t-1}}, \rho  &\sim \text{DEL}(\bold{m}_{N_{t-1}}, \rho) \\
-	\theta_{t} | \theta_{t-1}   &\sim 
+\bold{m}_{t} | \bold{m}_{t-1}, \rho  &\sim \text{DEL}(\bold{m}_{t-1}, \rho) \\
+c_{i} | \bold{m}_{t}, \alpha  &\sim  \text{CRP}(\bold{m}_{t}, \alpha) \\
+\theta_{c_{i}, t} | \theta_{c_{i}, t-1}   &\sim
 \begin{cases}
-	P(\theta_{t} | \theta_{t-1}) \\
+	P(\theta_{c_{i}, t} | \theta_{c_{i}, t-1}) \\
 	\mathbb{G}_{0}
-\end{cases}
+\end{cases} \\
+\bold{x}_{i} | c_{i}, \theta_{c_{i}, t} &\sim F(\theta_{c_{i}, t})
 \end{split}
 \end{align}
-where $\text{CRP}(\cdot, \cdot)$ is given in \eqref{crp_rep}. The graphical model corresponding with this formulation is shown in Figure \ref{fig:gpuddpm_gm_1}.
+where $\text{CRP}(\cdot, \alpha)$ is given in \eqref{crp_rep}.
+The graphical model corresponding with this formulation is shown in Figure \ref{fig:gpuddpm_gm_1}.
+The above formulation holds if there is exactly one observation at each time step.
+Often, the data extracted from objects in videos consists of multiple observations at each time step, each with an assignment variable ($c_{i}$ for $i \in \{ 1, \ldots, N_{t} \}$, where $N_{t}$ is the number of observations at time step $t$).
+If we allow for multiple observations $i \in \{ 1, \ldots, N_{t} \}$ at each time $t$, this formulation becomes,
+\begin{align}
+\begin{split}
+\bold{m}_{t}^{1} | \bold{m}_{t-1}^{N_{t-1}}, \rho  &\sim \text{DEL}(\bold{m}_{t-1}^{N_{t-1}}, \rho) \\
+c_{i} | \bold{m}_{t}^{i-1}, \alpha  &\sim  \text{CRP}(\bold{m}_{t}^{i-1}, \alpha) \\
+\theta_{c_{i}, t} | \theta_{c_{i}, t-1}   &\sim
+\begin{cases}
+	P(\theta_{c_{i}, t} | \theta_{c_{i}, t-1}) \\
+	\mathbb{G}_{0}
+\end{cases} \\
+\bold{x}_{i} | c_{i}, \theta_{c_{i}, t} &\sim F(\theta_{c_{i}, t})
+\end{split}
+\end{align}
+where $\bold{m}_{t}^{i}$ represents the $i^{\text{th}}$ observation at time $t$.
 
@@
   \subfloat[]{\label{fig:pets2009_img2}\includegraphics[width=0.333\textwidth]{../img/pets2009_frame2.png}}
-  \subfloat[]{\label{fig:pets2009_img_framediff}\includegraphics[width=0.333\textwidth]{../img/pets2009_framediff.png}}
-  \caption{(a., b.) Consecutive frames of a video containing four ants. (c.) Extraction results produced by taking the pixel-wise difference between these frames}
+  \subfloat[]{\label{fig:pets2009_img_framediff}\includegraphics[width=0.333\textwidth]{../img/pets2009_framediff.png}}\\
+  \subfloat[]{\label{fig:longimg}\includegraphics[width=1.05\textwidth]{../img/longimg_6.pdf}} 
+  \caption{Pairs of consecutive frames and the results produced by taking the pixel-wise difference between these frames.
+\willie{Want to add colorbar that explains how color represents frame number red = frame 1, blue = frame 100}}
   \label{fig:img_and_framediff}

### 2011-12-24 ltx/paper.tex
new edits. better svjour3 formatting. first draft of abstract. more author details.

@@
-\documentclass[smallcondensed, final]{svjour3}
+% \documentclass[smallcondensed, final, numbook]{svjour3}[12/23/2011]
+\documentclass[smallcondensed, final]{svjour3}[12/23/2011]
 % \documentclass[twocolumn, final]{svjour3}
@@
 \title{Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures}
-\author{Willie Neiswanger and Frank Wood}
+\titlerunning{Unsupervised Object Detection and Tracking with Dependent Dirichlet Processes}
+% \subtitle{A Bayesian Nonparametric Approach to Tracking Multiple Objects in Videos}
+% \subtitle{Subtitle?}
+\author{Willie Neiswanger$^{*}$ and Frank Wood$^{\dagger}$}
+\institute{
+	$^{*}$Columbia University, Department of Applied Math and Applied Physics. 
+	Tel.: 503-464-6152.
+	\email{wdn2101@columbia.edu}
+	\and
+	$^{\dagger}$Columbia University, Department of Statistics. 
+	Tel.: 212-851-2150. 
+	\email{fwood@stat.columbia.edu}}
+\date{}
 \maketitle
-\mbox{}
-
-\abstract{Main points.}
+% \mbox{}
+
+\begin{abstract}
+This paper proposes an unsupervised technique for the detection and tracking of arbitrary objects in videos.
+This method is intended to reduce the need for explicit target localization and serve as a general framework applicable to videos with varied object types, backgrounds, and film qualities.
+The technique uses a dependent Dirichlet process mixture (DDPM) known as the Generalized Polya Urn (GPUDDPM) to model image pixel data that can be extracted in a general manner from the regions in a video that represent objects.
+This paper describes a specific implementation of the model using spatial and color pixel data extracted via frame differencing and gives two algorithms for performing inference on the model to accomplish detection and tracking.
+This technique is demonstrated on multiple synthetic and benchmark video datasets that illustrate its ability to, without modification, detect and track variety of objects with diverse physical charactersitics over non-uniform backgrounds and through occlusion.
+% \keywords{Object Detection \and Tracking \and Bayesian Nonparametrics \and Multiple Target Tracking}
+\end{abstract}
 
@@
 Research in the past ten years involving the development of methods for general detection and tracking of objects has tended to focus on one of either extraction, localization, or tracking; attempts to integrate all three tasks and produce a system aimed at multiple arbitrary objects and diverse video types was not often a primary focus, or was attempted in an ad-hoc manner to test or apply a more specific technique.
-Research aimed at developing methods to accomplish the three tasks in a cohesive manner has been pursued more recently to produce algorithms that can, in a fully automated way, accomplish unsupervised detection and tracking--or in short, that can determine the spatiotemporal region of a video occupied by each distinct object.
+Research aimed at developing methods to accomplish the three tasks in a cohesive manner has been pursued more recently to produce algorithms that can, in a fully automated way, accomplish unsupervised detection and tracking---or in short, that can determine the spatiotemporal region of a video occupied by each distinct object.
 A widely applied, though unsophisticated, example of general detection and tracking is the technique of blob tracking.
-Blob tracking typically refers to methods that perform extraction and localization of objects at each frame independently--for example, by assuming that pixels representing an object are distinguishable from pixels representing the background via color, motion, or texture information, and segmenting the extracted data at each frame into ``blobs'' to accomplish localization.
+Blob tracking typically refers to methods that perform extraction and localization of objects at each frame independently---for example, by assuming that pixels representing an object are distinguishable from pixels representing the background via color, motion, or texture information, and segmenting the extracted data at each frame into ``blobs'' to accomplish localization.
 A number of tracking techniques based on finding blobs in consecutive frames of a video with similar locations and appearances have been attempted to maintain the identity of detected blobs over time.
@@
 This paper presents the use of a Bayesian nonparametric mixture model for general, unsupervised multiple object detection and tracking.
-This method is intended to provide a widely applicable way to detect arbitrary objects in videos--particularly in cases where frame-by-frame segmentation is difficult or video quality is low and extraction is noisy--and better maintain the isolation of independent objects during tracking.
+This method is intended to provide a widely applicable way to detect arbitrary objects in videos---particularly in cases where frame-by-frame segmentation is difficult or video quality is low and extraction is noisy---and better maintain the isolation of independent objects during tracking.
 Specifically, this work describes how a Dirichlet process mixture model can accomplish these goals when it is applied to data that may be extracted in a general manner from videos.
@@
 Additionally, color has been successfully applied as the feature in a technique known a the mean-shift procedure (or the kernel-based method) for tracking \citep{comaniciu_2003, perez_2002, nummiaro_2003, lee_2011}.
-This procedure--a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--optimizes the search for the region in a frame which is most similar to a target region in the previous frame.
+This procedure---a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--optimizes the search for the region in a frame which is most similar to a target region in the previous frame.
 It involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges for each data point.
@@
 the search for the region in a frame which is most similar to a target region in the previous frame).
-This procedure--a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges for each data point.
+This procedure---a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges for each data point.
 Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors.
@@
 Recall that the distribution over the cluster assignments $c_{i}$ (one for each observation $i$) is a function of the cluster sizes (see \eqref{crp_rep}).
-Hence, to account for varying numbers of clusters--and in particular, to allow clusters to diminish in size (i.e.
-reduce the number of assigned observations) and even die--there is a deletion procedure (described in \cite{caron_2007}) by which observations are considered ``removed'' from their assigned clusters at a given time step and value of $m_{k}$ (defined in \eqref{crp_rep} to be the size of cluster $k$) is reduced.
+Hence, to account for varying numbers of clusters---and in particular, to allow clusters to diminish in size (i.e.
+reduce the number of assigned observations) and even die---there is a deletion procedure (described in \cite{caron_2007}) by which observations are considered ``removed'' from their assigned clusters at a given time step and value of $m_{k}$ (defined in \eqref{crp_rep} to be the size of cluster $k$) is reduced.
 
@@
   \subfloat[]{\label{fig:pets2009_img_framediff}\includegraphics[width=0.333\textwidth]{../img/pets2009_framediff.png}}\\
-  \subfloat[]{\label{fig:longimg}\includegraphics[width=1.05\textwidth]{../img/longimg_6.pdf}} 
+  %\subfloat[]{\label{fig:longimg}\includegraphics[width=1.05\textwidth]{../img/longimg_6.pdf}} \\
+  \subfloat[]{\label{fig:longimg_all}\includegraphics[width=1\textwidth]{../img/longimg_7_detail.pdf}} 
   \caption{Pairs of consecutive frames and the results produced by taking the pixel-wise difference between these frames.
@@
 Both the multivariate normal and multinomial distributions were chosen because they are sufficiently simple (and well studied) to allow for tractable inference and sufficiently flexible to provide a reasonable approximation to the data gained during extraction.
-In particular, the multivariate normal distribution over the spatial features $\bold{x}^{s}$ can be thought to represent the shape of each object as an oval; likewise, data generated by each moving object during extraction are often ovular--as noisy extraction procedures cause some smoothing of edges and corners, producing blob-like shapes even when objects are not particularly round--and centered on a given object.
+In particular, the multivariate normal distribution over the spatial features $\bold{x}^{s}$ can be thought to represent the shape of each object as an oval; likewise, data generated by each moving object during extraction are often ovular---as noisy extraction procedures cause some smoothing of edges and corners, producing blob-like shapes even when objects are not particularly round---and centered on a given object.
 Furthermore, this model is justified as the maximum likelihood parameter estimate of a normal distribution corresponds to the least squares fit of data relative to the mean of the distribution.
@@
 This section details a Markov Chain Monte Carlo (MCMC) sampler used to perform inference, which provides a method to carry out unsupervised detection and tracking.
-It operates by iteratively sampling from conditional distributions of elements in the joint probability distribution--i.e.
-it implements Gibbs sampling--in order to gain approximate samples from the posterior distribution of the cluster parameters given the data.
+It operates by iteratively sampling from conditional distributions of elements in the joint probability distribution---i.e.
+it implements Gibbs sampling---in order to gain approximate samples from the posterior distribution of the cluster parameters given the data.
 Additionally, this sampler makes use of the Metropolis-Hastings (M-H) algorithm to sample from a few distributions.
@@
 
-\begin{algorithm}[h]
+\begin{algorithm}[h!]
 \label{alg:MH}
@@
 
-This section details a Sequential Monte Carlo (SMC) sampler--also known as a particle filter--used to perform inference, which provides another method to carry out unsupervised detection and tracking.
-An overview of SMC inference is given in Algorithm~\ref{alg:SMC}.
-\willie{general overview here: at each time step, perform sampling, and then show algorithm figure / thing}
+This section details a Sequential Monte Carlo (SMC) sampler---also known as a particle filter---used to perform inference, which provides another method to carry out unsupervised detection and tracking.
+The SMC inference algorithm is given in Algorithm~\ref{alg:SMC}.
+The algorithm operates by
 
@@
 \STATE Draw $c_{i} \sim P(c_{i} | m_{1:K_{t}, i-1}, \theta_{1:K_{t}, t} )$
+\IF{$c_{i} = K_{t} + 1$}
+\STATE Draw a new cluster $\theta_{K_{t}+1,t} \sim \mathbb{G}_{0}(\text{priors})$   \willie{distro after algo?}
+\STATE $K_{t} \leftarrow K_{t} + 1$
+\ENDIF
 \ENDFOR
@@
 
+A few key distributions used in the SMC algorithm are shown below.
+
 \subsubsection{Proposal Distribution for Assignments $c_{i}$}
@@
 
-Noisy extraction procedures may result in erroneous data points that do not correspond to any foreground object.
+Noisy extraction procedures may result in spurrious data points that do not correspond to any foreground object.
 We would like to incorporate these points into our model.
@@
 
-%, where a numerical search algorith--the Hungarian algorithm \cite{munkres_1957}--is used to find an optimal mapping from ground-truth to output tracks without requiring the performance metric value of all possible mapping combinations to be computed (reducing the time complexity from factorial to polynomial time)
+%, where a numerical search algorith---the Hungarian algorithm \cite{munkres_1957}--is used to find an optimal mapping from ground-truth to output tracks without requiring the performance metric value of all possible mapping combinations to be computed (reducing the time complexity from factorial to polynomial time)
 
@@
 The second half of the two videos differ in that both squares in the first video continue in the same direction and end at the other's starting positions at $f=200$, and both squares in the second video reverse directions and end at their initial starting positions at $f=200$.
-Frame difference extraction on each video yields a dataset with identical spatial features; hence, successful tracking depends fully on the reasonable incorporation of color information into the model.
+The frame difference extraction yields identical spatial features in both videos; hence, successful tracking depends fully on the incorporation of color information into the model.
 
@@
 
+Groundtruth for the synthetic videos
+
 \begin{figure}
   \centering               
-  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/cross_final}}
-  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/bounce_final}}
+  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/synth1_bounce.pdf}}
+  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/synth1_cross.pdf}}
   \caption{Each plot shows a sample from the posterior distribution of the state for one of the videos in synthetic experiment one, where the time axis ascends vertically, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.
@@
 Benchmark video datasets for object tracking and detection have been produced to provide standard scenes on which researchers can compare detection and tracking results.
-These videos have been primarily produced for surveillance-related workshops--notably, for the International Workshop on Performance Evaluation of Tracking and Surveillance (PETS)--which provide researchers with video datasets and algorithmic goals on which to focus.
-Three commonly used benchmark videos from PETS workshops--one used in PETS2000, one in PETS2001, and one used both in PETS2009 and PETS2010--were chosen to demonstrate the efficacy of the method presented in this study.
+These videos have been primarily produced for surveillance-related workshops---notably, for the International Workshop on Performance Evaluation of Tracking and Surveillance (PETS)--which provide researchers with video datasets and algorithmic goals on which to focus.
+Three commonly used benchmark videos from PETS workshops---one used in PETS2000, one in PETS2001, and one used both in PETS2009 and PETS2010---were chosen to demonstrate the efficacy of the method presented in this study.
 The performance metrics and benchmark datasets allow the methods developed in this paper to be quantitatively compared against other detection and tracking algorithms.

### 2011-12-25 ltx/paper.tex
edits. more figures. better formating. new image overlay figures (for PETS2001 dataset)

@@
+\documentclass[smallcondensed, final]{svjour3}
 % \documentclass[smallcondensed, final, numbook]{svjour3}[12/23/2011]
-\documentclass[smallcondensed, final]{svjour3}[12/23/2011]
 % \documentclass[twocolumn, final]{svjour3}
@@
 \titlerunning{Unsupervised Object Detection and Tracking with Dependent Dirichlet Processes}
+\authorrunning{Willie Neiswanger and Frank Wood}
 % \subtitle{A Bayesian Nonparametric Approach to Tracking Multiple Objects in Videos}
@@
 \date{}
+
 \maketitle
-% \mbox{}
 
@@
 This paper proposes an unsupervised technique for the detection and tracking of arbitrary objects in videos.
-This method is intended to reduce the need for explicit target localization and serve as a general framework applicable to videos with varied object types, backgrounds, and film qualities.
+It is intended to reduce the need for detection or localization methods tailored to specific objects and serve as a general framework applicable to videos with varied object types, backgrounds, and film qualities.
 The technique uses a dependent Dirichlet process mixture (DDPM) known as the Generalized Polya Urn (GPUDDPM) to model image pixel data that can be extracted in a general manner from the regions in a video that represent objects.
@@
   \centering               
-  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/synth1_bounce.pdf}}
-  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/synth1_cross.pdf}}
+  \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/synth1_cross.pdf}}
+  \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/synth1_bounce.pdf}}
+  \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/synth4.pdf}}
   \caption{Each plot shows a sample from the posterior distribution of the state for one of the videos in synthetic experiment one, where the time axis ascends vertically, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.
-In both cases, the two objects are successfully tracked through occlusion, though both objects reverse their directions in the second video.
-\willie{will show assignments by marker type in addition to color}}
+In both cases, the two objects are successfully tracked through occlusion, though both objects reverse their directions in the second video.}
   \label{fig:synth_one_plot}
@@
 
+\begin{figure}
+  \centering
+  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2000_front.pdf}}             
+  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2001_back.pdf}}
+  \caption{PETS2000 dataset results (a) and PETS2001 dataset results (b).
+Both plots show a sample from the posterior distribution of the state, where the vertical axis denots time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.}
+  \label{fig:benchmark_results}
+\end{figure}
+
+% \begin{figure}
+%   \centering               
+%   \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2000_front.pdf}}
+%   \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2000_back.pdf}}
+%   \caption{PETS2000 dataset results, from a front view (a) and a back view (b).
+Both plots show a sample from the posterior distribution of the state, where the vertical axis denots time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.}
+%   \label{fig:pets2000_results}
+% \end{figure}
+
+\begin{figure}
+  \centering               
+  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_f1.pdf}}
+  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_f3.pdf}}
+  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_f5.pdf}}
+  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_f7.pdf}}
+  \caption{PETS2001 results.  \willie{add tails. different color ovals for different objects.}}
+  \label{fig:pets2001_overlay}
+\end{figure}
+
 To calculate performance metrics (both SFDA and ATA), one must specify a confidence value that allows the oval representing the region occupied by an object to be computed from the inferred covariance matrix of each cluster (as discussed in Section~\ref{sec:inference_to_results}).

### 2011-12-29 ltx/paper.tex
more edits. new figures. re-do figures for pets2000 and pets2001 datasets

@@
 
+% frank says the following worm figures are a no go
 % \begin{figure}[h]
@@
 In particular, the Metropolis-Hastings (M-H) algorithm is used to gain samples from the posterior distribution for the auxiliary variables; it involves constructing a markov chain with the sought posterior as its stationary distribution, and sampling sequentially from this constructed chain to get an approximate posterior sample.
-The M-H algorithm is outlined in Algorithm~\ref{alg:MH}.
+The M-H algorithm is outlined in Algorithm 1.%Algorithm~\ref{alg:MH}.
 \willie{am i citing wrong?}
@@
 Before the samples can be iteratively drawn from each element of the state given all other elements, the state must be initialized.
-While the state of the sampler can theoretically be initialized at any value, beginning at a value closer to the maximum joint probability (over the state) allows for quicker inference.
+While the state of the sampler can be initialized at any value, beginning at a value closer to the maximum joint probability (over the state) allows for quicker inference.
 
@@
 These metrics were developed during VACE Phase II to provide a single, comprehensive metric to describe detection, and one to describe tracking.
-
-The following terms are used in the definitions of the performance metrics:
+The following are used in the definitions of the performance metrics:
 \begin{itemize}
@@
 \item $G_{i}$ denotes the spatiotemporal region occupied by the $i$th ground-truth object in a video, and $G_{i}^{(t)}$ denotes the region occupied by the $i$th ground-truth object in frame $t$.
+\vspace{2mm}
 \item $D_{i}$ denotes the spatiotemporal region occupied by the $i$th detected object in a video, and $D_{i}^{(t)}$ denotes the region occupied by the $i$th detected object in frame $t$.
+\vspace{2mm}
 \item $N_{G}$ denotes the total number of unique ground-truth objects in a video, and $N_{G}^{(t)}$ denotes the number of unique ground-truth objects present at frame $t$.
+\vspace{2mm}
 \item $N_{D}$ denotes the total number of unique detected objects in a video, and $N_{D}^{(t)}$ denotes the number of unique detected objects present at frame $t$.
+\vspace{2mm}
 \item $N_{\text{frames}}$ denotes the total number of frames in a video, and $N_{\text{frames}}^{(i)}$ denotes the number of frames in which an object $i$ (which can be a ground-truth or detected object, depending on the context) is present in a video.
+\vspace{2mm}
 \item $N_{\text{mapped}}$ denotes the number of mapped ground-truth/detect pairs in a video, and $N_{\text{mapped}}^{(t)}$ denotes the number of mapped ground-truth/detect pairs present at frame $t$.
@@
 
-\subsubsection{Tracking by Appearance}
+% \subsubsection{Tracking by Appearance}
 
@@
 
-Groundtruth for the synthetic videos
-
 \begin{figure}
   \centering               
-  \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/synth1_cross.pdf}}
-  \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/synth1_bounce.pdf}}
-  \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/synth4.pdf}}
+  \subfloat[]{\includegraphics[width=0.32\textwidth]{../img/synth1_cross.pdf}} \hspace{0.5mm}
+  \subfloat[]{\includegraphics[width=0.32\textwidth]{../img/synth1_bounce.pdf}} \hspace{0.5mm}
+  \subfloat[]{\includegraphics[width=0.32\textwidth]{../img/synth4.pdf}}
   \caption{Each plot shows a sample from the posterior distribution of the state for one of the videos in synthetic experiment one, where the time axis ascends vertically, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.
@@
 
-\subsubsection{Tracking though Occlusion and Appearance Shift}
+% \subsubsection{Tracking though Occlusion and Appearance Shift}
 
@@
 \willie{still need to include performance metric results for this synthetic dataset}.
+
+Groundtruth for the synthetic videos
 
@@
 
-\begin{figure}[h]
-        \center{\includegraphics[width=65mm]{../img/synth2_alphaimg_hd.png}}
-        \caption{\label{fig:synth2_alphaimg} Sample from posterior of the state for synthetic experiment two, where color represents assignment and the mean and standard deviation are shown.
+% \begin{figure}[h]
+%         \center{\includegraphics[width=65mm]{../img/synth2_alphaimg_hd.png}}
+%         \caption{\label{fig:synth2_alphaimg} Sample from posterior of the state for synthetic experiment two, where color represents assignment and the mean and standard deviation are shown.
 Both the sample and data are overlaid on three images from synthetic experiment.
@@
 \willie{will show assignments by marker type in addition to color}}
-\end{figure}
+% \end{figure}
 
@@
 The MCMC sampler was successful for both benchmark videos; each object was detected, tracked, and its shape estimated in manner very consistent with the ground-truth.
-The results for the PETS2000 dataset are displayed in Figure~\ref{fig:pets2000_results} and for the PETS2001 dataset in Figure~\ref{fig:pets2001_alphaimg}; in these figure, a sample from the posterior of the cluster parameters is overlayed the data and video frames at three points in time, where the data is colored by assignment.
-\willie{also show pets 2001 result and per-frame overlays for both}.
-
-\begin{figure}[h]
-        \center{\includegraphics[width=65mm]{../img/pets2001_alphaimg_hr.png}}
-        \caption{\label{fig:pets2001_alphaimg} Extracted data and sample from posterior of the model overlaid on three frames from the PETS2001 video.}
-\end{figure}
+The results for the PETS2000 dataset are displayed in Figure~\ref{fig:pets2000_results} and for the PETS2001 dataset in Figure~\ref{fig:pets2001_results}; in these figure, a sample from the posterior of the cluster parameters is overlayed on the extracted data over a sequence of frames, where the assignment of each data point is represented by color and marker type.
 
@@
   \centering
-	\subfloat[]{\includegraphics[width=0.58\textwidth]{../img/pets2000_results_1}}
-	\subfloat[]{\includegraphics[width=0.42\textwidth]{../img/pets2000_4objects.png}}
-	\caption{Extracted data and a sample from posterior of the model (a) overlaid on a frame from the PETS2000 video (b).}
-	\label{fig:pets2000_results}
+  \subfloat[]{\label{fig:pets2000_results} \includegraphics[width=0.48\textwidth]{../img/pets2000_front.pdf}} \hspace{1mm}           
+  \subfloat[]{\label{fig:pets2001_results} \includegraphics[width=0.48\textwidth]{../img/pets2001_back.pdf}}
+  \caption{PETS2000 dataset results (a) and PETS2001 dataset results (b).
+Both plots show a sample from the posterior distribution of the state, where the vertical axis denotes time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.}
+  \label{fig:benchmark_results_1}
+\end{figure}
+
+\begin{figure}[h]
+  \centering             
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f1}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f4}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f9}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f11}}
+  \caption{Four frames from the PETS2000 sequence with one posterior sample mean and covariance matrix representation shown for each frame (and for the previous 20 frames).}
+  \label{fig:pets2001_imgs}
 \end{figure}
@@
   \centering               
-  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2000_front.pdf}}             
-  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2001_back.pdf}}
-  \caption{PETS2000 dataset results (a) and PETS2001 dataset results (b).
-Both plots show a sample from the posterior distribution of the state, where the vertical axis denots time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.}
-  \label{fig:benchmark_results}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f1.pdf}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f3.pdf}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f5.pdf}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f8.pdf}}
+  \caption{Four frames from the PETS2001 sequence with one posterior sample mean and covariance matrix representation shown for each frame (and for the previous 20 frames).}
+  \label{fig:pets2001_overlay}
 \end{figure}
+
+To calculate performance metrics (both SFDA and ATA), one must specify a confidence value that allows the oval representing the region occupied by an object to be computed from the inferred covariance matrix of each cluster (as discussed in Section~\ref{sec:inference_to_results}).
+The performance metrics were found for a range of confidence intervals, and these curves, for the PETS2000 and PETS2001 video are shown in Figure~\ref{fig:pm_conf}.
+
+\begin{figure}[h]
+  \centering             
+  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2000_pm}}
+  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2001_pm}}
+  \caption{The SFDA (blue solid line) and ATA (red dashed line are) vs  confidence values from which an object's oval region is computed for (a) PETS2000 and (b) PETS2001 video datasets.}
+  \label{fig:pm_conf}
+\end{figure}
+
+\begin{figure}[h]
+	\centering
+    \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/synth2_alphaimg_hd.png}}
+    \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2001_alphaimg_hr.png}}
+	\subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2000_4objects.png}}
+    \caption{\label{fig:synth2_alphaimg} Sample from posterior of the state for synthetic experiment two (a), and benchmark datasets PETS2001 (b) and PETS2000 (c), where color represents assignment and the mean and standard deviation are shown.
+In each plot, the sample and data are overlaid on three images from the experiment.}
+\end{figure}
+
+% old figure ideas:
+
+% \begin{figure}[h]
+%         \center{\includegraphics[width=65mm]{../img/pets2001_alphaimg_hr.png}}
+%         \caption{\label{fig:pets2001_alphaimg} Extracted data and sample from posterior of the model overlaid on three frames from the PETS2001 video.}
+% \end{figure}
+
+% \begin{figure}[h]
+% 	\centering
+% 	\subfloat[]{\includegraphics[width=0.58\textwidth]{../img/pets2000_results_1}}
+% 	\subfloat[]{\includegraphics[width=0.42\textwidth]{../img/pets2000_4objects.png}}
+% 	\caption{Extracted data and a sample from posterior of the model (a) overlaid on a frame from the PETS2000 video (b).}
+% 	\label{fig:pets2000_results}
+% \end{figure}
 
@@
 
-\begin{figure}
-  \centering               
-  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_f1.pdf}}
-  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_f3.pdf}}
-  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_f5.pdf}}
-  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_f7.pdf}}
-  \caption{PETS2001 results.  \willie{add tails. different color ovals for different objects.}}
-  \label{fig:pets2001_overlay}
-\end{figure}
-
-To calculate performance metrics (both SFDA and ATA), one must specify a confidence value that allows the oval representing the region occupied by an object to be computed from the inferred covariance matrix of each cluster (as discussed in Section~\ref{sec:inference_to_results}).
-The performance metrics were found for a range of confidence intervals, and these curves, for the PETS2000 and PETS2001 video are shown in Figure~\ref{fig:pm_conf}.
-
-\begin{figure}[h]
-  \centering             
-  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2000_pm}}
-  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2001_pm}}
-  \caption{The SFDA (blue solid line) and ATA (red dashed line are) vs  confidence values from which an object's oval region is computed for (a) PETS2000 and (b) PETS2001 video datasets.}
-  \label{fig:pm_conf}
-\end{figure}
-
-\begin{figure}[h]
-  \centering             
-  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_1}}
-  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_2}}
-  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_3}}
-  \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2001_4}}
-  \caption{Four frames $f=25, 35, 45, 55$ from the PETS2001 sequence with one posterior sample mean and covarance matrix overlaid.
-\willie{will show path trail over prev 10 frames.} \willie{These are old images!...
-haven't replaced yet}}
-  \label{fig:pets2001_imgs}
-\end{figure}
-
 \subsubsection{PETS2009/2010}
@@
 The SMC inference algorithm yielded a sample from the posterior distribution, from which the object detection and tracking results were obtained (as described in Section~\ref{sec:inference_to_results}).
-Figure \willie{provide figure showing PETS2009 results} shows a sample from the posterior of the cluster parameters overlayed on the data, and video frames at three points in time, where the data is colored by assignment
-\willie{show SFDA and ATA vs cov mat conf for this sample that I show figures for.}
+In Figure \ref{fig:pets2009_results}, a sample from the posterior of the cluster parameters is overlayed on the extracted data over a sequence of frames, where the assignment of each data point is represented by color and marker type.
+
+\begin{figure}[h]
+  \centering
+  % \subfloat[]{\label{fig:pets2009_results} \includegraphics[width=0.7\textwidth]{../img/pets2009_1-50_front.pdf}}
+  \includegraphics[width=0.7\textwidth]{../img/pets2009_1-50_front.pdf}
+  \caption{Results on frames 1-50 of the PETS2009 dataset, showing a sample from the posterior distribution of the state, where the vertical axis denotes time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.}
+  \label{fig:pets2009_results}
+  % \label{fig:benchmark_results_2}
+\end{figure}
 

### 2011-12-30 ltx/paper.tex
re-did performance metrics for pets2000 and pets2001. re-wrote GPUDDPM part of paper.

@@
 \documentclass[smallcondensed, final]{svjour3}
-% \documentclass[smallcondensed, final, numbook]{svjour3}[12/23/2011]
-% \documentclass[twocolumn, final]{svjour3}
+% \documentclass[smallcondensed, final]{svjour3}[1/1/2012]
 
@@
 \titlerunning{Unsupervised Object Detection and Tracking with Dependent Dirichlet Processes}
+% \subtitle{A Bayesian Nonparametric Approach to Tracking Multiple Objects in Videos}
+
+\author{Willie Neiswanger$^{*}$ and Frank Wood$^{\dagger}$}
 \authorrunning{Willie Neiswanger and Frank Wood}
-% \subtitle{A Bayesian Nonparametric Approach to Tracking Multiple Objects in Videos}
-% \subtitle{Subtitle?}
-\author{Willie Neiswanger$^{*}$ and Frank Wood$^{\dagger}$}
-\institute{
-	$^{*}$Columbia University, Department of Applied Math and Applied Physics. 
+
+\institute{$^{*}$Columbia University, Department of Applied Math and Applied Physics. 
 Tel.: 503-464-6152.
@@
 \email{fwood@stat.columbia.edu}}
+
 \date{}
@@
 \begin{abstract}
-This paper proposes an unsupervised technique for the detection and tracking of arbitrary objects in videos.
-It is intended to reduce the need for detection or localization methods tailored to specific objects and serve as a general framework applicable to videos with varied object types, backgrounds, and film qualities.
+This paper proposes a technique for the unsupervised detection and tracking of arbitrary objects in videos.
+It is intended to reduce the need for detection or localization methods tailored to specific object types and serve as a general framework applicable to videos with varied objects, backgrounds, and film qualities.
 The technique uses a dependent Dirichlet process mixture (DDPM) known as the Generalized Polya Urn (GPUDDPM) to model image pixel data that can be extracted in a general manner from the regions in a video that represent objects.
 This paper describes a specific implementation of the model using spatial and color pixel data extracted via frame differencing and gives two algorithms for performing inference on the model to accomplish detection and tracking.
-This technique is demonstrated on multiple synthetic and benchmark video datasets that illustrate its ability to, without modification, detect and track variety of objects with diverse physical charactersitics over non-uniform backgrounds and through occlusion.
+This technique is demonstrated on multiple synthetic and benchmark video datasets that illustrate its ability to, without modification, detect and track objects with diverse physical charactersitics moving over non-uniform backgrounds and through occlusion.
 % \keywords{Object Detection \and Tracking \and Bayesian Nonparametrics \and Multiple Target Tracking}
@@
 \section{Introduction}
-% % \subsection{Topic Overview and Terminology}
 
@@
 The main challenges, however, associated with blob tracking involve the difficulty of segmenting neighboring or occluding objects in a given frame \cite{zhao2004tracking} and preventing multiple tracked blobs from collapsing onto and tracking a single object \cite{vermaak_2003}.
-These difficulties stem from the typical blob tracker's need to perform detection via segmentation on individual frames and lack of a principled way to perform tracking while resisting the collapse of independent tracks.
+These difficulties stem from the typical blob tracker's need to perform detection via segmentation on individual frames and lack of a principled way to perform tracking of multiple objects while resisting the collapse of independent tracks.
 
@@
 Specifically, this work describes how a Dirichlet process mixture model can accomplish these goals when it is applied to data that may be extracted in a general manner from videos.
-We will demonstrate how to extract data from objects in videos and how this generative model allows for automated detection and tracking.
+We will demonstrate how to perform this extraction and how this generative model allows for automated detection and tracking.
 The model is first presented in an abstract form, to allow it to be used with arbitrary extraction procedures and object appearance representations that may be desired in future studies.
 We then formulate a specific model for data gained during a basic extraction procedure and describe inference algorithms that allow object localization and tracking to be carried out.
-We demonstrate our implementation on a number of synthetic and benchmark datasets and describe and compute performance metrics to quantify results.
+We demonstrate our implementation on multiple synthetic and benchmark datasets and compute a few standard performance metrics for each to quantify results.
 
@@
 Heuristic-based methods perform extracting by looking for temporal changes in pixel value characteristics.
-These methods are often less accurate than model-based methods, but also far less computationally expensive.
+These methods are often less accurate than model-based methods, but also less computationally expensive.
 Frame differencing and background subtraction are two such methods, and involve comparing pixel values, or groups of pixel values, between two images, to detect a change surpassing some threshold.
 The goal of these techniques is to find locations in a scene that display motion, and to deem these moving areas the foreground of a scene.
-Often, background subtraction refers to comparing an image containing targets with an image of the background without targets or with some model of the background that is learned as the video progresses (in these cases, objects that do not move throughout a scene will be treated as the background), while frame or image differencing refers to comparing two subsequent images in a video.
+Often, background subtraction refers to comparing an image containing targets with an image of the background without targets or with some model of the background that is learned as the video progresses (in these cases, objects that do not move throughout a scene will be treated as the background), while frame or image differencing refers to comparing pairs of consecutive images in a video.
 Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking schemes \citep{perez_2002}.
-In addition to the two techniques mentioned above, there exist other ways of extracting movement from an image, such as through techniques involving optical flow (a calculation based on the `generalized gradient model' of an image, which attempts to capture the speed and direction of movement over areas in the image) \citep{horn_1981, bobick_2001}.
+In addition, there exist other ways of extracting movement from an image, such as through techniques involving optical flow (a calculation based on the `generalized gradient model' of an image, which attempts to capture the speed and direction of movement over areas in the image) \citep{horn_1981, bobick_2001}.
 In \citep{black_2000}, a motion detection and object tracking method is developed based on the relative motion of moving objects' boundaries, which are again found through analysis of optical flow.
 
-Localization is closely related to the topics of object detection, position-finding, segmentation, and target recognition.
+Localization is related to the topics of object detection, position-finding, segmentation, and target recognition.
 These methods are mainly concerned with finding the positions and/or shapes of objects in a given video frame, or with discerning between objects similar in position or appearance.
 There has been a focus in this field on developing techniques for representing an object's appearance, both to allow for localization of objects on a per-frame basis and to allow for consistent localization of a given object over time.
-Research into the unsupervised detection of arbitrary objects (which we define as methods that accomplish both extraction and localization) has included work involving image segmentation, parsing, and pixel clustering \willie{must cite here}.
+Research into the unsupervised detection of arbitrary objects (which we define as methods that accomplish both extraction and localization) has included work involving image segmentation and pixel clustering \willie{must cite here}.
 These topics usually relate to the automated partitioning of images into meaningful regions.
@@
 Areas of interest in this field involve developing metrics for judging the similarity between two object representations (for example, potential representations of the same object at adjacent frames in a video), techniques that provide a way to conduct efficient searches over a frame for regions that are similar to a given object, and methods for predicting the future position or appearance of an object.
-
 Methods for tracking single, non-rigid, objects are numerous.
-Color, for example, has long been used for tracking.
-These techniques often operate by modeling an object with some color-based appearance model, and using this model to find the object in subsequent frames.
+Color, for example, has long been used for tracking; these techniques often operate by modeling an object with some color-based appearance model, and using this model to find the object in subsequent frames.
 One classic approach involves modeling the color of a target (in particular, a distribution over the hue-saturation space of a target region) with a Gaussian mixture model (GMM), and choosing subsequent target positions by searching surrounding areas for regions that yield similar GMMs.
-Furthermore, the GMM is often allowed to adapt over time, and slowly change to model changes in lighting or other smooth time-based variations in the target's color \citep{raja_1998, mckenna_1999, jepson_2003}.
+Furthermore, the GMM is often allowed to adapt over time, and slowly change to model changes in lighting or other time-based variations in the target's color \citep{raja_1998, mckenna_1999, jepson_2003}.
 Additionally, color has been successfully applied as the feature in a technique known a the mean-shift procedure (or the kernel-based method) for tracking \citep{comaniciu_2003, perez_2002, nummiaro_2003, lee_2011}.
@@
 Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors.
-It can be shown that, through this procedure, each data point becomes associated with a local point of high density (dependent upon the underlying weight distribution specified by the kernel) which naturally allows for clustering \citep{cheng_1995}.
+It can be shown that, through this procedure, each data point becomes associated with a local point of high density (dependent upon the underlying weight distribution specified by the kernel) which allows for clustering \citep{cheng_1995}.
 The mean shift algorithm has been implemented successfully to allow for a sort of kernel-based object tracking \citep{comaniciu_2003, comaniciu_1999, comaniciu_2000}.
@@
 % Extraction
-
 %
@@
 \label{sec:modeloverview}
-%
-This section provides details on the data to be modeled, the Dirichlet process mixture (DPM) model, and the dependent Dirichlet process mixture (DDPM) model.
+
+This section provides details on the extracted data, Dirichlet process mixture (DPM) models, dependent Dirichlet process mixture (DDPM) models, and the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM) model.
 We make explicit the way in which characteristics of video objects can be represented by elements of these models.
-We formulate the model in a sufficently abstract manner to allow for implementations tailored for specified objects in future studies.
-
-\subsection{Data}
+We formulate the model in a sufficently abstract manner to allow for implementations using specific object representations and extraction procedures in future studies.
+
+\subsection{Data Extraction}
 \label{sec:data}
-%
+
 We defined the task of extraction to be the unsupervised determination of which spatiotemporal regions of a video constitute objects, and outlined a number of extraction methods in Section~\ref{sec:relatedwork}.
@@
 An extraction dataset of this form is comprised of the three-dimensional points $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $x_{1}$ and $x_{2}$ are the two spatial dimensions of a pixel contained in an extracted region in a given frame, and $t$ is the discrete time index of the frame.
-These three features are always available once regions containing objects are discerned, and are useful features for modeling (as position information is key for reasonable tracking and time information is necessary for maintaining the temporal dependencies in the data).
-
-Examples of additional features that could potentially be extracted from object-regions in video frames include color information, pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, texture representations, transform calculations (such as results of hough or fourier transforms), and quantiative region descriptions derived from image segmentation procedures; the main idea is to choose features which are able to be easily extracted or computed, capture variability in the appearance of objects, are applicable to a wide variety of object types, and can be represented by distributions that allow for tractable inference (discussed further in Section~\ref{sec:inference}).
+These three features are always available once regions containing objects are discerned, and are useful features for modeling (as position and time information is key for reasonable tracking).
+
+Examples of additional features that could be extracted from object-regions in video frames include color information, pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, texture representations, transform calculations (such as results of hough or fourier transforms), and quantiative region descriptions derived from image segmentation procedures; the main idea is to choose features which are able to be easily extracted or computed, capture variability in the appearance of objects, are applicable to a wide variety of object types, and can be represented by distributions that allow for tractable inference (discussed further in Section~\ref{sec:inference}).
 
 We will use the phrase ``baseline'' three-dimensional extraction data to refer to the features $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$ of pixels present in regions containing objects.
-These data are distributed roughly as worm-like shapes that follow the paths of individual objects as they move, when they are plotted in three dimensions (where we let the frame number $t$, our discrete representation of time, ascend the vertical axis and the spatial positions occupy the horizontal axes).
-An example of data with such a distribution can been seen in Figure~\ref{fig:baselinedata}.
-The data consists of the three-dimensional baseline features and was gained by extraction peformed on a 50 frame sequence showing five ants moving across the video scene.
-
-% frank says the following worm figures are a no go
+These data are distributed roughly as worm-like shapes that follow the paths of individual objects as they move, when the data are plotted in three dimensions (where we let the frame number $t$, our discrete representation of time, ascend the vertical axis and the spatial positions occupy the horizontal axes).
+%An example of data with such a distribution can been seen in Figure~\ref{fig:baselinedata}...The data consists of the three-dimensional baseline features and was gained by extraction peformed on a 50 frame sequence showing five ants moving across the video scene.
+
+% frank says the following worm figures are a no go.
 % \begin{figure}[h]
@@
 \subsection{Dirichlet Process Mixtures}
+\label{sec:dpm}
 
@@
 They are, in particular, useful for estimating the number of latent classes (clusters) in mixture models.
-In this work, estimating the number of clusters in the data generated via extraction is shown in the following sections to be equivalent to estimating the number of distinct objects in a video.
-Consequentially, a model that is able to perform robust nonparametric estimation of the clusters in an extraction dataset may be applied to perform unsupervised object detection and tracking.
-
-The following sections are intended to provide a solid introduction to DPMs without a major foray into the theoretical aspects of the model that require a more rigorus technical setup.
-We give background on finite mixture models, Bayesian finite mixture models, and Dirichlet processes, which allows us to subsequently describe DPMs in context.
+In this work, estimating the number of clusters in the data generated via extraction is equivalent to estimating the number of distinct objects in a video.
+Consequentially, a model that is able to perform robust nonparametric estimation of the clusters in data gained during extraction may be applied to perform unsupervised object detection and tracking.
+
+The following sections are intended to provide an introduction to DPMs without a major foray into the theoretical aspects of the model that require a more rigorus technical setup.
+We provide background on finite mixture models, Bayesian finite mixture models, and Dirichlet processes, which allows us to subsequently describe DPMs in context.
 
 \subsubsection{Finite Mixture Model}
-
-A finite mixture model can be thought of as a probability distribution for an observation $x_{i}$ formulated as a linear combination of K mixture components (also called `clusters'), where the coefficients of the linear combination sum to one; each mixture component, in turn, is a probability distribution for $x_{i}$, given some parametric form.
+\label{sec:finite_mixture}
+
+A finite mixture model can be thought of as a probability distribution for an observation $x_{i}$ formulated as a linear combination of K mixture components (which we also refer to as `clusters'), where each mixture component is a probability distribution for $x_{i}$ with some parametric form, and the coefficients of the linear combination sum to one.
 The finite mixture model can be written as
 \begin{equation}
-P(x) = \sum_{k=1}^{K} P(c_{i} = k)P(x|\theta_{k})
-\end{equation}
-where $c_{i} \in \{ 1, \ldots, K \}$ denotes the assignment of $x_{i}$ to a given mixture (which allows the coefficients of the linear combination to be written as $P(c_{i} = k)$), and $\theta_{k}$ denotes the parametric form of the $k^{\text{th}}$ mixture component.
-We also define $p_{k} = P(c_{i} = k)$ for $k \in \{ 1, \ldots, K \} $.
+P(x_{i}) = \sum_{k=1}^{K} P(c_{i} = k)P(x_{i}|\theta_{k})
+\end{equation}
+$\forall i \in \{ 1, \ldots, N \}$, where $c_{i} \in \{ 1, \ldots, K \}$ denotes the assignment of $x_{i}$ to a given mixture and $\theta_{k}$ denotes the parameters of the $k^{\text{th}}$ mixture component.
+Note that by choosing $P(c_{i} = k)$ as coefficients of the linear combination, it is ensured that these coefficients sum to one.
+We also define $p_{k} := P(c_{i} = k)$ for $k \in \{ 1, \ldots, K \} $.
 We can therefore write this model generatively as
@@
 \end{align}
-where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $c_{i}$ are the mixture component assignments associated with each observation, and the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e.
-the distribution to be mixed, $F(\theta_{c_{i}})$).
+$\forall i \in \{ 1, \ldots, N \}$, where the $x_{i}$ are observations, the $c_{i}$ are the mixture component assignments associated with each observation, the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e.
+the distribution to be mixed, $F(\theta_{c_{i}})$), and the ``Discrete'' distribution refers to a multinomial distribution whose parameters are a 1-of-K vector (i.e.
+a vector of counts that sums to one).
 
 \subsubsection{Bayesian (Finite) Mixture Model}
-
-The finite mixture model of the previous section can be extended to a Bayesian mixture model by viewing the distribution parameters that were previously point values, $\theta_{c_{i}}$ (the mixture component parameters) and $p_{1}, \ldots, p_{K}$ (the mixture component assignment weights), as random variables and providing each with a prior distribution.
-In this case, the prior distribution $\mathbb{G}_{0}$ is placed on the mixture component parameters, and the prior distribution $\text{Dir}(\alpha/K, \ldots, \alpha/K)$ is placed on the K mixture component assignment weights.
-The resulting Bayesian mixture model can be formulated in a generative representation as
+\label{sec:bayesian_mixture}
+
+The finite mixture model of the previous section can be extended to a Bayesian mixture model by viewing parameters that were previously point values, $\theta_{c_{i}}$ (the mixture component parameters) and $p_{1}, \ldots, p_{K}$ (the mixture component assignment weights), as random variables and providing each with a prior distribution.
+In this case, the prior distribution $\mathbb{G}_{0}$ is placed on the mixture component parameters, and the prior distribution $\text{Dir}(\alpha/K, \ldots, \alpha/K)$ is placed on the mixture component assignment weights.
+The resulting Bayesian mixture model can be formulated generatively as
 \begin{align}
@@
 	p_{1}, \ldots, p_{K}  &\sim  \text{Dir}(\alpha/K, \ldots, \alpha/K)\\
+		\theta_{k}  &\sim  \mathbb{G}_{0}, \hspace{2mm} \forall k \in \{1, \ldots, K\} \\
 	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
-	\theta_{c_{i}}  &\sim  \mathbb{G}_{0} \\
 	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}})
@@
 \end{align}
-where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $c_{i}$ are the mixture component assignments associated with each observation, the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e.
-the distribution to be mixed, $F(\theta_{c_{i}})$), the $\theta_{c_{i}}$ are drawn from the prior distribution $\mathbb{G}_{0}$, and $p_{1}, \ldots, p_{K}$ are drawn from a Dirichlet prior parameterized by $\alpha/K, \ldots, \alpha/K$.
+$\forall i \in \{ 1, \ldots, N \}$, where the $x_{i}$ are observations, the $c_{i}$ are the mixture component assignments associated with each observation, the $\theta_{k}$ are parameters defining the $k^{\text{th}}$ mixture component (i.e.
+the distribution to be mixed, $F(\theta_{k})$), the $\theta_{k}$ are drawn from a prior distribution $\mathbb{G}_{0}$, and $p_{1}, \ldots, p_{K}$ are drawn from a Dirichlet prior parameterized by $\alpha/K, \ldots, \alpha/K$.
 
 \subsubsection{Dirichlet Process}
+\label{sec:dirichlet_process}
 
@@
 \subsubsection{Dirichlet Process (Infinite) Mixture Model}
+\label{sec:dirichlet_process_mixture}
 
 A Dirichlet process mixture model, also refered to as an infinite mixture model, is an extension of the Bayesian mixture model described previously. 
-When using a DP as a prior in a Bayesian mixture model, $\Theta$ can be viewed as the set of parameters of the component mixture distributions.
-	A DPM is formulated by allowing the prior distribution over component weights in a standard mixture model to be distributed according to a DP \willie{need to make sure am saying this right}; this allows for modeling data where the true number of latent mixture components is unknown and arbitrarily large by assuming an infinite number of components, of which only a finite amount are expressed by the data.
+When using a DP as a prior in a Bayesian mixture model, $\Theta$ represents the set of parameters of the component mixture distributions.
+A DPM may be viewed as allowing the prior distribution over the mixture component parameters in a standard mixture model to be distributed according to a DP; this allows for modeling data where the true number of latent mixture components is unknown and arbitrarily large by letting the number of components remain unbounded (note that only a finite number of these components are assigned to the data).
 In particular, the DPM can be defined generatively as
@@
 \end{align}
-where the $x_{i}$ (for $i \in \{ 1, \ldots, N \}$) are observations, the $\phi_{i}$ are parameters defining the mixture component from which the $i_{th}$ observation is drawn (i.e.
-the distribution to be mixed, $F(\phi_{i})$), and the $\phi_{i}$ are drawn from $\mathbb{G}$, which is in turn drawn from a DP with base distribution $\mathbb{G}_{0}$ and parameter $\alpha$.
+$\forall i \in \{ 1, \ldots, N \}$, where the $x_{i}$ are observations, the $\phi_{i}$ are parameters defining the mixture component from which the $i_{th}$ observation is drawn (i.e.
+the distribution to be mixed, $F(\phi_{i})$), and the $\phi_{i}$ are drawn from a prior distribution $\mathbb{G}$, which is in turn drawn from a DP with base distribution $\mathbb{G}_{0}$ and parameter $\alpha$.
 See \cite{gasthaus_2008} and \cite{gasthaus_thesis} for more details on this formulation.
 Note the difference between the indexing of the clusters in this model and the indexing in the previous two models.
-This formulation can be shown to be equivalent to the Bayesian mixture model defined in \eqref{bayesian_mixture_model}, when K is taken to be unbounded.
-This is the origin for references to this model as the infinite mixture model.
-
-An alternative formultion of the DPM is known as the Chinese resturant process (CRP) sampling representation.
-If we let $K$ be the number of distinct mixture components in the above model, we can write the distinct mixture components as $\theta_{1}, \ldots, \theta_{K}$.
+This formulation can be shown to be equivalent to the Bayesian mixture model defined in \eqref{bayesian_mixture_model}, when K is taken to be unbounded; as a result, this model is often called an infinite mixture model.
+
+% An alternative formultion of the DPM is known as the Chinese resturant process (CRP) sampling representation. 
+If we let $K$ be the number of distinct mixture components assigned to observations using the above model, we can write the mixture components as $\theta_{1}, \ldots, \theta_{K}$.
 Let $c_{1}, \ldots, c_{N}$ (where $c_{i} \in \{1, \ldots, K \}$) be class assignment variables that indicate the cluster to which observation $x_{i}$ is assigned.
-This allows the DPM to be represented by the CRP, a discrete-time stochastic process that defines a partition of the set $\{ x_{1}, \ldots, x_{N} \}$ (via the elements' assignments $c_{1}, \ldots, c_{N}$).
-The CRP allows samples to be drawn from the conditional distribution of the indictor variables $c_{i}$, and can be expressed as
+We can now formulate the Chinese restaurant process, a discrete-time stochastic process that defines a partition of the set $\{ x_{1}, \ldots, x_{N} \}$ (via the elements' assignments $c_{1}, \ldots, c_{N}$).
+The Chinese restaurant process allows samples to be drawn from the conditional distribution over the assignment variables $c_{i}$, and can be expressed as
 \begin{align}
@@
 where $m_{k}$ is the cardinality of the set $\{ c_{j} | (c_{j}=c_{i}=k)  \wedge  (j < i) \}$ and $\alpha$ is the parameter of the DP prior on $\mathbb{G}$.
+A modified Chinese restaurant process will be used to draw samples from the conditional distribution over the assignment variables when performing inference on the GPUDDPM (Section~\ref{sec:gpuddpm}).
 
 \subsection{Dependent Dirichlet Process Mixtures}
-
-The goal of dependent Dirichlet process mixtures (DDPMs) is to allow modeling of data that is not independent and identically distributed (i.i.d) but instead has some underlying dependencies.
-For example, data generated during extraction procedures from videos have some associated temporal structure, since tracked objects display time dependent characteristics (in the features that denote position, as well as those which denote other appearance characteristics).
-
-To account for the dependent behavior of data, research has been conducted on developing models involving a sequence of DPMs, where components of the mixtures are dependent upon (or may be considered `tied to') corresponding components at neighboring positions in the sequence.
-For example, if the data shows temporal dependence, the goal might be to create a sequence of DPMs, one for each time-step, where the components of the mixture at each step are dependent upon corresponding components in the both the following and previous time steps \willie{cite some of these}.
+\label{sec:ddpm}
+
+The goal of dependent Dirichlet process mixtures is to allow modeling of data that is not independent and identically distributed (i.i.d) but instead has some underlying dependencies.
+For example, data generated during extraction procedures from videos have some associated temporal dependencies, since tracked objects display time dependent characteristics (i.e.
+there are similarities among features, such as the spatial positions or appearances of objects, for data at nearby time steps).
+
+To account for the dependent behavior of data, there has been research on developing models involving a sequence of DPMs, where components of the mixtures are dependent upon (or are sometimes said to be ``tied to'') corresponding components at neighboring positions in the sequence.
+For example, if the data shows temporal dependence, the goal might be to create a sequence of DPMs, one for each time-step, where the components of the mixture at each step are dependent upon corresponding components in the both the following and previous time steps.
+\willie{Discuss and cite DDPMs}.
 
@@
 \subsubsection{Generalized Polya Urn Dependent Dirichlet Process Mixture}
-\label{sec:gpudpm}
+\label{sec:gpuddpm}
 
@@
 
-A transition kernel $P(\phi_{k}^{t} | \phi_{k}^{t-1})$ specifies how mixture component parameters in time step $t$ are dependent upon associated mixture component parameters in time step $t-1$.
-One caveat is that each mixture component must be drawn independently from $\mathbb{G}_{0}$ (the base distribution of the DP, which acts as a prior distribution for the cluster parameters) which we achieve by making $\mathbb{G}_{0}$ the invariant distribution of $P(\phi_{k}^{t} | \phi_{k}^{t-1})$ (note that the transition kernel is a markov chain).
+A transition kernel $P(\theta_{k,t} | \theta_{k,t-1})$ specifies how mixture component parameters in time step $t$ are dependent upon associated mixture component parameters in time step $t-1$.
+One caveat is that each mixture component must be drawn independently from $\mathbb{G}_{0}$ (the base distribution of the DP, which acts as a prior distribution for the cluster parameters) which we achieve by making $\mathbb{G}_{0}$ the invariant distribution of the transition kernel $P(\theta_{k,t} | \theta_{k,t-1})$ (which, one should note, is also a markov chain).
 
@@
 Specifically, each (remaining) assignment is removed from its cluster with probability $\rho$.
-We can consider the size $m_{k,t}$ of cluster $k$ at time $t$ to be dependent upon the cluster size at time $t-1$ via a deletion step.
+The size $m_{k,t}$ of cluster $k$ at time $t$ can be computed by removing elements from cluster $k$ at the previous time and thus diminishing the cluster's size at time $t-1$.
 It can be shown that performing deletion in this way is equivalent to, for each cluster $k$, drawing $r \sim \text{Binomial}(m_{k,t-1}, \rho)$, and reducing the cluster size $m_{k,t}$ by $r$.
-This can be written as
-\begin{equation}
+% This can be written as
+% \begin{equation}
+% \label{del_step}
+% m_{k,t} | m_{k,t-1}   \sim   m_{k,t-1}-\text{Binomial}(m_{k,t-1}, \rho)
+% \end{equation}
+For each time step, after all observations are given an assignment, this deletion procedure is carried out.
+Hence, the size $m_{k,t}$ of a given cluster $k$ at time $t$, is dependent upon the cluster size at the previous time step and on the assignments $c_{i}$ for all observations $i \in \{ 1, \ldots, N_{t} \}$ at time $t$.
+We refer to the conditional distribution over $m_{k,t} | m_{k,t-1}, c_{i}, \rho$ as DEL, which we define to be
+\begin{equation}
+\begin{split}
 \label{del_step}
-m_{k,t} | m_{k,t-1}   \sim   m_{k,t-1}-\text{Binomial}(m_{k,t-1}, \rho)
-\end{equation}
-with deletion parameter $\rho$.
-This process adheres to what \cite{caron_2007} calls a `uniform deletion strategy' over all the observations' assignments (since assignments to each cluster have equal probability of being deleted), though a more complex deletion strategy, dependent upon cluster size, can also be implemented \cite{caron_2007}.
-For a given time $t$ and cluster $k$, the size $m_{k,t}$ is dependent upon the cluster size at the previous time step (via \eqref{del_step}) and on the assignments $c_{i}$ for observations at time $t$.
-We refer to this distribution over $m_{k,t} | m_{k,t-1}, c_{i}, \rho$ as DEL.
-Using the cluster size and observation deletion terms introduced above, we can define the GPUDDPM generatively as, for each time step $t \in \{1, \ldots, T\}$,  % and observation $i \in \{ 1, \ldots, N_{t} \}$ at $t$,
+\text{DEL} & (\cdot | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) :=
+\left( m_{k,t-1} - \text{Binomial}( \cdot | {m_{k,t-1}, \hspace{1mm} \rho}) \right) + \sum_{i=1}^{N_{t}} \mathbb{I}(c_{i,t} = k)
+\end{split}
+\end{equation}
+$\forall k \in \{1, \ldots, K_{t} \}$, where $\rho$ is a deletion parameter, and $\mathbb{I}(c_{i,t} = k)$ is an indicator function whose value is 1 if $c_{i,t} = k$ and 0 otherwise.
+This process adheres to what \cite{caron_2007} calls a `uniform deletion strategy' over all the observations' assignments (since assignments to each cluster have equal probability of being deleted), though a more complex deletion strategy, dependent upon cluster size, can also be implemented and is described in \cite{caron_2007}.
+The Chinese restaurant process is used to sample assignment variables during inference procedures (described in Section~\ref{sec:inference}).
+We define a distribution used in these procedures, CRP, to be
+\begin{equation}
+\begin{split}
+\text{CRP} & (\cdot | m_{1:K_{t},t}, \hspace{1mm} \alpha) :=
+\text{Discrete}(\cdot | P(c_{i,t}=1 | m_{1:K_{t},t}), \ldots, P(c_{i,t}=K_{t+1}  | m_{1:K_{t},t}))
+\end{split}
+\end{equation}
+where
+\begin{equation}
+P(c_{i,t}=k  | m_{1:K_{t},t}) = 
+\begin{cases}
+\frac{m_{k,t}}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{2mm} \text{if} \hspace{2mm} k \in \{ 1, \ldots, K_{t} \} \\
+\frac{\alpha}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{2mm} \text{if} \hspace{2mm} k = K_{t} + 1
+\end{cases}
+\end{equation}
+$\forall i \in \{1, \ldots, N \}$, where there exists $K_{t}$ clusters at time $t$, and we give a newly created cluster the index $K_{t+1}$.
+Note the similarity between the above definition and the Chinese resturant process of the DPM given by \eqref{crp_rep}.
+The deletion and Chinese resturant procedures together comprise what \cite{caron_2007} refers to as the ``Generalized Polya Urn''.
+Using the DEL and CRP distributions given above, we can define the GPUDDPM generatively as, for each time step $t \in \{1, \ldots, T\}$, and each cluster $k \in \{ 1, \ldots, K_{t} \}$ at time $t$,  % and observation $i \in \{ 1, \ldots, N_{t} \}$ at $t$,
 \begin{align}
 \begin{split}
-\bold{m}_{t} | \bold{m}_{t-1}, \rho  &\sim \text{DEL}(\bold{m}_{t-1}, \rho) \\
-c_{i} | \bold{m}_{t}, \alpha  &\sim  \text{CRP}(\bold{m}_{t}, \alpha) \\
-\theta_{c_{i}, t} | \theta_{c_{i}, t-1}   &\sim
+m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho  &\sim \text{DEL}(m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) \\
+\theta_{k, t} | \theta_{k, t-1}   &\sim
 \begin{cases}
-	P(\theta_{c_{i}, t} | \theta_{c_{i}, t-1}) \\
-	\mathbb{G}_{0}
+	P(\theta_{k, t} | \theta_{k, t-1}) \hspace{2mm} \text{if} \hspace{2mm} k \in \{ 1, \ldots, K_{t} \} \\
+	\mathbb{G}_{0}   \hspace{2mm} \text{if} \hspace{2mm} k = K_{t+1}
 \end{cases} \\
-\bold{x}_{i} | c_{i}, \theta_{c_{i}, t} &\sim F(\theta_{c_{i}, t})
+c_{i,t} | m_{1:K_{t},t}, \hspace{1mm} \alpha  &\sim  \text{CRP} (m_{1:K_{t},t}, \hspace{1mm} \alpha) \\
+\bold{x}_{i} | c_{i}, \theta_{1:K_{t}, t} &\sim F(\theta_{c_{i}, t})
 \end{split}
 \end{align}
-where $\text{CRP}(\cdot, \alpha)$ is given in \eqref{crp_rep}.
+The graphical model corresponding with this formulation is shown in Figure \ref{fig:gpuddpm_gm_1}.
+\begin{figure}[h]
+        \center{\includegraphics[width=90mm]{../img/gpuddp_gm_1.pdf}}
+        \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Polya Urn Dependent Dirichlet Process Mixture.
+Note that all observations at time $t$, $\bold{x}_{1:N,t}$, and their associated assignments $c_{1:N,t}$ are denoted respectively as $x_{t}$ and $c_{t}$ (this also holds for time step $t-1$).}
+\end{figure}
+% \begin{align}
+% \begin{split}
+% \bold{m}_{t} | \bold{m}_{t-1}, \rho  &\sim \text{DEL}(\bold{m}_{t-1}, \rho) \\
+% c_{i} | \bold{m}_{t}, \alpha  &\sim  \text{CRP}(\bold{m}_{t}, \alpha) \\
+% \theta_{c_{i}, t} | \theta_{c_{i}, t-1}   &\sim
+% \begin{cases}
+% 	P(\theta_{c_{i}, t} | \theta_{c_{i}, t-1}) \\
+% 	\mathbb{G}_{0}
+% \end{cases} \\
+% \bold{x}_{i} | c_{i}, \theta_{c_{i}, t} &\sim F(\theta_{c_{i}, t})
+% \end{split}
+% \end{align}
+% where $\text{CRP}(\cdot, \alpha)$ is given in \eqref{crp_rep}.
 The graphical model corresponding with this formulation is shown in Figure \ref{fig:gpuddpm_gm_1}.
@@
 If we allow for multiple observations $i \in \{ 1, \ldots, N_{t} \}$ at each time $t$, this formulation becomes,
-\begin{align}
-\begin{split}
-\bold{m}_{t}^{1} | \bold{m}_{t-1}^{N_{t-1}}, \rho  &\sim \text{DEL}(\bold{m}_{t-1}^{N_{t-1}}, \rho) \\
-c_{i} | \bold{m}_{t}^{i-1}, \alpha  &\sim  \text{CRP}(\bold{m}_{t}^{i-1}, \alpha) \\
-\theta_{c_{i}, t} | \theta_{c_{i}, t-1}   &\sim
-\begin{cases}
-	P(\theta_{c_{i}, t} | \theta_{c_{i}, t-1}) \\
-	\mathbb{G}_{0}
-\end{cases} \\
-\bold{x}_{i} | c_{i}, \theta_{c_{i}, t} &\sim F(\theta_{c_{i}, t})
-\end{split}
-\end{align}
-where $\bold{m}_{t}^{i}$ represents the $i^{\text{th}}$ observation at time $t$.
-
-\begin{figure}[h]
-        \center{\includegraphics[width=90mm]{../img/gpuddp_gm_1.pdf}}
-        \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Poly Urn Dependent Dirichlet Process Mixture}
-\end{figure}
+% \begin{align}
+% \begin{split}
+% \bold{m}_{t}^{1} | \bold{m}_{t-1}^{N_{t-1}}, \rho  &\sim \text{DEL}(\bold{m}_{t-1}^{N_{t-1}}, \rho) \\
+% c_{i} | \bold{m}_{t}^{i-1}, \alpha  &\sim  \text{CRP}(\bold{m}_{t}^{i-1}, \alpha) \\
+% \theta_{c_{i}, t} | \theta_{c_{i}, t-1}   &\sim
+% \begin{cases}
+% 	P(\theta_{c_{i}, t} | \theta_{c_{i}, t-1}) \\
+% 	\mathbb{G}_{0}
+% \end{cases} \\
+% \bold{x}_{i} | c_{i}, \theta_{c_{i}, t} &\sim F(\theta_{c_{i}, t})
+% \end{split}
+% \end{align}
+% where $\bold{m}_{t}^{i}$ represents the $i^{\text{th}}$ observation at time $t$.
 
@@
 
-\subsection{Extraction}
+\subsection{Data Extraction via Frame Differencing}
 \label{sec:modelspec_extraction}
@@
 
-As described in Section~\ref{sec:gpudpm}, $\mathbb{G}_{0}$ must be the invariant distribution of $P(\theta_{t} | \theta_{t-1})$ in order for the the cluster parameters to remain marginally distributed according to the base distribution.
+As described in Section~\ref{sec:gpuddpm}, $\mathbb{G}_{0}$ must be the invariant distribution of $P(\theta_{t} | \theta_{t-1})$ in order for the the cluster parameters to remain marginally distributed according to the base distribution.
 In other words, the transition kernel must satisfy
@@
 
-The generative model associated with the GPUDDPM is illustrated in Section~\ref{sec:gpudpm}.
+The generative model associated with the GPUDDPM is illustrated in Section~\ref{sec:gpuddpm}.
 This representation includes cluster size $m$ (the number of observations assigned to a given cluster), which is standard in DPM constructions.
@@
 This formulation of the GPUDDPM is also used in \cite{gasthaus_thesis} and \cite{caron_2007}.
-The associated graphical model for this formulation is given in \ref{fig:gpuddpm_gm_2}.
+The associated graphical model for this formulation is given in Figure~\ref{fig:gpuddpm_gm_2}.
 \begin{figure}[h]
         \center{\includegraphics[width=100mm]{../img/gpuddp_gm_3.pdf}}
-        \caption{\label{fig:gpuddpm_gm_2} Graphical model of the deletion-variable representation of the GPUDDPM}
+        \caption{Graphical model of the deletion variable and auxiliary variable transition kernel representation of the GPUDDPM}
+        \label{fig:gpuddpm_gm_2}
 \end{figure}
@@
   \centering             
-  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2000_pm}}
-  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2001_pm}}
+  \subfloat[]{\includegraphics[width=0.49\textwidth]{../img/pets2000_pm.pdf}} \hspace{1mm}
+  \subfloat[]{\includegraphics[width=0.49\textwidth]{../img/pets2001_pm.pdf}}
   \caption{The SFDA (blue solid line) and ATA (red dashed line are) vs  confidence values from which an object's oval region is computed for (a) PETS2000 and (b) PETS2001 video datasets.}
@@
   \includegraphics[width=0.7\textwidth]{../img/pets2009_1-50_front.pdf}
-  \caption{Results on frames 1-50 of the PETS2009 dataset, showing a sample from the posterior distribution of the state, where the vertical axis denotes time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.}
+  \caption{Results for the PETS2009 dataset, showing a sample from the posterior distribution of the state for frames 1-50, where the vertical axis denotes time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.}
   \label{fig:pets2009_results}

### 2012-01-04 ltx/paper.tex
more work on SMC inference algorithm

@@
-\documentclass[smallcondensed, final]{svjour3}
+\documentclass[smallcondensed, draft]{svjour3}
 % \documentclass[smallcondensed, final]{svjour3}[1/1/2012]
@@
 \begin{equation}
+\label{CRP}
 \begin{split}
@@
 c_{i,t} | m_{1:K_{t},t}, \hspace{1mm} \alpha  &\sim  \text{CRP} (m_{1:K_{t},t}, \hspace{1mm} \alpha) \\
-\bold{x}_{i} | c_{i}, \theta_{1:K_{t}, t} &\sim F(\theta_{c_{i}, t})
+\bold{x}_{i,t} | c_{i,t}, \theta_{1:K_{t}, t} &\sim F(\theta_{c_{i,t}, t})
 \end{split}
@@
         \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Polya Urn Dependent Dirichlet Process Mixture.
-Note that all observations at time $t$, $\bold{x}_{1:N,t}$, and their associated assignments $c_{1:N,t}$ are denoted respectively as $x_{t}$ and $c_{t}$ (this also holds for time step $t-1$).}
+Note that all observations at time $t$, $\bold{x}_{1:N,t}$, and their associated assignments $c_{1:N,t}$ are denoted respectively as $\bold{x}_{t}$ and $c_{t}$ (this also holds for time step $t-1$).}
 \end{figure}
@@
 \begin{equation}
+\label{likelihood}
 P(\bold{x}|\theta) = \mathcal{N}(\bold{x}^{s} | \boldsymbol{\mu}, \Sigma)  \mathcal{M}n(\bold{x}^{c} | \bold{p})
@@
 Specifically, in the experiments carried out in this paper, a normal-inverse-Wishart prior is placed on the multivariate normal parameters $\{ \boldsymbol{\mu}, \Sigma \}$ , and a Dirichlet prior is placed on the multinomial parameter $ \{  \bold{p}  \} $ (where the normal-inverse-Wishart is a conjugate prior for the multivariate normal component of the likelihood and the Dirichlet is a conjugate prior for the multinomial component).
-The prior can thus be written
-\begin{equation} \label{basedistro}
+The prior can therefore be written
+\begin{equation}
+\label{basedistro}
 \mathbb{G}_{0}(\theta) = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}, \Sigma | \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0})  \mathcal{D}ir( \bold{p} | \bold{q}_{0})
@@
 \end{equation}
-for a given cluster $\theta$.
+for a given cluster with parameters $\theta$.
 One way to achieve this is through the use of auxiliary variables.
-Auxiliary variables are a set of $M$ variables $\bold{z}_{t} = (z_{t,1}, \ldots, z_{t,M})$ associated with a each cluster $\theta$ at time $t$ that satisfy
+Auxiliary variables are a set of $M$ variables $\bold{z}_{t} = (z_{t,1}, \ldots, z_{t,M})$ associated with a each cluster at time $t$ that satisfy
 \begin{eqnarray}
@@
 For each cluster, we introduce $M$ auxiliary variables $z_{t, 1}, \ldots, z_{t, M}$ at time $t$ that are each drawn from the product of a multivariate normal and multinomial when conditioned on the associated cluster parameters $\theta_{t} = \{ \boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t} \}$:
-\begin{equation} \label{transker1}
+\begin{equation}
+\label{transker1}
 z_{t, m} | \boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t}  \sim  \mathcal{N}(\boldsymbol{\mu}_{t}, \Sigma_{t}) \mathcal{M}n(\bold{p}_{t})   \hspace{15pt}   
@@
 To satisfy the above criteria for auxiliary variables, at each time $t$ we specify the dependencies of a given cluster on its associated set of auxiliary variables by
-\begin{equation} \label{transker2}
+\begin{equation}
+\label{transker2}
 \boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t} | \bold{z}_t  \sim  \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}_{M}, \kappa_{M}, \nu_{M}, \Lambda_{M})  \mathcal{D}ir(\bold{q}_{M})
@@
 where $\boldsymbol{\mu}_{M}, \kappa_{M}, \nu_{M}, \Lambda_{M},$ and $\bold{q}_{M}$ are parameters given by
-\willie{these are given in next section, need to consider if I actually want to repeat them here}
+\begin{eqnarray}
+\kappa_{M} &=& \kappa_{0} + M \\
+\nu_{M} &=& \nu_{0} + M \\
+\boldsymbol{\mu}_{M} &=& \frac{\kappa_{0}}{\kappa_{0}+M} \boldsymbol{\mu}_{0}  +  \frac{M}{\kappa_{0}+M} \overline{\bold{z}_t}^{s}\\
+\Lambda_{M} &=& \Lambda_{0} + S_{\bold{z}_{t}^{s}}\\
+\bold{q}_{M} &=& \bold{q}_{0} + \sum_{m=1}^{M} z_{t,m}^{c}
+\end{eqnarray}
+where $M$ is the number of auxiliary variables, $\{ \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0} \}$ are the $\mathcal{N}i\mathcal{W}$ prior parameters, $\bold{q}_{0}$ is the $\mathcal{D}ir$ prior parameter, $\bold{z}^{s}$ and $\bold{z}^{c}$ respectively denote the spatial and color features of an auxiliary variable $\bold{z}$, and $\overline{\bold{z}}$ and $S_{\bold{z}}$ respectively denote the sample mean and sample covariance for a set $\bold{z} = \{ z_{1}, \ldots, z_{M} \}$ (of auxiliary variables, in this case), which we can define as
+\begin{eqnarray}
+\label{samplemean}
+\overline{\bold{z}}  &=&  \left( \sum_{m=1}^{M} z_{m} \right) / M\\
+\label{samplecov}
+S_{\bold{z}}  &=&  \sum_{m=1}^{M} (z_{m} - \overline{\bold{z}}) (z_{m} - \overline{\bold{z}})^{T}
+\end{eqnarray}
 
@@
 $\Lambda_{0}$  &  shape factor of the prior on the covariance\\
-$q_{0}$  &  scale factor of the prior on the multinomial counts
+$\bold{q}_{0}$  &  scale factor of the prior on the multinomial counts
 \end{tabular}
@@
 \end{equation}
-The MCMC sampler described in this section operates in this alternate representation, which we can formulate, for each time step $t \in \{1, \ldots, T\}$ and observation $i \in \{ 1, \ldots, N_{t} \}$ at time $t$, as
+The MCMC sampler described in this section operates in this alternate representation, which we can write, for each time step $t \in \{1, \ldots, T\}$ and clusters $k \in \{ 1, \ldots, K_{t} \} $ at time $t$, as
+% \begin{align}
+% \begin{split}
+% 	 \bold{x}_{i} | c_{i}, \theta_{c_{i},t} &\sim \mathcal{N}(\boldsymbol{\mu}_{c_{i},t}, \Sigma_{c_{i},t})  \mathcal{M}n(\bold{p}_{c_{i},t}) \\
+% 	c_{i} | c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha  &\sim  \text{CRP}(c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha) \\
+% 	d_{i} | \rho  &\sim \text{Geo}(\rho) + i + 1\\
+% 	\theta_{k, t} | \theta_{k, t-1}, \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0}  &\sim 
+% \begin{cases}
+% 	\text{TK} \hspace{12pt} \text{if } k \in \{ 1, \ldots, K_{t-1}  \} \\
+% 	\mathbb{G}_{0}(\boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0})  \hspace{15pt} \text{if } k = K_{t-1} + 1
+% \end{cases}
+% \end{split}
+% \end{align}
+% where TK stands for the transition kernel, given in \eqref{transker1} and \eqref{transker2}, and the base distribution $\mathbb{G}_{0}$ is given in \eqref{basedistro}.
 \begin{align}
 \begin{split}
-	 \bold{x}_{i} | c_{i}, \theta_{c_{i},t} &\sim \mathcal{N}(\boldsymbol{\mu}_{c_{i},t}, \Sigma_{c_{i},t})  \mathcal{M}n(\bold{p}_{c_{i},t}) \\
-	c_{i} | c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha  &\sim  \text{CRP}(c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha) \\
-	d_{i} | \rho  &\sim \text{Geo}(\rho) + i + 1\\
-	\theta_{k, t} | \theta_{k, t-1}, \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0}  &\sim 
+d_{i,t} | \rho  &\sim \text{Geo}(\rho) + t + 1 \\
+\theta_{k, t} | \theta_{k, t-1}  &\sim
 \begin{cases}
-	\text{TK} \hspace{12pt} \text{if } k \in \{ 1, \ldots, K_{t-1}  \} \\
-	\mathbb{G}_{0}(\boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0})  \hspace{15pt} \text{if } k = K_{t-1} + 1
-\end{cases}
+P(\theta_{k,t} | \theta_{k,t-1}) \hspace{2mm} \text{if} \hspace{2mm} k \in \{ 1, \ldots, K_{t} \} \\
+\mathbb{G}_{0} \hspace{2mm} \text{if} \hspace{2mm} k = K_{t} + 1
+\end{cases} \\
+c_{i,t} | \bold{c}_{1}, \ldots, \bold{c}_{t-1}, \bold{d}_{1}, \ldots, \bold{d}_{t-1}, \alpha  &\sim  \text{CRP}(\bold{c}_{1}, \ldots, \bold{c}_{t-1}, \bold{d}_{1}, \ldots, \bold{d}_{t-1}, \alpha) \\
+\bold{x}_{i,t} | c_{i,t}, \theta_{c_{i,t},t}  &\sim  F(\theta_{c_{i,t}, t})
 \end{split}
 \end{align}
-where TK stands for the transition kernel, given in \eqref{transker1} and \eqref{transker2}, and the base distribution $\mathbb{G}_{0}$ is given in \eqref{basedistro}.
-This formulation of the GPUDDPM is also used in \cite{gasthaus_thesis} and \cite{caron_2007}.
-The associated graphical model for this formulation is given in Figure~\ref{fig:gpuddpm_gm_2}.
+where $P(\theta_{k,t} | \theta_{k,t-1})$ is defined in \eqref{transker1} and \eqref{transker2}, $\mathbb{G}_{0}$ is defined in \eqref{basedistro}, CRP is defined in \eqref{CRP}, and $F(\theta_{c_{i,t}, t})$ is defined in \eqref{likelihood}.
+
+This formulation of the GPUDDPM is also used in \cite{gasthaus_thesis} and \cite{caron_2007}. The associated graphical model for this formulation is given in Figure~\ref{fig:gpuddpm_gm_2}.
 \begin{figure}[h]
@@
 \begin{eqnarray}
-k_{N} &=& k_{0} + N \\
-v_{N} &=& v_{0} + N \\
-\boldsymbol{\mu}_{N} &=& \frac{k_{0}}{k_{0}+N} \boldsymbol{\mu}_{0}  +  \frac{N}{k_{0}+N} \overline{\bold{x}}^{s}\\
+\kappa_{N} &=& \kappa_{0} + N \\
+\nu_{N} &=& \nu_{0} + N \\
+\boldsymbol{\mu}_{N} &=& \frac{\kappa_{0}}{\kappa_{0}+N} \boldsymbol{\mu}_{0}  +  \frac{N}{\kappa_{0}+N} \overline{\bold{x}}^{s}\\
 \Lambda_{N} &=& \Lambda_{0} + S_{\bold{x}^{s}}\\
-\bold{q}_{N} &=& \bold{q}_{0} + \sum_{i=1}^{N} \bold{x}^{c}
+\bold{q}_{N} &=& \bold{q}_{0} + \sum_{i=1}^{N} \bold{x}_{i}^{c}
 \end{eqnarray}
-where $N$ is the number of observations, $\{ \boldsymbol{\mu}_{0}, k_{0}, v_{0}, \Lambda_{N} \}$ are the $\mathcal{N}i\mathcal{W}$ prior parameters, $\bold{q}_{0}$ is the $\mathcal{D}ir$ prior parameter, $\bold{x}^{s}$ and $\bold{x}^{c}$ respectively denote the spatial and color features of the observations, and $\overline{\bold{x}}$ and $S_{\bold{x}}$ respectively denote the sample mean and sample covariance for a set of observations $\bold{x}$, defined to be
-\begin{eqnarray}
-\overline{\bold{x}}  &=&  \left( \sum_{i=1}^{N} \bold{x}_{i} \right) / N\\
-S_{\bold{x}}  &=&  \sum_{i=1}^{N} (\bold{x}_{i} - \overline{\bold{x}}) (\bold{x}_{i} - \overline{\bold{x}})^{T}
-\end{eqnarray}
+where $N$ is the number of observations, $\{ \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0} \}$ are the $\mathcal{N}i\mathcal{W}$ prior parameters, $\bold{q}_{0}$ is the $\mathcal{D}ir$ prior parameter, $\bold{x}^{s}$ and $\bold{x}^{c}$ respectively denote the spatial and color features of the observations, and $\overline{\bold{x}}$ and $S_{\bold{x}}$ respectively denote the sample mean and sample covariance for a set of observations $\bold{x}$, defined in \eqref{samplemean} and \eqref{samplecov}.
+% \begin{eqnarray}
+% \overline{\bold{x}}  &=&  \left( \sum_{i=1}^{N} \bold{x}_{i} \right) / N\\
+% S_{\bold{x}}  &=&  \sum_{i=1}^{N} (\bold{x}_{i} - \overline{\bold{x}}) (\bold{x}_{i} - \overline{\bold{x}})^{T}
+% \end{eqnarray}
 
@@
 The SMC inference algorithm is given in Algorithm~\ref{alg:SMC}.
-The algorithm operates by
+In this algorithm, at each time step $t \in \{ 1, \ldots, T \}$, a number of samples refered to as ``particles'' are generated; each particle consists of sample of the assignment for each observation $\bold{x}_{1,t}, \ldots, \bold{x}_{N_{t}, t}$, parameters for each cluster $\theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and size after deletion for each cluster $m_{1,t}, \ldots, m_{K_{t},t}$.
+A set of particles is sampled at each time step from relevant importance distributions (described in Sections~\ref{sec:smc_proposal_1}, \ref{sec:smc_proposal_2}, and \ref{sec:smc_proposal_3}), a weight is computed for each particle, and a subset of the particles are resampled and kept.
+\willie{Gibbs sampling is used to generate the samples associated with each particle at each time step.}
 
@@
 
+\begin{algorithm}
+\caption{Sequential Monte Carlo Inference for the GPUDDPM}
+\label{alg:SMC}
+\begin{algorithmic}[1]
+\FOR{$l = 1 : L$}
+\STATE $w_{0}^{(l)} \leftarrow 1/L$ \hfill $\triangleright$ initialize weights
+\ENDFOR
+\FOR{$t = 1 : T$}
+\FOR{$l = 1 : L$}
+\STATE $K_{t}^{(l)}  \leftarrow K_{t-1}^{(l)}$
+\STATE $m_{1:K_{t}^{(l)}, t}^{(l)} \leftarrow m_{1:K_{t-1}^{(l)}, t-1}^{(l)}$
+\FOR{$s = 1 : S$}
+\FOR{$i = 1 : N_{t}$}
+\IF{$s=1$}
+\STATE Sample $c_{i,t}^{(l)} \sim P(c_{i,t}^{(l)} | m_{1:K_{t-1}^{(l)}, t-1}^{(l)}, \theta_{1:K_{t-1}^{(l)}, t-1}^{(l)} )$
+\STATE $m_{c_{i,t}^{(l)},t}^{(l)} \leftarrow m_{c_{i,t}^{(l)},t}^{(l)} + 1$
+\ELSE
+\STATE $m_{c_{i,t}^{(l)},t}^{(l)} \leftarrow m_{c_{i,t}^{(l)},t}^{(l)} - 1$
+\STATE Sample $c_{i,t}^{(l)} \sim P(c_{i,t}^{(l)} | m_{1:K_{t}^{(l)}, t}^{(l)}, \theta_{1:K_{t}^{(l)}, t}^{(l)} )$  \hfill $\triangleright$ eq.~\eqref{smc_assig_eqn}
+\STATE $m_{c_{i,t}^{(l)},t}^{(l)} \leftarrow m_{c_{i,t}^{(l)},t}^{(l)} + 1$
+\ENDIF
+\IF{$c_{i,t}^{(l)} = K_{t}^{(l)} + 1$}
+\STATE Sample $\theta_{c_{i,t}^{(l)}, t}^{(l)} \sim q_{1}(\bold{x}_{i,t})$ \hfill $\triangleright$ eq.~\eqref{smc_q1_eqn}
+\STATE $K_{t}^{(l)} \leftarrow K_{t}^{(l)} + 1$
+\STATE $m_{K_{t}^{(l)},t}^{(l)} \leftarrow 1$
+\ENDIF
+\ENDFOR
+\FOR{$k = 1 : K_{t}^{(l)}$}
+\IF{$k > K_{t-1}^{(l)}$}
+\STATE Sample
+\ELSIF{$k \leq K_{t-1}^{(l)}$ and $ \#\{ \bold{x}_{1:N_{t}, t} = k \} > 0$}
+\STATE Sample
+\ELSIF{$m_{k,t}^{(l)} > 0$}
+\STATE Sample
+\ENDIF
+\IF{$s=S$}
+\STATE Sample
+\ENDIF
+\ENDFOR
+\ENDFOR
+\STATE Calculate weight of particle
+\ENDFOR
+\STATE Resample (only keep some of the $L$ particles).
+\ENDFOR
+\end{algorithmic}
+\end{algorithm}
+
 A few key distributions used in the SMC algorithm are shown below.
@@
 \subsubsection{Proposal Distribution for Assignments $c_{i}$}
+\label{sec:smc_proposal_1}
+
 The probability of assignments given current cluster sizes and cluster parameters can be written as
 \begin{equation}
+\label{smc_assig_eqn}
 P(c_{i} = k) = P(c_{i} | m_{1:K_{t}, i-1}, \theta_{1:K_{t}, t} )
@@
 \subsubsection{Proposal Distribution $q_{1}$}
-
-\begin{equation}
+\label{sec:smc_proposal_2}
+
+\begin{equation}
+\label{smc_q1_eqn}
 P(\theta_{k,t}) = P(c_{i} | m_{1:K_{t}, i-1}, \theta_{1:K_{t}, t} )
@@
 \subsubsection{Proposal Distribution $q_{2}$}
-
-\begin{equation}
+\label{sec:smc_proposal_3}
+
+\begin{equation}
+\label{smc_q2_eqn}
 P(\theta_{k,t}) = 	P(c_{i} | m_{1:K_{t}, i-1}, \theta_{1:K_{t}, t} )

### 2012-01-06 ltx/paper.tex
added more to relevant work

@@
-\documentclass[smallcondensed, draft]{svjour3}
+% header stuff
+% ------------
+
+\documentclass[smallcondensed, final]{svjour3}
 % \documentclass[smallcondensed, final]{svjour3}[1/1/2012]
-
 \usepackage[square,numbers]{natbib}
@@
 
+% title and author related
+% ------------------------
+
 \title{Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures}
+
 \titlerunning{Unsupervised Object Detection and Tracking with Dependent Dirichlet Processes}
-% \subtitle{A Bayesian Nonparametric Approach to Tracking Multiple Objects in Videos}
+
+% \subtitle{}
 
 \author{Willie Neiswanger$^{*}$ and Frank Wood$^{\dagger}$}
+
 \authorrunning{Willie Neiswanger and Frank Wood}
 
-\institute{$^{*}$Columbia University, Department of Applied Math and Applied Physics. 
+\institute{ $^{*}$Columbia University, Department of Applied Math and Applied Physics.
 Tel.: 503-464-6152.
-\email{wdn2101@columbia.edu}
-\and
-$^{\dagger}$Columbia University, Department of Statistics. 
+\email{wdn2101@columbia.edu} \and $^{\dagger}$Columbia University, Department of Statistics.
 Tel.: 212-851-2150.
@@
 
-\date{}
+\date{}  % add submit date when submitted
 
 \maketitle
+
+% Abstract
+% --------
 
@@
 
+% introduction
+% ------------
+
 \section{Introduction}
@@
 We demonstrate our implementation on multiple synthetic and benchmark datasets and compute a few standard performance metrics for each to quantify results.
+
+% related work
+% ------------
 
@@
 There has been a focus in this field on developing techniques for representing an object's appearance, both to allow for localization of objects on a per-frame basis and to allow for consistent localization of a given object over time.
-Research into the unsupervised detection of arbitrary objects (which we define as methods that accomplish both extraction and localization) has included work involving image segmentation and pixel clustering \willie{must cite here}.
-These topics usually relate to the automated partitioning of images into meaningful regions.
-Note that the popular phrase `object detection' is also typically used to refer to procedures that carry out extraction and localization; however, it is often reserved to denote methods that are tailored to find the position of specified objects or object types (as opposed to arbitrary object detection), or to denote supervised methods, where extraction is peformed implicitely during searches for specified objects and localization procedures.
+Research into the unsupervised detection of arbitrary objects (which we define as methods that accomplish both extraction and localization) has included work involving model-based clustering of individual pixels, local image texture representations, or other patterns found in image regions \cite{jain1997object, fei2005bayesian, sivic2005discovering}.
+These works usually aim to automatically segment images into meaningful regions.
+Note that the popular phrase `object detection' is sometimes used to denote methods that are tailored to find the position of specified objects or object types (as opposed to arbitrary object detection, which incorporates both the task of determining what constitutes distinct objects and localizing these objects), or to denote supervised methods, where extraction is peformed implicitely during searches for specified objects and localization procedures.
 
@@
 In addition to single object tracking, a great deal of research in this field over the past decade has focused on the development of algorithms to track multiple objects simultaneously.
-There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera), complex object interactions, object coalescence (where one object is tracked by multiple times/redundantly), objects with similar appearances, variable (and potentially high) numbers of objects, and objects that enter and exit a field of view at different times \willie{cite paper(s) for each of these}.
-
-In the past five years, there has been a push for methods that incorporate detection and tracking of multiple objects to allow for full systems capable of autonomously handling real-world video datasets.
+There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera), complex object interactions, object coalescence (where one object is tracked by multiple times/redundantly), objects with similar appearances, variable (and potentially high) numbers of objects, and objects that enter and exit a field of view at different times.
+Multiple single-object trackers running simultaneously have been shown to be ineffective, as distinct single-object trackers will tend to coalesce and track the same object.
+To remedy this problem, methods such as \cite{maccormick1999probabilistic} have developed probabilistic principles for maintaining isolation of object trackers.
+An approach to this problem involving the use of mixture models has also been studied by \cite{vermaak_2003}; this paper found success in maintaining isolation of distinct objects through the use of a nonparametric mixture model.
+
+In the past five years, there has been a push for methods that attempt to provide general algorithms for the detection and tracking of multiple objects in videos.
+Blob tracking, a basic method attempting to carry out this goal, has been in use for many years, and has found success in arenas where objects are easily discerned from the background of a video and where localization and segmentation of distinct objects is possible.
+Blob tracking-based methods run into problems when faced with videos where detection is difficult, object appearance or orientation varies heavily, there exists object occlusion,
 \willie{Describe tracking by detection, more on blob tracking, the other few methods i've found for automated joint detection and tracking}
+
+% extra background that I decided to cut (for now)
+% ------------------------------------------------
 
@@
 
-% %  ------------------------------------------------------------------------------------
+% more background that I decided to cut (for now)
+% -----------------------------------------------
+
 % % Localization-focused. Clustering methods, segmentation methods, object appearance models and representations (allows for consistent object localization over time).
@@
 Does cluster tracking on whichever extraction data, using a time dependent Dirichlet Process mixture model (to better estimate and maintain the number of targets, as was shown to be effective in vermaak and fox), which takes the form of a sequence of clusters, where the objects' appearances are modeled in a principled manner by the cluster component distributons, and the tracking behavior is modeled by time-dependencies between adjacent elements in the cluster-sequences.
-% %  ------------------------------------------------------------------------------------
+
+% Contributions (decided to cut for now)
+% --------------------------------------
 
@@
 
+% general model overview section
+% ------------------------------
+
 \section{Model Overview}
@@
 We formulate the model in a sufficently abstract manner to allow for implementations using specific object representations and extraction procedures in future studies.
+
+% general data extraction section
+% -------------------------------
 
@@
 These data are distributed roughly as worm-like shapes that follow the paths of individual objects as they move, when the data are plotted in three dimensions (where we let the frame number $t$, our discrete representation of time, ascend the vertical axis and the spatial positions occupy the horizontal axes).
+
+% this is the pretty worms figure that we decided to cut
+% ------------------------------------------------------
+
 %An example of data with such a distribution can been seen in Figure~\ref{fig:baselinedata}...The data consists of the three-dimensional baseline features and was gained by extraction peformed on a 50 frame sequence showing five ants moving across the video scene.
@@
 
+% section providing background to introduce DPMs
+% ----------------------------------------------
+
 \subsection{Dirichlet Process Mixtures}
@@
 We provide background on finite mixture models, Bayesian finite mixture models, and Dirichlet processes, which allows us to subsequently describe DPMs in context.
+
+% defining the finite mixture model
+% ---------------------------------
 
@@
 a vector of counts that sums to one).
+
+% defining the Bayesian mixture model
+% -----------------------------------
 
@@
 
+% defining Dirichlet processes
+% ----------------------------
+
 \subsubsection{Dirichlet Process}
@@
 Hence, the expectation of $G(A)$ is $\mathbb{G}_{0}$, the variance of $G(A) \rightarrow 0$ as $\alpha \rightarrow \infty$, and $G$ converges pointwise to $\mathbb{G}_{0}$ when $\alpha$ is unbounded.
+
+% defining the DPM
+% ----------------
 
@@
 
-% An alternative formultion of the DPM is known as the Chinese resturant process (CRP) sampling representation. 
 If we let $K$ be the number of distinct mixture components assigned to observations using the above model, we can write the mixture components as $\theta_{1}, \ldots, \theta_{K}$.
@@
 
+% section providing background to introduce DDPMs
+% -----------------------------------------------
+
 \subsection{Dependent Dirichlet Process Mixtures}
@@
 
+% defining the GPUDDPM
+% --------------------
+
 \subsubsection{Generalized Polya Urn Dependent Dirichlet Process Mixture}
@@
 
-Recall that the distribution over the cluster assignments $c_{i}$ (one for each observation $i$) is a function of the cluster sizes (see \eqref{crp_rep}).
+Recall that the distribution over the cluster assignments $c_{i}$ (one for each observation $i$) is a function of the cluster sizes (as per \eqref{crp_rep}).
 Hence, to account for varying numbers of clusters---and in particular, to allow clusters to diminish in size (i.e.
-reduce the number of assigned observations) and even die---there is a deletion procedure (described in \cite{caron_2007}) by which observations are considered ``removed'' from their assigned clusters at a given time step and value of $m_{k}$ (defined in \eqref{crp_rep} to be the size of cluster $k$) is reduced.
+to diminish in number of assigned observations) and even ``die off"---there is a deletion procedure that allows observations to be considered removed from their assigned clusters at a given time step.
+This procedure is introduced and described in detail in \cite{caron_2007}.
 
 The deletion procedure operates in the following way: at each time step, all previous assignments (that have not yet been deleted) are independently considered for deletion.
-Specifically, each (remaining) assignment is removed from its cluster with probability $\rho$.
-The size $m_{k,t}$ of cluster $k$ at time $t$ can be computed by removing elements from cluster $k$ at the previous time and thus diminishing the cluster's size at time $t-1$.
-It can be shown that performing deletion in this way is equivalent to, for each cluster $k$, drawing $r \sim \text{Binomial}(m_{k,t-1}, \rho)$, and reducing the cluster size $m_{k,t}$ by $r$.
-% This can be written as
-% \begin{equation}
-% \label{del_step}
-% m_{k,t} | m_{k,t-1}   \sim   m_{k,t-1}-\text{Binomial}(m_{k,t-1}, \rho)
-% \end{equation}
-For each time step, after all observations are given an assignment, this deletion procedure is carried out.
-Hence, the size $m_{k,t}$ of a given cluster $k$ at time $t$, is dependent upon the cluster size at the previous time step and on the assignments $c_{i}$ for all observations $i \in \{ 1, \ldots, N_{t} \}$ at time $t$.
+Specifically, each remaining assignment is removed from its cluster with probability $\rho$.
+It can be shown that performing deletion in this way is equivalent to, for each cluster $k$, drawing $r \sim \text{Binomial}(m_{k,t-1}, \rho)$, and reducing the previous-time cluster size $m_{k,t-1}$ by $r$.
+At each time step, this deletion procedure is carried out on all existing assignments, including those at the current time.
+Hence, the size $m_{k,t}$ of a given cluster $k$ at time $t$ is dependent upon the cluster size at the previous time step, $m_{k,t-1}$, and on the assignments $c_{i}$ for all observations $i \in \{ 1, \ldots, N_{t} \}$ at time $t$.
 We refer to the conditional distribution over $m_{k,t} | m_{k,t-1}, c_{i}, \rho$ as DEL, which we define to be
@@
 \end{equation}
-$\forall k \in \{1, \ldots, K_{t} \}$, where $\rho$ is a deletion parameter, and $\mathbb{I}(c_{i,t} = k)$ is an indicator function whose value is 1 if $c_{i,t} = k$ and 0 otherwise.
-This process adheres to what \cite{caron_2007} calls a `uniform deletion strategy' over all the observations' assignments (since assignments to each cluster have equal probability of being deleted), though a more complex deletion strategy, dependent upon cluster size, can also be implemented and is described in \cite{caron_2007}.
-The Chinese restaurant process is used to sample assignment variables during inference procedures (described in Section~\ref{sec:inference}).
-We define a distribution used in these procedures, CRP, to be
+$\forall k \in \{1, \ldots, K_{t} \}$, where $\rho$ is a deletion parameter, $K_{t}$ is the number of clusters at time $t$, and $\mathbb{I}(c_{i,t} = k)$ is an indicator function whose value is 1 if $c_{i,t} = k$ and 0 otherwise.
+This process adheres to what \cite{caron_2007} calls a ``uniform deletion strategy'' over all the observations' assignments (since assignments to each cluster have equal probability of being deleted), though a more complex deletion strategy, dependent upon cluster size, can also be implemented and is described in \cite{caron_2007}.
+
+The Chinese restaurant process, defined in \eqref{crp_rep}, is used to sample assignment variables during inference procedures (described in Section~\ref{sec:inference}).
+Based on this definition, we define the distribution CRP to be
 \begin{equation}
@@
 \end{equation}
-$\forall i \in \{1, \ldots, N \}$, where there exists $K_{t}$ clusters at time $t$, and we give a newly created cluster the index $K_{t+1}$.
+$\forall i \in \{1, \ldots, N_{t} \}$, where there exists $K_{t}$ clusters at time $t$, and we give a newly created cluster the index $K_{t+1}$.
 Note the similarity between the above definition and the Chinese resturant process of the DPM given by \eqref{crp_rep}.
@@
 
+% section making explicit connection between model and video tracking
+% -------------------------------------------------------------------
+
 \subsubsection{Object Characteristics Modeled}
@@
 
+% section providing specific model
+% --------------------------------
+
 \section{Model Specification}
@@
 
-In the following sections, a particular extraction method for generating data from videos and a specific formulation of the model described in Section~\ref{sec:modeloverview} are given.
+In the following sections, a particular extraction method for generating data from videos and a specific formulation of the general model described in Section~\ref{sec:modeloverview} are given.
 Inference schemes on this model, which when carried out allow for unsupervised detection and tracking of multiple arbitrary objects, are also described.
+
+% gives specific data extraction method (frame differencing)
+% ----------------------------------------------------------
 
@@
 
+% introduce specific model for experiments
+% ----------------------------------------
+
 \subsection{Model Implementation in Experiments}
@@
 
+% data in experiments
+% -------------------
+
 \subsubsection{Data}
@@
 
-In the following experiments, the hue component of a pixel's h-s-v value is recorded from each pixel surrounding a given $\bold{x}^{s}$ in the manner described in Section~\ref{sec:modelspec_extraction} (where we choose $L=5$).
+In the following experiments, the hue component of a pixel's hue-saturation-value (h-s-v) tuple is recorded from each pixel surrounding a given $\bold{x}^{s}$ in the manner described in Section~\ref{sec:modelspec_extraction} (where we choose $L=5$).
 Additionally, the set of possible hue values is partitioned into 10 bins, and the number of pixels with a hue value lying in each of the 10 bins is recorded to yield the vector of `color counts' $\bold{x}^{c} = x^{c_{1}}, \ldots, x^{c_{10}}$.
 Hue is chosen to represent object color since it has been demonstrated in previous work as a simple representation of object appearance that allows for distinct objects to be well differentiated (\willie{cite mckenna, other papers here}).
+
+% likelihood / object appearance model
+% ------------------------------------
 
@@
 
+% base distribution / appearance prior
+% ------------------------------------
+
 \subsubsection{Base Distribution $\mathbb{G}_{0}$ and Appearance Prior}
@@
 \willie{reference text with good background on NiW and Dir distributions}
+
+% transition kernel
+% -----------------
 
@@
 
+% recap of parameters
+% -------------------
+
 \subsection{Recap of Parameters}
@@
 
-The multivariate normal-multinomial GPUDDPM implemented in this study has a number of parameters, which can be tuned to increase the efficacy of object detection and tracking.
-Object apperance parameters include\\\\
+The multivariate normal-multinomial GPUDDPM implemented in this study has a number of parameters, which can be tuned to increase the efficacy of object detection and tracking. \\ \\
+Object appearance parameters include:
 \begin{center}
@@
 \end{tabular}
-\end{center} \vspace{5mm}
+\end{center} \vspace{3mm}
 The following parameter dictates characteristics involving the movement of objects
@@
 \end{tabular}
-\end{center} \vspace{5mm}
+\end{center} \vspace{3mm}
 Additionally, one can tune the model's tendency to detect new objects and maintain the existence of these objects (or the rate at which objects become inactive or leave the video) with the following parameters
@@
 
+% inference section
+% -----------------
+
 \section{Inference}
@@
 The first is a type of Markov Chain Monte Carlo (MCMC) batch inference, which uses Gibbs sampling to generate samples from a posterior distribution over a model, and the second is a type of Sequential Monte Carlo (SMC) inference, also known as a particle filter, which also generates samples from a posterior information, but in this case in a sequential manner that is allowed due to the time-dependent nature of the model.
+
+% MCMC inference section
+% ----------------------
 
@@
 \end{cases} \\
-c_{i,t} | \bold{c}_{1}, \ldots, \bold{c}_{t-1}, \bold{d}_{1}, \ldots, \bold{d}_{t-1}, \alpha  &\sim  \text{CRP}(\bold{c}_{1}, \ldots, \bold{c}_{t-1}, \bold{d}_{1}, \ldots, \bold{d}_{t-1}, \alpha) \\
+% c_{i,t} | \bold{c}_{1}, \ldots, \bold{c}_{t-1}, \bold{d}_{1}, \ldots, \bold{d}_{t-1}, \alpha  &\sim  \text{CRP}(\bold{c}_{1}, \ldots, \bold{c}_{t-1}, \bold{d}_{1}, \ldots, \bold{d}_{t-1}, \alpha) \\
+c_{i,t} | \bold{c}_{1:t-1}, \bold{d}_{1:t-1}, \alpha  &\sim  \text{CRP}(\bold{c}_{1:t-1}, \bold{d}_{1:t-1}, \alpha) \\
 \bold{x}_{i,t} | c_{i,t}, \theta_{c_{i,t},t}  &\sim  F(\theta_{c_{i,t}, t})
@@
 \end{align}
-where $P(\theta_{k,t} | \theta_{k,t-1})$ is defined in \eqref{transker1} and \eqref{transker2}, $\mathbb{G}_{0}$ is defined in \eqref{basedistro}, CRP is defined in \eqref{CRP}, and $F(\theta_{c_{i,t}, t})$ is defined in \eqref{likelihood}.
-
-This formulation of the GPUDDPM is also used in \cite{gasthaus_thesis} and \cite{caron_2007}. The associated graphical model for this formulation is given in Figure~\ref{fig:gpuddpm_gm_2}.
+where $\bold{c}_{t} = c_{1,t}, \ldots, c_{N_{t}, t}$ (for $N_{t}$ observations at time $t$), $\bold{d}_{t} = d_{1,t}, \ldots, d_{N_{t}, t}$, $P(\theta_{k,t} | \theta_{k,t-1})$ is defined in \eqref{transker1} and \eqref{transker2}, $\mathbb{G}_{0}$ is defined in \eqref{basedistro}, CRP is defined in \eqref{CRP}, and $F(\theta_{c_{i,t}, t})$ is defined in \eqref{likelihood}.
+This formulation of the GPUDDPM is also used in \cite{gasthaus_thesis} and \cite{caron_2007}.
+The associated graphical model for this formulation is given in Figure~\ref{fig:gpuddpm_gm_2}.
 \begin{figure}[h]
@@
 
-We define the state of the sampler to consist of the assignment variables for all observations, $c_{1}, \ldots, c_{N}$, the deletion variables for all observations, $d_{1}, \dots, d_{N}$, the sequence of cluster parameters for each active cluster $k$, $\theta_{k,1}, \ldots, \theta_{k, T}$, and the $M$ associated auxiliary variables for each cluster $\theta_{k,t}$, $z_{k,t,1}, \ldots, z_{k, t, M}$ (where $N$ denotes the number of observations and $T$ denotes the number of time steps).
-
-The sampler performs Gibbs sampling; it operates by iteratively sampling each member of the state given all other members.
+We define the state of the sampler to consist of the assignment variables for all observations, $c_{1:N_{1},1}, \ldots, c_{1:N_{T}, T}$, the deletion variables for all observations, $d_{1:N_{1},1}, \ldots, d_{1:N_{T}, T}$, the sequence of cluster parameters for each active cluster, $\theta_{k,1}, \ldots, \theta_{k, T}$ and the $M$ associated auxiliary variables for each cluster $\theta_{k,t}$ at time $t$, $z_{k,t,1}, \ldots, z_{k, t, M}$.
+
+The sampler performs an MCMC method known as Gibbs sampling; it operates by iteratively sampling each member of the state given all other members.
 In particular, it moves sequentially through each of the $N$ observations, sampling the assignment $c_{i}$ and the deletion variable $d_{i}$ at each step.
@@
 \end{equation}
-which is taken under the situation where $c = k$.
-If a new cluster is sampled as an assignment, the cluster parameters and auxiliary variables must be initialized for all time steps.
-\willie{finish writing the few tricky technical stuff here}
+which is taken under the situation where $c = k$. Note that the above integral can be determined analytically to give
+\begin{equation}
+\label{marginal_obs_prob}
+\int P(\bold{x}_{i} | \theta) \mathbb{G}_{0}(\theta)d\theta = \text{Student}()
+\end{equation}
+where ``Student'' denotes the Student's t distribution, where we follow the three-value parameterization (location parameter, scale parameter, and degrees of freedom) given in \willie{cite}.
+
+If a new cluster is sampled as an assignment, the cluster parameters and auxiliary variables must be initialized for all time steps. \willie{finish writing the few tricky technical stuff here}
 
 \subsubsection{Sampling Cluster Parameters $\theta_{k, t}$}
-The conjugacy present in chosen distributions and auxiliary variables in the transition kernel allow for posterior samples of the cluster parameters to be easily gained, and gives us
+The conjugacy of the chosen distributions in the appearance model and transition kernel allow for posterior samples of the cluster parameters to be easily gained, and gives us
 % \begin{eqnarray}
@@
 \begin{eqnarray}
+\label{bayesian_update_1}
 \kappa_{N} &=& \kappa_{0} + N \\
+\label{bayesian_update_2}
 \nu_{N} &=& \nu_{0} + N \\
+\label{bayesian_update_3}
 \boldsymbol{\mu}_{N} &=& \frac{\kappa_{0}}{\kappa_{0}+N} \boldsymbol{\mu}_{0}  +  \frac{N}{\kappa_{0}+N} \overline{\bold{x}}^{s}\\
 \Lambda_{N} &=& \Lambda_{0} + S_{\bold{x}^{s}}\\
+\label{bayesian_update_4}
 \bold{q}_{N} &=& \bold{q}_{0} + \sum_{i=1}^{N} \bold{x}_{i}^{c}
@@
 % \end{eqnarray}
-
-% \subsubsection{Metropolis-Hastings Algorithm}
 
@@
 
+% SMC inference section
+% ---------------------
+
 \subsection{SMC Inference}
@@
 In this algorithm, at each time step $t \in \{ 1, \ldots, T \}$, a number of samples refered to as ``particles'' are generated; each particle consists of sample of the assignment for each observation $\bold{x}_{1,t}, \ldots, \bold{x}_{N_{t}, t}$, parameters for each cluster $\theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and size after deletion for each cluster $m_{1,t}, \ldots, m_{K_{t},t}$.
-A set of particles is sampled at each time step from relevant importance distributions (described in Sections~\ref{sec:smc_proposal_1}, \ref{sec:smc_proposal_2}, and \ref{sec:smc_proposal_3}), a weight is computed for each particle, and a subset of the particles are resampled and kept.
+A set of particles is sampled at each time step from relevant proposal distributions (described in Sections~\ref{sec:smc_proposal_1}, \ref{sec:smc_proposal_2}, and \ref{sec:smc_proposal_3}), a weight is computed for each particle, and a subset of the particles are resampled and kept.
 \willie{Gibbs sampling is used to generate the samples associated with each particle at each time step.}
 
-\begin{algorithm}
-\caption{Sequential Monte Carlo Inference for the GPUDDPM}
-\label{alg:SMC}
-\begin{algorithmic}[1]
-\FOR{$t = 1 : T$}
-\FOR{$l = 1 : L$}
-\STATE $K_{t}^{(l)}  \leftarrow K_{t-1}^{(l)}$
-\FOR{$i = 1 : N_{t}$}
-\STATE Sample assigments from previous time sizes and cluster parameters $\theta_{k,t-1}$
-\STATE Draw $c_{i} \sim P(c_{i} | m_{1:K_{t-1}, i-1}, \theta_{1:K_{t-1}, t-1} )$
-\IF{$c_{i} = K_{t} + 1$}
-\STATE Draw a new cluster $\theta_{K_{t}+1,t} \sim \mathbb{G}_{0}(\text{priors})$   \willie{distro after algo?}
-\STATE $K_{t} \leftarrow K_{t} + 1$
-\ENDIF
-\ENDFOR
-\FOR{$s = 1 : S$}
-\FOR{$k = 1 : K$}
-\STATE Sample new cluster params via aux (given curent assignments and prev clusters)
-\ENDFOR
-\FOR{$i = 1 : N_{t}$}
-\STATE Sample new assignments (given current cluster params)
-\STATE Draw $c_{i} \sim P(c_{i} | m_{1:K_{t}, i-1}, \theta_{1:K_{t}, t} )$
-\IF{$c_{i} = K_{t} + 1$}
-\STATE Draw a new cluster $\theta_{K_{t}+1,t} \sim \mathbb{G}_{0}(\text{priors})$   \willie{distro after algo?}
-\STATE $K_{t} \leftarrow K_{t} + 1$
-\ENDIF
-\ENDFOR
-\STATE Sample new cluster sizes.
-\ENDFOR
-\ENDFOR
-\STATE Only keep some of the $L$ samples.
-\ENDFOR
-\end{algorithmic}
-\end{algorithm}
-
-\begin{algorithm}
+The sequence of target distributions for the SMC algorithm may be written as
+\begin{equation}
+\begin{split}
+%\pi_{t}(c_{1:N_{1}, 1}, \ldots, c_{1:N_{t}, t}, & \theta_{1:K_{1}, 1}, \ldots, \theta_{1:K_{t}, t}, m_{1:K_{1}, 1}, \ldots, m_{1:K_{t}, t}) = \\
+%\pi_{t-1}(c_{1:N_{1}, 1}, \ldots, c_{1:N_{t-1}, t-1}, & \theta_{1:K_{1}, 1}, \ldots, \theta_{1:K_{t-1}, t-1}, m_{1:K_{1}, 1}, \ldots, m_{1:K_{t-1}, t-1}) \\
+\pi_{t}(\bold{c}_{1:t}, \Theta_{1:t}, \bold{m}_{1:t+1}) & =  \pi_{t-1}(\bold{c}_{1:t-1}, \Theta_{1:t-1}, \bold{m}_{1:t})\\
+& \times \prod_{i=1}^{N_{t}} P(c_{i,t} | \bold{m}_{t}, \Theta_{t}, \bold{c}_{1:t}, \bold{x}_{1:N_{t}}) \\
+& \times \prod_{k=1}^{K} 
+\begin{cases}
+\begin{split}
+P(\theta_{k,t} | \theta_{k,t-1}) \hspace{2mm} & \text{if} \hspace{2mm} k \leq K_{t-1} \\
+\mathbb{G}_{0} \hspace{18mm} & \text{if} \hspace{2mm} k > K_{t-1}
+\end{split}
+\end{cases} \\
+& \times \prod_{k=1}^{K} \text{DEL}(m_{k,t+1} | m_{k,t}, c_{1:N_{t-1}, t-1}, \rho)
+\end{split}
+\end{equation}
+where $\bold{c}_{t} = c_{1,t}, \ldots, c_{N_{t}, t}$, $\Theta_{t} = \theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and $\bold{m}_{t} = m_{1,t}, \ldots, m_{K_{t}, t}$ A few distributions referenced in the SMC algorithm are described in detail in the following sections.
+
+% \begin{algorithm}
+% \caption{Sequential Monte Carlo Inference for the GPUDDPM}
+% \label{alg:SMC}
+% \begin{algorithmic}[1]
+% \FOR{$t = 1 : T$}
+% \FOR{$l = 1 : L$}
+% \STATE $K_{t}^{(l)}  \leftarrow K_{t-1}^{(l)}$
+% \FOR{$i = 1 : N_{t}$}
+% \STATE Sample assigments from previous time sizes and cluster parameters $\theta_{k,t-1}$
+% \STATE Draw $c_{i} \sim P(c_{i} | m_{1:K_{t-1}, i-1}, \theta_{1:K_{t-1}, t-1} )$
+% \IF{$c_{i} = K_{t} + 1$}
+% \STATE Draw a new cluster $\theta_{K_{t}+1,t} \sim \mathbb{G}_{0}(\text{priors})$   \willie{distro after algo?}
+% \STATE $K_{t} \leftarrow K_{t} + 1$
+% \ENDIF
+% \ENDFOR
+% \FOR{$s = 1 : S$}
+% \FOR{$k = 1 : K$}
+% \STATE Sample new cluster params via aux (given curent assignments and prev clusters)
+% \ENDFOR
+% \FOR{$i = 1 : N_{t}$}
+% \STATE Sample new assignments (given current cluster params)
+% \STATE Draw $c_{i} \sim P(c_{i} | m_{1:K_{t}, i-1}, \theta_{1:K_{t}, t} )$
+% \IF{$c_{i} = K_{t} + 1$}
+% \STATE Draw a new cluster $\theta_{K_{t}+1,t} \sim \mathbb{G}_{0}(\text{priors})$   \willie{distro after algo?}
+% \STATE $K_{t} \leftarrow K_{t} + 1$
+% \ENDIF
+% \ENDFOR
+% \STATE Sample new cluster sizes.
+% \ENDFOR
+% \ENDFOR
+% \STATE Only keep some of the $L$ samples.
+% \ENDFOR
+% \end{algorithmic}
+% \end{algorithm}
+
+\begin{algorithm}[!]
 \caption{Sequential Monte Carlo Inference for the GPUDDPM}
@@
 \ENDFOR
+\STATE $K_{0}^{(l)} \leftarrow 0$
 \FOR{$t = 1 : T$}
@@
 \IF{$s=1$}
-\STATE Sample $c_{i,t}^{(l)} \sim P(c_{i,t}^{(l)} | m_{1:K_{t-1}^{(l)}, t-1}^{(l)}, \theta_{1:K_{t-1}^{(l)}, t-1}^{(l)} )$
+\STATE Sample $c_{i,t}^{(l)} \sim P \left( c_{i,t}^{(l)} | m_{1:K_{t-1}^{(l)}, t-1}^{(l)}, \theta_{1:K_{t-1}^{(l)}, t-1}^{(l)}, \alpha \right)$  \hfill $\triangleright$ eq.~\eqref{smc_assig_eqn}
 \STATE $m_{c_{i,t}^{(l)},t}^{(l)} \leftarrow m_{c_{i,t}^{(l)},t}^{(l)} + 1$
@@
 \STATE $m_{c_{i,t}^{(l)},t}^{(l)} \leftarrow m_{c_{i,t}^{(l)},t}^{(l)} - 1$
-\STATE Sample $c_{i,t}^{(l)} \sim P(c_{i,t}^{(l)} | m_{1:K_{t}^{(l)}, t}^{(l)}, \theta_{1:K_{t}^{(l)}, t}^{(l)} )$  \hfill $\triangleright$ eq.~\eqref{smc_assig_eqn}
+\STATE Sample $c_{i,t}^{(l)} \sim P \left( c_{i,t}^{(l)} | m_{1:K_{t}^{(l)}, t}^{(l)}, \theta_{1:K_{t}^{(l)}, t}^{(l)}, \alpha \right)$  \hfill $\triangleright$ eq.~\eqref{smc_assig_eqn}
 \STATE $m_{c_{i,t}^{(l)},t}^{(l)} \leftarrow m_{c_{i,t}^{(l)},t}^{(l)} + 1$
@@
 \IF{$k > K_{t-1}^{(l)}$}
-\STATE Sample
+\STATE Sample $\theta_{k,t}^{(l)} \sim q_{1}(\{ \bold{x}_{1:N_{t}, t} = k \})$  \hfill $\triangleright$ eq.~\eqref{smc_q1_eqn}
 \ELSIF{$k \leq K_{t-1}^{(l)}$ and $ \#\{ \bold{x}_{1:N_{t}, t} = k \} > 0$}
-\STATE Sample
+\STATE Sample $\theta_{k,t}^{(l)} \sim q_{2}(\theta_{k,t-1}^{(l)}, \{ \bold{x}_{1:N_{t}, t} = k \})$  \hfill $\triangleright$ eq.~\eqref{smc_q2_eqn}
 \ELSIF{$m_{k,t}^{(l)} > 0$}
-\STATE Sample
+\STATE Sample $\theta_{k,t}^{(l)} \sim P(\theta_{k,t}^{(l)} | \theta_{k,t-1}^{(l)})$  \hfill $\triangleright$ eqs.~\eqref{transker1} $\&$ \eqref{transker2}
 \ENDIF
 \IF{$s=S$}
-\STATE Sample
+\STATE Sample $m_{k,t+1}^{(l)} \sim \text{DEL}(m_{k,t}^{(l)}, c_{1:N_{t}, t}^{(l)}, \rho)$  \hfill $\triangleright$ eq.~\eqref{del_step}
 \ENDIF
@@
 \ENDFOR
-\STATE Calculate weight of particle
+\STATE $\tilde{w}_{t}^{(l)} \leftarrow w_{t-1}^{(l)} \times \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})}	{P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) } \times r $  \hfill $\triangleright$ eq.~\eqref{smc_weight_eqn}
 \ENDFOR
-\STATE Resample (only keep some of the $L$ particles).
+\FOR {l = 1 : L}
+\STATE $w_{t}^{(l)} \leftarrow \frac{\tilde{w}_{t}^{(l)}}{\sum_{l=1}^{L} \tilde{w}_{t}^{(l)} }$
+\ENDFOR
+\STATE Resample particles $1, \ldots, L$ and weights $w^{(1)}, \ldots, w^{(L)}$
 \ENDFOR
@@
 
-A few key distributions used in the SMC algorithm are shown below.
-
 \subsubsection{Proposal Distribution for Assignments $c_{i}$}
@@
 
-The probability of assignments given current cluster sizes and cluster parameters can be written as
+The probability of assignments given current cluster sizes, cluster parameters, and the Dirichlet Process granularity parameter $\alpha$ can be written as
 \begin{equation}
 \label{smc_assig_eqn}
-P(c_{i} = k) = P(c_{i} | m_{1:K_{t}, i-1}, \theta_{1:K_{t}, t} )
-\end{equation}
+\begin{split}
+P \left( c_{i,t} | m_{1:K_{t}, t}, \theta_{1:K_{t}, t}, \alpha \right) &\propto\\
+\text{CRP}(m_{1:K_{t}, t}, \alpha) &\times
+\begin{cases}
+P(\bold{x}_{i,t} | \theta_{c_{i,t},t}) \hspace{2mm} \text{if} \hspace{2mm} k \leq K_{t-1} \\
+\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta)d\theta \hspace{2mm} k > K_{t-1}
+\end{cases}
+\end{split}
+\end{equation}
+The integral can $\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta)d\theta$ can be determined analytically, and is given in \eqref{marginal_obs_prob}.
 
@@
 
+The following is a distribution over cluster parameters $\theta_{k,t}$ given a set of $N$ observations $\bold{x}_{1:N, t}$. We define $q_{1}$ to be
 \begin{equation}
 \label{smc_q1_eqn}
-P(\theta_{k,t}) = P(c_{i} | m_{1:K_{t}, i-1}, \theta_{1:K_{t}, t} )
-\end{equation}
+q_{1}(\theta_{k,t} | \bold{x}_{1:N,t}) = P(\theta_{k,t} | \bold{x}_{1:N,t})
+\end{equation}
+where $P(\theta_{k,t} | \bold{x}_{1:N,t})$ is given by the bayesian updates found in \eqref{bayesian_update_1}, \eqref{bayesian_update_2}, \eqref{bayesian_update_3}, and \eqref{bayesian_update_4}, where the $\bold{x}_{1:N,t}$ are taken to be the observations.
 
@@
 
+The following is a distribution over cluster parameters $\theta_{k,t}$ given a set of $N$ observations $\bold{x}_{1:N, t}$ and the cluster parameters at a previous time, $\theta_{k,t-1}$.
+We define $q_{2}$ to be
 \begin{equation}
 \label{smc_q2_eqn}
-P(\theta_{k,t}) = 	P(c_{i} | m_{1:K_{t}, i-1}, \theta_{1:K_{t}, t} )
-\end{equation}
+q_{2}(\theta_{k,t} | \theta_{k,t-1},\bold{x}_{1:N,t}) = P(\theta_{k,t} | \theta_{k,t-1} \bold{x}_{1:N,t})
+\end{equation}
+where $P(\theta_{k,t} | \theta_{k,t-1} \bold{x}_{1:N,t})$ is given by the bayesian updates found in \eqref{bayesian_update_1}, \eqref{bayesian_update_2}, \eqref{bayesian_update_3}, and \eqref{bayesian_update_4}, where both the $\bold{x}_{1:N,t}$ and auxiliary variables $z_{k,t,1:M}$ are taken to be the observations.
+
+\subsection{Calculation of particle weights $\tilde{w}_{t}^{(l)}$}
+In Algorithm~\ref{alg:SMC}, the weight $\tilde{w}_{t}^{(l)}$ of each particle $l \in \{ 1. \ldots, L \}$ at time $t$ is given in terms of the particle's weight at time $t-1$ by the equation
+\begin{equation}
+\tilde{w}_{t}^{(l)} \leftarrow w_{t-1}^{(l)} \times \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})}	{P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) } \times r
+\end{equation}
+where $r$ denotes the computed ratio for the particle, and we define
+\begin{equation}
+\label{smc_weight_eqn}
+\begin{split}
+& \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})}	{P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) }  = \\
+% & \frac{P \left( \bold{x}_{1, t} | \theta_{c_{1, t}^{(l)}, t}^{(l)} \right) \times \ldots \times P \left( \bold{x}_{N_{t}, t} | \theta_{c_{N_{t}, t}^{(l)}, t}^{(l)} \right) \times
+% P \left( c_{1, t}^{(l)} | m_{1:K_{t}^{(l)}, t}^{(l)} \right) \times \ldots \times P \left( c_{N_{t}, t}^{(l)} | m_{1:K_{t}^{(l)}, t}^{(l)} \right)}{b}
+\end{split}
+\end{equation}
+
+\subsection{Resampling Particles and Particle Weights}
+
+At each time step $t \in \{ 1, \ldots, T \}$, after all of the $L$ particles have been sampled and their associated weights computed, a resampling step is carried out.
+In this step, $L$ new particles are sampled from currnt set of $L$ particles.
 
@@
 \item $N_{\text{mapped}}$ denotes the number of mapped ground-truth/detect pairs in a video, and $N_{\text{mapped}}^{(t)}$ denotes the number of mapped ground-truth/detect pairs present at frame $t$.
+\vspace{1mm}
 \end{itemize}
@@
 
-\begin{figure}[h]
+\begin{figure}[!]
   \centering
@@
 \begin{center}
-\begin{tabular}[c]{l l  l  l  l  l  l}
+\begin{tabular}[!]{l l  l  l  l  l  l}
   & & & $\hspace{8mm}\rho$ & & & \\
@@
 \begin{center}
-\begin{tabular}[c]{l l  l  l  l  l  l}
+\begin{tabular}[!]{l l  l  l  l  l  l}
 		& & & $\hspace{8mm}\rho$ & & & \\
@@
 
-\begin{figure}[h]
+\begin{figure}[!]
   \centering             

### 2012-01-07 ltx/paper.tex
lot of edits. still need to add to inference and results sections. look for next push after dinner

@@
 
-\documentclass[smallcondensed, final]{svjour3}
+\documentclass[smallcondensed, draft]{svjour3}
 % \documentclass[smallcondensed, final]{svjour3}[1/1/2012]
@@
 Areas of interest in this field involve developing metrics for judging the similarity between two object representations (for example, potential representations of the same object at adjacent frames in a video), techniques that provide a way to conduct efficient searches over a frame for regions that are similar to a given object, and methods for predicting the future position or appearance of an object.
+An overview of object tracking methods can be found in \cite{yilmaz2006object}.
 Methods for tracking single, non-rigid, objects are numerous.
@@
 This procedure---a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--optimizes the search for the region in a frame which is most similar to a target region in the previous frame.
-It involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges for each data point.
+It involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges to a number of modes.
 Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors.
@@
 In addition to single object tracking, a great deal of research in this field over the past decade has focused on the development of algorithms to track multiple objects simultaneously.
-There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera), complex object interactions, object coalescence (where one object is tracked by multiple times/redundantly), objects with similar appearances, variable (and potentially high) numbers of objects, and objects that enter and exit a field of view at different times.
+There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera) \cite{senior2006appearance, cucchiara2004probabilistic, zhou2003background}, complex object interactions \cite{khan_2004, mckenna2000tracking, dockstader2001multiple}, objects with similar appearances \cite{maccormick1999probabilistic, jepson_2003}, variable (and potentially high) numbers of objects \cite{reilly2010detection}, and objects that enter and exit a field of view at different times \cite{stauffer2003estimating, nedrich2010learning}.
 Multiple single-object trackers running simultaneously have been shown to be ineffective, as distinct single-object trackers will tend to coalesce and track the same object.
 To remedy this problem, methods such as \cite{maccormick1999probabilistic} have developed probabilistic principles for maintaining isolation of object trackers.
-An approach to this problem involving the use of mixture models has also been studied by \cite{vermaak_2003}; this paper found success in maintaining isolation of distinct objects through the use of a nonparametric mixture model.
-
-In the past five years, there has been a push for methods that attempt to provide general algorithms for the detection and tracking of multiple objects in videos.
-Blob tracking, a basic method attempting to carry out this goal, has been in use for many years, and has found success in arenas where objects are easily discerned from the background of a video and where localization and segmentation of distinct objects is possible.
-Blob tracking-based methods run into problems when faced with videos where detection is difficult, object appearance or orientation varies heavily, there exists object occlusion,
-\willie{Describe tracking by detection, more on blob tracking, the other few methods i've found for automated joint detection and tracking}
+An approach to this problem involving the use of a nonparametric mixture model has also found success in maintaining isolation of distinct objects \cite{vermaak_2003}.
+
+Over the past decade, there have been attempts to provide general algorithms for the fully unsupervised detection and tracking of arbitrary objects in videos.
+Blob tracking, a basic method attempting to carry out this goal, has been used for many years, and has found success in arenas where objects are easily discerned from the background of a video and where localization and segmentation of distinct objects is possible \cite{francois2004real, isard_2001}.
+Methods involving blob tracking, however, run into problems when faced with videos where detection is difficult, object appearance or orientation varies heavily, and there exists object occlusion \cite{song2005model}.
+To aid the accuracy of these methods, techniques have been developed for performing extraction and segmentation in a joint manner, incorporating statistical methods for maintaining hypotheses of different numbers of detected objects, and introducing some of the tracking methods described previously to track distinct blobs after they have been segmented \cite{collins2003mean, isard_2001}.
+Another family of methods related to the task of unsupervised detection and tracking of video objects goes under the heading of video segmentation algorithms---these methods extend single-frame image segmentation to maintain coherence of image segments over time, and have had some success when used for the explicit purpose of detecting and tracking foreground objects of videos \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}.
+Other attempts to perform unsupervised detection and tracking include methods for clustering short sequences of positions that can be gained by detecting the motion of objects \cite{brostow2006unsupervised, brox2010object}, which aim to return full-length distinct object tracks, and a graph based method that carries out a similar task using spectral clustering \cite{fragkiadaki2011detection}.
+Another approach based on a Gaussian mixture model has also been used to cluster data gained from moving objects \cite{pece_2002}; this method also develops heuristics for the initialization and elimination of new tracks.
+The technique we introduce in this paper falls into this category of methods, as it aims to provide a general unsupervised detection and tracking method for arbitrary objects in videos via clustering.
+Differing from previous work, we use a type of time dependent Bayesian nonparametric mixture model, and show how it can be applied to a variety of extraction data to perform detection and tracking.
 
@@
 
-This section provides details on the extracted data, Dirichlet process mixture (DPM) models, dependent Dirichlet process mixture (DDPM) models, and the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM) model.
-We make explicit the way in which characteristics of video objects can be represented by elements of these models.
-We formulate the model in a sufficently abstract manner to allow for implementations using specific object representations and extraction procedures in future studies.
+This section provides details on the data gained during extraction procedures, and on Dirichlet process mixture (DPM) models, dependent Dirichlet process mixture (DDPM) models, and the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM) model.
+We make explicit the way in which characteristics of video objects can be represented by elements of the GPUDDPM.
+We formulate the model in a sufficently abstract manner to allow for implementations in future studies that may use more-specific object representations or extraction procedures.
 
@@
 
-Inference performed on the time dependent Dirichlet process mixture model (Section~\ref{sec:inference}) carries out clustering of the data, which isolates each worm-like data cloud and allows each object to be localized and tracked.
+Modeling this data with a time dependent Dirichlet process mixture model (Section~\ref{sec:gpuddpm}) allows us to perform inference (Section~\ref{sec:inference}), which carries out clustering of the data.
+By choosing this type of model, we aim to provide a clustering result that isolates each worm-like data cloud and consequently allows each object to be localized and tracked.
 
@@
 	p_{1}, \ldots, p_{K}  &\sim  \text{Dir}(\alpha/K, \ldots, \alpha/K)\\
-		\theta_{k}  &\sim  \mathbb{G}_{0}, \hspace{2mm} \forall k \in \{1, \ldots, K\} \\
+	\theta_{1}, \ldots, \theta_{K}  &\sim  \mathbb{G}_{0} \\
+	% \theta_{k}  &\sim  \mathbb{G}_{0}, \hspace{2mm} \forall k \in \{1, \ldots, K\} \\
 	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
@@
 Note the difference between the indexing of the clusters in this model and the indexing in the previous two models.
-This formulation can be shown to be equivalent to the Bayesian mixture model defined in \eqref{bayesian_mixture_model}, when K is taken to be unbounded; as a result, this model is often called an infinite mixture model.
+This formulation can be shown to be equivalent to the Bayesian mixture model defined in \eqref{bayesian_mixture_model}, when K is taken to be unbounded; as a result, this model is sometimes called an infinite mixture model.
 
@@
 For example, if the data shows temporal dependence, the goal might be to create a sequence of DPMs, one for each time-step, where the components of the mixture at each step are dependent upon corresponding components in the both the following and previous time steps.
-\willie{Discuss and cite DDPMs}.
 
@@
 Hence, a time-dependent DDPM is a model which remains a Dirichlet process, marginally, at each time step, yet allows cluster parameters at a given time step to vary from (and remain dependent upon) the parameters in neighboring time steps.
+
+%%%%%
+% add more info here on other DDPMs, why we chose the GPUDDPM, how it has an ability to model birth and death of clusters (unlike some sort of sophisticated hidden markov model)
+%%%%%
 
@@
 We desire an extraction procedure that is as unsophisticated as possible, both to gauge the robustness of this method on potentially noisy extraction data and to ensure that the procedure is applicable to a wide range of videos.
-For these reasons we choose the method of frame-differencing for all extraction performed in this paper, which involves recording the positions of pixels (or pixel groups) that have exhibited differences in intensity or value in succesive frames beyond a given threshold.
+For these reasons we choose the method of frame-differencing for all extraction performed for experiments in this paper (Section~\ref{sec:experiments}), which involves recording the positions of pixels that have exhibited differences in intensity or value in succesive frames beyond a given threshold.
 In particular, if we let $I_{t}$ be the image difference obtained by subtracting the value of the $(i,j)^{th}$ pixel in video frame $t$ from the value of the $(i,j)^{th}$ pixel in video frame $t+1$, we can define extraction on frame $t$ to be the process that returns the dataset
 \begin{equation}
-	\Omega_{t} = \{ (i,j) | I_{t}(i,j) > \phi \}
+	\Omega_{t} = \{ (i,j,t) \hspace{2mm} | \hspace{2mm} I_{t}(i,j) > \phi \}
 \end{equation}
@@
 \begin{equation}
-	\Omega = \bigcup_{t=1}^{T} \Omega_{t} = \{ (i,j,t) | I_{t}(i,j) > \phi \}
+	\Omega = \bigcup_{t=1}^{T} \Omega_{t} = \{ (i,j,t) \hspace{2mm} | \hspace{2mm} I_{t}(i,j) > \phi,  \hspace{2mm}  \forall t \in \{1, \ldots, T \} \}
 \end{equation}
@@
 
-Frame differencing is unsophisticated, computationally inexpensive, able to be applied to a wide range of static, single-camera videos (note that videos used in the experiments described in the following sections were chosen to be static; moving-camera videos should be used in conjunction with applicable extraction methods that allow for camera movement).
+Frame differencing is unsophisticated, computationally inexpensive, able to be applied to a wide range of static, single-camera videos (note that videos used in the experiments described in the following sections were chosen to be static; moving-camera videos should be used in conjunction with applicable extraction methods that provide accurate foreground segmentation for non-stationary video).
 We have found that, when correctly implemented, frame differencing is sufficiently general to extract the desired worm-like data clouds from a variety of videos containing multiple moving objects.
@@
   \subfloat[]{\label{fig:longimg_all}\includegraphics[width=1\textwidth]{../img/longimg_7_detail.pdf}} 
-  \caption{Pairs of consecutive frames and the results produced by taking the pixel-wise difference between these frames.
-\willie{Want to add colorbar that explains how color represents frame number red = frame 1, blue = frame 100}}
+  \caption{Pairs of consecutive frames and the results produced by taking the pixel-wise difference between these frames (a - f).
+The final image shows the result of frame differencing over a sequence of images (from the PETS2009 dataset), where color denotes frame number.}
   \label{fig:img_and_framediff}
@@
 For each pixel $x = (i, j, t) \in \Omega$ (defined above), we specify a square, $L$ pixels in length, centered on $(i, j)$, that selects a set of pixels surrounding $\bold{x}$ in frame $t$.
-We also specify a scalar color value that is able to be computed for each pixel; this value could, for example, be some function of the r-g-b or h-s-v value of a pixel.
+We also specify a scalar color value that is able to be computed for each pixel; this value could, for example, be some function of the red-green-blue (r-g-b) or hue-saturation-value (h-s-v) characteristic of a pixel.
 The color value of each of the selected pixels surrounding $\bold{x}$ is recorded.
@@
 
-\willie{Show example color-count vector (histogram) for different objects in an image. give more-formal definition like for frame differencing above.}
+% perhaps: Show example color-count vector (histogram) for different objects in an image. give more-formal definition like for frame differencing above.
 
@@
 
-The frame differencing and color counting extraction outlined in Section~\ref{sec:modelspec_extraction} yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \mathbb{Z}_{+}$, where each element $\bold{x} \in \bold{X}$ can be written
+The frame differencing and color counting extraction outlined in Section~\ref{sec:modelspec_extraction} yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \{1, \ldots, T \}$, where each element $\bold{x} \in \bold{X}$ can be written
 \begin{equation}
@@
 \end{equation}
-where $\bold{x}^{s} \in \mathbb{R}^{2}$ is the two dimensional vector of spatial coordinates,  $\bold{x}^{c} \in \mathbb{Z}_{+}^{V}$ is the $V$ discrete dimensional vector of color counts, and $t \in \mathbb{Z}_{+}$ is the discrete time index.
-
-In the following experiments, the hue component of a pixel's hue-saturation-value (h-s-v) tuple is recorded from each pixel surrounding a given $\bold{x}^{s}$ in the manner described in Section~\ref{sec:modelspec_extraction} (where we choose $L=5$).
+where $\bold{x}^{s} \in \mathbb{R}^{2}$ is the two dimensional vector of spatial coordinates,  $\bold{x}^{c} \in \mathbb{Z}_{+}^{V}$ is the $V$ discrete dimensional vector of color counts, and $t \in \{1, \ldots, T \}$ is the discrete time index.
+
+In the following experiments, the hue component of the h-s-v value is recorded from all pixels that surround each extracted pixel $\bold{x}$ in the manner described in Section~\ref{sec:modelspec_extraction} (where we choose $L=5$).
 Additionally, the set of possible hue values is partitioned into 10 bins, and the number of pixels with a hue value lying in each of the 10 bins is recorded to yield the vector of `color counts' $\bold{x}^{c} = x^{c_{1}}, \ldots, x^{c_{10}}$.
-Hue is chosen to represent object color since it has been demonstrated in previous work as a simple representation of object appearance that allows for distinct objects to be well differentiated (\willie{cite mckenna, other papers here}).
+Hue is chosen to represent object color since it has been demonstrated in previous work as a simple representation of object appearance that allows for distinct objects to be well differentiated \cite{perez_2002, raja_1998, mckenna_1999}.
 
@@
 In particular, the multivariate normal distribution over the spatial features $\bold{x}^{s}$ can be thought to represent the shape of each object as an oval; likewise, data generated by each moving object during extraction are often ovular---as noisy extraction procedures cause some smoothing of edges and corners, producing blob-like shapes even when objects are not particularly round---and centered on a given object.
-Furthermore, this model is justified as the maximum likelihood parameter estimate of a normal distribution corresponds to the least squares fit of data relative to the mean of the distribution.
-\willie{needs to be said more accurately and explained why this helps.} Modeling the color features $\bold{x}^{c}$ as draws from a multinomial distribution (equivalently, as draws from a product of discrete distributions), is justified since distinct object tend to generate pixels whose hue values are noisy but yield consistent counts in discrete hue bins.
+Furthermore, this model is justified as the maximum likelihood parameter estimate of a normal distribution corresponds to the least squares fit of data relative to the normal distribution mean; since extraction data produced by an object clusters around the centroid of the object, the estimated normal distribution mean should provide a reasonable reflection of the object's centroid.
+Modeling the color features $\bold{x}^{c}$ as draws from a multinomial distribution (equivalently, as draws from a product of discrete distributions), is justified since we have observed that distinct object tend to generate pixels whose hue values are noisy but yield consistent counts in discrete hue bins.
 
@@
 where $\mathcal{N}i\mathcal{W}$ denotes the normal-inverse-Wishart distribution, $\mathcal{D}ir$ denotes the Dirichlet distribution, and the prior has the hyperparameters $\boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0}$ and $\bold{q}_{0}$.
-\willie{reference text with good background on NiW and Dir distributions}
+
+% perhaps: reference text with good background on NiW and Dir distributions
 
@@
 
-We wish to formulate a transition kernel $P(\theta_{t} | \theta_{t-1})$ that provides a reasonable representation of how we expect tracked objects to behave.
-In the interest of formulating a model with the intention of using it to track arbitrary objects, we do not wish to make many assumptions about the long-term behavior of objects (though transition kernels that incorporate, for example, dynamics could potentially be implemented if one is aware beforehand of an object's behavior).
+We wish to formulate a transition kernel $P(\theta_{t} | \theta_{t-1})$ that provides a reasonable representation of how we expect tracked objects to move over time.
+In the interest of formulating a model with the intention of using it to track arbitrary objects, we do not wish to make many assumptions about the long-term behavior of objects.
+For example, we choose not to adopt a transition kernel that incorporates object dynamics, though a transition kernel that models this or other sophisticated behavioral tendencies could potentially be implemented in future works if one knows that they are consistent characteristics of an object's movement.
 
@@
 \end{eqnarray}
+
+% note:
+% don't i have to include more to really define the transition kernel?? I feel there is something that is not fully specified.
 
@@
 \end{center}
-\willie{Discuss choosing parameters}
+
+We have not found an easily justifiable way to choose parameters.
+For the appearance and movement parameters, samples were drawn given a range of parameter values, and those that yielded samples similar in appearance to the experimental data were kept (a technique also performed in \cite{gasthaus_thesis}).
+The parameters chosen for use in experiments (Section~\ref{sec:experiments}) were accepted based on the fact that they yielded reasonable results.
+Additionally, a sensitivity analysis was carried out for the granularity and deletion parameters ($\alpha$ and $\rho$), where a range of values for both were chosen, and inference performed for each; results of this analysis are discussion in Section~\ref{sec:sens_analysis}.
 
@@
 By formulating the detection and tracking task in terms of a generative model, a variety of methods previously developed for statistical inference can be readily applied; we separate presentations of model and inference so that multiple inference methods, each with its own benefits in terms of precision and speed, can be formulated and applied independently to gain detection and tracking results.
-This section provides background on the two inference procedures implemented in this study, used to perform Bayesian inference on the model described in Section~\ref{sec:modelspec}.
-The first is a type of Markov Chain Monte Carlo (MCMC) batch inference, which uses Gibbs sampling to generate samples from a posterior distribution over a model, and the second is a type of Sequential Monte Carlo (SMC) inference, also known as a particle filter, which also generates samples from a posterior information, but in this case in a sequential manner that is allowed due to the time-dependent nature of the model.
+This section provides background on the two inference procedures implemented in this study, both used to perform Bayesian inference on the model described in Section~\ref{sec:modelspec}.
+The first is a type of Markov Chain Monte Carlo (MCMC) batch inference, which uses Gibbs sampling to generate samples from a posterior distribution over a model, and the second is a type of Sequential Monte Carlo (SMC) inference, also known as a particle filter, which also generates samples from a posterior information, but in a sequential manner that operates by making use of the time-dependent nature of the model.
 
@@
 We define the state of the sampler to consist of the assignment variables for all observations, $c_{1:N_{1},1}, \ldots, c_{1:N_{T}, T}$, the deletion variables for all observations, $d_{1:N_{1},1}, \ldots, d_{1:N_{T}, T}$, the sequence of cluster parameters for each active cluster, $\theta_{k,1}, \ldots, \theta_{k, T}$ and the $M$ associated auxiliary variables for each cluster $\theta_{k,t}$ at time $t$, $z_{k,t,1}, \ldots, z_{k, t, M}$.
-
+Formally, we can write
+\begin{equation}
+\text{State} = \{ \bold{c}_{1:T}, \bold{d}_{1:T}, \Theta_{1:T}, \bold{m}_{1:T} \}
+\end{equation}
+where $\bold{c}_{t} = c_{1:N_{t},t}$, $\bold{d}_{t} = d_{1:N_{t},t}, \Theta_{t} = \theta_{1:K_{t}, t}$, and $m_{1:K_{t},t}$.
 The sampler performs an MCMC method known as Gibbs sampling; it operates by iteratively sampling each member of the state given all other members.
-In particular, it moves sequentially through each of the $N$ observations, sampling the assignment $c_{i}$ and the deletion variable $d_{i}$ at each step.
-After the final observation at a given time $t$ step is sampled, the cluster parameters for all active clusters at $t$ are sampled, and a M-H step is employed to sample the $M$ auxiliary variables for each active cluster.
-The following sections detail how each sample is generated will be explicitely shown in the following sections.
-
-\subsubsection{Sampling Assignment Variables $c_{i}$}
+In particular, at each time $t$ it moves sequentially through each of the $N_{t}$ observations, sampling the assignment $c_{i,t}$ and the deletion variable $d_{i,t}$.
+After the final observation at a given time is sampled, the cluster parameters for all active clusters at $t$ are sampled, and a M-H step is employed to sample the $M$ auxiliary variables for each active cluster.
+The following sections detail how each sample in the model is generated.
+
+% MCMC inference : sampling assignment variables
+% ----------------------------------------------
+
+\subsubsection{Sampling Assignment Variables $c_{i,t}$}
 \label{sec:sample_assignments}
 
-A value proportional to the posterior probability can be computed for each possible value that $c_{i}$ may take on; these values allow us to construct a discrete probability distribution that may be sampled from to get a sample from the posterior distribution over assignments.
-The possible values that the assignment may take on are $c_{i} = k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*}\}$, where $K_{t}^{'}$ is the number clusters with a non-zero size at any point between observation $i$ and its deletion point $d_{i}$, and $k^{*}$ is a potential new cluster.
-The probability that $c_{i}$ is assigned to a cluster $k$ can be computed for each $k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*} \}$, given all other members of the state, by
-\begin{equation}
-p(c_{i} = k | state) \propto \prod_{i' = i}^{d_{i}} P(c_{i'} | m_{i'-1}, c_{1:i'-1}) \newline \times
+A value proportional to the posterior probability can be computed for each possible value that $c_{i,t}$ may take on; these values allow us to construct a discrete probability distribution that may be drawn from to get a sample from the posterior distribution over assignments.
+The possible values that the assignment may take on are $c_{i,t} = k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*}\}$, where $K_{t}^{'}$ is the number clusters with a non-zero size at any point between time $t$ and the assignment's associated deletion time $d_{i,t}$, and $k^{*}$ is a potential new cluster.
+The probability that $c_{i,t}$ is assigned to a cluster $k$ can be computed for each $k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*} \}$, given all other members of the state, by
+\begin{equation}
+\begin{split}
+p(c_{i,t} = k | \text{state}) & \propto \prod_{t' = t}^{d_{i,t}} P(c_{i,t'} | m_{k,t'-1}, \bold{c}_{1:t'-1}) \\
+ & \times
 \begin{cases}
-	P(\bold{x}_{i} | \theta_{k,t})                                        \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
-	\int P(\bold{x}_{i} | \theta) \mathbb{G}_{0}(\theta) d\theta    \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
+	P(\bold{x}_{i,t} | \theta_{k,t})                                        \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
+	\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta) d\theta    \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
 \end{cases}
-\end{equation}
-\willie{$d_{i}$ is actually a time; product needs to actually go up to last observation at time $d_{i}$} where $t$ is the time of observation $i$, and the distribution under the product $P(c_{i'} | m_{1:K_{t}^{'}, i'-1}, c_{1:i'-1})$ is given by
+\end{split}
+\end{equation}
+where $t$ is the time of observation $i$, and the distribution under the product $P(c_{i'} | m_{1:K_{t}^{'}, i'-1}, c_{1:i'-1})$ is given by 
 \begin{equation}
@@
 If a new cluster is sampled as an assignment, the cluster parameters and auxiliary variables must be initialized for all time steps. \willie{finish writing the few tricky technical stuff here}
+
+% MCMC inference : sampling cluster parameters
+% --------------------------------------------
 
@@
 
+% MCMC inference : sampling auxiliary variables
+% ---------------------------------------------
+
 \subsubsection{Sampling Auxiliary Variables $z_{k,t,m}$}
@@
 
+% MCMC inference : sampling deletion variables
+% --------------------------------------------
+
 \subsubsection{Sampling Deletion Variables $d_{i}$}
@@
 and accept or reject this sample using the standard M-H acceptance ratio \willie{write computation of ratio}.
+
+% MCMC inference : initializing the sampler
+% -----------------------------------------
 
@@
 
+% experiments section
+% -------------------
+
 \section{Experiments}
@@
 \subsubsection{Sensitivity Analysis}
+\label{sec:sens_analysis}
 

### 2012-01-07 ltx/paper.tex
pretty much final first draft

@@
 
-\documentclass[smallcondensed, draft]{svjour3}
+\documentclass[smallcondensed, final]{svjour3}
 % \documentclass[smallcondensed, final]{svjour3}[1/1/2012]
@@
 \begin{align}
+\label{gpuddpm_def}
 \begin{split}
@@
 
-\subsubsection{Transition Kernels and Motion Behavior}
+\subsubsection{Transition Kernel and Motion Model}
 
@@
 \label{sec:inference}
-Algorithms that infer latent model parameters in a Bayesian manner given the data gained during extraction provide detection and tracking results.
-By formulating the detection and tracking task in terms of a generative model, a variety of methods previously developed for statistical inference can be readily applied; we separate presentations of model and inference so that multiple inference methods, each with its own benefits in terms of precision and speed, can be formulated and applied independently to gain detection and tracking results.
-This section provides background on the two inference procedures implemented in this study, both used to perform Bayesian inference on the model described in Section~\ref{sec:modelspec}.
-The first is a type of Markov Chain Monte Carlo (MCMC) batch inference, which uses Gibbs sampling to generate samples from a posterior distribution over a model, and the second is a type of Sequential Monte Carlo (SMC) inference, also known as a particle filter, which also generates samples from a posterior information, but in a sequential manner that operates by making use of the time-dependent nature of the model.
+Algorithms that infer latent model parameters via Bayesian inference given the data gained during extraction provide detection and tracking results.
+By formulating the detection and tracking task in terms of a generative model, a variety of methods previously developed for statistical inference can be readily applied; we separate presentations of model and inference so that multiple inference methods, each with its own benefits in terms of precision and speed, can be formulated as independent algorithms for detection and tracking.
+This section provides background on the two inference algorithms implemented in this study, which are both used to perform Bayesian inference on the model described in Section~\ref{sec:modelspec}.
+The first is a type of Markov Chain Monte Carlo (MCMC) batch inference, which uses Gibbs sampling to generate samples from the posterior distribution of the model.
+The second is a type of Sequential Monte Carlo (SMC) inference, also known as a particle filter, which also generates samples from the posterior distribution of the model, but in a sequential manner that makes use of the time-dependent nature of the data.
 
@@
 The generative model associated with the GPUDDPM is illustrated in Section~\ref{sec:gpuddpm}.
-This representation includes cluster size $m$ (the number of observations assigned to a given cluster), which is standard in DPM constructions.
-To capture time-dependence, the GPUDDPM allows clusters to `lose assigned observations' in a deletion step, which modifies the sizes $m$ over time.
-Instead of incorporating cluster size variables $m$ directly, we can formulate an equivalent model which makes use of deletion variables $d$, which represent the time when observations are lost from a cluster .
+To allow for modeling time-dependent data, the GPUDDPM allows clusters to `lose assigned observations' in a deletion step, which modifies the values of the cluster sizes over time.
+Instead of incorporating the cluster size variable $m$ directly (as it was in the definition of the GPUDDPM given in \eqref{gpuddpm_def}), we can formulate an equivalent model which makes use of deletion variables $d$, which represent the time at which observations are lost from a cluster.
+Specifically, we introduce variables $d_{1}, \ldots, d_{N}$ which denote the times at which the assignments of given observations are considered to be removed.
 At each time step, cluster sizes can be recontructed from all previous assignment and deletion variables.
-In particular, we introduce variables $d_{1}, \ldots, d_{N}$ which denote the times at which the assignments of given observations are considered to be removed.
-With these variables, we can now express the $m_{k}$ value at time $t$ as
+With these we can express the size of cluster $k$ at time $t$, $m_{k,t}$, as
 \begin{equation}
@@
 where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise.
-Additionally, for an observation $x_{i}$ at a given time-step $t$, the deletion time $d_{i}$ can be defined to be $d_{i} = t + l_{i}$, where $l_{i}$ is considered the lifespan of an assignment, is distributed geometrically, and can be expressed as
+Additionally, for an observation $\bold{x}_{i}$ at a given time-step $t$, the deletion time $d_{i}$ can be defined to be $d_{i} = t + l_{i}$, where $l_{i}$ is considered the lifespan of an assignment, is distributed geometrically, and can be expressed as
 \begin{equation}
@@
 where $\bold{c}_{t} = c_{1,t}, \ldots, c_{N_{t}, t}$ (for $N_{t}$ observations at time $t$), $\bold{d}_{t} = d_{1,t}, \ldots, d_{N_{t}, t}$, $P(\theta_{k,t} | \theta_{k,t-1})$ is defined in \eqref{transker1} and \eqref{transker2}, $\mathbb{G}_{0}$ is defined in \eqref{basedistro}, CRP is defined in \eqref{CRP}, and $F(\theta_{c_{i,t}, t})$ is defined in \eqref{likelihood}.
-This formulation of the GPUDDPM is also used in \cite{gasthaus_thesis} and \cite{caron_2007}.
+This formulation of the GPUDDPM is also used by \cite{gasthaus_thesis} and \cite{caron_2007}.
 The associated graphical model for this formulation is given in Figure~\ref{fig:gpuddpm_gm_2}.
@@
 
-% MCMC inference : sampling assignment variables
-% ----------------------------------------------
+% mcmc: sampling assignment variables
+% -----------------------------------
 
@@
 \begin{split}
-p(c_{i,t} = k | \text{state}) & \propto \prod_{t' = t}^{d_{i,t}} P(c_{i,t'} | m_{k,t'-1}, \bold{c}_{1:t'-1}) \\
+p(c_{i,t} = k | \text{State}) & \propto
+\prod_{i'=1}^{N_{t}}  \text{CRP}(c_{i',t} | \bold{m}_{t-1}, \alpha) \\
+& \times \prod_{t' = t+1}^{d_{i,t}}  \prod_{i'=1}^{N_{t'}}   \text{CRP}(c_{i',t'} | \bold{m}_{t'-1}, \alpha) \\
  & \times
@@
 \end{equation}
-where $t$ is the time of observation $i$, and the distribution under the product $P(c_{i'} | m_{1:K_{t}^{'}, i'-1}, c_{1:i'-1})$ is given by 
-\begin{equation}
-P(c_{i} = k | m_{1:K_{t}^{'}, i-1}, c_{1:i-1}, \alpha) \propto 
-\begin{cases}
-m_{k, i-1} \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
-\alpha \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
-\end{cases}
-\end{equation}
-which is taken under the situation where $c = k$. Note that the above integral can be determined analytically to give
+% where $t$ is the time of observation $i$, and the distribution under the product $P(c_{i'} | m_{1:K_{t}^{'}, i'-1}, c_{1:i'-1})$ is given by 
+%\begin{equation}
+%P(c_{i} = k | m_{1:K_{t}^{'}, i-1}, c_{1:i-1}, \alpha) \propto 
+%\begin{cases}
+%m_{k, i-1} \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
+%\alpha \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
+%\end{cases}
+%\end{equation}
+where the distribution CRP is given by \eqref{crp_rep}, and the sizes of the clusters $\bold{m}_{t'-1}$ are calculated given the fact that $c_{i,t} = k$.
+Note that the above integral can be determined analytically to give
 \begin{equation}
 \label{marginal_obs_prob}
-\int P(\bold{x}_{i} | \theta) \mathbb{G}_{0}(\theta)d\theta = \text{Student}()
-\end{equation}
-where ``Student'' denotes the Student's t distribution, where we follow the three-value parameterization (location parameter, scale parameter, and degrees of freedom) given in \willie{cite}.
-
-If a new cluster is sampled as an assignment, the cluster parameters and auxiliary variables must be initialized for all time steps. \willie{finish writing the few tricky technical stuff here}
-
-% MCMC inference : sampling cluster parameters
-% --------------------------------------------
+\int P(\bold{x}_{i} | \theta) \mathbb{G}_{0}(\theta)d\theta = \text{Student}_{\nu_{n} - 1}(\mu_{n}, \Lambda_{n}/(\kappa_{n}(\nu_{n}-1)))
+\end{equation}
+where ``Student'' denotes the Student's t distribution, where we follow the three-value parameterization (location parameter, scale parameter, and degrees of freedom) given in \cite{gelman2004bayesian}.
+
+%%%%%
+% Need to include ratio of Dirichlet also in the above posterior predictive distro of a single observation
+%%%%%
+
+If a new cluster is sampled as an assignment, the cluster parameters and auxiliary variables for this new cluster must be initialized for all time steps before sampling can proceed.
+In experiments, this initialization was carried out by iteratively samping forward to time T and backwards to time 1 via the transition kernel.
+
+% mcmc: sampling cluster parameters
+% ---------------------------------
 
@@
 \begin{split}
-P(\theta_{k,t} | state) & = P(\bold{x}_{i} | \theta_{k,t}) P(z_{k,t+1,1:M} | \theta_{k,t})  P(\theta_{k,t} | z_{k,t,1:M}) \\
+P(\theta_{k,t} | \text{State}) & = P(\bold{x}_{i,t} | \theta_{k,t}) P(z_{k,t+1,1:M} | \theta_{k,t})  P(\theta_{k,t} | z_{k,t,1:M}) \\
  & = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}_{k,t}, \Sigma_{k,t} | \boldsymbol{\mu}_{N}, k_{N}, v_{N}, \Lambda_{N})  \mathcal{D}ir( \bold{p}_{k,t} | \bold{q}_{N})
@@
 \end{align}
-Where the parameters in the above distribution are given when the data at time $t$, auxiliary variables $z_{k,t-1,1:M}$ at time $t-1$, and auxiliary variables $z_{k,t,1:M}$ at time $t$, are taken to be the observations in the following Bayesian updates
+Where the parameters in the above distribution are given when the data at time $t$, auxiliary variables $z_{k,t-1,1:M}$ at time $t-1$, and auxiliary variables $z_{k,t,1:M}$ at time $t$, are taken to be the observations for the following Bayesian updates
 \begin{eqnarray}
@@
 
-% MCMC inference : sampling auxiliary variables
-% ---------------------------------------------
+% mcmc: sampling auxiliary variables
+% ----------------------------------
 
@@
 \begin{equation}
-P(z_{k,t,m} | state) \propto  P(z_{k,t,m} | \theta_{k,t-1}) P(\theta_{k,t} | z_{k,t,1:M})
+P(z_{k,t,m} | \text{State}) \propto  P(z_{k,t,m} | \theta_{k,t-1}) P(\theta_{k,t} | z_{k,t,1:M})
 \end{equation}
@@
 
-% MCMC inference : sampling deletion variables
-% --------------------------------------------
+% mcmc: sampling deletion variables
+% ---------------------------------
 
@@
 \end{align}
-and accept or reject this sample using the standard M-H acceptance ratio \willie{write computation of ratio}.
-
-% MCMC inference : initializing the sampler
-% -----------------------------------------
+and accept or reject this sample using the standard M-H acceptance ratio.
+
+%%%%%
+% need to include the "standard M-H acceptance ratio"
+%%%%%
+
+% mcmc: initializing the sampler
+% ------------------------------
 
@@
 The SMC inference algorithm is given in Algorithm~\ref{alg:SMC}.
-In this algorithm, at each time step $t \in \{ 1, \ldots, T \}$, a number of samples refered to as ``particles'' are generated; each particle consists of sample of the assignment for each observation $\bold{x}_{1,t}, \ldots, \bold{x}_{N_{t}, t}$, parameters for each cluster $\theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and size after deletion for each cluster $m_{1,t}, \ldots, m_{K_{t},t}$.
-A set of particles is sampled at each time step from relevant proposal distributions (described in Sections~\ref{sec:smc_proposal_1}, \ref{sec:smc_proposal_2}, and \ref{sec:smc_proposal_3}), a weight is computed for each particle, and a subset of the particles are resampled and kept.
-\willie{Gibbs sampling is used to generate the samples associated with each particle at each time step.}
+In this algorithm, at each time step $t \in \{ 1, \ldots, T \}$, a number of samples refered to as ``particles'' are generated; each particle consists of a sample from the posterior distribution over the assignment for each observation, $c_{1,t}, \ldots, c_{N_{t}, t}$, parameters for each cluster $\theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and size after deletion for each cluster $m_{1,t}, \ldots, m_{K_{t},t}$.
+A set of particles is sampled at each time step from relevant proposal distributions (described in Sections~\ref{sec:smc_proposal_1}, \ref{sec:smc_proposal_2}, and \ref{sec:smc_proposal_3}), a weight is computed for each particle, and a subset of the particles are resampled and kept via a resampling process.
+Furthermore, within each time step, Gibbs sampling is used to generate the samples associated with each particle.
 
@@
 
+%%%%%
+% in SMC algorithm: still need to include computing r and make sure things are correct
+%%%%%
+
 \begin{algorithm}[!]
@@
 \ENDFOR
-\STATE $K_{0}^{(l)} \leftarrow 0$
-\FOR{$t = 1 : T$}
-\FOR{$l = 1 : L$}
+\STATE $K_{0}^{(l)} \leftarrow 0$ \hfill $\triangleright$ initialize \# of clusters
+\FOR{$t = 1 : T  \text{  (\# of frames})$}
+\FOR{$l = 1 : L  \text{  (\# of particles})$}
 \STATE $K_{t}^{(l)}  \leftarrow K_{t-1}^{(l)}$
 \STATE $m_{1:K_{t}^{(l)}, t}^{(l)} \leftarrow m_{1:K_{t-1}^{(l)}, t-1}^{(l)}$
-\FOR{$s = 1 : S$}
-\FOR{$i = 1 : N_{t}$}
+\FOR{$s = 1 : S \text{  (\# of Gibbs samples})$}
+\FOR{$i = 1 : N_{t} \text{  (\# of observations at frame $t$})$}
 \IF{$s=1$}
@@
 \ENDIF
-\IF{$c_{i,t}^{(l)} = K_{t}^{(l)} + 1$}
+\IF{$c_{i,t}^{(l)} = K_{t}^{(l)} + 1  \text{  (a new cluster)}$}
 \STATE Sample $\theta_{c_{i,t}^{(l)}, t}^{(l)} \sim q_{1}(\bold{x}_{i,t})$ \hfill $\triangleright$ eq.~\eqref{smc_q1_eqn}
@@
 \ENDFOR
-\FOR{$k = 1 : K_{t}^{(l)}$}
+\FOR{$k = 1 : K_{t}^{(l)}  \text{  (\# of clusters at frame $t$})$}
 \IF{$k > K_{t-1}^{(l)}$}
@@
 \FOR {l = 1 : L}
-\STATE $w_{t}^{(l)} \leftarrow \frac{\tilde{w}_{t}^{(l)}}{\sum_{l=1}^{L} \tilde{w}_{t}^{(l)} }$
+\STATE $w_{t}^{(l)} \leftarrow \frac{\tilde{w}_{t}^{(l)}}{\sum_{l=1}^{L} \tilde{w}_{t}^{(l)} }$   \hfill $\triangleright$ normalize weights
 \ENDFOR
-\STATE Resample particles $1, \ldots, L$ and weights $w^{(1)}, \ldots, w^{(L)}$
+\STATE Resample particles $1, \ldots, L$ and weights $w^{(1)}, \ldots, w^{(L)}$   \hfill $\triangleright$ Section~\ref{sec:resample}
 \ENDFOR
@@
 \subsection{Resampling Particles and Particle Weights}
+\label{sec:resample}
 
@@
 We would like to incorporate these points into our model.
-To do this, we modify our base distribution $\mathbb{G}_{0}$ to be a two component mixture, where one component is the base distribution and the other is a multivariate normal distribution with a fixed `flat' covarance $\sigma = [100, 0; 0, 100]$ multiplied with an `even' multinomial distribution \willie{need more rigorus terms here} with a fixed parameter $\bold{p} = (1/K, \ldots, 1/K)$.
+To do this, we modify our base distribution $\mathbb{G}_{0}$ to be a two component mixture, where one component is the base distribution and the other is a multivariate normal distribution with a fixed `flat' covarance $\sigma = [100, 0; 0, 100]$ multiplied with an `even' multinomial distribution with a fixed parameter $\bold{p} = (1/K, \ldots, 1/K)$.
+
+%%%%%
+% define more rigorously, and then say why we don't include in experiments?
+%%%%%
 
@@
 Inference was carried out using the MCMC algorithm (Section~\ref{sec:MCMC}); the posterior sample taken as the result correctly tracked both colored squares through occlusion in both videos, and is shown in Figure~\ref{fig:synth_one_plot}.
-\willie{still need to include performance metric results for these synthetic datasets.}.
+
+%%%%%
+% I still need to include performance metric results for these synthetic datasets.}.
+%%%%%
 
@@
 The MCMC inference algorithm correctly tracked all three objects through occlusion and inferred the appearance and size shifts.
-Figure~\ref{fig:synth2_alphaimg} shows a sample from the posterior distribution of the cluster parameters, where the mean and oval representation of the covariance matrix (with $0.5$ confidence value) are overlayed on the data.
-The data is plotted in three dimensional space, where time is represented on the vertical axis, and each data point is colored by its assignment to one of the three inferred clusters.
-Both the cluster parameter sample and data are overlaid on three semi-transparent planes showing the objecs in the video in three frames.
+Figure~\ref{fig:synth_one_plot} shows a sample from the posterior distribution of the cluster parameters, where the mean and oval representation of the covariance matrix (with $0.5$ confidence value) are overlayed on the data.
+The data is plotted in three dimensional space, where time is represented on the vertical axis, and the assignment of each data point to one of the three inferred clusters is denoted by color and marker type.
+
+%Both the cluster parameter sample and data are overlaid on three semi-transparent planes showing the objecs in the video in three frames.
 \willie{still need to include performance metric results for this synthetic dataset}.
-
 Groundtruth for the synthetic videos
@@
 
-\begin{figure}[h]
-	\centering
-    \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/synth2_alphaimg_hd.png}}
-    \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2001_alphaimg_hr.png}}
-	\subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2000_4objects.png}}
-    \caption{\label{fig:synth2_alphaimg} Sample from posterior of the state for synthetic experiment two (a), and benchmark datasets PETS2001 (b) and PETS2000 (c), where color represents assignment and the mean and standard deviation are shown.
+% 3-panel alpha-img figure that I removed
+
+% \begin{figure}[h]
+% 	\centering
+%     \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/synth2_alphaimg_hd.png}}
+%     \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2001_alphaimg_hr.png}}
+% 	\subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2000_4objects.png}}
+%     \caption{\label{fig:synth2_alphaimg} Sample from posterior of the state for synthetic experiment two (a), and benchmark datasets PETS2001 (b) and PETS2000 (c), where color represents assignment and the mean and standard deviation are shown.
 In each plot, the sample and data are overlaid on three images from the experiment.}
-\end{figure}
+% \end{figure}
 
@@
 
-A video dataset used in both the PETS2009 and PETS2010 conferences, called `S2.L1 at time sequence 12.34' was chosen for experimentation due to its prominence in a number of studies \willie{ref papers using this dataset}.
-This dataset consists of a monocular, stationary camera, 794 frame video sequence.
-The entire video sequence was used in this experiment.
+A video dataset used in both the PETS2009 and PETS2010 conferences, called `S2.L1 at time sequence 12.34' was chosen for experimentation due to its prominence in a number of studies \cite{ellis_2010}.
+%%%%%
+% cite other papers that use this dataset?
+%%%%%
+This dataset consists of a monocular, stationary camera, 794 frame video sequence. The entire video sequence was used in this experiment.
 
@@
 
-Discuss parameters somewhere in here.
-to add in discussion: thoughts about: deletion param vs.
-dataset size (and tracking through occlusion vs clusters taking over), mvn cluster size vs.
+We have presented a new technique for the unsupervised detection and tracking of arbitrary objects in videos.
+The primary intention of this technique is to reduce the need for detection or localization methods tailored to specific object types and serve as a general framework applicable to videos with varied objects, backgrounds, and film qualities.
+The GPUDDPM, a time-dependent Dirichlet process mixture, has been introduced, and we have shown how inference on this model allows for the desired clustering of extraction data.
+Furthermore, we have demonstrated a specific implementation of the model using spatial and color pixel data extracted via frame differencing and provided two algorithms for performing inference on the model to accomplish detection and tracking.
+Both algorithms were carried out on multiple synthetic and benchmark multi-object video datasets in order to demonstrate an ability to accomplish unsupervised detection and tracking of arbitrary objects in both manufactured and real world settings.
+We have described and computed standard performance metrics for our method's detection and tracking results, and found it to be comparable with other techniques that were designed for the specific object-types present in the videos in question.
+These results illustrate the ability of the technique described in this paper to, without modification, detect and track objects with diverse physical charactersitics moving over non-uniform backgrounds and through occlusion
+
+More to include: more discussion of parameter choices, thoughts about deletion params vs.
+dataset size (and the tradeoff between tracking through occlusion vs clusters rapidly gaining size), mvn cluster size vs.
 tracking in occlusion vs.
 sampling correct number of objects.
-we detect movement, which isn’t exactly same as perf.
-metrics (ie our tracking stops if people stop moving).
+discuss how we detect movement, which isn’t exactly same as authors in the perf.
+metric paper (ie our tracking is designed to stop if people stop moving).
 objects hard to track through occlusion becuase they are small, similar appearance, and our algorithm doesn’t assume dynamics to keep robust (we could, probably for better accuracy but less generality).

### 2012-01-08 ltx/paper.tex
a few quick edits on inference section

@@
 p(c_{i,t} = k | \text{State}) & \propto
-\prod_{i'=1}^{N_{t}}  \text{CRP}(c_{i',t} | \bold{m}_{t-1}, \alpha) \\
+\prod_{i'=i}^{N_{t}}  \text{CRP}(c_{i',t} | \bold{m}_{t-1}, \alpha) \\
 & \times \prod_{t' = t+1}^{d_{i,t}}  \prod_{i'=1}^{N_{t'}}   \text{CRP}(c_{i',t'} | \bold{m}_{t'-1}, \alpha) \\
@@
 \begin{equation}
+\label{stationary_pdf}
 P(z_{k,t,m} | \text{State}) \propto  P(z_{k,t,m} | \theta_{k,t-1}) P(\theta_{k,t} | z_{k,t,1:M})
@@
 \begin{algorithmic}[1]
-\STATE $f(x)$ is the desired stationary PDF
-\STATE $P(x_{i} | x_{i-1})$ is the proposal pdf
-\STATE Initialize $x_{0}$
-\FOR{$i = 1 : burn\hspace{1pt}in + N$}
-\STATE Draw $x^{*} \sim P(x_{i} | x_{i-1})$
-\STATE Draw $\alpha \sim U(0,1)$
-\STATE $r_{accept} \leftarrow \frac{f(x^{*})P(x_{i-1} | x^{*})} {f(x_{i-1})P(x^{*} | x_{i-1})} $
-\IF{$\alpha < min \{ 1, r_{accept} \} $}
+\STATE $f(x)$ is the stationary PDF from which samples are desired \hfill $\triangleright$ eq. \eqref{stationary_pdf}
+\STATE $q( \cdot | x_{i-1})$ is the proposal pdf \hfill $\triangleright$ eq. \eqref{proposal_distro}
+\STATE $x_{0} \leftarrow$ initialize
+\STATE B $\leftarrow$ \# of burn in samples
+\FOR{$i = 1 : \text{B} + N$}
+\STATE Sample $x^{*} \sim q( \cdot | x_{i-1})$
+\STATE Sample $\alpha \sim \text{Uniform}(0,1)$
+\STATE $r_{accept} \leftarrow \frac{f(x^{*})q(x_{i-1} | x^{*})} {f(x_{i-1})q(x^{*} | x_{i-1})} $ \hfill $\triangleright$  eq. \eqref{accept_ratio}
+\IF{$\alpha < \text{min} \{ 1, r_{accept} \} $}
 \STATE $x_{i} \leftarrow x^{*}$
@@
 \ENDFOR
-\STATE Discard samples $x_{1}, \ldots, x_{burn\hspace{1pt}in}$
+\STATE Discard samples $x_{1}, \ldots, x_{\text{B}}$
 \STATE \textbf{Output}: Remaining samples $x_{1}, \ldots, x_{N}$
@@
 \begin{equation}
+\label{proposal_distro}
 z_{k,t,m}^{*}  \sim  P(z_{k,t,m} | \theta_{k,t}) = \mathcal{N}(z_{k,t,m}^{s} | \boldsymbol{\mu}_{k,t}, \Sigma_{k,t}) \mathcal{M}n(z_{k,t,m}^{c} | \bold{p}_{k,t})
@@
 \begin{equation}
+\label{accept_ratio}
 r_{accept} = \frac{P(\theta_{k,t-1} | z_{k,t,m}^{*})}{P(\theta_{k,t-1} | z_{k,t,m})}
 \end{equation}
-and keep this sample with probability $min(1, r_{accept})$.
+and keep this sample with probability $\text{min}(1, r_{accept})$.
 
@@
 A set of particles is sampled at each time step from relevant proposal distributions (described in Sections~\ref{sec:smc_proposal_1}, \ref{sec:smc_proposal_2}, and \ref{sec:smc_proposal_3}), a weight is computed for each particle, and a subset of the particles are resampled and kept via a resampling process.
-Furthermore, within each time step, Gibbs sampling is used to generate the samples associated with each particle.
+Within each time step, Gibbs sampling is used to generate the samples associated with each particle.
 
@@
 %\pi_{t-1}(c_{1:N_{1}, 1}, \ldots, c_{1:N_{t-1}, t-1}, & \theta_{1:K_{1}, 1}, \ldots, \theta_{1:K_{t-1}, t-1}, m_{1:K_{1}, 1}, \ldots, m_{1:K_{t-1}, t-1}) \\
-\pi_{t}(\bold{c}_{1:t}, \Theta_{1:t}, \bold{m}_{1:t+1}) & =  \pi_{t-1}(\bold{c}_{1:t-1}, \Theta_{1:t-1}, \bold{m}_{1:t})\\
+\pi_{t}(\bold{c}_{1:t}, \Theta_{1:t}, \bold{m}_{1:t}) & =  \pi_{t-1}(\bold{c}_{1:t-1}, \Theta_{1:t-1}, \bold{m}_{1:t-1})\\
 & \times \prod_{i=1}^{N_{t}} P(c_{i,t} | \bold{m}_{t}, \Theta_{t}, \bold{c}_{1:t}, \bold{x}_{1:N_{t}}) \\
@@
 \end{cases} \\
-& \times \prod_{k=1}^{K} \text{DEL}(m_{k,t+1} | m_{k,t}, c_{1:N_{t-1}, t-1}, \rho)
+& \times \prod_{k=1}^{K} \text{DEL}(m_{k,t} | m_{k,t-1}, c_{1:N_{t-1}, t-1}, \rho)
 \end{split}
@@
 \begin{split}
-P \left( c_{i,t} | m_{1:K_{t}, t}, \theta_{1:K_{t}, t}, \alpha \right) &\propto\\
-\text{CRP}(m_{1:K_{t}, t}, \alpha) &\times
+P \left( c_{i,t} | m_{1:K_{t}, t}, \theta_{1:K_{t}, t}, \alpha \right) &\propto
+\text{CRP}(m_{1:K_{t}, t}, \alpha) \\ &\times
 \begin{cases}
@@
 
-\section*{Discussion}
+\section*{Discussion and Conclusion}
 \label{sec:discussion}
@@
 
-\section*{Conclusion}
+%\section*{Conclusion}
 

### 2012-01-10 ltx/paper.tex
a small change
received full performance metric dataset from Ellis/Ferryman, included bar graph of this data to show comparison of our technique with object-specific trackers

@@
 In the following experiments, the hue component of the h-s-v value is recorded from all pixels that surround each extracted pixel $\bold{x}$ in the manner described in Section~\ref{sec:modelspec_extraction} (where we choose $L=5$).
-Additionally, the set of possible hue values is partitioned into 10 bins, and the number of pixels with a hue value lying in each of the 10 bins is recorded to yield the vector of `color counts' $\bold{x}^{c} = x^{c_{1}}, \ldots, x^{c_{10}}$.
+Additionally, the set of possible hue values is partitioned into 10 bins, and the number of pixels with a hue value lying in each of the 10 bins is recorded to yield the vector of ``color counts'' $\bold{x}^{c} = x^{c_{1}}, \ldots, x^{c_{10}}$.
 Hue is chosen to represent object color since it has been demonstrated in previous work as a simple representation of object appearance that allows for distinct objects to be well differentiated \cite{perez_2002, raja_1998, mckenna_1999}.
@@
 The multivariate normal-multinomial model on which inference is described in this section can be used to gain object detection and tracking results.
-The nature of object detection and tracking results may vary depending on the application or performance metrics used (see Section~\ref{sec:performance_metrics}).
-In this study, we desired the centroid position and approximate spatial region for each moving object, for each frame that this object can be seen within the video.
-We chose to make our results depend completely on the inferred parameters of the multivariate normal distribution of each cluster.
+The nature of object detection and tracking results may vary depending on the application or performance metrics used (see Section~\ref{sec:performance_metrics}); in this study, we desire the centroid position and approximate spatial region for each moving object, for each frame that this object can be seen within the video.
+We choose to make our detection and tracking results depend completely on the inferred parameters of the multivariate normal distribution of each cluster.
 Each inferred cluster was taken to be a distinct detected object, and the sequence of means and covariance matrices for a given cluster were used, respectively, to determine the position and spatial region of a given object over a sequence of time steps.
@@
 
-Due to the computation required for the MCMC batch inference method (discussed further in Section~\ref{sec:discussion}), only the final 1000 frames of the video were used.
+Due to the computation required for the MCMC batch inference method (discussed further in Section~\ref{sec:discussion}), only the final 1000 frames of the video were used from both datasets.
 Extraction was performed with frame differencing as described in Section~\ref{sec:modelspec_extraction}, using $L=3$.
@@
 The MCMC sampler was successful for both benchmark videos; each object was detected, tracked, and its shape estimated in manner very consistent with the ground-truth.
-The results for the PETS2000 dataset are displayed in Figure~\ref{fig:pets2000_results} and for the PETS2001 dataset in Figure~\ref{fig:pets2001_results}; in these figure, a sample from the posterior of the cluster parameters is overlayed on the extracted data over a sequence of frames, where the assignment of each data point is represented by color and marker type.
+The results for the PETS2000 dataset are displayed in Figure~\ref{fig:pets2000_results} and for the PETS2001 dataset in Figure~\ref{fig:pets2001_results}; in these figures, a sample from the posterior distribution of the cluster parameters is overlayed on the extracted data over a sequence of frames, where the assignment of each data point is represented by color and marker type.
 
@@
 
-The SMC inference algorithm yielded a sample from the posterior distribution, from which the object detection and tracking results were obtained (as described in Section~\ref{sec:inference_to_results}).
-In Figure \ref{fig:pets2009_results}, a sample from the posterior of the cluster parameters is overlayed on the extracted data over a sequence of frames, where the assignment of each data point is represented by color and marker type.
+The SMC inference algorithm yielded a sample from the posterior distribution of the model, from which the object detection and tracking results were obtained (as described in Section~\ref{sec:inference_to_results}).
+In Figure \ref{fig:pets2009_results}, a sample from the posterior distribution over the cluster parameters is overlayed on the extracted data over a sequence of frames, where the assignment of each data point is represented by color and marker type.
 
@@
   \centering
-  % \subfloat[]{\label{fig:pets2009_results} \includegraphics[width=0.7\textwidth]{../img/pets2009_1-50_front.pdf}}
-  \includegraphics[width=0.7\textwidth]{../img/pets2009_1-50_front.pdf}
-  \caption{Results for the PETS2009 dataset, showing a sample from the posterior distribution of the state for frames 1-50, where the vertical axis denotes time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.}
+  \subfloat[]{\includegraphics[width=0.43\textwidth]{../img/pets2009_1-50_front.pdf}}
+  \hspace{4mm}
+  \subfloat[]{\includegraphics[width=0.53\textwidth]{../img/pets2009full_pm_b.pdf}}
+  \caption{ (a) Results for the PETS2009 dataset, showing a sample from the posterior distribution of the state for frames 1-50, where the vertical axis denotes time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.
+(b) performance metrics vs.
+covariance confidence interval threshold for tracking results showing the SFDA (blue solid line) and ATA (red dashed line).}
   \label{fig:pets2009_results}
-  % \label{fig:benchmark_results_2}
+\end{figure}
+
+In \cite{ellis_2010}, performance metrics were computed for a number of studies that carried out object detection and tracking for the PETS2009/2010 dataset.
+As this dataset consists solely of humans, many of these algorithms were developed for the specific purpose of people tracking (i.e.
+not for general detection and tracking of arbitrary objects); in consequence, many of these studies involve appearance models specifically developed for humans, sometimes making use of their orientation in this specific dataset, and motion models based on assumptions about human motion.
+We compare SFDA and ATA results of our strategy with these methods to show that our general method can yield comparable results even when compared with object-specific trackers.
+The following table shows performance metric results for comparison (data published with permission from the authors of \cite{ellis_2010}).
+\begin{figure}[!]
+  \centering
+  \subfloat[]{\includegraphics[width=0.49\textwidth]{../img/pets2009full_comp_sfda.pdf}}
+  \hspace{1mm}
+  \subfloat[]{\includegraphics[width=0.49\textwidth]{../img/pets2009full_comp_ata.pdf}}
+  \caption{Performance metric comparison showing the (a) SFDA and (b) ATA results for submissions to the PETS2009/PETS2010 workshops shows comparable performance with state-of-the-art, object-specific detection and tracking strategies.
+Results for the technique in this paper are shown in red, and were computed using a covariance confidence interval threshold of 0.35.}
+  \label{fig:sfda_and_ata_comparison}
 \end{figure}

### 2012-01-10 ltx/paper.tex
a few minor edits

@@
 We compare SFDA and ATA results of our strategy with these methods to show that our general method can yield comparable results even when compared with object-specific trackers.
-The following table shows performance metric results for comparison (data published with permission from the authors of \cite{ellis_2010}).
+Figure~\ref{fig:sfda_and_ata_comparison} shows performance metric results for comparison (data published with permission from the authors of \cite{ellis_2010}).
 \begin{figure}[!]
@@
   \subfloat[]{\includegraphics[width=0.49\textwidth]{../img/pets2009full_comp_ata.pdf}}
-  \caption{Performance metric comparison showing the (a) SFDA and (b) ATA results for submissions to the PETS2009/PETS2010 workshops shows comparable performance with state-of-the-art, object-specific detection and tracking strategies.
-Results for the technique in this paper are shown in red, and were computed using a covariance confidence interval threshold of 0.35.}
+  \caption{Performance metric comparison showing the (a) SFDA and (b) ATA results for submissions to the PETS2009/PETS2010 workshops (blue bars) and for the technique introduced in this paper (red bars).
+These graphs show comparable performance for the general technique in this paper with state-of-the-art, object-specific detection and tracking strategies.
+Our results were computed using a covariance confidence interval threshold of 0.35.}
   \label{fig:sfda_and_ata_comparison}

### 2012-01-11 ltx/paper.tex
some more edits

@@
 After the final observation at a given time is sampled, the cluster parameters for all active clusters at $t$ are sampled, and a M-H step is employed to sample the $M$ auxiliary variables for each active cluster.
-The following sections detail how each sample in the model is generated.
+Before the samples can be iteratively drawn from each element of the state given all other elements, the state must be initialized.
+While the state of the sampler can be initialized at any value, beginning at a value closer to the maximum joint probability (over the state) allows for quicker inference.
+The following sections detail how each sample in the model is drawn.
 
@@
 
-A value proportional to the posterior probability can be computed for each possible value that $c_{i,t}$ may take on; these values allow us to construct a discrete probability distribution that may be drawn from to get a sample from the posterior distribution over assignments.
+A value proportional to the posterior probability can be computed for each possible value that $c_{i,t}$ may take on; these values allow us to construct a discrete probability distribution from which we can draw samples from the posterior distribution over assignments.
 The possible values that the assignment may take on are $c_{i,t} = k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*}\}$, where $K_{t}^{'}$ is the number clusters with a non-zero size at any point between time $t$ and the assignment's associated deletion time $d_{i,t}$, and $k^{*}$ is a potential new cluster.
@@
 %%%%%
-% Need to include ratio of Dirichlet also in the above posterior predictive distro of a single observation
+% Need to include ratio of Dirichlet also in the above posterior predictive distro of a single observation, and discuss how to compute any parameters in here that aren't yet specified
 %%%%%
@@
 \begin{algorithmic}[1]
-\STATE $f(x)$ is the stationary PDF from which samples are desired \hfill $\triangleright$ eq. \eqref{stationary_pdf}
+\STATE $f(x)$ is the stationary PDF from which $N$ samples are desired \hfill $\triangleright$ eq. \eqref{stationary_pdf}
 \STATE $q( \cdot | x_{i-1})$ is the proposal pdf \hfill $\triangleright$ eq. \eqref{proposal_distro}
+\STATE B is the number of burn in samples
 \STATE $x_{0} \leftarrow$ initialize
-\STATE B $\leftarrow$ \# of burn in samples
+%\STATE B $\leftarrow$ \# of burn in samples
 \FOR{$i = 1 : \text{B} + N$}
@@
 
-The M-H algorithm is used to sample auxiliary variables for each alive cluster.
+The M-H algorithm is used to sample auxiliary variables for each alive (positive size) cluster.
 In short, we sample a new value for all $M$ auxiliary variables, denoted $z_{k,t,m}^{*}$, according to the following proposal distribution
@@
 \end{equation}
-and keep this sample with probability $\text{min}(1, r_{accept})$.
+and choose these auxiliary variables as the next sample in the constructed sequence if $\alpha < \text{min}(1, r_{accept})$, where $\alpha$ is a sample from a $\text{Uniform}(0,1)$ distribution.
 
@@
 \subsubsection{Sampling Deletion Variables $d_{i}$}
-Computing the posterior distribution (or a value proportional to this distribution) for each possible value that a given deletion variable may take on (in a manner similar to how samples from the posterior distribution of the assignment variables $c_{i}$ are gained in Section~\ref{sec:sample_assignments}) may be computationally expensive due to the large number of potential deletion times.
+Computing the posterior distribution (or a value proportional to this distribution) for each possible value that a given deletion variable may take on, may be accomplished in a manner similar to how samples from the posterior distribution of the assignment variables $c_{i}$ are gained in Section~\ref{sec:sample_assignments}; however, this may be computationally expensive due to the large number of potential deletion times.
 To remedy this, the M-H algorithm may again be used to sample from the posterior, where the prior distribution for the deletion times can be used to propose new samples $d_{i}^{*}$.
-As the `lifetime' of each assignment can be shown to be geometrically distributed, we can propose a new deletion time sample with
+As the ``lifetime'' of each assignment can be shown to be geometrically distributed, we can propose a new deletion time sample with
 \begin{align}
@@
 
-% mcmc: initializing the sampler
-% ------------------------------
-
-\subsubsection{Initializing the Sampler}
-Before the samples can be iteratively drawn from each element of the state given all other elements, the state must be initialized.
-While the state of the sampler can be initialized at any value, beginning at a value closer to the maximum joint probability (over the state) allows for quicker inference.
-
 % SMC inference section
@@
 In this algorithm, at each time step $t \in \{ 1, \ldots, T \}$, a number of samples refered to as ``particles'' are generated; each particle consists of a sample from the posterior distribution over the assignment for each observation, $c_{1,t}, \ldots, c_{N_{t}, t}$, parameters for each cluster $\theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and size after deletion for each cluster $m_{1,t}, \ldots, m_{K_{t},t}$.
-A set of particles is sampled at each time step from relevant proposal distributions (described in Sections~\ref{sec:smc_proposal_1}, \ref{sec:smc_proposal_2}, and \ref{sec:smc_proposal_3}), a weight is computed for each particle, and a subset of the particles are resampled and kept via a resampling process.
+A set of particles is sampled at each time step from relevant proposal distributions (described in Sections~\ref{sec:smc_proposal_1}, \ref{sec:smc_proposal_2}, and \ref{sec:smc_proposal_3}), a weight is computed for each particle, and a new set of particles are sampled from the set of weighted particles via a resampling process.
 Within each time step, Gibbs sampling is used to generate the samples associated with each particle.
@@
 \end{equation}
-where $\bold{c}_{t} = c_{1,t}, \ldots, c_{N_{t}, t}$, $\Theta_{t} = \theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and $\bold{m}_{t} = m_{1,t}, \ldots, m_{K_{t}, t}$ A few distributions referenced in the SMC algorithm are described in detail in the following sections.
+where $\bold{c}_{t} = c_{1,t}, \ldots, c_{N_{t}, t}$, $\Theta_{t} = \theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and $\bold{m}_{t} = m_{1,t}, \ldots, m_{K_{t}, t}$ The distributions used in the SMC algorithm are described in detail in the following sections.
 
@@
 \ENDFOR
-\STATE Resample particles $1, \ldots, L$ and weights $w^{(1)}, \ldots, w^{(L)}$   \hfill $\triangleright$ Section~\ref{sec:resample}
+\STATE Resample particles $1, \ldots, L$ and weights $w_{t}^{(1)}, \ldots, w_{t}^{(L)}$   \hfill $\triangleright$ Section~\ref{sec:resample}
 \ENDFOR
@@
 \end{equation}
-The integral can $\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta)d\theta$ can be determined analytically, and is given in \eqref{marginal_obs_prob}.
+where CRP is defined in \eqref{CRP}, and the integral $\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta)d\theta$ can be determined analytically, and is given in \eqref{marginal_obs_prob}.
 
@@
 \end{equation}
-where $P(\theta_{k,t} | \bold{x}_{1:N,t})$ is given by the bayesian updates found in \eqref{bayesian_update_1}, \eqref{bayesian_update_2}, \eqref{bayesian_update_3}, and \eqref{bayesian_update_4}, where the $\bold{x}_{1:N,t}$ are taken to be the observations.
+where samples can be drawn from $P(\theta_{k,t} | \bold{x}_{1:N,t})$ using the bayesian updates found in \eqref{bayesian_update_1}, \eqref{bayesian_update_2}, \eqref{bayesian_update_3}, and \eqref{bayesian_update_4}, where the $\bold{x}_{1:N,t}$ are taken to be the observations.
 
@@
 \end{equation}
-where $P(\theta_{k,t} | \theta_{k,t-1} \bold{x}_{1:N,t})$ is given by the bayesian updates found in \eqref{bayesian_update_1}, \eqref{bayesian_update_2}, \eqref{bayesian_update_3}, and \eqref{bayesian_update_4}, where both the $\bold{x}_{1:N,t}$ and auxiliary variables $z_{k,t,1:M}$ are taken to be the observations.
+where samples can be drawn from $P(\theta_{k,t} | \theta_{k,t-1} \bold{x}_{1:N,t})$ using the bayesian updates found in \eqref{bayesian_update_1}, \eqref{bayesian_update_2}, \eqref{bayesian_update_3}, and \eqref{bayesian_update_4}, where both the $\bold{x}_{1:N,t}$ and auxiliary variables $z_{k,t,1:M}$ are taken to be the observations.
 
@@
 \end{equation}
-where $r$ denotes the computed ratio for the particle, and we define
+where $r$ denotes the computed ratio for the particle (defined in Algorithm~\ref{alg:SMC}), and we define
 \begin{equation}
@@
 
+%%%%%
+% need to complete this equation
+%%%%%
+
 \subsection{Resampling Particles and Particle Weights}
@@
 At each time step $t \in \{ 1, \ldots, T \}$, after all of the $L$ particles have been sampled and their associated weights computed, a resampling step is carried out.
-In this step, $L$ new particles are sampled from currnt set of $L$ particles.
+In this step, $L$ new particles are sampled from current set of $L$ particles.
+We will describe a specific, simple method known as Multinomial Resampling.
+In this method of resampling, we draw
+\begin{equation}
+\Upsilon \sim \mathcal{M}n((w_{t}^{(1)}, \ldots, w_{t}^{(L)}))
+\end{equation}
+where $\mathcal{M}n$ denotes the multinomial distribution, and $\Upsilon$ represents a vector of counts showing the multiplicity of each of the old particles in the new particle set (i.e.
+$\Upsilon_{i}$ denotes how many times particle $i$ is duplicated in the new particle set).
+More sophisticated resampling strategies are described for the SMC algorithm in \cite{gasthaus_thesis} and \cite{douc2005comparison}.
+
+%%%%%
+% need to include particle resampling strategies
+%%%%%
 
@@
 
+% Noisy extraction procedures may result in spurrious data points that do not correspond to any foreground object.
+We would like to incorporate these points into our model.
+To do this, we modify our base distribution $\mathbb{G}_{0}$ to be a two component mixture, where one component is the base distribution and the other is a multivariate normal distribution with a fixed `flat' covarance $\sigma = [100, 0; 0, 100]$ multiplied with an `even' multinomial distribution with a fixed parameter $\bold{p} = (1/K, \ldots, 1/K)$.
 Noisy extraction procedures may result in spurrious data points that do not correspond to any foreground object.
 We would like to incorporate these points into our model.
-To do this, we modify our base distribution $\mathbb{G}_{0}$ to be a two component mixture, where one component is the base distribution and the other is a multivariate normal distribution with a fixed `flat' covarance $\sigma = [100, 0; 0, 100]$ multiplied with an `even' multinomial distribution with a fixed parameter $\bold{p} = (1/K, \ldots, 1/K)$.
+To do this, we modify our model to be a two component mixture model, where one component is the previous model (the GPUDDPM) and the other is a multivariate normal distribution with with a fixed `flat' covarance $\sigma = [100, 0; 0, 100]$ multiplied with an `even' multinomial distribution with a fixed parameter $\bold{p} = (1/K, \ldots, 1/K)$.
+This normal-multinomial component is left intentionally broad, and only the mean parameter of the normal distribution is inferred.
+We can write this generative model as
+\begin{align}
+\label{background_model}
+\begin{split}
+\gamma_{i,t} | p_{1}, p_{2} &\sim \text{Discrete}(p_{1}, p_{2})\\
+m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho  &\sim \text{DEL}(m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) \\
+\theta_{k, t} | \theta_{k, t-1}   &\sim
+\begin{cases}
+	P(\theta_{k, t} | \theta_{k, t-1}) \hspace{2mm} \text{if} \hspace{2mm} k \in \{ 1, \ldots, K_{t} \} \\
+	\mathbb{G}_{0}   \hspace{2mm} \text{if} \hspace{2mm} k = K_{t+1}
+\end{cases} \\
+c_{i,t} | m_{1:K_{t},t}, \hspace{1mm} \alpha  &\sim  \text{CRP} (m_{1:K_{t},t}, \hspace{1mm} \alpha) \\
+\bold{x}_{i,t} | c_{i,t}, \theta_{1:K_{t}, t}, \gamma_{i,t}, \phi &\sim
+\begin{cases}
+	F(\theta_{c_{i,t}, t}) \hspace{2mm} \text{if} \hspace{2mm} \gamma_{i,t} = 1 \\
+	\phi \hspace{2mm} \text{if} \hspace{2mm} \gamma_{i,t} = 2
+\end{cases} \\
+\end{split}
+\end{align}
+where $\phi$ represents the normal-multinomial component distribution defined above.
 
 %%%%%
-% define more rigorously, and then say why we don't include in experiments?
+% define more rigorously, and then say why we don't include in experiments?--- or perhaps include for an example in experiments
 %%%%%
@@
   \subfloat[]{\includegraphics[width=0.32\textwidth]{../img/synth4.pdf}}
-  \caption{Each plot shows a sample from the posterior distribution of the state for one of the videos in synthetic experiment one, where the time axis ascends vertically, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.
-In both cases, the two objects are successfully tracked through occlusion, though both objects reverse their directions in the second video.}
+  \caption{Each plot shows a sample from the posterior distribution of the model for synthetic experiments one (a and b) and two (c), where the vertical axis represents frame number, the horizontal axes represent spatial position, objects are denoted by marker colors and marker types, and the mean and standard deviation are shown.
+In all cases, the objects are successfully tracked through occlusion, whether they travel in a straight line (a), reverse direction (b), or do a combination of both (c).}
   \label{fig:synth_one_plot}
@@
   \caption{Performance metric comparison showing the (a) SFDA and (b) ATA results for submissions to the PETS2009/PETS2010 workshops (blue bars) and for the technique introduced in this paper (red bars).
-These graphs show comparable performance for the general technique in this paper with state-of-the-art, object-specific detection and tracking strategies.
+These graphs show that the general technique in this paper yields comparable performance with state-of-the-art, object-specific detection and tracking strategies.
 Our results were computed using a covariance confidence interval threshold of 0.35.}

### 2012-01-12 ltx/paper.tex
some edits

@@
 We will use the phrase ``baseline'' three-dimensional extraction data to refer to the features $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$ of pixels present in regions containing objects.
-These data are distributed roughly as worm-like shapes that follow the paths of individual objects as they move, when the data are plotted in three dimensions (where we let the frame number $t$, our discrete representation of time, ascend the vertical axis and the spatial positions occupy the horizontal axes).
+These data are distributed roughly as worm-like shapes that follow the paths of individual objects as they move, when the data are plotted in three dimensions (where we represent the frame number $t$, our discrete representation of time, on the vertical axis and spatial position on the horizontal axes).
+Modeling this data with a time dependent Dirichlet process mixture model (Section~\ref{sec:gpuddpm}) allows us to perform inference (Section~\ref{sec:inference}), which carries out clustering of the data.
+By choosing this type of model, we aim to provide a clustering result that isolates each worm-like data cloud and infers a sequence of distributions for each, which we use to gain localization and tracking results for each object (Section~\ref{sec:inference_to_results}).
 
@@
 % \end{figure}
-
-Modeling this data with a time dependent Dirichlet process mixture model (Section~\ref{sec:gpuddpm}) allows us to perform inference (Section~\ref{sec:inference}), which carries out clustering of the data.
-By choosing this type of model, we aim to provide a clustering result that isolates each worm-like data cloud and consequently allows each object to be localized and tracked.
 
@@
 \begin{equation}
+\label{mcmc_assig_vars}
 \begin{split}
@@
 \begin{split}
-& \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})}	{P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) }  = \\
+& \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})}	{P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) }  = 
+\frac{P(\bold{x}_{1:N_{t},t} | \theta_{1:K_{t}^{(l)}}) P(c_{1:N_{t}, t}^{(l)} | m_{1:K_{t},t}^{(l)}) )}{P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t})}
+\\
+%
 % & \frac{P \left( \bold{x}_{1, t} | \theta_{c_{1, t}^{(l)}, t}^{(l)} \right) \times \ldots \times P \left( \bold{x}_{N_{t}, t} | \theta_{c_{N_{t}, t}^{(l)}, t}^{(l)} \right) \times
 % P \left( c_{1, t}^{(l)} | m_{1:K_{t}^{(l)}, t}^{(l)} \right) \times \ldots \times P \left( c_{N_{t}, t}^{(l)} | m_{1:K_{t}^{(l)}, t}^{(l)} \right)}{b}
+%
 \end{split}
 \end{equation}
+where $P(\bold{x}_{1:N_{t},t} | \theta_{1:K_{t}^{(l)}})$ is defined by \eqref{likelihood}, $P(c_{1:N_{t}, t}^{(l)} | m_{1:K_{t},t}^{(l)}))$ is defined by \eqref{CRP}, and $P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t})$ is defined by \eqref{mcmc_assig_vars}.
 
@@
 
-The following shows the SFDA sensitivity analysis results. The largest value is in bold.
+The following shows the SFDA sensitivity analysis results. The largest two values are shown in bold.
 \begin{center}
@@
 
-The following shows the ATA sensitivity analysis results. The largest value is in bold.
+The following shows the ATA sensitivity analysis results. The largest two values are shown in bold.
 \begin{center}
@@
 
+Figure~\ref{fig:sens_surf} shows a surface plot of the sensitivity analysis results.
+From this figure, we can see that the SFDA and ATA achieve their maximal values at different $\alpha$ and $\rho$ parameters, though both achieve a reasonably optimal value at the intermediate parameter values $\alpha = 10$ and $\rho = 0.85$.
+We can also see that detection and tracking performance is fairly robust to minor variations in these parameter values.
+
 % By inspecting the figures, one can see that there is a point where the SFDA and ATA both achieve a maximal value.
@@
 
-\section*{Discussion and Conclusion}
+\section*{Conclusion}
 \label{sec:discussion}
@@
 The GPUDDPM, a time-dependent Dirichlet process mixture, has been introduced, and we have shown how inference on this model allows for the desired clustering of extraction data.
-Furthermore, we have demonstrated a specific implementation of the model using spatial and color pixel data extracted via frame differencing and provided two algorithms for performing inference on the model to accomplish detection and tracking.
+Furthermore, we have demonstrated a specific implementation of the model using spatial and color pixel data extracted via frame differencing and provided two algorithms for performing Bayesian inference on the model to accomplish detection and tracking.
 Both algorithms were carried out on multiple synthetic and benchmark multi-object video datasets in order to demonstrate an ability to accomplish unsupervised detection and tracking of arbitrary objects in both manufactured and real world settings.
-We have described and computed standard performance metrics for our method's detection and tracking results, and found it to be comparable with other techniques that were designed for the specific object-types present in the videos in question.
-These results illustrate the ability of the technique described in this paper to, without modification, detect and track objects with diverse physical charactersitics moving over non-uniform backgrounds and through occlusion
-
-More to include: more discussion of parameter choices, thoughts about deletion params vs.
+We have described and computed standard performance metrics for our technique's detection and tracking results, and found it to be comparable with state-of-the-art object-specific detection and tracking methods designed for people tracking in the PETS2009/2010 video dataset.
+Results from the synthetic and benchmark video datasets illustrate the ability of the technique described in this paper to, without modification, perform completely unsupervised detection and tracking of objects with diverse physical charactersitics moving over non-uniform backgrounds and through occlusion
+
+% More to include: more discussion of parameter choices, thoughts about deletion params vs.
 dataset size (and the tradeoff between tracking through occlusion vs clusters rapidly gaining size), mvn cluster size vs.

### 2012-02-01 ltx/paper_brainstorm.tex
re did format of whole paper. IMPORTANT: the next .tex file of interest is: paper_short.tex

--- 2012-01-12 ltx/paper.tex
@@
 
-\documentclass[smallcondensed, final]{svjour3}
+\documentclass[twocolumn, final]{svjour3}
 % \documentclass[smallcondensed, final]{svjour3}[1/1/2012]
@@
 
-We define the task of automated detection and tracking of arbitrary objects in videos to be the fulfillment of three actions: determining the spatiotemporal regions of a video that constitute objects (`extraction'), finding the positions and/or shapes of distinct objects (`localization'), and maintaining the identities of the detected objects over time (`tracking').
+We define the task of automated detection and tracking of arbitrary objects in videos to \willie{consist of three actions} be the fulfillment of three actions: determining the spatiotemporal regions of a video that constitute objects (`extraction'), finding the positions and/or shapes of distinct objects (`localization'), and maintaining the identities of the detected objects over time (`tracking').
 An ability to carry out these tasks in an automated manner is useful for many fields that make use of video data, including robotics, video surveillance, time-lapse microscopy, and video summarization.
 Algorithms able to be applied to a variety of objects and video types are particularly desirable.
-This paper serves to introduce a new technique involving the use of dependent Dirichlet processes mixture models, which we hope will provide the foundation for a class of unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.
+This paper \willie{introduces a new framework involving the use of dependent Dirichlet process mixture models.
+This framework provides the foundation for a class of unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.}
+% serves to introduce a new technique involving the use of dependent Dirichlet process mixture models, which we hope will provide the foundation for a class of unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.
 

### 2012-02-01 ltx/paper_short.tex
re did format of whole paper. IMPORTANT: the next .tex file of interest is: paper_short.tex

--- 2012-01-11 ltx/paper.tex
@@
 
-\documentclass[smallcondensed, final]{svjour3}
-% \documentclass[smallcondensed, final]{svjour3}[1/1/2012]
+\documentclass[twocolumn, final]{svjour3}
 \usepackage[square,numbers]{natbib}
@@
 
-\title{Unsupervised Detection and Tracking of Multiple Objects with Dependent Dirichlet Process Mixtures}
-
-\titlerunning{Unsupervised Object Detection and Tracking with Dependent Dirichlet Processes}
-
-% \subtitle{}
-
+\title{Unsupervised Detection and Tracking of Arbitrary Objects with Dependent Dirichlet Process Mixtures}
+\titlerunning{Unsupervised Detection and Tracking of Arbitrary Objects with Dependent Dirichlet Process Mixtures}
 \author{Willie Neiswanger$^{*}$ and Frank Wood$^{\dagger}$}
-
 \authorrunning{Willie Neiswanger and Frank Wood}
-
 \institute{ $^{*}$Columbia University, Department of Applied Math and Applied Physics.
@@
 \email{fwood@stat.columbia.edu}}
-
 \date{}  % add submit date when submitted
-
 \maketitle
@@
 \section{Introduction}
+\label{sec:introduction}
 
-We define the task of automated detection and tracking of arbitrary objects in videos to be the fulfillment of three actions: determining the spatiotemporal regions of a video that constitute objects (`extraction'), finding the positions and/or shapes of distinct objects (`localization'), and maintaining the identities of the detected objects over time (`tracking').
+We define the task of automated detection and tracking of arbitrary objects in videos to consist of three actions: determining the spatiotemporal regions of a video that constitute objects (``extraction''), finding the positions and/or shapes of distinct objects (``localization''), and maintaining the identities of the detected objects over time (``tracking'').
 An ability to carry out these tasks in an automated manner is useful for many fields that make use of video data, including robotics, video surveillance, time-lapse microscopy, and video summarization.
 Algorithms able to be applied to a variety of objects and video types are particularly desirable.
-This paper serves to introduce a new technique involving the use of dependent Dirichlet processes mixture models, which we hope will provide the foundation for a class of unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.
+This paper introduces a new framework involving the use of dependent Dirichlet process mixture models.
+This framework provides a foundation for a class of unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.
 
-Research in the past ten years involving the development of methods for general detection and tracking of objects has tended to focus on one of either extraction, localization, or tracking; attempts to integrate all three tasks and produce a system aimed at multiple arbitrary objects and diverse video types was not often a primary focus, or was attempted in an ad-hoc manner to test or apply a more specific technique.
-Research aimed at developing methods to accomplish the three tasks in a cohesive manner has been pursued more recently to produce algorithms that can, in a fully automated way, accomplish unsupervised detection and tracking---or in short, that can determine the spatiotemporal region of a video occupied by each distinct object.
-A widely applied, though unsophisticated, example of general detection and tracking is the technique of blob tracking.
-Blob tracking typically refers to methods that perform extraction and localization of objects at each frame independently---for example, by assuming that pixels representing an object are distinguishable from pixels representing the background via color, motion, or texture information, and segmenting the extracted data at each frame into ``blobs'' to accomplish localization.
-A number of tracking techniques based on finding blobs in consecutive frames of a video with similar locations and appearances have been attempted to maintain the identity of detected blobs over time.
-The main challenges, however, associated with blob tracking involve the difficulty of segmenting neighboring or occluding objects in a given frame \cite{zhao2004tracking} and preventing multiple tracked blobs from collapsing onto and tracking a single object \cite{vermaak_2003}.
-These difficulties stem from the typical blob tracker's need to perform detection via segmentation on individual frames and lack of a principled way to perform tracking of multiple objects while resisting the collapse of independent tracks.
+Research on methods for general detection and tracking of objects has focused on one of either extraction, localization, or tracking.
+Integrating all three tasks in a system for multiple arbitrary objects and diverse video types was not often a primary focus.
+Accomplishing the three tasks in a cohesive manner has been pursued more recently and has produced algorithms that can, in a fully automated way, accomplish unsupervised detection and tracking---or in short, that can determine the spatiotemporal region of a video occupied by each distinct object.
+An example of general detection and tracking is called ``blob tracking'', which typically refers to methods that perform extraction and localization of objects in each frame independently.
+Tracking involves finding blobs with similar locations and appearances over time.
+The main blob tracking challenges are the difficulty of segmenting neighboring or occluding objects in a given frame \cite{zhao2004tracking} and preventing multiple tracked blobs from merging into a single object \cite{vermaak_2003}.
+These difficulties stem from performing detection via segmentation on individual frames.
 
 This paper presents the use of a Bayesian nonparametric mixture model for general, unsupervised multiple object detection and tracking.
-This method is intended to provide a widely applicable way to detect arbitrary objects in videos---particularly in cases where frame-by-frame segmentation is difficult or video quality is low and extraction is noisy---and better maintain the isolation of independent objects during tracking.
-Specifically, this work describes how a Dirichlet process mixture model can accomplish these goals when it is applied to data that may be extracted in a general manner from videos.
-We will demonstrate how to perform this extraction and how this generative model allows for automated detection and tracking.
-The model is first presented in an abstract form, to allow it to be used with arbitrary extraction procedures and object appearance representations that may be desired in future studies.
-We then formulate a specific model for data gained during a basic extraction procedure and describe inference algorithms that allow object localization and tracking to be carried out.
-We demonstrate our implementation on multiple synthetic and benchmark datasets and compute a few standard performance metrics for each to quantify results.
+This method provides both a widely applicable way to detect arbitrary objects in videos---particularly in cases where frame-by-frame segmentation is difficult or video quality is low and extraction is noisy---and a general framework for identifying and isolating objects during tracking.
+Specifically, this work describes how a dependent Dirichlet process mixture model can accomplish these goals.
+The DDP is first reviewed in the abstract.
+This model can be used with arbitrary extraction procedures and object appearance representations that may be desired in future studies.
+We then present our specific model for data obtained via a basic extraction procedure.
+We describe inference algorithms for the DDP, and specifically our specific model, that allow object localization and tracking to be carried out.
+We demonstrate our implementation on multiple synthetic and benchmark datasets.
+Standard performance metrics for each supports our hypothesis that general extraction combined withe sophisticated inference can perform detection and tracking at a level comperable to state-of-the-art, object specific algorithms.
 
-% related work
-% ------------
+% Prior Work
+% ----------
 
-\subsection{Related Work}
-\label{sec:relatedwork}
+\section{Prior Work}
+\label{sec:priorwork}
 
-Related work involves studies related to extraction procedures, localization techniques, single and multiple object tracking methods, and combination methods for joint object detection and tracking.
+Many methods in the fields of image processing, signal processing, and computer vision have been developed to solve aspects of the problem of unsupervised detection and tracking of arbitrary objects.
+These methods might be placed into a few broad categories: those that aim to distinguish the foreground regions of images from the background, segment images into distinct regions, localize specified objects or features in images, track an object over a sequence of images (after its position has been specified in an initial image), track multiple objects over a sequence of images (especially when the objects interact or occlude one another), segment a sequence of images into distinct spatiotemporal regions, and combine the previous methods in some way to create systems capable of both detecting and tracking specified objects, or of discerning which regions of a video constitute distinct, arbitrary objects and tracking these.
 
-Extraction is related to the tasks of foregound segmentation, background modeling, motion detection, and appearance feature extraction.
-In general, these methods aim to extract features from regions of a video frame that are considered objects or are distinguished in some way from the background of a scene.
-Areas of research in this field include attempts to handle situations where the background displays motion or other dynamic characteristics, reduce noise caused by the false extraction of backgrounds, handle abberant image effects present in some videos, and operate accurately on scenes involving illumination changes.
-These procedures can usually be classified as either model-based methods or heuristic-based methods.
-Model-based methods emphasize the detection of shapes of regions containing objects accurately and reducing background noise.
-For example, foreground extraction has been robustly carried out by modeling each pixel over time in order to infer a probability distribution over pixel values; if the pixel takes on a very unlikely value at a future time, it is marked as a foreground pixel at that point \citep{stauffer_1999, elgammal_2000, elgammal_2002}.
-Note that these techniques treat each pixel independently; this has a tendency to cause errors in foreground detection, resulting in fragmentation of foreground images (which is problematic as many of these methods intend to blob the foreground objects after detection in order to perform segmentation and locate objects' centroids).
-Heuristic-based methods perform extracting by looking for temporal changes in pixel value characteristics.
-These methods are often less accurate than model-based methods, but also less computationally expensive.
-Frame differencing and background subtraction are two such methods, and involve comparing pixel values, or groups of pixel values, between two images, to detect a change surpassing some threshold.
-The goal of these techniques is to find locations in a scene that display motion, and to deem these moving areas the foreground of a scene.
-Often, background subtraction refers to comparing an image containing targets with an image of the background without targets or with some model of the background that is learned as the video progresses (in these cases, objects that do not move throughout a scene will be treated as the background), while frame or image differencing refers to comparing pairs of consecutive images in a video.
-Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking schemes \citep{perez_2002}.
-In addition, there exist other ways of extracting movement from an image, such as through techniques involving optical flow (a calculation based on the `generalized gradient model' of an image, which attempts to capture the speed and direction of movement over areas in the image) \citep{horn_1981, bobick_2001}.
-In \citep{black_2000}, a motion detection and object tracking method is developed based on the relative motion of moving objects' boundaries, which are again found through analysis of optical flow.
+% Overview
+% --------
 
-Localization is related to the topics of object detection, position-finding, segmentation, and target recognition.
-These methods are mainly concerned with finding the positions and/or shapes of objects in a given video frame, or with discerning between objects similar in position or appearance.
-There has been a focus in this field on developing techniques for representing an object's appearance, both to allow for localization of objects on a per-frame basis and to allow for consistent localization of a given object over time.
-Research into the unsupervised detection of arbitrary objects (which we define as methods that accomplish both extraction and localization) has included work involving model-based clustering of individual pixels, local image texture representations, or other patterns found in image regions \cite{jain1997object, fei2005bayesian, sivic2005discovering}.
-These works usually aim to automatically segment images into meaningful regions.
-Note that the popular phrase `object detection' is sometimes used to denote methods that are tailored to find the position of specified objects or object types (as opposed to arbitrary object detection, which incorporates both the task of determining what constitutes distinct objects and localizing these objects), or to denote supervised methods, where extraction is peformed implicitely during searches for specified objects and localization procedures.
+\section{Overview of Method}
+\label{sec:overviewofmethod}
 
-Tracking is related to the topics of video tracking, data association, filtering, point matching, and `tracking of non-rigid objects'.
-In general, these tasks aim to maintain the localization of the position and/or shape of an object over time.
-Areas of interest in this field involve developing metrics for judging the similarity between two object representations (for example, potential representations of the same object at adjacent frames in a video), techniques that provide a way to conduct efficient searches over a frame for regions that are similar to a given object, and methods for predicting the future position or appearance of an object.
-An overview of object tracking methods can be found in \cite{yilmaz2006object}.
-Methods for tracking single, non-rigid, objects are numerous.
-Color, for example, has long been used for tracking; these techniques often operate by modeling an object with some color-based appearance model, and using this model to find the object in subsequent frames.
-One classic approach involves modeling the color of a target (in particular, a distribution over the hue-saturation space of a target region) with a Gaussian mixture model (GMM), and choosing subsequent target positions by searching surrounding areas for regions that yield similar GMMs.
-Furthermore, the GMM is often allowed to adapt over time, and slowly change to model changes in lighting or other time-based variations in the target's color \citep{raja_1998, mckenna_1999, jepson_2003}.
-Additionally, color has been successfully applied as the feature in a technique known a the mean-shift procedure (or the kernel-based method) for tracking \citep{comaniciu_2003, perez_2002, nummiaro_2003, lee_2011}.
-This procedure---a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--optimizes the search for the region in a frame which is most similar to a target region in the previous frame.
-It involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges to a number of modes.
-Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors.
-It can be shown that, through this procedure, each data point becomes associated with a local point of high density (dependent upon the underlying weight distribution specified by the kernel) which allows for clustering \citep{cheng_1995}.
-The mean shift algorithm has been implemented successfully to allow for a sort of kernel-based object tracking \citep{comaniciu_2003, comaniciu_1999, comaniciu_2000}.
+Successful methods for detecting and tracking objects in videos are often tailored for specific object types.
+These methods rely on ad-hoc detection criteria that exploit previous knowledge about the appearance or behavior of objects in a video.
+The method described in this paper is designed to track arbitrary objects without using any explicit detection criteria, and serve as a general strategy that can be used, without modification, to perform accurate detection and tracking of diverse objects in a wide range of videos.
+This method begins by performing a data extraction procedure (a consistent procedure for all videos), which can be carried out broadly but yields noisy data; we model this noisy extracted data with a sophisticated type of Bayesian mixture model.
+Inference can then be performed on this model to gain detection and tracking results.
 
-In addition to single object tracking, a great deal of research in this field over the past decade has focused on the development of algorithms to track multiple objects simultaneously.
-There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera) \cite{senior2006appearance, cucchiara2004probabilistic, zhou2003background}, complex object interactions \cite{khan_2004, mckenna2000tracking, dockstader2001multiple}, objects with similar appearances \cite{maccormick1999probabilistic, jepson_2003}, variable (and potentially high) numbers of objects \cite{reilly2010detection}, and objects that enter and exit a field of view at different times \cite{stauffer2003estimating, nedrich2010learning}.
-Multiple single-object trackers running simultaneously have been shown to be ineffective, as distinct single-object trackers will tend to coalesce and track the same object.
-To remedy this problem, methods such as \cite{maccormick1999probabilistic} have developed probabilistic principles for maintaining isolation of object trackers.
-An approach to this problem involving the use of a nonparametric mixture model has also found success in maintaining isolation of distinct objects \cite{vermaak_2003}.
+This paper describes the data extraction procedures (Section~\ref{sec:dataextraction}), the general form of the model (Section~\ref{sec:modeloverview}), the specific distributions used in our implementation of the model (Section~\ref{sec:modelspecification}), and inference procedures used to get tracking results (Section~\ref{sec:inference}).
 
-Over the past decade, there have been attempts to provide general algorithms for the fully unsupervised detection and tracking of arbitrary objects in videos.
-Blob tracking, a basic method attempting to carry out this goal, has been used for many years, and has found success in arenas where objects are easily discerned from the background of a video and where localization and segmentation of distinct objects is possible \cite{francois2004real, isard_2001}.
-Methods involving blob tracking, however, run into problems when faced with videos where detection is difficult, object appearance or orientation varies heavily, and there exists object occlusion \cite{song2005model}.
-To aid the accuracy of these methods, techniques have been developed for performing extraction and segmentation in a joint manner, incorporating statistical methods for maintaining hypotheses of different numbers of detected objects, and introducing some of the tracking methods described previously to track distinct blobs after they have been segmented \cite{collins2003mean, isard_2001}.
-Another family of methods related to the task of unsupervised detection and tracking of video objects goes under the heading of video segmentation algorithms---these methods extend single-frame image segmentation to maintain coherence of image segments over time, and have had some success when used for the explicit purpose of detecting and tracking foreground objects of videos \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}.
-Other attempts to perform unsupervised detection and tracking include methods for clustering short sequences of positions that can be gained by detecting the motion of objects \cite{brostow2006unsupervised, brox2010object}, which aim to return full-length distinct object tracks, and a graph based method that carries out a similar task using spectral clustering \cite{fragkiadaki2011detection}.
-Another approach based on a Gaussian mixture model has also been used to cluster data gained from moving objects \cite{pece_2002}; this method also develops heuristics for the initialization and elimination of new tracks.
-The technique we introduce in this paper falls into this category of methods, as it aims to provide a general unsupervised detection and tracking method for arbitrary objects in videos via clustering.
-Differing from previous work, we use a type of time dependent Bayesian nonparametric mixture model, and show how it can be applied to a variety of extraction data to perform detection and tracking.
+% Frame Diff and Longimg figure
+% -----------------------------
 
-% extra background that I decided to cut (for now)
-% ------------------------------------------------
+\begin{figure*}
+  \centering
+  \subfloat[]{\label{fig:ants_img}\includegraphics[width=0.335\textwidth]{../img/usbh_frame_723.pdf}} 
+  \subfloat[]{\label{fig:ants_img2}\includegraphics[width=0.335\textwidth]{../img/usbh_frame_724.pdf}}
+  \subfloat[]{\label{fig:ants_img_framediff}\includegraphics[width=0.323\textwidth]{../img/usbh_framediff.png}}\\
+  \subfloat[]{\label{fig:traffic2_img}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame1.png}} 
+  \subfloat[]{\label{fig:traffic2_img2}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame2.png}}
+  \subfloat[]{\label{fig:traffic2_img_framediff}\includegraphics[width=0.3273\textwidth]{../img/traffic2_framediff.png}}\\
+  \subfloat[]{\label{fig:pets2009_img}\includegraphics[width=0.333\textwidth]{../img/pets2009_frame1.png}} 
+  \subfloat[]{\label{fig:pets2009_img2}\includegraphics[width=0.333\textwidth]{../img/pets2009_frame2.png}}
+  \subfloat[]{\label{fig:pets2009_img_framediff}\includegraphics[width=0.333\textwidth]{../img/pets2009_framediff.png}}\\
+  %\subfloat[]{\label{fig:longimg}\includegraphics[width=1.05\textwidth]{../img/longimg_6.pdf}} \\
+  \subfloat[]{\label{fig:longimg_all}\includegraphics[width=1\textwidth]{../img/longimg_7_detail.pdf}} 
+  \caption{Pairs of consecutive frames and the results produced by taking the pixel-wise difference between these frames (a - f).
+The final image shows the result of frame differencing over a sequence of images (from the PETS2009 dataset), where color denotes frame number.}
+  \label{fig:img_and_framediff}
+\end{figure*}
 
-% Others have attempted to abstract this work with a non-parametric color modeling approach based on kernel density estimation, which does not assume a specific underlying distribution (such as the Gaussian mixture in the previous case) and instead converges to reasonable distribution that depends on the data \citep{elgammal_2001}.
+% Data Extraction
+% ---------------
 
-% Extraction
-%
+\section{Data Extraction}
+\label{sec:dataextraction}
 
-% The mean-shift procedure (also known as `kernel-based object tracking') attempts to provide a robust way for non-rigid objects to be tracked.
-This method is beneficial because it optimizes the search for the `next' position of an object (i.e.
-the search for the region in a frame which is most similar to a target region in the previous frame).
-This procedure---a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges for each data point.
-Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors.
-It can be shown that, through this procedure, each data point becomes associated with a local point of high density (dependent upon the underlying weight distribution specified by the kernel) which naturally allows for clustering \citep{cheng_1995}.
-The mean shift algorithm has been implemented successfully to allow for a sort of kernel-based object tracking \citep{comaniciu_2003, comaniciu_1999, comaniciu_2000}.
-Given the current position of an object at a given frame, the goal of tracking is often to find a nearby position in the next frame that has the most similar distribution over some common set of features.
-Usually this must be done through an exhaustive search, comparing the similarity of distributions at each nearby position with that at the current position.
-However, if a certain type of `isotropic kernel' (mean-shift kernel) known as the Bhattacharyya coefficient is chosen as a similarity metric between the feature distributions at two positions, it creates a smooth function where gradient descent techniques can be used to quickly converge upon an optimal subsequent position without using an exhaustive search.
-Many papers are concerned with applying this kernel-based object tracking scheme to different sets of features or with different kernels.
-Additionally, the scheme has been attempted with an adaptive kernel whose shape, scale, and orientation is influenced by a target being tracked \citep{wang_2004}, with a kernel adjusted by the estimated centroid of a tracked target \citep{mehmood_2009}, and with a heirarchical version of the mean-shift procedure \citep{dementhon_2002}.
+% General intuition/definition, possible extraction methods, example with frame differencing (cite figure).
+% \\previous 3 data sections:
+% \\a. general def, baseline data, other examples, idea that we are clustering to get results
+% \\b. frame differencing def (+ color counting)
+% \\c. specific form of data (specified frame diff params) used in experiments
+% \\\\new plan: general intuition, baseline data, general def, (idea that we are clustering?), list of extraction examples, frame differencing def, color counting def\\
 
-% Some have tried to extent single object tracking schemes to tracking multiple targets; at the most basic level, this involves initializing multiple targets and running an instance of (single) VOT for each, either simultaneously or in succession \citep{perez_2002}.
-Furthermore, if performed simultaneously, the tracking of each target can be made dependent upon characteristics of other targets (such as their proximity) to resolve errors and improve tracking (such as those caused by the incorrect merging of two targets) \citep{khan_2004, vermaak_2003} .
+We defined the task of extraction to be the unsupervised determination of which spatiotemporal regions of a video constitute objects, and outlined a number of extraction methods in Section~\ref{sec:priorwork}.
+The most basic dataset that can be gained from extraction is comprised of the three-dimensional points $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $x_{1}$ and $x_{2}$ are the two spatial dimensions of a pixel contained in a region found during extraction and $t$ is the frame in which the pixel resides.
+Once this dataset, which we will refer to as a ``baseline extraction dataset'', has been found, additional features may also be extracted from the video.
+Examples of additional features include color information, pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, and texture representations; the main idea is to choose features which are able to be easily extracted or computed, capture variability in the appearance of objects, and are applicable to a wide variety of object types.
 
-% more background that I decided to cut (for now)
-% -----------------------------------------------
+The baseline extraction dataset has the appearance of worm-like clusters that follow the paths of individual objects as they move (as illustrated in Figure~\ref{fig:longimg_all}).
+Modeling this data with a dependent Dirichlet process mixture model (Section~\ref{sec:gpuddpmixture}) allows us to perform inference (Section~\ref{sec:inference}), which carries out clustering of the data.
+By choosing this type of model, we aim to provide a clustering result that isolates each worm-like cluster and infers a sequence of distributions for each, which we use to gain localization and tracking results for each object (Section~\ref{sec:inferencetotrackingresults}).
 
-% % Localization-focused. Clustering methods, segmentation methods, object appearance models and representations (allows for consistent object localization over time).
+For implementation in experiments (Section~\ref{sec:experiments}), we desire an extraction procedure that is as unsophisticated as possible, both to gauge the robustness of this method on potentially noisy extraction data and to ensure that the procedure is applicable to a wide range of videos.
+Consequently, in all experiments we carry out a method called ``frame-differencing'', which locates motion by recording the positions of pixels that have exhibited differences in intensity or value in succesive frames beyond a given threshold.
+Frame differencing is unsophisticated, computationally inexpensive, able to be applied to a wide range of static, single-camera videos (note that videos used in all experiments were stationary; moving-camera videos require extraction methods that provide accurate foreground segmentation for non-stationary video).
+We have found that, when correctly implemented, frame differencing is sufficiently general to extract the desired worm-like clusters from a variety of videos containing multiple moving objects.
+A few examples of this extraction procedure on pairs of consecutive video frames can be found in Figure~\ref{fig:img_and_framediff}(a-i).
 
-% % Methods for tracking single objects in videos: filtering, similarity metrics / localization (and how they play a role), mean shift procedures.
+We furthermore wish to capture the color (or grayscale) value of each pixel and incorporate this color information into our model.
+For each pixel $x = (i, j, t)$ recorded during frame differencing, we specify a square, $L$ pixels in length, centered on $(i, j)$, that selects a set of pixels surrounding $\bold{x}$ in frame $t$.
+We also specify a scalar color value that is able to be computed for each pixel; this value could, for example, be some function of the red-green-blue (r-g-b) or hue-saturation-value (h-s-v) characteristic of a pixel.
+The color value of each of the selected pixels surrounding $\bold{x}$ is recorded.
+Afterwards, the set of possible color values (ie the range of color values to which a pixel may be assigned) is partitioned into $V$ bins, and the number of pixels with a color value lying in each of the bins yields a $V$ dimensional vector of `color counts'.
+We will refer to this extraction technique as `color counting'.
 
-% % extensions to multiple visual object tracking.
-problems with multiple independent single object trackers (coalesence, object interaction, object occlusion, objects entering and leaving scene [and, relatedly, variable numbers of objects]).
-% % Multiple object tracking attempts at overcoming various the various challenges specified above (include nonparametric varmaak/doucet for maintaining multi-modality here?---sure, and then mention again later with fox's work?)
-% % End with description of initialization problem.
+The frame differencing and color counting extraction yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \{1, \ldots, T \}$, where each element $\bold{x} \in \bold{X}$ can be written
+\begin{equation}
+\bold{x} = ( \bold{x}^{s}, \bold{x}^{c}, t ) = ( x^{s_{1}}, x^{s_{2}}, x^{c_{1}}, \ldots, x^{c_{V}}, t )
+\end{equation}
+where $\bold{x}^{s} \in \mathbb{R}^{2}$ is the two dimensional vector of spatial coordinates,  $\bold{x}^{c} \in \mathbb{Z}_{+}^{V}$ is the $V$ discrete dimensional vector of color counts, and $t \in \{1, \ldots, T \}$ is the discrete time index.
 
-% % Methods for detection and tracking of objects.
-Heuristic initialization at beginning and then carrying out above methods (problems with varying numbers of obects / objects entering or leaving scene).
-Supervised learning of specific targets to allow for initialization and provide better localization throughout.
-Various research has been towards integrating unsupervised extraction and localization into multiple object tracking methods.
-Particle filters with integrated detection (at least two of these).
-Introduce concept of blob tracking and define basic algorithm (some manner of extraction, independent/per-frame localization, tracking; note that each object is represented in each frame as a low-dimensional point [or higher dimensional thing like region/outline with sophisticated blob/contour detectors like bramble and others]).
-The key thing characterizing a blob detector is that the position of each object (or the spatially separate region associated with each object) is assigned/fixed at each frame without the aid of tracking (ie if localization/segmentation can be carried out at each frame independently), and tracking is done with this data.
-Schemes often are simply segmentation of extraction data into spatially disjoint blobs at each frame (where blobs  are not modeled in any principled manner), returning the position or region associated with each object at each frame.
+In the experiments described in Section~\ref{sec:experiments}, the hue component of the h-s-v value is recorded from all pixels that surround each extracted pixel $\bold{x}$ in the manner described above, choosing $L=5$.
+Additionally, the set of possible hue values is partitioned into 10 bins, and the number of pixels with a hue value lying in each of the 10 bins is recorded to yield the vector of ``color counts'' $\bold{x}^{c} = x^{c_{1}}, \ldots, x^{c_{10}}$.
+Hue is chosen to represent object color since it has been demonstrated in previous work as a simple representation of object appearance that allows for distinct objects to be well differentiated \cite{perez_2002, raja_1998, mckenna_1999}.
 
-% % When the state of objects (for example, position and/or other characteristics) can be found independently at each frame, well established methods in the field known as Multiple Target Tracking (MTT) can be carried out to perform tracking of multiple objects.
-Two classic algorithms for MTT are known as MHT and JPDA.
-In general, MTT methods are useful for procedures that can represent each object at each frame as a point or region without localization information from this object at a previous frame (which requires the use of tracking).
-MTT in the literature is often quite separate from fields of computer vision involving object detection and tracking research.
-In general, the situation arises as video data without modification is very high dimensional (discussion of difference between MTT and detection methods here?).
+% Model Overview
+% --------------
 
-% % Methods like blob tracking which attempt to find the state of each object independently at each frame have a hard time dealing with occlusion and other complexities (except for sophisticated blob trackers like bramble which isn't really blob tracking).
-Attempts have been made to perform time dependent clustering on extraction data (look at cluster tracking paper for good def), to carry out joint localization and tracking; this has been refered to as cluster tracking.
-Work of Pece, MTT of multiple targets work(?).
-The work presented in this paper most closely resembes the topics listed here.
-Does cluster tracking on whichever extraction data, using a time dependent Dirichlet Process mixture model (to better estimate and maintain the number of targets, as was shown to be effective in vermaak and fox), which takes the form of a sequence of clusters, where the objects' appearances are modeled in a principled manner by the cluster component distributons, and the tracking behavior is modeled by time-dependencies between adjacent elements in the cluster-sequences.
-
-% Contributions (decided to cut for now)
-% --------------------------------------
-
-% \subsection{Contributions}
-
-% This paper applies a time dependent Dirichlet process mixture to model data gained during extraction procedures, provides background as to how this Bayesian nonparametric model allows for the incorporation of prior knowledge about object appearances and motion behaviors, describes methods for unsupervised detection and tracking of arbitrary objects in terms of Bayesian inference procedures, and presents a specific implementation of the model and inference procedures to carry out detection and tracking on a number of target-types in different settings.
-
-% % List here the benefits of this idea that should be discussed/listed:
-% This paper is heavily model-based, which differs from many papers in the field of multiple object tracking that do not explicitely describe models which underlie various procedures that are carried out.
-Solutions in many papers are, instead, algorithm oriented.
-The model-based approach allows rigor to be kept as inference procedures are developed.
-Furthermore, development of inference procedures are performed separately from development of the model, and various well studied inference and approximate inference methods can be carried out to achieve algorithmic results that are (in some cases) known to be optimal.
-% %treated separately, which allows for multiple methods to be carried out in a unified framework and   these solutions are  but provide algorithm-based presentations.
-% %	This way allows rigor to be kept as inference procedures are developed.
-% %	Furthermore, inference is treated separately; this allows multiple detection and tracking algorithms to be developed (in a well-formed/rigorous fashion) by developing different inference techniques.
-Also, inference on graphical models is a well-structured field and allows for use of techniques in this field as well as better interpretations of the resulting detection/tracking algorithms.
-
-% % (related to previous?) models objects as distribution-components (of a mixture model) as opposed to point-targets or targets with noise that adheres to some distribution (is this truly a distinction and is it good?)
-
-% This idea presents a framework which allows for time-varying models of objects' appearances and motion behaviors to be incorporated, and is formulated in a Bayesian manner that allows distributions underlying the parameters of these models to be specified in a principled manner.
-It has been shown to work successfully in conjunction with simple (noisy) extraction methods.
-More robust, expensive, or in general accurate methods can be used for further accuracy.
-This method is able to handle the appearance and disappearance of objects from a scene, variable numbers of objects, cases where objects show time varying changes in behavior, and cases of occlusion.
-
-% Section two of this paper provides an overview of the model and describes how aspects can be modified for for modeling object appearance and motion behavior.
-Section three presents a specific implementation of the model and a number of inference procedures for carrying out object detection and tracking.
-Section four describes a number of experiments carried out with this implementation.
-Section five describes results of these experiments, comments on further directions of study, and provides concluding remarks.
-
-% general model overview section
-% ------------------------------
-
-\section{Model Overview}
+\section{Model Overview and General Form}
 \label{sec:modeloverview}
 
-This section provides details on the data gained during extraction procedures, and on Dirichlet process mixture (DPM) models, dependent Dirichlet process mixture (DDPM) models, and the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM) model.
-We make explicit the way in which characteristics of video objects can be represented by elements of the GPUDDPM.
-We formulate the model in a sufficently abstract manner to allow for implementations in future studies that may use more-specific object representations or extraction procedures.
+Dirichlet Process Mixture (DPM) models (Section~\ref{sec:dpmixture}) fall under the heading of Bayesian nonparametric (BNP) models.
+These models have been widely used in the past decade to perform nonparametric density estimation and cluster analysis.
+Since we have time series data and wish to find spatiotemporal clusters, we are interested in using a closely related class of models known as Dependent Dirichlet Process Mixture (DDPM) models (Section~\ref{sec:ddpmixture}).
+These models are particularly useful for estimating the number of latent classes (clusters) in time dependent data.
+In this work, estimating the number of clusters in video extraction data is equivalent to estimating the number of distinct objects in a video.
 
-% general data extraction section
-% -------------------------------
+This section provides background information that allows us to introduce DDPM models, describes the motivation behind our use of these models, and gives a definition of the particular DDPM chosen, known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM) model (Section~\ref{sec:gpuddpmixture}).
 
-\subsection{Data Extraction}
-\label{sec:data}
-
-We defined the task of extraction to be the unsupervised determination of which spatiotemporal regions of a video constitute objects, and outlined a number of extraction methods in Section~\ref{sec:relatedwork}.
-Extraction allows us to generate a collection of data points.
-Each data point can be thought of as a hypothesis about, or noisy version of, the representation of an object; a collection of hypotheses for a given object is assumed to be distributed in some way around the true object representation.
-A video of multiple objects is therefore assumed to yield data that contains multiple groups of object hypotheses with similar characteristics (i.e.
-clusters of similar hypotheses).
-
-Determining the regions of each video frame that contain objects allows us to generate these object hypotheses; namely, we let each hypothesis be a description of the pixel data over an area centered at a point in the extracted region.
-We can, for example, let each pixel in the extracted region be a hypothesis, and let the features of the representation consist of the spatial position of the pixel.
-An extraction dataset of this form is comprised of the three-dimensional points $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $x_{1}$ and $x_{2}$ are the two spatial dimensions of a pixel contained in an extracted region in a given frame, and $t$ is the discrete time index of the frame.
-These three features are always available once regions containing objects are discerned, and are useful features for modeling (as position and time information is key for reasonable tracking).
-
-Examples of additional features that could be extracted from object-regions in video frames include color information, pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, texture representations, transform calculations (such as results of hough or fourier transforms), and quantiative region descriptions derived from image segmentation procedures; the main idea is to choose features which are able to be easily extracted or computed, capture variability in the appearance of objects, are applicable to a wide variety of object types, and can be represented by distributions that allow for tractable inference (discussed further in Section~\ref{sec:inference}).
-
-We will use the phrase ``baseline'' three-dimensional extraction data to refer to the features $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$ of pixels present in regions containing objects.
-These data are distributed roughly as worm-like shapes that follow the paths of individual objects as they move, when the data are plotted in three dimensions (where we let the frame number $t$, our discrete representation of time, ascend the vertical axis and the spatial positions occupy the horizontal axes).
-
-% this is the pretty worms figure that we decided to cut
-% ------------------------------------------------------
-
-%An example of data with such a distribution can been seen in Figure~\ref{fig:baselinedata}...The data consists of the three-dimensional baseline features and was gained by extraction peformed on a 50 frame sequence showing five ants moving across the video scene.
-
-% frank says the following worm figures are a no go.
-% \begin{figure}[h]
-%   \centering
-%   \subfloat[]{\label{fig:baselinedata1}\includegraphics[width=0.33\textwidth]{../img/pets2001_data_1.png}} 
-%   \subfloat[]{\label{fig:baselinedata2}\includegraphics[width=0.33\textwidth]{../img/usbh_data_3.png}} 
-%   \subfloat[]{\label{fig:baselinedata3}\includegraphics[width=0.33\textwidth]{../img/traffic2_data_1.png}}\\
-%   \subfloat[]{\label{fig:baselinedata4}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_1.png}} 
-%   \subfloat[]{\label{fig:baselinedata5}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_2.png}} 
-%   \subfloat[]{\label{fig:baselinedata6}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_3.png}}
-%   \caption{Baseline extracted features from the (a.) PETS2001 Dataset, (b.) Ants Dataset, (c.) Traffic Dataset, and PETS2009/2010 benchmark dataset, frames (a.) 1-200, (b.) 201-400, (c.) 401-600.
-Time is shown on the vertical axis.
-Marker color corresponds with spatial coordinates and is shown for ease of viewing.}
-%   \label{fig:baselinedata}
-% \end{figure}
-
-Modeling this data with a time dependent Dirichlet process mixture model (Section~\ref{sec:gpuddpm}) allows us to perform inference (Section~\ref{sec:inference}), which carries out clustering of the data.
-By choosing this type of model, we aim to provide a clustering result that isolates each worm-like data cloud and consequently allows each object to be localized and tracked.
-
-% section providing background to introduce DPMs
-% ----------------------------------------------
-
-\subsection{Dirichlet Process Mixtures}
-\label{sec:dpm}
-
-Dirichlet process mixtures fall under the heading of Bayesian nonparametric (BNP) models.
-These models have been widely used over the past decade to perform nonparametric density estimation and cluster analysis.
-They are, in particular, useful for estimating the number of latent classes (clusters) in mixture models.
-In this work, estimating the number of clusters in the data generated via extraction is equivalent to estimating the number of distinct objects in a video.
-Consequentially, a model that is able to perform robust nonparametric estimation of the clusters in data gained during extraction may be applied to perform unsupervised object detection and tracking.
-
-The following sections are intended to provide an introduction to DPMs without a major foray into the theoretical aspects of the model that require a more rigorus technical setup.
-We provide background on finite mixture models, Bayesian finite mixture models, and Dirichlet processes, which allows us to subsequently describe DPMs in context.
-
-% defining the finite mixture model
-% ---------------------------------
-
-\subsubsection{Finite Mixture Model}
-\label{sec:finite_mixture}
+\subsection{Finite Mixture Model}
+\label{sec:finitemixture}
 
@@
 	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
-	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}})
+	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  \text{F}(\theta_{c_{i}})
 \end{split}
@@
 $\forall i \in \{ 1, \ldots, N \}$, where the $x_{i}$ are observations, the $c_{i}$ are the mixture component assignments associated with each observation, the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e.
-the distribution to be mixed, $F(\theta_{c_{i}})$), and the ``Discrete'' distribution refers to a multinomial distribution whose parameters are a 1-of-K vector (i.e.
+the distribution to be mixed, $\text{F}(\theta_{c_{i}})$), and the ``Discrete'' distribution refers to a multinomial distribution whose parameters are a 1-of-K vector (i.e.
 a vector of counts that sums to one).
 
-% defining the Bayesian mixture model
-% -----------------------------------
+\subsection{Bayesian (Finite) Mixture Model}
+\label{sec:bayesianmixture}
 
-\subsubsection{Bayesian (Finite) Mixture Model}
-\label{sec:bayesian_mixture}
-
-The finite mixture model of the previous section can be extended to a Bayesian mixture model by viewing parameters that were previously point values, $\theta_{c_{i}}$ (the mixture component parameters) and $p_{1}, \ldots, p_{K}$ (the mixture component assignment weights), as random variables and providing each with a prior distribution.
+The finite mixture model of Section~\ref{sec:finitemixture} can be extended to a Bayesian mixture model by viewing parameters that were previously point values, $\theta_{c_{i}}$ (the mixture component parameters) and $p_{1}, \ldots, p_{K}$ (the mixture component assignment weights), as random variables and providing each with a prior distribution.
 In this case, the prior distribution $\mathbb{G}_{0}$ is placed on the mixture component parameters, and the prior distribution $\text{Dir}(\alpha/K, \ldots, \alpha/K)$ is placed on the mixture component assignment weights.
@@
 $\forall i \in \{ 1, \ldots, N \}$, where the $x_{i}$ are observations, the $c_{i}$ are the mixture component assignments associated with each observation, the $\theta_{k}$ are parameters defining the $k^{\text{th}}$ mixture component (i.e.
-the distribution to be mixed, $F(\theta_{k})$), the $\theta_{k}$ are drawn from a prior distribution $\mathbb{G}_{0}$, and $p_{1}, \ldots, p_{K}$ are drawn from a Dirichlet prior parameterized by $\alpha/K, \ldots, \alpha/K$.
+the distribution to be mixed, $\text{F}(\theta_{k})$), the $\theta_{k}$ are drawn from a prior distribution $\mathbb{G}_{0}$, and $p_{1}, \ldots, p_{K}$ are drawn from a Dirichlet prior parameterized by $\alpha/K, \ldots, \alpha/K$.
 
-% defining Dirichlet processes
-% ----------------------------
-
-\subsubsection{Dirichlet Process}
-\label{sec:dirichlet_process}
-
-% Define Dirichlet distribution first?
-Each draw from a K-dimensional Dirichlet lies on the K-simplex (i.e.
-is a vector of K values which sum to 1) and can therefore be thought of as a K-dimensional discrete distribution.
+\subsection{Dirichlet Process}
+\label{sec:dirichletprocess}
 
@@
 
-% defining the DPM
-% ----------------
+\subsection{Dirichlet Process (Infinite) Mixture Model}
+\label{sec:dpmixture}
 
-\subsubsection{Dirichlet Process (Infinite) Mixture Model}
-\label{sec:dirichlet_process_mixture}
-
-A Dirichlet process mixture model, also refered to as an infinite mixture model, is an extension of the Bayesian mixture model described previously. 
+A DPM model, also refered to as an infinite mixture model, is an extension of the Bayesian mixture model described in Section~\ref{sec:bayesianmixture}.
 When using a DP as a prior in a Bayesian mixture model, $\Theta$ represents the set of parameters of the component mixture distributions.
@@
 	\phi_{i} | \mathbb{G}  &\sim  \mathbb{G} \\
-	x_{i}|\phi_{i} &\sim F(\phi_{i})
+	x_{i}|\phi_{i} &\sim \text{F}(\phi_{i})
 \end{split}
@@
 $\forall i \in \{ 1, \ldots, N \}$, where the $x_{i}$ are observations, the $\phi_{i}$ are parameters defining the mixture component from which the $i_{th}$ observation is drawn (i.e.
-the distribution to be mixed, $F(\phi_{i})$), and the $\phi_{i}$ are drawn from a prior distribution $\mathbb{G}$, which is in turn drawn from a DP with base distribution $\mathbb{G}_{0}$ and parameter $\alpha$.
+the distribution to be mixed, $\text{F}(\phi_{i})$), and the $\phi_{i}$ are drawn from a prior distribution $\mathbb{G}$, which is in turn drawn from a DP with base distribution $\mathbb{G}_{0}$ and parameter $\alpha$.
 See \cite{gasthaus_2008} and \cite{gasthaus_thesis} for more details on this formulation.
 Note the difference between the indexing of the clusters in this model and the indexing in the previous two models.
-This formulation can be shown to be equivalent to the Bayesian mixture model defined in \eqref{bayesian_mixture_model}, when K is taken to be unbounded; as a result, this model is sometimes called an infinite mixture model.
+This formulation can be shown to be equivalent to the Bayesian mixture model defined in \eqref{bayesian_mixture_model}, when $K$ is taken to be unbounded; as a result, this model is sometimes called an infinite mixture model.
+If we let $K$ be the number of distinct mixture components assigned to observations using the above model, we can write the mixture components as $\theta_{1}, \ldots, \theta_{K}$.
+We also let $c_{1}, \ldots, c_{N}$ (where $c_{i} \in \{1, \ldots, K \}$) be class assignment variables that indicate the cluster to which observation $x_{i}$ is assigned.
 
-If we let $K$ be the number of distinct mixture components assigned to observations using the above model, we can write the mixture components as $\theta_{1}, \ldots, \theta_{K}$.
-Let $c_{1}, \ldots, c_{N}$ (where $c_{i} \in \{1, \ldots, K \}$) be class assignment variables that indicate the cluster to which observation $x_{i}$ is assigned.
-We can now formulate the Chinese restaurant process, a discrete-time stochastic process that defines a partition of the set $\{ x_{1}, \ldots, x_{N} \}$ (via the elements' assignments $c_{1}, \ldots, c_{N}$).
-The Chinese restaurant process allows samples to be drawn from the conditional distribution over the assignment variables $c_{i}$, and can be expressed as
-\begin{align}
-\begin{split}
-\label{crp_rep}
-	P(c_{i} = c_{j} \text{  for some  } j<i) &= \frac{m_{k}}{i-1+\alpha}\\
-	P(c_{i} \neq c_{j} \text{  for all  } j<i) &= \frac{\alpha}{i-1+\alpha}
-\end{split}
-\end{align}
-where $m_{k}$ is the cardinality of the set $\{ c_{j} | (c_{j}=c_{i}=k)  \wedge  (j < i) \}$ and $\alpha$ is the parameter of the DP prior on $\mathbb{G}$.
-A modified Chinese restaurant process will be used to draw samples from the conditional distribution over the assignment variables when performing inference on the GPUDDPM (Section~\ref{sec:gpuddpm}).
+\subsection{Dependent Dirichlet Process Mixture Model}
+\label{sec:ddpmixture}
 
-% section providing background to introduce DDPMs
-% -----------------------------------------------
+The goal of DDPM models is to allow modeling of data that is not independent and identically distributed but instead has some underlying dependencies.
+For example, data generated during video extraction procedures have some associated temporal dependencies, since there exist similarities between features (such as those that encode the spatial positions or appearances of objects) of data at nearby time steps.
 
-\subsection{Dependent Dirichlet Process Mixtures}
-\label{sec:ddpm}
-
-The goal of dependent Dirichlet process mixtures is to allow modeling of data that is not independent and identically distributed (i.i.d) but instead has some underlying dependencies.
-For example, data generated during extraction procedures from videos have some associated temporal dependencies, since tracked objects display time dependent characteristics (i.e.
-there are similarities among features, such as the spatial positions or appearances of objects, for data at nearby time steps).
-
-To account for the dependent behavior of data, there has been research on developing models involving a sequence of DPMs, where components of the mixtures are dependent upon (or are sometimes said to be ``tied to'') corresponding components at neighboring positions in the sequence.
+To account for the dependent behavior of data, there has been research into models involving a sequence of DPMs, where components of the mixtures are dependent upon (or are sometimes said to be ``tied to'') corresponding components at neighboring positions in the sequence.
 For example, if the data shows temporal dependence, the goal might be to create a sequence of DPMs, one for each time-step, where the components of the mixture at each step are dependent upon corresponding components in the both the following and previous time steps.
@@
 
-%%%%%
-% add more info here on other DDPMs, why we chose the GPUDDPM, how it has an ability to model birth and death of clusters (unlike some sort of sophisticated hidden markov model)
-%%%%%
+\subsection{Generalized Polya Urn Dependent Dirichlet Process Mixture Model}
+\label{sec:gpuddpmixture}
 
-% defining the GPUDDPM
-% --------------------
-
-\subsubsection{Generalized Polya Urn Dependent Dirichlet Process Mixture}
-\label{sec:gpuddpm}
-
-Data gained by performing extraction on videos containing objects is significantly time-dependent.
-Futhermore, the clusters of data, each representing a video object, might change in number as time progresses (which is to say, clusters may get `born' and may `die' at some intermediate time step), since objects can enter and exit a scene.
+The time dependent video extraction data contains a collection of clusters, each representing a video object, which might change in number as time progresses (which is to say, clusters may be created, or be ``born'', and may disappear, or ``die'', at some intermediate time step), since objects can enter and exit a scene.
 To handle these challenges, the specific DDPM chosen to model extraction data in this work is known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM), introduced by \cite{caron_2007} in 2006.
 
-The GPUDDPM, when applied to data over T discrete time steps, can be viewed as a sequence of DPMs (one for each $t \in \{1, \ldots, T \}$), which are linked together by dependencies between cluster parameters in neighboring time steps.
+The GPUDDPM, when applied to data extracted from a video with T frames, can be viewed as a sequence of DPMs (one for each ``time-step'' $t \in \{1, \ldots, T \}$), which are linked together by dependencies between cluster parameters in neighboring time steps.
 More specifically, the parameters at a given time step are distributed as a function of the parameters in the previous time step, and the distribution and number of distinct cluster assignments at a given time step are distributed according to all previous cluster assignments and a deletion procedure, described below.
@@
 A transition kernel $P(\theta_{k,t} | \theta_{k,t-1})$ specifies how mixture component parameters in time step $t$ are dependent upon associated mixture component parameters in time step $t-1$.
-One caveat is that each mixture component must be drawn independently from $\mathbb{G}_{0}$ (the base distribution of the DP, which acts as a prior distribution for the cluster parameters) which we achieve by making $\mathbb{G}_{0}$ the invariant distribution of the transition kernel $P(\theta_{k,t} | \theta_{k,t-1})$ (which, one should note, is also a markov chain).
+One caveat is that each mixture component must be drawn independently from $\mathbb{G}_{0}$ (the base distribution of the DP, which acts as a prior distribution for the cluster parameters) which we achieve by making $\mathbb{G}_{0}$ the invariant distribution of the transition kernel $P(\theta_{k,t} | \theta_{k,t-1})$ (which, one should note, is a markov chain).
 
-Recall that the distribution over the cluster assignments $c_{i}$ (one for each observation $i$) is a function of the cluster sizes (as per \eqref{crp_rep}).
-Hence, to account for varying numbers of clusters---and in particular, to allow clusters to diminish in size (i.e.
-to diminish in number of assigned observations) and even ``die off"---there is a deletion procedure that allows observations to be considered removed from their assigned clusters at a given time step.
+Recall that each observation $i$ is given a cluster assignments $c_{i}$ (Section~\ref{sec:dpmixture}).
+Cluster size $m_{k}$ denotes the number of observations assigned to a given cluster $k$; in a DPM, cluster size can be found by counting the number of these assigned observations.
+In contrast, to account for varying numbers of clusters at different points in time---and in particular, to allow clusters to diminish in size (i.e.
+to diminish in number of assigned observations) and even die off---the GPUDDPM has a deletion procedure that allows observations to be considered unassigned (or ``removed'') from their assigned clusters at a later time step.
 This procedure is introduced and described in detail in \cite{caron_2007}.
@@
 Specifically, each remaining assignment is removed from its cluster with probability $\rho$.
-It can be shown that performing deletion in this way is equivalent to, for each cluster $k$, drawing $r \sim \text{Binomial}(m_{k,t-1}, \rho)$, and reducing the previous-time cluster size $m_{k,t-1}$ by $r$.
+It can be shown that performing deletion in this way is equivalent to, for each cluster $k$, drawing $r \sim \text{Binomial}(m_{k,t-1}, \rho)$, and reducing the previous-time cluster size $m_{k,t-1}$ by $r$ to find the next cluster size $m_{k,t}$.
 At each time step, this deletion procedure is carried out on all existing assignments, including those at the current time.
 Hence, the size $m_{k,t}$ of a given cluster $k$ at time $t$ is dependent upon the cluster size at the previous time step, $m_{k,t-1}$, and on the assignments $c_{i}$ for all observations $i \in \{ 1, \ldots, N_{t} \}$ at time $t$.
-We refer to the conditional distribution over $m_{k,t} | m_{k,t-1}, c_{i}, \rho$ as DEL, which we define to be
+We refer to the conditional distribution over $m_{k,t} | m_{k,t-1}, c_{i}, \rho$ as Del, which we define to be
 \begin{equation}
@@
 \label{del_step}
-\text{DEL} & (\cdot | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) :=
-\left( m_{k,t-1} - \text{Binomial}( \cdot | {m_{k,t-1}, \hspace{1mm} \rho}) \right) + \sum_{i=1}^{N_{t}} \mathbb{I}(c_{i,t} = k)
+\text{Del} &(\cdot | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) := m_{k,t-1} \\
+&- \text{Binomial}( \cdot | {m_{k,t-1}, \hspace{1mm} \rho}) + \sum_{i=1}^{N_{t}} \mathbb{I}(c_{i,t} = k)
 \end{split}
@@
 
-The Chinese restaurant process, defined in \eqref{crp_rep}, is used to sample assignment variables during inference procedures (described in Section~\ref{sec:inference}).
-Based on this definition, we define the distribution CRP to be
+We also define a distribution, which we refer to as C, that is useful during inference procedures (Section~\ref{sec:inference}) for inferring assigments of observations. We define C to be
 \begin{equation}
-\label{CRP}
+\label{C}
 \begin{split}
-\text{CRP} & (\cdot | m_{1:K_{t},t}, \hspace{1mm} \alpha) :=
-\text{Discrete}(\cdot | P(c_{i,t}=1 | m_{1:K_{t},t}), \ldots, P(c_{i,t}=K_{t+1}  | m_{1:K_{t},t}))
+\text{C} &(\cdot | m_{1:K_{t},t}, \hspace{1mm} \alpha) := \text{Discrete}(\cdot | P(c_{i,t}=1 | m_{1:K_{t},t}), \\
+&  \ldots, P(c_{i,t}=K_{t+1}  | m_{1:K_{t},t}))
 \end{split}
@@
 \begin{equation}
-P(c_{i,t}=k  | m_{1:K_{t},t}) = 
-\begin{cases}
-\frac{m_{k,t}}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{2mm} \text{if} \hspace{2mm} k \in \{ 1, \ldots, K_{t} \} \\
-\frac{\alpha}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{2mm} \text{if} \hspace{2mm} k = K_{t} + 1
+\begin{split}
+P (c_{i,t}=k  | m_{1:K_{t},t})
+= \begin{cases}
+\frac{m_{k,t}}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{1mm} \text{if} \hspace{1mm} k \in \{ 1, \ldots, K_{t} \} \\
+\frac{\alpha}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{1mm} \text{if} \hspace{1mm} k = K_{t} + 1
 \end{cases}
+\end{split}
 \end{equation}
 $\forall i \in \{1, \ldots, N_{t} \}$, where there exists $K_{t}$ clusters at time $t$, and we give a newly created cluster the index $K_{t+1}$.
-Note the similarity between the above definition and the Chinese resturant process of the DPM given by \eqref{crp_rep}.
-The deletion and Chinese resturant procedures together comprise what \cite{caron_2007} refers to as the ``Generalized Polya Urn''.
-Using the DEL and CRP distributions given above, we can define the GPUDDPM generatively as, for each time step $t \in \{1, \ldots, T\}$, and each cluster $k \in \{ 1, \ldots, K_{t} \}$ at time $t$,  % and observation $i \in \{ 1, \ldots, N_{t} \}$ at $t$,
+Distributions \eqref{del_step} and \eqref{C} together comprise what \cite{caron_2007} refers to as the ``Generalized Polya Urn''.
+Using the Del and C distributions given above, we can define the GPUDDPM generatively as, for each time step $t \in \{1, \ldots, T\}$, and each cluster $k \in \{ 1, \ldots, K_{t} \}$ at time $t$,
 \begin{align}
@@
 \begin{split}
-m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho  &\sim \text{DEL}(m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) \\
+m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho  &\sim \text{Del}(m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) \\
 \theta_{k, t} | \theta_{k, t-1}   &\sim
 \begin{cases}
-	P(\theta_{k, t} | \theta_{k, t-1}) \hspace{2mm} \text{if} \hspace{2mm} k \in \{ 1, \ldots, K_{t} \} \\
+	P(\theta_{k, t} | \theta_{k, t-1}) \hspace{1mm} \text{if} \hspace{1mm} k \in \{ 1, \ldots, K_{t} \} \\
 	\mathbb{G}_{0}   \hspace{2mm} \text{if} \hspace{2mm} k = K_{t+1}
 \end{cases} \\
-c_{i,t} | m_{1:K_{t},t}, \hspace{1mm} \alpha  &\sim  \text{CRP} (m_{1:K_{t},t}, \hspace{1mm} \alpha) \\
-\bold{x}_{i,t} | c_{i,t}, \theta_{1:K_{t}, t} &\sim F(\theta_{c_{i,t}, t})
+c_{i,t} | m_{1:K_{t},t}, \hspace{1mm} \alpha  &\sim  \text{C} (m_{1:K_{t},t}, \hspace{1mm} \alpha) \\
+\bold{x}_{i,t} | c_{i,t}, \theta_{1:K_{t}, t} &\sim \text{F}(\theta_{c_{i,t}, t})
 \end{split}
 \end{align}
+where we specify distributions for F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ in Section~\ref{sec:modelspecification}.
 The graphical model corresponding with this formulation is shown in Figure \ref{fig:gpuddpm_gm_1}.
 \begin{figure}[h]
-        \center{\includegraphics[width=90mm]{../img/gpuddp_gm_1.pdf}}
+        \center{\includegraphics[width=80mm]{../img/gpuddp_gm_1.pdf}}
         \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Polya Urn Dependent Dirichlet Process Mixture.
@@
 \end{figure}
-% \begin{align}
-% \begin{split}
-% \bold{m}_{t} | \bold{m}_{t-1}, \rho  &\sim \text{DEL}(\bold{m}_{t-1}, \rho) \\
-% c_{i} | \bold{m}_{t}, \alpha  &\sim  \text{CRP}(\bold{m}_{t}, \alpha) \\
-% \theta_{c_{i}, t} | \theta_{c_{i}, t-1}   &\sim
-% \begin{cases}
-% 	P(\theta_{c_{i}, t} | \theta_{c_{i}, t-1}) \\
-% 	\mathbb{G}_{0}
-% \end{cases} \\
-% \bold{x}_{i} | c_{i}, \theta_{c_{i}, t} &\sim F(\theta_{c_{i}, t})
-% \end{split}
-% \end{align}
-% where $\text{CRP}(\cdot, \alpha)$ is given in \eqref{crp_rep}.
-The graphical model corresponding with this formulation is shown in Figure \ref{fig:gpuddpm_gm_1}.
-The above formulation holds if there is exactly one observation at each time step.
-Often, the data extracted from objects in videos consists of multiple observations at each time step, each with an assignment variable ($c_{i}$ for $i \in \{ 1, \ldots, N_{t} \}$, where $N_{t}$ is the number of observations at time step $t$).
-If we allow for multiple observations $i \in \{ 1, \ldots, N_{t} \}$ at each time $t$, this formulation becomes,
-% \begin{align}
-% \begin{split}
-% \bold{m}_{t}^{1} | \bold{m}_{t-1}^{N_{t-1}}, \rho  &\sim \text{DEL}(\bold{m}_{t-1}^{N_{t-1}}, \rho) \\
-% c_{i} | \bold{m}_{t}^{i-1}, \alpha  &\sim  \text{CRP}(\bold{m}_{t}^{i-1}, \alpha) \\
-% \theta_{c_{i}, t} | \theta_{c_{i}, t-1}   &\sim
-% \begin{cases}
-% 	P(\theta_{c_{i}, t} | \theta_{c_{i}, t-1}) \\
-% 	\mathbb{G}_{0}
-% \end{cases} \\
-% \bold{x}_{i} | c_{i}, \theta_{c_{i}, t} &\sim F(\theta_{c_{i}, t})
-% \end{split}
-% \end{align}
-% where $\bold{m}_{t}^{i}$ represents the $i^{\text{th}}$ observation at time $t$.
 
-% section making explicit connection between model and video tracking
-% -------------------------------------------------------------------
+\subsection{Deletion Variable Formulation of the GPUDDPM}
 
-\subsubsection{Object Characteristics Modeled}
+To allow for modeling time-dependent data, the GPUDDPM allows clusters to ``lose'' assigned observations in a deletion step, which modifies the values of the cluster sizes at each time.
+Instead of incorporating the cluster size variable $m$ directly in the model (as it was in the definition of the GPUDDPM given by \eqref{gpuddpm_def}), we can formulate an equivalent model which makes use of new set of variables called the deletion variables (denoted as $d$), which represent the times at which assignments are deleted.
+More precisely, we introduce variables $d_{1}, \ldots, d_{N}$ to the model, which denote the times at which the assignments of observations $1, \ldots, N$ are considered removed from their assigned cluster.
+At each time step, cluster sizes can be recontructed from all previous assignment and deletion variables.
+With these deletion variables, we can compute the size of cluster $k$ at time $t$, $m_{k,t}$, by
+\begin{equation}
+\label{compute_clust_size}
+m_{k,t} = \sum_{t' = 1}^{t} \mathbb{I}[(c_{t'}=k) \wedge (t < d_{t'})]
+\end{equation}
+where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise.
+Additionally, for an observation $\bold{x}_{i}$ at a given time-step $t$, the deletion time $d_{i}$ can be defined to be $d_{i} = t + l_{i}$, where $l_{i}$ is considered the lifespan of an assignment, is distributed geometrically, and can be expressed as
+\begin{equation}
+\label{del_rho_form}
+l_{i} | \rho  \sim  \rho(1 - \rho)^{l_{i}}
+\end{equation}
+This alternate, deletion variable formulation of the GPUDDPM is useful for the MCMC inference algorithm described in Section~\ref{sec:MCMC}.
+We can write the new formulation of the GPUDDPM as, for each time step $t \in \{1, \ldots, T\}$ and clusters $k \in \{ 1, \ldots, K_{t} \} $ at time $t$,
+\begin{align}
+\begin{split}
+d_{i,t} | \rho  &\sim \text{Geo}(\rho) + t + 1 \\
+\theta_{k, t} | \theta_{k, t-1}  &\sim
+\begin{cases}
+P(\theta_{k,t} | \theta_{k,t-1}) \hspace{2mm} \text{if} \hspace{2mm} k \in \{ 1, \ldots, K_{t} \} \\
+\mathbb{G}_{0} \hspace{2mm} \text{if} \hspace{2mm} k = K_{t} + 1
+\end{cases} \\
+c_{i,t} | \bold{c}_{1:t-1}, \bold{d}_{1:t-1}, \alpha  &\sim  \text{C}(\bold{c}_{1:t-1}, \bold{d}_{1:t-1}, \alpha) \\
+\bold{x}_{i,t} | c_{i,t}, \theta_{c_{i,t},t}  &\sim  \text{F}(\theta_{c_{i,t}, t})
+\end{split}
+\end{align}
+where $\bold{c}_{t} = c_{1,t}, \ldots, c_{N_{t}, t}$ (for $N_{t}$ observations at time $t$), $\bold{d}_{t} = d_{1,t}, \ldots, d_{N_{t}, t}$, and we specify distributions for F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ in Section~\ref{sec:modelspecification}.
+This formulation of the GPUDDPM is also used by \cite{gasthaus_thesis} and \cite{caron_2007}.
+The associated graphical model for this formulation is given in Figure~\ref{fig:gpuddpm_gm_2}.
+\begin{figure}[h]
+        \center{\includegraphics[width=82mm]{../img/gpuddp_gm_3.pdf}}
+        \caption{Graphical model of the deletion variable and auxiliary variable transition kernel representation of the GPUDDPM}
+        \label{fig:gpuddpm_gm_2}
+\end{figure}
 
-We would like make explicit the connection between certain elements of the GPUDDPM model and the characteristics of objects in vidoes that these elements represent.
-
-First, the appearances of objects are moded by what we term the `likelihood and object appearance model', which we denoted by $F$.
-In the above modeling formulation, $F$ is the probability distribution for an observation, given that it is assigned to specified cluster (i.e.
-given that it is associated with a given object in a video).
-The likelihood and object appearance model is therefore
-\begin{equation}
-F(\theta_{c_{i}}) = P(x_{i}|\theta_{c_{i}})
-\end{equation}
-where $x_{i}$ is an observation, $c_{i}$ is its associated assignment, $\theta_{c_{i}}$ specifies the parameters of the parametric form of the $c_{i}^{th}$ cluster.
-The specific form of this model is dependent upon the observations $x_{i}$; recall that the observations, which as a baseline include spatial and temporal features, could include an arbitrary amount of additional features which need to be incorporated into this appearance model.
-A specific formulation of the appearance model can be seen in the proceeding section.
-
-Object behavior is modeled by the dependencies between corresponding clusters at adjacent time steps.
-In particular, this relationship is captured by the transition kernel, $P(\theta_{k}^{t} | \theta_{k}^{t-1})$ (which captures the dependence of cluster $k$ at time $t$ on the same cluster at time $t-1$), described previously.
-The motion behavior of an object, i.e.
-the specific dependence of the parameters of cluster $k$ at a time $t$ on its parameters at time $t-1$, is therefore
-\begin{equation}
-Tr(\theta_{k, t} | \theta_{k, t-1}) = P(\theta_{k, t} | \theta_{k, t-1})
-\end{equation}
-
-Additionally, the base distribution $\mathbb{G}_{0}$ of the DP acts as a prior on the cluster parameters.
-This allows one to place a prior probability over appearances of objects.
-It is important that this prior is not chosen in such a specific manner that it limits this method from performing unsupervised detection of arbitrary objects.
-
-% section providing specific model
-% --------------------------------
+% Model Specification
+% -------------------
 
 \section{Model Specification}
-\label{sec:modelspec}
+\label{sec:modelspecification}
 
-In the following sections, a particular extraction method for generating data from videos and a specific formulation of the general model described in Section~\ref{sec:modeloverview} are given.
-Inference schemes on this model, which when carried out allow for unsupervised detection and tracking of multiple arbitrary objects, are also described.
+Sections~\ref{sec:objectappearance}-\ref{sec:motionmodel} describe the distributions chosen to fully specify the general GPUDDPM model form given in Section~\ref{sec:modeloverview}.
+The implementation given below was used for all the experiments of Section~\ref{sec:experiments}.
 
-% gives specific data extraction method (frame differencing)
-% ----------------------------------------------------------
+\subsection{Object Appearance (Mixture Component)}
+\label{sec:objectappearance}
 
-\subsection{Data Extraction via Frame Differencing}
-\label{sec:modelspec_extraction}
-
-We desire an extraction procedure that is as unsophisticated as possible, both to gauge the robustness of this method on potentially noisy extraction data and to ensure that the procedure is applicable to a wide range of videos.
-For these reasons we choose the method of frame-differencing for all extraction performed for experiments in this paper (Section~\ref{sec:experiments}), which involves recording the positions of pixels that have exhibited differences in intensity or value in succesive frames beyond a given threshold.
-In particular, if we let $I_{t}$ be the image difference obtained by subtracting the value of the $(i,j)^{th}$ pixel in video frame $t$ from the value of the $(i,j)^{th}$ pixel in video frame $t+1$, we can define extraction on frame $t$ to be the process that returns the dataset
-\begin{equation}
-	\Omega_{t} = \{ (i,j,t) \hspace{2mm} | \hspace{2mm} I_{t}(i,j) > \phi \}
-\end{equation}
-where $i$ and $j$ respectively denote the two spatial positions of a pixel, and $\phi$ is a given pixel value threshold. Frame differencing on each frame $t =\{1, \ldots, T \}$ yields the dataset
-\begin{equation}
-	\Omega = \bigcup_{t=1}^{T} \Omega_{t} = \{ (i,j,t) \hspace{2mm} | \hspace{2mm} I_{t}(i,j) > \phi,  \hspace{2mm}  \forall t \in \{1, \ldots, T \} \}
-\end{equation}
-which is equivalent to data with the baseline features described in Section~\ref{sec:data}.
-
-Frame differencing is unsophisticated, computationally inexpensive, able to be applied to a wide range of static, single-camera videos (note that videos used in the experiments described in the following sections were chosen to be static; moving-camera videos should be used in conjunction with applicable extraction methods that provide accurate foreground segmentation for non-stationary video).
-We have found that, when correctly implemented, frame differencing is sufficiently general to extract the desired worm-like data clouds from a variety of videos containing multiple moving objects.
-A few examples of this extraction procedure on pairs of consecutive video frames can be found in Figure \ref{fig:img_and_framediff}.
-
-% \begin{figure}
-% \centering{
-% \includegraphics[width=45mm]{../img/antpic2.png}}
-% %\hspace{2mm}
-% {\includegraphics[width=60mm]{../img/frame_diff_ants_1.png}}
-% \caption{(a.) Frame of a video containing five ants.
-(b.) Frame differencing extraction results from the same frame \willie{this is just a demo of the type of image i will have here...
-these aren't actually from the same frame, and also formatting is screwed up)}}
-% \label{test}
-% \end{figure}
-
-\begin{figure}
-  \centering
-  \subfloat[]{\label{fig:ants_img}\includegraphics[width=0.335\textwidth]{../img/usbh_frame_723.pdf}} 
-  \subfloat[]{\label{fig:ants_img2}\includegraphics[width=0.335\textwidth]{../img/usbh_frame_724.pdf}}
-  \subfloat[]{\label{fig:ants_img_framediff}\includegraphics[width=0.323\textwidth]{../img/usbh_framediff.png}}\\
-  \subfloat[]{\label{fig:traffic2_img}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame1.png}} 
-  \subfloat[]{\label{fig:traffic2_img2}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame2.png}}
-  \subfloat[]{\label{fig:traffic2_img_framediff}\includegraphics[width=0.3273\textwidth]{../img/traffic2_framediff.png}}\\
-  \subfloat[]{\label{fig:pets2009_img}\includegraphics[width=0.333\textwidth]{../img/pets2009_frame1.png}} 
-  \subfloat[]{\label{fig:pets2009_img2}\includegraphics[width=0.333\textwidth]{../img/pets2009_frame2.png}}
-  \subfloat[]{\label{fig:pets2009_img_framediff}\includegraphics[width=0.333\textwidth]{../img/pets2009_framediff.png}}\\
-  %\subfloat[]{\label{fig:longimg}\includegraphics[width=1.05\textwidth]{../img/longimg_6.pdf}} \\
-  \subfloat[]{\label{fig:longimg_all}\includegraphics[width=1\textwidth]{../img/longimg_7_detail.pdf}} 
-  \caption{Pairs of consecutive frames and the results produced by taking the pixel-wise difference between these frames (a - f).
-The final image shows the result of frame differencing over a sequence of images (from the PETS2009 dataset), where color denotes frame number.}
-  \label{fig:img_and_framediff}
-\end{figure}
-
-We furthermore wish to capture the color (or grayscale) value of each pixel and incorporate this color information into our model.
-For each pixel $x = (i, j, t) \in \Omega$ (defined above), we specify a square, $L$ pixels in length, centered on $(i, j)$, that selects a set of pixels surrounding $\bold{x}$ in frame $t$.
-We also specify a scalar color value that is able to be computed for each pixel; this value could, for example, be some function of the red-green-blue (r-g-b) or hue-saturation-value (h-s-v) characteristic of a pixel.
-The color value of each of the selected pixels surrounding $\bold{x}$ is recorded.
-Afterwards, the set of possible color values (ie the range of color values to which a pixel may be assigned) is partitioned into $V$ bins, and the number of pixels with a color value lying in each of the bins yields a $V$ dimensional vector of `color counts'.
-We will refer to this extraction technique as `color counting'.
-
-% perhaps: Show example color-count vector (histogram) for different objects in an image. give more-formal definition like for frame differencing above.
-
-% introduce specific model for experiments
-% ----------------------------------------
-
-\subsection{Model Implementation in Experiments}
-
-The following sections provide details on the data and specific model formulation (determined from the general model formulation of Section~\ref{sec:modeloverview}) involved in the experiments described in Section~\ref{sec:experiments}.
-
-% data in experiments
-% -------------------
-
-\subsubsection{Data}
-
-The frame differencing and color counting extraction outlined in Section~\ref{sec:modelspec_extraction} yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \{1, \ldots, T \}$, where each element $\bold{x} \in \bold{X}$ can be written
-\begin{equation}
-\bold{x} = ( \bold{x}^{s}, \bold{x}^{c}, t ) = ( x^{s_{1}}, x^{s_{2}}, x^{c_{1}}, \ldots, x^{c_{V}}, t )
-\end{equation}
-where $\bold{x}^{s} \in \mathbb{R}^{2}$ is the two dimensional vector of spatial coordinates,  $\bold{x}^{c} \in \mathbb{Z}_{+}^{V}$ is the $V$ discrete dimensional vector of color counts, and $t \in \{1, \ldots, T \}$ is the discrete time index.
-
-In the following experiments, the hue component of the h-s-v value is recorded from all pixels that surround each extracted pixel $\bold{x}$ in the manner described in Section~\ref{sec:modelspec_extraction} (where we choose $L=5$).
-Additionally, the set of possible hue values is partitioned into 10 bins, and the number of pixels with a hue value lying in each of the 10 bins is recorded to yield the vector of ``color counts'' $\bold{x}^{c} = x^{c_{1}}, \ldots, x^{c_{10}}$.
-Hue is chosen to represent object color since it has been demonstrated in previous work as a simple representation of object appearance that allows for distinct objects to be well differentiated \cite{perez_2002, raja_1998, mckenna_1999}.
-
-% likelihood / object appearance model
-% ------------------------------------
-
-\subsubsection{Likelihood and Object Appearance Model}
-
-At a given time $t$, we model each $\bold{x} \in \bold{X}$ as a draw from the product of a multivariate normal and multinomial distribution
+At a given time $t$, we model each observation $\bold{x} \in \bold{X}$ as a draw from the product of a multivariate normal and multinomial distribution
 \begin{equation}
 \label{likelihood}
-P(\bold{x}|\theta) = \mathcal{N}(\bold{x}^{s} | \boldsymbol{\mu}, \Sigma)  \mathcal{M}n(\bold{x}^{c} | \bold{p})
+\text{F}(\bold{x}|\theta) = \mathcal{N}(\bold{x}^{s} | \boldsymbol{\mu}, \Sigma)  \mathcal{M}n(\bold{x}^{c} | \bold{p})
 \end{equation}
 where $\theta = \{ \boldsymbol{\mu}, \Sigma, \bold{p} \}$ denote the parameters of a cluster at time $t$, with mean $\boldsymbol{\mu} \in \mathbb{R}^{2}$, covariance matrix $\Sigma \in \mathbb{R}^{2} \times \mathbb{R}^{2}$, and discrete probability vector $\bold{p} = (p_{1}, \ldots, p_{V})$ such that $\sum_{i=1}^{V}p_{i} = 1$.
-Also, $\mathcal{N}$ denotes the multivariate normal distribution and $\mathcal{M}n$ denotes the multinomial distribution,
+Additionally, $\mathcal{N}$ denotes the multivariate normal distribution, $\mathcal{M}n$ denotes the multinomial distribution, and F is sometimes referred to as the likelihood.
 
@@
 
-% base distribution / appearance prior
-% ------------------------------------
+\subsection{Appearance Prior (Base Distribution)}
+\label{sec:appearanceprior}
 
-\subsubsection{Base Distribution $\mathbb{G}_{0}$ and Appearance Prior}
-
-$\mathbb{G}_{0}$ denotes the base distribution of the time-dependent Dirichlet process mixture; it also serves as a prior distribution for the parameters $\theta = \{ \boldsymbol{\mu}, \Sigma, \bold{p} \}$ present in the likelihood.
+$\mathbb{G}_{0}$ denotes the base distribution of the time-dependent Dirichlet process mixture; it also serves as a prior distribution for the parameters $\theta = \{ \boldsymbol{\mu}, \Sigma, \bold{p} \}$ present in the mixture components (i.e.
+the parameters we've used to model object appearance).
 We make use of conjugate priors in the base distribution to allow for more efficient computation.
-Specifically, in the experiments carried out in this paper, a normal-inverse-Wishart prior is placed on the multivariate normal parameters $\{ \boldsymbol{\mu}, \Sigma \}$ , and a Dirichlet prior is placed on the multinomial parameter $ \{  \bold{p}  \} $ (where the normal-inverse-Wishart is a conjugate prior for the multivariate normal component of the likelihood and the Dirichlet is a conjugate prior for the multinomial component).
+Specifically, in the experiments carried out in this paper, a normal-inverse-Wishart prior is placed on the multivariate normal parameters $\{ \boldsymbol{\mu}, \Sigma \}$, and a Dirichlet prior is placed on the multinomial parameter $\bold{p}$ (where the normal-inverse-Wishart is a conjugate prior for the multivariate normal component of the likelihood and the Dirichlet is a conjugate prior for the multinomial component).
 The prior can therefore be written
@@
 
-% perhaps: reference text with good background on NiW and Dir distributions
-
-% transition kernel
-% -----------------
-
-\subsubsection{Transition Kernel and Motion Model}
+\subsection{Motion Model (Transition Kernel)}
+\label{sec:motionmodel}
 
@@
 \end{eqnarray}
-where $M$ is the number of auxiliary variables, $\{ \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0} \}$ are the $\mathcal{N}i\mathcal{W}$ prior parameters, $\bold{q}_{0}$ is the $\mathcal{D}ir$ prior parameter, $\bold{z}^{s}$ and $\bold{z}^{c}$ respectively denote the spatial and color features of an auxiliary variable $\bold{z}$, and $\overline{\bold{z}}$ and $S_{\bold{z}}$ respectively denote the sample mean and sample covariance for a set $\bold{z} = \{ z_{1}, \ldots, z_{M} \}$ (of auxiliary variables, in this case), which we can define as
+where $M$ is the number of auxiliary variables, \\
+$\{ \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0} \}$ are the $\mathcal{N}i\mathcal{W}$ prior parameters, $\bold{q}_{0}$ is the $\mathcal{D}ir$ prior parameter, $\bold{z}^{s}$ and $\bold{z}^{c}$ respectively denote the spatial and color features of an auxiliary variable $\bold{z}$, and $\overline{\bold{z}}$ and $S_{\bold{z}}$ respectively denote the sample mean and sample covariance for a set $\bold{z} = \{ z_{1}, \ldots, z_{M} \}$ (of auxiliary variables, in this case), which we can define as
 \begin{eqnarray}
@@
 
-% note:
-% don't i have to include more to really define the transition kernel?? I feel there is something that is not fully specified.
-
-% recap of parameters
-% -------------------
-
-\subsection{Recap of Parameters}
-\label{sec:parameter_recap}
+\subsection{Recap of Parameters in Model}
+\label{sec:recapofparameters}
 
@@
 \begin{center}
-\begin{tabular}[c]{l p{10cm}}
+\begin{tabular}[c]{l p{7cm}}
 $\boldsymbol{\mu}_{0}$  &  The mean prior. In experiments performed in Section~\ref{sec:experiments} the data was recentered to the origin, and this parameter was set to $(0,0)$ \\
@@
 \begin{center}
-\begin{tabular}[c]{l p{10cm}}
+\begin{tabular}[c]{l p{7cm}}
 $M$  &  Number of auxiliary variables. A large number will produce lower variation in position (better for slower objects) and a low number will produce higher variation in position.
@@
 \begin{center}
-\begin{tabular}[c]{l p{10cm}}
+\begin{tabular}[c]{l p{7cm}}
 $\alpha$  &  The granularity parameter for the Dirichlet Process. A higher value will increase the tendency for new objects to be detected.\\
@@
 The parameters chosen for use in experiments (Section~\ref{sec:experiments}) were accepted based on the fact that they yielded reasonable results.
-Additionally, a sensitivity analysis was carried out for the granularity and deletion parameters ($\alpha$ and $\rho$), where a range of values for both were chosen, and inference performed for each; results of this analysis are discussion in Section~\ref{sec:sens_analysis}.
+Additionally, a sensitivity analysis was carried out for the granularity and deletion parameters ($\alpha$ and $\rho$), where a range of values for both were chosen, and inference performed for each; results of this analysis are discussion in Section~\ref{sec:sensitivityanalysis}.
 
-% inference section
-% -----------------
+% Inference
+% ---------
 
@@
 \label{sec:inference}
-Algorithms that infer latent model parameters via Bayesian inference given the data gained during extraction provide detection and tracking results.
-By formulating the detection and tracking task in terms of a generative model, a variety of methods previously developed for statistical inference can be readily applied; we separate presentations of model and inference so that multiple inference methods, each with its own benefits in terms of precision and speed, can be formulated as independent algorithms for detection and tracking.
-This section provides background on the two inference algorithms implemented in this study, which are both used to perform Bayesian inference on the model described in Section~\ref{sec:modelspec}.
-The first is a type of Markov Chain Monte Carlo (MCMC) batch inference, which uses Gibbs sampling to generate samples from the posterior distribution of the model.
-The second is a type of Sequential Monte Carlo (SMC) inference, also known as a particle filter, which also generates samples from the posterior distribution of the model, but in a sequential manner that makes use of the time-dependent nature of the data.
 
-% MCMC inference section
-% ----------------------
-
-\subsection{MCMC Batch Inference}
+\subsection{MCMC: Batch Inference}
 \label{sec:MCMC}
 
-This section details a Markov Chain Monte Carlo (MCMC) sampler used to perform inference, which provides a method to carry out unsupervised detection and tracking.
-It operates by iteratively sampling from conditional distributions of elements in the joint probability distribution---i.e.
-it implements Gibbs sampling---in order to gain approximate samples from the posterior distribution of the cluster parameters given the data.
-Additionally, this sampler makes use of the Metropolis-Hastings (M-H) algorithm to sample from a few distributions.
-
-The generative model associated with the GPUDDPM is illustrated in Section~\ref{sec:gpuddpm}.
-To allow for modeling time-dependent data, the GPUDDPM allows clusters to `lose assigned observations' in a deletion step, which modifies the values of the cluster sizes over time.
-Instead of incorporating the cluster size variable $m$ directly (as it was in the definition of the GPUDDPM given in \eqref{gpuddpm_def}), we can formulate an equivalent model which makes use of deletion variables $d$, which represent the time at which observations are lost from a cluster.
-Specifically, we introduce variables $d_{1}, \ldots, d_{N}$ which denote the times at which the assignments of given observations are considered to be removed.
-At each time step, cluster sizes can be recontructed from all previous assignment and deletion variables.
-With these we can express the size of cluster $k$ at time $t$, $m_{k,t}$, as
-\begin{equation}
-\label{compute_clust_size}
-m_{k,t} = \sum_{t' = 1}^{t} \mathbb{I}[(c_{t'}=k) \wedge (t < d_{t'})]
-\end{equation}
-where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise.
-Additionally, for an observation $\bold{x}_{i}$ at a given time-step $t$, the deletion time $d_{i}$ can be defined to be $d_{i} = t + l_{i}$, where $l_{i}$ is considered the lifespan of an assignment, is distributed geometrically, and can be expressed as
-\begin{equation}
-\label{del_rho_form}
-l_{i} | \rho  \sim  \rho(1 - \rho)^{l_{i}}
-\end{equation}
-The MCMC sampler described in this section operates in this alternate representation, which we can write, for each time step $t \in \{1, \ldots, T\}$ and clusters $k \in \{ 1, \ldots, K_{t} \} $ at time $t$, as
-% \begin{align}
-% \begin{split}
-% 	 \bold{x}_{i} | c_{i}, \theta_{c_{i},t} &\sim \mathcal{N}(\boldsymbol{\mu}_{c_{i},t}, \Sigma_{c_{i},t})  \mathcal{M}n(\bold{p}_{c_{i},t}) \\
-% 	c_{i} | c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha  &\sim  \text{CRP}(c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha) \\
-% 	d_{i} | \rho  &\sim \text{Geo}(\rho) + i + 1\\
-% 	\theta_{k, t} | \theta_{k, t-1}, \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0}  &\sim 
-% \begin{cases}
-% 	\text{TK} \hspace{12pt} \text{if } k \in \{ 1, \ldots, K_{t-1}  \} \\
-% 	\mathbb{G}_{0}(\boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0})  \hspace{15pt} \text{if } k = K_{t-1} + 1
-% \end{cases}
-% \end{split}
-% \end{align}
-% where TK stands for the transition kernel, given in \eqref{transker1} and \eqref{transker2}, and the base distribution $\mathbb{G}_{0}$ is given in \eqref{basedistro}.
-\begin{align}
-\begin{split}
-d_{i,t} | \rho  &\sim \text{Geo}(\rho) + t + 1 \\
-\theta_{k, t} | \theta_{k, t-1}  &\sim
-\begin{cases}
-P(\theta_{k,t} | \theta_{k,t-1}) \hspace{2mm} \text{if} \hspace{2mm} k \in \{ 1, \ldots, K_{t} \} \\
-\mathbb{G}_{0} \hspace{2mm} \text{if} \hspace{2mm} k = K_{t} + 1
-\end{cases} \\
-% c_{i,t} | \bold{c}_{1}, \ldots, \bold{c}_{t-1}, \bold{d}_{1}, \ldots, \bold{d}_{t-1}, \alpha  &\sim  \text{CRP}(\bold{c}_{1}, \ldots, \bold{c}_{t-1}, \bold{d}_{1}, \ldots, \bold{d}_{t-1}, \alpha) \\
-c_{i,t} | \bold{c}_{1:t-1}, \bold{d}_{1:t-1}, \alpha  &\sim  \text{CRP}(\bold{c}_{1:t-1}, \bold{d}_{1:t-1}, \alpha) \\
-\bold{x}_{i,t} | c_{i,t}, \theta_{c_{i,t},t}  &\sim  F(\theta_{c_{i,t}, t})
-\end{split}
-\end{align}
-where $\bold{c}_{t} = c_{1,t}, \ldots, c_{N_{t}, t}$ (for $N_{t}$ observations at time $t$), $\bold{d}_{t} = d_{1,t}, \ldots, d_{N_{t}, t}$, $P(\theta_{k,t} | \theta_{k,t-1})$ is defined in \eqref{transker1} and \eqref{transker2}, $\mathbb{G}_{0}$ is defined in \eqref{basedistro}, CRP is defined in \eqref{CRP}, and $F(\theta_{c_{i,t}, t})$ is defined in \eqref{likelihood}.
-This formulation of the GPUDDPM is also used by \cite{gasthaus_thesis} and \cite{caron_2007}.
-The associated graphical model for this formulation is given in Figure~\ref{fig:gpuddpm_gm_2}.
-\begin{figure}[h]
-        \center{\includegraphics[width=100mm]{../img/gpuddp_gm_3.pdf}}
-        \caption{Graphical model of the deletion variable and auxiliary variable transition kernel representation of the GPUDDPM}
-        \label{fig:gpuddpm_gm_2}
-\end{figure}
-
-We define the state of the sampler to consist of the assignment variables for all observations, $c_{1:N_{1},1}, \ldots, c_{1:N_{T}, T}$, the deletion variables for all observations, $d_{1:N_{1},1}, \ldots, d_{1:N_{T}, T}$, the sequence of cluster parameters for each active cluster, $\theta_{k,1}, \ldots, \theta_{k, T}$ and the $M$ associated auxiliary variables for each cluster $\theta_{k,t}$ at time $t$, $z_{k,t,1}, \ldots, z_{k, t, M}$.
-Formally, we can write
-\begin{equation}
-\text{State} = \{ \bold{c}_{1:T}, \bold{d}_{1:T}, \Theta_{1:T}, \bold{m}_{1:T} \}
-\end{equation}
-where $\bold{c}_{t} = c_{1:N_{t},t}$, $\bold{d}_{t} = d_{1:N_{t},t}, \Theta_{t} = \theta_{1:K_{t}, t}$, and $m_{1:K_{t},t}$.
-The sampler performs an MCMC method known as Gibbs sampling; it operates by iteratively sampling each member of the state given all other members.
-In particular, at each time $t$ it moves sequentially through each of the $N_{t}$ observations, sampling the assignment $c_{i,t}$ and the deletion variable $d_{i,t}$.
-After the final observation at a given time is sampled, the cluster parameters for all active clusters at $t$ are sampled, and a M-H step is employed to sample the $M$ auxiliary variables for each active cluster.
-Before the samples can be iteratively drawn from each element of the state given all other elements, the state must be initialized.
-While the state of the sampler can be initialized at any value, beginning at a value closer to the maximum joint probability (over the state) allows for quicker inference.
-The following sections detail how each sample in the model is drawn.
-
-% mcmc: sampling assignment variables
-% -----------------------------------
-
-\subsubsection{Sampling Assignment Variables $c_{i,t}$}
-\label{sec:sample_assignments}
-
-A value proportional to the posterior probability can be computed for each possible value that $c_{i,t}$ may take on; these values allow us to construct a discrete probability distribution from which we can draw samples from the posterior distribution over assignments.
-The possible values that the assignment may take on are $c_{i,t} = k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*}\}$, where $K_{t}^{'}$ is the number clusters with a non-zero size at any point between time $t$ and the assignment's associated deletion time $d_{i,t}$, and $k^{*}$ is a potential new cluster.
-The probability that $c_{i,t}$ is assigned to a cluster $k$ can be computed for each $k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*} \}$, given all other members of the state, by
-\begin{equation}
-\begin{split}
-p(c_{i,t} = k | \text{State}) & \propto
-\prod_{i'=i}^{N_{t}}  \text{CRP}(c_{i',t} | \bold{m}_{t-1}, \alpha) \\
-& \times \prod_{t' = t+1}^{d_{i,t}}  \prod_{i'=1}^{N_{t'}}   \text{CRP}(c_{i',t'} | \bold{m}_{t'-1}, \alpha) \\
- & \times
-\begin{cases}
-	P(\bold{x}_{i,t} | \theta_{k,t})                                        \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
-	\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta) d\theta    \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
-\end{cases}
-\end{split}
-\end{equation}
-% where $t$ is the time of observation $i$, and the distribution under the product $P(c_{i'} | m_{1:K_{t}^{'}, i'-1}, c_{1:i'-1})$ is given by 
-%\begin{equation}
-%P(c_{i} = k | m_{1:K_{t}^{'}, i-1}, c_{1:i-1}, \alpha) \propto 
-%\begin{cases}
-%m_{k, i-1} \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
-%\alpha \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
-%\end{cases}
-%\end{equation}
-where the distribution CRP is given by \eqref{crp_rep}, and the sizes of the clusters $\bold{m}_{t'-1}$ are calculated given the fact that $c_{i,t} = k$.
-Note that the above integral can be determined analytically to give
-\begin{equation}
-\label{marginal_obs_prob}
-\int P(\bold{x}_{i} | \theta) \mathbb{G}_{0}(\theta)d\theta = \text{Student}_{\nu_{n} - 1}(\mu_{n}, \Lambda_{n}/(\kappa_{n}(\nu_{n}-1)))
-\end{equation}
-where ``Student'' denotes the Student's t distribution, where we follow the three-value parameterization (location parameter, scale parameter, and degrees of freedom) given in \cite{gelman2004bayesian}.
-
-%%%%%
-% Need to include ratio of Dirichlet also in the above posterior predictive distro of a single observation, and discuss how to compute any parameters in here that aren't yet specified
-%%%%%
-
-If a new cluster is sampled as an assignment, the cluster parameters and auxiliary variables for this new cluster must be initialized for all time steps before sampling can proceed.
-In experiments, this initialization was carried out by iteratively samping forward to time T and backwards to time 1 via the transition kernel.
-
-% mcmc: sampling cluster parameters
-% ---------------------------------
-
-\subsubsection{Sampling Cluster Parameters $\theta_{k, t}$}
-The conjugacy of the chosen distributions in the appearance model and transition kernel allow for posterior samples of the cluster parameters to be easily gained, and gives us
-% \begin{eqnarray}
-% P(\theta_{k,t} | state) && = P(\bold{x}_{i} | \theta_{k,t}) P(z_{k,t+1,1:M} | \theta_{k,t})  P(\theta_{k,t} | z_{k,t,1:M}) \\
-%  && = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}, \Sigma | \boldsymbol{\mu}_{0}, k_{0}, v_{0}, \Lambda_{0})  \mathcal{D}ir( \bold{p} | \bold{q}_{0})
-% \end{eqnarray}
-\begin{align}
-\begin{split}
-P(\theta_{k,t} | \text{State}) & = P(\bold{x}_{i,t} | \theta_{k,t}) P(z_{k,t+1,1:M} | \theta_{k,t})  P(\theta_{k,t} | z_{k,t,1:M}) \\
- & = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}_{k,t}, \Sigma_{k,t} | \boldsymbol{\mu}_{N}, k_{N}, v_{N}, \Lambda_{N})  \mathcal{D}ir( \bold{p}_{k,t} | \bold{q}_{N})
-\end{split}
-\end{align}
-Where the parameters in the above distribution are given when the data at time $t$, auxiliary variables $z_{k,t-1,1:M}$ at time $t-1$, and auxiliary variables $z_{k,t,1:M}$ at time $t$, are taken to be the observations for the following Bayesian updates
-\begin{eqnarray}
-\label{bayesian_update_1}
-\kappa_{N} &=& \kappa_{0} + N \\
-\label{bayesian_update_2}
-\nu_{N} &=& \nu_{0} + N \\
-\label{bayesian_update_3}
-\boldsymbol{\mu}_{N} &=& \frac{\kappa_{0}}{\kappa_{0}+N} \boldsymbol{\mu}_{0}  +  \frac{N}{\kappa_{0}+N} \overline{\bold{x}}^{s}\\
-\Lambda_{N} &=& \Lambda_{0} + S_{\bold{x}^{s}}\\
-\label{bayesian_update_4}
-\bold{q}_{N} &=& \bold{q}_{0} + \sum_{i=1}^{N} \bold{x}_{i}^{c}
-\end{eqnarray}
-where $N$ is the number of observations, $\{ \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0} \}$ are the $\mathcal{N}i\mathcal{W}$ prior parameters, $\bold{q}_{0}$ is the $\mathcal{D}ir$ prior parameter, $\bold{x}^{s}$ and $\bold{x}^{c}$ respectively denote the spatial and color features of the observations, and $\overline{\bold{x}}$ and $S_{\bold{x}}$ respectively denote the sample mean and sample covariance for a set of observations $\bold{x}$, defined in \eqref{samplemean} and \eqref{samplecov}.
-% \begin{eqnarray}
-% \overline{\bold{x}}  &=&  \left( \sum_{i=1}^{N} \bold{x}_{i} \right) / N\\
-% S_{\bold{x}}  &=&  \sum_{i=1}^{N} (\bold{x}_{i} - \overline{\bold{x}}) (\bold{x}_{i} - \overline{\bold{x}})^{T}
-% \end{eqnarray}
-
-% mcmc: sampling auxiliary variables
-% ----------------------------------
-
-\subsubsection{Sampling Auxiliary Variables $z_{k,t,m}$}
-The posterior distribution of each of the auxiliary variables $z_{k,t,m}$ can be written as
-\begin{equation}
-\label{stationary_pdf}
-P(z_{k,t,m} | \text{State}) \propto  P(z_{k,t,m} | \theta_{k,t-1}) P(\theta_{k,t} | z_{k,t,1:M})
-\end{equation}
-This distribution does not allow for samples to be easily drawn and thus requires a more sophisticated MCMC technique.
-In particular, the Metropolis-Hastings (M-H) algorithm is used to gain samples from the posterior distribution for the auxiliary variables; it involves constructing a markov chain with the sought posterior as its stationary distribution, and sampling sequentially from this constructed chain to get an approximate posterior sample.
-The M-H algorithm is outlined in Algorithm 1.%Algorithm~\ref{alg:MH}.
-\willie{am i citing wrong?}
-
-\begin{algorithm}[h!]
-\label{alg:MH}
-\caption{Metropolis Hastings Algorithm}
-\begin{algorithmic}[1]
-\STATE $f(x)$ is the stationary PDF from which $N$ samples are desired \hfill $\triangleright$ eq. \eqref{stationary_pdf}
-\STATE $q( \cdot | x_{i-1})$ is the proposal pdf \hfill $\triangleright$ eq. \eqref{proposal_distro}
-\STATE B is the number of burn in samples
-\STATE $x_{0} \leftarrow$ initialize
-%\STATE B $\leftarrow$ \# of burn in samples
-\FOR{$i = 1 : \text{B} + N$}
-\STATE Sample $x^{*} \sim q( \cdot | x_{i-1})$
-\STATE Sample $\alpha \sim \text{Uniform}(0,1)$
-\STATE $r_{accept} \leftarrow \frac{f(x^{*})q(x_{i-1} | x^{*})} {f(x_{i-1})q(x^{*} | x_{i-1})} $ \hfill $\triangleright$  eq. \eqref{accept_ratio}
-\IF{$\alpha < \text{min} \{ 1, r_{accept} \} $}
-\STATE $x_{i} \leftarrow x^{*}$
-\ELSE
-\STATE $x_{i} \leftarrow x_{i-1}$
-\ENDIF
-\ENDFOR
-\STATE Discard samples $x_{1}, \ldots, x_{\text{B}}$
-\STATE \textbf{Output}: Remaining samples $x_{1}, \ldots, x_{N}$
-\end{algorithmic}
-%\caption{Say thing at end}
-\end{algorithm}
-
-The M-H algorithm is used to sample auxiliary variables for each alive (positive size) cluster.
-In short, we sample a new value for all $M$ auxiliary variables, denoted $z_{k,t,m}^{*}$, according to the following proposal distribution
-\begin{equation}
-\label{proposal_distro}
-z_{k,t,m}^{*}  \sim  P(z_{k,t,m} | \theta_{k,t}) = \mathcal{N}(z_{k,t,m}^{s} | \boldsymbol{\mu}_{k,t}, \Sigma_{k,t}) \mathcal{M}n(z_{k,t,m}^{c} | \bold{p}_{k,t})
-\end{equation}
-compute the standard M-H acceptance ratio, which in this case (due to the proposal distribution choice) simplifies to 
-\begin{equation}
-\label{accept_ratio}
-r_{accept} = \frac{P(\theta_{k,t-1} | z_{k,t,m}^{*})}{P(\theta_{k,t-1} | z_{k,t,m})}
-\end{equation}
-and choose these auxiliary variables as the next sample in the constructed sequence if $\alpha < \text{min}(1, r_{accept})$, where $\alpha$ is a sample from a $\text{Uniform}(0,1)$ distribution.
-
-% mcmc: sampling deletion variables
-% ---------------------------------
-
-\subsubsection{Sampling Deletion Variables $d_{i}$}
-Computing the posterior distribution (or a value proportional to this distribution) for each possible value that a given deletion variable may take on, may be accomplished in a manner similar to how samples from the posterior distribution of the assignment variables $c_{i}$ are gained in Section~\ref{sec:sample_assignments}; however, this may be computationally expensive due to the large number of potential deletion times.
-To remedy this, the M-H algorithm may again be used to sample from the posterior, where the prior distribution for the deletion times can be used to propose new samples $d_{i}^{*}$.
-As the ``lifetime'' of each assignment can be shown to be geometrically distributed, we can propose a new deletion time sample with
-\begin{align}
-\begin{split}
-l_{i}  &\sim  \text{Geo}(\rho)  \\
-d_{i}^{*}  &= l_{i} + t + 1
-\end{split}
-\end{align}
-and accept or reject this sample using the standard M-H acceptance ratio.
-
-%%%%%
-% need to include the "standard M-H acceptance ratio"
-%%%%%
-
-% SMC inference section
-% ---------------------
-
-\subsection{SMC Inference}
+\subsection{SMC: Sequential Inference}
 \label{sec:SMC}
 
-This section details a Sequential Monte Carlo (SMC) sampler---also known as a particle filter---used to perform inference, which provides another method to carry out unsupervised detection and tracking.
-The SMC inference algorithm is given in Algorithm~\ref{alg:SMC}.
-In this algorithm, at each time step $t \in \{ 1, \ldots, T \}$, a number of samples refered to as ``particles'' are generated; each particle consists of a sample from the posterior distribution over the assignment for each observation, $c_{1,t}, \ldots, c_{N_{t}, t}$, parameters for each cluster $\theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and size after deletion for each cluster $m_{1,t}, \ldots, m_{K_{t},t}$.
-A set of particles is sampled at each time step from relevant proposal distributions (described in Sections~\ref{sec:smc_proposal_1}, \ref{sec:smc_proposal_2}, and \ref{sec:smc_proposal_3}), a weight is computed for each particle, and a new set of particles are sampled from the set of weighted particles via a resampling process.
-Within each time step, Gibbs sampling is used to generate the samples associated with each particle.
+\subsection{From Inference to Tracking Results}
+\label{sec:inferencetotrackingresults}
 
-The sequence of target distributions for the SMC algorithm may be written as
-\begin{equation}
-\begin{split}
-%\pi_{t}(c_{1:N_{1}, 1}, \ldots, c_{1:N_{t}, t}, & \theta_{1:K_{1}, 1}, \ldots, \theta_{1:K_{t}, t}, m_{1:K_{1}, 1}, \ldots, m_{1:K_{t}, t}) = \\
-%\pi_{t-1}(c_{1:N_{1}, 1}, \ldots, c_{1:N_{t-1}, t-1}, & \theta_{1:K_{1}, 1}, \ldots, \theta_{1:K_{t-1}, t-1}, m_{1:K_{1}, 1}, \ldots, m_{1:K_{t-1}, t-1}) \\
-\pi_{t}(\bold{c}_{1:t}, \Theta_{1:t}, \bold{m}_{1:t}) & =  \pi_{t-1}(\bold{c}_{1:t-1}, \Theta_{1:t-1}, \bold{m}_{1:t-1})\\
-& \times \prod_{i=1}^{N_{t}} P(c_{i,t} | \bold{m}_{t}, \Theta_{t}, \bold{c}_{1:t}, \bold{x}_{1:N_{t}}) \\
-& \times \prod_{k=1}^{K} 
-\begin{cases}
-\begin{split}
-P(\theta_{k,t} | \theta_{k,t-1}) \hspace{2mm} & \text{if} \hspace{2mm} k \leq K_{t-1} \\
-\mathbb{G}_{0} \hspace{18mm} & \text{if} \hspace{2mm} k > K_{t-1}
-\end{split}
-\end{cases} \\
-& \times \prod_{k=1}^{K} \text{DEL}(m_{k,t} | m_{k,t-1}, c_{1:N_{t-1}, t-1}, \rho)
-\end{split}
-\end{equation}
-where $\bold{c}_{t} = c_{1,t}, \ldots, c_{N_{t}, t}$, $\Theta_{t} = \theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and $\bold{m}_{t} = m_{1,t}, \ldots, m_{K_{t}, t}$ The distributions used in the SMC algorithm are described in detail in the following sections.
-
-% \begin{algorithm}
-% \caption{Sequential Monte Carlo Inference for the GPUDDPM}
-% \label{alg:SMC}
-% \begin{algorithmic}[1]
-% \FOR{$t = 1 : T$}
-% \FOR{$l = 1 : L$}
-% \STATE $K_{t}^{(l)}  \leftarrow K_{t-1}^{(l)}$
-% \FOR{$i = 1 : N_{t}$}
-% \STATE Sample assigments from previous time sizes and cluster parameters $\theta_{k,t-1}$
-% \STATE Draw $c_{i} \sim P(c_{i} | m_{1:K_{t-1}, i-1}, \theta_{1:K_{t-1}, t-1} )$
-% \IF{$c_{i} = K_{t} + 1$}
-% \STATE Draw a new cluster $\theta_{K_{t}+1,t} \sim \mathbb{G}_{0}(\text{priors})$   \willie{distro after algo?}
-% \STATE $K_{t} \leftarrow K_{t} + 1$
-% \ENDIF
-% \ENDFOR
-% \FOR{$s = 1 : S$}
-% \FOR{$k = 1 : K$}
-% \STATE Sample new cluster params via aux (given curent assignments and prev clusters)
-% \ENDFOR
-% \FOR{$i = 1 : N_{t}$}
-% \STATE Sample new assignments (given current cluster params)
-% \STATE Draw $c_{i} \sim P(c_{i} | m_{1:K_{t}, i-1}, \theta_{1:K_{t}, t} )$
-% \IF{$c_{i} = K_{t} + 1$}
-% \STATE Draw a new cluster $\theta_{K_{t}+1,t} \sim \mathbb{G}_{0}(\text{priors})$   \willie{distro after algo?}
-% \STATE $K_{t} \leftarrow K_{t} + 1$
-% \ENDIF
-% \ENDFOR
-% \STATE Sample new cluster sizes.
-% \ENDFOR
-% \ENDFOR
-% \STATE Only keep some of the $L$ samples.
-% \ENDFOR
-% \end{algorithmic}
-% \end{algorithm}
-
-%%%%%
-% in SMC algorithm: still need to include computing r and make sure things are correct
-%%%%%
-
-\begin{algorithm}[!]
-\caption{Sequential Monte Carlo Inference for the GPUDDPM}
-\label{alg:SMC}
-\begin{algorithmic}[1]
-\FOR{$l = 1 : L$}
-\STATE $w_{0}^{(l)} \leftarrow 1/L$ \hfill $\triangleright$ initialize weights
-\ENDFOR
-\STATE $K_{0}^{(l)} \leftarrow 0$ \hfill $\triangleright$ initialize \# of clusters
-\FOR{$t = 1 : T  \text{  (\# of frames})$}
-\FOR{$l = 1 : L  \text{  (\# of particles})$}
-\STATE $K_{t}^{(l)}  \leftarrow K_{t-1}^{(l)}$
-\STATE $m_{1:K_{t}^{(l)}, t}^{(l)} \leftarrow m_{1:K_{t-1}^{(l)}, t-1}^{(l)}$
-\FOR{$s = 1 : S \text{  (\# of Gibbs samples})$}
-\FOR{$i = 1 : N_{t} \text{  (\# of observations at frame $t$})$}
-\IF{$s=1$}
-\STATE Sample $c_{i,t}^{(l)} \sim P \left( c_{i,t}^{(l)} | m_{1:K_{t-1}^{(l)}, t-1}^{(l)}, \theta_{1:K_{t-1}^{(l)}, t-1}^{(l)}, \alpha \right)$  \hfill $\triangleright$ eq.~\eqref{smc_assig_eqn}
-\STATE $m_{c_{i,t}^{(l)},t}^{(l)} \leftarrow m_{c_{i,t}^{(l)},t}^{(l)} + 1$
-\ELSE
-\STATE $m_{c_{i,t}^{(l)},t}^{(l)} \leftarrow m_{c_{i,t}^{(l)},t}^{(l)} - 1$
-\STATE Sample $c_{i,t}^{(l)} \sim P \left( c_{i,t}^{(l)} | m_{1:K_{t}^{(l)}, t}^{(l)}, \theta_{1:K_{t}^{(l)}, t}^{(l)}, \alpha \right)$  \hfill $\triangleright$ eq.~\eqref{smc_assig_eqn}
-\STATE $m_{c_{i,t}^{(l)},t}^{(l)} \leftarrow m_{c_{i,t}^{(l)},t}^{(l)} + 1$
-\ENDIF
-\IF{$c_{i,t}^{(l)} = K_{t}^{(l)} + 1  \text{  (a new cluster)}$}
-\STATE Sample $\theta_{c_{i,t}^{(l)}, t}^{(l)} \sim q_{1}(\bold{x}_{i,t})$ \hfill $\triangleright$ eq.~\eqref{smc_q1_eqn}
-\STATE $K_{t}^{(l)} \leftarrow K_{t}^{(l)} + 1$
-\STATE $m_{K_{t}^{(l)},t}^{(l)} \leftarrow 1$
-\ENDIF
-\ENDFOR
-\FOR{$k = 1 : K_{t}^{(l)}  \text{  (\# of clusters at frame $t$})$}
-\IF{$k > K_{t-1}^{(l)}$}
-\STATE Sample $\theta_{k,t}^{(l)} \sim q_{1}(\{ \bold{x}_{1:N_{t}, t} = k \})$  \hfill $\triangleright$ eq.~\eqref{smc_q1_eqn}
-\ELSIF{$k \leq K_{t-1}^{(l)}$ and $ \#\{ \bold{x}_{1:N_{t}, t} = k \} > 0$}
-\STATE Sample $\theta_{k,t}^{(l)} \sim q_{2}(\theta_{k,t-1}^{(l)}, \{ \bold{x}_{1:N_{t}, t} = k \})$  \hfill $\triangleright$ eq.~\eqref{smc_q2_eqn}
-\ELSIF{$m_{k,t}^{(l)} > 0$}
-\STATE Sample $\theta_{k,t}^{(l)} \sim P(\theta_{k,t}^{(l)} | \theta_{k,t-1}^{(l)})$  \hfill $\triangleright$ eqs.~\eqref{transker1} $\&$ \eqref{transker2}
-\ENDIF
-\IF{$s=S$}
-\STATE Sample $m_{k,t+1}^{(l)} \sim \text{DEL}(m_{k,t}^{(l)}, c_{1:N_{t}, t}^{(l)}, \rho)$  \hfill $\triangleright$ eq.~\eqref{del_step}
-\ENDIF
-\ENDFOR
-\ENDFOR
-\STATE $\tilde{w}_{t}^{(l)} \leftarrow w_{t-1}^{(l)} \times \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})}	{P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) } \times r $  \hfill $\triangleright$ eq.~\eqref{smc_weight_eqn}
-\ENDFOR
-\FOR {l = 1 : L}
-\STATE $w_{t}^{(l)} \leftarrow \frac{\tilde{w}_{t}^{(l)}}{\sum_{l=1}^{L} \tilde{w}_{t}^{(l)} }$   \hfill $\triangleright$ normalize weights
-\ENDFOR
-\STATE Resample particles $1, \ldots, L$ and weights $w_{t}^{(1)}, \ldots, w_{t}^{(L)}$   \hfill $\triangleright$ Section~\ref{sec:resample}
-\ENDFOR
-\end{algorithmic}
-\end{algorithm}
-
-\subsubsection{Proposal Distribution for Assignments $c_{i}$}
-\label{sec:smc_proposal_1}
-
-The probability of assignments given current cluster sizes, cluster parameters, and the Dirichlet Process granularity parameter $\alpha$ can be written as
-\begin{equation}
-\label{smc_assig_eqn}
-\begin{split}
-P \left( c_{i,t} | m_{1:K_{t}, t}, \theta_{1:K_{t}, t}, \alpha \right) &\propto
-\text{CRP}(m_{1:K_{t}, t}, \alpha) \\ &\times
-\begin{cases}
-P(\bold{x}_{i,t} | \theta_{c_{i,t},t}) \hspace{2mm} \text{if} \hspace{2mm} k \leq K_{t-1} \\
-\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta)d\theta \hspace{2mm} k > K_{t-1}
-\end{cases}
-\end{split}
-\end{equation}
-where CRP is defined in \eqref{CRP}, and the integral $\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta)d\theta$ can be determined analytically, and is given in \eqref{marginal_obs_prob}.
-
-\subsubsection{Proposal Distribution $q_{1}$}
-\label{sec:smc_proposal_2}
-
-The following is a distribution over cluster parameters $\theta_{k,t}$ given a set of $N$ observations $\bold{x}_{1:N, t}$. We define $q_{1}$ to be
-\begin{equation}
-\label{smc_q1_eqn}
-q_{1}(\theta_{k,t} | \bold{x}_{1:N,t}) = P(\theta_{k,t} | \bold{x}_{1:N,t})
-\end{equation}
-where samples can be drawn from $P(\theta_{k,t} | \bold{x}_{1:N,t})$ using the bayesian updates found in \eqref{bayesian_update_1}, \eqref{bayesian_update_2}, \eqref{bayesian_update_3}, and \eqref{bayesian_update_4}, where the $\bold{x}_{1:N,t}$ are taken to be the observations.
-
-\subsubsection{Proposal Distribution $q_{2}$}
-\label{sec:smc_proposal_3}
-
-The following is a distribution over cluster parameters $\theta_{k,t}$ given a set of $N$ observations $\bold{x}_{1:N, t}$ and the cluster parameters at a previous time, $\theta_{k,t-1}$.
-We define $q_{2}$ to be
-\begin{equation}
-\label{smc_q2_eqn}
-q_{2}(\theta_{k,t} | \theta_{k,t-1},\bold{x}_{1:N,t}) = P(\theta_{k,t} | \theta_{k,t-1} \bold{x}_{1:N,t})
-\end{equation}
-where samples can be drawn from $P(\theta_{k,t} | \theta_{k,t-1} \bold{x}_{1:N,t})$ using the bayesian updates found in \eqref{bayesian_update_1}, \eqref{bayesian_update_2}, \eqref{bayesian_update_3}, and \eqref{bayesian_update_4}, where both the $\bold{x}_{1:N,t}$ and auxiliary variables $z_{k,t,1:M}$ are taken to be the observations.
-
-\subsection{Calculation of particle weights $\tilde{w}_{t}^{(l)}$}
-In Algorithm~\ref{alg:SMC}, the weight $\tilde{w}_{t}^{(l)}$ of each particle $l \in \{ 1. \ldots, L \}$ at time $t$ is given in terms of the particle's weight at time $t-1$ by the equation
-\begin{equation}
-\tilde{w}_{t}^{(l)} \leftarrow w_{t-1}^{(l)} \times \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})}	{P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) } \times r
-\end{equation}
-where $r$ denotes the computed ratio for the particle (defined in Algorithm~\ref{alg:SMC}), and we define
-\begin{equation}
-\label{smc_weight_eqn}
-\begin{split}
-& \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})}	{P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) }  = \\
-% & \frac{P \left( \bold{x}_{1, t} | \theta_{c_{1, t}^{(l)}, t}^{(l)} \right) \times \ldots \times P \left( \bold{x}_{N_{t}, t} | \theta_{c_{N_{t}, t}^{(l)}, t}^{(l)} \right) \times
-% P \left( c_{1, t}^{(l)} | m_{1:K_{t}^{(l)}, t}^{(l)} \right) \times \ldots \times P \left( c_{N_{t}, t}^{(l)} | m_{1:K_{t}^{(l)}, t}^{(l)} \right)}{b}
-\end{split}
-\end{equation}
-
-%%%%%
-% need to complete this equation
-%%%%%
-
-\subsection{Resampling Particles and Particle Weights}
-\label{sec:resample}
-
-At each time step $t \in \{ 1, \ldots, T \}$, after all of the $L$ particles have been sampled and their associated weights computed, a resampling step is carried out.
-In this step, $L$ new particles are sampled from current set of $L$ particles.
-We will describe a specific, simple method known as Multinomial Resampling.
-In this method of resampling, we draw
-\begin{equation}
-\Upsilon \sim \mathcal{M}n((w_{t}^{(1)}, \ldots, w_{t}^{(L)}))
-\end{equation}
-where $\mathcal{M}n$ denotes the multinomial distribution, and $\Upsilon$ represents a vector of counts showing the multiplicity of each of the old particles in the new particle set (i.e.
-$\Upsilon_{i}$ denotes how many times particle $i$ is duplicated in the new particle set).
-More sophisticated resampling strategies are described for the SMC algorithm in \cite{gasthaus_thesis} and \cite{douc2005comparison}.
-
-%%%%%
-% need to include particle resampling strategies
-%%%%%
-
-\subsection{Modeling Background Noise}
-
-% Noisy extraction procedures may result in spurrious data points that do not correspond to any foreground object.
-We would like to incorporate these points into our model.
-To do this, we modify our base distribution $\mathbb{G}_{0}$ to be a two component mixture, where one component is the base distribution and the other is a multivariate normal distribution with a fixed `flat' covarance $\sigma = [100, 0; 0, 100]$ multiplied with an `even' multinomial distribution with a fixed parameter $\bold{p} = (1/K, \ldots, 1/K)$.
-Noisy extraction procedures may result in spurrious data points that do not correspond to any foreground object.
-We would like to incorporate these points into our model.
-To do this, we modify our model to be a two component mixture model, where one component is the previous model (the GPUDDPM) and the other is a multivariate normal distribution with with a fixed `flat' covarance $\sigma = [100, 0; 0, 100]$ multiplied with an `even' multinomial distribution with a fixed parameter $\bold{p} = (1/K, \ldots, 1/K)$.
-This normal-multinomial component is left intentionally broad, and only the mean parameter of the normal distribution is inferred.
-We can write this generative model as
-\begin{align}
-\label{background_model}
-\begin{split}
-\gamma_{i,t} | p_{1}, p_{2} &\sim \text{Discrete}(p_{1}, p_{2})\\
-m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho  &\sim \text{DEL}(m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) \\
-\theta_{k, t} | \theta_{k, t-1}   &\sim
-\begin{cases}
-	P(\theta_{k, t} | \theta_{k, t-1}) \hspace{2mm} \text{if} \hspace{2mm} k \in \{ 1, \ldots, K_{t} \} \\
-	\mathbb{G}_{0}   \hspace{2mm} \text{if} \hspace{2mm} k = K_{t+1}
-\end{cases} \\
-c_{i,t} | m_{1:K_{t},t}, \hspace{1mm} \alpha  &\sim  \text{CRP} (m_{1:K_{t},t}, \hspace{1mm} \alpha) \\
-\bold{x}_{i,t} | c_{i,t}, \theta_{1:K_{t}, t}, \gamma_{i,t}, \phi &\sim
-\begin{cases}
-	F(\theta_{c_{i,t}, t}) \hspace{2mm} \text{if} \hspace{2mm} \gamma_{i,t} = 1 \\
-	\phi \hspace{2mm} \text{if} \hspace{2mm} \gamma_{i,t} = 2
-\end{cases} \\
-\end{split}
-\end{align}
-where $\phi$ represents the normal-multinomial component distribution defined above.
-
-%%%%%
-% define more rigorously, and then say why we don't include in experiments?--- or perhaps include for an example in experiments
-%%%%%
-
-\subsection{From Inference to Tracking Results}
-\label{sec:inference_to_results}
-The multivariate normal-multinomial model on which inference is described in this section can be used to gain object detection and tracking results.
+% The multivariate normal-multinomial model on which inference is described in this section can be used to gain object detection and tracking results.
 The nature of object detection and tracking results may vary depending on the application or performance metrics used (see Section~\ref{sec:performance_metrics}); in this study, we desire the centroid position and approximate spatial region for each moving object, for each frame that this object can be seen within the video.
@@
 
-% experiments section
-% -------------------
+% Experiments
+% -----------
 
@@
 
-This section provides details on performance evaluation metrics that have been developed to quantify results in object detection and tracking studies (and adopted in this paper), synthetic video experiments that verify certain aspects of the developed technique, benchmark video experiments that demonstrate the performance of this technique in relation to other methods developed in recent years, and new video experiments to show the capability of this technique to perform unsupervised detection and tracking on a range of object and video types.
-
 \subsection{Performance Evaluation Metrics}
-\label{sec:performance_metrics}
-
-Performance evaluation metrics, which intend to provide a standardized way of quantifying the success of a detection and tracking procedure on a given video, have started to become well studied and consistently used in the past four years.
-The metrics developed in \cite{kasturi_2008} and used in \cite{ellis_2010, taj_2007, lee_2009} have become well established metrics for evaluating the performance of object detection and tracking in videos and have been adopted by the Video Analysis and Content Extraction (VACE) program and the Classification of Events, Activities, and Relationships (CLEAR) consortium, two large-scale efforts concerned with video tracking and interaction analysis.
-These metrics are used to quantify the experimental results in this study.
-
-The above metrics are dependent upon ground-truth data specifying the positions of each object in each frame throughout a video sequence.
-In the experiments described below, we gained the synthetic video ground-truth during construction of the videos (described in Section~\ref{sec:syntheticvideos}), and we used the Video Performance Evaluation Resource (ViPER) ground-truth software \cite{doermann_2000}, an open source tool commonly used in the video tracking community, to author ground-truth data for each of the benchmark datasets.
-
-\subsubsection{Mapping Ground-Truth to Output}
-\label{sec:mappingtoground}
-
-The problem of finding a mapping between a video's ground-truth tracks and an algorithm's output tracks is nontrivial, though necesary to solve, in order to compute the performance evaluation metrics used in this study.
-In short, the typical solution to this problem involves first specifying a performance metric and then choosing the mapping from ground-truth tracks to output tracks which yields the most favorable performance metric value.
-This process is described in detail by Kasturi et al.
-\cite{kasturi_2008}; we follow the method outlined in this paper to find an optimal mapping.
-Similiar to descriptions in \cite{kasturi_2008}, we implement the Hungarian algorithm \cite{munkres_1957} as a polynomial-time ($O(n^{3})$) solution to the problem of optimally mapping two sets of tracks once the similarity between any two tracks given some specified metric is established.
-Additionally, the method employed in \cite{kasturi_2008} allows erroneous and undetected tracks to be left unmapped, which is both desired and necessary in the case where there is a different number of ground-truth and output tracks.
-Note that once a mapping from a collection of ground-truth tracks to a collection of result tracks has been established, one can determine which result tracks are false positives (the result tracks to which no ground-truth track is assigned) and which ground-truth tracks are true negatives (the ground-truth tracks that are not assigned to a result track).
-The numbers of tracks displaying both of these failures are factors in the performance metrics used in this study.
-
-%, where a numerical search algorith---the Hungarian algorithm \cite{munkres_1957}--is used to find an optimal mapping from ground-truth to output tracks without requiring the performance metric value of all possible mapping combinations to be computed (reducing the time complexity from factorial to polynomial time)
-
-% \subsubsection{MODA and MOTA}
-
-% The two metrics used to quantify results in this study are refered to as the Multiple Object Detection Accuracy (MODA) and Multiple Object Tracking Accuracy (MOTA).
-These metrics, along with the the ground-truth tracks of all objects in the video sequence and the the mapping as defined in the previous section, can be used to evaluate detection and tracking accurancy of the video.
-For a given video frame $t$, we define MODA as
-% \begin{equation}
-% MODA = 1 - \frac{m^{t} + e^{t}}{N_{g}^{t}}
-% \end{equation}
-% where $m^{t}$ is the number of missed object detections (ie true-negative detections), $e^{t}$ is the number of erroneous object detections (ie false positive detections), and $N_{g}^{t}$ is the number of ground truth objects (all at frame $t$).
-
-% For a sequence of video frames $t \in \{ 1, \ldots, T \}$ we define MOTA as
-% \begin{equation}
-% MOTA = 1 - \frac{\sum_{t=1}^{T} m^{t} + e^{t} + s_{t}}{\sum_{t=1}^{T} N_{g}^{t}}
-% \end{equation}
-% where $s$ denotes the number of instances at frame $t$ where an object's identity switches from a previous frame (ie where the object's mapping differs between frame $t$ and frame $t-1$),  and the other terms are defined the same as above.
-
-\subsubsection{SFDA and ATA}
-
-The two metrics used to quantify performance in this study are known as the the Sequence Frame Detection Accuracy (SFDA) and the Average Tracking Accuracy (ATA).
-These metrics were developed during VACE Phase II to provide a single, comprehensive metric to describe detection, and one to describe tracking.
-The following are used in the definitions of the performance metrics:
-\begin{itemize}
-\renewcommand{\labelitemi}{$\bullet$}
-\item $G_{i}$ denotes the spatiotemporal region occupied by the $i$th ground-truth object in a video, and $G_{i}^{(t)}$ denotes the region occupied by the $i$th ground-truth object in frame $t$.
-\vspace{2mm}
-\item $D_{i}$ denotes the spatiotemporal region occupied by the $i$th detected object in a video, and $D_{i}^{(t)}$ denotes the region occupied by the $i$th detected object in frame $t$.
-\vspace{2mm}
-\item $N_{G}$ denotes the total number of unique ground-truth objects in a video, and $N_{G}^{(t)}$ denotes the number of unique ground-truth objects present at frame $t$.
-\vspace{2mm}
-\item $N_{D}$ denotes the total number of unique detected objects in a video, and $N_{D}^{(t)}$ denotes the number of unique detected objects present at frame $t$.
-\vspace{2mm}
-\item $N_{\text{frames}}$ denotes the total number of frames in a video, and $N_{\text{frames}}^{(i)}$ denotes the number of frames in which an object $i$ (which can be a ground-truth or detected object, depending on the context) is present in a video.
-\vspace{2mm}
-\item $N_{\text{mapped}}$ denotes the number of mapped ground-truth/detect pairs in a video, and $N_{\text{mapped}}^{(t)}$ denotes the number of mapped ground-truth/detect pairs present at frame $t$.
-\vspace{1mm}
-\end{itemize}
-
-The SFDA metric quantifies the performance of an object detection algorithm as a function of the number of correct detects, false positive detects, missed (true negative) detects, and spatial allignment of detects relative to the ground-truth.
-The SFDA is calculated by computing the Frame Detection Accuracy at frame $t$ ($\text{FDA}^{(t)}$) for each frame in a video sequence.
-The FDA provides a measure of the allignment between ground-truth and detected objects in a given frame via the overlap ratio of a ground-truth/detect pair, defined to be the ratio of the intersection of ground-truth and detect regions to the union of ground-truth and detect regions.
-Formally, we can write
-\begin{equation}
-\text{FDA}^{(t)} = \frac{\text{Overlap Ratio}}{\left(\frac{N_{G}^{(t)} + N_{D}^{(t)}}{2}\right)}
-\end{equation}
-where
-\begin{equation}
-\text{Overlap Ratio} = \sum_{i = 1}^{N_{\text{mapped}}^{(t)}} \frac{\left|G_{i}^{(t)} \cap D_{i}^{(t)}\right|}{\left|G_{i}^{(t)} \cup D_{i}^{(t)}\right|}
-\end{equation}
-The term $N_{\text{mapped}}^{(t)}$ refers to an optimal mapping between ground-truth and detects at frame $t$ as specified in section \ref{sec:mappingtoground} using the $\text{FDA}^{(t)}$ as the relevant metric.
-Given the $\text{FDA}^{(t)}$ at each frame, the SFDA can be computed; this metric may be viewed as the average FDA over all frames of a video sequence.
-We define
-\begin{equation}
-\text{SFDA} = \frac{\sum_{t=1}^{N_{\text{frames}}} \text{FDA}^{(t)}}{\sum_{t=1}^{N_{\text{frames}}} \exists \left( N_{G}^{(t)} \vee N_{D}^{(t)} \right)}
-\end{equation}
-where $\exists \left( N_{G}^{(t)} \vee N_{D}^{(t)} \right)$ yields a 1 if either a detected or ground-truth object is present in frame $t$ and a 0 otherwise.
-
-The ATA metric quantifies the performance of an object tracking algorithm as a function of the spatial overlap of a mapped set of sequences of detected object positions to a set of sequences of groundtruth object positions.
-The ATA is calculated by first computing the Sequence Track Detection Accuracy (STDA), which can be viewed as a tracking performance measure unnormalied in terms of the number of objects.
-We can write the STDA as
-\begin{equation}
-\text{STDA} = \sum_{i=1}^{N_{\text{mapped}}} \frac{\sum_{t=1}^{N_{\text{frames}}} \left( \frac{\left|G_{i}^{(t)} \cap D_{i}^{(t)}\right|}{\left|G_{i}^{(t)} \cup D_{i}^{(t)}\right|}  \right) }{ N_{(G_{i} \cup D_{i} \neq \emptyset)} }
-\end{equation}
-where $N_{\text{mapped}}$ refers to an optimal mapping between ground-truth and detected objects as specified in section \ref{sec:mappingtoground} using the STDA as the relevant metric, and $N_{(G_{i} \cup D_{i} \neq \emptyset)}$ denotes the number of frames in which a given tracked object, the ground truth object to which it is mapped, or both, are present.
-
-Given the STDA for a video sequence, the ATA can be computed by the formula
-\begin{equation}
-\text{ATA} = \frac{\text{STDA}}{\left( \frac{N_{G} + N_{D}}{2} \right)}
-\end{equation}
-
-The ground-truth authored by the ViPER tool took the form of bounding boxes denoting the spatial postion of each object at each time step.
-Consequentially, to find the spatial overlap between results and ground-truth, which is intrinsic to both metrics, a rectangular bounding box was needed per object per time step from the results of the algorithm.
-We took the maximal and minimal axially aligned values of the oval produced by our algorithm (as described in Section~\ref{sec:inference_to_results}) to be the sides of a representative bounding box for a given object at a given frame.
+\label{sec:performanceevaluationmetrics}
 
 \subsection{Synthetic Video Datasets}
-\label{sec:syntheticvideos}
-
-Synthetic videos provide controlled scenarios in which aspects of the technique presented in this paper can be tested.
-Each of the following synthetic videos consists of a sequence of 200 images (each of size $500$ x $500$ pixels) containing a number of smaller colored squares of different (and potentially time-varying) sizes moving at varied speeds and trajectories over a black background.
-The synthetic videos contain instances of occlusion (where one or more objects are briefly hidden) and objects with time-varying appearances and behaviors, as these are examples of occurances that notoriously decrease the accuracy of detection and tracking.
-After each video was constructed, the extraction procedure described in Section~\ref{sec:modelspec_extraction} (using $L=3$) and inference procedures described in \ref{sec:inference} were carried out to return a sequence of multivariate-normal-parameters (means and covariance matrices) from which a sequence of positions and ovals approximating the shape of a tracked object over each frame that it is present in the video can be determined (as outlined in Section~\ref{sec:inference_to_results}).
-
-% \subsubsection{Tracking by Appearance}
-
-The first synthetic video experiment aimed to test the ability of the model and inference procedure to maintain the identity of independent objects based on color information alone.
-Two videos were constructed, both containing a red square (rgb value $[255,0,0]$ and size $20$ x $20$ pixels) and a blue square (rgb value $[0,0,255]$ and size $20$ x $20$ pixels).
-In both videos, the squares begin at opposite sides of the scene at frame $f=1$ and travel towards each other, arriving at the same location at $f=100$ (where the blue square occludes the red square).
-The second half of the two videos differ in that both squares in the first video continue in the same direction and end at the other's starting positions at $f=200$, and both squares in the second video reverse directions and end at their initial starting positions at $f=200$.
-The frame difference extraction yields identical spatial features in both videos; hence, successful tracking depends fully on the incorporation of color information into the model.
-
-Parameters were set to the same values for inference on both videos, using the method outlined in Section~\ref{sec:parameter_recap} to choose parameters.
-The chosen values were $\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (5, \ldots, 5)$.
-Inference was carried out using the MCMC algorithm (Section~\ref{sec:MCMC}); the posterior sample taken as the result correctly tracked both colored squares through occlusion in both videos, and is shown in Figure~\ref{fig:synth_one_plot}.
-
-%%%%%
-% I still need to include performance metric results for these synthetic datasets.}.
-%%%%%
-
-\begin{figure}
-  \centering               
-  \subfloat[]{\includegraphics[width=0.32\textwidth]{../img/synth1_cross.pdf}} \hspace{0.5mm}
-  \subfloat[]{\includegraphics[width=0.32\textwidth]{../img/synth1_bounce.pdf}} \hspace{0.5mm}
-  \subfloat[]{\includegraphics[width=0.32\textwidth]{../img/synth4.pdf}}
-  \caption{Each plot shows a sample from the posterior distribution of the model for synthetic experiments one (a and b) and two (c), where the vertical axis represents frame number, the horizontal axes represent spatial position, objects are denoted by marker colors and marker types, and the mean and standard deviation are shown.
-In all cases, the objects are successfully tracked through occlusion, whether they travel in a straight line (a), reverse direction (b), or do a combination of both (c).}
-  \label{fig:synth_one_plot}
-\end{figure}
-
-% \subsubsection{Tracking though Occlusion and Appearance Shift}
-
-The second synthetic video experiment aimed to test tracking performance under occlusion, object appearance change, and motion change.
-A video was constructed, containing a red square (rgb value $[255,0,0]$), a green square (rgb value $[0,255,0]$), and a blue square (rgb value $[0,0,255]$).
-The red square was of size $20$ x $20$ pixels, the blue square was of size $15$ x $15$ pixels, and the green square began at size $50$ x $50$ pixels at frame $f=1$, linearly shrinks to $10$ x $10$ pixels at $f=100$, then linearly grows back to $50$ x $50$ pixels by the end of the video, $f=200$.
-Furthermore, the red and blue squares display the same behavior as in the second video of the first synthetic experiment (they begin at opposite sides of the scene traveling towards each other, cross at the center of the scene at $f=100$, and reverse direction, ending at their initial positions at $f=200$).
-The green square begins at a point equidistant from the other two squares, intersects with them as they overlap (causing the blue square to occlude the other two), and continues on in a direction at a 20 degree angle from its initial trajectory.
-
-Parameter values were chosen in the same way, and set to the same values, as the first synthetic experiment.
-The MCMC inference algorithm correctly tracked all three objects through occlusion and inferred the appearance and size shifts.
-Figure~\ref{fig:synth_one_plot} shows a sample from the posterior distribution of the cluster parameters, where the mean and oval representation of the covariance matrix (with $0.5$ confidence value) are overlayed on the data.
-The data is plotted in three dimensional space, where time is represented on the vertical axis, and the assignment of each data point to one of the three inferred clusters is denoted by color and marker type.
-
-%Both the cluster parameter sample and data are overlaid on three semi-transparent planes showing the objecs in the video in three frames.
-\willie{still need to include performance metric results for this synthetic dataset}.
-Groundtruth for the synthetic videos
-
-% \begin{figure}
-%   \centering               
-%   % \subfloat[]{\includegraphics[width=0.35\textwidth, angle=30]{../img/synth2_img}}
-%   \subfloat[]{\includegraphics[width=0.55\textwidth]{../img/synth2_img_b}}
-%   \subfloat[]{\includegraphics[height=0.3\textheight]{../img/synth2_alphaimg_hd.png}}
-%   \caption{(a.) Video frame $f=10$ from synthetic experiment two.
-(b.) Sample from posterior of the state for synthetic experiment two, where color represents assignment and the mean and standard deviation are shown.
-All are overlaid on three, time-varying images from synthetic experiment.
-The three objects of time varying size, speed, and position are tracked through simultaneous occlusion.
-\willie{will show assignments by marker type in addition to color}}
-%   \label{fig:synth_two_plot}
-% \end{figure}
-
-% \begin{figure}[h]
-%         \center{\includegraphics[width=65mm]{../img/synth2_alphaimg_hd.png}}
-%         \caption{\label{fig:synth2_alphaimg} Sample from posterior of the state for synthetic experiment two, where color represents assignment and the mean and standard deviation are shown.
-Both the sample and data are overlaid on three images from synthetic experiment.
-The three objects of time varying size, speed, and position are tracked through simultaneous occlusion.
-\willie{will show assignments by marker type in addition to color}}
-% \end{figure}
-
-% \begin{figure}[h]
-%         \center{\includegraphics[width=0.99\textwidth]{../img/synth2_alphaimg_hd.png}}
-%         \caption{\label{fig:synth2_alphaimg} Extracted data and sample from the posterior of the model overlaid on three, time-varying images from synthetic experiment 2.}
-% \end{figure}
+\label{sec:syntheticvideodatasets}
 
 \subsection{Benchmark Video Datasets}
-
-Benchmark video datasets for object tracking and detection have been produced to provide standard scenes on which researchers can compare detection and tracking results.
-These videos have been primarily produced for surveillance-related workshops---notably, for the International Workshop on Performance Evaluation of Tracking and Surveillance (PETS)--which provide researchers with video datasets and algorithmic goals on which to focus.
-Three commonly used benchmark videos from PETS workshops---one used in PETS2000, one in PETS2001, and one used both in PETS2009 and PETS2010---were chosen to demonstrate the efficacy of the method presented in this study.
-The performance metrics and benchmark datasets allow the methods developed in this paper to be quantitatively compared against other detection and tracking algorithms.
-
-%\subsubsection{PETS2000}
+\label{sec:benchmarkvideodatasets}
 
 \subsubsection{PETS2000 and PETS2001}
-\label{sec:pets2000_2001}
-
-The PETS2000 and PETS2001 video datasets both consist of a small number of humans and vehicles traveling across a parking lot, with video taken from above, emulating what might be recorded by standard outdoor surveillance equipment.
-The `Test Sequence', a set of images from a monocular, stationary camera, was used from the PETS2000 workshop, and View Two of Dataset 1, also taken via a monocular, stationary camera, was used from the PETS2001 workshop.
-The MCMC algorithm (described in Section~\ref{sec:MCMC}) was used for inference in these experiments.
-
-Due to the computation required for the MCMC batch inference method (discussed further in Section~\ref{sec:discussion}), only the final 1000 frames of the video were used from both datasets.
-Extraction was performed with frame differencing as described in Section~\ref{sec:modelspec_extraction}, using $L=3$.
-Parameter values were chosen in the same way, and set to the same values, as in the the synthetic experiments ($\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (5, \ldots, 5)$).
-The MCMC sampler was successful for both benchmark videos; each object was detected, tracked, and its shape estimated in manner very consistent with the ground-truth.
-The results for the PETS2000 dataset are displayed in Figure~\ref{fig:pets2000_results} and for the PETS2001 dataset in Figure~\ref{fig:pets2001_results}; in these figures, a sample from the posterior distribution of the cluster parameters is overlayed on the extracted data over a sequence of frames, where the assignment of each data point is represented by color and marker type.
-
-\begin{figure}[h]
-  \centering
-  \subfloat[]{\label{fig:pets2000_results} \includegraphics[width=0.48\textwidth]{../img/pets2000_front.pdf}} \hspace{1mm}           
-  \subfloat[]{\label{fig:pets2001_results} \includegraphics[width=0.48\textwidth]{../img/pets2001_back.pdf}}
-  \caption{PETS2000 dataset results (a) and PETS2001 dataset results (b).
-Both plots show a sample from the posterior distribution of the state, where the vertical axis denotes time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.}
-  \label{fig:benchmark_results_1}
-\end{figure}
-
-\begin{figure}[h]
-  \centering             
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f1}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f4}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f9}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f11}}
-  \caption{Four frames from the PETS2000 sequence with one posterior sample mean and covariance matrix representation shown for each frame (and for the previous 20 frames).}
-  \label{fig:pets2001_imgs}
-\end{figure}
-
-\begin{figure}
-  \centering               
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f1.pdf}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f3.pdf}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f5.pdf}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f8.pdf}}
-  \caption{Four frames from the PETS2001 sequence with one posterior sample mean and covariance matrix representation shown for each frame (and for the previous 20 frames).}
-  \label{fig:pets2001_overlay}
-\end{figure}
-
-To calculate performance metrics (both SFDA and ATA), one must specify a confidence value that allows the oval representing the region occupied by an object to be computed from the inferred covariance matrix of each cluster (as discussed in Section~\ref{sec:inference_to_results}).
-The performance metrics were found for a range of confidence intervals, and these curves, for the PETS2000 and PETS2001 video are shown in Figure~\ref{fig:pm_conf}.
-
-\begin{figure}[h]
-  \centering             
-  \subfloat[]{\includegraphics[width=0.49\textwidth]{../img/pets2000_pm.pdf}} \hspace{1mm}
-  \subfloat[]{\includegraphics[width=0.49\textwidth]{../img/pets2001_pm.pdf}}
-  \caption{The SFDA (blue solid line) and ATA (red dashed line are) vs  confidence values from which an object's oval region is computed for (a) PETS2000 and (b) PETS2001 video datasets.}
-  \label{fig:pm_conf}
-\end{figure}
-
-% 3-panel alpha-img figure that I removed
-
-% \begin{figure}[h]
-% 	\centering
-%     \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/synth2_alphaimg_hd.png}}
-%     \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2001_alphaimg_hr.png}}
-% 	\subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2000_4objects.png}}
-%     \caption{\label{fig:synth2_alphaimg} Sample from posterior of the state for synthetic experiment two (a), and benchmark datasets PETS2001 (b) and PETS2000 (c), where color represents assignment and the mean and standard deviation are shown.
-In each plot, the sample and data are overlaid on three images from the experiment.}
-% \end{figure}
-
-% old figure ideas:
-
-% \begin{figure}[h]
-%         \center{\includegraphics[width=65mm]{../img/pets2001_alphaimg_hr.png}}
-%         \caption{\label{fig:pets2001_alphaimg} Extracted data and sample from posterior of the model overlaid on three frames from the PETS2001 video.}
-% \end{figure}
-
-% \begin{figure}[h]
-% 	\centering
-% 	\subfloat[]{\includegraphics[width=0.58\textwidth]{../img/pets2000_results_1}}
-% 	\subfloat[]{\includegraphics[width=0.42\textwidth]{../img/pets2000_4objects.png}}
-% 	\caption{Extracted data and a sample from posterior of the model (a) overlaid on a frame from the PETS2000 video (b).}
-% 	\label{fig:pets2000_results}
-% \end{figure}
-
-% \begin{figure}
-%   \centering               
-%   \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2000_front.pdf}}
-%   \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2000_back.pdf}}
-%   \caption{PETS2000 dataset results, from a front view (a) and a back view (b).
-Both plots show a sample from the posterior distribution of the state, where the vertical axis denots time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.}
-%   \label{fig:pets2000_results}
-% \end{figure}
+\label{sec:pets2000and2001}
 
 \subsubsection{PETS2009/2010}
-
-A video dataset used in both the PETS2009 and PETS2010 conferences, called `S2.L1 at time sequence 12.34' was chosen for experimentation due to its prominence in a number of studies \cite{ellis_2010}.
-%%%%%
-% cite other papers that use this dataset?
-%%%%%
-This dataset consists of a monocular, stationary camera, 794 frame video sequence. The entire video sequence was used in this experiment.
-
-Due to the large number of frames and objects in this video, the SMC algorithm (described in Section~\ref{sec:SMC}) was used for inference.
-This method of sequential inference was observed, on this dataset, to converge to a better sample in a shorter period of time in comparison with the MCMC algorithm; for this reason, we found that it was better suited for this large dataset.
-
-Extraction was performed with frame differencing as described in Section~\ref{sec:modelspec_extraction}, using $L=3$ and parameters for the model were chosen, using the method outlined in Section~\ref{sec:parameter_recap}, to be $\alpha = 0.1, \rho = 0.8, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 6, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (3, \ldots, 3)$.
-Additionally, as with the other video datasets, ground-truth bounding boxes around each object were authored using the ViPER tool.
-
-The SMC inference algorithm yielded a sample from the posterior distribution of the model, from which the object detection and tracking results were obtained (as described in Section~\ref{sec:inference_to_results}).
-In Figure \ref{fig:pets2009_results}, a sample from the posterior distribution over the cluster parameters is overlayed on the extracted data over a sequence of frames, where the assignment of each data point is represented by color and marker type.
-
-\begin{figure}[!]
-  \centering
-  \subfloat[]{\includegraphics[width=0.43\textwidth]{../img/pets2009_1-50_front.pdf}}
-  \hspace{4mm}
-  \subfloat[]{\includegraphics[width=0.53\textwidth]{../img/pets2009full_pm_b.pdf}}
-  \caption{ (a) Results for the PETS2009 dataset, showing a sample from the posterior distribution of the state for frames 1-50, where the vertical axis denotes time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.
-(b) performance metrics vs.
-covariance confidence interval threshold for tracking results showing the SFDA (blue solid line) and ATA (red dashed line).}
-  \label{fig:pets2009_results}
-\end{figure}
-
-In \cite{ellis_2010}, performance metrics were computed for a number of studies that carried out object detection and tracking for the PETS2009/2010 dataset.
-As this dataset consists solely of humans, many of these algorithms were developed for the specific purpose of people tracking (i.e.
-not for general detection and tracking of arbitrary objects); in consequence, many of these studies involve appearance models specifically developed for humans, sometimes making use of their orientation in this specific dataset, and motion models based on assumptions about human motion.
-We compare SFDA and ATA results of our strategy with these methods to show that our general method can yield comparable results even when compared with object-specific trackers.
-Figure~\ref{fig:sfda_and_ata_comparison} shows performance metric results for comparison (data published with permission from the authors of \cite{ellis_2010}).
-\begin{figure}[!]
-  \centering
-  \subfloat[]{\includegraphics[width=0.49\textwidth]{../img/pets2009full_comp_sfda.pdf}}
-  \hspace{1mm}
-  \subfloat[]{\includegraphics[width=0.49\textwidth]{../img/pets2009full_comp_ata.pdf}}
-  \caption{Performance metric comparison showing the (a) SFDA and (b) ATA results for submissions to the PETS2009/PETS2010 workshops (blue bars) and for the technique introduced in this paper (red bars).
-These graphs show that the general technique in this paper yields comparable performance with state-of-the-art, object-specific detection and tracking strategies.
-Our results were computed using a covariance confidence interval threshold of 0.35.}
-  \label{fig:sfda_and_ata_comparison}
-\end{figure}
+\label{sec:pets20092010}
 
 \subsubsection{Sensitivity Analysis}
-\label{sec:sens_analysis}
+\label{sec:sensitivityanalysis}
 
-SMC inference on the PETS2009 video dataset was carried out for a range of two key parameter values, $\alpha$ and $\rho$.
-The performance metric measures, SFDA and ATA, were computed for each combination of these two parameters.
-This sensitivity investigation focused on these parameters due to their potential to have a large effect on object detection accuracies.
-The $\alpha$ values tested included $\{ 0.01, 0.1, 1, 10, 100 \}$, and the $\rho$ values tested included $\{ 0.7, 0.75, 0.8, 0.85, 0.9 \}$.
+% Conclusion
+% ----------
 
-The following shows the SFDA sensitivity analysis results. The largest value is in bold.
-\begin{center}
-\begin{tabular}[!]{l l  l  l  l  l  l}
-  & & & $\hspace{8mm}\rho$ & & & \\
-  SFDA  &  	& $0.7$ & $0.75$ & $0.8$ & $0.85$ & $0.9$  \\  
-  &$0.01$ 	& $0.4121$ & $0.4262$ & $0.4219$ & $\bold{0.4331}$ & $0.4277$  \\
-  &$0.1$  	& $\bold{0.4336}$ & $0.4184$ & $0.4200$ & $0.4316$ & $0.4321$  \\ 
-$\alpha$ 	& $1$    & $0.4084$ & $0.4209$ & $0.4205$ & $0.4290$ & $0.4330$  \\  
-  &$10$   	& $0.3992$ & $0.4130$ & $0.4289$ & $0.4232$ & $0.4226$  \\  
-  &$100$  	& $0.3655$ & $0.3739$ & $0.3636$ & $0.3755$ & $0.3672$  \\
-\end{tabular}
-\end{center}
-
-The following shows the ATA sensitivity analysis results. The largest value is in bold.
-\begin{center}
-\begin{tabular}[!]{l l  l  l  l  l  l}
-		& & & $\hspace{8mm}\rho$ & & & \\
- 	 ATA  & 	 	& $0.7$ & $0.75$ & $0.8$ & $0.85$ & $0.9$  \\  
-	  &$0.01$ 		& $0.2032$ & $0.2017$ & $0.2262$ & $0.1945$ & $0.1914$  \\
-		  &$0.1$ 	& $0.1964$ & $0.2042$ & $0.2401$ & $0.1990$ & $0.2017$  \\
-$\alpha$ & $1$  	& $0.2016$ & $0.2365$ & $0.2766$ & $0.2347$ & $0.2766$  \\   
-  		&$10$   	& $0.2277$ & $0.2284$ & $0.2613$ & $\bold{0.2781}$ & $0.2732$  \\  
- 		 &$100$ 	& $0.2468$ & $0.2645$ & $0.2725$ & $0.2670$ & $\bold{0.2843}$  \\
-\end{tabular}
-\end{center}
-
-% By inspecting the figures, one can see that there is a point where the SFDA and ATA both achieve a maximal value.
-
-\begin{figure}[!]
-  \centering             
-  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2009full_sens_sfda.png}}
-  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2009full_sens_ata.png}}
-  \caption{Surface plot of the sensitivity performance metric tables}
-  \label{fig:sens_surf}
-\end{figure}
-
-% \begin{figure}
-%   \centering             
-%   \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2009_result1.png}}
-%   \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2009_result2.png}}
-%   \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2009_result3.png}}
-%   \caption{Surface of the sensitivity performance metric tables}
-%   \label{fig:sens_surf}
-% \end{figure}
-
-% \begin{figure}
-%   \centering             
-%   \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_1}}
-%   \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_2}}
-%   \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_3}}
-%   \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_4}}
-%   \caption{Four frames $f=10, 25, 35, 75$ from the PETS2009 sequence with one posterior sample mean and covarance matrix overlaid.
-\willie{will show path trail over prev 10 frames}\willie{these are old images!...
-haven't replaced yet}}
-%   \label{fig:pets2009_imgs}
-% \end{figure}
-
-% \begin{figure}
-% 	\centering
-% 	\includegraphics[width=3in]{../img/pets2009_results}
-% 	% \subfloat[]{\includegraphics[width=0.52\textwidth]{../img/pets2001_results_1}}
-% 	% \subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2001_results_2}}
-% 	\caption{Sample from posterior of the state for the PETS2009 dataset, where color represents assignment and the mean and standard deviation are shown.
-\willie{will show assignments by marker type in addition to color}}
-% 	\label{fig:pets2009_results}
-% \end{figure}
-
-% \subsection{New Video Datasets}
-
-% The two video datasets in this section are video datasets first used in this paper. The datasets and groundtruth detailing the positions of each obect in each video will be published online.
-
-% \subsubsection{Ants Video?}
-
-% \subsubsection{Cells Video?}
-
-\section*{Discussion and Conclusion}
-\label{sec:discussion}
-
-We have presented a new technique for the unsupervised detection and tracking of arbitrary objects in videos.
-The primary intention of this technique is to reduce the need for detection or localization methods tailored to specific object types and serve as a general framework applicable to videos with varied objects, backgrounds, and film qualities.
-The GPUDDPM, a time-dependent Dirichlet process mixture, has been introduced, and we have shown how inference on this model allows for the desired clustering of extraction data.
-Furthermore, we have demonstrated a specific implementation of the model using spatial and color pixel data extracted via frame differencing and provided two algorithms for performing inference on the model to accomplish detection and tracking.
-Both algorithms were carried out on multiple synthetic and benchmark multi-object video datasets in order to demonstrate an ability to accomplish unsupervised detection and tracking of arbitrary objects in both manufactured and real world settings.
-We have described and computed standard performance metrics for our method's detection and tracking results, and found it to be comparable with other techniques that were designed for the specific object-types present in the videos in question.
-These results illustrate the ability of the technique described in this paper to, without modification, detect and track objects with diverse physical charactersitics moving over non-uniform backgrounds and through occlusion
-
-More to include: more discussion of parameter choices, thoughts about deletion params vs.
-dataset size (and the tradeoff between tracking through occlusion vs clusters rapidly gaining size), mvn cluster size vs.
-tracking in occlusion vs.
-sampling correct number of objects.
-discuss how we detect movement, which isn’t exactly same as authors in the perf.
-metric paper (ie our tracking is designed to stop if people stop moving).
-objects hard to track through occlusion becuase they are small, similar appearance, and our algorithm doesn’t assume dynamics to keep robust (we could, probably for better accuracy but less generality).
-background (appearance) is like a natural prior for remaining stationary
-
-%\section*{Conclusion}
+\section{Conclusion}
+\label{sec:conclusion}
 

### 2012-02-08 ltx/paper_short.tex
I really need to clean up this repo. sorry. the  main tex file is paper_short_2.tex

@@
 
-We define the task of automated detection and tracking of arbitrary objects in videos to consist of three actions: determining the spatiotemporal regions of a video that constitute objects (``extraction''), finding the positions and/or shapes of distinct objects (``localization''), and maintaining the identities of the detected objects over time (``tracking'').
+We define the task of automated detection and tracking of arbitrary objects in videos to be: extraction, localization, and tracking.
+Extraction is the task of determining the spatiotemporal regions of a video that constitute objects, localization is the task of finding the positions and/or shapes of distinct objects, and tracking is the task of maintaining the identities of the detected objects over time.
 An ability to carry out these tasks in an automated manner is useful for many fields that make use of video data, including robotics, video surveillance, time-lapse microscopy, and video summarization.
@@
 Integrating all three tasks in a system for multiple arbitrary objects and diverse video types was not often a primary focus.
-Accomplishing the three tasks in a cohesive manner has been pursued more recently and has produced algorithms that can, in a fully automated way, accomplish unsupervised detection and tracking---or in short, that can determine the spatiotemporal region of a video occupied by each distinct object.
+Accomplishing the three tasks in a cohesive manner has been pursued more recently and has produced algorithms that can, in a fully automated way, accomplish unsupervised detection and tracking---or in short, that can determine the spatiotemporal region of a video occupied by distinct objects.
 An example of general detection and tracking is called ``blob tracking'', which typically refers to methods that perform extraction and localization of objects in each frame independently.

### 2012-02-08 ltx/paper_short_2.tex
I really need to clean up this repo. sorry. the  main tex file is paper_short_2.tex

--- 2012-01-12 ltx/paper.tex
@@
 It is intended to reduce the need for detection or localization methods tailored to specific object types and serve as a general framework applicable to videos with varied objects, backgrounds, and film qualities.
-The technique uses a dependent Dirichlet process mixture (DDPM) known as the Generalized Polya Urn (GPUDDPM) to model image pixel data that can be extracted in a general manner from the regions in a video that represent objects.
+The technique uses a Dependent Dirichlet Process Mixture (DDPM) known as the Generalized Polya Urn (GPUDDPM) to model image pixel data that can be extracted in a general manner from the regions in a video that represent objects.
 This paper describes a specific implementation of the model using spatial and color pixel data extracted via frame differencing and gives two algorithms for performing inference on the model to accomplish detection and tracking.
@@
 \section{Introduction}
-
-We define the task of automated detection and tracking of arbitrary objects in videos to be the fulfillment of three actions: determining the spatiotemporal regions of a video that constitute objects (`extraction'), finding the positions and/or shapes of distinct objects (`localization'), and maintaining the identities of the detected objects over time (`tracking').
+\label{sec:introduction}
+
+We define the task of automated detection and tracking of arbitrary objects in videos to be three things: extraction, localization, and tracking.
+Extraction is the task of determining the spatiotemporal regions of a video that constitute objects, localization is the task of finding the positions and/or shapes of distinct objects, and tracking is the task of maintaining the identities of the detected objects over time.
 An ability to carry out these tasks in an automated manner is useful for many fields that make use of video data, including robotics, video surveillance, time-lapse microscopy, and video summarization.
 Algorithms able to be applied to a variety of objects and video types are particularly desirable.
-This paper serves to introduce a new technique involving the use of dependent Dirichlet processes mixture models, which we hope will provide the foundation for a class of unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.
-
-Research in the past ten years involving the development of methods for general detection and tracking of objects has tended to focus on one of either extraction, localization, or tracking; attempts to integrate all three tasks and produce a system aimed at multiple arbitrary objects and diverse video types was not often a primary focus, or was attempted in an ad-hoc manner to test or apply a more specific technique.
-Research aimed at developing methods to accomplish the three tasks in a cohesive manner has been pursued more recently to produce algorithms that can, in a fully automated way, accomplish unsupervised detection and tracking---or in short, that can determine the spatiotemporal region of a video occupied by each distinct object.
-A widely applied, though unsophisticated, example of general detection and tracking is the technique of blob tracking.
-Blob tracking typically refers to methods that perform extraction and localization of objects at each frame independently---for example, by assuming that pixels representing an object are distinguishable from pixels representing the background via color, motion, or texture information, and segmenting the extracted data at each frame into ``blobs'' to accomplish localization.
-A number of tracking techniques based on finding blobs in consecutive frames of a video with similar locations and appearances have been attempted to maintain the identity of detected blobs over time.
-The main challenges, however, associated with blob tracking involve the difficulty of segmenting neighboring or occluding objects in a given frame \cite{zhao2004tracking} and preventing multiple tracked blobs from collapsing onto and tracking a single object \cite{vermaak_2003}.
-These difficulties stem from the typical blob tracker's need to perform detection via segmentation on individual frames and lack of a principled way to perform tracking of multiple objects while resisting the collapse of independent tracks.
+This paper introduces a new framework involving the use of dependent Dirichlet process mixture models.
+This framework provides a foundation for a class of unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.
+
+Research on methods for general detection and tracking of objects tends to focus on one of either extraction, localization, or tracking.
+Integrating all three tasks in a system for multiple arbitrary objects and diverse video types is not often a primary focus.
+Accomplishing the three tasks in a cohesive manner has been pursued more recently and has produced algorithms that can, in a fully automated way, accomplish unsupervised detection, localization, and tracking.
+An example of general detection and tracking is called ``blob tracking'', which typically refers to methods that perform extraction and localization of objects in each frame independently.
+Tracking involves finding blobs with similar locations and appearances over time.
+The main blob tracking challenges are the difficulty of segmenting neighboring or occluding objects in a given frame \cite{zhao2004tracking} and preventing multiple tracked blobs from merging into a single object \cite{vermaak_2003}.
+These difficulties stem from performing detection via segmentation on individual frames.
 
 This paper presents the use of a Bayesian nonparametric mixture model for general, unsupervised multiple object detection and tracking.
-This method is intended to provide a widely applicable way to detect arbitrary objects in videos---particularly in cases where frame-by-frame segmentation is difficult or video quality is low and extraction is noisy---and better maintain the isolation of independent objects during tracking.
-Specifically, this work describes how a Dirichlet process mixture model can accomplish these goals when it is applied to data that may be extracted in a general manner from videos.
-We will demonstrate how to perform this extraction and how this generative model allows for automated detection and tracking.
-The model is first presented in an abstract form, to allow it to be used with arbitrary extraction procedures and object appearance representations that may be desired in future studies.
-We then formulate a specific model for data gained during a basic extraction procedure and describe inference algorithms that allow object localization and tracking to be carried out.
-We demonstrate our implementation on multiple synthetic and benchmark datasets and compute a few standard performance metrics for each to quantify results.
-
-% related work
-% ------------
-
-\subsection{Related Work}
-\label{sec:relatedwork}
-
-Related work involves studies related to extraction procedures, localization techniques, single and multiple object tracking methods, and combination methods for joint object detection and tracking.
-
-Extraction is related to the tasks of foregound segmentation, background modeling, motion detection, and appearance feature extraction.
-In general, these methods aim to extract features from regions of a video frame that are considered objects or are distinguished in some way from the background of a scene.
-Areas of research in this field include attempts to handle situations where the background displays motion or other dynamic characteristics, reduce noise caused by the false extraction of backgrounds, handle abberant image effects present in some videos, and operate accurately on scenes involving illumination changes.
-These procedures can usually be classified as either model-based methods or heuristic-based methods.
-Model-based methods emphasize the detection of shapes of regions containing objects accurately and reducing background noise.
-For example, foreground extraction has been robustly carried out by modeling each pixel over time in order to infer a probability distribution over pixel values; if the pixel takes on a very unlikely value at a future time, it is marked as a foreground pixel at that point \citep{stauffer_1999, elgammal_2000, elgammal_2002}.
-Note that these techniques treat each pixel independently; this has a tendency to cause errors in foreground detection, resulting in fragmentation of foreground images (which is problematic as many of these methods intend to blob the foreground objects after detection in order to perform segmentation and locate objects' centroids).
-Heuristic-based methods perform extracting by looking for temporal changes in pixel value characteristics.
-These methods are often less accurate than model-based methods, but also less computationally expensive.
+This method provides both a widely applicable way to detect arbitrary objects in videos---particularly in cases where frame-by-frame segmentation is difficult or video quality is low and extraction is noisy---and a general framework for identifying and isolating objects during tracking.
+Specifically, this work describes how a dependent Dirichlet process mixture model can accomplish these goals.
+The DDP is first reviewed in the abstract.
+This model can be used with arbitrary extraction procedures and object appearance representations that may be desired in future studies.
+We then present our specific model for data obtained via a basic extraction procedure.
+We describe inference algorithms for the DDP, and specifically our specific model, that allow object localization and tracking to be carried out.
+We demonstrate our implementation on multiple synthetic and benchmark datasets.
+Standard performance metrics for each supports our hypothesis that general extraction combined withe sophisticated inference can perform detection and tracking at a level comperable to state-of-the-art, object specific algorithms.
+
+% Prior Work
+% ----------
+
+\section{Background}
+\label{sec:priorwork}
+
+A variety of methods in the fields of image processing, signal processing, and computer vision have been developed to solve aspects of the problem of unsupervised detection and tracking of arbitrary objects.
+These methods might be placed into a few broad categories: those that aim to distinguish the foreground regions of images from the background \cite{hong2007real,chien2002efficient, zhang2007moving, kim2002fast}, segment images into distinct regions to perform localization \cite{jain1997object, fei2005bayesian, sivic2005discovering}, track an object over a sequence of images (after its position has been specified in an initial image) \cite{raja_1998, mckenna_1999, jepson_2003, comaniciu_2003, perez_2002}, track multiple objects over a sequence of images (especially when the objects interact or occlude one another) \cite{senior2006appearance, cucchiara2004probabilistic, zhou2003background, han_2004, mckenna2000tracking, dockstader2001multiple}, segment a sequence of images into distinct spatiotemporal regions \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}, and combine the previous methods in some way to create systems capable of both detecting and tracking specified objects, or of discerning which regions of a video constitute distinct, arbitrary objects and tracking these \cite{brostow2006unsupervised, brox2010object, fragkiadaki2011detection, pece_2002}.
+
+Extraction methods aim to discern the foreground regions of a video from the background.
 Frame differencing and background subtraction are two such methods, and involve comparing pixel values, or groups of pixel values, between two images, to detect a change surpassing some threshold.
@@
 Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking schemes \citep{perez_2002}.
-In addition, there exist other ways of extracting movement from an image, such as through techniques involving optical flow (a calculation based on the `generalized gradient model' of an image, which attempts to capture the speed and direction of movement over areas in the image) \citep{horn_1981, bobick_2001}.
-In \citep{black_2000}, a motion detection and object tracking method is developed based on the relative motion of moving objects' boundaries, which are again found through analysis of optical flow.
-
-Localization is related to the topics of object detection, position-finding, segmentation, and target recognition.
-These methods are mainly concerned with finding the positions and/or shapes of objects in a given video frame, or with discerning between objects similar in position or appearance.
-There has been a focus in this field on developing techniques for representing an object's appearance, both to allow for localization of objects on a per-frame basis and to allow for consistent localization of a given object over time.
-Research into the unsupervised detection of arbitrary objects (which we define as methods that accomplish both extraction and localization) has included work involving model-based clustering of individual pixels, local image texture representations, or other patterns found in image regions \cite{jain1997object, fei2005bayesian, sivic2005discovering}.
-These works usually aim to automatically segment images into meaningful regions.
-Note that the popular phrase `object detection' is sometimes used to denote methods that are tailored to find the position of specified objects or object types (as opposed to arbitrary object detection, which incorporates both the task of determining what constitutes distinct objects and localizing these objects), or to denote supervised methods, where extraction is peformed implicitely during searches for specified objects and localization procedures.
-
-Tracking is related to the topics of video tracking, data association, filtering, point matching, and `tracking of non-rigid objects'.
-In general, these tasks aim to maintain the localization of the position and/or shape of an object over time.
-Areas of interest in this field involve developing metrics for judging the similarity between two object representations (for example, potential representations of the same object at adjacent frames in a video), techniques that provide a way to conduct efficient searches over a frame for regions that are similar to a given object, and methods for predicting the future position or appearance of an object.
-An overview of object tracking methods can be found in \cite{yilmaz2006object}.
-Methods for tracking single, non-rigid, objects are numerous.
-Color, for example, has long been used for tracking; these techniques often operate by modeling an object with some color-based appearance model, and using this model to find the object in subsequent frames.
-One classic approach involves modeling the color of a target (in particular, a distribution over the hue-saturation space of a target region) with a Gaussian mixture model (GMM), and choosing subsequent target positions by searching surrounding areas for regions that yield similar GMMs.
-Furthermore, the GMM is often allowed to adapt over time, and slowly change to model changes in lighting or other time-based variations in the target's color \citep{raja_1998, mckenna_1999, jepson_2003}.
-Additionally, color has been successfully applied as the feature in a technique known a the mean-shift procedure (or the kernel-based method) for tracking \citep{comaniciu_2003, perez_2002, nummiaro_2003, lee_2011}.
-This procedure---a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--optimizes the search for the region in a frame which is most similar to a target region in the previous frame.
-It involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges to a number of modes.
-Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors.
-It can be shown that, through this procedure, each data point becomes associated with a local point of high density (dependent upon the underlying weight distribution specified by the kernel) which allows for clustering \citep{cheng_1995}.
-The mean shift algorithm has been implemented successfully to allow for a sort of kernel-based object tracking \citep{comaniciu_2003, comaniciu_1999, comaniciu_2000}.
-
-In addition to single object tracking, a great deal of research in this field over the past decade has focused on the development of algorithms to track multiple objects simultaneously.
+Another technique for discerning moving regions in an image involves optical flow \citep{horn_1981, bobick_2001}.
+In \cite{black_2000}, a motion detection and object tracking method is developed based on the relative motion of moving objects' boundaries, which are found through analysis of optical flow.
+
+A great deal of research has focused on developing algorithms to track multiple objects simultaneously.
 There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera) \cite{senior2006appearance, cucchiara2004probabilistic, zhou2003background}, complex object interactions \cite{khan_2004, mckenna2000tracking, dockstader2001multiple}, objects with similar appearances \cite{maccormick1999probabilistic, jepson_2003}, variable (and potentially high) numbers of objects \cite{reilly2010detection}, and objects that enter and exit a field of view at different times \cite{stauffer2003estimating, nedrich2010learning}.
@@
 Methods involving blob tracking, however, run into problems when faced with videos where detection is difficult, object appearance or orientation varies heavily, and there exists object occlusion \cite{song2005model}.
-To aid the accuracy of these methods, techniques have been developed for performing extraction and segmentation in a joint manner, incorporating statistical methods for maintaining hypotheses of different numbers of detected objects, and introducing some of the tracking methods described previously to track distinct blobs after they have been segmented \cite{collins2003mean, isard_2001}.
+To aid the accuracy of these methods, techniques have been developed for performing extraction and segmentation in a joint manner, incorporating statistical methods for maintaining hypotheses of different numbers of detected objects, and introducing some of the multi-object tracking methods described previously to track distinct blobs after they have been segmented \cite{collins2003mean, isard_2001}.
 Another family of methods related to the task of unsupervised detection and tracking of video objects goes under the heading of video segmentation algorithms---these methods extend single-frame image segmentation to maintain coherence of image segments over time, and have had some success when used for the explicit purpose of detecting and tracking foreground objects of videos \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}.
@@
 Another approach based on a Gaussian mixture model has also been used to cluster data gained from moving objects \cite{pece_2002}; this method also develops heuristics for the initialization and elimination of new tracks.
-The technique we introduce in this paper falls into this category of methods, as it aims to provide a general unsupervised detection and tracking method for arbitrary objects in videos via clustering.
+
+Nearly all methods capable of high accuracy detection and tracking of objects in videos are those tailored for specific object types.
+These methods rely on ad-hoc detection criteria that exploit previous knowledge about the appearance or behavior of objects in a video.
+The method described in this paper is designed to track arbitrary objects without using any explicit detection criteria, and serve as a general strategy that can be used, without modification, to perform accurate detection and tracking of diverse objects in a wide range of videos.
+The method we introduce falls into the category of clustering-based arbitrary object detection and tracking methods.
 Differing from previous work, we use a type of time dependent Bayesian nonparametric mixture model, and show how it can be applied to a variety of extraction data to perform detection and tracking.
-
-% extra background that I decided to cut (for now)
-% ------------------------------------------------
-
-% Others have attempted to abstract this work with a non-parametric color modeling approach based on kernel density estimation, which does not assume a specific underlying distribution (such as the Gaussian mixture in the previous case) and instead converges to reasonable distribution that depends on the data \citep{elgammal_2001}.
-
-% Extraction
-%
-
-% The mean-shift procedure (also known as `kernel-based object tracking') attempts to provide a robust way for non-rigid objects to be tracked.
-This method is beneficial because it optimizes the search for the `next' position of an object (i.e.
-the search for the region in a frame which is most similar to a target region in the previous frame).
-This procedure---a derivation and overview of which can be found in \citep{fukunaga_1975, cheng_1995}--involves an iterative algorithm that repeatedly shifts each data point to the weighted average of data points in its neighborhood; it has been proven that this process converges for each data point.
-Additionally, the process (in particular, the specification of the neighborhood and the weight-distribution when calculating the weighted mean of nearby data points) is generalized so that multiple kernels can be specified and used to allow for varied clustering behaviors.
-It can be shown that, through this procedure, each data point becomes associated with a local point of high density (dependent upon the underlying weight distribution specified by the kernel) which naturally allows for clustering \citep{cheng_1995}.
-The mean shift algorithm has been implemented successfully to allow for a sort of kernel-based object tracking \citep{comaniciu_2003, comaniciu_1999, comaniciu_2000}.
-Given the current position of an object at a given frame, the goal of tracking is often to find a nearby position in the next frame that has the most similar distribution over some common set of features.
-Usually this must be done through an exhaustive search, comparing the similarity of distributions at each nearby position with that at the current position.
-However, if a certain type of `isotropic kernel' (mean-shift kernel) known as the Bhattacharyya coefficient is chosen as a similarity metric between the feature distributions at two positions, it creates a smooth function where gradient descent techniques can be used to quickly converge upon an optimal subsequent position without using an exhaustive search.
-Many papers are concerned with applying this kernel-based object tracking scheme to different sets of features or with different kernels.
-Additionally, the scheme has been attempted with an adaptive kernel whose shape, scale, and orientation is influenced by a target being tracked \citep{wang_2004}, with a kernel adjusted by the estimated centroid of a tracked target \citep{mehmood_2009}, and with a heirarchical version of the mean-shift procedure \citep{dementhon_2002}.
-
-% Some have tried to extent single object tracking schemes to tracking multiple targets; at the most basic level, this involves initializing multiple targets and running an instance of (single) VOT for each, either simultaneously or in succession \citep{perez_2002}.
-Furthermore, if performed simultaneously, the tracking of each target can be made dependent upon characteristics of other targets (such as their proximity) to resolve errors and improve tracking (such as those caused by the incorrect merging of two targets) \citep{khan_2004, vermaak_2003} .
-
-% more background that I decided to cut (for now)
-% -----------------------------------------------
-
-% % Localization-focused. Clustering methods, segmentation methods, object appearance models and representations (allows for consistent object localization over time).
-
-% % Methods for tracking single objects in videos: filtering, similarity metrics / localization (and how they play a role), mean shift procedures.
-
-% % extensions to multiple visual object tracking.
-problems with multiple independent single object trackers (coalesence, object interaction, object occlusion, objects entering and leaving scene [and, relatedly, variable numbers of objects]).
-% % Multiple object tracking attempts at overcoming various the various challenges specified above (include nonparametric varmaak/doucet for maintaining multi-modality here?---sure, and then mention again later with fox's work?)
-% % End with description of initialization problem.
-
-% % Methods for detection and tracking of objects.
-Heuristic initialization at beginning and then carrying out above methods (problems with varying numbers of obects / objects entering or leaving scene).
-Supervised learning of specific targets to allow for initialization and provide better localization throughout.
-Various research has been towards integrating unsupervised extraction and localization into multiple object tracking methods.
-Particle filters with integrated detection (at least two of these).
-Introduce concept of blob tracking and define basic algorithm (some manner of extraction, independent/per-frame localization, tracking; note that each object is represented in each frame as a low-dimensional point [or higher dimensional thing like region/outline with sophisticated blob/contour detectors like bramble and others]).
-The key thing characterizing a blob detector is that the position of each object (or the spatially separate region associated with each object) is assigned/fixed at each frame without the aid of tracking (ie if localization/segmentation can be carried out at each frame independently), and tracking is done with this data.
-Schemes often are simply segmentation of extraction data into spatially disjoint blobs at each frame (where blobs  are not modeled in any principled manner), returning the position or region associated with each object at each frame.
-
-% % When the state of objects (for example, position and/or other characteristics) can be found independently at each frame, well established methods in the field known as Multiple Target Tracking (MTT) can be carried out to perform tracking of multiple objects.
-Two classic algorithms for MTT are known as MHT and JPDA.
-In general, MTT methods are useful for procedures that can represent each object at each frame as a point or region without localization information from this object at a previous frame (which requires the use of tracking).
-MTT in the literature is often quite separate from fields of computer vision involving object detection and tracking research.
-In general, the situation arises as video data without modification is very high dimensional (discussion of difference between MTT and detection methods here?).
-
-% % Methods like blob tracking which attempt to find the state of each object independently at each frame have a hard time dealing with occlusion and other complexities (except for sophisticated blob trackers like bramble which isn't really blob tracking).
-Attempts have been made to perform time dependent clustering on extraction data (look at cluster tracking paper for good def), to carry out joint localization and tracking; this has been refered to as cluster tracking.
-Work of Pece, MTT of multiple targets work(?).
-The work presented in this paper most closely resembes the topics listed here.
-Does cluster tracking on whichever extraction data, using a time dependent Dirichlet Process mixture model (to better estimate and maintain the number of targets, as was shown to be effective in vermaak and fox), which takes the form of a sequence of clusters, where the objects' appearances are modeled in a principled manner by the cluster component distributons, and the tracking behavior is modeled by time-dependencies between adjacent elements in the cluster-sequences.
-
-% Contributions (decided to cut for now)
-% --------------------------------------
-
-% \subsection{Contributions}
-
-% This paper applies a time dependent Dirichlet process mixture to model data gained during extraction procedures, provides background as to how this Bayesian nonparametric model allows for the incorporation of prior knowledge about object appearances and motion behaviors, describes methods for unsupervised detection and tracking of arbitrary objects in terms of Bayesian inference procedures, and presents a specific implementation of the model and inference procedures to carry out detection and tracking on a number of target-types in different settings.
-
-% % List here the benefits of this idea that should be discussed/listed:
-% This paper is heavily model-based, which differs from many papers in the field of multiple object tracking that do not explicitely describe models which underlie various procedures that are carried out.
-Solutions in many papers are, instead, algorithm oriented.
-The model-based approach allows rigor to be kept as inference procedures are developed.
-Furthermore, development of inference procedures are performed separately from development of the model, and various well studied inference and approximate inference methods can be carried out to achieve algorithmic results that are (in some cases) known to be optimal.
-% %treated separately, which allows for multiple methods to be carried out in a unified framework and   these solutions are  but provide algorithm-based presentations.
-% %	This way allows rigor to be kept as inference procedures are developed.
-% %	Furthermore, inference is treated separately; this allows multiple detection and tracking algorithms to be developed (in a well-formed/rigorous fashion) by developing different inference techniques.
-Also, inference on graphical models is a well-structured field and allows for use of techniques in this field as well as better interpretations of the resulting detection/tracking algorithms.
-
-% % (related to previous?) models objects as distribution-components (of a mixture model) as opposed to point-targets or targets with noise that adheres to some distribution (is this truly a distinction and is it good?)
-
-% This idea presents a framework which allows for time-varying models of objects' appearances and motion behaviors to be incorporated, and is formulated in a Bayesian manner that allows distributions underlying the parameters of these models to be specified in a principled manner.
-It has been shown to work successfully in conjunction with simple (noisy) extraction methods.
-More robust, expensive, or in general accurate methods can be used for further accuracy.
-This method is able to handle the appearance and disappearance of objects from a scene, variable numbers of objects, cases where objects show time varying changes in behavior, and cases of occlusion.
-
-% Section two of this paper provides an overview of the model and describes how aspects can be modified for for modeling object appearance and motion behavior.
-Section three presents a specific implementation of the model and a number of inference procedures for carrying out object detection and tracking.
-Section four describes a number of experiments carried out with this implementation.
-Section five describes results of these experiments, comments on further directions of study, and provides concluding remarks.
-
-% general model overview section
-% ------------------------------
-
-\section{Model Overview}
-\label{sec:modeloverview}
-
-This section provides details on the data gained during extraction procedures, and on Dirichlet process mixture (DPM) models, dependent Dirichlet process mixture (DDPM) models, and the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM) model.
-We make explicit the way in which characteristics of video objects can be represented by elements of the GPUDDPM.
-We formulate the model in a sufficently abstract manner to allow for implementations in future studies that may use more-specific object representations or extraction procedures.
-
-% general data extraction section
-% -------------------------------
-
-\subsection{Data Extraction}
-\label{sec:data}
-
-We defined the task of extraction to be the unsupervised determination of which spatiotemporal regions of a video constitute objects, and outlined a number of extraction methods in Section~\ref{sec:relatedwork}.
-Extraction allows us to generate a collection of data points.
-Each data point can be thought of as a hypothesis about, or noisy version of, the representation of an object; a collection of hypotheses for a given object is assumed to be distributed in some way around the true object representation.
-A video of multiple objects is therefore assumed to yield data that contains multiple groups of object hypotheses with similar characteristics (i.e.
-clusters of similar hypotheses).
-
-Determining the regions of each video frame that contain objects allows us to generate these object hypotheses; namely, we let each hypothesis be a description of the pixel data over an area centered at a point in the extracted region.
-We can, for example, let each pixel in the extracted region be a hypothesis, and let the features of the representation consist of the spatial position of the pixel.
-An extraction dataset of this form is comprised of the three-dimensional points $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $x_{1}$ and $x_{2}$ are the two spatial dimensions of a pixel contained in an extracted region in a given frame, and $t$ is the discrete time index of the frame.
-These three features are always available once regions containing objects are discerned, and are useful features for modeling (as position and time information is key for reasonable tracking).
-
-Examples of additional features that could be extracted from object-regions in video frames include color information, pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, texture representations, transform calculations (such as results of hough or fourier transforms), and quantiative region descriptions derived from image segmentation procedures; the main idea is to choose features which are able to be easily extracted or computed, capture variability in the appearance of objects, are applicable to a wide variety of object types, and can be represented by distributions that allow for tractable inference (discussed further in Section~\ref{sec:inference}).
-
-We will use the phrase ``baseline'' three-dimensional extraction data to refer to the features $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$ of pixels present in regions containing objects.
-These data are distributed roughly as worm-like shapes that follow the paths of individual objects as they move, when the data are plotted in three dimensions (where we represent the frame number $t$, our discrete representation of time, on the vertical axis and spatial position on the horizontal axes).
-Modeling this data with a time dependent Dirichlet process mixture model (Section~\ref{sec:gpuddpm}) allows us to perform inference (Section~\ref{sec:inference}), which carries out clustering of the data.
-By choosing this type of model, we aim to provide a clustering result that isolates each worm-like data cloud and infers a sequence of distributions for each, which we use to gain localization and tracking results for each object (Section~\ref{sec:inference_to_results}).
-
-% this is the pretty worms figure that we decided to cut
-% ------------------------------------------------------
-
-%An example of data with such a distribution can been seen in Figure~\ref{fig:baselinedata}...The data consists of the three-dimensional baseline features and was gained by extraction peformed on a 50 frame sequence showing five ants moving across the video scene.
-
-% frank says the following worm figures are a no go.
-% \begin{figure}[h]
-%   \centering
-%   \subfloat[]{\label{fig:baselinedata1}\includegraphics[width=0.33\textwidth]{../img/pets2001_data_1.png}} 
-%   \subfloat[]{\label{fig:baselinedata2}\includegraphics[width=0.33\textwidth]{../img/usbh_data_3.png}} 
-%   \subfloat[]{\label{fig:baselinedata3}\includegraphics[width=0.33\textwidth]{../img/traffic2_data_1.png}}\\
-%   \subfloat[]{\label{fig:baselinedata4}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_1.png}} 
-%   \subfloat[]{\label{fig:baselinedata5}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_2.png}} 
-%   \subfloat[]{\label{fig:baselinedata6}\includegraphics[width=0.33\textwidth]{../img/pets2009_data_3.png}}
-%   \caption{Baseline extracted features from the (a.) PETS2001 Dataset, (b.) Ants Dataset, (c.) Traffic Dataset, and PETS2009/2010 benchmark dataset, frames (a.) 1-200, (b.) 201-400, (c.) 401-600.
-Time is shown on the vertical axis.
-Marker color corresponds with spatial coordinates and is shown for ease of viewing.}
-%   \label{fig:baselinedata}
-% \end{figure}
-
-% section providing background to introduce DPMs
-% ----------------------------------------------
-
-\subsection{Dirichlet Process Mixtures}
-\label{sec:dpm}
-
-Dirichlet process mixtures fall under the heading of Bayesian nonparametric (BNP) models.
-These models have been widely used over the past decade to perform nonparametric density estimation and cluster analysis.
-They are, in particular, useful for estimating the number of latent classes (clusters) in mixture models.
-In this work, estimating the number of clusters in the data generated via extraction is equivalent to estimating the number of distinct objects in a video.
-Consequentially, a model that is able to perform robust nonparametric estimation of the clusters in data gained during extraction may be applied to perform unsupervised object detection and tracking.
-
-The following sections are intended to provide an introduction to DPMs without a major foray into the theoretical aspects of the model that require a more rigorus technical setup.
-We provide background on finite mixture models, Bayesian finite mixture models, and Dirichlet processes, which allows us to subsequently describe DPMs in context.
-
-% defining the finite mixture model
-% ---------------------------------
-
-\subsubsection{Finite Mixture Model}
-\label{sec:finite_mixture}
-
-A finite mixture model can be thought of as a probability distribution for an observation $x_{i}$ formulated as a linear combination of K mixture components (which we also refer to as `clusters'), where each mixture component is a probability distribution for $x_{i}$ with some parametric form, and the coefficients of the linear combination sum to one.
-The finite mixture model can be written as
-\begin{equation}
-P(x_{i}) = \sum_{k=1}^{K} P(c_{i} = k)P(x_{i}|\theta_{k})
-\end{equation}
-$\forall i \in \{ 1, \ldots, N \}$, where $c_{i} \in \{ 1, \ldots, K \}$ denotes the assignment of $x_{i}$ to a given mixture and $\theta_{k}$ denotes the parameters of the $k^{\text{th}}$ mixture component.
-Note that by choosing $P(c_{i} = k)$ as coefficients of the linear combination, it is ensured that these coefficients sum to one.
-We also define $p_{k} := P(c_{i} = k)$ for $k \in \{ 1, \ldots, K \} $.
-We can therefore write this model generatively as
-\begin{align}
-\begin{split}
-	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
-	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}})
-\end{split}
-\end{align}
-$\forall i \in \{ 1, \ldots, N \}$, where the $x_{i}$ are observations, the $c_{i}$ are the mixture component assignments associated with each observation, the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e.
-the distribution to be mixed, $F(\theta_{c_{i}})$), and the ``Discrete'' distribution refers to a multinomial distribution whose parameters are a 1-of-K vector (i.e.
-a vector of counts that sums to one).
-
-% defining the Bayesian mixture model
-% -----------------------------------
-
-\subsubsection{Bayesian (Finite) Mixture Model}
-\label{sec:bayesian_mixture}
-
-The finite mixture model of the previous section can be extended to a Bayesian mixture model by viewing parameters that were previously point values, $\theta_{c_{i}}$ (the mixture component parameters) and $p_{1}, \ldots, p_{K}$ (the mixture component assignment weights), as random variables and providing each with a prior distribution.
-In this case, the prior distribution $\mathbb{G}_{0}$ is placed on the mixture component parameters, and the prior distribution $\text{Dir}(\alpha/K, \ldots, \alpha/K)$ is placed on the mixture component assignment weights.
-The resulting Bayesian mixture model can be formulated generatively as
-\begin{align}
-\begin{split}
-\label{bayesian_mixture_model}
-	p_{1}, \ldots, p_{K}  &\sim  \text{Dir}(\alpha/K, \ldots, \alpha/K)\\
-	\theta_{1}, \ldots, \theta_{K}  &\sim  \mathbb{G}_{0} \\
-	% \theta_{k}  &\sim  \mathbb{G}_{0}, \hspace{2mm} \forall k \in \{1, \ldots, K\} \\
-	c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
-	x_{i}|c_{i}, \theta_{c_{i}}  &\sim  F(\theta_{c_{i}})
-\end{split}
-\end{align}
-$\forall i \in \{ 1, \ldots, N \}$, where the $x_{i}$ are observations, the $c_{i}$ are the mixture component assignments associated with each observation, the $\theta_{k}$ are parameters defining the $k^{\text{th}}$ mixture component (i.e.
-the distribution to be mixed, $F(\theta_{k})$), the $\theta_{k}$ are drawn from a prior distribution $\mathbb{G}_{0}$, and $p_{1}, \ldots, p_{K}$ are drawn from a Dirichlet prior parameterized by $\alpha/K, \ldots, \alpha/K$.
-
-% defining Dirichlet processes
-% ----------------------------
-
-\subsubsection{Dirichlet Process}
-\label{sec:dirichlet_process}
-
-% Define Dirichlet distribution first?
-Each draw from a K-dimensional Dirichlet lies on the K-simplex (i.e.
-is a vector of K values which sum to 1) and can therefore be thought of as a K-dimensional discrete distribution.
-
-The Dirichlet process (DP), first introduced by \cite{ferguson_1973} in 1973, may be intuitively viewed as a probability distribution over discrete probability distributions.
-Accordingly, draws from a DP are probability mass functions (PMFs).
-A DP is parameterized by a base distribution $\mathbb{G}_{0}$, which is a probability distribution over a set $\Theta$, and a concentration parameter $\alpha \in \mathbb{R}_{+}$.
-We say that $G$ is a random PMF distributed according to a DP, written $G \sim \text{DP}(\alpha, \mathbb{G}_{0})$, if the following holds for all finite partitions $A_{1}, \ldots, A_{p}$ of $\Theta$:
-\begin{equation}
-(G(A_{1}), \ldots, G(A_{p})) \sim \text{Dir}(\alpha \mathbb{G}_{0}(A_{1}), \ldots, \alpha \mathbb{G}_{0}(A_{p}))
-\end{equation}
-Where `Dir' denotes a Dirichlet distribution.
-The parameters $\mathbb{G}_{0}$ and $\alpha$ may be intuitively viewed as the mean and precision of the DP.
-This is due to the fact that if the base distribution $\mathbb{G}_{0}$ is a distribution over $\Theta$, $A \subset \Theta$, and $G \sim \text{DP}(\alpha, \mathbb{G}_{0})$, then the following holds:
-\begin{equation}
-\mathbb{E}[G(A)] = \mathbb{G}_{0}(A)
-\end{equation}
-\begin{equation}
-\text{Var}[G(A)] = \mathbb{G}_{0}(A) (1 - \mathbb{G}_{0}(A)) / (\alpha + 1)
-\end{equation}
-Hence, the expectation of $G(A)$ is $\mathbb{G}_{0}$, the variance of $G(A) \rightarrow 0$ as $\alpha \rightarrow \infty$, and $G$ converges pointwise to $\mathbb{G}_{0}$ when $\alpha$ is unbounded.
-
-% defining the DPM
-% ----------------
-
-\subsubsection{Dirichlet Process (Infinite) Mixture Model}
-\label{sec:dirichlet_process_mixture}
-
-A Dirichlet process mixture model, also refered to as an infinite mixture model, is an extension of the Bayesian mixture model described previously. 
-When using a DP as a prior in a Bayesian mixture model, $\Theta$ represents the set of parameters of the component mixture distributions.
-A DPM may be viewed as allowing the prior distribution over the mixture component parameters in a standard mixture model to be distributed according to a DP; this allows for modeling data where the true number of latent mixture components is unknown and arbitrarily large by letting the number of components remain unbounded (note that only a finite number of these components are assigned to the data).
-In particular, the DPM can be defined generatively as
-\begin{align}
-\begin{split}
-	\mathbb{G} | \alpha, \mathbb{G}_{0}  &\sim  \text{DP}(\alpha, \mathbb{G}_{0}) \\
-	\phi_{i} | \mathbb{G}  &\sim  \mathbb{G} \\
-	x_{i}|\phi_{i} &\sim F(\phi_{i})
-\end{split}
-\end{align}
-$\forall i \in \{ 1, \ldots, N \}$, where the $x_{i}$ are observations, the $\phi_{i}$ are parameters defining the mixture component from which the $i_{th}$ observation is drawn (i.e.
-the distribution to be mixed, $F(\phi_{i})$), and the $\phi_{i}$ are drawn from a prior distribution $\mathbb{G}$, which is in turn drawn from a DP with base distribution $\mathbb{G}_{0}$ and parameter $\alpha$.
-See \cite{gasthaus_2008} and \cite{gasthaus_thesis} for more details on this formulation.
-Note the difference between the indexing of the clusters in this model and the indexing in the previous two models.
-This formulation can be shown to be equivalent to the Bayesian mixture model defined in \eqref{bayesian_mixture_model}, when K is taken to be unbounded; as a result, this model is sometimes called an infinite mixture model.
-
-If we let $K$ be the number of distinct mixture components assigned to observations using the above model, we can write the mixture components as $\theta_{1}, \ldots, \theta_{K}$.
-Let $c_{1}, \ldots, c_{N}$ (where $c_{i} \in \{1, \ldots, K \}$) be class assignment variables that indicate the cluster to which observation $x_{i}$ is assigned.
-We can now formulate the Chinese restaurant process, a discrete-time stochastic process that defines a partition of the set $\{ x_{1}, \ldots, x_{N} \}$ (via the elements' assignments $c_{1}, \ldots, c_{N}$).
-The Chinese restaurant process allows samples to be drawn from the conditional distribution over the assignment variables $c_{i}$, and can be expressed as
-\begin{align}
-\begin{split}
-\label{crp_rep}
-	P(c_{i} = c_{j} \text{  for some  } j<i) &= \frac{m_{k}}{i-1+\alpha}\\
-	P(c_{i} \neq c_{j} \text{  for all  } j<i) &= \frac{\alpha}{i-1+\alpha}
-\end{split}
-\end{align}
-where $m_{k}$ is the cardinality of the set $\{ c_{j} | (c_{j}=c_{i}=k)  \wedge  (j < i) \}$ and $\alpha$ is the parameter of the DP prior on $\mathbb{G}$.
-A modified Chinese restaurant process will be used to draw samples from the conditional distribution over the assignment variables when performing inference on the GPUDDPM (Section~\ref{sec:gpuddpm}).
-
-% section providing background to introduce DDPMs
-% -----------------------------------------------
-
-\subsection{Dependent Dirichlet Process Mixtures}
-\label{sec:ddpm}
-
-The goal of dependent Dirichlet process mixtures is to allow modeling of data that is not independent and identically distributed (i.i.d) but instead has some underlying dependencies.
-For example, data generated during extraction procedures from videos have some associated temporal dependencies, since tracked objects display time dependent characteristics (i.e.
-there are similarities among features, such as the spatial positions or appearances of objects, for data at nearby time steps).
-
-To account for the dependent behavior of data, there has been research on developing models involving a sequence of DPMs, where components of the mixtures are dependent upon (or are sometimes said to be ``tied to'') corresponding components at neighboring positions in the sequence.
-For example, if the data shows temporal dependence, the goal might be to create a sequence of DPMs, one for each time-step, where the components of the mixture at each step are dependent upon corresponding components in the both the following and previous time steps.
-
-More rigorously, we take the definition of a DDPM to be a stochastic process defined on the space of probability distributions over a domain, which are indexed by time, space, or a selection of other covariates in such a way that the marginal distribution at any point in the domain follows a Dirichlet process (adapted from definitions found in \cite{gasthaus_thesis} and \cite{griffin2006order}).
-Hence, a time-dependent DDPM is a model which remains a Dirichlet process, marginally, at each time step, yet allows cluster parameters at a given time step to vary from (and remain dependent upon) the parameters in neighboring time steps.
-
-%%%%%
-% add more info here on other DDPMs, why we chose the GPUDDPM, how it has an ability to model birth and death of clusters (unlike some sort of sophisticated hidden markov model)
-%%%%%
-
-% defining the GPUDDPM
-% --------------------
-
-\subsubsection{Generalized Polya Urn Dependent Dirichlet Process Mixture}
-\label{sec:gpuddpm}
-
-Data gained by performing extraction on videos containing objects is significantly time-dependent.
-Futhermore, the clusters of data, each representing a video object, might change in number as time progresses (which is to say, clusters may get `born' and may `die' at some intermediate time step), since objects can enter and exit a scene.
-To handle these challenges, the specific DDPM chosen to model extraction data in this work is known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM), introduced by \cite{caron_2007} in 2006.
-
-The GPUDDPM, when applied to data over T discrete time steps, can be viewed as a sequence of DPMs (one for each $t \in \{1, \ldots, T \}$), which are linked together by dependencies between cluster parameters in neighboring time steps.
-More specifically, the parameters at a given time step are distributed as a function of the parameters in the previous time step, and the distribution and number of distinct cluster assignments at a given time step are distributed according to all previous cluster assignments and a deletion procedure, described below.
-
-A transition kernel $P(\theta_{k,t} | \theta_{k,t-1})$ specifies how mixture component parameters in time step $t$ are dependent upon associated mixture component parameters in time step $t-1$.
-One caveat is that each mixture component must be drawn independently from $\mathbb{G}_{0}$ (the base distribution of the DP, which acts as a prior distribution for the cluster parameters) which we achieve by making $\mathbb{G}_{0}$ the invariant distribution of the transition kernel $P(\theta_{k,t} | \theta_{k,t-1})$ (which, one should note, is also a markov chain).
-
-Recall that the distribution over the cluster assignments $c_{i}$ (one for each observation $i$) is a function of the cluster sizes (as per \eqref{crp_rep}).
-Hence, to account for varying numbers of clusters---and in particular, to allow clusters to diminish in size (i.e.
-to diminish in number of assigned observations) and even ``die off"---there is a deletion procedure that allows observations to be considered removed from their assigned clusters at a given time step.
-This procedure is introduced and described in detail in \cite{caron_2007}.
-
-The deletion procedure operates in the following way: at each time step, all previous assignments (that have not yet been deleted) are independently considered for deletion.
-Specifically, each remaining assignment is removed from its cluster with probability $\rho$.
-It can be shown that performing deletion in this way is equivalent to, for each cluster $k$, drawing $r \sim \text{Binomial}(m_{k,t-1}, \rho)$, and reducing the previous-time cluster size $m_{k,t-1}$ by $r$.
-At each time step, this deletion procedure is carried out on all existing assignments, including those at the current time.
-Hence, the size $m_{k,t}$ of a given cluster $k$ at time $t$ is dependent upon the cluster size at the previous time step, $m_{k,t-1}$, and on the assignments $c_{i}$ for all observations $i \in \{ 1, \ldots, N_{t} \}$ at time $t$.
-We refer to the conditional distribution over $m_{k,t} | m_{k,t-1}, c_{i}, \rho$ as DEL, which we define to be
-\begin{equation}
-\begin{split}
-\label{del_step}
-\text{DEL} & (\cdot | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) :=
-\left( m_{k,t-1} - \text{Binomial}( \cdot | {m_{k,t-1}, \hspace{1mm} \rho}) \right) + \sum_{i=1}^{N_{t}} \mathbb{I}(c_{i,t} = k)
-\end{split}
-\end{equation}
-$\forall k \in \{1, \ldots, K_{t} \}$, where $\rho$ is a deletion parameter, $K_{t}$ is the number of clusters at time $t$, and $\mathbb{I}(c_{i,t} = k)$ is an indicator function whose value is 1 if $c_{i,t} = k$ and 0 otherwise.
-This process adheres to what \cite{caron_2007} calls a ``uniform deletion strategy'' over all the observations' assignments (since assignments to each cluster have equal probability of being deleted), though a more complex deletion strategy, dependent upon cluster size, can also be implemented and is described in \cite{caron_2007}.
-
-The Chinese restaurant process, defined in \eqref{crp_rep}, is used to sample assignment variables during inference procedures (described in Section~\ref{sec:inference}).
-Based on this definition, we define the distribution CRP to be
-\begin{equation}
-\label{CRP}
-\begin{split}
-\text{CRP} & (\cdot | m_{1:K_{t},t}, \hspace{1mm} \alpha) :=
-\text{Discrete}(\cdot | P(c_{i,t}=1 | m_{1:K_{t},t}), \ldots, P(c_{i,t}=K_{t+1}  | m_{1:K_{t},t}))
-\end{split}
-\end{equation}
-where
-\begin{equation}
-P(c_{i,t}=k  | m_{1:K_{t},t}) = 
-\begin{cases}
-\frac{m_{k,t}}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{2mm} \text{if} \hspace{2mm} k \in \{ 1, \ldots, K_{t} \} \\
-\frac{\alpha}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{2mm} \text{if} \hspace{2mm} k = K_{t} + 1
-\end{cases}
-\end{equation}
-$\forall i \in \{1, \ldots, N_{t} \}$, where there exists $K_{t}$ clusters at time $t$, and we give a newly created cluster the index $K_{t+1}$.
-Note the similarity between the above definition and the Chinese resturant process of the DPM given by \eqref{crp_rep}.
-The deletion and Chinese resturant procedures together comprise what \cite{caron_2007} refers to as the ``Generalized Polya Urn''.
-Using the DEL and CRP distributions given above, we can define the GPUDDPM generatively as, for each time step $t \in \{1, \ldots, T\}$, and each cluster $k \in \{ 1, \ldots, K_{t} \}$ at time $t$,  % and observation $i \in \{ 1, \ldots, N_{t} \}$ at $t$,
-\begin{align}
-\label{gpuddpm_def}
-\begin{split}
-m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho  &\sim \text{DEL}(m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) \\
-\theta_{k, t} | \theta_{k, t-1}   &\sim
-\begin{cases}
-	P(\theta_{k, t} | \theta_{k, t-1}) \hspace{2mm} \text{if} \hspace{2mm} k \in \{ 1, \ldots, K_{t} \} \\
-	\mathbb{G}_{0}   \hspace{2mm} \text{if} \hspace{2mm} k = K_{t+1}
-\end{cases} \\
-c_{i,t} | m_{1:K_{t},t}, \hspace{1mm} \alpha  &\sim  \text{CRP} (m_{1:K_{t},t}, \hspace{1mm} \alpha) \\
-\bold{x}_{i,t} | c_{i,t}, \theta_{1:K_{t}, t} &\sim F(\theta_{c_{i,t}, t})
-\end{split}
-\end{align}
-The graphical model corresponding with this formulation is shown in Figure \ref{fig:gpuddpm_gm_1}.
-\begin{figure}[h]
-        \center{\includegraphics[width=90mm]{../img/gpuddp_gm_1.pdf}}
-        \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Polya Urn Dependent Dirichlet Process Mixture.
-Note that all observations at time $t$, $\bold{x}_{1:N,t}$, and their associated assignments $c_{1:N,t}$ are denoted respectively as $\bold{x}_{t}$ and $c_{t}$ (this also holds for time step $t-1$).}
-\end{figure}
-% \begin{align}
-% \begin{split}
-% \bold{m}_{t} | \bold{m}_{t-1}, \rho  &\sim \text{DEL}(\bold{m}_{t-1}, \rho) \\
-% c_{i} | \bold{m}_{t}, \alpha  &\sim  \text{CRP}(\bold{m}_{t}, \alpha) \\
-% \theta_{c_{i}, t} | \theta_{c_{i}, t-1}   &\sim
-% \begin{cases}
-% 	P(\theta_{c_{i}, t} | \theta_{c_{i}, t-1}) \\
-% 	\mathbb{G}_{0}
-% \end{cases} \\
-% \bold{x}_{i} | c_{i}, \theta_{c_{i}, t} &\sim F(\theta_{c_{i}, t})
-% \end{split}
-% \end{align}
-% where $\text{CRP}(\cdot, \alpha)$ is given in \eqref{crp_rep}.
-The graphical model corresponding with this formulation is shown in Figure \ref{fig:gpuddpm_gm_1}.
-The above formulation holds if there is exactly one observation at each time step.
-Often, the data extracted from objects in videos consists of multiple observations at each time step, each with an assignment variable ($c_{i}$ for $i \in \{ 1, \ldots, N_{t} \}$, where $N_{t}$ is the number of observations at time step $t$).
-If we allow for multiple observations $i \in \{ 1, \ldots, N_{t} \}$ at each time $t$, this formulation becomes,
-% \begin{align}
-% \begin{split}
-% \bold{m}_{t}^{1} | \bold{m}_{t-1}^{N_{t-1}}, \rho  &\sim \text{DEL}(\bold{m}_{t-1}^{N_{t-1}}, \rho) \\
-% c_{i} | \bold{m}_{t}^{i-1}, \alpha  &\sim  \text{CRP}(\bold{m}_{t}^{i-1}, \alpha) \\
-% \theta_{c_{i}, t} | \theta_{c_{i}, t-1}   &\sim
-% \begin{cases}
-% 	P(\theta_{c_{i}, t} | \theta_{c_{i}, t-1}) \\
-% 	\mathbb{G}_{0}
-% \end{cases} \\
-% \bold{x}_{i} | c_{i}, \theta_{c_{i}, t} &\sim F(\theta_{c_{i}, t})
-% \end{split}
-% \end{align}
-% where $\bold{m}_{t}^{i}$ represents the $i^{\text{th}}$ observation at time $t$.
-
-% section making explicit connection between model and video tracking
-% -------------------------------------------------------------------
-
-\subsubsection{Object Characteristics Modeled}
-
-We would like make explicit the connection between certain elements of the GPUDDPM model and the characteristics of objects in vidoes that these elements represent.
-
-First, the appearances of objects are moded by what we term the `likelihood and object appearance model', which we denoted by $F$.
-In the above modeling formulation, $F$ is the probability distribution for an observation, given that it is assigned to specified cluster (i.e.
-given that it is associated with a given object in a video).
-The likelihood and object appearance model is therefore
-\begin{equation}
-F(\theta_{c_{i}}) = P(x_{i}|\theta_{c_{i}})
-\end{equation}
-where $x_{i}$ is an observation, $c_{i}$ is its associated assignment, $\theta_{c_{i}}$ specifies the parameters of the parametric form of the $c_{i}^{th}$ cluster.
-The specific form of this model is dependent upon the observations $x_{i}$; recall that the observations, which as a baseline include spatial and temporal features, could include an arbitrary amount of additional features which need to be incorporated into this appearance model.
-A specific formulation of the appearance model can be seen in the proceeding section.
-
-Object behavior is modeled by the dependencies between corresponding clusters at adjacent time steps.
-In particular, this relationship is captured by the transition kernel, $P(\theta_{k}^{t} | \theta_{k}^{t-1})$ (which captures the dependence of cluster $k$ at time $t$ on the same cluster at time $t-1$), described previously.
-The motion behavior of an object, i.e.
-the specific dependence of the parameters of cluster $k$ at a time $t$ on its parameters at time $t-1$, is therefore
-\begin{equation}
-Tr(\theta_{k, t} | \theta_{k, t-1}) = P(\theta_{k, t} | \theta_{k, t-1})
-\end{equation}
-
-Additionally, the base distribution $\mathbb{G}_{0}$ of the DP acts as a prior on the cluster parameters.
-This allows one to place a prior probability over appearances of objects.
-It is important that this prior is not chosen in such a specific manner that it limits this method from performing unsupervised detection of arbitrary objects.
-
-% section providing specific model
-% --------------------------------
-
-\section{Model Specification}
-\label{sec:modelspec}
-
-In the following sections, a particular extraction method for generating data from videos and a specific formulation of the general model described in Section~\ref{sec:modeloverview} are given.
-Inference schemes on this model, which when carried out allow for unsupervised detection and tracking of multiple arbitrary objects, are also described.
-
-% gives specific data extraction method (frame differencing)
-% ----------------------------------------------------------
-
-\subsection{Data Extraction via Frame Differencing}
-\label{sec:modelspec_extraction}
-
-We desire an extraction procedure that is as unsophisticated as possible, both to gauge the robustness of this method on potentially noisy extraction data and to ensure that the procedure is applicable to a wide range of videos.
-For these reasons we choose the method of frame-differencing for all extraction performed for experiments in this paper (Section~\ref{sec:experiments}), which involves recording the positions of pixels that have exhibited differences in intensity or value in succesive frames beyond a given threshold.
-In particular, if we let $I_{t}$ be the image difference obtained by subtracting the value of the $(i,j)^{th}$ pixel in video frame $t$ from the value of the $(i,j)^{th}$ pixel in video frame $t+1$, we can define extraction on frame $t$ to be the process that returns the dataset
-\begin{equation}
-	\Omega_{t} = \{ (i,j,t) \hspace{2mm} | \hspace{2mm} I_{t}(i,j) > \phi \}
-\end{equation}
-where $i$ and $j$ respectively denote the two spatial positions of a pixel, and $\phi$ is a given pixel value threshold. Frame differencing on each frame $t =\{1, \ldots, T \}$ yields the dataset
-\begin{equation}
-	\Omega = \bigcup_{t=1}^{T} \Omega_{t} = \{ (i,j,t) \hspace{2mm} | \hspace{2mm} I_{t}(i,j) > \phi,  \hspace{2mm}  \forall t \in \{1, \ldots, T \} \}
-\end{equation}
-which is equivalent to data with the baseline features described in Section~\ref{sec:data}.
-
-Frame differencing is unsophisticated, computationally inexpensive, able to be applied to a wide range of static, single-camera videos (note that videos used in the experiments described in the following sections were chosen to be static; moving-camera videos should be used in conjunction with applicable extraction methods that provide accurate foreground segmentation for non-stationary video).
-We have found that, when correctly implemented, frame differencing is sufficiently general to extract the desired worm-like data clouds from a variety of videos containing multiple moving objects.
-A few examples of this extraction procedure on pairs of consecutive video frames can be found in Figure \ref{fig:img_and_framediff}.
-
-% \begin{figure}
-% \centering{
-% \includegraphics[width=45mm]{../img/antpic2.png}}
-% %\hspace{2mm}
-% {\includegraphics[width=60mm]{../img/frame_diff_ants_1.png}}
-% \caption{(a.) Frame of a video containing five ants.
-(b.) Frame differencing extraction results from the same frame \willie{this is just a demo of the type of image i will have here...
-these aren't actually from the same frame, and also formatting is screwed up)}}
-% \label{test}
-% \end{figure}
-
-\begin{figure}
+This method begins by performing a simple data extraction procedure (that is consistent for all videos), which can be carried out broadly but yields noisy data; we model this noisy extracted data with a sophisticated type of Bayesian mixture model.
+Inference can then be performed on this model to gain detection and tracking results.
+
+This paper describes the data extraction procedures (Section~\ref{sec:dataextraction}), the general form of the model (Section~\ref{sec:modeldefinition}), the specific distributions used in our implementation of the model (Section~\ref{sec:modelspecification}), inference procedures used to get tracking results (Section~\ref{sec:inference}), and experimental results (Section~\ref{sec:experiments}).
+
+% Overview
+% --------
+
+% \section{Overview of Method}
+% \label{sec:overviewofmethod}
+
+% Frame Diff and Longimg figure
+% -----------------------------
+
+\begin{figure*}
   \centering
@@
   %\subfloat[]{\label{fig:longimg}\includegraphics[width=1.05\textwidth]{../img/longimg_6.pdf}} \\
-  \subfloat[]{\label{fig:longimg_all}\includegraphics[width=1\textwidth]{../img/longimg_7_detail.pdf}} 
+  \subfloat[]{\label{fig:longimg_all}\includegraphics[width=1\textwidth]{../img/longimg_7_detail_4.pdf}} 
+  %../img/longimg_7_detail.pdf}} 
   \caption{Pairs of consecutive frames and the results produced by taking the pixel-wise difference between these frames (a - f).
@@
   \label{fig:img_and_framediff}
-\end{figure}
-
-We furthermore wish to capture the color (or grayscale) value of each pixel and incorporate this color information into our model.
-For each pixel $x = (i, j, t) \in \Omega$ (defined above), we specify a square, $L$ pixels in length, centered on $(i, j)$, that selects a set of pixels surrounding $\bold{x}$ in frame $t$.
+\end{figure*}
+
+% Data Extraction
+% ---------------
+
+\section{Data Extraction}
+\label{sec:dataextraction}
+
+A basic set of observations that can be extracted from video are three-dimensional points $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $x_{1}$ and $x_{2}$ are the two spatial locations of a pixel.
+Extraction procedures, such as those discussed in Section~\ref{sec:introduction}, are used to isolate pixels lying in the regions of a video that represent objects.
+For implementation in experiments (Section~\ref{sec:experiments}), we desire an extraction procedure that is as unsophisticated as possible, both to gauge the robustness of this method on potentially noisy extraction data and to ensure that the procedure is applicable to a wide range of videos.
+Accordingly, we carry out a method called ``frame differencing'' for all experiments, which locates regions within video frames that contain motion by recording the positions of pixels exhibiting large enough differences in intensity or value between succesive frames.
+Frame differencing is unsophisticated, computationally inexpensive, able to be applied to a wide range of static, single-camera videos (note that videos used in all experiments were stationary; moving-camera videos require extraction methods useful for non-stationary video, such as \willie{cite cite cite}).
+Examples of pixel location data extracted via frame differencing are shown in Fig.~\ref{fig:img_and_framediff}(a-j).
+
+Incorporating color information into our model allows us to infer object color, better maintain the identity of adjacent objects, and track objects through occlusion with greater accuracy.
+We therefore capture the color of each extracted pixel.
+Specifically, for each pixel $\bold{x} = ( x_{1}, x_{2}, t )$ recorded during frame differencing, we specify a square, $L$ pixels in length, centered on $(x_{1}, x_{2})$, that contains a set of pixels surrounding $\bold{x}$ in frame $t$.
+We choose a scalar color value that is able to be computed for each pixel; this value could, for example, be some function of the red-green-blue (r-g-b) or hue-saturation-value (h-s-v) characteristic of a pixel.
+The color value of each of the selected pixels surrounding $\bold{x}$ is recorded.
+Afterwards, the set of possible color values (ie the range of color values to which a pixel may be assigned) is partitioned into $V$ bins, and the number of pixels with a color value lying in each of the bins yields a $V$ dimensional vector of ``color counts''.
+This vector represents a discrete distribution over colors in the vicinity of a pixel.
+We will refer to this extraction technique as ``color counting''.
+
+The frame differencing and color counting extraction yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \{1, \ldots, T \}$, where each element $\bold{x} \in \bold{X}$ can be written
+\begin{equation}
+\bold{x} = ( \bold{x}^{s}, \bold{x}^{c}, t ) = ( x^{s_{1}}, x^{s_{2}}, x^{c_{1}}, \ldots, x^{c_{V}}, t )
+\end{equation}
+where $\bold{x}^{s} \in \mathbb{R}^{2}$ is the two dimensional vector of spatial coordinates,  $\bold{x}^{c} \in \mathbb{Z}_{+}^{V}$ is the $V$ discrete dimensional vector of color counts, and $t \in \{1, \ldots, T \}$ is the discrete time index.
+Additional features may be extracted from each pixel to attempt more sophisticated object representations, including pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, and texture representations.
+We leave these attempts for future studies.
+
+% We defined the task of extraction to be the unsupervised determination of which spatiotemporal regions of a video constitute objects, and outlined a number of extraction methods in Section~\ref{sec:priorwork}.
+The most basic dataset that can be gained from extraction is comprised of the three-dimensional points $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $x_{1}$ and $x_{2}$ are the two spatial dimensions of a pixel contained in a region found during extraction and $t$ is the frame in which the pixel resides.
+Once this dataset, which we will refer to as a ``baseline extraction dataset'', has been found, additional features may also be extracted from the video.
+Examples of additional features include color information, pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, and texture representations; the main idea is to choose features which are able to be easily extracted or computed, capture variability in the appearance of objects, and are applicable to a wide variety of object types.
+
+% The baseline extraction dataset has the appearance of worm-like clusters that follow the paths of individual objects as they move (as illustrated in Figure~\ref{fig:longimg_all}).
+Modeling this data with a dependent Dirichlet process mixture model (Section~\ref{sec:gpuddpmixture}) allows us to perform inference (Section~\ref{sec:inference}), which carries out clustering of the data.
+By choosing this type of model, we aim to provide a clustering result that isolates each worm-like cluster and infers a sequence of distributions for each, which we use to gain localization and tracking results for each object (Section~\ref{sec:inferencetotrackingresults}).
+
+% For implementation in experiments (Section~\ref{sec:experiments}), we desire an extraction procedure that is as unsophisticated as possible, both to gauge the robustness of this method on potentially noisy extraction data and to ensure that the procedure is applicable to a wide range of videos.
+Consequently, in all experiments we carry out a method called ``frame-differencing'', which locates motion by recording the positions of pixels that have exhibited differences in intensity or value in succesive frames beyond a given threshold.
+Frame differencing is unsophisticated, computationally inexpensive, able to be applied to a wide range of static, single-camera videos (note that videos used in all experiments were stationary; moving-camera videos require extraction methods that provide accurate foreground segmentation for non-stationary video).
+We have found that, when correctly implemented, frame differencing is sufficiently general to extract the desired worm-like clusters from a variety of videos containing multiple moving objects.
+A few examples of this extraction procedure on pairs of consecutive video frames can be found in Figure~\ref{fig:img_and_framediff}(a-i).
+
+% We furthermore wish to capture the color (or grayscale) value of each pixel and incorporate this color information into our model.
+For each pixel $x = (i, j, t)$ recorded during frame differencing, we specify a square, $L$ pixels in length, centered on $(i, j)$, that selects a set of pixels surrounding $\bold{x}$ in frame $t$.
 We also specify a scalar color value that is able to be computed for each pixel; this value could, for example, be some function of the red-green-blue (r-g-b) or hue-saturation-value (h-s-v) characteristic of a pixel.
@@
 
-% perhaps: Show example color-count vector (histogram) for different objects in an image. give more-formal definition like for frame differencing above.
-
-% introduce specific model for experiments
-% ----------------------------------------
-
-\subsection{Model Implementation in Experiments}
-
-The following sections provide details on the data and specific model formulation (determined from the general model formulation of Section~\ref{sec:modeloverview}) involved in the experiments described in Section~\ref{sec:experiments}.
-
-% data in experiments
-% -------------------
-
-\subsubsection{Data}
-
-The frame differencing and color counting extraction outlined in Section~\ref{sec:modelspec_extraction} yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \{1, \ldots, T \}$, where each element $\bold{x} \in \bold{X}$ can be written
-\begin{equation}
-\bold{x} = ( \bold{x}^{s}, \bold{x}^{c}, t ) = ( x^{s_{1}}, x^{s_{2}}, x^{c_{1}}, \ldots, x^{c_{V}}, t )
-\end{equation}
-where $\bold{x}^{s} \in \mathbb{R}^{2}$ is the two dimensional vector of spatial coordinates,  $\bold{x}^{c} \in \mathbb{Z}_{+}^{V}$ is the $V$ discrete dimensional vector of color counts, and $t \in \{1, \ldots, T \}$ is the discrete time index.
-
-In the following experiments, the hue component of the h-s-v value is recorded from all pixels that surround each extracted pixel $\bold{x}$ in the manner described in Section~\ref{sec:modelspec_extraction} (where we choose $L=5$).
+% The frame differencing and color counting extraction yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \{1, \ldots, T \}$, where each element $\bold{x} \in \bold{X}$ can be written
+% \begin{equation}
+% \bold{x} = ( \bold{x}^{s}, \bold{x}^{c}, t ) = ( x^{s_{1}}, x^{s_{2}}, x^{c_{1}}, \ldots, x^{c_{V}}, t )
+% \end{equation}
+% where $\bold{x}^{s} \in \mathbb{R}^{2}$ is the two dimensional vector of spatial coordinates,  $\bold{x}^{c} \in \mathbb{Z}_{+}^{V}$ is the $V$ discrete dimensional vector of color counts, and $t \in \{1, \ldots, T \}$ is the discrete time index.
+
+% In the experiments described in Section~\ref{sec:experiments}, the hue component of the h-s-v value is recorded from all pixels that surround each extracted pixel $\bold{x}$ in the manner described above, choosing $L=5$.
 Additionally, the set of possible hue values is partitioned into 10 bins, and the number of pixels with a hue value lying in each of the 10 bins is recorded to yield the vector of ``color counts'' $\bold{x}^{c} = x^{c_{1}}, \ldots, x^{c_{10}}$.
@@
 
-% likelihood / object appearance model
-% ------------------------------------
-
-\subsubsection{Likelihood and Object Appearance Model}
-
-At a given time $t$, we model each $\bold{x} \in \bold{X}$ as a draw from the product of a multivariate normal and multinomial distribution
+% Model Overview
+% --------------
+
+\section{Model Definition}
+\label{sec:modeldefinition}
+
+A type of Dependent Dirichlet Process Mixture (DDPM) model known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM) is used to provide a model-based, time dependent clustering of the extracted data.
+The model definition is provided in Section~\ref{sec:gpuddpm}, a secondary form of the model used in one of the two inference algorithms is given in Section~\ref{sec:MCMC}, and background information introducing mixture models and the Dirichlet process are given in Appendix~\ref{sec:modelbackground}.
+
+% GPUDDPM definition
+% ------------------
+
+\subsection{Generalized Polya Urn Dependent Dirichlet Process Mixture Model}
+\label{sec:gpuddpm}
+
+The extracted data contain a collection of spatiotemporal clusters, each representing an object in the video.
+The number of clusters present in each time step may not be constant (which is to say, clusters may be created, or be ``born'', and may disappear, or ``die'', at intermediate time steps), since objects can enter and exit a scene.
+To cluster data with these properties, we choose a DDPM known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM), introduced by \cite{caron_2007} in 2006.
+This model may be intuitively viewed as a sequence of DPMs, where the parameters and number of clusters are allowed to vary in a specified way between adjacent time steps.
+Like the DPM, the GPUDDPM allows for the number of clusters within a dataset---which, in this application, corresponds to the number of objects---to be found in an automated manner.
+
+As in typical mixture models, each observation $\bold{x}_{i,t}$ is associated with an assignment variable $c_{i,t}$ that represents its assignment to a cluster $\theta_{k,t}$.
+Usually, the size of each cluster $m_{k,t}$ is a function of the number of observations that have been assigned to it, but in this model, cluster sizes may decrease over time as observations become ``unassigned''; this allows clusters to decrease in strength and potentially die.
+We define a distribution over the size of cluster $k$ at time $t$ ($m_{k,t}$) conditioned on the previous cluster size ($m_{k,t-1}$), assignments at time $t$ ($c_{1:N_{t}, t}$), and deletion parameter $\rho$ to be
+\begin{equation}
+\begin{split}
+\label{D_distro}
+\text{D} &(m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) := \hspace{2mm} m_{k,t-1} \\
+&- \text{Binomial}( m_{k,t} | {m_{k,t-1}, \hspace{1mm} \rho}) + \sum_{i=1}^{N_{t}} \mathbb{I}(c_{i,t} = k)
+\end{split}
+\end{equation}
+$\forall k \in \{1, \ldots, K_{t} \}$, where $K_{t}$ is the number of clusters at time $t$ and $\mathbb{I}(c_{i,t} = k)$ is an indicator function whose value is 1 if $c_{i,t} = k$ and 0 otherwise.
+We also define a distribution over the assignment of an observation $i$ at time $t$ to a cluster $k$ ($c_{i,t} = k$) conditioned on the sizes of all clusters at time $t$ ($m_{1:K_{t},t}$) and a concentration parameter $\alpha$, to be
+\begin{equation}
+\begin{split}
+\label{C_distro}
+\text{C} ( c_{i,t} = k | &m_{1:K_{t},t}, \hspace{1mm} \alpha)\\ :=
+& \begin{cases}
+\frac{m_{k,t}}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{1mm} \text{if} \hspace{1mm} k \in \{ 1, \ldots, K_{t} \} \\
+\frac{\alpha}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{1mm} \text{if} \hspace{1mm} k = K_{t} + 1
+\end{cases}
+\end{split}
+\end{equation}
+% \begin{equation}
+% \label{C}
+% \begin{split}
+% \text{C} &(\cdot | m_{1:K_{t},t}, \hspace{1mm} \alpha) := \text{Discrete}(\cdot | P(c_{i,t}=1 | m_{1:K_{t},t}), \\
+% &  \ldots, P(c_{i,t}=K_{t+1}  | m_{1:K_{t},t}))
+% \end{split}
+% \end{equation}
+% where
+% \begin{equation}
+% \begin{split}
+% P (c_{i,t}=k  | m_{1:K_{t},t})
+% = \begin{cases}
+% \frac{m_{k,t}}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{1mm} \text{if} \hspace{1mm} k \in \{ 1, \ldots, K_{t} \} \\
+% \frac{\alpha}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{1mm} \text{if} \hspace{1mm} k = K_{t} + 1
+% \end{cases}
+% \end{split}
+% \end{equation}
+$\forall i \in \{1, \ldots, N_{t} \}$, where there exists $K_{t}$ clusters at time $t$ and we give a newly created cluster the index $K_{t+1}$.
+Distributions \eqref{D_distro} and \eqref{C_distro} together comprise what \cite{caron_2007} refers to as the ``Generalized Polya Urn''.
+We can now define the GPUDDPM generatively as, for each time step $t \in \{1, \ldots, T\}$, and each cluster $k \in \{ 1, \ldots, K_{t} \}$ at time $t$,
+\begin{align}
+\label{gpuddpm_def}
+\begin{split}
+m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho  &\sim \text{D}(m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) \\
+\theta_{k, t} | \theta_{k, t-1}   &\sim
+\begin{cases}
+  P(\theta_{k, t} | \theta_{k, t-1}) \hspace{1mm} \text{if} \hspace{1mm} k \in \{ 1, \ldots, K_{t} \} \\
+  \mathbb{G}_{0}   \hspace{2mm} \text{if} \hspace{2mm} k = K_{t+1}
+\end{cases} \\
+c_{i,t} | m_{1:K_{t},t}, \hspace{1mm} \alpha  &\sim  \text{C} (m_{1:K_{t},t}, \hspace{1mm} \alpha) \\
+\bold{x}_{i,t} | c_{i,t}, \theta_{1:K_{t}, t} &\sim \text{F}(\theta_{c_{i,t}, t})
+\end{split}
+\end{align}
+where we specify distributions for F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ in Section~\ref{sec:modelspecification}.
+The graphical model associated with this formulation is shown in Fig.~\ref{fig:gpuddpm_gm_1}.
+\begin{figure}[h]
+        \center{\includegraphics[width=80mm]{../img/gpuddp_gm_1.pdf}}
+        \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Polya Urn Dependent Dirichlet Process Mixture.
+Note that all observations at time $t$, $\bold{x}_{1:N,t}$, and their associated assignments $c_{1:N,t}$ are denoted respectively as $\bold{x}_{t}$ and $c_{t}$ (this also holds for time step $t-1$).}
+\end{figure}
+
+% Model Specification
+% -------------------
+
+\section{Model Specification}
+\label{sec:modelspecification}
+
+Sections~\ref{sec:objectappearance}-\ref{sec:motionmodel} describe the distributions chosen to fully specify the general GPUDDPM model form given in Section~\ref{sec:gpuddpm}.
+The implementation given below was used for all the experiments of Section~\ref{sec:experiments}.
+The distributions F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ represent object appearance, appearance prior, and object movement; different specifications allow one to choose a model biased toward a specific application or kept general for wide applicability.
+
+\subsection{Object Appearance (Mixture Component)}
+\label{sec:objectappearance}
+
+At a given time $t$, we model each observation $\bold{x} \in \bold{X}$ as a draw from the product of a multivariate normal and multinomial distribution
 \begin{equation}
 \label{likelihood}
-P(\bold{x}|\theta) = \mathcal{N}(\bold{x}^{s} | \boldsymbol{\mu}, \Sigma)  \mathcal{M}n(\bold{x}^{c} | \bold{p})
+\text{F}(\bold{x}|\theta) = \mathcal{N}(\bold{x}^{s} | \boldsymbol{\mu}, \Sigma)  \mathcal{M}n(\bold{x}^{c} | \bold{p})
 \end{equation}
 where $\theta = \{ \boldsymbol{\mu}, \Sigma, \bold{p} \}$ denote the parameters of a cluster at time $t$, with mean $\boldsymbol{\mu} \in \mathbb{R}^{2}$, covariance matrix $\Sigma \in \mathbb{R}^{2} \times \mathbb{R}^{2}$, and discrete probability vector $\bold{p} = (p_{1}, \ldots, p_{V})$ such that $\sum_{i=1}^{V}p_{i} = 1$.
-Also, $\mathcal{N}$ denotes the multivariate normal distribution and $\mathcal{M}n$ denotes the multinomial distribution,
+Additionally, $\mathcal{N}$ denotes the multivariate normal distribution, $\mathcal{M}n$ denotes the multinomial distribution, and F is sometimes referred to as the likelihood.
 
 The distribution families assumed to generate each $\bold{x}$ must be justified.
-Both the multivariate normal and multinomial distributions were chosen because they are sufficiently simple (and well studied) to allow for tractable inference and sufficiently flexible to provide a reasonable approximation to the data gained during extraction.
-In particular, the multivariate normal distribution over the spatial features $\bold{x}^{s}$ can be thought to represent the shape of each object as an oval; likewise, data generated by each moving object during extraction are often ovular---as noisy extraction procedures cause some smoothing of edges and corners, producing blob-like shapes even when objects are not particularly round---and centered on a given object.
-Furthermore, this model is justified as the maximum likelihood parameter estimate of a normal distribution corresponds to the least squares fit of data relative to the normal distribution mean; since extraction data produced by an object clusters around the centroid of the object, the estimated normal distribution mean should provide a reasonable reflection of the object's centroid.
+%Both the multivariate normal and multinomial distributions were chosen because they are sufficiently simple (and well studied) to allow for tractable inference and sufficiently flexible to provide a reasonable approximation to the data gained during extraction.
+The multivariate normal distribution over the spatial features $\bold{x}^{s}$ represents the shape of each object as an oval; likewise, data generated by each moving object during extraction are often ovular.
+%---as noisy extraction procedures cause some smoothing of edges and corners, producing blob-like shapes even when objects are not particularly round---and centered on a given object.
+% Furthermore, this model is justified as the maximum likelihood parameter estimate of a normal distribution corresponds to the least squares fit of data relative to the normal distribution mean; 
+Since extraction data produced by an object clusters around the centroid of the object, the estimated normal distribution mean should provide a reasonable reflection of the object's centroid.
 Modeling the color features $\bold{x}^{c}$ as draws from a multinomial distribution (equivalently, as draws from a product of discrete distributions), is justified since we have observed that distinct object tend to generate pixels whose hue values are noisy but yield consistent counts in discrete hue bins.
 
-% base distribution / appearance prior
-% ------------------------------------
-
-\subsubsection{Base Distribution $\mathbb{G}_{0}$ and Appearance Prior}
-
-$\mathbb{G}_{0}$ denotes the base distribution of the time-dependent Dirichlet process mixture; it also serves as a prior distribution for the parameters $\theta = \{ \boldsymbol{\mu}, \Sigma, \bold{p} \}$ present in the likelihood.
+\subsection{Appearance Prior (Base Distribution)}
+\label{sec:appearanceprior}
+
+$\mathbb{G}_{0}$ denotes the base distribution of the time-dependent Dirichlet process mixture; it also serves as a prior distribution for the parameters $\theta = \{ \boldsymbol{\mu}, \Sigma, \bold{p} \}$ present in the mixture components (i.e.
+the parameters we've used to model object appearance).
 We make use of conjugate priors in the base distribution to allow for more efficient computation.
-Specifically, in the experiments carried out in this paper, a normal-inverse-Wishart prior is placed on the multivariate normal parameters $\{ \boldsymbol{\mu}, \Sigma \}$ , and a Dirichlet prior is placed on the multinomial parameter $ \{  \bold{p}  \} $ (where the normal-inverse-Wishart is a conjugate prior for the multivariate normal component of the likelihood and the Dirichlet is a conjugate prior for the multinomial component).
+Specifically, in the experiments carried out in this paper, a normal-inverse-Wishart prior is placed on the multivariate normal parameters $\{ \boldsymbol{\mu}, \Sigma \}$, and a Dirichlet prior is placed on the multinomial parameter $\bold{p}$ (where the normal-inverse-Wishart is a conjugate prior for the multivariate normal component of the likelihood and the Dirichlet is a conjugate prior for the multinomial component).
 The prior can therefore be written
@@
 
-% perhaps: reference text with good background on NiW and Dir distributions
-
-% transition kernel
-% -----------------
-
-\subsubsection{Transition Kernel and Motion Model}
+\subsection{Motion Model (Transition Kernel)}
+\label{sec:motionmodel}
 
 We wish to formulate a transition kernel $P(\theta_{t} | \theta_{t-1})$ that provides a reasonable representation of how we expect tracked objects to move over time.
-In the interest of formulating a model with the intention of using it to track arbitrary objects, we do not wish to make many assumptions about the long-term behavior of objects.
-For example, we choose not to adopt a transition kernel that incorporates object dynamics, though a transition kernel that models this or other sophisticated behavioral tendencies could potentially be implemented in future works if one knows that they are consistent characteristics of an object's movement.
-
-As described in Section~\ref{sec:gpuddpm}, $\mathbb{G}_{0}$ must be the invariant distribution of $P(\theta_{t} | \theta_{t-1})$ in order for the the cluster parameters to remain marginally distributed according to the base distribution.
+Since we want our model to be used for tracking arbitrary objects, we do not wish to make any sophisticated assumptions about the behavior of objects.
+For example, we choose not to incorporate complex objects dynamics, though they are often used with success in certain object-specific tracking tasks, such as tracking walking people.
+We assume only that the position of an object at a given time is a function of its position at the previous time, and that the position varies in all directions equally between time steps.
+A transition kernel that incorporates more sophisticated behavioral tendencies is left for future studies (especially for those intending to use this model for a more specific object type or application).
+
+The base distribution $\mathbb{G}_{0}$ must be the invariant distribution of the transition kernel $P(\theta_{t} | \theta_{t-1})$ in order for the the cluster parameters to remain marginally distributed according to the base distribution.
 In other words, the transition kernel must satisfy
@@
 \begin{center}
-\begin{tabular}[c]{l p{10cm}}
-$\alpha$  &  The granularity parameter for the Dirichlet Process. A higher value will increase the tendency for new objects to be detected.\\
+\begin{tabular}[c]{l p{7cm}}
+$\alpha$  &  The concentration parameter for the Dirichlet Process. A higher value will increase the tendency for new objects to be detected.\\
 $\rho$  &  The deletion parameter. A higher value will give objects an increased tendency to die off.
@@
 The parameters chosen for use in experiments (Section~\ref{sec:experiments}) were accepted based on the fact that they yielded reasonable results.
-Additionally, a sensitivity analysis was carried out for the granularity and deletion parameters ($\alpha$ and $\rho$), where a range of values for both were chosen, and inference performed for each; results of this analysis are discussion in Section~\ref{sec:sens_analysis}.
-
-% inference section
-% -----------------
+Additionally, a sensitivity analysis was carried out for the concentration and deletion parameters ($\alpha$ and $\rho$), where a range of values for both were chosen, and inference performed for each; results of this analysis are discussion in Section~\ref{sec:sensitivityanalysis}.
+
+% Inference
+% ---------
 
@@
 \label{sec:inference}
-Algorithms that infer latent model parameters via Bayesian inference given the data gained during extraction provide detection and tracking results.
-By formulating the detection and tracking task in terms of a generative model, a variety of methods previously developed for statistical inference can be readily applied; we separate presentations of model and inference so that multiple inference methods, each with its own benefits in terms of precision and speed, can be formulated as independent algorithms for detection and tracking.
+
+Bayesian inference algorithms, which infer distributions over latent model parameters, are used to get detection and tracking results.
+A variety of methods previously developed for statistical inference can be applied to the generative model defined in Sections~\ref{sec:} and \ref{sec}.
+
+ we separate presentations of model and inference so that multiple inference methods, each with its own benefits in terms of precision and speed, can be formulated as independent algorithms for detection and tracking.
 This section provides background on the two inference algorithms implemented in this study, which are both used to perform Bayesian inference on the model described in Section~\ref{sec:modelspec}.
@@
 
-% MCMC inference section
-% ----------------------
-
-\subsection{MCMC Batch Inference}
+\subsection{MCMC: Batch Inference}
 \label{sec:MCMC}
 
-This section details a Markov Chain Monte Carlo (MCMC) sampler used to perform inference, which provides a method to carry out unsupervised detection and tracking.
+This section details a Markov Chain Monte Carlo (MCMC) sampler used to perform inference, which serves as an algorithm for unsupervised detection and tracking.
 It operates by iteratively sampling from conditional distributions of elements in the joint probability distribution---i.e.
@@
 Additionally, this sampler makes use of the Metropolis-Hastings (M-H) algorithm to sample from a few distributions.
-
-The generative model associated with the GPUDDPM is illustrated in Section~\ref{sec:gpuddpm}.
-To allow for modeling time-dependent data, the GPUDDPM allows clusters to `lose assigned observations' in a deletion step, which modifies the values of the cluster sizes over time.
-Instead of incorporating the cluster size variable $m$ directly (as it was in the definition of the GPUDDPM given in \eqref{gpuddpm_def}), we can formulate an equivalent model which makes use of deletion variables $d$, which represent the time at which observations are lost from a cluster.
-Specifically, we introduce variables $d_{1}, \ldots, d_{N}$ which denote the times at which the assignments of given observations are considered to be removed.
+A secondary formulation of the GPUDDPM, which we refer to as the ``Deletion Variable Formulation'', is used (defined in Section~\ref{sec:deletionvariableformulation}).
+This formulation is equivalent to the formulation given in Section~\ref{sec:gpuddpm}, but allows for easier Gibbs sampling.
+
+\subsubsection{Deletion Variable Formulation of the GPUDDPM}
+\label{sec:deletionvariableformulation}
+
+Instead of incorporating the cluster size variable $m_{k,t}$ directly in the GPUDDPM model (as it was in the definition given by \eqref{gpuddpm_def}), we can formulate an equivalent model which makes use of new set of variables called the deletion variables, which represent the times at which assignments are deleted.
+More precisely, each deletion variable $d_{i,t}$ denotes the time at which the associated observation $\bold{x}_{i,t}$ is considered removed from its assigned cluster $c_{i,t}$.
 At each time step, cluster sizes can be recontructed from all previous assignment and deletion variables.
-With these we can express the size of cluster $k$ at time $t$, $m_{k,t}$, as
+Specifically, we can compute the size of cluster $k$ at time $t$, $m_{k,t}$, by
 \begin{equation}
@@
 where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise.
-Additionally, for an observation $\bold{x}_{i}$ at a given time-step $t$, the deletion time $d_{i}$ can be defined to be $d_{i} = t + l_{i}$, where $l_{i}$ is considered the lifespan of an assignment, is distributed geometrically, and can be expressed as
+Additionally, for an observation $\bold{x}_{i,t}$, the deletion time $d_{i,t}$ can be defined to be $d_{i,t} = t + l_{i,t}$, where $l_{i,t}$ can be thought of as the lifespan of an assignment, is distributed geometrically, and can be expressed as
 \begin{equation}
 \label{del_rho_form}
-l_{i} | \rho  \sim  \rho(1 - \rho)^{l_{i}}
-\end{equation}
-The MCMC sampler described in this section operates in this alternate representation, which we can write, for each time step $t \in \{1, \ldots, T\}$ and clusters $k \in \{ 1, \ldots, K_{t} \} $ at time $t$, as
-% \begin{align}
-% \begin{split}
-% 	 \bold{x}_{i} | c_{i}, \theta_{c_{i},t} &\sim \mathcal{N}(\boldsymbol{\mu}_{c_{i},t}, \Sigma_{c_{i},t})  \mathcal{M}n(\bold{p}_{c_{i},t}) \\
-% 	c_{i} | c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha  &\sim  \text{CRP}(c_{1}, \ldots, c_{i-1}, d_{1}, \ldots, d_{i}, \alpha) \\
-% 	d_{i} | \rho  &\sim \text{Geo}(\rho) + i + 1\\
-% 	\theta_{k, t} | \theta_{k, t-1}, \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0}  &\sim 
-% \begin{cases}
-% 	\text{TK} \hspace{12pt} \text{if } k \in \{ 1, \ldots, K_{t-1}  \} \\
-% 	\mathbb{G}_{0}(\boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0})  \hspace{15pt} \text{if } k = K_{t-1} + 1
-% \end{cases}
-% \end{split}
-% \end{align}
-% where TK stands for the transition kernel, given in \eqref{transker1} and \eqref{transker2}, and the base distribution $\mathbb{G}_{0}$ is given in \eqref{basedistro}.
+l_{i,t} | \rho  \sim  \rho(1 - \rho)^{l_{i,t}}
+\end{equation}
+where the deletion parameter $\rho$ is the same as that given in \eqref{D_distro}.
+We can write the new formulation of the GPUDDPM as, for each time step $t \in \{1, \ldots, T\}$ and cluster $k \in \{ 1, \ldots, K_{t} \} $ at time $t$,
 \begin{align}
@@
 \end{cases} \\
-% c_{i,t} | \bold{c}_{1}, \ldots, \bold{c}_{t-1}, \bold{d}_{1}, \ldots, \bold{d}_{t-1}, \alpha  &\sim  \text{CRP}(\bold{c}_{1}, \ldots, \bold{c}_{t-1}, \bold{d}_{1}, \ldots, \bold{d}_{t-1}, \alpha) \\
-c_{i,t} | \bold{c}_{1:t-1}, \bold{d}_{1:t-1}, \alpha  &\sim  \text{CRP}(\bold{c}_{1:t-1}, \bold{d}_{1:t-1}, \alpha) \\
-\bold{x}_{i,t} | c_{i,t}, \theta_{c_{i,t},t}  &\sim  F(\theta_{c_{i,t}, t})
+c_{i,t} | \bold{c}_{1:t-1}, \bold{d}_{1:t-1}, \alpha  &\sim  \text{C}(\bold{c}_{1:t-1}, \bold{d}_{1:t-1}, \alpha) \\
+\bold{x}_{i,t} | c_{i,t}, \theta_{c_{i,t},t}  &\sim  \text{F}(\theta_{c_{i,t}, t})
 \end{split}
 \end{align}
-where $\bold{c}_{t} = c_{1,t}, \ldots, c_{N_{t}, t}$ (for $N_{t}$ observations at time $t$), $\bold{d}_{t} = d_{1,t}, \ldots, d_{N_{t}, t}$, $P(\theta_{k,t} | \theta_{k,t-1})$ is defined in \eqref{transker1} and \eqref{transker2}, $\mathbb{G}_{0}$ is defined in \eqref{basedistro}, CRP is defined in \eqref{CRP}, and $F(\theta_{c_{i,t}, t})$ is defined in \eqref{likelihood}.
+where $\bold{c}_{t} = c_{1,t}, \ldots, c_{N_{t}, t}$ (for $N_{t}$ observations at time $t$), $\bold{d}_{t} = d_{1,t}, \ldots, d_{N_{t}, t}$, and we specify distributions for F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ in Section~\ref{sec:modelspecification}.
 This formulation of the GPUDDPM is also used by \cite{gasthaus_thesis} and \cite{caron_2007}.
@@
 \begin{figure}[h]
-        \center{\includegraphics[width=100mm]{../img/gpuddp_gm_3.pdf}}
+        \center{\includegraphics[width=82mm]{../img/gpuddp_gm_3.pdf}}
         \caption{Graphical model of the deletion variable and auxiliary variable transition kernel representation of the GPUDDPM}
@@
 We define the state of the sampler to consist of the assignment variables for all observations, $c_{1:N_{1},1}, \ldots, c_{1:N_{T}, T}$, the deletion variables for all observations, $d_{1:N_{1},1}, \ldots, d_{1:N_{T}, T}$, the sequence of cluster parameters for each active cluster, $\theta_{k,1}, \ldots, \theta_{k, T}$ and the $M$ associated auxiliary variables for each cluster $\theta_{k,t}$ at time $t$, $z_{k,t,1}, \ldots, z_{k, t, M}$.
-Formally, we can write
+We can write
 \begin{equation}
@@
 p(c_{i,t} = k | \text{State}) & \propto
-\prod_{i'=i}^{N_{t}}  \text{CRP}(c_{i',t} | \bold{m}_{t-1}, \alpha) \\
-& \times \prod_{t' = t+1}^{d_{i,t}}  \prod_{i'=1}^{N_{t'}}   \text{CRP}(c_{i',t'} | \bold{m}_{t'-1}, \alpha) \\
+\prod_{i'=i}^{N_{t}}  \text{C}(c_{i',t} | \bold{m}_{t-1}, \alpha) \\
+& \times \prod_{t' = t+1}^{d_{i,t}}  \prod_{i'=1}^{N_{t'}}   \text{C}(c_{i',t'} | \bold{m}_{t'-1}, \alpha) \\
  & \times
 \begin{cases}
-	P(\bold{x}_{i,t} | \theta_{k,t})                                        \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
+  F(\bold{x}_{i,t} | \theta_{k,t})                                        \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
   \int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta) d\theta    \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
@@
 \end{equation}
-% where $t$ is the time of observation $i$, and the distribution under the product $P(c_{i'} | m_{1:K_{t}^{'}, i'-1}, c_{1:i'-1})$ is given by 
-%\begin{equation}
-%P(c_{i} = k | m_{1:K_{t}^{'}, i-1}, c_{1:i-1}, \alpha) \propto 
-%\begin{cases}
-%m_{k, i-1} \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
-%\alpha \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
-%\end{cases}
-%\end{equation}
-where the distribution CRP is given by \eqref{crp_rep}, and the sizes of the clusters $\bold{m}_{t'-1}$ are calculated given the fact that $c_{i,t} = k$.
+where the C is given by \eqref{C_distro}, and the sizes of the clusters $\bold{m}_{t'-1}$ are calculated under the assumption that $c_{i,t} = k$.
 Note that the above integral can be determined analytically to give
@@
 In particular, the Metropolis-Hastings (M-H) algorithm is used to gain samples from the posterior distribution for the auxiliary variables; it involves constructing a markov chain with the sought posterior as its stationary distribution, and sampling sequentially from this constructed chain to get an approximate posterior sample.
-The M-H algorithm is outlined in Algorithm 1.%Algorithm~\ref{alg:MH}.
-\willie{am i citing wrong?}
-
-\begin{algorithm}[h!]
-\label{alg:MH}
-\caption{Metropolis Hastings Algorithm}
-\begin{algorithmic}[1]
-\STATE $f(x)$ is the stationary PDF from which $N$ samples are desired \hfill $\triangleright$ eq. \eqref{stationary_pdf}
-\STATE $q( \cdot | x_{i-1})$ is the proposal pdf \hfill $\triangleright$ eq. \eqref{proposal_distro}
-\STATE B is the number of burn in samples
-\STATE $x_{0} \leftarrow$ initialize
-%\STATE B $\leftarrow$ \# of burn in samples
-\FOR{$i = 1 : \text{B} + N$}
-\STATE Sample $x^{*} \sim q( \cdot | x_{i-1})$
-\STATE Sample $\alpha \sim \text{Uniform}(0,1)$
-\STATE $r_{accept} \leftarrow \frac{f(x^{*})q(x_{i-1} | x^{*})} {f(x_{i-1})q(x^{*} | x_{i-1})} $ \hfill $\triangleright$  eq. \eqref{accept_ratio}
-\IF{$\alpha < \text{min} \{ 1, r_{accept} \} $}
-\STATE $x_{i} \leftarrow x^{*}$
-\ELSE
-\STATE $x_{i} \leftarrow x_{i-1}$
-\ENDIF
-\ENDFOR
-\STATE Discard samples $x_{1}, \ldots, x_{\text{B}}$
-\STATE \textbf{Output}: Remaining samples $x_{1}, \ldots, x_{N}$
-\end{algorithmic}
-%\caption{Say thing at end}
-\end{algorithm}
+% The M-H algorithm is outlined in Algorithm 1.%Algorithm~\ref{alg:MH}. \willie{am i citing wrong?}
+
+% \begin{algorithm}[h!]
+% \label{alg:MH}
+% \caption{Metropolis Hastings Algorithm}
+% \begin{algorithmic}[1]
+% \STATE $f(x)$ is the stationary PDF from which $N$ samples are desired \hfill $\triangleright$ eq. \eqref{stationary_pdf}
+% \STATE $q( \cdot | x_{i-1})$ is the proposal pdf \hfill $\triangleright$ eq. \eqref{proposal_distro}
+% \STATE B is the number of burn in samples
+% \STATE $x_{0} \leftarrow$ initialize
+% %\STATE B $\leftarrow$ \# of burn in samples
+% \FOR{$i = 1 : \text{B} + N$}
+% \STATE Sample $x^{*} \sim q( \cdot | x_{i-1})$
+% \STATE Sample $\alpha \sim \text{Uniform}(0,1)$
+% \STATE $r_{accept} \leftarrow \frac{f(x^{*})q(x_{i-1} | x^{*})} {f(x_{i-1})q(x^{*} | x_{i-1})} $ \hfill $\triangleright$  eq. \eqref{accept_ratio}
+% \IF{$\alpha < \text{min} \{ 1, r_{accept} \} $}
+% \STATE $x_{i} \leftarrow x^{*}$
+% \ELSE
+% \STATE $x_{i} \leftarrow x_{i-1}$
+% \ENDIF
+% \ENDFOR
+% \STATE Discard samples $x_{1}, \ldots, x_{\text{B}}$
+% \STATE \textbf{Output}: Remaining samples $x_{1}, \ldots, x_{N}$
+% \end{algorithmic}
+% %\caption{Say thing at end}
+% \end{algorithm}
 
@@
 In short, we sample a new value for all $M$ auxiliary variables, denoted $z_{k,t,m}^{*}$, according to the following proposal distribution
-\begin{equation}
+\begin{align}
+\begin{split}
 \label{proposal_distro}
-z_{k,t,m}^{*}  \sim  P(z_{k,t,m} | \theta_{k,t}) = \mathcal{N}(z_{k,t,m}^{s} | \boldsymbol{\mu}_{k,t}, \Sigma_{k,t}) \mathcal{M}n(z_{k,t,m}^{c} | \bold{p}_{k,t})
-\end{equation}
+z_{k,t,m}^{*}  &\sim  P(z_{k,t,m} | \theta_{k,t}) \\
+&= \mathcal{N}(z_{k,t,m}^{s} | \boldsymbol{\mu}_{k,t}, \Sigma_{k,t}) \mathcal{M}n(z_{k,t,m}^{c} | \bold{p}_{k,t})
+\end{split}
+\end{align}
 compute the standard M-H acceptance ratio, which in this case (due to the proposal distribution choice) simplifies to 
@@
 
-\subsubsection{Sampling Deletion Variables $d_{i}$}
-Computing the posterior distribution (or a value proportional to this distribution) for each possible value that a given deletion variable may take on, may be accomplished in a manner similar to how samples from the posterior distribution of the assignment variables $c_{i}$ are gained in Section~\ref{sec:sample_assignments}; however, this may be computationally expensive due to the large number of potential deletion times.
-To remedy this, the M-H algorithm may again be used to sample from the posterior, where the prior distribution for the deletion times can be used to propose new samples $d_{i}^{*}$.
-As the ``lifetime'' of each assignment can be shown to be geometrically distributed, we can propose a new deletion time sample with
+\subsubsection{Sampling Deletion Variables $d_{i,t}$}
+Computing a value proportional to the posterior distribution for each possible value that a given deletion variable may take on may be accomplished in a manner similar to how samples from the posterior distribution of the assignment variables $c_{i,t}$ are gained in Section~\ref{sec:sample_assignments}; however, this may be computationally expensive due to the large number of possible deletion times.
+To remedy this, the M-H algorithm may again be used to sample from the posterior, where the prior distribution for the deletion times can be used to propose new samples $d_{i,t}^{*}$.
+As the ``lifetime'' $l_{i,t}$ of each assignment can be shown to be geometrically distributed, we can propose a new deletion time sample with
 \begin{align}
 \begin{split}
-l_{i}  &\sim  \text{Geo}(\rho)  \\
-d_{i}^{*}  &= l_{i} + t + 1
+l_{i,t}  &\sim  \text{Geo}(\rho)  \\
+d_{i,t}^{*}  &= l_{i,t} + t + 1
 \end{split}
@@
 
-% SMC inference section
-% ---------------------
-
-\subsection{SMC Inference}
+\subsection{SMC: Sequential Inference}
 \label{sec:SMC}
 
-This section details a Sequential Monte Carlo (SMC) sampler---also known as a particle filter---used to perform inference, which provides another method to carry out unsupervised detection and tracking.
-The SMC inference algorithm is given in Algorithm~\ref{alg:SMC}.
-In this algorithm, at each time step $t \in \{ 1, \ldots, T \}$, a number of samples refered to as ``particles'' are generated; each particle consists of a sample from the posterior distribution over the assignment for each observation, $c_{1,t}, \ldots, c_{N_{t}, t}$, parameters for each cluster $\theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and size after deletion for each cluster $m_{1,t}, \ldots, m_{K_{t},t}$.
+This section details a Sequential Monte Carlo (SMC) sampler---also known as a particle filter---used to perform inference, which provides another method for unsupervised detection and tracking.
+SMC inference operates in the orginal GPUDDPM formulation (defined in Section~\ref{sec:gpuddpm}).
+The algorithm is shown in Algorithm~\ref{alg:SMC}.
+At each time step $t \in \{ 1, \ldots, T \}$, a number of samples refered to as ``particles'' are generated; each particle consists of a sample from the posterior distribution over the assignment for each observation, $c_{1,t}, \ldots, c_{N_{t}, t}$, parameters for each cluster $\theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and size after deletion for each cluster $m_{1,t}, \ldots, m_{K_{t},t}$.
 A set of particles is sampled at each time step from relevant proposal distributions (described in Sections~\ref{sec:smc_proposal_1}, \ref{sec:smc_proposal_2}, and \ref{sec:smc_proposal_3}), a weight is computed for each particle, and a new set of particles are sampled from the set of weighted particles via a resampling process.
@@
 \end{cases} \\
-& \times \prod_{k=1}^{K} \text{DEL}(m_{k,t} | m_{k,t-1}, c_{1:N_{t-1}, t-1}, \rho)
+& \times \prod_{k=1}^{K} \text{D}(m_{k,t} | m_{k,t-1}, c_{1:N_{t-1}, t-1}, \rho)
 \end{split}
@@
 
-% \begin{algorithm}
-% \caption{Sequential Monte Carlo Inference for the GPUDDPM}
-% \label{alg:SMC}
-% \begin{algorithmic}[1]
-% \FOR{$t = 1 : T$}
-% \FOR{$l = 1 : L$}
-% \STATE $K_{t}^{(l)}  \leftarrow K_{t-1}^{(l)}$
-% \FOR{$i = 1 : N_{t}$}
-% \STATE Sample assigments from previous time sizes and cluster parameters $\theta_{k,t-1}$
-% \STATE Draw $c_{i} \sim P(c_{i} | m_{1:K_{t-1}, i-1}, \theta_{1:K_{t-1}, t-1} )$
-% \IF{$c_{i} = K_{t} + 1$}
-% \STATE Draw a new cluster $\theta_{K_{t}+1,t} \sim \mathbb{G}_{0}(\text{priors})$   \willie{distro after algo?}
-% \STATE $K_{t} \leftarrow K_{t} + 1$
-% \ENDIF
-% \ENDFOR
-% \FOR{$s = 1 : S$}
-% \FOR{$k = 1 : K$}
-% \STATE Sample new cluster params via aux (given curent assignments and prev clusters)
-% \ENDFOR
-% \FOR{$i = 1 : N_{t}$}
-% \STATE Sample new assignments (given current cluster params)
-% \STATE Draw $c_{i} \sim P(c_{i} | m_{1:K_{t}, i-1}, \theta_{1:K_{t}, t} )$
-% \IF{$c_{i} = K_{t} + 1$}
-% \STATE Draw a new cluster $\theta_{K_{t}+1,t} \sim \mathbb{G}_{0}(\text{priors})$   \willie{distro after algo?}
-% \STATE $K_{t} \leftarrow K_{t} + 1$
-% \ENDIF
-% \ENDFOR
-% \STATE Sample new cluster sizes.
-% \ENDFOR
-% \ENDFOR
-% \STATE Only keep some of the $L$ samples.
-% \ENDFOR
-% \end{algorithmic}
-% \end{algorithm}
-
-%%%%%
-% in SMC algorithm: still need to include computing r and make sure things are correct
-%%%%%
-
-\begin{algorithm}[!]
+\begin{algorithm*}[!]
 \caption{Sequential Monte Carlo Inference for the GPUDDPM}
@@
 \IF{$s=S$}
-\STATE Sample $m_{k,t+1}^{(l)} \sim \text{DEL}(m_{k,t}^{(l)}, c_{1:N_{t}, t}^{(l)}, \rho)$  \hfill $\triangleright$ eq.~\eqref{del_step}
+\STATE Sample $m_{k,t+1}^{(l)} \sim \text{D}(m_{k,t}^{(l)}, c_{1:N_{t}, t}^{(l)}, \rho)$  \hfill $\triangleright$ eq.~\eqref{del_step}
 \ENDIF
@@
 \end{algorithmic}
-\end{algorithm}
-
-\subsubsection{Proposal Distribution for Assignments $c_{i}$}
+\end{algorithm*}
+
+\subsubsection{Proposal Distribution for Assignments $c_{i,t}$}
 \label{sec:smc_proposal_1}
 
-The probability of assignments given current cluster sizes, cluster parameters, and the Dirichlet Process granularity parameter $\alpha$ can be written as
+The probability of assignments given current cluster sizes, cluster parameters, and the Dirichlet Process concentration parameter $\alpha$ can be written as
 \begin{equation}
@@
 P \left( c_{i,t} | m_{1:K_{t}, t}, \theta_{1:K_{t}, t}, \alpha \right) &\propto
-\text{CRP}(m_{1:K_{t}, t}, \alpha) \\ &\times
+\text{C}(m_{1:K_{t}, t}, \alpha) \\ &\times
 \begin{cases}
-P(\bold{x}_{i,t} | \theta_{c_{i,t},t}) \hspace{2mm} \text{if} \hspace{2mm} k \leq K_{t-1} \\
+F(\bold{x}_{i,t} | \theta_{c_{i,t},t}) \hspace{2mm} \text{if} \hspace{2mm} k \leq K_{t-1} \\
 \int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta)d\theta \hspace{2mm} k > K_{t-1}
@@
 \end{equation}
-where CRP is defined in \eqref{CRP}, and the integral $\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta)d\theta$ can be determined analytically, and is given in \eqref{marginal_obs_prob}.
+where C is defined in \eqref{C_distro}, and the integral \\
+$\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta)d\theta$ can be determined analytically, and is given in \eqref{marginal_obs_prob}.
 
@@
 \begin{split}
-& \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})}	{P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) }  = 
-\frac{P(\bold{x}_{1:N_{t},t} | \theta_{1:K_{t}^{(l)}}) P(c_{1:N_{t}, t}^{(l)} | m_{1:K_{t},t}^{(l)}) )}{P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t})}
-\\
-%
-% & \frac{P \left( \bold{x}_{1, t} | \theta_{c_{1, t}^{(l)}, t}^{(l)} \right) \times \ldots \times P \left( \bold{x}_{N_{t}, t} | \theta_{c_{N_{t}, t}^{(l)}, t}^{(l)} \right) \times
-% P \left( c_{1, t}^{(l)} | m_{1:K_{t}^{(l)}, t}^{(l)} \right) \times \ldots \times P \left( c_{N_{t}, t}^{(l)} | m_{1:K_{t}^{(l)}, t}^{(l)} \right)}{b}
-%
-\end{split}
-\end{equation}
-where $P(\bold{x}_{1:N_{t},t} | \theta_{1:K_{t}^{(l)}})$ is defined by \eqref{likelihood}, $P(c_{1:N_{t}, t}^{(l)} | m_{1:K_{t},t}^{(l)}))$ is defined by \eqref{CRP}, and $P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t})$ is defined by \eqref{mcmc_assig_vars}.
+& \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})} {P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) }  \\
+& = \frac{P(\bold{x}_{1:N_{t},t} | \theta_{1:K_{t}^{(l)}}) P(c_{1:N_{t}, t}^{(l)} | m_{1:K_{t},t}^{(l)}) )}{P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t})}
+\end{split}
+\end{equation}
+where $P(\bold{x}_{1:N_{t},t} | \theta_{1:K_{t}^{(l)}})$ is defined by \eqref{likelihood}, $P(c_{1:N_{t}, t}^{(l)} | m_{1:K_{t},t}^{(l)}))$ is defined by \eqref{C_distro}, and $P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t})$ is defined by \eqref{mcmc_assig_vars}.
 
@@
 
-%%%%%
-% need to include particle resampling strategies
-%%%%%
-
-\subsection{Modeling Background Noise}
-
-% Noisy extraction procedures may result in spurrious data points that do not correspond to any foreground object.
-We would like to incorporate these points into our model.
-To do this, we modify our base distribution $\mathbb{G}_{0}$ to be a two component mixture, where one component is the base distribution and the other is a multivariate normal distribution with a fixed `flat' covarance $\sigma = [100, 0; 0, 100]$ multiplied with an `even' multinomial distribution with a fixed parameter $\bold{p} = (1/K, \ldots, 1/K)$.
-Noisy extraction procedures may result in spurrious data points that do not correspond to any foreground object.
-We would like to incorporate these points into our model.
-To do this, we modify our model to be a two component mixture model, where one component is the previous model (the GPUDDPM) and the other is a multivariate normal distribution with with a fixed `flat' covarance $\sigma = [100, 0; 0, 100]$ multiplied with an `even' multinomial distribution with a fixed parameter $\bold{p} = (1/K, \ldots, 1/K)$.
-This normal-multinomial component is left intentionally broad, and only the mean parameter of the normal distribution is inferred.
-We can write this generative model as
-\begin{align}
-\label{background_model}
-\begin{split}
-\gamma_{i,t} | p_{1}, p_{2} &\sim \text{Discrete}(p_{1}, p_{2})\\
-m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho  &\sim \text{DEL}(m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) \\
-\theta_{k, t} | \theta_{k, t-1}   &\sim
-\begin{cases}
-	P(\theta_{k, t} | \theta_{k, t-1}) \hspace{2mm} \text{if} \hspace{2mm} k \in \{ 1, \ldots, K_{t} \} \\
-	\mathbb{G}_{0}   \hspace{2mm} \text{if} \hspace{2mm} k = K_{t+1}
-\end{cases} \\
-c_{i,t} | m_{1:K_{t},t}, \hspace{1mm} \alpha  &\sim  \text{CRP} (m_{1:K_{t},t}, \hspace{1mm} \alpha) \\
-\bold{x}_{i,t} | c_{i,t}, \theta_{1:K_{t}, t}, \gamma_{i,t}, \phi &\sim
-\begin{cases}
-	F(\theta_{c_{i,t}, t}) \hspace{2mm} \text{if} \hspace{2mm} \gamma_{i,t} = 1 \\
-	\phi \hspace{2mm} \text{if} \hspace{2mm} \gamma_{i,t} = 2
-\end{cases} \\
-\end{split}
-\end{align}
-where $\phi$ represents the normal-multinomial component distribution defined above.
-
-%%%%%
-% define more rigorously, and then say why we don't include in experiments?--- or perhaps include for an example in experiments
-%%%%%
-
 \subsection{From Inference to Tracking Results}
-\label{sec:inference_to_results}
+\label{sec:inferencetotrackingresults}
+
 The multivariate normal-multinomial model on which inference is described in this section can be used to gain object detection and tracking results.
-The nature of object detection and tracking results may vary depending on the application or performance metrics used (see Section~\ref{sec:performance_metrics}); in this study, we desire the centroid position and approximate spatial region for each moving object, for each frame that this object can be seen within the video.
+The nature of object detection and tracking results may vary depending on the application or performance metrics used (see Section~\ref{sec:performanceevaluationmetrics}); in this study, we desire the centroid position and approximate spatial region for each moving object, for each frame that this object can be seen within the video.
 We choose to make our detection and tracking results depend completely on the inferred parameters of the multivariate normal distribution of each cluster.
 Each inferred cluster was taken to be a distinct detected object, and the sequence of means and covariance matrices for a given cluster were used, respectively, to determine the position and spatial region of a given object over a sequence of time steps.
-In particular, the mean parameter was taken to be the centroid of an object, and a 2-dimensional oval denoting a boundary around the mean that contains a specified percentage of the data (where the specification for this oval is given in Section~\ref{sec:pets2000_2001}) was taken to be the spatial region of the object.
+In particular, the mean parameter was taken to be the centroid of an object, and a 2-dimensional oval denoting a boundary around the mean that contains a specified percentage of the data (where the specification for this oval is given in Section~\ref{sec:pets2000and2001}) was taken to be the spatial region of the object.
 Furthermore, as both provided inference methods use sampling, and each sample represents a result, we must choose a way to narrow these down to a single result.
@@
 \subsection{Performance Evaluation Metrics}
-\label{sec:performance_metrics}
+\label{sec:performanceevaluationmetrics}
 
@@
 The numbers of tracks displaying both of these failures are factors in the performance metrics used in this study.
-
-%, where a numerical search algorith---the Hungarian algorithm \cite{munkres_1957}--is used to find an optimal mapping from ground-truth to output tracks without requiring the performance metric value of all possible mapping combinations to be computed (reducing the time complexity from factorial to polynomial time)
-
-% \subsubsection{MODA and MOTA}
-
-% The two metrics used to quantify results in this study are refered to as the Multiple Object Detection Accuracy (MODA) and Multiple Object Tracking Accuracy (MOTA).
-These metrics, along with the the ground-truth tracks of all objects in the video sequence and the the mapping as defined in the previous section, can be used to evaluate detection and tracking accurancy of the video.
-For a given video frame $t$, we define MODA as
-% \begin{equation}
-% MODA = 1 - \frac{m^{t} + e^{t}}{N_{g}^{t}}
-% \end{equation}
-% where $m^{t}$ is the number of missed object detections (ie true-negative detections), $e^{t}$ is the number of erroneous object detections (ie false positive detections), and $N_{g}^{t}$ is the number of ground truth objects (all at frame $t$).
-
-% For a sequence of video frames $t \in \{ 1, \ldots, T \}$ we define MOTA as
-% \begin{equation}
-% MOTA = 1 - \frac{\sum_{t=1}^{T} m^{t} + e^{t} + s_{t}}{\sum_{t=1}^{T} N_{g}^{t}}
-% \end{equation}
-% where $s$ denotes the number of instances at frame $t$ where an object's identity switches from a previous frame (ie where the object's mapping differs between frame $t$ and frame $t-1$),  and the other terms are defined the same as above.
 
@@
 Consequentially, to find the spatial overlap between results and ground-truth, which is intrinsic to both metrics, a rectangular bounding box was needed per object per time step from the results of the algorithm.
-We took the maximal and minimal axially aligned values of the oval produced by our algorithm (as described in Section~\ref{sec:inference_to_results}) to be the sides of a representative bounding box for a given object at a given frame.
+We took the maximal and minimal axially aligned values of the oval produced by our algorithm (as described in Section~\ref{sec:inferencetotrackingresults}) to be the sides of a representative bounding box for a given object at a given frame.
 
 \subsection{Synthetic Video Datasets}
-\label{sec:syntheticvideos}
-
-Synthetic videos provide controlled scenarios in which aspects of the technique presented in this paper can be tested.
-Each of the following synthetic videos consists of a sequence of 200 images (each of size $500$ x $500$ pixels) containing a number of smaller colored squares of different (and potentially time-varying) sizes moving at varied speeds and trajectories over a black background.
-The synthetic videos contain instances of occlusion (where one or more objects are briefly hidden) and objects with time-varying appearances and behaviors, as these are examples of occurances that notoriously decrease the accuracy of detection and tracking.
-After each video was constructed, the extraction procedure described in Section~\ref{sec:modelspec_extraction} (using $L=3$) and inference procedures described in \ref{sec:inference} were carried out to return a sequence of multivariate-normal-parameters (means and covariance matrices) from which a sequence of positions and ovals approximating the shape of a tracked object over each frame that it is present in the video can be determined (as outlined in Section~\ref{sec:inference_to_results}).
-
-% \subsubsection{Tracking by Appearance}
-
-The first synthetic video experiment aimed to test the ability of the model and inference procedure to maintain the identity of independent objects based on color information alone.
-Two videos were constructed, both containing a red square (rgb value $[255,0,0]$ and size $20$ x $20$ pixels) and a blue square (rgb value $[0,0,255]$ and size $20$ x $20$ pixels).
-In both videos, the squares begin at opposite sides of the scene at frame $f=1$ and travel towards each other, arriving at the same location at $f=100$ (where the blue square occludes the red square).
-The second half of the two videos differ in that both squares in the first video continue in the same direction and end at the other's starting positions at $f=200$, and both squares in the second video reverse directions and end at their initial starting positions at $f=200$.
-The frame difference extraction yields identical spatial features in both videos; hence, successful tracking depends fully on the incorporation of color information into the model.
-
-Parameters were set to the same values for inference on both videos, using the method outlined in Section~\ref{sec:parameter_recap} to choose parameters.
-The chosen values were $\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (5, \ldots, 5)$.
-Inference was carried out using the MCMC algorithm (Section~\ref{sec:MCMC}); the posterior sample taken as the result correctly tracked both colored squares through occlusion in both videos, and is shown in Figure~\ref{fig:synth_one_plot}.
+\label{sec:syntheticvideodatasets}
 
 %%%%%
-% I still need to include performance metric results for these synthetic datasets.}.
+% I still need to include performance metric results for these synthetic datasets...maybe
 %%%%%
-
-\begin{figure}
+\begin{figure*}[!]
   \centering               
@@
   \label{fig:synth_one_plot}
-\end{figure}
-
-% \subsubsection{Tracking though Occlusion and Appearance Shift}
+\end{figure*}
+
+Synthetic videos provide controlled scenarios in which aspects of the technique presented in this paper can be tested.
+Each of the following synthetic videos consists of a sequence of 200 images (each of size $500$ x $500$ pixels) containing a number of smaller colored squares of different (and potentially time-varying) sizes moving at varied speeds and trajectories over a black background.
+The synthetic videos contain instances of occlusion (where one or more objects are briefly hidden) and objects with time-varying appearances and behaviors, as these are examples of occurances that notoriously decrease the accuracy of detection and tracking.
+After each video was constructed, the extraction procedure described in Section~\ref{sec:dataextraction} (using $L=3$) and inference procedures described in \ref{sec:inference} were carried out to return a sequence of multivariate-normal-parameters (means and covariance matrices) from which a sequence of positions and ovals approximating the shape of a tracked object over each frame that it is present in the video can be determined (as outlined in Section~\ref{sec:inferencetotrackingresults}).
+
+The first synthetic video experiment aimed to test the ability of the model and inference procedure to maintain the identity of independent objects based on color information alone.
+Two videos were constructed, both containing a red square (rgb value $[255,0,0]$ and size $20$ x $20$ pixels) and a blue square (rgb value $[0,0,255]$ and size $20$ x $20$ pixels).
+In both videos, the squares begin at opposite sides of the scene at frame $f=1$ and travel towards each other, arriving at the same location at $f=100$ (where the blue square occludes the red square).
+The second half of the two videos differ in that both squares in the first video continue in the same direction and end at the other's starting positions at $f=200$, and both squares in the second video reverse directions and end at their initial starting positions at $f=200$.
+The frame difference extraction yields identical spatial features in both videos; hence, successful tracking depends fully on the incorporation of color information into the model.
+
+Parameters were set to the same values for inference on both videos, using the method outlined in Section~\ref{sec:parameter_recap} to choose parameters.
+The chosen values were $\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (5, \ldots, 5)$.
+Inference was carried out using the MCMC algorithm (Section~\ref{sec:MCMC}); the posterior sample taken as the result correctly tracked both colored squares through occlusion in both videos, and is shown in Figure~\ref{fig:synth_one_plot}.
 
@@
 
-%Both the cluster parameter sample and data are overlaid on three semi-transparent planes showing the objecs in the video in three frames.
-\willie{still need to include performance metric results for this synthetic dataset}.
-Groundtruth for the synthetic videos
-
-% \begin{figure}
-%   \centering               
-%   % \subfloat[]{\includegraphics[width=0.35\textwidth, angle=30]{../img/synth2_img}}
-%   \subfloat[]{\includegraphics[width=0.55\textwidth]{../img/synth2_img_b}}
-%   \subfloat[]{\includegraphics[height=0.3\textheight]{../img/synth2_alphaimg_hd.png}}
-%   \caption{(a.) Video frame $f=10$ from synthetic experiment two.
-(b.) Sample from posterior of the state for synthetic experiment two, where color represents assignment and the mean and standard deviation are shown.
-All are overlaid on three, time-varying images from synthetic experiment.
-The three objects of time varying size, speed, and position are tracked through simultaneous occlusion.
-\willie{will show assignments by marker type in addition to color}}
-%   \label{fig:synth_two_plot}
-% \end{figure}
-
-% \begin{figure}[h]
-%         \center{\includegraphics[width=65mm]{../img/synth2_alphaimg_hd.png}}
-%         \caption{\label{fig:synth2_alphaimg} Sample from posterior of the state for synthetic experiment two, where color represents assignment and the mean and standard deviation are shown.
-Both the sample and data are overlaid on three images from synthetic experiment.
-The three objects of time varying size, speed, and position are tracked through simultaneous occlusion.
-\willie{will show assignments by marker type in addition to color}}
-% \end{figure}
-
-% \begin{figure}[h]
-%         \center{\includegraphics[width=0.99\textwidth]{../img/synth2_alphaimg_hd.png}}
-%         \caption{\label{fig:synth2_alphaimg} Extracted data and sample from the posterior of the model overlaid on three, time-varying images from synthetic experiment 2.}
-% \end{figure}
-
 \subsection{Benchmark Video Datasets}
+\label{sec:benchmarkvideodatasets}
 
@@
 
-%\subsubsection{PETS2000}
-
 \subsubsection{PETS2000 and PETS2001}
-\label{sec:pets2000_2001}
+\label{sec:pets2000and2001}
 
@@
 Due to the computation required for the MCMC batch inference method (discussed further in Section~\ref{sec:discussion}), only the final 1000 frames of the video were used from both datasets.
-Extraction was performed with frame differencing as described in Section~\ref{sec:modelspec_extraction}, using $L=3$.
+Extraction was performed with frame differencing as described in Section~\ref{sec:dataextraction}, using $L=3$.
 Parameter values were chosen in the same way, and set to the same values, as in the the synthetic experiments ($\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (5, \ldots, 5)$).
@@
 
-\begin{figure}[h]
+\begin{figure*}[!]
   \centering
@@
   \label{fig:benchmark_results_1}
-\end{figure}
-
-\begin{figure}[h]
+\end{figure*}
+
+\begin{figure*}[!]
   \centering             
@@
   \label{fig:pets2001_imgs}
-\end{figure}
-
-\begin{figure}
+\end{figure*}
+
+\begin{figure*}[!]
   \centering               
@@
   \label{fig:pets2001_overlay}
-\end{figure}
-
-To calculate performance metrics (both SFDA and ATA), one must specify a confidence value that allows the oval representing the region occupied by an object to be computed from the inferred covariance matrix of each cluster (as discussed in Section~\ref{sec:inference_to_results}).
+\end{figure*}
+
+To calculate performance metrics (both SFDA and ATA), one must specify a confidence value that allows the oval representing the region occupied by an object to be computed from the inferred covariance matrix of each cluster (as discussed in Section~\ref{sec:inferencetotrackingresults}).
 The performance metrics were found for a range of confidence intervals, and these curves, for the PETS2000 and PETS2001 video are shown in Figure~\ref{fig:pm_conf}.
 
-\begin{figure}[h]
+\begin{figure*}[!]
   \centering             
@@
   \label{fig:pm_conf}
-\end{figure}
-
-% 3-panel alpha-img figure that I removed
-
-% \begin{figure}[h]
-% 	\centering
-%     \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/synth2_alphaimg_hd.png}}
-%     \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2001_alphaimg_hr.png}}
-% 	\subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2000_4objects.png}}
-%     \caption{\label{fig:synth2_alphaimg} Sample from posterior of the state for synthetic experiment two (a), and benchmark datasets PETS2001 (b) and PETS2000 (c), where color represents assignment and the mean and standard deviation are shown.
-In each plot, the sample and data are overlaid on three images from the experiment.}
-% \end{figure}
-
-% old figure ideas:
-
-% \begin{figure}[h]
-%         \center{\includegraphics[width=65mm]{../img/pets2001_alphaimg_hr.png}}
-%         \caption{\label{fig:pets2001_alphaimg} Extracted data and sample from posterior of the model overlaid on three frames from the PETS2001 video.}
-% \end{figure}
-
-% \begin{figure}[h]
-% 	\centering
-% 	\subfloat[]{\includegraphics[width=0.58\textwidth]{../img/pets2000_results_1}}
-% 	\subfloat[]{\includegraphics[width=0.42\textwidth]{../img/pets2000_4objects.png}}
-% 	\caption{Extracted data and a sample from posterior of the model (a) overlaid on a frame from the PETS2000 video (b).}
-% 	\label{fig:pets2000_results}
-% \end{figure}
-
-% \begin{figure}
-%   \centering               
-%   \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2000_front.pdf}}
-%   \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2000_back.pdf}}
-%   \caption{PETS2000 dataset results, from a front view (a) and a back view (b).
-Both plots show a sample from the posterior distribution of the state, where the vertical axis denots time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.}
-%   \label{fig:pets2000_results}
-% \end{figure}
+\end{figure*}
 
 \subsubsection{PETS2009/2010}
-
-A video dataset used in both the PETS2009 and PETS2010 conferences, called `S2.L1 at time sequence 12.34' was chosen for experimentation due to its prominence in a number of studies \cite{ellis_2010}.
+\label{sec:pets20092010}
+
+A video dataset used in both the PETS2009 and PETS2010 conferences, called ``S2.L1 at time sequence 12.34'' was chosen for experimentation due to its prominence in a number of studies \cite{ellis_2010}.
 %%%%%
@@
 
-Extraction was performed with frame differencing as described in Section~\ref{sec:modelspec_extraction}, using $L=3$ and parameters for the model were chosen, using the method outlined in Section~\ref{sec:parameter_recap}, to be $\alpha = 0.1, \rho = 0.8, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 6, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (3, \ldots, 3)$.
+Extraction was performed with frame differencing as described in Section~\ref{sec:dataextraction}, using $L=3$ and parameters for the model were chosen, using the method outlined in Section~\ref{sec:recapofparameters}, to be $\alpha = 0.1, \rho = 0.8, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 6, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (3, \ldots, 3)$.
 Additionally, as with the other video datasets, ground-truth bounding boxes around each object were authored using the ViPER tool.
 
-The SMC inference algorithm yielded a sample from the posterior distribution of the model, from which the object detection and tracking results were obtained (as described in Section~\ref{sec:inference_to_results}).
+The SMC inference algorithm yielded a sample from the posterior distribution of the model, from which the object detection and tracking results were obtained (as described in Section~\ref{sec:inferencetotrackingresults}).
 In Figure \ref{fig:pets2009_results}, a sample from the posterior distribution over the cluster parameters is overlayed on the extracted data over a sequence of frames, where the assignment of each data point is represented by color and marker type.
 
-\begin{figure}[!]
+\begin{figure*}[!]
   \centering
@@
   \label{fig:pets2009_results}
-\end{figure}
+\end{figure*}
+
+\begin{figure*}[!]
+  \centering               
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2009_f3.pdf}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2009_f6.pdf}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2009_f13.pdf}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2009_f14.pdf}}
+  \caption{Four frames from the PETS2009/2010 sequence with one posterior sample mean and covariance matrix representation shown for each frame (and for the previous 20 frames).}
+  \label{fig:pets2009_overlay}
+\end{figure*}
 
@@
 Figure~\ref{fig:sfda_and_ata_comparison} shows performance metric results for comparison (data published with permission from the authors of \cite{ellis_2010}).
-\begin{figure}[!]
+\begin{figure*}[!]
   \centering
@@
   \label{fig:sfda_and_ata_comparison}
-\end{figure}
+\end{figure*}
 
 \subsubsection{Sensitivity Analysis}
-\label{sec:sens_analysis}
+\label{sec:sensitivityanalysis}
 
@@
 
-The following shows the SFDA sensitivity analysis results. The largest two values are shown in bold.
-\begin{center}
-\begin{tabular}[!]{l l  l  l  l  l  l}
-  & & & $\hspace{8mm}\rho$ & & & \\
-  SFDA  &  	& $0.7$ & $0.75$ & $0.8$ & $0.85$ & $0.9$  \\  
-  &$0.01$ 	& $0.4121$ & $0.4262$ & $0.4219$ & $\bold{0.4331}$ & $0.4277$  \\
-  &$0.1$  	& $\bold{0.4336}$ & $0.4184$ & $0.4200$ & $0.4316$ & $0.4321$  \\ 
-$\alpha$ 	& $1$    & $0.4084$ & $0.4209$ & $0.4205$ & $0.4290$ & $0.4330$  \\  
-  &$10$   	& $0.3992$ & $0.4130$ & $0.4289$ & $0.4232$ & $0.4226$  \\  
-  &$100$  	& $0.3655$ & $0.3739$ & $0.3636$ & $0.3755$ & $0.3672$  \\
-\end{tabular}
-\end{center}
-
-The following shows the ATA sensitivity analysis results. The largest two values are shown in bold.
-\begin{center}
-\begin{tabular}[!]{l l  l  l  l  l  l}
-		& & & $\hspace{8mm}\rho$ & & & \\
- 	 ATA  & 	 	& $0.7$ & $0.75$ & $0.8$ & $0.85$ & $0.9$  \\  
-	  &$0.01$ 		& $0.2032$ & $0.2017$ & $0.2262$ & $0.1945$ & $0.1914$  \\
-		  &$0.1$ 	& $0.1964$ & $0.2042$ & $0.2401$ & $0.1990$ & $0.2017$  \\
-$\alpha$ & $1$  	& $0.2016$ & $0.2365$ & $0.2766$ & $0.2347$ & $0.2766$  \\   
-  		&$10$   	& $0.2277$ & $0.2284$ & $0.2613$ & $\bold{0.2781}$ & $0.2732$  \\  
- 		 &$100$ 	& $0.2468$ & $0.2645$ & $0.2725$ & $0.2670$ & $\bold{0.2843}$  \\
-\end{tabular}
-\end{center}
+% The following shows the SFDA sensitivity analysis results. The largest two values are shown in bold.
+% \begin{center}
+% \begin{tabular}[!]{l l  l  l  l  l  l}
+%   & & & $\hspace{8mm}\rho$ & & & \\
+%   SFDA  &   & $0.7$ & $0.75$ & $0.8$ & $0.85$ & $0.9$  \\  
+%   &$0.01$   & $0.4121$ & $0.4262$ & $0.4219$ & $\bold{0.4331}$ & $0.4277$  \\
+%   &$0.1$    & $\bold{0.4336}$ & $0.4184$ & $0.4200$ & $0.4316$ & $0.4321$  \\ 
+% $\alpha$  & $1$    & $0.4084$ & $0.4209$ & $0.4205$ & $0.4290$ & $0.4330$  \\  
+%   &$10$     & $0.3992$ & $0.4130$ & $0.4289$ & $0.4232$ & $0.4226$  \\  
+%   &$100$    & $0.3655$ & $0.3739$ & $0.3636$ & $0.3755$ & $0.3672$  \\
+% \end{tabular}
+% \end{center}
+
+% The following shows the ATA sensitivity analysis results. The largest two values are shown in bold.
+% \begin{center}
+% \begin{tabular}[!]{l l  l  l  l  l  l}
+%     & & & $\hspace{8mm}\rho$ & & & \\
+%    ATA  &     & $0.7$ & $0.75$ & $0.8$ & $0.85$ & $0.9$  \\  
+%     &$0.01$     & $0.2032$ & $0.2017$ & $0.2262$ & $0.1945$ & $0.1914$  \\
+%       &$0.1$  & $0.1964$ & $0.2042$ & $0.2401$ & $0.1990$ & $0.2017$  \\
+% $\alpha$ & $1$    & $0.2016$ & $0.2365$ & $0.2766$ & $0.2347$ & $0.2766$  \\   
+%       &$10$     & $0.2277$ & $0.2284$ & $0.2613$ & $\bold{0.2781}$ & $0.2732$  \\  
+%      &$100$   & $0.2468$ & $0.2645$ & $0.2725$ & $0.2670$ & $\bold{0.2843}$  \\
+% \end{tabular}
+% \end{center}
 
@@
 
-% By inspecting the figures, one can see that there is a point where the SFDA and ATA both achieve a maximal value.
-
-\begin{figure}[!]
+\begin{figure*}[!]
   \centering             
@@
   \label{fig:sens_surf}
-\end{figure}
-
-% \begin{figure}
-%   \centering             
-%   \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2009_result1.png}}
-%   \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2009_result2.png}}
-%   \subfloat[]{\includegraphics[width=0.33\textwidth]{../img/pets2009_result3.png}}
-%   \caption{Surface of the sensitivity performance metric tables}
-%   \label{fig:sens_surf}
-% \end{figure}
-
-% \begin{figure}
-%   \centering             
-%   \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_1}}
-%   \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_2}}
-%   \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_3}}
-%   \subfloat[]{\includegraphics[width=0.25\textwidth]{../img/pets2009_4}}
-%   \caption{Four frames $f=10, 25, 35, 75$ from the PETS2009 sequence with one posterior sample mean and covarance matrix overlaid.
-\willie{will show path trail over prev 10 frames}\willie{these are old images!...
-haven't replaced yet}}
-%   \label{fig:pets2009_imgs}
-% \end{figure}
-
-% \begin{figure}
-% 	\centering
-% 	\includegraphics[width=3in]{../img/pets2009_results}
-% 	% \subfloat[]{\includegraphics[width=0.52\textwidth]{../img/pets2001_results_1}}
-% 	% \subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2001_results_2}}
-% 	\caption{Sample from posterior of the state for the PETS2009 dataset, where color represents assignment and the mean and standard deviation are shown.
-\willie{will show assignments by marker type in addition to color}}
-% 	\label{fig:pets2009_results}
-% \end{figure}
-
-% \subsection{New Video Datasets}
-
-% The two video datasets in this section are video datasets first used in this paper. The datasets and groundtruth detailing the positions of each obect in each video will be published online.
-
-% \subsubsection{Ants Video?}
-
-% \subsubsection{Cells Video?}
-
-\section*{Conclusion}
-\label{sec:discussion}
+\end{figure*}
+
+% Conclusion
+% ----------
+
+\section{Conclusion}
+\label{sec:conclusion}
 
@@
 objects hard to track through occlusion becuase they are small, similar appearance, and our algorithm doesn’t assume dynamics to keep robust (we could, probably for better accuracy but less generality).
-background (appearance) is like a natural prior for remaining stationary
-
-%\section*{Conclusion}
+background (appearance) is like a natural prior
 
@@
 
+% Model Overview
+% --------------
+
+\appendix
+
+\section{Model Background}
+\label{sec:modelbackground}
+
+Dirichlet Process Mixture (DPM) models (Section~\ref{sec:dpmixture}) fall under the heading of Bayesian nonparametric (BNP) models.
+These models have been widely used in the past decade to perform nonparametric density estimation and cluster analysis.
+Since we have time series data and wish to find spatiotemporal clusters, we are interested in using a closely related class of models known as Dependent Dirichlet Process Mixture (DDPM) models (Section~\ref{sec:ddpmixture}).
+These models are particularly useful for estimating the number of latent classes (clusters) in time dependent data.
+In this work, estimating the number of clusters in video extraction data is equivalent to estimating the number of distinct objects in a video.
+
+This section provides background information that allows us to introduce DDPM models, describes the motivation behind our use of these models, and gives a definition of the particular DDPM chosen, known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM) model (Section~\ref{sec:gpuddpmixture}).
+
+\subsection{Finite Mixture Model}
+\label{sec:finitemixture}
+
+A finite mixture model can be thought of as a probability distribution for an observation $x_{i}$ formulated as a linear combination of K mixture components (which we also refer to as `clusters'), where each mixture component is a probability distribution for $x_{i}$ with some parametric form, and the coefficients of the linear combination sum to one.
+The finite mixture model can be written as
+\begin{equation}
+P(x_{i}) = \sum_{k=1}^{K} P(c_{i} = k)P(x_{i}|\theta_{k})
+\end{equation}
+$\forall i \in \{ 1, \ldots, N \}$, where $c_{i} \in \{ 1, \ldots, K \}$ denotes the assignment of $x_{i}$ to a given mixture and $\theta_{k}$ denotes the parameters of the $k^{\text{th}}$ mixture component.
+Note that by choosing $P(c_{i} = k)$ as coefficients of the linear combination, it is ensured that these coefficients sum to one.
+We also define $p_{k} := P(c_{i} = k)$ for $k \in \{ 1, \ldots, K \} $.
+We can therefore write this model generatively as
+\begin{align}
+\begin{split}
+  c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
+  x_{i}|c_{i}, \theta_{c_{i}}  &\sim  \text{F}(\theta_{c_{i}})
+\end{split}
+\end{align}
+$\forall i \in \{ 1, \ldots, N \}$, where the $x_{i}$ are observations, the $c_{i}$ are the mixture component assignments associated with each observation, the $\theta_{c_{i}}$ are parameters defining the $c_{i}^{\text{th}}$ mixture component (i.e.
+the distribution to be mixed, $\text{F}(\theta_{c_{i}})$), and the ``Discrete'' distribution refers to a multinomial distribution whose parameters are a 1-of-K vector (i.e.
+a vector of counts that sums to one).
+
+\subsection{Bayesian (Finite) Mixture Model}
+\label{sec:bayesianmixture}
+
+The finite mixture model of Section~\ref{sec:finitemixture} can be extended to a Bayesian mixture model by viewing parameters that were previously point values, $\theta_{c_{i}}$ (the mixture component parameters) and $p_{1}, \ldots, p_{K}$ (the mixture component assignment weights), as random variables and providing each with a prior distribution.
+In this case, the prior distribution $\mathbb{G}_{0}$ is placed on the mixture component parameters, and the prior distribution $\text{Dir}(\alpha/K, \ldots, \alpha/K)$ is placed on the mixture component assignment weights.
+The resulting Bayesian mixture model can be formulated generatively as
+\begin{align}
+\begin{split}
+\label{bayesian_mixture_model}
+  p_{1}, \ldots, p_{K}  &\sim  \text{Dir}(\alpha/K, \ldots, \alpha/K)\\
+  \theta_{1}, \ldots, \theta_{K}  &\sim  \mathbb{G}_{0} \\
+  % \theta_{k}  &\sim  \mathbb{G}_{0}, \hspace{2mm} \forall k \in \{1, \ldots, K\} \\
+  c_{i}|p_{1}, \ldots, p_{K}  &\sim  \text{Discrete}(p_{1}, \ldots, p_{K}) \\
+  x_{i}|c_{i}, \theta_{c_{i}}  &\sim  \text{F}(\theta_{c_{i}})
+\end{split}
+\end{align}
+$\forall i \in \{ 1, \ldots, N \}$, where the $x_{i}$ are observations, the $c_{i}$ are the mixture component assignments associated with each observation, the $\theta_{k}$ are parameters defining the $k^{\text{th}}$ mixture component (i.e.
+the distribution to be mixed, $\text{F}(\theta_{k})$), the $\theta_{k}$ are drawn from a prior distribution $\mathbb{G}_{0}$, and $p_{1}, \ldots, p_{K}$ are drawn from a Dirichlet prior parameterized by $\alpha/K, \ldots, \alpha/K$.
+
+\subsection{Dirichlet Process}
+\label{sec:dirichletprocess}
+
+The Dirichlet process (DP), first introduced by \cite{ferguson_1973} in 1973, may be intuitively viewed as a probability distribution over discrete probability distributions.
+Accordingly, draws from a DP are probability mass functions (PMFs).
+A DP is parameterized by a base distribution $\mathbb{G}_{0}$, which is a probability distribution over a set $\Theta$, and a concentration parameter $\alpha \in \mathbb{R}_{+}$.
+We say that $G$ is a random PMF distributed according to a DP, written $G \sim \text{DP}(\alpha, \mathbb{G}_{0})$, if the following holds for all finite partitions $A_{1}, \ldots, A_{p}$ of $\Theta$:
+\begin{equation}
+(G(A_{1}), \ldots, G(A_{p})) \sim \text{Dir}(\alpha \mathbb{G}_{0}(A_{1}), \ldots, \alpha \mathbb{G}_{0}(A_{p}))
+\end{equation}
+Where `Dir' denotes a Dirichlet distribution.
+The parameters $\mathbb{G}_{0}$ and $\alpha$ may be intuitively viewed as the mean and precision of the DP.
+This is due to the fact that if the base distribution $\mathbb{G}_{0}$ is a distribution over $\Theta$, $A \subset \Theta$, and $G \sim \text{DP}(\alpha, \mathbb{G}_{0})$, then the following holds:
+\begin{equation}
+\mathbb{E}[G(A)] = \mathbb{G}_{0}(A)
+\end{equation}
+\begin{equation}
+\text{Var}[G(A)] = \mathbb{G}_{0}(A) (1 - \mathbb{G}_{0}(A)) / (\alpha + 1)
+\end{equation}
+Hence, the expectation of $G(A)$ is $\mathbb{G}_{0}$, the variance of $G(A) \rightarrow 0$ as $\alpha \rightarrow \infty$, and $G$ converges pointwise to $\mathbb{G}_{0}$ when $\alpha$ is unbounded.
+
+\subsection{Dirichlet Process (Infinite) Mixture Model}
+\label{sec:dpmixture}
+
+A DPM model, also refered to as an infinite mixture model, is an extension of the Bayesian mixture model described in Section~\ref{sec:bayesianmixture}.
+When using a DP as a prior in a Bayesian mixture model, $\Theta$ represents the set of parameters of the component mixture distributions.
+A DPM may be viewed as allowing the prior distribution over the mixture component parameters in a standard mixture model to be distributed according to a DP; this allows for modeling data where the true number of latent mixture components is unknown and arbitrarily large by letting the number of components remain unbounded (note that only a finite number of these components are assigned to the data).
+In particular, the DPM can be defined generatively as
+\begin{align}
+\begin{split}
+  \mathbb{G} | \alpha, \mathbb{G}_{0}  &\sim  \text{DP}(\alpha, \mathbb{G}_{0}) \\
+  \phi_{i} | \mathbb{G}  &\sim  \mathbb{G} \\
+  x_{i}|\phi_{i} &\sim \text{F}(\phi_{i})
+\end{split}
+\end{align}
+$\forall i \in \{ 1, \ldots, N \}$, where the $x_{i}$ are observations, the $\phi_{i}$ are parameters defining the mixture component from which the $i_{th}$ observation is drawn (i.e.
+the distribution to be mixed, $\text{F}(\phi_{i})$), and the $\phi_{i}$ are drawn from a prior distribution $\mathbb{G}$, which is in turn drawn from a DP with base distribution $\mathbb{G}_{0}$ and parameter $\alpha$.
+See \cite{gasthaus_2008} and \cite{gasthaus_thesis} for more details on this formulation.
+Note the difference between the indexing of the clusters in this model and the indexing in the previous two models.
+This formulation can be shown to be equivalent to the Bayesian mixture model defined in \eqref{bayesian_mixture_model}, when $K$ is taken to be unbounded; as a result, this model is sometimes called an infinite mixture model.
+If we let $K$ be the number of distinct mixture components assigned to observations using the above model, we can write the mixture components as $\theta_{1}, \ldots, \theta_{K}$.
+We also let $c_{1}, \ldots, c_{N}$ (where $c_{i} \in \{1, \ldots, K \}$) be class assignment variables that indicate the cluster to which observation $x_{i}$ is assigned.
+
+\subsection{Dependent Dirichlet Process Mixture Model}
+\label{sec:ddpmixture}
+
+The goal of DDPM models is to allow modeling of data that is not independent and identically distributed but instead has some underlying dependencies.
+For example, data generated during video extraction procedures have some associated temporal dependencies, since there exist similarities between features (such as those that encode the spatial positions or appearances of objects) of data at nearby time steps.
+
+To account for the dependent behavior of data, there has been research into models involving a sequence of DPMs, where components of the mixtures are dependent upon (or are sometimes said to be ``tied to'') corresponding components at neighboring positions in the sequence.
+For example, if the data shows temporal dependence, the goal might be to create a sequence of DPMs, one for each time-step, where the components of the mixture at each step are dependent upon corresponding components in the both the following and previous time steps.
+
+More rigorously, we take the definition of a DDPM to be a stochastic process defined on the space of probability distributions over a domain, which are indexed by time, space, or a selection of other covariates in such a way that the marginal distribution at any point in the domain follows a Dirichlet process (adapted from definitions found in \cite{gasthaus_thesis} and \cite{griffin2006order}).
+Hence, a time-dependent DDPM is a model which remains a Dirichlet process, marginally, at each time step, yet allows cluster parameters at a given time step to vary from (and remain dependent upon) the parameters in neighboring time steps.
+
+% \subsection{Generalized Polya Urn Dependent Dirichlet Process Mixture Model}
+% \label{sec:gpuddpmixture}
+
+% The time dependent video extraction data contains a collection of clusters, each representing a video object, which might change in number as time progresses (which is to say, clusters may be created, or be ``born'', and may disappear, or ``die'', at some intermediate time step), since objects can enter and exit a scene.
+To handle these challenges, the specific DDPM chosen to model extraction data in this work is known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM), introduced by \cite{caron_2007} in 2006.
+
+% The GPUDDPM, when applied to data extracted from a video with T frames, can be viewed as a sequence of DPMs (one for each ``time-step'' $t \in \{1, \ldots, T \}$), which are linked together by dependencies between cluster parameters in neighboring time steps.
+More specifically, the parameters at a given time step are distributed as a function of the parameters in the previous time step, and the distribution and number of distinct cluster assignments at a given time step are distributed according to all previous cluster assignments and a deletion procedure, described below.
+
+% A transition kernel $P(\theta_{k,t} | \theta_{k,t-1})$ specifies how mixture component parameters in time step $t$ are dependent upon associated mixture component parameters in time step $t-1$.
+One caveat is that each mixture component must be drawn independently from $\mathbb{G}_{0}$ (the base distribution of the DP, which acts as a prior distribution for the cluster parameters) which we achieve by making $\mathbb{G}_{0}$ the invariant distribution of the transition kernel $P(\theta_{k,t} | \theta_{k,t-1})$ (which, one should note, is a markov chain).
+
+% Recall that each observation $i$ is given a cluster assignments $c_{i}$ (Section~\ref{sec:dpmixture}).
+Cluster size $m_{k}$ denotes the number of observations assigned to a given cluster $k$; in a DPM, cluster size can be found by counting the number of these assigned observations.
+In contrast, to account for varying numbers of clusters at different points in time---and in particular, to allow clusters to diminish in size (i.e.
+to diminish in number of assigned observations) and even die off---the GPUDDPM has a deletion procedure that allows observations to be considered unassigned (or ``removed'') from their assigned clusters at a later time step.
+This procedure is introduced and described in detail in \cite{caron_2007}.
+
+% The deletion procedure operates in the following way: at each time step, all previous assignments (that have not yet been deleted) are independently considered for deletion.
+Specifically, each remaining assignment is removed from its cluster with probability $\rho$.
+It can be shown that performing deletion in this way is equivalent to, for each cluster $k$, drawing $r \sim \text{Binomial}(m_{k,t-1}, \rho)$, and reducing the previous-time cluster size $m_{k,t-1}$ by $r$ to find the next cluster size $m_{k,t}$.
+At each time step, this deletion procedure is carried out on all existing assignments, including those at the current time.
+Hence, the size $m_{k,t}$ of a given cluster $k$ at time $t$ is dependent upon the cluster size at the previous time step, $m_{k,t-1}$, and on the assignments $c_{i}$ for all observations $i \in \{ 1, \ldots, N_{t} \}$ at time $t$.
+We refer to the conditional distribution over $m_{k,t} | m_{k,t-1}, c_{i}, \rho$ as Del, which we define to be
+% \begin{equation}
+% \begin{split}
+% \label{del_step}
+% \text{Del} &(\cdot | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) := m_{k,t-1} \\
+% &- \text{Binomial}( \cdot | {m_{k,t-1}, \hspace{1mm} \rho}) + \sum_{i=1}^{N_{t}} \mathbb{I}(c_{i,t} = k)
+% \end{split}
+% \end{equation}
+% $\forall k \in \{1, \ldots, K_{t} \}$, where $\rho$ is a deletion parameter, $K_{t}$ is the number of clusters at time $t$, and $\mathbb{I}(c_{i,t} = k)$ is an indicator function whose value is 1 if $c_{i,t} = k$ and 0 otherwise.
+This process adheres to what \cite{caron_2007} calls a ``uniform deletion strategy'' over all the observations' assignments (since assignments to each cluster have equal probability of being deleted), though a more complex deletion strategy, dependent upon cluster size, can also be implemented and is described in \cite{caron_2007}.
+
+% We also define a distribution, which we refer to as C, that is useful during inference procedures (Section~\ref{sec:inference}) for inferring assigments of observations. We define C to be
+% \begin{equation}
+% \label{C}
+% \begin{split}
+% \text{C} &(\cdot | m_{1:K_{t},t}, \hspace{1mm} \alpha) := \text{Discrete}(\cdot | P(c_{i,t}=1 | m_{1:K_{t},t}), \\
+% &  \ldots, P(c_{i,t}=K_{t+1}  | m_{1:K_{t},t}))
+% \end{split}
+% \end{equation}
+% where
+% \begin{equation}
+% \begin{split}
+% P (c_{i,t}=k  | m_{1:K_{t},t})
+% = \begin{cases}
+% \frac{m_{k,t}}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{1mm} \text{if} \hspace{1mm} k \in \{ 1, \ldots, K_{t} \} \\
+% \frac{\alpha}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{1mm} \text{if} \hspace{1mm} k = K_{t} + 1
+% \end{cases}
+% \end{split}
+% \end{equation}
+% $\forall i \in \{1, \ldots, N_{t} \}$, where there exists $K_{t}$ clusters at time $t$, and we give a newly created cluster the index $K_{t+1}$.
+Distributions \eqref{del_step} and \eqref{C} together comprise what \cite{caron_2007} refers to as the ``Generalized Polya Urn''.
+Using the Del and C distributions given above, we can define the GPUDDPM generatively as, for each time step $t \in \{1, \ldots, T\}$, and each cluster $k \in \{ 1, \ldots, K_{t} \}$ at time $t$,
+% \begin{align}
+% \label{gpuddpm_def}
+% \begin{split}
+% m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho  &\sim \text{Del}(m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) \\
+% \theta_{k, t} | \theta_{k, t-1}   &\sim
+% \begin{cases}
+%   P(\theta_{k, t} | \theta_{k, t-1}) \hspace{1mm} \text{if} \hspace{1mm} k \in \{ 1, \ldots, K_{t} \} \\
+%   \mathbb{G}_{0}   \hspace{2mm} \text{if} \hspace{2mm} k = K_{t+1}
+% \end{cases} \\
+% c_{i,t} | m_{1:K_{t},t}, \hspace{1mm} \alpha  &\sim  \text{C} (m_{1:K_{t},t}, \hspace{1mm} \alpha) \\
+% \bold{x}_{i,t} | c_{i,t}, \theta_{1:K_{t}, t} &\sim \text{F}(\theta_{c_{i,t}, t})
+% \end{split}
+% \end{align}
+% where we specify distributions for F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ in Section~\ref{sec:modelspecification}.
+The graphical model corresponding with this formulation is shown in Figure \ref{fig:gpuddpm_gm_1}.
+% \begin{figure}[h]
+%         \center{\includegraphics[width=80mm]{../img/gpuddp_gm_1.pdf}}
+%         \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Polya Urn Dependent Dirichlet Process Mixture.
+Note that all observations at time $t$, $\bold{x}_{1:N,t}$, and their associated assignments $c_{1:N,t}$ are denoted respectively as $\bold{x}_{t}$ and $c_{t}$ (this also holds for time step $t-1$).}
+% \end{figure}
+
+% \subsection{Deletion Variable Formulation of the GPUDDPM}
+
+% To allow for modeling time-dependent data, the GPUDDPM allows clusters to ``lose'' assigned observations in a deletion step, which modifies the values of the cluster sizes at each time.
+Instead of incorporating the cluster size variable $m$ directly in the model (as it was in the definition of the GPUDDPM given by \eqref{gpuddpm_def}), we can formulate an equivalent model which makes use of new set of variables called the deletion variables (denoted as $d$), which represent the times at which assignments are deleted.
+More precisely, we introduce variables $d_{1}, \ldots, d_{N}$ to the model, which denote the times at which the assignments of observations $1, \ldots, N$ are considered removed from their assigned cluster.
+At each time step, cluster sizes can be recontructed from all previous assignment and deletion variables.
+With these deletion variables, we can compute the size of cluster $k$ at time $t$, $m_{k,t}$, by
+% \begin{equation}
+% \label{compute_clust_size}
+% m_{k,t} = \sum_{t' = 1}^{t} \mathbb{I}[(c_{t'}=k) \wedge (t < d_{t'})]
+% \end{equation}
+% where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise.
+Additionally, for an observation $\bold{x}_{i}$ at a given time-step $t$, the deletion time $d_{i}$ can be defined to be $d_{i} = t + l_{i}$, where $l_{i}$ is considered the lifespan of an assignment, is distributed geometrically, and can be expressed as
+% \begin{equation}
+% \label{del_rho_form}
+% l_{i} | \rho  \sim  \rho(1 - \rho)^{l_{i}}
+% \end{equation}
+% This alternate, deletion variable formulation of the GPUDDPM is useful for the MCMC inference algorithm described in Section~\ref{sec:MCMC}.
+We can write the new formulation of the GPUDDPM as, for each time step $t \in \{1, \ldots, T\}$ and clusters $k \in \{ 1, \ldots, K_{t} \} $ at time $t$,
+% \begin{align}
+% \begin{split}
+% d_{i,t} | \rho  &\sim \text{Geo}(\rho) + t + 1 \\
+% \theta_{k, t} | \theta_{k, t-1}  &\sim
+% \begin{cases}
+% P(\theta_{k,t} | \theta_{k,t-1}) \hspace{2mm} \text{if} \hspace{2mm} k \in \{ 1, \ldots, K_{t} \} \\
+% \mathbb{G}_{0} \hspace{2mm} \text{if} \hspace{2mm} k = K_{t} + 1
+% \end{cases} \\
+% c_{i,t} | \bold{c}_{1:t-1}, \bold{d}_{1:t-1}, \alpha  &\sim  \text{C}(\bold{c}_{1:t-1}, \bold{d}_{1:t-1}, \alpha) \\
+% \bold{x}_{i,t} | c_{i,t}, \theta_{c_{i,t},t}  &\sim  \text{F}(\theta_{c_{i,t}, t})
+% \end{split}
+% \end{align}
+% where $\bold{c}_{t} = c_{1,t}, \ldots, c_{N_{t}, t}$ (for $N_{t}$ observations at time $t$), $\bold{d}_{t} = d_{1,t}, \ldots, d_{N_{t}, t}$, and we specify distributions for F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ in Section~\ref{sec:modelspecification}.
+This formulation of the GPUDDPM is also used by \cite{gasthaus_thesis} and \cite{caron_2007}.
+The associated graphical model for this formulation is given in Figure~\ref{fig:gpuddpm_gm_2}.
+% \begin{figure}[h]
+%         \center{\includegraphics[width=82mm]{../img/gpuddp_gm_3.pdf}}
+%         \caption{Graphical model of the deletion variable and auxiliary variable transition kernel representation of the GPUDDPM}
+%         \label{fig:gpuddpm_gm_2}
+% \end{figure}
+
 \end{document}

### 2012-02-08 ltx/paper_short_2.tex
a few more edits. as before, main file is: paper_short_2.tex

@@
 Accordingly, we carry out a method called ``frame differencing'' for all experiments, which locates regions within video frames that contain motion by recording the positions of pixels exhibiting large enough differences in intensity or value between succesive frames.
-Frame differencing is unsophisticated, computationally inexpensive, able to be applied to a wide range of static, single-camera videos (note that videos used in all experiments were stationary; moving-camera videos require extraction methods useful for non-stationary video, such as \willie{cite cite cite}).
+Frame differencing is unsophisticated, computationally inexpensive, able to be applied to a wide range of static, single-camera videos (note that videos used in all experiments were stationary; moving-camera videos require extraction methods useful for non-stationary video, such as \cite{chien2002efficient, zhang2007moving}).
 Examples of pixel location data extracted via frame differencing are shown in Fig.~\ref{fig:img_and_framediff}(a-j).
@@
 
-We have not found an easily justifiable way to choose parameters.
-For the appearance and movement parameters, samples were drawn given a range of parameter values, and those that yielded samples similar in appearance to the experimental data were kept (a technique also performed in \cite{gasthaus_thesis}).
 The parameters chosen for use in experiments (Section~\ref{sec:experiments}) were accepted based on the fact that they yielded reasonable results.
@@
 Bayesian inference algorithms, which infer distributions over latent model parameters, are used to get detection and tracking results.
-A variety of methods previously developed for statistical inference can be applied to the generative model defined in Sections~\ref{sec:} and \ref{sec}.
-
- we separate presentations of model and inference so that multiple inference methods, each with its own benefits in terms of precision and speed, can be formulated as independent algorithms for detection and tracking.
-This section provides background on the two inference algorithms implemented in this study, which are both used to perform Bayesian inference on the model described in Section~\ref{sec:modelspec}.
+A variety of methods previously developed for statistical inference can be applied to the generative model defined in Sections~\ref{sec:gpuddpm} and \ref{sec:modelspecification}.
+We've separated descriptions of the model and inference algorithms so that multiple inference methods, each with its own benefits in terms of precision and speed, can be formulated as independent algorithms for detection and tracking.
+This section provides background on the two Bayesian inference algorithms implemented in this study.
 The first is a type of Markov Chain Monte Carlo (MCMC) batch inference, which uses Gibbs sampling to generate samples from the posterior distribution of the model.
@@
 and accept or reject this sample using the standard M-H acceptance ratio.
-
 %%%%%

### 2012-02-27 ltx/paper_short_2.tex
first final draft added (paper_short_2.tex). lots of new and adjusted images

@@
 This paper proposes a technique for the unsupervised detection and tracking of arbitrary objects in videos.
-It is intended to reduce the need for detection or localization methods tailored to specific object types and serve as a general framework applicable to videos with varied objects, backgrounds, and film qualities.
-The technique uses a Dependent Dirichlet Process Mixture (DDPM) known as the Generalized Polya Urn (GPUDDPM) to model image pixel data that can be extracted in a general manner from the regions in a video that represent objects.
+It is intended to reduce the need for detection or localization methods tailored to specific object types and serve as a general framework applicable to videos with varied objects, backgrounds, and image qualities.
+The technique uses a Dependent Dirichlet Process Mixture (DDPM) known as the Generalized Polya Urn (GPUDDPM) to model image pixel data that can be easily and efficiently extracted from the regions in a video that represent objects.
 This paper describes a specific implementation of the model using spatial and color pixel data extracted via frame differencing and gives two algorithms for performing inference on the model to accomplish detection and tracking.
@@
 
-We define the task of automated detection and tracking of arbitrary objects in videos to be three things: extraction, localization, and tracking.
-Extraction is the task of determining the spatiotemporal regions of a video that constitute objects, localization is the task of finding the positions and/or shapes of distinct objects, and tracking is the task of maintaining the identities of the detected objects over time.
-An ability to carry out these tasks in an automated manner is useful for many fields that make use of video data, including robotics, video surveillance, time-lapse microscopy, and video summarization.
-Algorithms able to be applied to a variety of objects and video types are particularly desirable.
-This paper introduces a new framework involving the use of dependent Dirichlet process mixture models.
+We define unsupervised detection and tracking of arbitrary objects in videos to be the task of automatically identifying the distinct objects present in a sequence of images and determining the path each object follows over time.
+Techniques that accomplish this task are useful in many fields that make use of video data, including robotics, video surveillance, time-lapse microscopy, and video summarization.
+By studying this task, we hope to help make progress towards general machine vision algorithms that can learn the positions, appearances, and number of objects present in any video scene.
+
+We break-down this task into three parts: data extraction, localization, and tracking.
+Data extraction is the act of extracting features from regions of video-frames that constitute objects, localization is the act of finding the positions and/or shapes of distinct objects, and tracking is the act of maintaining the identities of the detected objects over time.
+This paper introduces a new framework for carrying out these three actions based on a type of dependent Dirichlet process mixture model.
 This framework provides a foundation for a class of unsupervised algorithms that can detect and track arbitrary objects in a wide range of videos.
 
-Research on methods for general detection and tracking of objects tends to focus on one of either extraction, localization, or tracking.
+Research related to general detection and tracking of objects tends to focus on one of either extraction, localization, or tracking.
 Integrating all three tasks in a system for multiple arbitrary objects and diverse video types is not often a primary focus.
-Accomplishing the three tasks in a cohesive manner has been pursued more recently and has produced algorithms that can, in a fully automated way, accomplish unsupervised detection, localization, and tracking.
-An example of general detection and tracking is called ``blob tracking'', which typically refers to methods that perform extraction and localization of objects in each frame independently.
+A few attempts at accomplishing the three tasks in a cohesive manner have been studied in recent years \cite{brostow2006unsupervised, brox2010object, fragkiadaki2011detection, pece_2002}.
+%\willie{An example of general detection and tracking, called blob tracking, typically refers to methods that extract data and localize objects in each frame independently.
 Tracking involves finding blobs with similar locations and appearances over time.
 The main blob tracking challenges are the difficulty of segmenting neighboring or occluding objects in a given frame \cite{zhao2004tracking} and preventing multiple tracked blobs from merging into a single object \cite{vermaak_2003}.
-These difficulties stem from performing detection via segmentation on individual frames.
-
-This paper presents the use of a Bayesian nonparametric mixture model for general, unsupervised multiple object detection and tracking.
-This method provides both a widely applicable way to detect arbitrary objects in videos---particularly in cases where frame-by-frame segmentation is difficult or video quality is low and extraction is noisy---and a general framework for identifying and isolating objects during tracking.
-Specifically, this work describes how a dependent Dirichlet process mixture model can accomplish these goals.
-The DDP is first reviewed in the abstract.
-This model can be used with arbitrary extraction procedures and object appearance representations that may be desired in future studies.
-We then present our specific model for data obtained via a basic extraction procedure.
-We describe inference algorithms for the DDP, and specifically our specific model, that allow object localization and tracking to be carried out.
-We demonstrate our implementation on multiple synthetic and benchmark datasets.
-Standard performance metrics for each supports our hypothesis that general extraction combined withe sophisticated inference can perform detection and tracking at a level comperable to state-of-the-art, object specific algorithms.
+These difficulties stem from performing detection via segmentation on individual frames.}
+This paper furthers this line of work by providing a model that gives rise to a number of algorithms to detect arbitrary objects in videos---particularly in cases where frame-by-frame segmentation is difficult, video quality is low, and extraction is noisy---and maintain the isolation of distinct objects during tracking and through occlusion.
+
+We begin by describing characteristics of the extracted data (Section~\ref{sec:dataextraction}), and giving the generalized form of the model (Section~\ref{sec:modeldefinition}).
+To implement this model, one must specify a data extraction procedure and distributions for representing objects, which may be chosen to allow for general object tracking or tailored to a specific object type for a given application.
+In our implementation, we give the specific distributions for representing objects in a general manner (Section~\ref{sec:modelspecification}), which we apply to data extracted via a basic frame differencing procedure.
+We describe inference algorithms for our model, and show how output of these algorithms can be interpreted as object localization and tracking results (Section~\ref{sec:inference}).
+Our implementation is demonstrated on multiple synthetic and benchmark datasets, and standard performance metric values are computed to quantify results on the benchmark datasets (Section~\ref{sec:experiments}).
+We compare our performance metrics with those yielded by specialized detection and tracking algorithms tailored to specific objects.
+Our results support our hypothesis that simple data extraction combined with inference on our model can perform detection and tracking of arbitrary objects at a level comparable to state-of-the-art, object specific algorithms.
 
@@
 
-Extraction methods aim to discern the foreground regions of a video from the background.
-Frame differencing and background subtraction are two such methods, and involve comparing pixel values, or groups of pixel values, between two images, to detect a change surpassing some threshold.
-The goal of these techniques is to find locations in a scene that display motion, and to deem these moving areas the foreground of a scene.
-Often, background subtraction refers to comparing an image containing targets with an image of the background without targets or with some model of the background that is learned as the video progresses (in these cases, objects that do not move throughout a scene will be treated as the background), while frame or image differencing refers to comparing pairs of consecutive images in a video.
+Methods that discern between the foreground and background regions of a video allow for data to be extracted from the areas in each frame where objects exist.
+Frame differencing and background subtraction are two such methods.
+Both record locations that exhibit motion relative to the background.
+Often, background subtraction refers to methods that compare an image containing targets with an image of the background only or with some model of the background that is learned as the video progresses \cite{piccardi2004background}, while frame differencing refers to methods that compare pairs of consecutive images in a video \cite{zhang2001segmentation}.
 Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking schemes \citep{perez_2002}.
-Another technique for discerning moving regions in an image involves optical flow \citep{horn_1981, bobick_2001}.
-In \cite{black_2000}, a motion detection and object tracking method is developed based on the relative motion of moving objects' boundaries, which are found through analysis of optical flow.
 
@@
 Over the past decade, there have been attempts to provide general algorithms for the fully unsupervised detection and tracking of arbitrary objects in videos.
-Blob tracking, a basic method attempting to carry out this goal, has been used for many years, and has found success in arenas where objects are easily discerned from the background of a video and where localization and segmentation of distinct objects is possible \cite{francois2004real, isard_2001}.
+Blob tracking, a basic method to carry out this goal, has been used for many years, and has found success in videos where objects are easily isolated from the background and where localization and segmentation of distinct objects is possible \cite{francois2004real, isard_2001}.
 Methods involving blob tracking, however, run into problems when faced with videos where detection is difficult, object appearance or orientation varies heavily, and there exists object occlusion \cite{song2005model}.
 To aid the accuracy of these methods, techniques have been developed for performing extraction and segmentation in a joint manner, incorporating statistical methods for maintaining hypotheses of different numbers of detected objects, and introducing some of the multi-object tracking methods described previously to track distinct blobs after they have been segmented \cite{collins2003mean, isard_2001}.
-Another family of methods related to the task of unsupervised detection and tracking of video objects goes under the heading of video segmentation algorithms---these methods extend single-frame image segmentation to maintain coherence of image segments over time, and have had some success when used for the explicit purpose of detecting and tracking foreground objects of videos \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}.
-Other attempts to perform unsupervised detection and tracking include methods for clustering short sequences of positions that can be gained by detecting the motion of objects \cite{brostow2006unsupervised, brox2010object}, which aim to return full-length distinct object tracks, and a graph based method that carries out a similar task using spectral clustering \cite{fragkiadaki2011detection}.
-Another approach based on a Gaussian mixture model has also been used to cluster data gained from moving objects \cite{pece_2002}; this method also develops heuristics for the initialization and elimination of new tracks.
+Another family of methods related to the task of unsupervised detection and tracking of video objects goes under the heading of video segmentation algorithms---these methods extend single-frame image segmentation to maintain coherence of image segments over time, and have had some success when used for the explicit purpose of detecting and tracking foreground objects in videos \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}.
+Other attempts to perform unsupervised detection and tracking include methods for clustering short sequences of positions extracted by detecting the motion of objects \cite{brostow2006unsupervised, brox2010object}, which aim to return full-length distinct object tracks, and graph based method that carry out a similar task using spectral clustering \cite{fragkiadaki2011detection}.
+Another approach based on a Gaussian mixture model has also been used to cluster data extracted from moving objects \cite{pece_2002}; this method also develops heuristics for the initialization and elimination of new object tracks.
 
 Nearly all methods capable of high accuracy detection and tracking of objects in videos are those tailored for specific object types.
-These methods rely on ad-hoc detection criteria that exploit previous knowledge about the appearance or behavior of objects in a video.
+These methods rely on ad-hoc detection criteria that exploit previous knowledge about the appearance or behavior of the given objects in a video.
 The method described in this paper is designed to track arbitrary objects without using any explicit detection criteria, and serve as a general strategy that can be used, without modification, to perform accurate detection and tracking of diverse objects in a wide range of videos.
 The method we introduce falls into the category of clustering-based arbitrary object detection and tracking methods.
-Differing from previous work, we use a type of time dependent Bayesian nonparametric mixture model, and show how it can be applied to a variety of extraction data to perform detection and tracking.
-This method begins by performing a simple data extraction procedure (that is consistent for all videos), which can be carried out broadly but yields noisy data; we model this noisy extracted data with a sophisticated type of Bayesian mixture model.
-Inference can then be performed on this model to gain detection and tracking results.
-
-This paper describes the data extraction procedures (Section~\ref{sec:dataextraction}), the general form of the model (Section~\ref{sec:modeldefinition}), the specific distributions used in our implementation of the model (Section~\ref{sec:modelspecification}), inference procedures used to get tracking results (Section~\ref{sec:inference}), and experimental results (Section~\ref{sec:experiments}).
-
-% Overview
-% --------
-
-% \section{Overview of Method}
-% \label{sec:overviewofmethod}
+Differing from previous work, we use a type of time dependent Bayesian nonparametric mixture model, and show how it can be applied to a variety of easily extracted data to perform detection and tracking.
+This method begins by performing a simple data extraction procedure, which can be carried out broadly but yields noisy data; our model is then applied to this noisy extracted data.
+Inference is performed on this model to achieve detection and tracking results.
 
@@
   \centering
-  \subfloat[]{\label{fig:ants_img}\includegraphics[width=0.335\textwidth]{../img/usbh_frame_723.pdf}} 
-  \subfloat[]{\label{fig:ants_img2}\includegraphics[width=0.335\textwidth]{../img/usbh_frame_724.pdf}}
+  \subfloat[]{\label{fig:ants_img}\includegraphics[width=0.33\textwidth]{../img/usbh_frame_723.pdf}} \hspace{0.1mm}
+  \subfloat[]{\label{fig:ants_img2}\includegraphics[width=0.33\textwidth]{../img/usbh_frame_724.pdf}} \hspace{0.1mm}
   \subfloat[]{\label{fig:ants_img_framediff}\includegraphics[width=0.323\textwidth]{../img/usbh_framediff.png}}\\
-  \subfloat[]{\label{fig:traffic2_img}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame1.png}} 
-  \subfloat[]{\label{fig:traffic2_img2}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame2.png}}
+  \subfloat[]{\label{fig:traffic2_img}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame1.png}} \hspace{0.1mm}
+  \subfloat[]{\label{fig:traffic2_img2}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame2.png}} \hspace{0.1mm}
   \subfloat[]{\label{fig:traffic2_img_framediff}\includegraphics[width=0.3273\textwidth]{../img/traffic2_framediff.png}}\\
-  \subfloat[]{\label{fig:pets2009_img}\includegraphics[width=0.333\textwidth]{../img/pets2009_frame1.png}} 
-  \subfloat[]{\label{fig:pets2009_img2}\includegraphics[width=0.333\textwidth]{../img/pets2009_frame2.png}}
-  \subfloat[]{\label{fig:pets2009_img_framediff}\includegraphics[width=0.333\textwidth]{../img/pets2009_framediff.png}}\\
+  \subfloat[]{\label{fig:pets2009_img}\includegraphics[width=0.328\textwidth]{../img/pets2009_frame1.png}} \hspace{0.1mm}
+  \subfloat[]{\label{fig:pets2009_img2}\includegraphics[width=0.328\textwidth]{../img/pets2009_frame2.png}} \hspace{0.1mm}
+  \subfloat[]{\label{fig:pets2009_img_framediff}\includegraphics[width=0.328\textwidth]{../img/pets2009_framediff.png}}\\
   %\subfloat[]{\label{fig:longimg}\includegraphics[width=1.05\textwidth]{../img/longimg_6.pdf}} \\
-  \subfloat[]{\label{fig:longimg_all}\includegraphics[width=1\textwidth]{../img/longimg_7_detail_4.pdf}} 
+  \subfloat[]{\label{fig:longimg_all}\includegraphics[width=1\textwidth]{../img/longimg_7_detail_7.pdf}} 
   %../img/longimg_7_detail.pdf}} 
-  \caption{Pairs of consecutive frames and the results produced by taking the pixel-wise difference between these frames (a - f).
-The final image shows the result of frame differencing over a sequence of images (from the PETS2009 dataset), where color denotes frame number.}
+  \caption{Three pairs of consecutive frames and the results produced by taking the pixel-wise frame difference between each pair (a - i).
+The final image shows the results of frame differencing over a sequence of images (from the PETS2009/2010 dataset.}
   \label{fig:img_and_framediff}
@@
 
-A basic set of observations that can be extracted from video are three-dimensional points $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $x_{1}$ and $x_{2}$ are the two spatial locations of a pixel.
-Extraction procedures, such as those discussed in Section~\ref{sec:introduction}, are used to isolate pixels lying in the regions of a video that represent objects.
-For implementation in experiments (Section~\ref{sec:experiments}), we desire an extraction procedure that is as unsophisticated as possible, both to gauge the robustness of this method on potentially noisy extraction data and to ensure that the procedure is applicable to a wide range of videos.
-Accordingly, we carry out a method called ``frame differencing'' for all experiments, which locates regions within video frames that contain motion by recording the positions of pixels exhibiting large enough differences in intensity or value between succesive frames.
-Frame differencing is unsophisticated, computationally inexpensive, able to be applied to a wide range of static, single-camera videos (note that videos used in all experiments were stationary; moving-camera videos require extraction methods useful for non-stationary video, such as \cite{chien2002efficient, zhang2007moving}).
-Examples of pixel location data extracted via frame differencing are shown in Fig.~\ref{fig:img_and_framediff}(a-j).
-
-Incorporating color information into our model allows us to infer object color, better maintain the identity of adjacent objects, and track objects through occlusion with greater accuracy.
-We therefore capture the color of each extracted pixel.
-Specifically, for each pixel $\bold{x} = ( x_{1}, x_{2}, t )$ recorded during frame differencing, we specify a square, $L$ pixels in length, centered on $(x_{1}, x_{2})$, that contains a set of pixels surrounding $\bold{x}$ in frame $t$.
-We choose a scalar color value that is able to be computed for each pixel; this value could, for example, be some function of the red-green-blue (r-g-b) or hue-saturation-value (h-s-v) characteristic of a pixel.
-The color value of each of the selected pixels surrounding $\bold{x}$ is recorded.
-Afterwards, the set of possible color values (ie the range of color values to which a pixel may be assigned) is partitioned into $V$ bins, and the number of pixels with a color value lying in each of the bins yields a $V$ dimensional vector of ``color counts''.
-This vector represents a discrete distribution over colors in the vicinity of a pixel.
-We will refer to this extraction technique as ``color counting''.
-
+The basic set of observations extracted from video are the three-dimensional points
+\begin{equation}
+  \centering
+  \bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \{ 1, \ldots, T \}
+\end{equation}
+where $x_{1}$ and $x_{2}$ are the two spatial locations of a pixel and $t$ is the time index.
+These data are extracted by recording the positions of pixels lying in regions that contain objects.
+For our implementation in experiments (Section~\ref{sec:experiments}), we desire a data extraction procedure that is as unsophisticated as possible, to verify that we can perform accurate inference on noisy data extracted using a broadly applicable procedure.
+Consequentially, we carry out simple frame differencing (introduced in Section~\ref{sec:introduction}) for all experiments, to find pixels at locations where there is motion.
+This data extraction procedure is simple, computationally inexpensive, and able to be applied to a wide range of static, single-camera videos (note that videos used in all experiments were stationary; moving-camera videos require data extraction methods useful for non-stationary video, such as \cite{chien2002efficient, zhang2007moving}).
+Examples of pixel location data extracted via frame differencing are shown in Figure~\ref{fig:img_and_framediff}(a-j).
+
+We can extract additional features from videos and incorporate this new information into our model.
+For example, we can extract color information, and modify our model to allow for object color to be inferred during tracking.
+Adding this feature allows our method to better distingush between adjacent objects and more accurately track objects through occlusion.
+For each extracted pixel, we compute a $V$ dimensional vector that represents the color distribution in the pixel's immediate vicinity (details on this vector and how it is computed are given in Section~\ref{sec:dataextractioninexperiments}).
+We call the components of this vector the ``color counts'' of the pixel.
 The frame differencing and color counting extraction yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \{1, \ldots, T \}$, where each element $\bold{x} \in \bold{X}$ can be written
@@
 \end{equation}
-where $\bold{x}^{s} \in \mathbb{R}^{2}$ is the two dimensional vector of spatial coordinates,  $\bold{x}^{c} \in \mathbb{Z}_{+}^{V}$ is the $V$ discrete dimensional vector of color counts, and $t \in \{1, \ldots, T \}$ is the discrete time index.
-Additional features may be extracted from each pixel to attempt more sophisticated object representations, including pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, and texture representations.
-We leave these attempts for future studies.
+where $\bold{x}^{s} \in \mathbb{R}^{2}$ is the two dimensional vector of spatial coordinates,  $\bold{x}^{c} \in \mathbb{Z}_{+}^{V}$ is the $V$ dimensional vector of color counts, and $t \in \{1, \ldots, T \}$ is the time index.
+Other features may also be extracted from each pixel to attempt more sophisticated object representations, including pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, and texture representations.
 
@@
 
-% Model Overview
+% Model Definition
 % --------------
@@
 
-A type of Dependent Dirichlet Process Mixture (DDPM) model known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM) is used to provide a model-based, time dependent clustering of the extracted data.
-The model definition is provided in Section~\ref{sec:gpuddpm}, a secondary form of the model used in one of the two inference algorithms is given in Section~\ref{sec:MCMC}, and background information introducing mixture models and the Dirichlet process are given in Appendix~\ref{sec:modelbackground}.
+A type of Dependent Dirichlet Process Mixture (DDPM) model known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM) is applied to the extracted data.
+We define the generalized form of this model in Section~\ref{sec:gpuddpm}, and specify the distributions used in our implementation of this model in Section~\ref{sec:modelspecification}.
+We also define a secondary form of this model in Section~\ref{sec:MCMC}, which is used in one of the two inference algorithms.
+Background information providing a brief introduction to mixture models and the Dirichlet process is contained in Appendix~\ref{sec:modelbackground}.
 
@@
 
-The extracted data contain a collection of spatiotemporal clusters, each representing an object in the video.
-The number of clusters present in each time step may not be constant (which is to say, clusters may be created, or be ``born'', and may disappear, or ``die'', at intermediate time steps), since objects can enter and exit a scene.
-To cluster data with these properties, we choose a DDPM known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM), introduced by \cite{caron_2007} in 2006.
-This model may be intuitively viewed as a sequence of DPMs, where the parameters and number of clusters are allowed to vary in a specified way between adjacent time steps.
-Like the DPM, the GPUDDPM allows for the number of clusters within a dataset---which, in this application, corresponds to the number of objects---to be found in an automated manner.
-
-As in typical mixture models, each observation $\bold{x}_{i,t}$ is associated with an assignment variable $c_{i,t}$ that represents its assignment to a cluster $\theta_{k,t}$.
-Usually, the size of each cluster $m_{k,t}$ is a function of the number of observations that have been assigned to it, but in this model, cluster sizes may decrease over time as observations become ``unassigned''; this allows clusters to decrease in strength and potentially die.
-We define a distribution over the size of cluster $k$ at time $t$ ($m_{k,t}$) conditioned on the previous cluster size ($m_{k,t-1}$), assignments at time $t$ ($c_{1:N_{t}, t}$), and deletion parameter $\rho$ to be
+The extracted data comprise a collection of spatiotemporal clusters, each corresponding with an object in the video.
+Since objects can enter and exit a scene, the number of clusters present in each time step may not be constant (i.e., clusters may be created, or be ``born'', and may disappear, or ``die'', at intermediate time steps).
+To cluster data with these properties, we choose to use a DDPM known as the Generalized Polya Urn Dependent Dirichlet Process Mixture, introduced by \cite{caron_2007}.
+This model may be intuitively viewed as a sequence of DPMs, where there exist dependencies between the parameters and number of clusters in adjacent time steps.
+Like the DPM, the GPUDDPM allows for the number of clusters within a dataset---which, in this application, corresponds to the number of objects in a video---to be found in an automated manner.
+
+Similar to a typical mixture model, in this model, each observation $\bold{x}_{i,t}$ is associated with an assignment variable $c_{i,t}$ that represents its assignment to a cluster $\theta_{k,t}$.
+In a DPM, the probability of assignment to a cluster $k$ is a function of the cluster size $m_{k,t}$, which is equal to the number of observations that have been assigned to it.
+In contrast, cluster sizes in the GPUDDPM may decrease over time as observations become ``unassigned''; this allows clusters to decrease in strength and potentially die.
+We define a distribution over the size of cluster $k$ at time $t$ ($m_{k,t}$), conditioned on the cluster's previous size ($m_{k,t-1}$), the assignments at time $t$ ($c_{1:N_{t}, t}$), and a deletion parameter ($\rho$) to be
 \begin{equation}
@@
 $\forall k \in \{1, \ldots, K_{t} \}$, where $K_{t}$ is the number of clusters at time $t$ and $\mathbb{I}(c_{i,t} = k)$ is an indicator function whose value is 1 if $c_{i,t} = k$ and 0 otherwise.
-We also define a distribution over the assignment of an observation $i$ at time $t$ to a cluster $k$ ($c_{i,t} = k$) conditioned on the sizes of all clusters at time $t$ ($m_{1:K_{t},t}$) and a concentration parameter $\alpha$, to be
+
+We also define a distribution over the assignment of observation $i$ at time $t$ ($c_{i,t}$), conditioned on the sizes of all clusters at time $t$ ($m_{1:K_{t},t}$) and a concentration parameter ($\alpha$), to be
 \begin{equation}
@@
 \end{equation}
-% \begin{equation}
-% \label{C}
-% \begin{split}
-% \text{C} &(\cdot | m_{1:K_{t},t}, \hspace{1mm} \alpha) := \text{Discrete}(\cdot | P(c_{i,t}=1 | m_{1:K_{t},t}), \\
-% &  \ldots, P(c_{i,t}=K_{t+1}  | m_{1:K_{t},t}))
-% \end{split}
-% \end{equation}
-% where
-% \begin{equation}
-% \begin{split}
-% P (c_{i,t}=k  | m_{1:K_{t},t})
-% = \begin{cases}
-% \frac{m_{k,t}}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{1mm} \text{if} \hspace{1mm} k \in \{ 1, \ldots, K_{t} \} \\
-% \frac{\alpha}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{1mm} \text{if} \hspace{1mm} k = K_{t} + 1
-% \end{cases}
-% \end{split}
-% \end{equation}
 $\forall i \in \{1, \ldots, N_{t} \}$, where there exists $K_{t}$ clusters at time $t$ and we give a newly created cluster the index $K_{t+1}$.
-Distributions \eqref{D_distro} and \eqref{C_distro} together comprise what \cite{caron_2007} refers to as the ``Generalized Polya Urn''.
-We can now define the GPUDDPM generatively as, for each time step $t \in \{1, \ldots, T\}$, and each cluster $k \in \{ 1, \ldots, K_{t} \}$ at time $t$,
+
+Distributions \eqref{D_distro} and \eqref{C_distro} together comprise what \cite{caron_2007} refers to as the ``Generalized Polya Urn''. We can now define the GPUDDPM generatively to be
 \begin{align}
@@
 m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho  &\sim \text{D}(m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) \\
-\theta_{k, t} | \theta_{k, t-1}   &\sim
+\theta_{k, t} | \theta_{k, t-1}   & \sim
 \begin{cases}
-  P(\theta_{k, t} | \theta_{k, t-1}) \hspace{1mm} \text{if} \hspace{1mm} k \in \{ 1, \ldots, K_{t} \} \\
-  \mathbb{G}_{0}   \hspace{2mm} \text{if} \hspace{2mm} k = K_{t+1}
+  P(\theta_{k, t} | \theta_{k, t-1}) \hspace{1mm} \text{if} \hspace{1mm} k \leq K_{t} \\ % \in \{ 1, \ldots, K_{t} \} \\
+  \mathbb{G}_{0} \hspace{1mm} \text{if} \hspace{1mm} k = K_{t+1}
 \end{cases} \\
@@
 \end{align}
-where we specify distributions for F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ in Section~\ref{sec:modelspecification}.
-The graphical model associated with this formulation is shown in Fig.~\ref{fig:gpuddpm_gm_1}.
+$\forall$ times $t \in \{1, \ldots, T\}$ and each cluster $k \in \{ 1, \ldots, K_{t} \}$ at time $t$, where we specify distributions for F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ in Section~\ref{sec:modelspecification}.
+The graphical model associated with this formulation is shown in Figure~\ref{fig:gpuddpm_gm_1}.
 \begin{figure}[h]
@@
         \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Polya Urn Dependent Dirichlet Process Mixture.
-Note that all observations at time $t$, $\bold{x}_{1:N,t}$, and their associated assignments $c_{1:N,t}$ are denoted respectively as $\bold{x}_{t}$ and $c_{t}$ (this also holds for time step $t-1$).}
+The observations at time $t$, $\bold{x}_{1:N_{t},t}$, and their associated assignments $c_{1:N_{t},t}$ are denoted respectively as $\bold{x}_{t}$ and $c_{t}$ (likewise for those at time $t-1$).}
 \end{figure}
@@
 
-Sections~\ref{sec:objectappearance}-\ref{sec:motionmodel} describe the distributions chosen to fully specify the general GPUDDPM model form given in Section~\ref{sec:gpuddpm}.
-The implementation given below was used for all the experiments of Section~\ref{sec:experiments}.
-The distributions F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ represent object appearance, appearance prior, and object movement; different specifications allow one to choose a model biased toward a specific application or kept general for wide applicability.
-
-\subsection{Object Appearance (Mixture Component)}
+Sections~\ref{sec:objectappearance}-\ref{sec:motionmodel} detail the object representation distributions chosen to fully specify our implementation of the GPUDDPM.
+This specification is used for all experiments in Section~\ref{sec:experiments}.
+The distributions F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ represent object appearance, the appearance prior, and object movement.
+Our specification is kept general to allow for wide applicability, though it could also be chosen to incorporate known object information for a specific tracking application.
+
+\subsection{Object Appearance and Mixture Component F}
 \label{sec:objectappearance}
@@
 \end{equation}
-where $\theta = \{ \boldsymbol{\mu}, \Sigma, \bold{p} \}$ denote the parameters of a cluster at time $t$, with mean $\boldsymbol{\mu} \in \mathbb{R}^{2}$, covariance matrix $\Sigma \in \mathbb{R}^{2} \times \mathbb{R}^{2}$, and discrete probability vector $\bold{p} = (p_{1}, \ldots, p_{V})$ such that $\sum_{i=1}^{V}p_{i} = 1$.
-Additionally, $\mathcal{N}$ denotes the multivariate normal distribution, $\mathcal{M}n$ denotes the multinomial distribution, and F is sometimes referred to as the likelihood.
-
-The distribution families assumed to generate each $\bold{x}$ must be justified.
+where $\theta = \{ \boldsymbol{\mu}, \Sigma, \bold{p} \}$ denotes the parameters of a cluster at time $t$, with mean $\boldsymbol{\mu} \in \mathbb{R}^{2}$, covariance matrix $\Sigma \in \mathbb{R}^{2} \times \mathbb{R}^{2}$, and discrete probability vector $\bold{p} = (p_{1}, \ldots, p_{V})$ such that $\sum_{i=1}^{V}p_{i} = 1$.
+Additionally, $\mathcal{N}$ denotes the multivariate normal distribution and $\mathcal{M}n$ denotes the multinomial distribution.
+The mixture component F is sometimes referred to as the likelihood.
+
+%The distribution families assumed to generate each $\bold{x}$ must be justified.
 %Both the multivariate normal and multinomial distributions were chosen because they are sufficiently simple (and well studied) to allow for tractable inference and sufficiently flexible to provide a reasonable approximation to the data gained during extraction.
-The multivariate normal distribution over the spatial features $\bold{x}^{s}$ represents the shape of each object as an oval; likewise, data generated by each moving object during extraction are often ovular.
+The multivariate normal distribution over the spatial features $\bold{x}^{s}$ can be intuitively thought to represent the shape of each object as an oval; likewise, data extracted from each moving object tend to cluster around the object's centroid in an oval shape.
 %---as noisy extraction procedures cause some smoothing of edges and corners, producing blob-like shapes even when objects are not particularly round---and centered on a given object.
 % Furthermore, this model is justified as the maximum likelihood parameter estimate of a normal distribution corresponds to the least squares fit of data relative to the normal distribution mean;
-Since extraction data produced by an object clusters around the centroid of the object, the estimated normal distribution mean should provide a reasonable reflection of the object's centroid.
-Modeling the color features $\bold{x}^{c}$ as draws from a multinomial distribution (equivalently, as draws from a product of discrete distributions), is justified since we have observed that distinct object tend to generate pixels whose hue values are noisy but yield consistent counts in discrete hue bins.
-
-\subsection{Appearance Prior (Base Distribution)}
+Modeling the color features $\bold{x}^{c}$ as draws from a multinomial distribution is justified since we have observed that the pixels associated with a distinct object tend to have similar color count vectors.
+
+\subsection{Appearance Prior and Base Distribution $\mathbb{G}_{0}$}
 \label{sec:appearanceprior}
 
-$\mathbb{G}_{0}$ denotes the base distribution of the time-dependent Dirichlet process mixture; it also serves as a prior distribution for the parameters $\theta = \{ \boldsymbol{\mu}, \Sigma, \bold{p} \}$ present in the mixture components (i.e.
-the parameters we've used to model object appearance).
-We make use of conjugate priors in the base distribution to allow for more efficient computation.
-Specifically, in the experiments carried out in this paper, a normal-inverse-Wishart prior is placed on the multivariate normal parameters $\{ \boldsymbol{\mu}, \Sigma \}$, and a Dirichlet prior is placed on the multinomial parameter $\bold{p}$ (where the normal-inverse-Wishart is a conjugate prior for the multivariate normal component of the likelihood and the Dirichlet is a conjugate prior for the multinomial component).
+$\mathbb{G}_{0}$ denotes the base distribution of the DDPM; it also serves as a prior distribution for the parameters $\theta = \{ \boldsymbol{\mu}, \Sigma, \bold{p} \}$ of the mixture components (i.e.
+of the object appearance distributions).
+We use conjugate priors in the base distribution to allow for more efficient computation.
+Specifically, in our implementation, a normal-inverse-Wishart prior is placed on the multivariate normal parameters $\{ \boldsymbol{\mu}, \Sigma \}$, and a Dirichlet prior is placed on the multinomial parameter $\bold{p}$ (where the normal-inverse-Wishart is a conjugate prior for the multivariate normal component of the likelihood and the Dirichlet is a conjugate prior for the multinomial component).
 The prior can therefore be written
@@
 
-\subsection{Motion Model (Transition Kernel)}
+\subsection{Motion Model and Transition Kernel $P(\theta_{t} | \theta_{t-1})$}
 \label{sec:motionmodel}
 
-We wish to formulate a transition kernel $P(\theta_{t} | \theta_{t-1})$ that provides a reasonable representation of how we expect tracked objects to move over time.
-Since we want our model to be used for tracking arbitrary objects, we do not wish to make any sophisticated assumptions about the behavior of objects.
-For example, we choose not to incorporate complex objects dynamics, though they are often used with success in certain object-specific tracking tasks, such as tracking walking people.
+The transition kernel $P(\theta_{t} | \theta_{t-1})$ represents how we expect tracked objects to move over time.
+Since our implementation is intended for tracking arbitrary objects, we do not wish to make sophisticated assumptions about object motion.
+For example, we choose not to incorporate complex objects dynamics, though they are often used with success in certain object-specific tracking tasks, such as people tracking.
 We assume only that the position of an object at a given time is a function of its position at the previous time, and that the position varies in all directions equally between time steps.
-A transition kernel that incorporates more sophisticated behavioral tendencies is left for future studies (especially for those intending to use this model for a more specific object type or application).
 
@@
 One way to achieve this is through the use of auxiliary variables.
-Auxiliary variables are a set of $M$ variables $\bold{z}_{t} = (z_{t,1}, \ldots, z_{t,M})$ associated with a each cluster at time $t$ that satisfy
+Auxiliary variables are a set of $M$ variables $\bold{z}_{t} = (z_{t,1}, \ldots, z_{t,M})$ associated with each cluster at each time $t$ that satisfy
 \begin{eqnarray}
-% P(\theta_{k,t} | \theta_{k,t-1}) = \int P(\theta_{k,t} | \bold{z}_{k,t}) P(\bold{z}_{k,t} | \theta_{k,t-1}) d \bold{z}_{k,t}
-P(\theta_{t} | \theta_{t-1}) = \int P(\theta_{t} | \bold{z}_{t}) P(\bold{z}_{t} | \theta_{t-1}) d \bold{z}_{t}  %\\
-% P(\theta', \bold{z}_{t}) = p(\bold{z}_{t}|\theta') \mathbb{G}_{0}(\theta')
+\label{auxiliaryvariablecriteria}
+P(\theta_{t} | \theta_{t-1}) = \int P(\theta_{t} | \bold{z}_{t}) P(\bold{z}_{t} | \theta_{t-1}) d \bold{z}_{t}
 \end{eqnarray}
 
-In this way, the parameters of a cluster at a given time do not depend directly on their value at the previous time; they are instead dependent upon an intermediate sequence of auxiliary variables chosen to satisfy the above criteria, which allows the cluster parameters at each time step to be marginally distributed according to the base distribution $\mathbb{G}_{0}$.
-
-For each cluster, we introduce $M$ auxiliary variables $z_{t, 1}, \ldots, z_{t, M}$ at time $t$ that are each drawn from the product of a multivariate normal and multinomial when conditioned on the associated cluster parameters $\theta_{t} = \{ \boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t} \}$:
+With the addition of these variables, the parameters of a cluster at a given time do not depend directly on their value at the previous time; they are instead dependent upon an intermediate sequence of variables.
+This allows the cluster parameters at each time step to be marginally distributed according to the base distribution $\mathbb{G}_{0}$ while maintaining simple time varying behavior.
+
+For each cluster, we introduce $M$ auxiliary variables $z_{t, 1}, \ldots, z_{t, M}$ at time $t$.
+Each auxiliary variable is drawn from the product of a multivariate normal and multinomial with the associated cluster parameters $\theta_{t} = \{ \boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t} \}$
 \begin{equation}
@@
 z_{t, m} | \boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t}  \sim  \mathcal{N}(\boldsymbol{\mu}_{t}, \Sigma_{t}) \mathcal{M}n(\bold{p}_{t})   \hspace{15pt}   
-\forall m \in \{ 1, \ldots, M \}
-\end{equation}
-To satisfy the above criteria for auxiliary variables, at each time $t$ we specify the dependencies of a given cluster on its associated set of auxiliary variables by
+\end{equation}
+$\forall m \in \{ 1, \ldots, M \}$. To satisfy \eqref{auxiliaryvariablecriteria}, we specify the dependencies of a given cluster on its associated set of auxiliary variables at each time $t$ by
 \begin{equation}
@@
 where $M$ is the number of auxiliary variables, \\
-$\{ \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0} \}$ are the $\mathcal{N}i\mathcal{W}$ prior parameters, $\bold{q}_{0}$ is the $\mathcal{D}ir$ prior parameter, $\bold{z}^{s}$ and $\bold{z}^{c}$ respectively denote the spatial and color features of an auxiliary variable $\bold{z}$, and $\overline{\bold{z}}$ and $S_{\bold{z}}$ respectively denote the sample mean and sample covariance for a set $\bold{z} = \{ z_{1}, \ldots, z_{M} \}$ (of auxiliary variables, in this case), which we can define as
+$\{ \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0} \}$ are the $\mathcal{N}i\mathcal{W}$ prior parameters, $\bold{q}_{0}$ is the $\mathcal{D}ir$ prior parameter, $\bold{z}^{s}$ and $\bold{z}^{c}$ respectively denote the spatial and color features of an auxiliary variable $\bold{z}$, and $\overline{\bold{z}}$ and $S_{\bold{z}}$ respectively denote the sample mean and sample covariance for a set $\bold{z} = \{ z_{1}, \ldots, z_{M} \}$ (of auxiliary variables, in this case), which we can write as
 \begin{eqnarray}
@@
 
-The multivariate normal-multinomial GPUDDPM implemented in this study has a number of parameters, which can be tuned to increase the efficacy of object detection and tracking. \\ \\
-Object appearance parameters include:
+The multivariate normal-multinomial GPUDDPM implemented in this study has a number of parameters, which are used to specify the object appearance prior distribution, transition kernel, and Generalized Polya Urn distributions (C and D).\newpage %remove if formatting changes
+\hspace{-5mm}The object appearance prior parameters include:
 \begin{center}
-\begin{tabular}[c]{l p{7cm}}
-$\boldsymbol{\mu}_{0}$  &  The mean prior. In experiments performed in Section~\ref{sec:experiments} the data was recentered to the origin, and this parameter was set to $(0,0)$ \\
-$\kappa_{0}$  &  scale factor of the variance of the prior on the mean\\
-$\nu_{0}$  &  scale factor of the variance of the prior on the covariance\\
-$\Lambda_{0}$  &  shape factor of the prior on the covariance\\
-$\bold{q}_{0}$  &  scale factor of the prior on the multinomial counts
+\begin{tabular}[c]{| l |  p{7cm} | }
+\hline
+$\boldsymbol{\mu}_{0}$  &  Mean position prior. In experiments performed in Section~\ref{sec:experiments} the data was recentered to the origin, and this parameter was set to $(0,0)$. \\ \hline
+$\kappa_{0}$  &  Scale factor of the mean prior.\\ \hline
+$\Lambda_{0}$  &  Shape factor of the covariance prior.\\ \hline
+$\nu_{0}$  &  Scale factor of the covariance prior.\\ \hline
+$\bold{q}_{0}$  &  Scale factor of the multinomial prior. \\
+\hline
 \end{tabular}
 \end{center} \vspace{3mm}
-The following parameter dictates characteristics involving the movement of objects
+The following parameter dictates characteristics of object movement.
 \begin{center}
-\begin{tabular}[c]{l p{7cm}}
-$M$  &  Number of auxiliary variables. A large number will produce lower variation in position (better for slower objects) and a low number will produce higher variation in position.
+\begin{tabular}[c]{ | l | p{7cm} | }
+\hline
+$M$  &  Number of auxiliary variables. A larger number will produce lower variation in position (better for slower objects) and a smaller number will produce higher variation in position. \\
+\hline
 \end{tabular}
 \end{center} \vspace{3mm}
-Additionally, one can tune the model's tendency to detect new objects and maintain the existence of these objects (or the rate at which objects become inactive or leave the video) with the following parameters
+Additionally, one can tune the model's tendency to detect new objects and maintain the existence of these objects (both dictated by distributions C and D) with the following parameters.
 \begin{center}
-\begin{tabular}[c]{l p{7cm}}
-$\alpha$  &  The concentration parameter for the Dirichlet Process. A higher value will increase the tendency for new objects to be detected.\\
-$\rho$  &  The deletion parameter. A higher value will give objects an increased tendency to die off.
+\begin{tabular}[c]{ | l | p{7cm} | }
+\hline
+$\alpha$  &  The concentration parameter for the Dirichlet Process. A higher value will increase the tendency for new objects to be detected.\\ \hline
+$\rho$  &  The deletion parameter. A higher value will give objects an increased tendency to die off. \\
+\hline
 \end{tabular}
@@
 The parameters chosen for use in experiments (Section~\ref{sec:experiments}) were accepted based on the fact that they yielded reasonable results.
-Additionally, a sensitivity analysis was carried out for the concentration and deletion parameters ($\alpha$ and $\rho$), where a range of values for both were chosen, and inference performed for each; results of this analysis are discussion in Section~\ref{sec:sensitivityanalysis}.
+Additionally, a sensitivity analysis was carried out for the concentration and deletion parameters ($\alpha$ and $\rho$), where inference was performed for range of both values (Section~\ref{sec:sensitivityanalysis}).
 
@@
 
-Bayesian inference algorithms, which infer distributions over latent model parameters, are used to get detection and tracking results.
-A variety of methods previously developed for statistical inference can be applied to the generative model defined in Sections~\ref{sec:gpuddpm} and \ref{sec:modelspecification}.
-We've separated descriptions of the model and inference algorithms so that multiple inference methods, each with its own benefits in terms of precision and speed, can be formulated as independent algorithms for detection and tracking.
-This section provides background on the two Bayesian inference algorithms implemented in this study.
+Bayesian inference algorithms, which infer distributions over latent model parameters, are used to achieve detection and tracking results.
+Previously developed inference strategies can be applied to the generative model defined in Section~\ref{sec:gpuddpm} and Section~\ref{sec:modelspecification}.
+We provide details on the two Bayesian inference algorithms implemented in this study.
 The first is a type of Markov Chain Monte Carlo (MCMC) batch inference, which uses Gibbs sampling to generate samples from the posterior distribution of the model.
-The second is a type of Sequential Monte Carlo (SMC) inference, also known as a particle filter, which also generates samples from the posterior distribution of the model, but in a sequential manner that makes use of the time-dependent nature of the data.
+The second is a type of Sequential Monte Carlo (SMC) inference, also known as a particle filter, which exploits the time-series data to generate samples from the posterior distribution of the model in a sequential manner.
 
@@
 
-This section details a Markov Chain Monte Carlo (MCMC) sampler used to perform inference, which serves as an algorithm for unsupervised detection and tracking.
-It operates by iteratively sampling from conditional distributions of elements in the joint probability distribution---i.e.
-it implements Gibbs sampling---in order to gain approximate samples from the posterior distribution of the cluster parameters given the data.
-Additionally, this sampler makes use of the Metropolis-Hastings (M-H) algorithm to sample from a few distributions.
-A secondary formulation of the GPUDDPM, which we refer to as the ``Deletion Variable Formulation'', is used (defined in Section~\ref{sec:deletionvariableformulation}).
+This section details the MCMC sampler used to perform inference, which serves as our first algorithm for unsupervised detection and tracking.
+It iteratively samples from the conditional distribution of each element in our model given all the rest---i.e.
+implements Gibbs sampling---to generate approximate samples from the posterior distribution of the model.
+Additionally, we implement the Metropolis-Hastings (MH) algorithm to sample from a few of these conditional distributions.
+A secondary formulation of the GPUDDPM, which we refer to as the ``Deletion Variable Formulation'' (defined in Section~\ref{sec:deletionvariableformulation}), is used here.
 This formulation is equivalent to the formulation given in Section~\ref{sec:gpuddpm}, but allows for easier Gibbs sampling.
@@
 
-Instead of incorporating the cluster size variable $m_{k,t}$ directly in the GPUDDPM model (as it was in the definition given by \eqref{gpuddpm_def}), we can formulate an equivalent model which makes use of new set of variables called the deletion variables, which represent the times at which assignments are deleted.
-More precisely, each deletion variable $d_{i,t}$ denotes the time at which the associated observation $\bold{x}_{i,t}$ is considered removed from its assigned cluster $c_{i,t}$.
-At each time step, cluster sizes can be recontructed from all previous assignment and deletion variables.
-Specifically, we can compute the size of cluster $k$ at time $t$, $m_{k,t}$, by
+Instead of incorporating the cluster size variable $m_{k,t}$ directly in the GPUDDPM model (as it was in the definition given by \eqref{gpuddpm_def}), we can formulate an equivalent model which makes use of new set of variables called the deletion variables.
+We introduce a deletion variable $d_{i,t}$ for each observation $\bold{x}_{i,t}$, which denotes the time at which the observation is removed from its assigned cluster.
+At each time, cluster sizes $m_{k,t}$ can be reconstructed from all previous assignments and deletion variables by
 \begin{equation}
@@
 where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise.
-Additionally, for an observation $\bold{x}_{i,t}$, the deletion time $d_{i,t}$ can be defined to be $d_{i,t} = t + l_{i,t}$, where $l_{i,t}$ can be thought of as the lifespan of an assignment, is distributed geometrically, and can be expressed as
+Additionally, we can define the deletion variable $d_{i,t}$ to be $d_{i,t} = t + l_{i,t}$, where $l_{i,t}$ can be thought of as the lifespan of an assignment.
+From the definition of distribution D (given by \eqref{D_distro}), the lifespan can be shown to be distributed geometrically, and can be written as
 \begin{equation}
@@
 \end{equation}
-where the deletion parameter $\rho$ is the same as that given in \eqref{D_distro}.
-We can write the new formulation of the GPUDDPM as, for each time step $t \in \{1, \ldots, T\}$ and cluster $k \in \{ 1, \ldots, K_{t} \} $ at time $t$,
+where the parameter $\rho$ is the same as that in \eqref{D_distro}.
+
+We can now define the Deletion Variable Formulation of the GPUDDPM generatively to be
 \begin{align}
@@
 \begin{cases}
-P(\theta_{k,t} | \theta_{k,t-1}) \hspace{2mm} \text{if} \hspace{2mm} k \in \{ 1, \ldots, K_{t} \} \\
-\mathbb{G}_{0} \hspace{2mm} \text{if} \hspace{2mm} k = K_{t} + 1
+P(\theta_{k,t} | \theta_{k,t-1}) \hspace{1mm} \text{if} \hspace{1mm} k \leq K_{t} \\ % \in \{ 1, \ldots, K_{t} \} \\
+\mathbb{G}_{0} \hspace{1mm} \text{if} \hspace{1mm} k = K_{t} + 1
 \end{cases} \\
@@
 \end{align}
-where $\bold{c}_{t} = c_{1,t}, \ldots, c_{N_{t}, t}$ (for $N_{t}$ observations at time $t$), $\bold{d}_{t} = d_{1,t}, \ldots, d_{N_{t}, t}$, and we specify distributions for F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ in Section~\ref{sec:modelspecification}.
+$\forall$ times $t \in \{1, \ldots, T\}$ and clusters $k \in \{ 1, \ldots, K_{t} \} $ at time $t$,
+where $\bold{c}_{t} = c_{1:N_{t},t}$, $\bold{d}_{t} = d_{1:N_{t},t}$, and the distributions F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ are described in Section~\ref{sec:modelspecification}.
 This formulation of the GPUDDPM is also used by \cite{gasthaus_thesis} and \cite{caron_2007}.
@@
         \center{\includegraphics[width=82mm]{../img/gpuddp_gm_3.pdf}}
-        \caption{Graphical model of the deletion variable and auxiliary variable transition kernel representation of the GPUDDPM}
+        \caption{Graphical model of the Deletion Variable Formulation of the GPUDDPM, showing auxiliary variables.}
         \label{fig:gpuddpm_gm_2}
@@
 
-We define the state of the sampler to consist of the assignment variables for all observations, $c_{1:N_{1},1}, \ldots, c_{1:N_{T}, T}$, the deletion variables for all observations, $d_{1:N_{1},1}, \ldots, d_{1:N_{T}, T}$, the sequence of cluster parameters for each active cluster, $\theta_{k,1}, \ldots, \theta_{k, T}$ and the $M$ associated auxiliary variables for each cluster $\theta_{k,t}$ at time $t$, $z_{k,t,1}, \ldots, z_{k, t, M}$.
+We define the set of variables, S, to consist of all observations, assignment variables, deletion variables, cluster parameters (for all active clusters), and auxiliary variables (for all active clusters).
 We can write
 \begin{equation}
-\text{State} = \{ \bold{c}_{1:T}, \bold{d}_{1:T}, \Theta_{1:T}, \bold{m}_{1:T} \}
-\end{equation}
-where $\bold{c}_{t} = c_{1:N_{t},t}$, $\bold{d}_{t} = d_{1:N_{t},t}, \Theta_{t} = \theta_{1:K_{t}, t}$, and $m_{1:K_{t},t}$.
-The sampler performs an MCMC method known as Gibbs sampling; it operates by iteratively sampling each member of the state given all other members.
-In particular, at each time $t$ it moves sequentially through each of the $N_{t}$ observations, sampling the assignment $c_{i,t}$ and the deletion variable $d_{i,t}$.
-After the final observation at a given time is sampled, the cluster parameters for all active clusters at $t$ are sampled, and a M-H step is employed to sample the $M$ auxiliary variables for each active cluster.
-Before the samples can be iteratively drawn from each element of the state given all other elements, the state must be initialized.
-While the state of the sampler can be initialized at any value, beginning at a value closer to the maximum joint probability (over the state) allows for quicker inference.
-The following sections detail how each sample in the model is drawn.
+\text{S} = \{ \boldsymbol{x}_{1:T}, \bold{c}_{1:T}, \bold{d}_{1:T}, \Theta_{1:T}, \bold{z}_{1:T} \}
+\end{equation}
+where $\boldsymbol{x}_{t} = \bold{x}_{1:N_{t},t}$, $\bold{c}_{t} = c_{1:N_{t},t}$, $\bold{d}_{t} = d_{1:N_{t},t}, \Theta_{t} = \theta_{1:K_{t}, t}$, and $\bold{z}_{t} = z_{1:K_{t},t,1:M}$.
+At each time $t$, the sampler moves sequentially through the $N_{t}$ observations, sampling the assignment $c_{i,t}$ and the deletion variable $d_{i,t}$ for each.
+After, the cluster parameters for all active clusters at $t$ are sampled, and the $M$ auxiliary variables for all active clusters at $t$ are sampled (using the MH algorithm).
+Note that before we can iteratively sample each of these elements, each variable in S must be initialized.
+The following sections detail the posterior distributions from which each of these samples is drawn.
 
@@
 
-A value proportional to the posterior probability can be computed for each possible value that $c_{i,t}$ may take on; these values allow us to construct a discrete probability distribution from which we can draw samples from the posterior distribution over assignments.
-The possible values that the assignment may take on are $c_{i,t} = k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*}\}$, where $K_{t}^{'}$ is the number clusters with a non-zero size at any point between time $t$ and the assignment's associated deletion time $d_{i,t}$, and $k^{*}$ is a potential new cluster.
-The probability that $c_{i,t}$ is assigned to a cluster $k$ can be computed for each $k \in \{ 1 , \ldots ,  K_{t}^{'}, k^{*} \}$, given all other members of the state, by
+A value proportional to the posterior probability can be computed for each possible value that $c_{i,t}$ may take on.
+These values allow us to construct a discrete probability distribution from which we can draw samples from the posterior distribution over assignments.
+The possible values that the $c_{i,t}$ may take on are $k \in \{ 1 , \ldots ,  K_{t}^{'}+1 \}$, where $K_{t}^{'}$ denotes the number clusters with a non-zero size at any time $t' \in \{ t, \ldots, d_{i,t} \}$, and $K_{t}'+1$ denotes a newly sampled cluster.
+The probability that assignment $c_{i,t}$ equals a potential cluster $k$, given all other members of S (denoted S$'$), is given by
 \begin{equation}
@@
 \begin{split}
-p(c_{i,t} = k | \text{State}) & \propto
-\prod_{i'=i}^{N_{t}}  \text{C}(c_{i',t} | \bold{m}_{t-1}, \alpha) \\
-& \times \prod_{t' = t+1}^{d_{i,t}}  \prod_{i'=1}^{N_{t'}}   \text{C}(c_{i',t'} | \bold{m}_{t'-1}, \alpha) \\
+p(c_{i,t} = k | \text{S}') & \propto
+\prod_{i'=i}^{N_{t}}  \text{C}(c_{i',t} | \bold{m}_{t}, \alpha) \\
+& \times \prod_{t' = t+1}^{d_{i,t}}  \prod_{i'=1}^{N_{t'}}   \text{C}(c_{i',t'} | \bold{m}_{t'}, \alpha) \\
  & \times
 \begin{cases}
-  F(\bold{x}_{i,t} | \theta_{k,t})                                        \hspace{12pt} \text{if} \hspace{4pt} k \in \{ 1, \ldots, K_{t}^{'} \} \\
-  \int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta) d\theta    \hspace{12pt} \text{if}  \hspace{4pt}  k = k^{*}
+  F(\bold{x}_{i,t} | \theta_{k,t}) \hspace{2mm} \text{if} \hspace{1mm} k \leq K_{t}' \\ % \in \{ 1, \ldots, K_{t}' \} \\
+  \int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta) d\theta \hspace{2mm} \text{if} \hspace{1mm}  k = K_{t}'+1
 \end{cases}
@@
 \end{equation}
-where the C is given by \eqref{C_distro}, and the sizes of the clusters $\bold{m}_{t'-1}$ are calculated under the assumption that $c_{i,t} = k$.
+where the C is given by \eqref{C_distro}, and the cluster sizes $\bold{m}_{t'}$ are calculated under the assumption that $c_{i,t} = k$.
 Note that the above integral can be determined analytically to give
 \begin{equation}
+\begin{split}
 \label{marginal_obs_prob}
-\int P(\bold{x}_{i} | \theta) \mathbb{G}_{0}(\theta)d\theta = \text{Student}_{\nu_{n} - 1}(\mu_{n}, \Lambda_{n}/(\kappa_{n}(\nu_{n}-1)))
-\end{equation}
-where ``Student'' denotes the Student's t distribution, where we follow the three-value parameterization (location parameter, scale parameter, and degrees of freedom) given in \cite{gelman2004bayesian}.
+\int & P(\bold{x}_{i} | \theta) \mathbb{G}_{0}(\theta)d\theta \\
+&= \text{Student}_{\nu_{n} - 1}(\mu_{n}, \Lambda_{n}/(\kappa_{n}(\nu_{n}-1)))
+\end{split}
+\end{equation}
+where Student denotes the t-distribution, where we follow the three-value parameterization (location parameter, scale parameter, and degrees of freedom) given in \cite{gelman2004bayesian}.
 
@@
 If a new cluster is sampled as an assignment, the cluster parameters and auxiliary variables for this new cluster must be initialized for all time steps before sampling can proceed.
-In experiments, this initialization was carried out by iteratively samping forward to time T and backwards to time 1 via the transition kernel.
+In our implementation, newly sampled clusters were initialized by iteratively sampling forward to time T and backwards to time 1 via the transition kernel.
 
@@
 \subsubsection{Sampling Cluster Parameters $\theta_{k, t}$}
-The conjugacy of the chosen distributions in the appearance model and transition kernel allow for posterior samples of the cluster parameters to be easily gained, and gives us
-% \begin{eqnarray}
-% P(\theta_{k,t} | state) && = P(\bold{x}_{i} | \theta_{k,t}) P(z_{k,t+1,1:M} | \theta_{k,t})  P(\theta_{k,t} | z_{k,t,1:M}) \\
-%  && = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}, \Sigma | \boldsymbol{\mu}_{0}, k_{0}, v_{0}, \Lambda_{0})  \mathcal{D}ir( \bold{p} | \bold{q}_{0})
-% \end{eqnarray}
+
+The conjugacy of chosen distributions in the appearance model and transition kernel allow us to easily sample from the posterior distribution over the cluster parameters, which we can write
 \begin{align}
 \begin{split}
-P(\theta_{k,t} | \text{State}) & = P(\bold{x}_{i,t} | \theta_{k,t}) P(z_{k,t+1,1:M} | \theta_{k,t})  P(\theta_{k,t} | z_{k,t,1:M}) \\
- & = \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}_{k,t}, \Sigma_{k,t} | \boldsymbol{\mu}_{N}, k_{N}, v_{N}, \Lambda_{N})  \mathcal{D}ir( \bold{p}_{k,t} | \bold{q}_{N})
+P(\theta_{k,t} | \text{S}') = & \hspace{1mm} P(\bold{x}_{i,t} | \theta_{k,t}) P(z_{k,t+1,1:M} | \theta_{k,t})
+\\ & \times P(\theta_{k,t} | z_{k,t,1:M}) \\
+ = & \hspace{1mm} \mathcal{N}i\mathcal{W}(\boldsymbol{\mu}_{k,t}, \Sigma_{k,t} | \boldsymbol{\mu}_{N}, k_{N}, v_{N}, \Lambda_{N}) \\
+  & \times \mathcal{D}ir( \bold{p}_{k,t} | \bold{q}_{N})
 \end{split}
 \end{align}
-Where the parameters in the above distribution are given when the data at time $t$, auxiliary variables $z_{k,t-1,1:M}$ at time $t-1$, and auxiliary variables $z_{k,t,1:M}$ at time $t$, are taken to be the observations for the following Bayesian updates
+Where the parameters in the above distribution are given when the observations $\bold{x}_{1:N_{t},t}$, auxiliary variables $z_{k,t-1:t,1:M}$ for cluster k at time $t-1$ and $t$, are taken to be the ``observations'' for the following Bayesian updates
 \begin{eqnarray}
@@
 \end{eqnarray}
-where $N$ is the number of observations, $\{ \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0} \}$ are the $\mathcal{N}i\mathcal{W}$ prior parameters, $\bold{q}_{0}$ is the $\mathcal{D}ir$ prior parameter, $\bold{x}^{s}$ and $\bold{x}^{c}$ respectively denote the spatial and color features of the observations, and $\overline{\bold{x}}$ and $S_{\bold{x}}$ respectively denote the sample mean and sample covariance for a set of observations $\bold{x}$, defined in \eqref{samplemean} and \eqref{samplecov}.
-% \begin{eqnarray}
-% \overline{\bold{x}}  &=&  \left( \sum_{i=1}^{N} \bold{x}_{i} \right) / N\\
-% S_{\bold{x}}  &=&  \sum_{i=1}^{N} (\bold{x}_{i} - \overline{\bold{x}}) (\bold{x}_{i} - \overline{\bold{x}})^{T}
-% \end{eqnarray}
+where $N$ is the number of observations, $\{ \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0} \}$ are the $\mathcal{N}i\mathcal{W}$ prior parameters, $\bold{q}_{0}$ is the $\mathcal{D}ir$ prior parameter, $\bold{x}^{s}$ and $\bold{x}^{c}$ respectively denote the spatial and color features of the observations, and $\overline{\bold{x}}$ and $S_{\bold{x}}$ respectively denote the sample mean and sample covariance for the set of observations $\bold{x}$, defined in \eqref{samplemean} and \eqref{samplecov}.
 
@@
 \subsubsection{Sampling Auxiliary Variables $z_{k,t,m}$}
-The posterior distribution of each of the auxiliary variables $z_{k,t,m}$ can be written as
+\label{sec:sample_aux_vars}
+
+The posterior distribution over each of the auxiliary variables $z_{k,t,m}$ can be written as
 \begin{equation}
 \label{stationary_pdf}
-P(z_{k,t,m} | \text{State}) \propto  P(z_{k,t,m} | \theta_{k,t-1}) P(\theta_{k,t} | z_{k,t,1:M})
-\end{equation}
-This distribution does not allow for samples to be easily drawn and thus requires a more sophisticated MCMC technique.
-In particular, the Metropolis-Hastings (M-H) algorithm is used to gain samples from the posterior distribution for the auxiliary variables; it involves constructing a markov chain with the sought posterior as its stationary distribution, and sampling sequentially from this constructed chain to get an approximate posterior sample.
-% The M-H algorithm is outlined in Algorithm 1.%Algorithm~\ref{alg:MH}. \willie{am i citing wrong?}
-
+P(z_{k,t,m} | \text{S}') \propto  P(z_{k,t,m} | \theta_{k,t-1}) P(\theta_{k,t} | z_{k,t,1:M})
+\end{equation}
+Samples cannot easily be drawn from this distribution.
+We therefore use the MH algorithm to sample auxiliary variables from the posterior distribution (for each living cluster).
+The MH algorithm involves constructing a markov chain with the sought posterior as its stationary distribution, and sampling sequentially from this constructed chain to get an approximate posterior sample.
+% The MH algorithm is outlined in Algorithm 1.%Algorithm~\ref{alg:MH}. \willie{am i citing wrong?}
 % \begin{algorithm}[h!]
@@
 % \end{algorithm}
-
-The M-H algorithm is used to sample auxiliary variables for each alive (positive size) cluster.
 In short, we sample a new value for all $M$ auxiliary variables, denoted $z_{k,t,m}^{*}$, according to the following proposal distribution
@@
 \end{align}
-compute the standard M-H acceptance ratio, which in this case (due to the proposal distribution choice) simplifies to 
+compute the standard MH acceptance ratio, which in this case simplifies to 
 \begin{equation}
@@
 \end{equation}
-and choose these auxiliary variables as the next sample in the constructed sequence if $\alpha < \text{min}(1, r_{accept})$, where $\alpha$ is a sample from a $\text{Uniform}(0,1)$ distribution.
+and choose $z_{k,t,m}^{*}$ as the next sample in the constructed sequence if $\alpha < \text{min}(1, r_{accept})$, where $\alpha$ is a sample from a $\text{Uniform}(0,1)$ distribution.
 
@@
 \subsubsection{Sampling Deletion Variables $d_{i,t}$}
-Computing a value proportional to the posterior distribution for each possible value that a given deletion variable may take on may be accomplished in a manner similar to how samples from the posterior distribution of the assignment variables $c_{i,t}$ are gained in Section~\ref{sec:sample_assignments}; however, this may be computationally expensive due to the large number of possible deletion times.
-To remedy this, the M-H algorithm may again be used to sample from the posterior, where the prior distribution for the deletion times can be used to propose new samples $d_{i,t}^{*}$.
-As the ``lifetime'' $l_{i,t}$ of each assignment can be shown to be geometrically distributed, we can propose a new deletion time sample with
+Sampling deletion variables could be performed in a manner similar to how we sample the assignment variables in Section~\ref{sec:sample_assignments}, but this may be computationally expensive due to the large number of possible deletion times.
+To remedy this, the MH algorithm may again be used to generate samples from the posterior distribution over possible deletion times, where the prior distribution for the deletion times can be used to propose new samples $d_{i,t}^{*}$.
+As the ``lifetime'' $l_{i,t}$ of each assignment is geometrically distributed, we can propose each new deletion time sample in our constructed chain with
 \begin{align}
@@
 \end{align}
-and accept or reject this sample using the standard M-H acceptance ratio.
+and accept or reject this sample using the process described in Section~\ref{sec:sample_aux_vars}.
 %%%%%
-% need to include the "standard M-H acceptance ratio"
+% need to include the "standard MH acceptance ratio"
 %%%%%
@@
 & \times \prod_{i=1}^{N_{t}} P(c_{i,t} | \bold{m}_{t}, \Theta_{t}, \bold{c}_{1:t}, \bold{x}_{1:N_{t}}) \\
-& \times \prod_{k=1}^{K} 
+& \times \prod_{k=1}^{K_{t}} 
 \begin{cases}
@@
 \end{cases} \\
-& \times \prod_{k=1}^{K} \text{D}(m_{k,t} | m_{k,t-1}, c_{1:N_{t-1}, t-1}, \rho)
-\end{split}
-\end{equation}
-where $\bold{c}_{t} = c_{1,t}, \ldots, c_{N_{t}, t}$, $\Theta_{t} = \theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and $\bold{m}_{t} = m_{1,t}, \ldots, m_{K_{t}, t}$ The distributions used in the SMC algorithm are described in detail in the following sections.
+& \times \prod_{k=1}^{K_{t}} \text{D}(m_{k,t} | m_{k,t-1}, c_{1:N_{t-1}, t-1}, \rho)
+\end{split}
+\end{equation}
+where $\bold{c}_{t} = c_{1:N_{t},t}$, $\Theta_{t} = \theta_{1:K_{t},t}$, and $\bold{m}_{t} = m_{1:K_{t},t} $, and D is given by \eqref{D_distro}.
+%The distributions used in the SMC algorithm are described in detail in the following sections.
 
@@
 \IF{$s=S$}
-\STATE Sample $m_{k,t+1}^{(l)} \sim \text{D}(m_{k,t}^{(l)}, c_{1:N_{t}, t}^{(l)}, \rho)$  \hfill $\triangleright$ eq.~\eqref{del_step}
+\STATE Sample $m_{k,t+1}^{(l)} \sim \text{D}(m_{k,t}^{(l)}, c_{1:N_{t}, t}^{(l)}, \rho)$  \hfill $\triangleright$ eq.~\eqref{D_distro}
 \ENDIF
@@
 \ENDFOR
-\STATE $\tilde{w}_{t}^{(l)} \leftarrow w_{t-1}^{(l)} \times \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})} {P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) } \times r $  \hfill $\triangleright$ eq.~\eqref{smc_weight_eqn}
+\STATE $\tilde{w}_{t}^{(l)} \leftarrow w_{t-1}^{(l)} \times \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})} {P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) } \times r $  % \hfill $\triangleright$ eq.~\eqref{smc_weight_eqn}
 \ENDFOR
@@
 The probability of assignments given current cluster sizes, cluster parameters, and the Dirichlet Process concentration parameter $\alpha$ can be written as
-\begin{equation}
+\begin{align}
+\begin{split}
 \label{smc_assig_eqn}
-\begin{split}
-P \left( c_{i,t} | m_{1:K_{t}, t}, \theta_{1:K_{t}, t}, \alpha \right) &\propto
-\text{C}(m_{1:K_{t}, t}, \alpha) \\ &\times
+P &\left( c_{i,t} | m_{1:K_{t}, t}, \theta_{1:K_{t}, t}, \alpha \right) \propto
+\text{C}(m_{1:K_{t}, t}, \alpha) \\ 
+& \hspace{15mm}\times
 \begin{cases}
@@
 \end{split}
-\end{equation}
+\end{align}
 where C is defined in \eqref{C_distro}, and the integral \\
@@
 
-\subsection{Calculation of particle weights $\tilde{w}_{t}^{(l)}$}
-In Algorithm~\ref{alg:SMC}, the weight $\tilde{w}_{t}^{(l)}$ of each particle $l \in \{ 1. \ldots, L \}$ at time $t$ is given in terms of the particle's weight at time $t-1$ by the equation
-\begin{equation}
-\tilde{w}_{t}^{(l)} \leftarrow w_{t-1}^{(l)} \times \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})} {P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) } \times r
-\end{equation}
-where $r$ denotes the computed ratio for the particle (defined in Algorithm~\ref{alg:SMC}), and we define
-\begin{equation}
-\label{smc_weight_eqn}
-\begin{split}
-& \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})} {P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) }  \\
-& = \frac{P(\bold{x}_{1:N_{t},t} | \theta_{1:K_{t}^{(l)}}) P(c_{1:N_{t}, t}^{(l)} | m_{1:K_{t},t}^{(l)}) )}{P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t})}
-\end{split}
-\end{equation}
-where $P(\bold{x}_{1:N_{t},t} | \theta_{1:K_{t}^{(l)}})$ is defined by \eqref{likelihood}, $P(c_{1:N_{t}, t}^{(l)} | m_{1:K_{t},t}^{(l)}))$ is defined by \eqref{C_distro}, and $P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t})$ is defined by \eqref{mcmc_assig_vars}.
+% \subsection{Calculation of particle weights $\tilde{w}_{t}^{(l)}$}
+% In Algorithm~\ref{alg:SMC}, the weight $\tilde{w}_{t}^{(l)}$ of each particle $l \in \{ 1. \ldots, L \}$ at time $t$ is given in terms of the particle's weight at time $t-1$ by the equation
+% \begin{equation}
+% \tilde{w}_{t}^{(l)} \leftarrow w_{t-1}^{(l)} \times \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})} {P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) } \times r
+% \end{equation}
+% where $r$ denotes the computed ratio for the particle (defined in Algorithm~\ref{alg:SMC}), and we define
+% \begin{equation}
+% \label{smc_weight_eqn}
+% \begin{split}
+% & \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})} {P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) }  \\
+% & = \frac{P(\bold{x}_{1:N_{t},t} | \theta_{1:K_{t}^{(l)}}) P(c_{1:N_{t}, t}^{(l)} | m_{1:K_{t},t}^{(l)}) )}{P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t})}
+% \end{split}
+% \end{equation}
+% where $P(\bold{x}_{1:N_{t},t} | \theta_{1:K_{t}^{(l)}})$ is defined by \eqref{likelihood}, $P(c_{1:N_{t}, t}^{(l)} | m_{1:K_{t},t}^{(l)}))$ is defined by \eqref{C_distro}, and $P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t})$ is defined by \eqref{mcmc_assig_vars}.
 
@@
 
-The multivariate normal-multinomial model on which inference is described in this section can be used to gain object detection and tracking results.
-The nature of object detection and tracking results may vary depending on the application or performance metrics used (see Section~\ref{sec:performanceevaluationmetrics}); in this study, we desire the centroid position and approximate spatial region for each moving object, for each frame that this object can be seen within the video.
-We choose to make our detection and tracking results depend completely on the inferred parameters of the multivariate normal distribution of each cluster.
-Each inferred cluster was taken to be a distinct detected object, and the sequence of means and covariance matrices for a given cluster were used, respectively, to determine the position and spatial region of a given object over a sequence of time steps.
-In particular, the mean parameter was taken to be the centroid of an object, and a 2-dimensional oval denoting a boundary around the mean that contains a specified percentage of the data (where the specification for this oval is given in Section~\ref{sec:pets2000and2001}) was taken to be the spatial region of the object.
-Furthermore, as both provided inference methods use sampling, and each sample represents a result, we must choose a way to narrow these down to a single result.
-We have found that calculating the joint log probability for each sample and choosing the sample that yields the highest value as a result works reasonably well.
+The nature of object detection and tracking results may vary depending on the application; in this study, due to the performance metrics we wish to compute (see Section~\ref{sec:performanceevaluationmetrics}), we desire the approximate spatial region for each moving object, for each frame that this object can be seen within the video.
+We choose to make our detection and tracking results depend completely on the inferred parameters of the multivariate normal factor of each cluster.
+Each inferred cluster is taken to be a distinct detected object, and the sequence of means and covariance matrices for a given cluster are used, respectively, to determine the position and spatial region of a given object over a sequence of time steps.
+In particular, the mean parameter is taken to be the centroid of an object, and a 2-dimensional oval centered on the mean that contains a specified percentage of the normal distribution mass (where we refer to the specified percentage as the confidence value) is taken to be the spatial region of the object.
+Furthermore, as both inference procedures described in this paper involve sampling, and each sample represents a result, we must determine a way to narrow these samples down to get a single result.
+We have found that calculating a value proportional to joint probability of the model for each sample and choosing the sample that yields the highest value gives reasonable results.
 
@@
 
-This section provides details on performance evaluation metrics that have been developed to quantify results in object detection and tracking studies (and adopted in this paper), synthetic video experiments that verify certain aspects of the developed technique, benchmark video experiments that demonstrate the performance of this technique in relation to other methods developed in recent years, and new video experiments to show the capability of this technique to perform unsupervised detection and tracking on a range of object and video types.
+This section provides details on performance evaluation metrics that have been developed to quantify results in object detection and tracking studies (and adopted in this paper), synthetic video experiments that verify aspects of the developed technique, and benchmark video experiments that demonstrate the performance of this technique in relation to other strategies (including state-of-the-art, object-specific strategies) that have been developed in recent years.
 
@@
 The above metrics are dependent upon ground-truth data specifying the positions of each object in each frame throughout a video sequence.
-In the experiments described below, we gained the synthetic video ground-truth during construction of the videos (described in Section~\ref{sec:syntheticvideos}), and we used the Video Performance Evaluation Resource (ViPER) ground-truth software \cite{doermann_2000}, an open source tool commonly used in the video tracking community, to author ground-truth data for each of the benchmark datasets.
+In the experiments described below, we recorded the synthetic video ground-truth during construction of the videos (described in Section~\ref{sec:syntheticvideodatasets}), and we used the Video Performance Evaluation Resource (ViPER) ground-truth software \cite{doermann_2000}, an open source tool commonly used in the video tracking community, to author ground-truth data for each of the benchmark datasets.
 
@@
 
+\subsection{Data Extraction in Experiments}
+\label{sec:dataextractioninexperiments}
+
+Frame differencing is used in all experiments to identify pixels exhibiting motion. 
+For each pixel $\bold{x} = ( x_{1}, x_{2}, t )$ recorded during frame differencing, we also extract color information.
+Specifically, we specify a square, $L$ pixels in length, centered on $(x_{1}, x_{2})$, that contains a set of pixels surrounding $\bold{x}$ in frame $t$.
+We choose a scalar color value that is able to be computed for each pixel; this value could, for example, be some function of the red-green-blue (r-g-b) or hue-saturation-value (h-s-v) characteristic of a pixel.
+The color value for each of the selected pixels surrounding $\bold{x}$ is recorded.
+Afterwards, the set of possible color values (ie the range of color values to which a pixel may be assigned) is partitioned into $V$ bins, and the number of pixels with a color value lying in each of the bins yields the $V$ dimensional vector of color counts.
+
 \subsection{Synthetic Video Datasets}
@@
 
-Parameters were set to the same values for inference on both videos, using the method outlined in Section~\ref{sec:parameter_recap} to choose parameters.
-The chosen values were $\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (5, \ldots, 5)$.
+Parameters were set to the same values for inference on both videos.
+The chosen values were $\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right)$, and $q_{0} = (5, \ldots, 5)$.
 Inference was carried out using the MCMC algorithm (Section~\ref{sec:MCMC}); the posterior sample taken as the result correctly tracked both colored squares through occlusion in both videos, and is shown in Figure~\ref{fig:synth_one_plot}.
@@
 
-Parameter values were chosen in the same way, and set to the same values, as the first synthetic experiment.
+Parameters were set to the same values chosen in the first synthetic experiment.
 The MCMC inference algorithm correctly tracked all three objects through occlusion and inferred the appearance and size shifts.
@@
 The data is plotted in three dimensional space, where time is represented on the vertical axis, and the assignment of each data point to one of the three inferred clusters is denoted by color and marker type.
+
+\begin{figure*}[!]
+  \centering
+  \subfloat[]{\label{fig:pets2000_results} \includegraphics[width=0.48\textwidth]{../img/pets2000_front.pdf}} \hspace{1mm}           
+  \subfloat[]{\label{fig:pets2001_results} \includegraphics[width=0.48\textwidth]{../img/pets2001_back.pdf}} \\
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f1}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f4}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f9}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f11}} \\           
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f1.pdf}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f3.pdf}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f5.pdf}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f8.pdf}} \\
+  \caption{PETS2000 dataset results (a) and PETS2001 dataset results (b).
+Both plots show a sample from the posterior distribution of the state, where the vertical axis denotes time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.
+Below are four frames from the PETS2000 (c-f) and PETS2001 (g-j) sequence with one posterior sample mean and covariance matrix representation shown for each frame (and one sample mean shown for the previous 20 frames).}
+  \label{fig:pets2001_overlay}
+\end{figure*}
 
@@
 
+\begin{figure*}[!]
+  \centering             
+  \subfloat[]{\includegraphics[width=0.49\textwidth]{../img/pets2000_pm_v2.pdf}} \hspace{1mm}
+  \subfloat[]{\includegraphics[width=0.49\textwidth]{../img/pets2001_pm_v2.pdf}}
+  \caption{The SFDA (blue solid line) and ATA (red dashed line) vs  confidence values from which an object's oval region is computed for (a) PETS2000 and (b) PETS2001 video datasets.}
+  \label{fig:pm_conf}
+\end{figure*}
+
 \subsubsection{PETS2000 and PETS2001}
@@
 
-Due to the computation required for the MCMC batch inference method (discussed further in Section~\ref{sec:discussion}), only the final 1000 frames of the video were used from both datasets.
-Extraction was performed with frame differencing as described in Section~\ref{sec:dataextraction}, using $L=3$.
-Parameter values were chosen in the same way, and set to the same values, as in the the synthetic experiments ($\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (5, \ldots, 5)$).
+Due to the computation required for the MCMC batch inference method (discussed further in Section~\ref{sec:conclusion}), only the final 1000 frames of the video were used from both datasets.
+Extraction was performed with frame differencing as described in Section~\ref{sec:dataextractioninexperiments}, using $L=3$.
+Parameter values were set to the same values as in the the synthetic experiments ($\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (5, \ldots, 5)$).
 The MCMC sampler was successful for both benchmark videos; each object was detected, tracked, and its shape estimated in manner very consistent with the ground-truth.
@@
 
-\begin{figure*}[!]
-  \centering
-  \subfloat[]{\label{fig:pets2000_results} \includegraphics[width=0.48\textwidth]{../img/pets2000_front.pdf}} \hspace{1mm}           
-  \subfloat[]{\label{fig:pets2001_results} \includegraphics[width=0.48\textwidth]{../img/pets2001_back.pdf}}
-  \caption{PETS2000 dataset results (a) and PETS2001 dataset results (b).
-Both plots show a sample from the posterior distribution of the state, where the vertical axis denotes time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.}
-  \label{fig:benchmark_results_1}
-\end{figure*}
-
-\begin{figure*}[!]
-  \centering             
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f1}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f4}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f9}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f11}}
-  \caption{Four frames from the PETS2000 sequence with one posterior sample mean and covariance matrix representation shown for each frame (and for the previous 20 frames).}
-  \label{fig:pets2001_imgs}
-\end{figure*}
-
-\begin{figure*}[!]
-  \centering               
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f1.pdf}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f3.pdf}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f5.pdf}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f8.pdf}}
-  \caption{Four frames from the PETS2001 sequence with one posterior sample mean and covariance matrix representation shown for each frame (and for the previous 20 frames).}
-  \label{fig:pets2001_overlay}
-\end{figure*}
-
 To calculate performance metrics (both SFDA and ATA), one must specify a confidence value that allows the oval representing the region occupied by an object to be computed from the inferred covariance matrix of each cluster (as discussed in Section~\ref{sec:inferencetotrackingresults}).
-The performance metrics were found for a range of confidence intervals, and these curves, for the PETS2000 and PETS2001 video are shown in Figure~\ref{fig:pm_conf}.
-
-\begin{figure*}[!]
-  \centering             
-  \subfloat[]{\includegraphics[width=0.49\textwidth]{../img/pets2000_pm.pdf}} \hspace{1mm}
-  \subfloat[]{\includegraphics[width=0.49\textwidth]{../img/pets2001_pm.pdf}}
-  \caption{The SFDA (blue solid line) and ATA (red dashed line are) vs  confidence values from which an object's oval region is computed for (a) PETS2000 and (b) PETS2001 video datasets.}
-  \label{fig:pm_conf}
-\end{figure*}
+The performance metrics were found for a range of confidence intervals, and the resulting curves for both the PETS2000 and PETS2001 video are shown in Figure~\ref{fig:pm_conf}.
 
@@
 \label{sec:pets20092010}
-
-A video dataset used in both the PETS2009 and PETS2010 conferences, called ``S2.L1 at time sequence 12.34'' was chosen for experimentation due to its prominence in a number of studies \cite{ellis_2010}.
-%%%%%
-% cite other papers that use this dataset?
-%%%%%
-This dataset consists of a monocular, stationary camera, 794 frame video sequence. The entire video sequence was used in this experiment.
-
-Due to the large number of frames and objects in this video, the SMC algorithm (described in Section~\ref{sec:SMC}) was used for inference.
-This method of sequential inference was observed, on this dataset, to converge to a better sample in a shorter period of time in comparison with the MCMC algorithm; for this reason, we found that it was better suited for this large dataset.
-
-Extraction was performed with frame differencing as described in Section~\ref{sec:dataextraction}, using $L=3$ and parameters for the model were chosen, using the method outlined in Section~\ref{sec:recapofparameters}, to be $\alpha = 0.1, \rho = 0.8, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 6, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (3, \ldots, 3)$.
-Additionally, as with the other video datasets, ground-truth bounding boxes around each object were authored using the ViPER tool.
-
-The SMC inference algorithm yielded a sample from the posterior distribution of the model, from which the object detection and tracking results were obtained (as described in Section~\ref{sec:inferencetotrackingresults}).
-In Figure \ref{fig:pets2009_results}, a sample from the posterior distribution over the cluster parameters is overlayed on the extracted data over a sequence of frames, where the assignment of each data point is represented by color and marker type.
 
@@
   \hspace{4mm}
-  \subfloat[]{\includegraphics[width=0.53\textwidth]{../img/pets2009full_pm_b.pdf}}
-  \caption{ (a) Results for the PETS2009 dataset, showing a sample from the posterior distribution of the state for frames 1-50, where the vertical axis denotes time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.
-(b) performance metrics vs.
-covariance confidence interval threshold for tracking results showing the SFDA (blue solid line) and ATA (red dashed line).}
-  \label{fig:pets2009_results}
-\end{figure*}
-
-\begin{figure*}[!]
-  \centering               
+  \subfloat[]{\includegraphics[width=0.53\textwidth]{../img/pets2009full_pm_b_v2.pdf}} \\
   \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2009_f3.pdf}} \hspace{1pt}
@@
   \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2009_f14.pdf}}
-  \caption{Four frames from the PETS2009/2010 sequence with one posterior sample mean and covariance matrix representation shown for each frame (and for the previous 20 frames).}
-  \label{fig:pets2009_overlay}
+  \caption{ (a) Results for the PETS2009/2010 dataset, showing a sample from the posterior distribution of the state for frames 1-50, where the vertical axis denotes time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.
+(b) performance metrics vs.
+covariance confidence interval threshold.
+Below (c-f) are four frames with one posterior sample mean and covariance matrix representation shown for each frame (and one sample mean shown for the previous 20 frames).}
+  \label{fig:pets2009_results}
 \end{figure*}
+
+A video dataset used in both the PETS2009 and PETS-\\2010 conferences, called ``S2.L1 at time sequence 12.34'' was chosen for experimentation due to its prominence in a number of studies \cite{ellis_2010,arsic2009multi,berclaz2009multiple,conte2010performance,bolme2009simple,breitenstein2009markovian,ge2009evaluation,alahi2009sparsity,yang2009probabilistic}.
+This dataset consists of a monocular, stationary camera, 794 frame video sequence.
+The entire video sequence was used in this experiment.
+
+Due to the large number of frames and objects in this video, the SMC algorithm (described in Section~\ref{sec:SMC}) was used for inference.
+This method of sequential inference was observed, on this dataset, to converge to a better sample in a shorter period of time in comparison with the MCMC algorithm; for this reason, we found that it was better suited for this large dataset.
+
+Extraction was performed with frame differencing as described in Section~\ref{sec:dataextractioninexperiments}, using $L=3$, and parameters for the model were chosen to be $\alpha = 0.1, \rho = 0.8, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 6, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (3, \ldots, 3)$.
+Additionally, as with the other video datasets, ground-truth bounding boxes around each object were authored using the ViPER tool.
+
+The SMC inference algorithm yielded a sample from the posterior distribution of the model, from which the object detection and tracking results were obtained (as described in Section~\ref{sec:inferencetotrackingresults}).
+In Figure \ref{fig:pets2009_results}, a sample from the posterior distribution over the cluster parameters is overlayed on the extracted data over a sequence of frames, where the assignment of each data point is represented by color and marker type.
 
@@
 We compare SFDA and ATA results of our strategy with these methods to show that our general method can yield comparable results even when compared with object-specific trackers.
-Figure~\ref{fig:sfda_and_ata_comparison} shows performance metric results for comparison (data published with permission from the authors of \cite{ellis_2010}).
-\begin{figure*}[!]
-  \centering
-  \subfloat[]{\includegraphics[width=0.49\textwidth]{../img/pets2009full_comp_sfda.pdf}}
-  \hspace{1mm}
-  \subfloat[]{\includegraphics[width=0.49\textwidth]{../img/pets2009full_comp_ata.pdf}}
-  \caption{Performance metric comparison showing the (a) SFDA and (b) ATA results for submissions to the PETS2009/PETS2010 workshops (blue bars) and for the technique introduced in this paper (red bars).
-These graphs show that the general technique in this paper yields comparable performance with state-of-the-art, object-specific detection and tracking strategies.
-Our results were computed using a covariance confidence interval threshold of 0.35.}
-  \label{fig:sfda_and_ata_comparison}
-\end{figure*}
+Table~\ref{benchmark_results_table} shows performance metric results for comparison (data published with permission from the authors of \cite{ellis_2010}).
+
+%Arsic BerclazLP Conte   Bolme_ASEF  Bolme_Cascade Bolme_Parts    Breitenstein   GE  Alahi_Ogreedy   Alahi_Olasso Anon   Yang GPUDDP4MTT(sens_t25_0.35conf)
+
+\begin{table}
+\begin{tabular}[!] {| l | l | l |}
+  \hline
+  \textsc{Method Name} & \textsc{SFDA}  & \textsc{ATA} \\ \hline \hline
+  GPUDDPM & $ \bold{0.5079} $ & $\bold{0.2996}$ \\ \hline
+  \cite{arsic2009multi} & $ 0.177768 $ & $0.016111$ \\ \hline
+  \cite{berclaz2009multiple} & $ 0.482081 $ & $0.145195$ \\ \hline
+  \cite{conte2010performance} & $ 0.52966  $ & $0.058003$ \\ \hline
+  \cite{bolme2009simple} & $ 0.413812 $ & NA \\ \hline
+  \cite{bolme2009simple} & $ 0.338979 $ & NA \\ \hline
+  \cite{bolme2009simple} & $ 0.338979 $ & NA \\ \hline
+  \cite{breitenstein2009markovian} & $ \emph{0.567265} $ & $0.30044$ \\ \hline
+  \cite{ge2009evaluation} & $ 0.383235 $ & $0.040118$ \\ \hline
+  \cite{alahi2009sparsity} & $ 0.430043 $ & $0.041053$ \\ \hline
+  \cite{alahi2009sparsity} & $ 0.419141 $ & $0.045609$ \\ \hline
+  Anonymous & $ 0.52311  $ & $0.246275$ \\ \hline
+  \cite{yang2009probabilistic} & $ 0.552439 $ & $\emph{0.448237}$ \\
+  \hline
+\end{tabular}
+\caption{SFDA and ATA performance metric results shown for 13 algorithms on the PETS2009/2010 benchmark dataset.
+Results for our method (GPUDDPM) are in bold.
+The largest ATA and SFDA values are italicized.
+Data given by the authors of \cite{ellis_2010}.}
+\label{benchmark_results_table}
+\end{table}
 
@@
 
-SMC inference on the PETS2009 video dataset was carried out for a range of two key parameter values, $\alpha$ and $\rho$.
+SMC inference on the PETS2009/2010 video dataset was carried out for a range of two key parameter values, $\alpha$ and $\rho$.
 The performance metric measures, SFDA and ATA, were computed for each combination of these two parameters.
@@
 
-% The following shows the SFDA sensitivity analysis results. The largest two values are shown in bold.
-% \begin{center}
-% \begin{tabular}[!]{l l  l  l  l  l  l}
-%   & & & $\hspace{8mm}\rho$ & & & \\
-%   SFDA  &   & $0.7$ & $0.75$ & $0.8$ & $0.85$ & $0.9$  \\  
-%   &$0.01$   & $0.4121$ & $0.4262$ & $0.4219$ & $\bold{0.4331}$ & $0.4277$  \\
-%   &$0.1$    & $\bold{0.4336}$ & $0.4184$ & $0.4200$ & $0.4316$ & $0.4321$  \\ 
-% $\alpha$  & $1$    & $0.4084$ & $0.4209$ & $0.4205$ & $0.4290$ & $0.4330$  \\  
-%   &$10$     & $0.3992$ & $0.4130$ & $0.4289$ & $0.4232$ & $0.4226$  \\  
-%   &$100$    & $0.3655$ & $0.3739$ & $0.3636$ & $0.3755$ & $0.3672$  \\
-% \end{tabular}
-% \end{center}
-
-% The following shows the ATA sensitivity analysis results. The largest two values are shown in bold.
-% \begin{center}
-% \begin{tabular}[!]{l l  l  l  l  l  l}
-%     & & & $\hspace{8mm}\rho$ & & & \\
-%    ATA  &     & $0.7$ & $0.75$ & $0.8$ & $0.85$ & $0.9$  \\  
-%     &$0.01$     & $0.2032$ & $0.2017$ & $0.2262$ & $0.1945$ & $0.1914$  \\
-%       &$0.1$  & $0.1964$ & $0.2042$ & $0.2401$ & $0.1990$ & $0.2017$  \\
-% $\alpha$ & $1$    & $0.2016$ & $0.2365$ & $0.2766$ & $0.2347$ & $0.2766$  \\   
-%       &$10$     & $0.2277$ & $0.2284$ & $0.2613$ & $\bold{0.2781}$ & $0.2732$  \\  
-%      &$100$   & $0.2468$ & $0.2645$ & $0.2725$ & $0.2670$ & $\bold{0.2843}$  \\
-% \end{tabular}
-% \end{center}
-
 Figure~\ref{fig:sens_surf} shows a surface plot of the sensitivity analysis results.
@@
   \centering             
-  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2009full_sens_sfda.png}}
-  \subfloat[]{\includegraphics[width=0.5\textwidth]{../img/pets2009full_sens_ata.png}}
-  \caption{Surface plot of the sensitivity performance metric tables}
+  \subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2009full_sens_sfda_heat.pdf}}
+  \hspace{6mm}
+  \subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2009full_sens_ata_heat.pdf}}
+  \caption{Heatmap showing sensitivity of performance metrics SFDA and ATA to variation of parameters $\alpha$ and $\rho$.}
   \label{fig:sens_surf}
@@
 The primary intention of this technique is to reduce the need for detection or localization methods tailored to specific object types and serve as a general framework applicable to videos with varied objects, backgrounds, and film qualities.
-The GPUDDPM, a time-dependent Dirichlet process mixture, has been introduced, and we have shown how inference on this model allows for the desired clustering of extraction data.
+The GPUDDPM, a time-dependent Dirichlet process mixture, has been introduced, and we have shown how inference on this model allows us to achieve detection and tracking results.
 Furthermore, we have demonstrated a specific implementation of the model using spatial and color pixel data extracted via frame differencing and provided two algorithms for performing Bayesian inference on the model to accomplish detection and tracking.
@@
 We have described and computed standard performance metrics for our technique's detection and tracking results, and found it to be comparable with state-of-the-art object-specific detection and tracking methods designed for people tracking in the PETS2009/2010 video dataset.
-Results from the synthetic and benchmark video datasets illustrate the ability of the technique described in this paper to, without modification, perform completely unsupervised detection and tracking of objects with diverse physical charactersitics moving over non-uniform backgrounds and through occlusion
+Results from the synthetic and benchmark video datasets illustrate the ability of the technique described in this paper to, without modification, perform completely unsupervised detection and tracking of objects with diverse physical charactersitics moving over non-uniform backgrounds and through occlusion.
 
@@
 
-\section{Model Background}
+\section{-  Appendix}
 \label{sec:modelbackground}
@@
 
-This section provides background information that allows us to introduce DDPM models, describes the motivation behind our use of these models, and gives a definition of the particular DDPM chosen, known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM) model (Section~\ref{sec:gpuddpmixture}).
+This section provides background information that allows us to introduce DDPM models and describes the motivation behind our use of these models.
+This background is helpful for the definition of the particular DDPM chosen, the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM) model, given in Section~\ref{sec:gpuddpm}.
 

### 2012-02-29 ltx/paper_short_2.tex
newest draft (paper_short_2.tex) is here

@@
 This paper proposes a technique for the unsupervised detection and tracking of arbitrary objects in videos.
-It is intended to reduce the need for detection or localization methods tailored to specific object types and serve as a general framework applicable to videos with varied objects, backgrounds, and image qualities.
-The technique uses a Dependent Dirichlet Process Mixture (DDPM) known as the Generalized Polya Urn (GPUDDPM) to model image pixel data that can be easily and efficiently extracted from the regions in a video that represent objects.
+It is intended to reduce the need for detection and localization methods tailored to specific object types and serve as a general framework applicable to videos with varied objects, backgrounds, and image qualities.
+The technique uses a dependent Dirichlet process mixture (DDPM) known as the Generalized Polya Urn (GPUDDPM) to model image pixel data that can be easily and efficiently extracted from the regions in a video that represent objects.
 This paper describes a specific implementation of the model using spatial and color pixel data extracted via frame differencing and gives two algorithms for performing inference on the model to accomplish detection and tracking.
@@
 We begin by describing characteristics of the extracted data (Section~\ref{sec:dataextraction}), and giving the generalized form of the model (Section~\ref{sec:modeldefinition}).
-To implement this model, one must specify a data extraction procedure and distributions for representing objects, which may be chosen to allow for general object tracking or tailored to a specific object type for a given application.
-In our implementation, we give the specific distributions for representing objects in a general manner (Section~\ref{sec:modelspecification}), which we apply to data extracted via a basic frame differencing procedure.
+To implement this model, one must specify a data extraction procedure and distributions for representing objects, which may be chosen to allow for arbitrary object tracking or tailored to a specific object type for a given application.
+In our implementation, we extract data via a basic frame differencing procedure and specify distributions useful for representing arbitrary objects (Section~\ref{sec:modelspecification}).
 We describe inference algorithms for our model, and show how output of these algorithms can be interpreted as object localization and tracking results (Section~\ref{sec:inference}).
@@
 A variety of methods in the fields of image processing, signal processing, and computer vision have been developed to solve aspects of the problem of unsupervised detection and tracking of arbitrary objects.
-These methods might be placed into a few broad categories: those that aim to distinguish the foreground regions of images from the background \cite{hong2007real,chien2002efficient, zhang2007moving, kim2002fast}, segment images into distinct regions to perform localization \cite{jain1997object, fei2005bayesian, sivic2005discovering}, track an object over a sequence of images (after its position has been specified in an initial image) \cite{raja_1998, mckenna_1999, jepson_2003, comaniciu_2003, perez_2002}, track multiple objects over a sequence of images (especially when the objects interact or occlude one another) \cite{senior2006appearance, cucchiara2004probabilistic, zhou2003background, han_2004, mckenna2000tracking, dockstader2001multiple}, segment a sequence of images into distinct spatiotemporal regions \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}, and combine the previous methods in some way to create systems capable of both detecting and tracking specified objects, or of discerning which regions of a video constitute distinct, arbitrary objects and tracking these \cite{brostow2006unsupervised, brox2010object, fragkiadaki2011detection, pece_2002}.
+These methods might be placed into a few broad categories: those that aim to distinguish the foreground regions of images from the background \cite{hong2007real,chien2002efficient, zhang2007moving, kim2002fast}, segment images into distinct regions to perform localization \cite{jain1997object, fei2005bayesian, sivic2005discovering}, track an object over a sequence of images (after its position has been specified in an initial image) \cite{raja_1998, mckenna_1999, jepson_2003, comaniciu_2003, perez_2002}, track multiple objects over a sequence of images (especially when the objects interact or occlude one another) \cite{senior2006appearance, cucchiara2004probabilistic, zhou2003background, han_2004, mckenna2000tracking, dockstader2001multiple}, segment a sequence of images into distinct spatiotemporal regions \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}, and combine the previous methods in some way to create systems capable of both detecting and tracking specified objects \cite{Okuma04aboosted,eth_biwi_00633,4036928, khan_2004, leibe2008coupled}, or of discerning which regions of a video constitute distinct, arbitrary objects and tracking these \cite{brostow2006unsupervised, brox2010object, fragkiadaki2011detection, pece_2002,paragios2000geodesic}.
 
@@
 There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera) \cite{senior2006appearance, cucchiara2004probabilistic, zhou2003background}, complex object interactions \cite{khan_2004, mckenna2000tracking, dockstader2001multiple}, objects with similar appearances \cite{maccormick1999probabilistic, jepson_2003}, variable (and potentially high) numbers of objects \cite{reilly2010detection}, and objects that enter and exit a field of view at different times \cite{stauffer2003estimating, nedrich2010learning}.
-Multiple single-object trackers running simultaneously have been shown to be ineffective, as distinct single-object trackers will tend to coalesce and track the same object.
+Multiple independent single-object trackers running simultaneously have been shown to be ineffective, as they will tend to coalesce and track the same object \cite{khan_2004}.
 To remedy this problem, methods such as \cite{maccormick1999probabilistic} have developed probabilistic principles for maintaining isolation of object trackers.
@@
 Over the past decade, there have been attempts to provide general algorithms for the fully unsupervised detection and tracking of arbitrary objects in videos.
-Blob tracking, a basic method to carry out this goal, has been used for many years, and has found success in videos where objects are easily isolated from the background and where localization and segmentation of distinct objects is possible \cite{francois2004real, isard_2001}.
+Blob tracking, a basic method to carry out this goal, has found success in videos where objects are easily isolated from the background and where localization and segmentation of distinct objects is possible \cite{francois2004real, isard_2001}.
 Methods involving blob tracking, however, run into problems when faced with videos where detection is difficult, object appearance or orientation varies heavily, and there exists object occlusion \cite{song2005model}.
@@
 Other attempts to perform unsupervised detection and tracking include methods for clustering short sequences of positions extracted by detecting the motion of objects \cite{brostow2006unsupervised, brox2010object}, which aim to return full-length distinct object tracks, and graph based method that carry out a similar task using spectral clustering \cite{fragkiadaki2011detection}.
-Another approach based on a Gaussian mixture model has also been used to cluster data extracted from moving objects \cite{pece_2002}; this method also develops heuristics for the initialization and elimination of new object tracks.
-
-Nearly all methods capable of high accuracy detection and tracking of objects in videos are those tailored for specific object types.
-These methods rely on ad-hoc detection criteria that exploit previous knowledge about the appearance or behavior of the given objects in a video.
-The method described in this paper is designed to track arbitrary objects without using any explicit detection criteria, and serve as a general strategy that can be used, without modification, to perform accurate detection and tracking of diverse objects in a wide range of videos.
+Another approach uses a Gaussian mixture model to cluster data extracted from moving objects \cite{pece_2002}; this method also develops heuristics for the initialization and elimination of new object tracks.
+
+Nearly all high accuracy object detection and tracking methods are those tailored for specific object types.
+These methods rely on ad-hoc detection criteria that exploit knowledge about the appearance or behavior of the given objects in a video.
+They often make use of state-of-the-art detectors designed to locate the specified objects of interest.
+In contrast, the method described in this paper is designed to track arbitrary objects without using any explicit detection criteria, and serve as a general strategy that can be used, without modification, to perform accurate detection and tracking of diverse objects in a wide range of videos.
 The method we introduce falls into the category of clustering-based arbitrary object detection and tracking methods.
 Differing from previous work, we use a type of time dependent Bayesian nonparametric mixture model, and show how it can be applied to a variety of easily extracted data to perform detection and tracking.
-This method begins by performing a simple data extraction procedure, which can be carried out broadly but yields noisy data; our model is then applied to this noisy extracted data.
-Inference is performed on this model to achieve detection and tracking results.
+This method begins by performing a simple data extraction procedure, which can be carried out broadly but yields noisy data.
+Our model of this data serves as a general framework for which we can choose a variety of object appearance distributions and inference algorithms; each choice provides a new method for unsupervised detection and tracking of arbitrary objects in videos.
 
@@
   \subfloat[]{\label{fig:pets2009_img_framediff}\includegraphics[width=0.328\textwidth]{../img/pets2009_framediff.png}}\\
-  %\subfloat[]{\label{fig:longimg}\includegraphics[width=1.05\textwidth]{../img/longimg_6.pdf}} \\
   \subfloat[]{\label{fig:longimg_all}\includegraphics[width=1\textwidth]{../img/longimg_7_detail_7.pdf}}
-  %../img/longimg_7_detail.pdf}} 
   \caption{Three pairs of consecutive frames and the results produced by taking the pixel-wise frame difference between each pair (a - i).
@@
 
-The basic set of observations extracted from video are the three-dimensional points
-\begin{equation}
-  \centering
-  \bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \{ 1, \ldots, T \}
-\end{equation}
-where $x_{1}$ and $x_{2}$ are the two spatial locations of a pixel and $t$ is the time index.
+% The basic set of observations extracted from video are the three-dimensional points
+% \begin{equation}
+%   \centering
+%   \bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \{ 1, \ldots, T \}
+% \end{equation}
+% where $x_{1}$ and $x_{2}$ are the two spatial locations of a pixel and $t$ is the time index.
 These data are extracted by recording the positions of pixels lying in regions that contain objects.
@@
 
-We can extract additional features from videos and incorporate this new information into our model.
-For example, we can extract color information, and modify our model to allow for object color to be inferred during tracking.
-Adding this feature allows our method to better distingush between adjacent objects and more accurately track objects through occlusion.
-For each extracted pixel, we compute a $V$ dimensional vector that represents the color distribution in the pixel's immediate vicinity (details on this vector and how it is computed are given in Section~\ref{sec:dataextractioninexperiments}).
-We call the components of this vector the ``color counts'' of the pixel.
-The frame differencing and color counting extraction yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \{1, \ldots, T \}$, where each element $\bold{x} \in \bold{X}$ can be written
-\begin{equation}
+We desire a data extraction procedure that yields observations of the form
+\begin{equation}
+\centering
 \bold{x} = ( \bold{x}^{s}, \bold{x}^{c}, t ) = ( x^{s_{1}}, x^{s_{2}}, x^{c_{1}}, \ldots, x^{c_{V}}, t )
 \end{equation}
-where $\bold{x}^{s} \in \mathbb{R}^{2}$ is the two dimensional vector of spatial coordinates,  $\bold{x}^{c} \in \mathbb{Z}_{+}^{V}$ is the $V$ dimensional vector of color counts, and $t \in \{1, \ldots, T \}$ is the time index.
-Other features may also be extracted from each pixel to attempt more sophisticated object representations, including pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, and texture representations.
+where each $\bold{x}$ corresponds to a point within an image region where an object (or foreground element)  is believed to exist, $\bold{x}^{s} \in \mathbb{R}^{2}$ denotes the spatial location of this point, $\bold{x}^{c} \in S_{1} \times \ldots \times S_{V}$ denotes some collection of local image characteristics in the vicinity of this point, and $t \in \{1, \ldots, T \}$ denotes the time index.
+
+We'd like to use an extraction procedure that is as unsophisticated as possible.
+Consequently, we use frame differencing.
+This procedure locates the pixels in image regions where there is motion relative to a previous frame.
+Here, each pixel corresponds to an observation $\bold{x}$.
+Frame differencing is simple, computationally inexpensive, and able to be applied to a wide range of static, single-camera videos (the videos used in experiments are stationary; moving-camera videos require data extraction methods useful for non-stationary video \cite{chien2002efficient, zhang2007moving}).
+Examples of pixel location data extracted via frame differencing are shown in Figure~\ref{fig:img_and_framediff}(a-j).
+
+We also extract features $\bold{x}^{c}$ that capture image information in the vicinity of each extracted pixel.
+Examples of possible features include color distributions, pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, and texture representations.
+In principle, we can extract any image features that may be used to characterize the appearance of objects.
+In our implementation, we choose to extract only color information in the vicinity of each pixel.
+Incorporating color features allows our method to infer a distribution over color for each detected object; this improves its ability to distingush between adjacent objects and track objects through occlusion.
+To add this information, we let $\bold{x}^{c}$ represent a $V$ dimensional discrete distribution over some aspect of color (such as the hue) in the pixel's immediate vicinity.
+Details on this vector and how it is computed are given in Section~\ref{sec:dataextractioninexperiments}.
+We refer to the components of this vector as the ``color counts'' of the pixel.
+
+% old discussion on incorporating color features
+
+% The frame differencing and color counting extraction yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \{1, \ldots, T \}$, where each element $\bold{x} \in \bold{X}$ can be written
+% \begin{equation}
+% \bold{x} = ( \bold{x}^{s}, \bold{x}^{c}, t ) = ( x^{s_{1}}, x^{s_{2}}, x^{c_{1}}, \ldots, x^{c_{V}}, t )
+% \end{equation}
+% where $\bold{x}^{s} \in \mathbb{R}^{2}$ is the two dimensional vector of spatial coordinates,  $\bold{x}^{c} \in \mathbb{Z}_{+}^{V}$ is the $V$ dimensional vector of color counts, and $t \in \{1, \ldots, T \}$ is the time index.
+% Other features may also be extracted from each pixel to attempt more sophisticated object representations, including pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, and texture representations.
+
+% old (long) data extraction - parts grabbed for new version above
+% ----------------------------------------------------------------
 
@@
 
-\section{Model Definition}
+\section{Model Framework}
 \label{sec:modeldefinition}
 
-A type of Dependent Dirichlet Process Mixture (DDPM) model known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM) is applied to the extracted data.
-We define the generalized form of this model in Section~\ref{sec:gpuddpm}, and specify the distributions used in our implementation of this model in Section~\ref{sec:modelspecification}.
+We use a type of dependent Dirichlet process mixture (DDPM) model known as the Generalized Polya Urn dependent Dirichlet process mixture (GPUDDPM).
+We define a general form of this model in Section~\ref{sec:gpuddpm}, and specify the distributions used in our implementation of this model in Section~\ref{sec:modelspecification}.
 We also define a secondary form of this model in Section~\ref{sec:MCMC}, which is used in one of the two inference algorithms.
-Background information providing a brief introduction to mixture models and the Dirichlet process is contained in Appendix~\ref{sec:modelbackground}.
+We provide a brief introduction to mixture models and the Dirichlet process in Appendix~\ref{sec:modelbackground}.
+
+Dirichlet process mixture (DPM) models (Section \ref{sec:dpmixture}) fall under the heading of Bayesian nonparametric models.
+These models have been widely used in the past decade to perform nonparametric density estimation and cluster analysis.
+Data extracted from videos comprises spatiotemporal clusters, each corresponding with a distinct object.
+Consequently, we are interested in using a class of models known as dependent Dirichlet process mixture (DDPM) models (Section~\ref{sec:ddpmixture}), which are particularly useful for estimating the number of latent classes (clusters) in time dependent data.
+%In this work, estimating the number of clusters in video extraction data is equivalent to estimating the number of distinct objects in a video.
+
+Since objects can enter and exit a scene, the number of clusters present throughout the video may not be constant (i.e., clusters may be created, or be ``born'', and may disappear, or ``die'', at intermediate time steps).
+To cluster data with these properties, we choose to use a DDPM known as the Generalized Polya Urn dependent Dirichlet process mixture, introduced by \cite{caron_2007}.
+This model may be intuitively viewed as a sequence of DPMs, where there exist dependencies between the parameters and number of clusters in adjacent time steps.
+Like the DPM, the GPUDDPM allows for the number of clusters within a dataset---which, in this work, corresponds to the number of objects in a video---to be found in an automated manner.
 
@@
 
-The extracted data comprise a collection of spatiotemporal clusters, each corresponding with an object in the video.
-Since objects can enter and exit a scene, the number of clusters present in each time step may not be constant (i.e., clusters may be created, or be ``born'', and may disappear, or ``die'', at intermediate time steps).
-To cluster data with these properties, we choose to use a DDPM known as the Generalized Polya Urn Dependent Dirichlet Process Mixture, introduced by \cite{caron_2007}.
-This model may be intuitively viewed as a sequence of DPMs, where there exist dependencies between the parameters and number of clusters in adjacent time steps.
-Like the DPM, the GPUDDPM allows for the number of clusters within a dataset---which, in this application, corresponds to the number of objects in a video---to be found in an automated manner.
-
-Similar to a typical mixture model, in this model, each observation $\bold{x}_{i,t}$ is associated with an assignment variable $c_{i,t}$ that represents its assignment to a cluster $\theta_{k,t}$.
-In a DPM, the probability of assignment to a cluster $k$ is a function of the cluster size $m_{k,t}$, which is equal to the number of observations that have been assigned to it.
-In contrast, cluster sizes in the GPUDDPM may decrease over time as observations become ``unassigned''; this allows clusters to decrease in strength and potentially die.
-We define a distribution over the size of cluster $k$ at time $t$ ($m_{k,t}$), conditioned on the cluster's previous size ($m_{k,t-1}$), the assignments at time $t$ ($c_{1:N_{t}, t}$), and a deletion parameter ($\rho$) to be
+In this model, each observation $\bold{x}_{i,t}$ is associated with an assignment variable $c_{i,t}$ that represents its assignment to a cluster $\theta_{k,t}$. 
+%In a DPM, the probability of assignment to a cluster $k$ is a function of the cluster size $m_{k,t}$, which is equal to the number of observations that have been assigned to it. In contrast, 
+The size (i.e.
+number of assigned observations) of each cluster in the GPUDDPM may decrease over time as observations become ``unassigned''.
+We define a distribution over the size of cluster $k$ at time $t$ ($m_{k,t}$), conditioned on the cluster's previous size ($m_{k,t-1}$), the assignments at time $t$ ($c_{1:N_{t}, t}$), and a deletion parameter ($\rho$),
 \begin{equation}
@@
 
-We also define a distribution over the assignment of observation $i$ at time $t$ ($c_{i,t}$), conditioned on the sizes of all clusters at time $t$ ($m_{1:K_{t},t}$) and a concentration parameter ($\alpha$), to be
+We also define a distribution over the assignment of observation $i$ at time $t$ ($c_{i,t}$), conditioned on the sizes of all clusters at time $t$ ($m_{1:K_{t},t}$) and a concentration parameter ($\alpha$),
 \begin{equation}
@@
 
-Distributions \eqref{D_distro} and \eqref{C_distro} together comprise what \cite{caron_2007} refers to as the ``Generalized Polya Urn''. We can now define the GPUDDPM generatively to be
+Distributions \eqref{D_distro} and \eqref{C_distro} together comprise what \cite{caron_2007} refers to as the ``Generalized Polya Urn''. We can now define the GPUDDPM generatively as
 \begin{align}
@@
         \center{\includegraphics[width=80mm]{../img/gpuddp_gm_1.pdf}}
-        \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Polya Urn Dependent Dirichlet Process Mixture.
+        \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Polya Urn dependent Dirichlet process mixture.
 The observations at time $t$, $\bold{x}_{1:N_{t},t}$, and their associated assignments $c_{1:N_{t},t}$ are denoted respectively as $\bold{x}_{t}$ and $c_{t}$ (likewise for those at time $t-1$).}
@@
 
-\section{Model Specification}
+\section{Model Specifics}
 \label{sec:modelspecification}
@@
 This specification is used for all experiments in Section~\ref{sec:experiments}.
-The distributions F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ represent object appearance, the appearance prior, and object movement.
-Our specification is kept general to allow for wide applicability, though it could also be chosen to incorporate known object information for a specific tracking application.
+The distributions F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ represent object appearance, the appearance prior, and object movement, respectively.
+Our specification is kept general to allow for wide applicability, though one could choose to incorporate known object appearance or motion information for a specific tracking application in future studies.
 
@@
 
-%The distribution families assumed to generate each $\bold{x}$ must be justified.
-%Both the multivariate normal and multinomial distributions were chosen because they are sufficiently simple (and well studied) to allow for tractable inference and sufficiently flexible to provide a reasonable approximation to the data gained during extraction.
-The multivariate normal distribution over the spatial features $\bold{x}^{s}$ can be intuitively thought to represent the shape of each object as an oval; likewise, data extracted from each moving object tend to cluster around the object's centroid in an oval shape.
-%---as noisy extraction procedures cause some smoothing of edges and corners, producing blob-like shapes even when objects are not particularly round---and centered on a given object.
-% Furthermore, this model is justified as the maximum likelihood parameter estimate of a normal distribution corresponds to the least squares fit of data relative to the normal distribution mean;
-Modeling the color features $\bold{x}^{c}$ as draws from a multinomial distribution is justified since we have observed that the pixels associated with a distinct object tend to have similar color count vectors.
+The multivariate normal distribution over the spatial features $\bold{x}^{s}$ models the position and spatial extent of an object.
+It can intuitively be thought to represent the shape of each object as an oval.
+We model the color features $\bold{x}^{c}$ as draws from a multinomial distribution.
+Incorporating this distribution into our cluster likelihood allows us to exploit our observation that the pixels associated with distinct objects tend to have similar color count vectors.
 
@@
 
-The multivariate normal-multinomial GPUDDPM implemented in this study has a number of parameters, which are used to specify the object appearance prior distribution, transition kernel, and Generalized Polya Urn distributions (C and D).\newpage %remove if formatting changes
-\hspace{-5mm}The object appearance prior parameters include:
+The multivariate normal-multinomial GPUDDPM implemented in this study has a number of parameters, which are used to specify the object appearance prior distribution, transition kernel, and Generalized Polya Urn distributions (C and D).
+%\newpage %remove if formatting changes
+%\hspace{-5mm}
+The object appearance prior parameters include:
 \begin{center}
@@
 \hline
-$M$  &  Number of auxiliary variables. A larger number will produce lower variation in position (better for slower objects) and a smaller number will produce higher variation in position. \\
+$M$  &  Number of auxiliary variables. A larger number will produce a smoother object path.\\
 \hline
@@
 \hline
-$\alpha$  &  The concentration parameter for the Dirichlet Process. A higher value will increase the tendency for new objects to be detected.\\ \hline
+$\alpha$  &  The concentration parameter for the Dirichlet process. A higher value will increase the tendency for new objects to be detected.\\ \hline
 $\rho$  &  The deletion parameter. A higher value will give objects an increased tendency to die off. \\
@@
 
-The parameters chosen for use in experiments (Section~\ref{sec:experiments}) were accepted based on the fact that they yielded reasonable results.
+%The parameters chosen for use in experiments (Section~\ref{sec:experiments}) were accepted based on the fact that they yielded reasonable results.
 Additionally, a sensitivity analysis was carried out for the concentration and deletion parameters ($\alpha$ and $\rho$), where inference was performed for range of both values (Section~\ref{sec:sensitivityanalysis}).
@@
 
-Bayesian inference algorithms, which infer distributions over latent model parameters, are used to achieve detection and tracking results.
+Bayesian inference, which infers distributions over latent model parameters, is used to achieve detection and tracking results.
 Previously developed inference strategies can be applied to the generative model defined in Section~\ref{sec:gpuddpm} and Section~\ref{sec:modelspecification}.
@@
 
-\subsubsection{Deletion Variable Formulation of the GPUDDPM}
+\subsubsection{Deletion Variable Formulation}
 \label{sec:deletionvariableformulation}
@@
 
-We define the set of variables, S, to consist of all observations, assignment variables, deletion variables, cluster parameters (for all active clusters), and auxiliary variables (for all active clusters).
-We can write
-\begin{equation}
-\text{S} = \{ \boldsymbol{x}_{1:T}, \bold{c}_{1:T}, \bold{d}_{1:T}, \Theta_{1:T}, \bold{z}_{1:T} \}
-\end{equation}
-where $\boldsymbol{x}_{t} = \bold{x}_{1:N_{t},t}$, $\bold{c}_{t} = c_{1:N_{t},t}$, $\bold{d}_{t} = d_{1:N_{t},t}, \Theta_{t} = \theta_{1:K_{t}, t}$, and $\bold{z}_{t} = z_{1:K_{t},t,1:M}$.
+To allow for easier notation, we define $\boldsymbol{x}_{t} = \bold{x}_{1:N_{t},t}$, $\bold{c}_{t} = c_{1:N_{t},t}$, $\bold{d}_{t} = d_{1:N_{t},t}, \Theta_{t} = \theta_{1:K_{t}, t}$, and $\bold{z}_{t} = z_{1:K_{t},t,1:M}$.
 At each time $t$, the sampler moves sequentially through the $N_{t}$ observations, sampling the assignment $c_{i,t}$ and the deletion variable $d_{i,t}$ for each.
@@
 
+% from older version where defined a set S
+
+% We define the set of variables, S, to consist of all observations, assignment variables, deletion variables, cluster parameters (for all active clusters), and auxiliary variables (for all active clusters).
+We can write
+% \begin{equation}
+% \text{S} = \{ \boldsymbol{x}_{1:T}, \bold{c}_{1:T}, \bold{d}_{1:T}, \Theta_{1:T}, \bold{z}_{1:T} \}
+% \end{equation}
+% where
+
 % mcmc: sampling assignment variables
@@
 The possible values that the $c_{i,t}$ may take on are $k \in \{ 1 , \ldots ,  K_{t}^{'}+1 \}$, where $K_{t}^{'}$ denotes the number clusters with a non-zero size at any time $t' \in \{ t, \ldots, d_{i,t} \}$, and $K_{t}'+1$ denotes a newly sampled cluster.
-The probability that assignment $c_{i,t}$ equals a potential cluster $k$, given all other members of S (denoted S$'$), is given by
+The probability that assignment $c_{i,t}$ equals a potential cluster $k$, given all other variables in the model (which we denote as ``$\ldots$"), is given by
 \begin{equation}
@@
 \begin{split}
-p(c_{i,t} = k | \text{S}') & \propto
+p(c_{i,t} = k | \ldots ) & \propto
 \prod_{i'=i}^{N_{t}}  \text{C}(c_{i',t} | \bold{m}_{t}, \alpha) \\
@@
 where the C is given by \eqref{C_distro}, and the cluster sizes $\bold{m}_{t'}$ are calculated under the assumption that $c_{i,t} = k$.
-Note that the above integral can be determined analytically to give
+Note that the above integral has an analytic solution for our specific model
 \begin{equation}
@@
 \begin{split}
-P(\theta_{k,t} | \text{S}') = & \hspace{1mm} P(\bold{x}_{i,t} | \theta_{k,t}) P(z_{k,t+1,1:M} | \theta_{k,t})
+P(\theta_{k,t} | \ldots ) = & \hspace{1mm} P(\bold{x}_{i,t} | \theta_{k,t}) P(z_{k,t+1,1:M} | \theta_{k,t})
 \\ & \times P(\theta_{k,t} | z_{k,t,1:M}) \\
@@
 \label{stationary_pdf}
-P(z_{k,t,m} | \text{S}') \propto  P(z_{k,t,m} | \theta_{k,t-1}) P(\theta_{k,t} | z_{k,t,1:M})
-\end{equation}
-Samples cannot easily be drawn from this distribution.
+P(z_{k,t,m} | \ldots) \propto  P(z_{k,t,m} | \theta_{k,t-1}) P(\theta_{k,t} | z_{k,t,1:M})
+\end{equation}
+%fw says to remove:
+%
+%Samples cannot easily be drawn from this distribution.
 We therefore use the MH algorithm to sample auxiliary variables from the posterior distribution (for each living cluster).
 The MH algorithm involves constructing a markov chain with the sought posterior as its stationary distribution, and sampling sequentially from this constructed chain to get an approximate posterior sample.
+%
+%
 % The MH algorithm is outlined in Algorithm 1.%Algorithm~\ref{alg:MH}. \willie{am i citing wrong?}
@@
 % \end{algorithm}
-In short, we sample a new value for all $M$ auxiliary variables, denoted $z_{k,t,m}^{*}$, according to the following proposal distribution
+%
+We sample a new value for all $M$ auxiliary variables, denoted $z_{k,t,m}^{*}$, using MH, with the proposal distribution
 \begin{align}
@@
 \end{align}
-compute the standard MH acceptance ratio, which in this case simplifies to 
+and compute the standard MH acceptance ratio, which in this case simplifies to 
 \begin{equation}
@@
 \end{equation}
-and choose $z_{k,t,m}^{*}$ as the next sample in the constructed sequence if $\alpha < \text{min}(1, r_{accept})$, where $\alpha$ is a sample from a $\text{Uniform}(0,1)$ distribution.
+% and choose $z_{k,t,m}^{*}$ as the next sample in the constructed sequence if $\alpha < \text{min}(1, r_{accept})$, where $\alpha$ is a sample from a $\text{Uniform}(0,1)$ distribution.
 
@@
 Sampling deletion variables could be performed in a manner similar to how we sample the assignment variables in Section~\ref{sec:sample_assignments}, but this may be computationally expensive due to the large number of possible deletion times.
-To remedy this, the MH algorithm may again be used to generate samples from the posterior distribution over possible deletion times, where the prior distribution for the deletion times can be used to propose new samples $d_{i,t}^{*}$.
-As the ``lifetime'' $l_{i,t}$ of each assignment is geometrically distributed, we can propose each new deletion time sample in our constructed chain with
+To remedy this, the MH algorithm may again be used to generate samples $d_{i,t}^{*}$ from the posterior distribution over possible deletion times, where we use the proposal distribution
 \begin{align}
@@
 \end{align}
-and accept or reject this sample using the process described in Section~\ref{sec:sample_aux_vars}.
-%%%%%
-% need to include the "standard MH acceptance ratio"
-%%%%%
+where $l_{i,t}$ denotes the geometrically distributed ``lifetime'', and we accept or reject this sample using the process described in Section~\ref{sec:sample_aux_vars}.
 
@@
 
-The probability of assignments given current cluster sizes, cluster parameters, and the Dirichlet Process concentration parameter $\alpha$ can be written as
+The probability of assignments given current cluster sizes, cluster parameters, and the Dirichlet process concentration parameter $\alpha$ can be written as
 \begin{align}
@@
 \end{equation}
-where samples can be drawn from $P(\theta_{k,t} | \bold{x}_{1:N,t})$ using the bayesian updates found in \eqref{bayesian_update_1}, \eqref{bayesian_update_2}, \eqref{bayesian_update_3}, and \eqref{bayesian_update_4}, where the $\bold{x}_{1:N,t}$ are taken to be the observations.
+where samples can be drawn from $P(\theta_{k,t} | \bold{x}_{1:N,t})$ using the Bayesian updates found in \eqref{bayesian_update_1}, \eqref{bayesian_update_2}, \eqref{bayesian_update_3}, and \eqref{bayesian_update_4}, where the $\bold{x}_{1:N,t}$ are taken to be the observations.
 
@@
 \end{equation}
-where samples can be drawn from $P(\theta_{k,t} | \theta_{k,t-1} \bold{x}_{1:N,t})$ using the bayesian updates found in \eqref{bayesian_update_1}, \eqref{bayesian_update_2}, \eqref{bayesian_update_3}, and \eqref{bayesian_update_4}, where both the $\bold{x}_{1:N,t}$ and auxiliary variables $z_{k,t,1:M}$ are taken to be the observations.
+where samples can be drawn from $P(\theta_{k,t} | \theta_{k,t-1} \bold{x}_{1:N,t})$ using the Bayesian updates found in \eqref{bayesian_update_1}, \eqref{bayesian_update_2}, \eqref{bayesian_update_3}, and \eqref{bayesian_update_4}, where both the $\bold{x}_{1:N,t}$ and auxiliary variables $z_{k,t,1:M}$ are taken to be the observations.
 
@@
 
-The nature of object detection and tracking results may vary depending on the application; in this study, due to the performance metrics we wish to compute (see Section~\ref{sec:performanceevaluationmetrics}), we desire the approximate spatial region for each moving object, for each frame that this object can be seen within the video.
+%The nature of object detection and tracking results may vary depending on the application; in this study, due to the performance metrics we wish to compute (see Section~\ref{sec:performanceevaluationmetrics}), we desire the approximate spatial region for each moving object, for each frame that this object can be seen within the video.
 We choose to make our detection and tracking results depend completely on the inferred parameters of the multivariate normal factor of each cluster.
@@
 In particular, the mean parameter is taken to be the centroid of an object, and a 2-dimensional oval centered on the mean that contains a specified percentage of the normal distribution mass (where we refer to the specified percentage as the confidence value) is taken to be the spatial region of the object.
-Furthermore, as both inference procedures described in this paper involve sampling, and each sample represents a result, we must determine a way to narrow these samples down to get a single result.
+We report the maximum a posteriori (MAP) sample as our result.
+%Furthermore, as both inference procedures described in this paper involve sampling, and each sample represents a result, we must determine a way to narrow these samples down to get a single result.
 We have found that calculating a value proportional to joint probability of the model for each sample and choosing the sample that yields the highest value gives reasonable results.
@@
 
-Performance evaluation metrics, which intend to provide a standardized way of quantifying the success of a detection and tracking procedure on a given video, have started to become well studied and consistently used in the past four years.
-The metrics developed in \cite{kasturi_2008} and used in \cite{ellis_2010, taj_2007, lee_2009} have become well established metrics for evaluating the performance of object detection and tracking in videos and have been adopted by the Video Analysis and Content Extraction (VACE) program and the Classification of Events, Activities, and Relationships (CLEAR) consortium, two large-scale efforts concerned with video tracking and interaction analysis.
-These metrics are used to quantify the experimental results in this study.
+Performance evaluation metrics, which intend to provide a standardized way of quantifying the success of a detection and tracking procedure on a given video, have started to become consistently used in the past four years.
+The metrics presented in \cite{kasturi_2008} and used in \cite{ellis_2010, taj_2007, lee_2009} have become well established metrics for evaluating the performance of object detection and tracking in videos and have been adopted by the Video Analysis and Content Extraction (VACE) program and the Classification of Events, Activities, and Relationships (CLEAR) consortium, two large-scale efforts concerned with video tracking and interaction analysis.
+The two metrics used to quantify the experimental results in this study are known as the Sequence Frame Detection Accuracy (SFDA) and Average Tracking Accuracy (ATA).
+Details on how these metrics are defined and computed are given in Appendix~\ref{sec:performancemetricdetails}.
 
@@
 
-\subsubsection{Mapping Ground-Truth to Output}
-\label{sec:mappingtoground}
-
-The problem of finding a mapping between a video's ground-truth tracks and an algorithm's output tracks is nontrivial, though necesary to solve, in order to compute the performance evaluation metrics used in this study.
-In short, the typical solution to this problem involves first specifying a performance metric and then choosing the mapping from ground-truth tracks to output tracks which yields the most favorable performance metric value.
-This process is described in detail by Kasturi et al.
-\cite{kasturi_2008}; we follow the method outlined in this paper to find an optimal mapping.
-Similiar to descriptions in \cite{kasturi_2008}, we implement the Hungarian algorithm \cite{munkres_1957} as a polynomial-time ($O(n^{3})$) solution to the problem of optimally mapping two sets of tracks once the similarity between any two tracks given some specified metric is established.
-Additionally, the method employed in \cite{kasturi_2008} allows erroneous and undetected tracks to be left unmapped, which is both desired and necessary in the case where there is a different number of ground-truth and output tracks.
-Note that once a mapping from a collection of ground-truth tracks to a collection of result tracks has been established, one can determine which result tracks are false positives (the result tracks to which no ground-truth track is assigned) and which ground-truth tracks are true negatives (the ground-truth tracks that are not assigned to a result track).
-The numbers of tracks displaying both of these failures are factors in the performance metrics used in this study.
-
-\subsubsection{SFDA and ATA}
-
-The two metrics used to quantify performance in this study are known as the the Sequence Frame Detection Accuracy (SFDA) and the Average Tracking Accuracy (ATA).
-These metrics were developed during VACE Phase II to provide a single, comprehensive metric to describe detection, and one to describe tracking.
-The following are used in the definitions of the performance metrics:
-\begin{itemize}
-\renewcommand{\labelitemi}{$\bullet$}
-\item $G_{i}$ denotes the spatiotemporal region occupied by the $i$th ground-truth object in a video, and $G_{i}^{(t)}$ denotes the region occupied by the $i$th ground-truth object in frame $t$.
-\vspace{2mm}
-\item $D_{i}$ denotes the spatiotemporal region occupied by the $i$th detected object in a video, and $D_{i}^{(t)}$ denotes the region occupied by the $i$th detected object in frame $t$.
-\vspace{2mm}
-\item $N_{G}$ denotes the total number of unique ground-truth objects in a video, and $N_{G}^{(t)}$ denotes the number of unique ground-truth objects present at frame $t$.
-\vspace{2mm}
-\item $N_{D}$ denotes the total number of unique detected objects in a video, and $N_{D}^{(t)}$ denotes the number of unique detected objects present at frame $t$.
-\vspace{2mm}
-\item $N_{\text{frames}}$ denotes the total number of frames in a video, and $N_{\text{frames}}^{(i)}$ denotes the number of frames in which an object $i$ (which can be a ground-truth or detected object, depending on the context) is present in a video.
-\vspace{2mm}
-\item $N_{\text{mapped}}$ denotes the number of mapped ground-truth/detect pairs in a video, and $N_{\text{mapped}}^{(t)}$ denotes the number of mapped ground-truth/detect pairs present at frame $t$.
-\vspace{1mm}
-\end{itemize}
-
-The SFDA metric quantifies the performance of an object detection algorithm as a function of the number of correct detects, false positive detects, missed (true negative) detects, and spatial allignment of detects relative to the ground-truth.
-The SFDA is calculated by computing the Frame Detection Accuracy at frame $t$ ($\text{FDA}^{(t)}$) for each frame in a video sequence.
-The FDA provides a measure of the allignment between ground-truth and detected objects in a given frame via the overlap ratio of a ground-truth/detect pair, defined to be the ratio of the intersection of ground-truth and detect regions to the union of ground-truth and detect regions.
-Formally, we can write
-\begin{equation}
-\text{FDA}^{(t)} = \frac{\text{Overlap Ratio}}{\left(\frac{N_{G}^{(t)} + N_{D}^{(t)}}{2}\right)}
-\end{equation}
-where
-\begin{equation}
-\text{Overlap Ratio} = \sum_{i = 1}^{N_{\text{mapped}}^{(t)}} \frac{\left|G_{i}^{(t)} \cap D_{i}^{(t)}\right|}{\left|G_{i}^{(t)} \cup D_{i}^{(t)}\right|}
-\end{equation}
-The term $N_{\text{mapped}}^{(t)}$ refers to an optimal mapping between ground-truth and detects at frame $t$ as specified in section \ref{sec:mappingtoground} using the $\text{FDA}^{(t)}$ as the relevant metric.
-Given the $\text{FDA}^{(t)}$ at each frame, the SFDA can be computed; this metric may be viewed as the average FDA over all frames of a video sequence.
-We define
-\begin{equation}
-\text{SFDA} = \frac{\sum_{t=1}^{N_{\text{frames}}} \text{FDA}^{(t)}}{\sum_{t=1}^{N_{\text{frames}}} \exists \left( N_{G}^{(t)} \vee N_{D}^{(t)} \right)}
-\end{equation}
-where $\exists \left( N_{G}^{(t)} \vee N_{D}^{(t)} \right)$ yields a 1 if either a detected or ground-truth object is present in frame $t$ and a 0 otherwise.
-
-The ATA metric quantifies the performance of an object tracking algorithm as a function of the spatial overlap of a mapped set of sequences of detected object positions to a set of sequences of groundtruth object positions.
-The ATA is calculated by first computing the Sequence Track Detection Accuracy (STDA), which can be viewed as a tracking performance measure unnormalied in terms of the number of objects.
-We can write the STDA as
-\begin{equation}
-\text{STDA} = \sum_{i=1}^{N_{\text{mapped}}} \frac{\sum_{t=1}^{N_{\text{frames}}} \left( \frac{\left|G_{i}^{(t)} \cap D_{i}^{(t)}\right|}{\left|G_{i}^{(t)} \cup D_{i}^{(t)}\right|}  \right) }{ N_{(G_{i} \cup D_{i} \neq \emptyset)} }
-\end{equation}
-where $N_{\text{mapped}}$ refers to an optimal mapping between ground-truth and detected objects as specified in section \ref{sec:mappingtoground} using the STDA as the relevant metric, and $N_{(G_{i} \cup D_{i} \neq \emptyset)}$ denotes the number of frames in which a given tracked object, the ground truth object to which it is mapped, or both, are present.
-
-Given the STDA for a video sequence, the ATA can be computed by the formula
-\begin{equation}
-\text{ATA} = \frac{\text{STDA}}{\left( \frac{N_{G} + N_{D}}{2} \right)}
-\end{equation}
-
 The ground-truth authored by the ViPER tool took the form of bounding boxes denoting the spatial postion of each object at each time step.
 Consequentially, to find the spatial overlap between results and ground-truth, which is intrinsic to both metrics, a rectangular bounding box was needed per object per time step from the results of the algorithm.
-We took the maximal and minimal axially aligned values of the oval produced by our algorithm (as described in Section~\ref{sec:inferencetotrackingresults}) to be the sides of a representative bounding box for a given object at a given frame.
+We took the maximal and minimal axially aligned values of the oval inferred by our algorithm (as described in Section~\ref{sec:inferencetotrackingresults}) to be the sides of a representative bounding box for a given object at a given frame.
 
@@
 Specifically, we specify a square, $L$ pixels in length, centered on $(x_{1}, x_{2})$, that contains a set of pixels surrounding $\bold{x}$ in frame $t$.
-We choose a scalar color value that is able to be computed for each pixel; this value could, for example, be some function of the red-green-blue (r-g-b) or hue-saturation-value (h-s-v) characteristic of a pixel.
-The color value for each of the selected pixels surrounding $\bold{x}$ is recorded.
-Afterwards, the set of possible color values (ie the range of color values to which a pixel may be assigned) is partitioned into $V$ bins, and the number of pixels with a color value lying in each of the bins yields the $V$ dimensional vector of color counts.
+We choose to capture the hue for each pixel; consequently, the hue value for each of the selected pixels surrounding $\bold{x}$ is recorded.
+Afterwards, the set of possible color values (i.e.
+the range of color values to which a pixel may be assigned) is partitioned into $V$ bins, and the number of pixels with a color value lying in each of the bins yields the $V$ dimensional vector of color counts.
 
@@
 
-Synthetic videos provide controlled scenarios in which aspects of the technique presented in this paper can be tested.
 Each of the following synthetic videos consists of a sequence of 200 images (each of size $500$ x $500$ pixels) containing a number of smaller colored squares of different (and potentially time-varying) sizes moving at varied speeds and trajectories over a black background.
@@
 The chosen values were $\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right)$, and $q_{0} = (5, \ldots, 5)$.
-Inference was carried out using the MCMC algorithm (Section~\ref{sec:MCMC}); the posterior sample taken as the result correctly tracked both colored squares through occlusion in both videos, and is shown in Figure~\ref{fig:synth_one_plot}.
+Inference was carried out using the MCMC algorithm (Section~\ref{sec:MCMC}); the MAP sample taken as the result correctly tracked both colored squares through occlusion in both videos, and is shown in Figure~\ref{fig:synth_one_plot}.
 
@@
   \subfloat[]{\label{fig:pets2001_results} \includegraphics[width=0.48\textwidth]{../img/pets2001_back.pdf}} \\
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f1}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f4}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f9}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f11}} \\           
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f1.pdf}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f3.pdf}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f5.pdf}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f8.pdf}} \\
-  \caption{PETS2000 dataset results (a) and PETS2001 dataset results (b).
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f1_w}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f4_w}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f9_w}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f11_w}} \\           
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f1_w.pdf}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f3_w.pdf}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f5_w.pdf}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f8_w.pdf}} \\
+  \caption{Results from the PETS2000 (a) and PETS2001 (b) dataset.
 Both plots show a sample from the posterior distribution of the state, where the vertical axis denotes time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.
@@
 These videos have been primarily produced for surveillance-related workshops---notably, for the International Workshop on Performance Evaluation of Tracking and Surveillance (PETS)--which provide researchers with video datasets and algorithmic goals on which to focus.
-Three commonly used benchmark videos from PETS workshops---one used in PETS2000, one in PETS2001, and one used both in PETS2009 and PETS2010---were chosen to demonstrate the efficacy of the method presented in this study.
+Three commonly used benchmark videos from PETS workshops---one used in PETS2000, one in PETS2001, and one used both in PETS2009 and PETS2010---were chosen to demonstrate the method presented in this study.
 The performance metrics and benchmark datasets allow the methods developed in this paper to be quantitatively compared against other detection and tracking algorithms.
@@
   \subfloat[]{\includegraphics[width=0.53\textwidth]{../img/pets2009full_pm_b_v2.pdf}} \\
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2009_f3.pdf}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2009_f6.pdf}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2009_f13.pdf}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2009_f14.pdf}}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2009_f3_w.pdf}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2009_f6_w.pdf}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2009_f13_w.pdf}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2009_f14_w.pdf}}
   \caption{ (a) Results for the PETS2009/2010 dataset, showing a sample from the posterior distribution of the state for frames 1-50, where the vertical axis denotes time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.
@@
 
-In \cite{ellis_2010}, performance metrics were computed for a number of studies that carried out object detection and tracking for the PETS2009/2010 dataset.
-As this dataset consists solely of humans, many of these algorithms were developed for the specific purpose of people tracking (i.e.
-not for general detection and tracking of arbitrary objects); in consequence, many of these studies involve appearance models specifically developed for humans, sometimes making use of their orientation in this specific dataset, and motion models based on assumptions about human motion.
-We compare SFDA and ATA results of our strategy with these methods to show that our general method can yield comparable results even when compared with object-specific trackers.
+\subsubsection{Comparison with Other Methods}
+\label{sec:comparisonwithothermethods}
+
+In \cite{ellis_2010}, performance metrics (including the SFDA and ATA) were computed for a number of studies that carried out object detection and tracking for the PETS20-\\09/2010 dataset.
+As this dataset consists solely of humans, all ten of the algorithms presented for comparison were developed for the specific purpose of people tracking (i.e.
+not for general detection and tracking of arbitrary objects).
+In consequence, many of these studies use externally developed (and trained) state-of-the-art human detectors, exploit the orientation of the humans in this specific dataset, or apply motion models based on assumptions about human motion.
+We compare SFDA and ATA results of our strategy with these methods to show that our arbitrary object framework can yield comparable results even when compared with object-specific trackers.
 Table~\ref{benchmark_results_table} shows performance metric results for comparison (data published with permission from the authors of \cite{ellis_2010}).
@@
   \textsc{Method Name} & \textsc{SFDA}  & \textsc{ATA} \\ \hline \hline
-  GPUDDPM & $ \bold{0.5079} $ & $\bold{0.2996}$ \\ \hline
-  \cite{arsic2009multi} & $ 0.177768 $ & $0.016111$ \\ \hline
-  \cite{berclaz2009multiple} & $ 0.482081 $ & $0.145195$ \\ \hline
-  \cite{conte2010performance} & $ 0.52966  $ & $0.058003$ \\ \hline
-  \cite{bolme2009simple} & $ 0.413812 $ & NA \\ \hline
-  \cite{bolme2009simple} & $ 0.338979 $ & NA \\ \hline
-  \cite{bolme2009simple} & $ 0.338979 $ & NA \\ \hline
-  \cite{breitenstein2009markovian} & $ \emph{0.567265} $ & $0.30044$ \\ \hline
-  \cite{ge2009evaluation} & $ 0.383235 $ & $0.040118$ \\ \hline
-  \cite{alahi2009sparsity} & $ 0.430043 $ & $0.041053$ \\ \hline
-  \cite{alahi2009sparsity} & $ 0.419141 $ & $0.045609$ \\ \hline
-  Anonymous & $ 0.52311  $ & $0.246275$ \\ \hline
-  \cite{yang2009probabilistic} & $ 0.552439 $ & $\emph{0.448237}$ \\
+  Breitenstein \cite{breitenstein2009markovian} & $ 0.567265 $ & $0.30044$ \\ \hline
+  Yang \cite{yang2009probabilistic} & $ 0.552439 $ & $0.448237$ \\ \hline
+  Conte \cite{conte2010performance} & $ 0.52966  $ & $0.058003$ \\ \hline
+  \textbf{GPUDDPM} & $ \bold{0.5079} $ & $\bold{0.2996}$ \\ \hline
+  Berclaz \cite{berclaz2009multiple} & $ 0.482081 $ & $0.145195$ \\ \hline
+  Alahi 1 \cite{alahi2009sparsity} & $ 0.430043 $ & $0.041053$ \\ \hline
+  Alahi 2 \cite{alahi2009sparsity} & $ 0.419141 $ & $0.045609$ \\ \hline
+  Bolme 1 \cite{bolme2009simple} & $ 0.413812 $ & NA \\ \hline
+  Ge \cite{ge2009evaluation} & $ 0.383235 $ & $0.040118$ \\ \hline
+  Bolme 2 \cite{bolme2009simple} & $ 0.338979 $ & NA \\ \hline
+  Arsic \cite{arsic2009multi} & $ 0.177768 $ & $0.016111$ \\
+  % Anonymous & $ 0.52311  $ & $0.246275$ \\ \hline
   \hline
 \end{tabular}
-\caption{SFDA and ATA performance metric results shown for 13 algorithms on the PETS2009/2010 benchmark dataset.
-Results for our method (GPUDDPM) are in bold.
-The largest ATA and SFDA values are italicized.
-Data given by the authors of \cite{ellis_2010}.}
+\caption{SFDA and ATA performance metric results are shown for our method (in bold) and for ten other algorithms on the PETS2009/2010 benchmark dataset.
+Results are listed in descending order of the SFDA value.
+Results are provided by the authors of \cite{ellis_2010}.}
 \label{benchmark_results_table}
@@
 
-Figure~\ref{fig:sens_surf} shows a surface plot of the sensitivity analysis results.
+Figure~\ref{fig:sens_surf} shows a heatmap and contour plot of the sensitivity analysis results.
 From this figure, we can see that the SFDA and ATA achieve their maximal values at different $\alpha$ and $\rho$ parameters, though both achieve a reasonably optimal value at the intermediate parameter values $\alpha = 10$ and $\rho = 0.85$.
@@
   \centering             
-  \subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2009full_sens_sfda_heat.pdf}}
+  \subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2009full_sens_sfda_contour5.pdf}}
   \hspace{6mm}
-  \subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2009full_sens_ata_heat.pdf}}
-  \caption{Heatmap showing sensitivity of performance metrics SFDA and ATA to variation of parameters $\alpha$ and $\rho$.}
+  \subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2009full_sens_ata_contour5.pdf}}
+  \caption{Heatmap and contour plot showing the sensitivity of performance metrics SFDA and ATA to variation of parameters $\alpha$ and $\rho$.}
   \label{fig:sens_surf}
@@
 
-We have presented a new technique for the unsupervised detection and tracking of arbitrary objects in videos.
+We have presented a new framework for the unsupervised detection and tracking of arbitrary objects in vid-\\eos.
 The primary intention of this technique is to reduce the need for detection or localization methods tailored to specific object types and serve as a general framework applicable to videos with varied objects, backgrounds, and film qualities.
@@
 
-\section{-  Appendix}
+\section{-  Appendix: Model Background}
 \label{sec:modelbackground}
 
-Dirichlet Process Mixture (DPM) models (Section~\ref{sec:dpmixture}) fall under the heading of Bayesian nonparametric (BNP) models.
+Dirichlet process mixture (DPM) models (Section~\ref{sec:dpmixture}) fall under the heading of Bayesian nonparametric (BNP) models.
 These models have been widely used in the past decade to perform nonparametric density estimation and cluster analysis.
-Since we have time series data and wish to find spatiotemporal clusters, we are interested in using a closely related class of models known as Dependent Dirichlet Process Mixture (DDPM) models (Section~\ref{sec:ddpmixture}).
+Since we have time series data and wish to find spatiotemporal clusters, we are interested in using a closely related class of models known as dependent Dirichlet process mixture (DDPM) models (Section~\ref{sec:ddpmixture}).
 These models are particularly useful for estimating the number of latent classes (clusters) in time dependent data.
@@
 This section provides background information that allows us to introduce DDPM models and describes the motivation behind our use of these models.
-This background is helpful for the definition of the particular DDPM chosen, the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM) model, given in Section~\ref{sec:gpuddpm}.
+This background is helpful for the definition of the particular DDPM chosen, the Generalized Polya Urn dependent Dirichlet process mixture (GPUDDPM) model, given in Section~\ref{sec:gpuddpm}.
 
@@
 % The time dependent video extraction data contains a collection of clusters, each representing a video object, which might change in number as time progresses (which is to say, clusters may be created, or be ``born'', and may disappear, or ``die'', at some intermediate time step), since objects can enter and exit a scene.
-To handle these challenges, the specific DDPM chosen to model extraction data in this work is known as the Generalized Polya Urn Dependent Dirichlet Process Mixture (GPUDDPM), introduced by \cite{caron_2007} in 2006.
+To handle these challenges, the specific DDPM chosen to model extraction data in this work is known as the Generalized Polya Urn dependent Dirichlet process mixture (GPUDDPM), introduced by \cite{caron_2007} in 2006.
 
@@
 %         \center{\includegraphics[width=80mm]{../img/gpuddp_gm_1.pdf}}
-%         \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Polya Urn Dependent Dirichlet Process Mixture.
+%         \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Polya Urn dependent Dirichlet process mixture.
 Note that all observations at time $t$, $\bold{x}_{1:N,t}$, and their associated assignments $c_{1:N,t}$ are denoted respectively as $\bold{x}_{t}$ and $c_{t}$ (this also holds for time step $t-1$).}
@@
 
+\section{-  Appendix: Performance Metric Details}
+\label{sec:performancemetricdetails}
+This section provides details on the definition and calculation of the performance evaluation metrics, SFDA and ATA, used to quantify detection and tracking results in this study.
+
+\subsection{Mapping Ground-Truth to Output}
+\label{sec:mappingtoground}
+
+The problem of finding a mapping between a video's ground-truth tracks and an algorithm's output tracks is nontrivial, though necesary to solve, in order to compute the performance evaluation metrics used in this study.
+In short, the typical solution to this problem involves first specifying a performance metric and then choosing the mapping from ground-truth tracks to output tracks which yields the most favorable performance metric value.
+This process is described in detail by Kasturi et al.
+\cite{kasturi_2008}; we follow the method outlined in this paper to find an optimal mapping.
+Similiar to descriptions in \cite{kasturi_2008}, we implement the Hungarian algorithm \cite{munkres_1957} as a polynomial-time ($O(n^{3})$) solution to the problem of optimally mapping two sets of tracks once the similarity between any two tracks given some specified metric is established.
+Additionally, the method employed in \cite{kasturi_2008} allows erroneous and undetected tracks to be left unmapped, which is both desired and necessary in the case where there is a different number of ground-truth and output tracks.
+Note that once a mapping from a collection of ground-truth tracks to a collection of result tracks has been established, one can determine which result tracks are false positives (the result tracks to which no ground-truth track is assigned) and which ground-truth tracks are true negatives (the ground-truth tracks that are not assigned to a result track).
+The numbers of tracks displaying both of these failures are factors in the performance metrics used in this study.
+
+\subsection{SFDA and ATA}
+
+The two metrics used to quantify performance in this study are known as the the Sequence Frame Detection Accuracy (SFDA) and the Average Tracking Accuracy (ATA).
+These metrics were developed during VACE Phase II to provide a single, comprehensive metric to describe detection, and one to describe tracking.
+The following are used in the definitions of the performance metrics:
+\begin{itemize}
+\renewcommand{\labelitemi}{$\bullet$}
+\item $G_{i}$ denotes the spatiotemporal region occupied by the $i$th ground-truth object in a video, and $G_{i}^{(t)}$ denotes the region occupied by the $i$th ground-truth object in frame $t$.
+\vspace{2mm}
+\item $D_{i}$ denotes the spatiotemporal region occupied by the $i$th detected object in a video, and $D_{i}^{(t)}$ denotes the region occupied by the $i$th detected object in frame $t$.
+\vspace{2mm}
+\item $N_{G}$ denotes the total number of unique ground-truth objects in a video, and $N_{G}^{(t)}$ denotes the number of unique ground-truth objects present at frame $t$.
+\vspace{2mm}
+\item $N_{D}$ denotes the total number of unique detected objects in a video, and $N_{D}^{(t)}$ denotes the number of unique detected objects present at frame $t$.
+\vspace{2mm}
+\item $N_{\text{frames}}$ denotes the total number of frames in a video, and $N_{\text{frames}}^{(i)}$ denotes the number of frames in which an object $i$ (which can be a ground-truth or detected object, depending on the context) is present in a video.
+\vspace{2mm}
+\item $N_{\text{mapped}}$ denotes the number of mapped ground-truth/detect pairs in a video, and $N_{\text{mapped}}^{(t)}$ denotes the number of mapped ground-truth/detect pairs present at frame $t$.
+\vspace{1mm}
+\end{itemize}
+
+The SFDA metric quantifies the performance of an object detection algorithm as a function of the number of correct detects, false positive detects, missed (true negative) detects, and spatial allignment of detects relative to the ground-truth.
+The SFDA is calculated by computing the Frame Detection Accuracy at frame $t$ ($\text{FDA}^{(t)}$) for each frame in a video sequence.
+The FDA provides a measure of the allignment between ground-truth and detected objects in a given frame via the overlap ratio of a ground-truth/detect pair, defined to be the ratio of the intersection of ground-truth and detect regions to the union of ground-truth and detect regions.
+Formally, we can write
+\begin{equation}
+\text{FDA}^{(t)} = \frac{\text{Overlap Ratio}}{\left(\frac{N_{G}^{(t)} + N_{D}^{(t)}}{2}\right)}
+\end{equation}
+where
+\begin{equation}
+\text{Overlap Ratio} = \sum_{i = 1}^{N_{\text{mapped}}^{(t)}} \frac{\left|G_{i}^{(t)} \cap D_{i}^{(t)}\right|}{\left|G_{i}^{(t)} \cup D_{i}^{(t)}\right|}
+\end{equation}
+The term $N_{\text{mapped}}^{(t)}$ refers to an optimal mapping between ground-truth and detects at frame $t$ as specified in section \ref{sec:mappingtoground} using the $\text{FDA}^{(t)}$ as the relevant metric.
+Given the $\text{FDA}^{(t)}$ at each frame, the SFDA can be computed; this metric may be viewed as the average FDA over all frames of a video sequence.
+We define
+\begin{equation}
+\text{SFDA} = \frac{\sum_{t=1}^{N_{\text{frames}}} \text{FDA}^{(t)}}{\sum_{t=1}^{N_{\text{frames}}} \exists \left( N_{G}^{(t)} \vee N_{D}^{(t)} \right)}
+\end{equation}
+where $\exists \left( N_{G}^{(t)} \vee N_{D}^{(t)} \right)$ yields a 1 if either a detected or ground-truth object is present in frame $t$ and a 0 otherwise.
+
+The ATA metric quantifies the performance of an object tracking algorithm as a function of the spatial overlap of a mapped set of sequences of detected object positions to a set of sequences of groundtruth object positions.
+The ATA is calculated by first computing the Sequence Track Detection Accuracy (STDA), which can be viewed as a tracking performance measure unnormalied in terms of the number of objects.
+We can write the STDA as
+\begin{equation}
+\text{STDA} = \sum_{i=1}^{N_{\text{mapped}}} \frac{\sum_{t=1}^{N_{\text{frames}}} \left( \frac{\left|G_{i}^{(t)} \cap D_{i}^{(t)}\right|}{\left|G_{i}^{(t)} \cup D_{i}^{(t)}\right|}  \right) }{ N_{(G_{i} \cup D_{i} \neq \emptyset)} }
+\end{equation}
+where $N_{\text{mapped}}$ refers to an optimal mapping between ground-truth and detected objects as specified in section \ref{sec:mappingtoground} using the STDA as the relevant metric, and $N_{(G_{i} \cup D_{i} \neq \emptyset)}$ denotes the number of frames in which a given tracked object, the ground truth object to which it is mapped, or both, are present.
+
+Given the STDA for a video sequence, the ATA can be computed by the formula
+\begin{equation}
+\text{ATA} = \frac{\text{STDA}}{\left( \frac{N_{G} + N_{D}}{2} \right)}
+\end{equation}
+
 \end{document}

### 2012-03-15 ltx/paper_uai.tex
newest version. main difference: removal of sensitivity analysis figures, lots of edits

--- 2012-02-29 ltx/paper_short_2.tex
@@
 
-\documentclass[twocolumn, final]{svjour3}
+% \documentclass[twocolumn, final]{svjour3}
+\documentclass{article}
+\usepackage{proceed}
 \usepackage[square,numbers]{natbib}
@@
 \title{Unsupervised Detection and Tracking of Arbitrary Objects with Dependent Dirichlet Process Mixtures}
-\titlerunning{Unsupervised Detection and Tracking of Arbitrary Objects with Dependent Dirichlet Process Mixtures}
+% \titlerunning{Unsupervised Detection and Tracking of Arbitrary Objects with Dependent Dirichlet Process Mixtures}
 \author{Willie Neiswanger$^{*}$ and Frank Wood$^{\dagger}$}
-\authorrunning{Willie Neiswanger and Frank Wood}
-\institute{ $^{*}$Columbia University, Department of Applied Math and Applied Physics.
+% \authorrunning{Willie Neiswanger and Frank Wood}
+% \institute{ $^{*}$Columbia University, Department of Applied Math and Applied Physics.
 Tel.: 503-464-6152.
@@
 \email{fwood@stat.columbia.edu}}
-\date{}  % add submit date when submitted
+% \date{}  % add submit date when submitted
 \maketitle
@@
 \end{equation}
-where each $\bold{x}$ corresponds to a point within an image region where an object (or foreground element)  is believed to exist, $\bold{x}^{s} \in \mathbb{R}^{2}$ denotes the spatial location of this point, $\bold{x}^{c} \in S_{1} \times \ldots \times S_{V}$ denotes some collection of local image characteristics in the vicinity of this point, and $t \in \{1, \ldots, T \}$ denotes the time index.
+where each $\bold{x}$ corresponds to a point within an image region where an object (or foreground element)  is believed to exist, $\bold{x}^{s} \in \mathbb{R}^{2}$ denotes the spatial location of this point, $\bold{x}^{c} \in S_{1} \times \ldots \times S_{V}$ denotes some collection of local image features in the vicinity of this point, and $t \in \{1, \ldots, T \}$ denotes the time index.
 
@@
 Table~\ref{benchmark_results_table} shows performance metric results for comparison (data published with permission from the authors of \cite{ellis_2010}).
+Our method achieves the fourth best SFDA and third best ATA.
+
+The algorithm in \cite{breitenstein2009markovian} is based around output from an externally made human detector; in \cite{yang2009probabilistic}, assumptions about the orientation and appearance of clothing on humans is used; and in
 

### 2012-03-23 ltx/paper_short_2.tex
added newStuff_ltx folder to hold lit review for new things; made some final edits on paper

@@
 The technique uses a dependent Dirichlet process mixture (DDPM) known as the Generalized Polya Urn (GPUDDPM) to model image pixel data that can be easily and efficiently extracted from the regions in a video that represent objects.
-This paper describes a specific implementation of the model using spatial and color pixel data extracted via frame differencing and gives two algorithms for performing inference on the model to accomplish detection and tracking.
+This paper describes a specific implementation of the model using spatial and color pixel data extracted via frame differencing and gives two algorithms for performing inference in the model to accomplish detection and tracking.
 This technique is demonstrated on multiple synthetic and benchmark video datasets that illustrate its ability to, without modification, detect and track objects with diverse physical charactersitics moving over non-uniform backgrounds and through occlusion.
@@
 
-We break-down this task into three parts: data extraction, localization, and tracking.
+This task can be broken down into three parts: data extraction, localization, and tracking.
 Data extraction is the act of extracting features from regions of video-frames that constitute objects, localization is the act of finding the positions and/or shapes of distinct objects, and tracking is the act of maintaining the identities of the detected objects over time.
@@
 These difficulties stem from performing detection via segmentation on individual frames.}
-This paper furthers this line of work by providing a model that gives rise to a number of algorithms to detect arbitrary objects in videos---particularly in cases where frame-by-frame segmentation is difficult, video quality is low, and extraction is noisy---and maintain the isolation of distinct objects during tracking and through occlusion.
+This paper furthers this line of work by providing a model that could give rise to a number of algorithms to detect arbitrary objects in videos---particularly in cases where frame-by-frame segmentation is difficult, video quality is low, and extraction is noisy---and maintain the isolation of distinct objects during tracking and through occlusion.
 
@@
 In our implementation, we extract data via a basic frame differencing procedure and specify distributions useful for representing arbitrary objects (Section~\ref{sec:modelspecification}).
-We describe inference algorithms for our model, and show how output of these algorithms can be interpreted as object localization and tracking results (Section~\ref{sec:inference}).
-Our implementation is demonstrated on multiple synthetic and benchmark datasets, and standard performance metric values are computed to quantify results on the benchmark datasets (Section~\ref{sec:experiments}).
+We describe inference algorithms for our model and show how the output of these algorithms can be interpreted as object localization and tracking results (Section~\ref{sec:inference}).
+Our implementation is demonstrated on multiple synthetic and benchmark datasets.
+Standard performance metric values are computed to quantify results on the benchmark datasets (Section~\ref{sec:experiments}).
 We compare our performance metrics with those yielded by specialized detection and tracking algorithms tailored to specific objects.
-Our results support our hypothesis that simple data extraction combined with inference on our model can perform detection and tracking of arbitrary objects at a level comparable to state-of-the-art, object specific algorithms.
-
-% Prior Work
-% ----------
+Our results support our hypothesis that approaches combining simple data extraction and a powerful model can perform detection and tracking of arbitrary objects at a level comparable to state-of-the-art, object specific algorithms.
+
+% Prior Work / Background
+% -----------------------
 
@@
 Often, background subtraction refers to methods that compare an image containing targets with an image of the background only or with some model of the background that is learned as the video progresses \cite{piccardi2004background}, while frame differencing refers to methods that compare pairs of consecutive images in a video \cite{zhang2001segmentation}.
-Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking schemes \citep{perez_2002}.
+Frame differencing has been used as the sole extraction method for object localization or tracking schemes with success \citep{pece_2002, beleznai_2006, chu_2007}, and also as a secondary data extraction method to help improve the accuracy of object tracking \citep{perez_2002}.
 
 A great deal of research has focused on developing algorithms to track multiple objects simultaneously.
-There has been a particular emphasis on developing ways to deal with problems such as object occlusions (where one object blocks another from the view of a video camera) \cite{senior2006appearance, cucchiara2004probabilistic, zhou2003background}, complex object interactions \cite{khan_2004, mckenna2000tracking, dockstader2001multiple}, objects with similar appearances \cite{maccormick1999probabilistic, jepson_2003}, variable (and potentially high) numbers of objects \cite{reilly2010detection}, and objects that enter and exit a field of view at different times \cite{stauffer2003estimating, nedrich2010learning}.
+There has been a particular emphasis on developing ways to deal with problems such as object occlusion (where one object blocks another from the view of a video camera) \cite{senior2006appearance, cucchiara2004probabilistic, zhou2003background}, complex object interactions \cite{khan_2004, mckenna2000tracking, dockstader2001multiple}, objects with similar appearances \cite{maccormick1999probabilistic, jepson_2003}, variable (and potentially high) numbers of objects \cite{reilly2010detection}, and objects that enter and exit a field of view at different times \cite{stauffer2003estimating, nedrich2010learning}.
 Multiple independent single-object trackers running simultaneously have been shown to be ineffective, as they will tend to coalesce and track the same object \cite{khan_2004}.
-To remedy this problem, methods such as \cite{maccormick1999probabilistic} have developed probabilistic principles for maintaining isolation of object trackers.
+To remedy this problem, methods have incorporated probabilistic principles for maintaining isolation of object trackers \cite{maccormick1999probabilistic}.
 An approach to this problem involving the use of a nonparametric mixture model has also found success in maintaining isolation of distinct objects \cite{vermaak_2003}.
@@
 Over the past decade, there have been attempts to provide general algorithms for the fully unsupervised detection and tracking of arbitrary objects in videos.
-Blob tracking, a basic method to carry out this goal, has found success in videos where objects are easily isolated from the background and where localization and segmentation of distinct objects is possible \cite{francois2004real, isard_2001}.
-Methods involving blob tracking, however, run into problems when faced with videos where detection is difficult, object appearance or orientation varies heavily, and there exists object occlusion \cite{song2005model}.
+Blob tracking, a basic method for carrying out this goal, has found success in videos where objects are easily isolated from the background and where localization and segmentation of distinct objects is possible \cite{francois2004real, isard_2001}.
+Blob tracking methods, however, run into problems when faced with videos where detection is difficult, object appearance or orientation varies heavily, and there exists object occlusion \cite{song2005model}.
 To aid the accuracy of these methods, techniques have been developed for performing extraction and segmentation in a joint manner, incorporating statistical methods for maintaining hypotheses of different numbers of detected objects, and introducing some of the multi-object tracking methods described previously to track distinct blobs after they have been segmented \cite{collins2003mean, isard_2001}.
 Another family of methods related to the task of unsupervised detection and tracking of video objects goes under the heading of video segmentation algorithms---these methods extend single-frame image segmentation to maintain coherence of image segments over time, and have had some success when used for the explicit purpose of detecting and tracking foreground objects in videos \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}.
-Other attempts to perform unsupervised detection and tracking include methods for clustering short sequences of positions extracted by detecting the motion of objects \cite{brostow2006unsupervised, brox2010object}, which aim to return full-length distinct object tracks, and graph based method that carry out a similar task using spectral clustering \cite{fragkiadaki2011detection}.
+Other attempts to perform unsupervised detection and tracking include methods for clustering short sequences of positions extracted by detecting the motion of objects \cite{brostow2006unsupervised, brox2010object}, which aim to return full-length distinct object tracks, and graph based methods that carry out a similar task using spectral clustering \cite{fragkiadaki2011detection}.
 Another approach uses a Gaussian mixture model to cluster data extracted from moving objects \cite{pece_2002}; this method also develops heuristics for the initialization and elimination of new object tracks.
@@
 Nearly all high accuracy object detection and tracking methods are those tailored for specific object types.
-These methods rely on ad-hoc detection criteria that exploit knowledge about the appearance or behavior of the given objects in a video.
-They often make use of state-of-the-art detectors designed to locate the specified objects of interest.
+These methods rely on detection criteria that exploit knowledge about the appearance or behavior of the objects in a video.
+Some of these methods make use of state-of-the-art detectors designed to locate the specified objects of interest.
 In contrast, the method described in this paper is designed to track arbitrary objects without using any explicit detection criteria, and serve as a general strategy that can be used, without modification, to perform accurate detection and tracking of diverse objects in a wide range of videos.
-The method we introduce falls into the category of clustering-based arbitrary object detection and tracking methods.
+The technique we introduce falls into the category of clustering-based arbitrary object detection and tracking methods.
 Differing from previous work, we use a type of time dependent Bayesian nonparametric mixture model, and show how it can be applied to a variety of easily extracted data to perform detection and tracking.
-This method begins by performing a simple data extraction procedure, which can be carried out broadly but yields noisy data.
+This method begins by performing a simple data extraction procedure that yields noisy data.
 Our model of this data serves as a general framework for which we can choose a variety of object appearance distributions and inference algorithms; each choice provides a new method for unsupervised detection and tracking of arbitrary objects in videos.
@@
 Consequently, we use frame differencing.
-This procedure locates the pixels in image regions where there is motion relative to a previous frame.
+This procedure locates the pixels in image regions that undergo change.
+Specifically, at each frame the pixels that differ from the previous frame beyond some threshold are recorded.
 Here, each pixel corresponds to an observation $\bold{x}$.
@@
 
-% Model Definition
-% --------------
+% Model Definition / Framework
+% ----------------------------
 
@@
 These models have been widely used in the past decade to perform nonparametric density estimation and cluster analysis.
-Data extracted from videos comprises spatiotemporal clusters, each corresponding with a distinct object.
+Data extracted from videos comprises spatiotemporal clusters, each corresponding to a distinct object.
 Consequently, we are interested in using a class of models known as dependent Dirichlet process mixture (DDPM) models (Section~\ref{sec:ddpmixture}), which are particularly useful for estimating the number of latent classes (clusters) in time dependent data.
@@
 Since objects can enter and exit a scene, the number of clusters present throughout the video may not be constant (i.e., clusters may be created, or be ``born'', and may disappear, or ``die'', at intermediate time steps).
-To cluster data with these properties, we choose to use a DDPM known as the Generalized Polya Urn dependent Dirichlet process mixture, introduced by \cite{caron_2007}.
-This model may be intuitively viewed as a sequence of DPMs, where there exist dependencies between the parameters and number of clusters in adjacent time steps.
-Like the DPM, the GPUDDPM allows for the number of clusters within a dataset---which, in this work, corresponds to the number of objects in a video---to be found in an automated manner.
+To cluster data with these properties, we choose to use a DDPM known as the Generalized Polya Urn dependent Dirichlet process mixture \cite{caron_2007}.
+This model may be viewed intuitively as a sequence of DPMs, where there exist dependencies between the parameters and number of clusters in adjacent time steps.
+Like the DPM, the GPUDDPM allows for a distribution over the number of clusters within a dataset---which, in this work, corresponds to the number of objects in a video---to be inferred.
 
@@
 
-In this model, each observation $\bold{x}_{i,t}$ is associated with an assignment variable $c_{i,t}$ that represents its assignment to a cluster $\theta_{k,t}$. 
+In a GPUDDPM, each observation $\bold{x}_{i,t}$ is associated with an assignment variable $c_{i,t}$ that represents its assignment to a cluster $\theta_{k,t}$. 
 %In a DPM, the probability of assignment to a cluster $k$ is a function of the cluster size $m_{k,t}$, which is equal to the number of observations that have been assigned to it. In contrast, 
-The size (i.e.
-number of assigned observations) of each cluster in the GPUDDPM may decrease over time as observations become ``unassigned''.
+The sizes of clusters in the GPUDDPM increase when observations are assigned to them, and decrease at later time points when these observations become ``unassigned''.
 We define a distribution over the size of cluster $k$ at time $t$ ($m_{k,t}$), conditioned on the cluster's previous size ($m_{k,t-1}$), the assignments at time $t$ ($c_{1:N_{t}, t}$), and a deletion parameter ($\rho$),
@@
 
-Distributions \eqref{D_distro} and \eqref{C_distro} together comprise what \cite{caron_2007} refers to as the ``Generalized Polya Urn''. We can now define the GPUDDPM generatively as
+Distributions \eqref{D_distro} and \eqref{C_distro} together comprise what is referred to as the ``Generalized Polya Urn'' \cite{caron_2007}. We can now define the GPUDDPM generatively as
 \begin{align}
@@
 \end{align}
-$\forall$ times $t \in \{1, \ldots, T\}$ and each cluster $k \in \{ 1, \ldots, K_{t} \}$ at time $t$, where we specify distributions for F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ in Section~\ref{sec:modelspecification}.
-The graphical model associated with this formulation is shown in Figure~\ref{fig:gpuddpm_gm_1}.
+$\forall$ times $t \in \{1, \ldots, T\}$ and each cluster $k \in \{ 1, \ldots, K_{t} \}$ at time $t$, where we choose application-specific distributions for F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ in Section~\ref{sec:modelspecification}.
+The graphical model associated with this formulation of the GPUDDPM is shown in Figure~\ref{fig:gpuddpm_gm_1}.
 \begin{figure}[h]
@@
 
-\subsection{Object Appearance and Mixture Component F}
+\subsection{Object Appearance and Mixture Component, F}
 \label{sec:objectappearance}
@@
 \end{equation}
-where $\theta = \{ \boldsymbol{\mu}, \Sigma, \bold{p} \}$ denotes the parameters of a cluster at time $t$, with mean $\boldsymbol{\mu} \in \mathbb{R}^{2}$, covariance matrix $\Sigma \in \mathbb{R}^{2} \times \mathbb{R}^{2}$, and discrete probability vector $\bold{p} = (p_{1}, \ldots, p_{V})$ such that $\sum_{i=1}^{V}p_{i} = 1$.
+where $\theta = \{ \boldsymbol{\mu}, \Sigma, \bold{p} \}$ denotes the parameters of a cluster at time $t$, with mean $\boldsymbol{\mu} \in \mathbb{R}^{2}$, covariance matrix $\Sigma \in \mathbb{R}^{2\times2}$, and discrete probability vector $\bold{p} = (p_{1}, \ldots, p_{V})$ such that $\sum_{i=1}^{V}p_{i} = 1$.
 Additionally, $\mathcal{N}$ denotes the multivariate normal distribution and $\mathcal{M}n$ denotes the multinomial distribution.
-The mixture component F is sometimes referred to as the likelihood.
+Note that we withhold writing the subscripts specifying the cluster number k and time t in this section when writing them is unnecessary.
 
 The multivariate normal distribution over the spatial features $\bold{x}^{s}$ models the position and spatial extent of an object.
-It can intuitively be thought to represent the shape of each object as an oval.
+It can be thought intuitively to represent the shape of each object as an oval.
 We model the color features $\bold{x}^{c}$ as draws from a multinomial distribution.
@@
 
-\subsection{Appearance Prior and Base Distribution $\mathbb{G}_{0}$}
+\subsection{Appearance Prior and Base Distribution, $\mathbb{G}_{0}$}
 \label{sec:appearanceprior}
@@
 We use conjugate priors in the base distribution to allow for more efficient computation.
-Specifically, in our implementation, a normal-inverse-Wishart prior is placed on the multivariate normal parameters $\{ \boldsymbol{\mu}, \Sigma \}$, and a Dirichlet prior is placed on the multinomial parameter $\bold{p}$ (where the normal-inverse-Wishart is a conjugate prior for the multivariate normal component of the likelihood and the Dirichlet is a conjugate prior for the multinomial component).
+Specifically, in our implementation, a normal-inverse-Wishart prior is placed on the multivariate normal parameters $\{ \boldsymbol{\mu}, \Sigma \}$, and a Dirichlet prior is placed on the multinomial parameter $\bold{p}$.
 The prior can therefore be written
@@
 
-\subsection{Motion Model and Transition Kernel $P(\theta_{t} | \theta_{t-1})$}
+\subsection{Motion Model and Transition Kernel, $P(\theta_{t} | \theta_{t-1})$}
 \label{sec:motionmodel}
@@
 Since our implementation is intended for tracking arbitrary objects, we do not wish to make sophisticated assumptions about object motion.
-For example, we choose not to incorporate complex objects dynamics, though they are often used with success in certain object-specific tracking tasks, such as people tracking.
-We assume only that the position of an object at a given time is a function of its position at the previous time, and that the position varies in all directions equally between time steps.
-
-The base distribution $\mathbb{G}_{0}$ must be the invariant distribution of the transition kernel $P(\theta_{t} | \theta_{t-1})$ in order for the the cluster parameters to remain marginally distributed according to the base distribution.
+For example, we choose not to incorporate complex objects dynamics, though they are often used with success in certain object-specific tracking tasks, such as people tracking \cite{choo2001people, conte2010performance}.
+We assume only that the position of an object at a given time is close to its position at the previous time, and that the position varies in all directions equally between time steps.
+
+The base distribution $\mathbb{G}_{0}$ must be the invariant distribution of the transition kernel $P(\theta_{t} | \theta_{t-1})$ in order for the the cluster parameters to remain marginally distributed according to the base distribution, and for the model to be a valid GPUDDPM.
 In other words, the transition kernel must satisfy
@@
 One way to achieve this is through the use of auxiliary variables.
-Auxiliary variables are a set of $M$ variables $\bold{z}_{t} = (z_{t,1}, \ldots, z_{t,M})$ associated with each cluster at each time $t$ that satisfy
+These are a set of $M$ variables $\bold{z}_{t} = (z_{t,1}, \ldots, z_{t,M})$ associated with each cluster at each time $t$ that satisfy
 \begin{eqnarray}
@@
 
-With the addition of these variables, the parameters of a cluster at a given time do not depend directly on their value at the previous time; they are instead dependent upon an intermediate sequence of variables.
+With the addition of these variables, the parameters of a cluster at a given time do not depend directly on their value at the previous time; they are instead dependent through an intermediate sequence of variables.
 This allows the cluster parameters at each time step to be marginally distributed according to the base distribution $\mathbb{G}_{0}$ while maintaining simple time varying behavior.
 
-For each cluster, we introduce $M$ auxiliary variables $z_{t, 1}, \ldots, z_{t, M}$ at time $t$.
-Each auxiliary variable is drawn from the product of a multivariate normal and multinomial with the associated cluster parameters $\theta_{t} = \{ \boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t} \}$
+Each of the auxiliary variables $z_{t,m}$ is drawn from the product of a multivariate normal and multinomial with the associated cluster parameters $\theta_{t} = \{ \boldsymbol{\mu}_{t}, \Sigma_{t}, \bold{p}_{t} \}$
 \begin{equation}
@@
 \end{equation}
-where $\boldsymbol{\mu}_{M}, \kappa_{M}, \nu_{M}, \Lambda_{M},$ and $\bold{q}_{M}$ are parameters given by
+where $\boldsymbol{\mu}_{M}, \kappa_{M}, \nu_{M}, \Lambda_{M},$ and $\bold{q}_{M}$ are
 \begin{eqnarray}
@@
 \end{eqnarray}
-where $M$ is the number of auxiliary variables, \\
-$\{ \boldsymbol{\mu}_{0}, \kappa_{0}, \nu_{0}, \Lambda_{0} \}$ are the $\mathcal{N}i\mathcal{W}$ prior parameters, $\bold{q}_{0}$ is the $\mathcal{D}ir$ prior parameter, $\bold{z}^{s}$ and $\bold{z}^{c}$ respectively denote the spatial and color features of an auxiliary variable $\bold{z}$, and $\overline{\bold{z}}$ and $S_{\bold{z}}$ respectively denote the sample mean and sample covariance for a set $\bold{z} = \{ z_{1}, \ldots, z_{M} \}$ (of auxiliary variables, in this case), which we can write as
+and where $M$ is the number of auxiliary variables, $\{ \boldsymbol{\mu}_{0},$ $\kappa_{0}$, $\nu_{0}$, $\Lambda_{0} \}$ are the $\mathcal{N}i\mathcal{W}$ prior parameters, and $\bold{q}_{0}$ is the $\mathcal{D}ir$ prior parameter.
+We use $\bold{z}^{s}$ and $\bold{z}^{c}$ to respectively denote the spatial and color features of an auxiliary variable $\bold{z}$, and $\overline{\bold{z}}$ and $S_{\bold{z}}$ to respectively denote the sample mean and sample covariance for a set $\bold{z} = \{ z_{1}, \ldots, z_{M} \}$ (of auxiliary variables, in this case), which we can write as
 \begin{eqnarray}
@@
 
-\subsection{Recap of Parameters in Model}
+\subsection{Recap of Model Parameters}
 \label{sec:recapofparameters}
 
-The multivariate normal-multinomial GPUDDPM implemented in this study has a number of parameters, which are used to specify the object appearance prior distribution, transition kernel, and Generalized Polya Urn distributions (C and D).
+The multivariate normal-multinomial GPUDDPM for object tracking has a number of parameters, which are used to specify the object appearance prior distribution, transition kernel, and Generalized Polya Urn distributions (C and D).
 %\newpage %remove if formatting changes
@@
 \begin{center}
-\begin{tabular}[c]{| l |  p{7cm} | }
+\begin{tabular}[c]{| p{2cm} |  p{5.5cm} | }
+%\begin{tabular}[c]{| l |  p{7cm} | }
 \hline
-$\boldsymbol{\mu}_{0}$  &  Mean position prior. In experiments performed in Section~\ref{sec:experiments} the data was recentered to the origin, and this parameter was set to $(0,0)$. \\ \hline
-$\kappa_{0}$  &  Scale factor of the mean prior.\\ \hline
-$\Lambda_{0}$  &  Shape factor of the covariance prior.\\ \hline
-$\nu_{0}$  &  Scale factor of the covariance prior.\\ \hline
-$\bold{q}_{0}$  &  Scale factor of the multinomial prior. \\
+$\boldsymbol{\mu}_{0} \in \mathbb{R}^{2}$  &  Mean position prior.
+In experiments performed in Section~\ref{sec:experiments} the data was recentered to the origin, and this parameter was set to $(0,0)$.
+\\ \hline
+$\kappa_{0} \in \mathbb{R}$  &  Scale factor of the mean prior.\\ \hline
+$\Lambda_{0} \in \mathbb{R}^{2 \times 2}$  &  Shape factor of the covariance prior.\\ \hline
+$\nu_{0} \in \mathbb{Z}_{+} > 1$  &  Scale factor of the covariance prior.\\ \hline
+$\bold{q}_{0} \in \mathbb{R}_{+}^{V}$  &  Scale factor of the multinomial prior. \\
 \hline
@@
 \begin{center}
-\begin{tabular}[c]{ | l | p{7cm} | }
+\begin{tabular}[c]{ | p{2cm} | p{5.5cm} | }
 \hline
-$M$  &  Number of auxiliary variables. A larger number will produce a smoother object path.\\
+$M \in \mathbb{Z}_{+}$  &  Number of auxiliary variables. A larger number will produce a smoother object path.\\
 \hline
@@
 \begin{center}
-\begin{tabular}[c]{ | l | p{7cm} | }
+\begin{tabular}[c]{ | p{2cm} | p{5.5cm} | }
 \hline
-$\alpha$  &  The concentration parameter for the Dirichlet process. A higher value will increase the tendency for new objects to be detected.\\ \hline
-$\rho$  &  The deletion parameter. A higher value will give objects an increased tendency to die off. \\
+$\alpha \in \mathbb{R}_{+}$  &  The concentration parameter for the Dirichlet process. A higher value will increase the tendency for new objects to be detected.\\ \hline
+$\rho \in ( 0, 1 ]$  &  The deletion parameter. A higher value will give objects an increased tendency to die off. \\
 \hline
@@
 
-Bayesian inference, which infers distributions over latent model parameters, is used to achieve detection and tracking results.
+Bayesian inference is used to achieve detection and tracking results.
 Previously developed inference strategies can be applied to the generative model defined in Section~\ref{sec:gpuddpm} and Section~\ref{sec:modelspecification}.
-We provide details on the two Bayesian inference algorithms implemented in this study.
+We provide details on the two Bayesian inference algorithms implemented in this stu-\\dy.
 The first is a type of Markov Chain Monte Carlo (MCMC) batch inference, which uses Gibbs sampling to generate samples from the posterior distribution of the model.
-The second is a type of Sequential Monte Carlo (SMC) inference, also known as a particle filter, which exploits the time-series data to generate samples from the posterior distribution of the model in a sequential manner.
+The second is a type of Sequential Monte Carlo (SMC) inference, also known as a particle filter, which generates samples from the posterior distribution of the model in a sequential manner.
 
@@
 
-This section details the MCMC sampler used to perform inference, which serves as our first algorithm for unsupervised detection and tracking.
+This section details the MCMC sampler used to perform inference.
+% which serves as our first algorithm for unsupervised detection and tracking.
 It iteratively samples from the conditional distribution of each element in our model given all the rest---i.e.
@@
 A secondary formulation of the GPUDDPM, which we refer to as the ``Deletion Variable Formulation'' (defined in Section~\ref{sec:deletionvariableformulation}), is used here.
-This formulation is equivalent to the formulation given in Section~\ref{sec:gpuddpm}, but allows for easier Gibbs sampling.
+This formulation is equivalent to the formulation given in Section~\ref{sec:gpuddpm}, but allows for easier sampling.
 
@@
 
-Instead of incorporating the cluster size variable $m_{k,t}$ directly in the GPUDDPM model (as it was in the definition given by \eqref{gpuddpm_def}), we can formulate an equivalent model which makes use of new set of variables called the deletion variables.
+Instead of incorporating the cluster size variable $m_{k,t}$ directly in the GPUDDPM model (as it was in the definition given by \eqref{gpuddpm_def}), we can formulate an equivalent model which makes use of new set of variables called deletion variables.
 We introduce a deletion variable $d_{i,t}$ for each observation $\bold{x}_{i,t}$, which denotes the time at which the observation is removed from its assigned cluster.
@@
 
-We can now define the Deletion Variable Formulation of the GPUDDPM generatively to be
+We can now define the Deletion Variable Formulation of the GPUDDPM generatively as
 \begin{align}
@@
 
-To allow for easier notation, we define $\boldsymbol{x}_{t} = \bold{x}_{1:N_{t},t}$, $\bold{c}_{t} = c_{1:N_{t},t}$, $\bold{d}_{t} = d_{1:N_{t},t}, \Theta_{t} = \theta_{1:K_{t}, t}$, and $\bold{z}_{t} = z_{1:K_{t},t,1:M}$.
+For easier notation, we define $\boldsymbol{x}_{t} = \bold{x}_{1:N_{t},t}$, $\bold{c}_{t} = c_{1:N_{t},t}$, $\bold{d}_{t} = d_{1:N_{t},t}, \Theta_{t} = \theta_{1:K_{t}, t}$, and $\bold{z}_{t} = z_{1:K_{t},t,1:M}$.
 At each time $t$, the sampler moves sequentially through the $N_{t}$ observations, sampling the assignment $c_{i,t}$ and the deletion variable $d_{i,t}$ for each.
-After, the cluster parameters for all active clusters at $t$ are sampled, and the $M$ auxiliary variables for all active clusters at $t$ are sampled (using the MH algorithm).
-Note that before we can iteratively sample each of these elements, each variable in S must be initialized.
-The following sections detail the posterior distributions from which each of these samples is drawn.
+Afterwards, the cluster parameters for all active clusters at $t$ are sampled, and the $M$ auxiliary variables for all active clusters at $t$ are sampled using Metropolis Hastings (MH).
+The following sections detail the distributions from which each of these samples is drawn.
 
@@
 These values allow us to construct a discrete probability distribution from which we can draw samples from the posterior distribution over assignments.
-The possible values that the $c_{i,t}$ may take on are $k \in \{ 1 , \ldots ,  K_{t}^{'}+1 \}$, where $K_{t}^{'}$ denotes the number clusters with a non-zero size at any time $t' \in \{ t, \ldots, d_{i,t} \}$, and $K_{t}'+1$ denotes a newly sampled cluster.
-The probability that assignment $c_{i,t}$ equals a potential cluster $k$, given all other variables in the model (which we denote as ``$\ldots$"), is given by
+The possible values that the $c_{i,t}$ may take on are $k \in \{ 1 , \ldots ,  K_{t}^{'}+1 \}$, where $K_{t}^{'}$ denotes the number clusters with a non-zero size at any time $t' \in \{ t, \ldots, d_{i,t} \}$, and $K_{t}'+1$ denotes a ``new'' cluster.
+The probability that $x_{i,t}$ is assigned to cluster $k$, i.e.
+that $c_{i,t}=k$, given values for all other variables in the model (which we denote as ``$\ldots$"), is given by
 \begin{equation}
@@
 \end{equation}
-where the C is given by \eqref{C_distro}, and the cluster sizes $\bold{m}_{t'}$ are calculated under the assumption that $c_{i,t} = k$.
+where C is given by \eqref{C_distro}, and the cluster sizes $\bold{m}_{t'}$ are calculated under the assumption that $c_{i,t} = k$.
 Note that the above integral has an analytic solution for our specific model
@@
 \int & P(\bold{x}_{i} | \theta) \mathbb{G}_{0}(\theta)d\theta \\
-&= \text{Student}_{\nu_{n} - 1}(\mu_{n}, \Lambda_{n}/(\kappa_{n}(\nu_{n}-1)))
+% &= \text{Student}_{\nu_{n} - 1}(\mu_{n}, \Lambda_{n}/(\kappa_{n}(\nu_{n}-1)))
+&= t_{\nu_{0} - 1}  \left ( \bold{x}_{i}^{s}  \hspace{1mm} \big{|} \hspace{1mm} \boldsymbol{\mu}_{0}, \frac{\Lambda_{0}(\kappa_{0}+1)}{\kappa_{0}(\nu_{0}-1)} \right ) \\
+& \times \prod_{v=1}^{V} \frac{ \Gamma(\bold{x}_{i}^{c}) }  { \Gamma(\bold{q}_{0})}
+\times \frac{\Gamma(\sum_{j=1}^{V} \bold{q}_{0}) }{\Gamma(\sum_{j=1}^{V} \bold{x}_{i}^{c}) }
 \end{split}
 \end{equation}
-where Student denotes the t-distribution, where we follow the three-value parameterization (location parameter, scale parameter, and degrees of freedom) given in \cite{gelman2004bayesian}.
+where t denotes the multivariate t-distribution, where we follow the three-value parameterization (location parameter, scale parameter, and degrees of freedom) given in \cite{gelman2004bayesian, Wood2008nonparametric}, and $\{ \boldsymbol{\mu}_{0},$ $\kappa_{0},$ $\Lambda_{0},$ $\nu_{0},$ $\bold{q}_{0} \}$ are prior parameters.
 
@@
 
-The conjugacy of chosen distributions in the appearance model and transition kernel allow us to easily sample from the posterior distribution over the cluster parameters, which we can write
+The conjugacy of appearance model and transition kernel distributions allow us to easily sample from the posterior distribution over the cluster parameters, which we can write
 \begin{align}
@@
 
-This section details a Sequential Monte Carlo (SMC) sampler---also known as a particle filter---used to perform inference, which provides another method for unsupervised detection and tracking.
-SMC inference operates in the orginal GPUDDPM formulation (defined in Section~\ref{sec:gpuddpm}).
+This section details a Sequential Monte Carlo (SMC) sampler---also known as a particle filter---used to perform inference.
+SMC inference operates in the orginal GPUDDPM formulation (Section~\ref{sec:gpuddpm}).
 The algorithm is shown in Algorithm~\ref{alg:SMC}.
-At each time step $t \in \{ 1, \ldots, T \}$, a number of samples refered to as ``particles'' are generated; each particle consists of a sample from the posterior distribution over the assignment for each observation, $c_{1,t}, \ldots, c_{N_{t}, t}$, parameters for each cluster $\theta_{1,t}, \ldots, \theta_{K_{t}, t}$, and size after deletion for each cluster $m_{1,t}, \ldots, m_{K_{t},t}$.
+At each time step $t \in \{ 1, \ldots, T \}$, a number of samples refered to as ``particles'' are generated; each particle consists of a sample from the posterior distribution over the assignment for each observation, $c_{1,t}, \ldots, c_{N_{t}, t}$, parameters for each cluster $\theta_{1,t}, \ldots,$ $\theta_{K_{t}, t}$, and size after deletion for each cluster $m_{1,t},$ $\ldots,$ $m_{K_{t},t}$.
 A set of particles is sampled at each time step from relevant proposal distributions (described in Sections~\ref{sec:smc_proposal_1}, \ref{sec:smc_proposal_2}, and \ref{sec:smc_proposal_3}), a weight is computed for each particle, and a new set of particles are sampled from the set of weighted particles via a resampling process.
@@
 \ENDFOR
-\STATE $\tilde{w}_{t}^{(l)} \leftarrow w_{t-1}^{(l)} \times \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})} {P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) } \times r $  % \hfill $\triangleright$ eq.~\eqref{smc_weight_eqn}
+\STATE $\tilde{w}_{t}^{(l)} \leftarrow w_{t-1}^{(l)} \times \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})} {P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) } $% \times r $  % \hfill $\triangleright$ eq.~\eqref{smc_weight_eqn}
 \ENDFOR
@@
 \end{align}
-where C is defined in \eqref{C_distro}, and the integral \\
+where C is defined in \eqref{C_distro}, and
 $\int P(\bold{x}_{i,t} | \theta) \mathbb{G}_{0}(\theta)d\theta$ can be determined analytically, and is given in \eqref{marginal_obs_prob}.
@@
 \label{smc_q2_eqn}
-q_{2}(\theta_{k,t} | \theta_{k,t-1},\bold{x}_{1:N,t}) = P(\theta_{k,t} | \theta_{k,t-1} \bold{x}_{1:N,t})
+q_{2}(\theta_{k,t} | \theta_{k,t-1}, \bold{x}_{1:N,t}) = P(\theta_{k,t} | \theta_{k,t-1}, \bold{x}_{1:N,t})
 \end{equation}
@@
 In this step, $L$ new particles are sampled from current set of $L$ particles.
-We will describe a specific, simple method known as Multinomial Resampling.
-In this method of resampling, we draw
-\begin{equation}
-\Upsilon \sim \mathcal{M}n((w_{t}^{(1)}, \ldots, w_{t}^{(L)}))
-\end{equation}
-where $\mathcal{M}n$ denotes the multinomial distribution, and $\Upsilon$ represents a vector of counts showing the multiplicity of each of the old particles in the new particle set (i.e.
+% We will describe a specific, simple method known as Multinomial Resampling. In this method of resampling, we draw
+% \begin{equation}
+% \Upsilon \sim \mathcal{M}n((w_{t}^{(1)}, \ldots, w_{t}^{(L)}))
+% \end{equation}
+% where $\mathcal{M}n$ denotes the multinomial distribution, and $\Upsilon$ represents a vector of counts showing the multiplicity of each of the old particles in the new particle set (i.e.
 $\Upsilon_{i}$ denotes how many times particle $i$ is duplicated in the new particle set).
-More sophisticated resampling strategies are described for the SMC algorithm in \cite{gasthaus_thesis} and \cite{douc2005comparison}.
+Resampling strategies such as those described in \cite{gasthaus_thesis} and \cite{douc2005comparison} can be used.
 
@@
 We choose to make our detection and tracking results depend completely on the inferred parameters of the multivariate normal factor of each cluster.
-Each inferred cluster is taken to be a distinct detected object, and the sequence of means and covariance matrices for a given cluster are used, respectively, to determine the position and spatial region of a given object over a sequence of time steps.
+Each inferred cluster is taken to be a distinct object, and the sequence of means and covariance matrices for a given cluster are used to determine the position and spatial region, respectively, of a given object over a sequence of time steps.
 In particular, the mean parameter is taken to be the centroid of an object, and a 2-dimensional oval centered on the mean that contains a specified percentage of the normal distribution mass (where we refer to the specified percentage as the confidence value) is taken to be the spatial region of the object.
@@
 
-Performance evaluation metrics, which intend to provide a standardized way of quantifying the success of a detection and tracking procedure on a given video, have started to become consistently used in the past four years.
+Performance evaluation metrics, which provide a standardized way of quantifying the success of a detection and tracking procedure on a given video, have started to become consistently used in the past four years.
 The metrics presented in \cite{kasturi_2008} and used in \cite{ellis_2010, taj_2007, lee_2009} have become well established metrics for evaluating the performance of object detection and tracking in videos and have been adopted by the Video Analysis and Content Extraction (VACE) program and the Classification of Events, Activities, and Relationships (CLEAR) consortium, two large-scale efforts concerned with video tracking and interaction analysis.
@@
 
-Frame differencing is used in all experiments to identify pixels exhibiting motion. 
-For each pixel $\bold{x} = ( x_{1}, x_{2}, t )$ recorded during frame differencing, we also extract color information.
-Specifically, we specify a square, $L$ pixels in length, centered on $(x_{1}, x_{2})$, that contains a set of pixels surrounding $\bold{x}$ in frame $t$.
-We choose to capture the hue for each pixel; consequently, the hue value for each of the selected pixels surrounding $\bold{x}$ is recorded.
-Afterwards, the set of possible color values (i.e.
-the range of color values to which a pixel may be assigned) is partitioned into $V$ bins, and the number of pixels with a color value lying in each of the bins yields the $V$ dimensional vector of color counts.
+Frame differencing was used in all experiments to identify pixels exhibiting motion. 
+For each pixel $\bold{x} = ( x_{1},$ $x_{2},$ $t )$ recorded during frame differencing, we also extracted color information.
+Specifically, we specified a square, $L$ pixels in length, centered on $(x_{1}, x_{2})$, that contained a set of pixels surrounding $\bold{x}$ in frame $t$.
+We chose to capture the hue for each pixel.
+The set of possible hue values (i.e.
+the range of hues to which a pixel may be assigned) was partitioned into $V$ bins, and the number of pixels with a color value lying in each of the bins yielded the $V$ dimensional vector of color counts.
 
@@
 Each of the following synthetic videos consists of a sequence of 200 images (each of size $500$ x $500$ pixels) containing a number of smaller colored squares of different (and potentially time-varying) sizes moving at varied speeds and trajectories over a black background.
-The synthetic videos contain instances of occlusion (where one or more objects are briefly hidden) and objects with time-varying appearances and behaviors, as these are examples of occurances that notoriously decrease the accuracy of detection and tracking.
-After each video was constructed, the extraction procedure described in Section~\ref{sec:dataextraction} (using $L=3$) and inference procedures described in \ref{sec:inference} were carried out to return a sequence of multivariate-normal-parameters (means and covariance matrices) from which a sequence of positions and ovals approximating the shape of a tracked object over each frame that it is present in the video can be determined (as outlined in Section~\ref{sec:inferencetotrackingresults}).
+The synthetic videos contain instances of occlusion (where one or more objects are briefly hidden) and objects with time-varying appearances and behaviors, as these notoriously decrease the accuracy of detection and tracking.
+After each video was constructed, the extraction procedure described in Section~\ref{sec:dataextraction} (using $L=3$) and inference procedures described in Section~\ref{sec:inference} were carried out to return a sequence of multivariate-normal-parameters (means and covariance matrices), which are used to determine a sequence of positions and ovals approximating, respectively, the locations and shapes of a tracked object over each frame that it is present in the video (as outlined in Section~\ref{sec:inferencetotrackingresults}).
 
@@
 
-Parameters were set to the same values for inference on both videos.
-The chosen values were $\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right)$, and $q_{0} = (5, \ldots, 5)$.
-Inference was carried out using the MCMC algorithm (Section~\ref{sec:MCMC}); the MAP sample taken as the result correctly tracked both colored squares through occlusion in both videos, and is shown in Figure~\ref{fig:synth_one_plot}.
+Parameters were set to the same values for inference on both videos: $\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right)$, and $q_{0} = (5, \ldots, 5)$.
+Inference was carried out using the MCMC algorithm (Section~\ref{sec:MCMC}); the MAP sample correctly tracked both colored squares through occlusion in both videos, and is shown in Figure~\ref{fig:synth_one_plot}.
 
 The second synthetic video experiment aimed to test tracking performance under occlusion, object appearance change, and motion change.
-A video was constructed, containing a red square (rgb value $[255,0,0]$), a green square (rgb value $[0,255,0]$), and a blue square (rgb value $[0,0,255]$).
+A video was constructed showing a red square (rgb value $[255,0,0]$), a green square (rgb value $[0,255,0]$), and a blue square (rgb value $[0,0,255]$).
 The red square was of size $20$ x $20$ pixels, the blue square was of size $15$ x $15$ pixels, and the green square began at size $50$ x $50$ pixels at frame $f=1$, linearly shrinks to $10$ x $10$ pixels at $f=100$, then linearly grows back to $50$ x $50$ pixels by the end of the video, $f=200$.
@@
 Figure~\ref{fig:synth_one_plot} shows a sample from the posterior distribution of the cluster parameters, where the mean and oval representation of the covariance matrix (with $0.5$ confidence value) are overlayed on the data.
-The data is plotted in three dimensional space, where time is represented on the vertical axis, and the assignment of each data point to one of the three inferred clusters is denoted by color and marker type.
+The data is plotted with time on the vertical axis, and the assignment of each data point to one of the three inferred clusters is denoted by color and marker type.
 
@@
 Due to the large number of frames and objects in this video, the SMC algorithm (described in Section~\ref{sec:SMC}) was used for inference.
-This method of sequential inference was observed, on this dataset, to converge to a better sample in a shorter period of time in comparison with the MCMC algorithm; for this reason, we found that it was better suited for this large dataset.
+This method of sequential inference was observed, on this dataset, to converge to a better sample in a shorter period of time in comparison with the MCMC algorithm.
 
@@
 
-The SMC inference algorithm yielded a sample from the posterior distribution of the model, from which the object detection and tracking results were obtained (as described in Section~\ref{sec:inferencetotrackingresults}).
-In Figure \ref{fig:pets2009_results}, a sample from the posterior distribution over the cluster parameters is overlayed on the extracted data over a sequence of frames, where the assignment of each data point is represented by color and marker type.
+The SMC inference algorithm yielded an estimate of the posterior distribution of the model, from which the object detection and tracking results were obtained (as described in Section~\ref{sec:inferencetotrackingresults}).
+In Figure \ref{fig:pets2009_results}, the MAP sample from the posterior distribution over the cluster parameters is overlayed on the extracted data over a sequence of frames, where the assignment of each data point is represented by color and marker type.
 
@@
 not for general detection and tracking of arbitrary objects).
-In consequence, many of these studies use externally developed (and trained) state-of-the-art human detectors, exploit the orientation of the humans in this specific dataset, or apply motion models based on assumptions about human motion.
+As a consequence, many of these studies use externally developed (and trained) state-of-the-art human detectors, exploit the orientation of the humans in this specific dataset, or apply motion models based on assumptions about human motion.
+In particular, Breitenstein et al.
+\cite{breitenstein2009markovian} base their tracking on ouput from an externally trained human-specific detector; Yang et al.
+\cite{yang2009probabilistic} assume they are tracking an upright person, and perform feet and head detection; Conte et al.
+\cite{conte2010performance} group foreground fragments based on geometry of the human shape to be recognized and look for shadows often present in human surveillance scenarios; Berclaz et al.
+\cite{berclaz2009multiple} use an external detector that makes use of multiple camera views and models each human as a cylinder; Alahi et al.
+\cite{alahi2009sparsity} base their method on modelling the silhouettes of humans; Bolme et al.
+\cite{bolme2009simple} train a human specific detector; Ge et al.
+\cite{ge2009evaluation} provide their algorithm with estimates of typical human size and orientation; and Arsic et al.
+\cite{arsic2009multi} localize human feet positions.
+
 We compare SFDA and ATA results of our strategy with these methods to show that our arbitrary object framework can yield comparable results even when compared with object-specific trackers.
 Table~\ref{benchmark_results_table} shows performance metric results for comparison (data published with permission from the authors of \cite{ellis_2010}).
-
-%Arsic BerclazLP Conte   Bolme_ASEF  Bolme_Cascade Bolme_Parts    Breitenstein   GE  Alahi_Ogreedy   Alahi_Olasso Anon   Yang GPUDDP4MTT(sens_t25_0.35conf)
+Our method achieves the fourth best SFDA and third best ATA.
 
@@
   \textsc{Method Name} & \textsc{SFDA}  & \textsc{ATA} \\ \hline \hline
-  Breitenstein \cite{breitenstein2009markovian} & $ 0.567265 $ & $0.30044$ \\ \hline
-  Yang \cite{yang2009probabilistic} & $ 0.552439 $ & $0.448237$ \\ \hline
-  Conte \cite{conte2010performance} & $ 0.52966  $ & $0.058003$ \\ \hline
-  \textbf{GPUDDPM} & $ \bold{0.5079} $ & $\bold{0.2996}$ \\ \hline
-  Berclaz \cite{berclaz2009multiple} & $ 0.482081 $ & $0.145195$ \\ \hline
-  Alahi 1 \cite{alahi2009sparsity} & $ 0.430043 $ & $0.041053$ \\ \hline
-  Alahi 2 \cite{alahi2009sparsity} & $ 0.419141 $ & $0.045609$ \\ \hline
-  Bolme 1 \cite{bolme2009simple} & $ 0.413812 $ & NA \\ \hline
-  Ge \cite{ge2009evaluation} & $ 0.383235 $ & $0.040118$ \\ \hline
-  Bolme 2 \cite{bolme2009simple} & $ 0.338979 $ & NA \\ \hline
-  Arsic \cite{arsic2009multi} & $ 0.177768 $ & $0.016111$ \\
+  Breitenstein \cite{breitenstein2009markovian} & $ 0.57 $ & $0.30$ \\ \hline
+  Yang \cite{yang2009probabilistic} & $ 0.55 $ & $0.45$ \\ \hline
+  Conte \cite{conte2010performance} & $ 0.53  $ & $0.06$ \\ \hline
+  \textbf{GPUDDPM} & $ \bold{0.51} $ & $\bold{0.30}$ \\ \hline
+  Berclaz \cite{berclaz2009multiple} & $ 0.48 $ & $0.15$ \\ \hline
+  Alahi 1 \cite{alahi2009sparsity} & $ 0.43 $ & $0.04$ \\ \hline
+  Alahi 2 \cite{alahi2009sparsity} & $ 0.42 $ & $0.05$ \\ \hline
+  Bolme 1 \cite{bolme2009simple} & $ 0.41 $ & NA \\ \hline
+  Ge \cite{ge2009evaluation} & $ 0.38 $ & $0.04$ \\ \hline
+  Bolme 2 \cite{bolme2009simple} & $ 0.34 $ & NA \\ \hline
+  Arsic \cite{arsic2009multi} & $ 0.18 $ & $0.02$ \\
   % Anonymous & $ 0.52311  $ & $0.246275$ \\ \hline
@@
 \end{table}
+%
+% table with more sig figs
+%
+% \begin{table}
+% \begin{tabular}[!] {| l | l | l |}
+%   \hline
+%   \textsc{Method Name} & \textsc{SFDA}  & \textsc{ATA} \\ \hline \hline
+%   Breitenstein \cite{breitenstein2009markovian} & $ 0.567265 $ & $0.30044$ \\ \hline
+%   Yang \cite{yang2009probabilistic} & $ 0.552439 $ & $0.448237$ \\ \hline
+%   Conte \cite{conte2010performance} & $ 0.52966  $ & $0.058003$ \\ \hline
+%   \textbf{GPUDDPM} & $ \bold{0.5079} $ & $\bold{0.2996}$ \\ \hline
+%   Berclaz \cite{berclaz2009multiple} & $ 0.482081 $ & $0.145195$ \\ \hline
+%   Alahi 1 \cite{alahi2009sparsity} & $ 0.430043 $ & $0.041053$ \\ \hline
+%   Alahi 2 \cite{alahi2009sparsity} & $ 0.419141 $ & $0.045609$ \\ \hline
+%   Bolme 1 \cite{bolme2009simple} & $ 0.413812 $ & NA \\ \hline
+%   Ge \cite{ge2009evaluation} & $ 0.383235 $ & $0.040118$ \\ \hline
+%   Bolme 2 \cite{bolme2009simple} & $ 0.338979 $ & NA \\ \hline
+%   Arsic \cite{arsic2009multi} & $ 0.177768 $ & $0.016111$ \\
+%   % Anonymous & $ 0.52311  $ & $0.246275$ \\ \hline
+%   \hline
+% \end{tabular}
+% \caption{SFDA and ATA performance metric results are shown for our method (in bold) and for ten other algorithms on the PETS2009/2010 benchmark dataset.
+Results are listed in descending order of the SFDA value.
+Results are provided by the authors of \cite{ellis_2010}.}
+% \label{benchmark_results_table}
+% \end{table}
 
@@
 
-SMC inference on the PETS2009/2010 video dataset was carried out for a range of two key parameter values, $\alpha$ and $\rho$.
+SMC inference on the PETS2009/2010 video dataset was carried out for a range of the Generalized Polya Urn parameter values, $\alpha$ and $\rho$.
 The performance metric measures, SFDA and ATA, were computed for each combination of these two parameters.
@@
 
-Figure~\ref{fig:sens_surf} shows a heatmap and contour plot of the sensitivity analysis results.
+The SFDA and ATA achieve their maximal values at different $\alpha$ and $\rho$ parameters, though both achieve a reasonably optimal value at the intermediate parameter values $\alpha = 10$ and $\rho = 0.85$.
+Detection and tracking performance was also shown to be fairly robust to minor variations in these parameter values.
+%Figure~\ref{fig:sens_surf} shows a heatmap and contour plot of the sensitivity analysis results.
 From this figure, we can see that the SFDA and ATA achieve their maximal values at different $\alpha$ and $\rho$ parameters, though both achieve a reasonably optimal value at the intermediate parameter values $\alpha = 10$ and $\rho = 0.85$.
@@
 
-\begin{figure*}[!]
-  \centering             
-  \subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2009full_sens_sfda_contour5.pdf}}
-  \hspace{6mm}
-  \subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2009full_sens_ata_contour5.pdf}}
-  \caption{Heatmap and contour plot showing the sensitivity of performance metrics SFDA and ATA to variation of parameters $\alpha$ and $\rho$.}
-  \label{fig:sens_surf}
-\end{figure*}
+% \begin{figure*}[!]
+%   \centering             
+%   \subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2009full_sens_sfda_contour5.pdf}}
+%   \hspace{6mm}
+%   \subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2009full_sens_ata_contour5.pdf}}
+%   \caption{Heatmap and contour plot showing the sensitivity of performance metrics SFDA and ATA to variation of parameters $\alpha$ and $\rho$.}
+%   \label{fig:sens_surf}
+% \end{figure*}
 
@@
 
-We have presented a new framework for the unsupervised detection and tracking of arbitrary objects in vid-\\eos.
+We have presented a new model for the unsupervised detection and tracking of arbitrary objects in videos.
 The primary intention of this technique is to reduce the need for detection or localization methods tailored to specific object types and serve as a general framework applicable to videos with varied objects, backgrounds, and film qualities.
@@
 
+% to push this onto new page--should probably change if draft changes significantly
+\newpage
+
 \section{-  Appendix: Model Background}
@@
 
-Dirichlet process mixture (DPM) models (Section~\ref{sec:dpmixture}) fall under the heading of Bayesian nonparametric (BNP) models.
+% Dirichlet process mixture (DPM) models (Section~\ref{sec:dpmixture}) fall under the heading of Bayesian nonparametric (BNP) models.
 These models have been widely used in the past decade to perform nonparametric density estimation and cluster analysis.
@@
 
-This section provides background information that allows us to introduce DDPM models and describes the motivation behind our use of these models.
+% This section provides background information that allows us to introduce DDPM models and describes the motivation behind our use of these models.
 This background is helpful for the definition of the particular DDPM chosen, the Generalized Polya Urn dependent Dirichlet process mixture (GPUDDPM) model, given in Section~\ref{sec:gpuddpm}.
+
+We give background on mixture models, Bayesian mixture models, the Dirichlet process, Dirichlet process (infinite) mixture models, and dependent Dirichlet process mixture models.
 
@@
 \vspace{2mm}
-\item $N_{\text{mapped}}$ denotes the number of mapped ground-truth/detect pairs in a video, and $N_{\text{mapped}}^{(t)}$ denotes the number of mapped ground-truth/detect pairs present at frame $t$.
+\item $N_{\text{mapped}}$ denotes the number of mapped ground-truth/ detect pairs in a video, and $N_{\text{mapped}}^{(t)}$ denotes the number of mapped ground-truth/detect pairs present at frame $t$.
 \vspace{1mm}

### 2012-03-23 newStuff_ltx/litReview.tex
added newStuff_ltx folder to hold lit review for new things; made some final edits on paper

% header stuff
% ------------

\documentclass{article}
\usepackage[]{natbib}
\usepackage{amsmath, epsfig}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{easybmat}
\usepackage{footmisc}
\usepackage[usenames,dvipsnames]{color}
\usepackage{subfig}
%\usepackage{fullpage} % Include this if you want to cram lots of things on a page 
\renewcommand\algorithmiccomment[1]{// \textit{#1}}
\newcommand{\ignore}[1]{}
\newcommand{\comment}[1]{}
\newcommand{\frank}[1]{\textcolor{red}{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}}}
\newcommand{\willie}[1]{\textcolor{green}{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}}}
\DeclareMathOperator*{\argmax}{arg\,max}

\begin{document}

% title and author related
% ------------------------

\title{Literature Review: Graphical Models for Articulated Forms, Humans, and Other Objects}
\author{Willie}
\maketitle

% introduction
% ------------

\begin{abstract}
This is a review of graphical models for representing humans and other objects. Models of both deformable and non-deformable objects allow us, given image and video data, to infer the pose and/or orientation of objects. We aim to either learn the structure and pose of objects, or assume a structure and learn the pose of objects, and incorporate this appearance model into the GPUDDPM detection and tracking scheme.
\end{abstract}

\section{Data}
We take the data to be the same as in the previous paper: frame difference (i.e. foreground, i.e. silhouette) pixel positions, and additional local appearance features

\section{Sigal Thesis}

This thesis (\cite{Sigal:2008:CGM:1467812}) involves articulated pose estimation and tracking using graphical models.
Nodes of the graph correspond to parts or limbs of the body.
Edges correspond to ``kinematic, inter-penetration, and occlusion constraints imposed by the structure of the body and the imaging process".
The model allows one to infer 3D pose from multiple synchronized views, or 2D pose from a single monocular image. 
Sigal also develops a hierarchical model where 2D pose is inferred from a monocular source, and then 3D pose is inferred given this 2D pose; finally, tracking is carried out in 3D.
Inference is carried out using Particle Message Passing (PaMPas).

\subsection{Chapter 4: graphical models for rigid objects}

Sigal states, ``we combine techniques from discriminative learning, graphical models, belief propagation, and particle filtering to achieve reliable multi-component object detection and tracking." [p98]

\subsection{Chapter 5: graphical models for loose-limbed objects}
\label{sec:sigal_chap5}

Each node represents the 3D position and orientation of a body part (in a 6D continuous domain).
This approach also incorporates bottom-up information from object detectors, which in this case are used for detection of body parts (i.e. face, heads, and limbs).
The probabilistic relationships between joints (nodes) in the loose-limbed graphical model are specified using mixture models learned from a database of motion capture sequences.
The model also incorporates an image likelihood for each limb. The domain of the likelihood is foreground silhouette and edge features.
Sigal presents two loose-limbed models for humans--a 10 node model, and a finer, 15 node model (which also has more dependencies).

Kinematic constraints refer to the geometry of and position between limbs. Kinematic constraints between limb positions [p123] and between limb orientations [p124] are modeled by mixtures of Gaussians.
For limb orientations, this mixtures of Gaussians representation can be converted back into the rotation group, SO(3), space in which the orientation representation resides.
These contraints are learned from ground truth motion capture sequences.

Penetration contraints refer to the inability for solid limbs to penetrate each other.
Sigal only constrains limbs that are likely to come into contact in his model.
Penetration constraints are learned computationally by making a model human out of basic shapes and and computing how much given parts have a tendency to intersect.

Sigal defines image likelihoods which give the probability of image data given the pose of a limb.
A main component of the domain over which this likelihood is defined is binary foreground / silhouette.
For each limb, he assumes three groups of pixels are present in the binary foreground image: those corresponding with a limb, those correlated with a limb, and those completely uncorrelated with the limb; these groups are used to define his limb likelihood on foreground pixel positions.
He also defines an edge likelihood (edge data gathered from Canny edge transformation of the image) that is a function of distances between edges.
Note that this likelihood is non-clothes-specific.

Sigal also incorporates part detectors (such as head and limb detection), referred to as ``shouters", which give noises guesses as to the positions of parts of the body.

Belief propagation is used for inference; in particular, Sigal uses a nonparametric algorithm called Particle Message Passing (PaMPas), which he also says is a generalization of a particle filter (that allows for inference over arbitrary graphs and not necessarily chains).

\subsection{Chapter 6: hierarchical model for 3D pose from monocular source}

Inferred 2D pose from monocular sources gives proposals in inference of 3D pose.

A mixture of experts model is used to learn a mapping from inferred 2D poses to 3D poses (including joint angles and foreshortening / perspective information). In particular, Sigal uses a mixture of regularized linear regression models that are trained from a set of 2D-3D pose pairs obtained from motion capture data.

The main difference between the 2D model used here and the loose-limbed model described in Section~\ref{sec:sigal_chap5} is that Sigal develops ``Occlusion-Sensitive Local Likelihoods'', which incorporates whether foreground / edge pixels could be explained by other body parts in the image.

\section{Sudderth Paper}

In this paper (\cite{Sudderth04distributedocclusion}), Sudderth presents a three-dimensional geometric hand model that incorporates geometric contraints, has an image likelihood defined on color and edge data, and infers a model of the hand using a nonparametric inference algorithm (similar to that used by Sigal). Various information about hand geometry and results of edge detection on videos of hands are gained from training data. This model incorporates kinematic contraints between parts of the hand, hand dynamics, and occlusion constraints / variables.

\section{Ying Wu Paper}

In this paper (\cite{wu2003tracking}), articulated bodies are modeled and tracked via a dynamic markov random field. This paper begins by defining how general constraints between articulated parts are encoded. In the experimental section, Wu defines a 2D ``cardboard" model where each subpart of the human body is represented as a planar object. For his motion model, there are dependencies between adjacent time steps for each limb. The image observation associated with each limb is taken to be the detected edges of the shape contour of the limb. A sequential mean field monte carlo algorithm is used to perform inference in this model.

\section{Plan of Attack}

The papers reviewed here provide models that encode information about certain articulated bodies (humans, hands) and give strategies for inferring correct positions of these bodies from image data (either foreground pixels, edges, or colors). Dependencies in these papers were primarily defined using undirected graphical models.

For incorporation into the GPUDDPM, we need to define some articulated body likelihood structure with parameters $\theta_{k,t}$ for frame difference pixel positions $\bold{x}_{i,t}$, and the distributions
\begin{equation}
\begin{split}
\bold{x}_{i,t} &\sim \text{F}(\theta_{k,t}) \\
\theta_{k,t} &\sim \mathbb{G}_{0}(\theta_{k,t}) \\
P(&\theta_{k,t} | \theta_{k,t-1})
\end{split}
\end{equation}

Additionally, it would be cool if we could
\begin{enumerate}
\item create a generative pose model to sample the pose of articulated bodies. I'm not sure if these undirected graphical model setups allow us to do this.
\item have a model which learns articulated structure from frame differencing data. Might be possible since we have an unsupervised system which can segment individual objects. It would probably be hard to capture specific dependencies / occlusion relations between limbs.
\item have a model that learns types of objects based on the different structures it infers.
\end{enumerate}

\begin{small}
\bibliographystyle{plainnat}
\bibliography{../ltx/paper_refs} 
\end{small}

\end{document}

### 2012-03-27 ltx/paper_short_2.tex
submitted to IJCV. files in submitted_ijcv_20120327 directory

@@
 This paper describes a specific implementation of the model using spatial and color pixel data extracted via frame differencing and gives two algorithms for performing inference in the model to accomplish detection and tracking.
-This technique is demonstrated on multiple synthetic and benchmark video datasets that illustrate its ability to, without modification, detect and track objects with diverse physical charactersitics moving over non-uniform backgrounds and through occlusion.
+This technique is demonstrated on multiple synthetic and benchmark video datasets that illustrate its ability to, without modification, detect and track objects with diverse physical characteristics moving over non-uniform backgrounds and through occlusion.
 % \keywords{Object Detection \and Tracking \and Bayesian Nonparametrics \and Multiple Target Tracking}
@@
 Standard performance metric values are computed to quantify results on the benchmark datasets (Section~\ref{sec:experiments}).
-We compare our performance metrics with those yielded by specialized detection and tracking algorithms tailored to specific objects.
+We compare the performance metrics of our method with those yielded by specialized detection and tracking algorithms tailored to specific objects in the benchmark datasets.
 Our results support our hypothesis that approaches combining simple data extraction and a powerful model can perform detection and tracking of arbitrary objects at a level comparable to state-of-the-art, object specific algorithms.
@@
 
-Methods that discern between the foreground and background regions of a video allow for data to be extracted from the areas in each frame where objects exist.
+Methods that discern between the foreground and background regions of a video allow for data to be extracted from the areas in each frame where objects reside.
 Frame differencing and background subtraction are two such methods.
@@
 Blob tracking methods, however, run into problems when faced with videos where detection is difficult, object appearance or orientation varies heavily, and there exists object occlusion \cite{song2005model}.
-To aid the accuracy of these methods, techniques have been developed for performing extraction and segmentation in a joint manner, incorporating statistical methods for maintaining hypotheses of different numbers of detected objects, and introducing some of the multi-object tracking methods described previously to track distinct blobs after they have been segmented \cite{collins2003mean, isard_2001}.
+To improve the accuracy of these methods, techniques have been developed for performing extraction and segmentation in a joint manner, incorporating statistical methods for maintaining hypotheses of different numbers of detected objects, and introducing some of the multi-object tracking methods described previously to track distinct blobs after they have been segmented \cite{collins2003mean, isard_2001}.
 Another family of methods related to the task of unsupervised detection and tracking of video objects goes under the heading of video segmentation algorithms---these methods extend single-frame image segmentation to maintain coherence of image segments over time, and have had some success when used for the explicit purpose of detecting and tracking foreground objects in videos \cite{brox2003unsupervised, sista2000unsupervised, wang1998unsupervised}.
@@
 \end{equation}
-where each $\bold{x}$ corresponds to a point within an image region where an object (or foreground element)  is believed to exist, $\bold{x}^{s} \in \mathbb{R}^{2}$ denotes the spatial location of this point, $\bold{x}^{c} \in S_{1} \times \ldots \times S_{V}$ denotes some collection of local image features in the vicinity of this point, and $t \in \{1, \ldots, T \}$ denotes the time index.
+where each $\bold{x}$ corresponds to a point within an image region where an object (or foreground element)  is believed to reside, $\bold{x}^{s} \in \mathbb{R}^{2}$ denotes the spatial location of this point, $\bold{x}^{c} \in S_{1} \times \ldots \times S_{V}$ denotes some collection of local image features in the vicinity of this point, and $t \in \{1, \ldots, T \}$ denotes the time index.
 
@@
 In our implementation, we choose to extract only color information in the vicinity of each pixel.
-Incorporating color features allows our method to infer a distribution over color for each detected object; this improves its ability to distingush between adjacent objects and track objects through occlusion.
+Incorporating color features allows our method to infer a distribution over color for each detected object; this improves its ability to distinguish between adjacent objects and track objects through occlusion.
 To add this information, we let $\bold{x}^{c}$ represent a $V$ dimensional discrete distribution over some aspect of color (such as the hue) in the pixel's immediate vicinity.
@@
 where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise.
-Additionally, we can define the deletion variable $d_{i,t}$ to be $d_{i,t} = t + l_{i,t}$, where $l_{i,t}$ can be thought of as the lifespan of an assignment.
-From the definition of distribution D (given by \eqref{D_distro}), the lifespan can be shown to be distributed geometrically, and can be written as
+Additionally, we can define the deletion variable $d_{i,t}$ to be $d_{i,t} = t + l_{i,t}$, where $l_{i,t}$ can be thought of as the lifetime of an assignment.
+From the definition of distribution D (given by \eqref{D_distro}), the lifetime can be shown to be distributed geometrically, and can be written as
 \begin{equation}
@@
 
-\subsubsection{Sampling Assignment Variables $c_{i,t}$}
+\subsubsection{Sampling Assignment Variables, $c_{i,t}$}
 \label{sec:sample_assignments}
@@
 
-\subsubsection{Sampling Cluster Parameters $\theta_{k, t}$}
+\subsubsection{Sampling Cluster Parameters, $\theta_{k, t}$}
 
@@
 \end{align}
-Where the parameters in the above distribution are given when the observations $\bold{x}_{1:N_{t},t}$, auxiliary variables $z_{k,t-1:t,1:M}$ for cluster k at time $t-1$ and $t$, are taken to be the ``observations'' for the following Bayesian updates
+Where the parameters in the above distribution are given when the observations $\bold{x}_{1:N_{t},t}$, and auxiliary variables $z_{k,t-1:t,1:M}$ for cluster k at time $t-1$ and $t$, are taken to be the ``observations'' for the following Bayesian updates
 \begin{eqnarray}
@@
 
-\subsubsection{Sampling Auxiliary Variables $z_{k,t,m}$}
+\subsubsection{Sampling Auxiliary Variables, $z_{k,t,m}$}
 \label{sec:sample_aux_vars}
@@
 
-\subsubsection{Sampling Deletion Variables $d_{i,t}$}
+\subsubsection{Sampling Deletion Variables, $d_{i,t}$}
 Sampling deletion variables could be performed in a manner similar to how we sample the assignment variables in Section~\ref{sec:sample_assignments}, but this may be computationally expensive due to the large number of possible deletion times.
@@
 The algorithm is shown in Algorithm~\ref{alg:SMC}.
-At each time step $t \in \{ 1, \ldots, T \}$, a number of samples refered to as ``particles'' are generated; each particle consists of a sample from the posterior distribution over the assignment for each observation, $c_{1,t}, \ldots, c_{N_{t}, t}$, parameters for each cluster $\theta_{1,t}, \ldots,$ $\theta_{K_{t}, t}$, and size after deletion for each cluster $m_{1,t},$ $\ldots,$ $m_{K_{t},t}$.
+At each time step $t \in \{ 1, \ldots, T \}$, a number of samples referred to as ``particles'' are generated; each particle consists of a sample from the posterior distribution over the assignment for each observation, $c_{1,t}, \ldots, c_{N_{t}, t}$, parameters for each cluster $\theta_{1,t}, \ldots,$ $\theta_{K_{t}, t}$, and size after deletion for each cluster $m_{1,t},$ $\ldots,$ $m_{K_{t},t}$.
 A set of particles is sampled at each time step from relevant proposal distributions (described in Sections~\ref{sec:smc_proposal_1}, \ref{sec:smc_proposal_2}, and \ref{sec:smc_proposal_3}), a weight is computed for each particle, and a new set of particles are sampled from the set of weighted particles via a resampling process.
@@
 
-\subsubsection{Proposal Distribution for Assignments $c_{i,t}$}
+\subsubsection{Proposal Distribution for Assignments}
 \label{sec:smc_proposal_1}
@@
 Performance evaluation metrics, which provide a standardized way of quantifying the success of a detection and tracking procedure on a given video, have started to become consistently used in the past four years.
-The metrics presented in \cite{kasturi_2008} and used in \cite{ellis_2010, taj_2007, lee_2009} have become well established metrics for evaluating the performance of object detection and tracking in videos and have been adopted by the Video Analysis and Content Extraction (VACE) program and the Classification of Events, Activities, and Relationships (CLEAR) consortium, two large-scale efforts concerned with video tracking and interaction analysis.
+The metrics presented in \cite{kasturi_2008} and used in \cite{ellis_2010, taj_2007, lee_2009} have become well established for evaluating the performance of object detection and tracking in videos and have been adopted by the Video Analysis and Content Extraction (VACE) program and the Classification of Events, Activities, and Relationships (CLEAR) consortium, two large-scale efforts concerned with video tracking and interaction analysis.
 The two metrics used to quantify the experimental results in this study are known as the Sequence Frame Detection Accuracy (SFDA) and Average Tracking Accuracy (ATA).
@@
 
-The ground-truth authored by the ViPER tool took the form of bounding boxes denoting the spatial postion of each object at each time step.
+The ground-truth authored by the ViPER tool took the form of bounding boxes denoting the spatial position of each object at each time step.
 Consequentially, to find the spatial overlap between results and ground-truth, which is intrinsic to both metrics, a rectangular bounding box was needed per object per time step from the results of the algorithm.
@@
 the range of hues to which a pixel may be assigned) was partitioned into $V$ bins, and the number of pixels with a color value lying in each of the bins yielded the $V$ dimensional vector of color counts.
+For all experiments, we chose $V=10$.
 
@@
 
-Each of the following synthetic videos consists of a sequence of 200 images (each of size $500$ x $500$ pixels) containing a number of smaller colored squares of different (and potentially time-varying) sizes moving at varied speeds and trajectories over a black background.
+Each of the following synthetic videos consists of a sequence of 200 images (each of size $500$ $\times$ $500$ pixels) containing a number of smaller colored squares of different (and potentially time-varying) sizes moving at varied speeds and trajectories over a black background.
 The synthetic videos contain instances of occlusion (where one or more objects are briefly hidden) and objects with time-varying appearances and behaviors, as these notoriously decrease the accuracy of detection and tracking.
-After each video was constructed, the extraction procedure described in Section~\ref{sec:dataextraction} (using $L=3$) and inference procedures described in Section~\ref{sec:inference} were carried out to return a sequence of multivariate-normal-parameters (means and covariance matrices), which are used to determine a sequence of positions and ovals approximating, respectively, the locations and shapes of a tracked object over each frame that it is present in the video (as outlined in Section~\ref{sec:inferencetotrackingresults}).
+After each video was constructed, the extraction procedure described in Section~\ref{sec:dataextractioninexperiments} (using $L=3$) and inference procedures described in Section~\ref{sec:inference} were carried out to return a sequence of multivariate-normal-parameters (means and covariance matrices), which are used to determine a sequence of positions and ovals approximating, respectively, the locations and shapes of a tracked object over each frame that it is present in the video (as outlined in Section~\ref{sec:inferencetotrackingresults}).
 
 The first synthetic video experiment aimed to test the ability of the model and inference procedure to maintain the identity of independent objects based on color information alone.
-Two videos were constructed, both containing a red square (rgb value $[255,0,0]$ and size $20$ x $20$ pixels) and a blue square (rgb value $[0,0,255]$ and size $20$ x $20$ pixels).
+Two videos were constructed, both containing a red square (rgb value $[255,0,0]$ and size $20$ $\times$ $20$ pixels) and a blue square (rgb value $[0,0,255]$ and size $20$ $\times$ $20$ pixels).
 In both videos, the squares begin at opposite sides of the scene at frame $f=1$ and travel towards each other, arriving at the same location at $f=100$ (where the blue square occludes the red square).
-The second half of the two videos differ in that both squares in the first video continue in the same direction and end at the other's starting positions at $f=200$, and both squares in the second video reverse directions and end at their initial starting positions at $f=200$.
+The second half of the two videos differ in that both squares in the first video continue in the same direction and end at the other's starting position at $f=200$, and both squares in the second video reverse directions and end at their initial starting positions at $f=200$.
 The frame difference extraction yields identical spatial features in both videos; hence, successful tracking depends fully on the incorporation of color information into the model.
 
-Parameters were set to the same values for inference on both videos: $\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right)$, and $q_{0} = (5, \ldots, 5)$.
+Parameters were set to the same values for inference on both videos: $\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right)$, and $\bold{q}_{0} = (5, \ldots, 5)$.
 Inference was carried out using the MCMC algorithm (Section~\ref{sec:MCMC}); the MAP sample correctly tracked both colored squares through occlusion in both videos, and is shown in Figure~\ref{fig:synth_one_plot}.
@@
 A video was constructed showing a red square (rgb value $[255,0,0]$), a green square (rgb value $[0,255,0]$), and a blue square (rgb value $[0,0,255]$).
-The red square was of size $20$ x $20$ pixels, the blue square was of size $15$ x $15$ pixels, and the green square began at size $50$ x $50$ pixels at frame $f=1$, linearly shrinks to $10$ x $10$ pixels at $f=100$, then linearly grows back to $50$ x $50$ pixels by the end of the video, $f=200$.
+The red square was of size $20$ $\times$ $20$ pixels, the blue square was of size $15$ $\times$ $15$ pixels, and the green square began at size $50$ $\times$ $50$ pixels at frame $f=1$, linearly shrinks to $10$ $\times$ $10$ pixels at $f=100$, then linearly grows back to $50$ $\times$ $50$ pixels by the end of the video, $f=200$.
 Furthermore, the red and blue squares display the same behavior as in the second video of the first synthetic experiment (they begin at opposite sides of the scene traveling towards each other, cross at the center of the scene at $f=100$, and reverse direction, ending at their initial positions at $f=200$).
@@
 Benchmark video datasets for object tracking and detection have been produced to provide standard scenes on which researchers can compare detection and tracking results.
-These videos have been primarily produced for surveillance-related workshops---notably, for the International Workshop on Performance Evaluation of Tracking and Surveillance (PETS)--which provide researchers with video datasets and algorithmic goals on which to focus.
+These videos have been primarily produced for surveillance-related workshops---notably, for the International Workshop on Performance Evaluation of Tracking and Surveillance (PETS)---which provides researchers with video datasets and algorithmic goals on which to focus.
 Three commonly used benchmark videos from PETS workshops---one used in PETS2000, one in PETS2001, and one used both in PETS2009 and PETS2010---were chosen to demonstrate the method presented in this study.
@@
 The PETS2000 and PETS2001 video datasets both consist of a small number of humans and vehicles traveling across a parking lot, with video taken from above, emulating what might be recorded by standard outdoor surveillance equipment.
-The `Test Sequence', a set of images from a monocular, stationary camera, was used from the PETS2000 workshop, and View Two of Dataset 1, also taken via a monocular, stationary camera, was used from the PETS2001 workshop.
+The ``Test Sequence'', a set of images from a monocular, stationary camera, was used from the PETS2000 workshop, and ``View Two of Dataset 1", also taken via a monocular, stationary camera, was used from the PETS2001 workshop.
 The MCMC algorithm (described in Section~\ref{sec:MCMC}) was used for inference in these experiments.
@@
 Extraction was performed with frame differencing as described in Section~\ref{sec:dataextractioninexperiments}, using $L=3$.
-Parameter values were set to the same values as in the the synthetic experiments ($\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (5, \ldots, 5)$).
+Parameter values were set to the same values as in the synthetic experiments ($\alpha = 0.1, \rho = 0.3, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 5, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } \bold{q}_{0}$ $=$ $(5,$ $\ldots,5)$).
 The MCMC sampler was successful for both benchmark videos; each object was detected, tracked, and its shape estimated in manner very consistent with the ground-truth.
@@
 
-Extraction was performed with frame differencing as described in Section~\ref{sec:dataextractioninexperiments}, using $L=3$, and parameters for the model were chosen to be $\alpha = 0.1, \rho = 0.8, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 6, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } q_{0} = (3, \ldots, 3)$.
+Extraction was performed with frame differencing as described in Section~\ref{sec:dataextractioninexperiments}, using $L=3$, and parameters for the model were chosen to be $\alpha = 0.1, \rho = 0.8, M = 10, \boldsymbol{\mu}_{0} = (0,0), \kappa_{0} = 0.05, \nu_{0} = 6, \Lambda_{0} = \left( \begin{smallmatrix} 1&0\\ 0&1 \end{smallmatrix} \right), \text{ and } \bold{q}_{0} = (3, \ldots, 3)$.
 Additionally, as with the other video datasets, ground-truth bounding boxes around each object were authored using the ViPER tool.
@@
 Results are listed in descending order of the SFDA value.
-Results are provided by the authors of \cite{ellis_2010}.}
+The results were provided by the authors of \cite{ellis_2010}.}
 \label{benchmark_results_table}
@@
 We have described and computed standard performance metrics for our technique's detection and tracking results, and found it to be comparable with state-of-the-art object-specific detection and tracking methods designed for people tracking in the PETS2009/2010 video dataset.
-Results from the synthetic and benchmark video datasets illustrate the ability of the technique described in this paper to, without modification, perform completely unsupervised detection and tracking of objects with diverse physical charactersitics moving over non-uniform backgrounds and through occlusion.
+Results from the synthetic and benchmark video datasets illustrate the ability of the technique described in this paper to, without modification, perform completely unsupervised detection and tracking of objects with diverse physical characteristics moving over non-uniform backgrounds and through occlusion.
 

### 2012-03-27 submitted_ijcv_20120327/neiswanger_wood_unsupervisedDetectionWithDDPM.tex
submitted to IJCV. files in submitted_ijcv_20120327 directory

--- 2012-03-27 ltx/paper_short_2.tex
@@
 A few attempts at accomplishing the three tasks in a cohesive manner have been studied in recent years \cite{brostow2006unsupervised, brox2010object, fragkiadaki2011detection, pece_2002}.
-%\willie{An example of general detection and tracking, called blob tracking, typically refers to methods that extract data and localize objects in each frame independently.
-Tracking involves finding blobs with similar locations and appearances over time.
-The main blob tracking challenges are the difficulty of segmenting neighboring or occluding objects in a given frame \cite{zhao2004tracking} and preventing multiple tracked blobs from merging into a single object \cite{vermaak_2003}.
-These difficulties stem from performing detection via segmentation on individual frames.}
 This paper furthers this line of work by providing a model that could give rise to a number of algorithms to detect arbitrary objects in videos---particularly in cases where frame-by-frame segmentation is difficult, video quality is low, and extraction is noisy---and maintain the isolation of distinct objects during tracking and through occlusion.
@@
   \centering
-  \subfloat[]{\label{fig:ants_img}\includegraphics[width=0.33\textwidth]{../img/usbh_frame_723.pdf}} \hspace{0.1mm}
-  \subfloat[]{\label{fig:ants_img2}\includegraphics[width=0.33\textwidth]{../img/usbh_frame_724.pdf}} \hspace{0.1mm}
-  \subfloat[]{\label{fig:ants_img_framediff}\includegraphics[width=0.323\textwidth]{../img/usbh_framediff.png}}\\
-  \subfloat[]{\label{fig:traffic2_img}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame1.png}} \hspace{0.1mm}
-  \subfloat[]{\label{fig:traffic2_img2}\includegraphics[width=0.328\textwidth]{../img/traffic2_frame2.png}} \hspace{0.1mm}
-  \subfloat[]{\label{fig:traffic2_img_framediff}\includegraphics[width=0.3273\textwidth]{../img/traffic2_framediff.png}}\\
-  \subfloat[]{\label{fig:pets2009_img}\includegraphics[width=0.328\textwidth]{../img/pets2009_frame1.png}} \hspace{0.1mm}
-  \subfloat[]{\label{fig:pets2009_img2}\includegraphics[width=0.328\textwidth]{../img/pets2009_frame2.png}} \hspace{0.1mm}
-  \subfloat[]{\label{fig:pets2009_img_framediff}\includegraphics[width=0.328\textwidth]{../img/pets2009_framediff.png}}\\
-  \subfloat[]{\label{fig:longimg_all}\includegraphics[width=1\textwidth]{../img/longimg_7_detail_7.pdf}}
+  \subfloat[]{\label{fig:ants_img}\includegraphics[width=0.33\textwidth]{usbh_frame_723.pdf}} \hspace{0.1mm}
+  \subfloat[]{\label{fig:ants_img2}\includegraphics[width=0.33\textwidth]{usbh_frame_724.pdf}} \hspace{0.1mm}
+  \subfloat[]{\label{fig:ants_img_framediff}\includegraphics[width=0.323\textwidth]{usbh_framediff.png}}\\
+  \subfloat[]{\label{fig:traffic2_img}\includegraphics[width=0.328\textwidth]{traffic2_frame1.png}} \hspace{0.1mm}
+  \subfloat[]{\label{fig:traffic2_img2}\includegraphics[width=0.328\textwidth]{traffic2_frame2.png}} \hspace{0.1mm}
+  \subfloat[]{\label{fig:traffic2_img_framediff}\includegraphics[width=0.3273\textwidth]{traffic2_framediff.png}}\\
+  \subfloat[]{\label{fig:pets2009_img}\includegraphics[width=0.328\textwidth]{pets2009_frame1.png}} \hspace{0.1mm}
+  \subfloat[]{\label{fig:pets2009_img2}\includegraphics[width=0.328\textwidth]{pets2009_frame2.png}} \hspace{0.1mm}
+  \subfloat[]{\label{fig:pets2009_img_framediff}\includegraphics[width=0.328\textwidth]{pets2009_framediff.png}}\\
+  \subfloat[]{\label{fig:longimg_all}\includegraphics[width=1\textwidth]{longimg_7_detail_7.pdf}}
   \caption{Three pairs of consecutive frames and the results produced by taking the pixel-wise frame difference between each pair (a - i).
@@
 \label{sec:dataextraction}
-
-% The basic set of observations extracted from video are the three-dimensional points
-% \begin{equation}
-%   \centering
-%   \bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \{ 1, \ldots, T \}
-% \end{equation}
-% where $x_{1}$ and $x_{2}$ are the two spatial locations of a pixel and $t$ is the time index.
-These data are extracted by recording the positions of pixels lying in regions that contain objects.
-For our implementation in experiments (Section~\ref{sec:experiments}), we desire a data extraction procedure that is as unsophisticated as possible, to verify that we can perform accurate inference on noisy data extracted using a broadly applicable procedure.
-Consequentially, we carry out simple frame differencing (introduced in Section~\ref{sec:introduction}) for all experiments, to find pixels at locations where there is motion.
-This data extraction procedure is simple, computationally inexpensive, and able to be applied to a wide range of static, single-camera videos (note that videos used in all experiments were stationary; moving-camera videos require data extraction methods useful for non-stationary video, such as \cite{chien2002efficient, zhang2007moving}).
-Examples of pixel location data extracted via frame differencing are shown in Figure~\ref{fig:img_and_framediff}(a-j).
 
@@
 
-% old discussion on incorporating color features
-
-% The frame differencing and color counting extraction yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \{1, \ldots, T \}$, where each element $\bold{x} \in \bold{X}$ can be written
-% \begin{equation}
-% \bold{x} = ( \bold{x}^{s}, \bold{x}^{c}, t ) = ( x^{s_{1}}, x^{s_{2}}, x^{c_{1}}, \ldots, x^{c_{V}}, t )
-% \end{equation}
-% where $\bold{x}^{s} \in \mathbb{R}^{2}$ is the two dimensional vector of spatial coordinates,  $\bold{x}^{c} \in \mathbb{Z}_{+}^{V}$ is the $V$ dimensional vector of color counts, and $t \in \{1, \ldots, T \}$ is the time index.
-% Other features may also be extracted from each pixel to attempt more sophisticated object representations, including pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, and texture representations.
-
-% old (long) data extraction - parts grabbed for new version above
-% ----------------------------------------------------------------
-
-% We defined the task of extraction to be the unsupervised determination of which spatiotemporal regions of a video constitute objects, and outlined a number of extraction methods in Section~\ref{sec:priorwork}.
-The most basic dataset that can be gained from extraction is comprised of the three-dimensional points $\bold{x} = ( x_{1}, x_{2}, t ) \in \mathbb{R}^{2} \times \mathbb{Z}_{+}$, where $x_{1}$ and $x_{2}$ are the two spatial dimensions of a pixel contained in a region found during extraction and $t$ is the frame in which the pixel resides.
-Once this dataset, which we will refer to as a ``baseline extraction dataset'', has been found, additional features may also be extracted from the video.
-Examples of additional features include color information, pixel intensity values, feature point (such as corner, shape, or edge) locations or spatial characteristics, and texture representations; the main idea is to choose features which are able to be easily extracted or computed, capture variability in the appearance of objects, and are applicable to a wide variety of object types.
-
-% The baseline extraction dataset has the appearance of worm-like clusters that follow the paths of individual objects as they move (as illustrated in Figure~\ref{fig:longimg_all}).
-Modeling this data with a dependent Dirichlet process mixture model (Section~\ref{sec:gpuddpmixture}) allows us to perform inference (Section~\ref{sec:inference}), which carries out clustering of the data.
-By choosing this type of model, we aim to provide a clustering result that isolates each worm-like cluster and infers a sequence of distributions for each, which we use to gain localization and tracking results for each object (Section~\ref{sec:inferencetotrackingresults}).
-
-% For implementation in experiments (Section~\ref{sec:experiments}), we desire an extraction procedure that is as unsophisticated as possible, both to gauge the robustness of this method on potentially noisy extraction data and to ensure that the procedure is applicable to a wide range of videos.
-Consequently, in all experiments we carry out a method called ``frame-differencing'', which locates motion by recording the positions of pixels that have exhibited differences in intensity or value in succesive frames beyond a given threshold.
-Frame differencing is unsophisticated, computationally inexpensive, able to be applied to a wide range of static, single-camera videos (note that videos used in all experiments were stationary; moving-camera videos require extraction methods that provide accurate foreground segmentation for non-stationary video).
-We have found that, when correctly implemented, frame differencing is sufficiently general to extract the desired worm-like clusters from a variety of videos containing multiple moving objects.
-A few examples of this extraction procedure on pairs of consecutive video frames can be found in Figure~\ref{fig:img_and_framediff}(a-i).
-
-% We furthermore wish to capture the color (or grayscale) value of each pixel and incorporate this color information into our model.
-For each pixel $x = (i, j, t)$ recorded during frame differencing, we specify a square, $L$ pixels in length, centered on $(i, j)$, that selects a set of pixels surrounding $\bold{x}$ in frame $t$.
-We also specify a scalar color value that is able to be computed for each pixel; this value could, for example, be some function of the red-green-blue (r-g-b) or hue-saturation-value (h-s-v) characteristic of a pixel.
-The color value of each of the selected pixels surrounding $\bold{x}$ is recorded.
-Afterwards, the set of possible color values (ie the range of color values to which a pixel may be assigned) is partitioned into $V$ bins, and the number of pixels with a color value lying in each of the bins yields a $V$ dimensional vector of `color counts'.
-We will refer to this extraction technique as `color counting'.
-
-% The frame differencing and color counting extraction yields a set of data $\bold{X} \subset \mathbb{R}^{2} \times \mathbb{Z}_{+}^{V} \times \{1, \ldots, T \}$, where each element $\bold{x} \in \bold{X}$ can be written
-% \begin{equation}
-% \bold{x} = ( \bold{x}^{s}, \bold{x}^{c}, t ) = ( x^{s_{1}}, x^{s_{2}}, x^{c_{1}}, \ldots, x^{c_{V}}, t )
-% \end{equation}
-% where $\bold{x}^{s} \in \mathbb{R}^{2}$ is the two dimensional vector of spatial coordinates,  $\bold{x}^{c} \in \mathbb{Z}_{+}^{V}$ is the $V$ discrete dimensional vector of color counts, and $t \in \{1, \ldots, T \}$ is the discrete time index.
-
-% In the experiments described in Section~\ref{sec:experiments}, the hue component of the h-s-v value is recorded from all pixels that surround each extracted pixel $\bold{x}$ in the manner described above, choosing $L=5$.
-Additionally, the set of possible hue values is partitioned into 10 bins, and the number of pixels with a hue value lying in each of the 10 bins is recorded to yield the vector of ``color counts'' $\bold{x}^{c} = x^{c_{1}}, \ldots, x^{c_{10}}$.
-Hue is chosen to represent object color since it has been demonstrated in previous work as a simple representation of object appearance that allows for distinct objects to be well differentiated \cite{perez_2002, raja_1998, mckenna_1999}.
-
 % Model Definition / Framework
@@
 Consequently, we are interested in using a class of models known as dependent Dirichlet process mixture (DDPM) models (Section~\ref{sec:ddpmixture}), which are particularly useful for estimating the number of latent classes (clusters) in time dependent data.
-%In this work, estimating the number of clusters in video extraction data is equivalent to estimating the number of distinct objects in a video.
 
@@
 In a GPUDDPM, each observation $\bold{x}_{i,t}$ is associated with an assignment variable $c_{i,t}$ that represents its assignment to a cluster $\theta_{k,t}$.
-%In a DPM, the probability of assignment to a cluster $k$ is a function of the cluster size $m_{k,t}$, which is equal to the number of observations that have been assigned to it. In contrast, 
 The sizes of clusters in the GPUDDPM increase when observations are assigned to them, and decrease at later time points when these observations become ``unassigned''.
@@
 \begin{figure}[h]
-        \center{\includegraphics[width=80mm]{../img/gpuddp_gm_1.pdf}}
+        \center{\includegraphics[width=80mm]{gpuddp_gm_1.pdf}}
         \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Polya Urn dependent Dirichlet process mixture.
@@
 
-%The parameters chosen for use in experiments (Section~\ref{sec:experiments}) were accepted based on the fact that they yielded reasonable results.
-Additionally, a sensitivity analysis was carried out for the concentration and deletion parameters ($\alpha$ and $\rho$), where inference was performed for range of both values (Section~\ref{sec:sensitivityanalysis}).
-
 % Inference
@@
 This section details the MCMC sampler used to perform inference.
-% which serves as our first algorithm for unsupervised detection and tracking.
-It iteratively samples from the conditional distribution of each element in our model given all the rest---i.e.
-implements Gibbs sampling---to generate approximate samples from the posterior distribution of the model.
-Additionally, we implement the Metropolis-Hastings (MH) algorithm to sample from a few of these conditional distributions.
 A secondary formulation of the GPUDDPM, which we refer to as the ``Deletion Variable Formulation'' (defined in Section~\ref{sec:deletionvariableformulation}), is used here.
@@
 \begin{figure}[h]
-        \center{\includegraphics[width=82mm]{../img/gpuddp_gm_3.pdf}}
+        \center{\includegraphics[width=82mm]{gpuddp_gm_3.pdf}}
         \caption{Graphical model of the Deletion Variable Formulation of the GPUDDPM, showing auxiliary variables.}
@@
 The following sections detail the distributions from which each of these samples is drawn.
-
-% from older version where defined a set S
-
-% We define the set of variables, S, to consist of all observations, assignment variables, deletion variables, cluster parameters (for all active clusters), and auxiliary variables (for all active clusters).
-We can write
-% \begin{equation}
-% \text{S} = \{ \boldsymbol{x}_{1:T}, \bold{c}_{1:T}, \bold{d}_{1:T}, \Theta_{1:T}, \bold{z}_{1:T} \}
-% \end{equation}
-% where
 
@@
 where t denotes the multivariate t-distribution, where we follow the three-value parameterization (location parameter, scale parameter, and degrees of freedom) given in \cite{gelman2004bayesian, Wood2008nonparametric}, and $\{ \boldsymbol{\mu}_{0},$ $\kappa_{0},$ $\Lambda_{0},$ $\nu_{0},$ $\bold{q}_{0} \}$ are prior parameters.
-
-%%%%%
-% Need to include ratio of Dirichlet also in the above posterior predictive distro of a single observation, and discuss how to compute any parameters in here that aren't yet specified
-%%%%%
 
@@
 \end{equation}
-%fw says to remove:
-%
-%Samples cannot easily be drawn from this distribution.
-We therefore use the MH algorithm to sample auxiliary variables from the posterior distribution (for each living cluster).
-The MH algorithm involves constructing a markov chain with the sought posterior as its stationary distribution, and sampling sequentially from this constructed chain to get an approximate posterior sample.
-%
-%
-% The MH algorithm is outlined in Algorithm 1.%Algorithm~\ref{alg:MH}. \willie{am i citing wrong?}
-% \begin{algorithm}[h!]
-% \label{alg:MH}
-% \caption{Metropolis Hastings Algorithm}
-% \begin{algorithmic}[1]
-% \STATE $f(x)$ is the stationary PDF from which $N$ samples are desired \hfill $\triangleright$ eq. \eqref{stationary_pdf}
-% \STATE $q( \cdot | x_{i-1})$ is the proposal pdf \hfill $\triangleright$ eq. \eqref{proposal_distro}
-% \STATE B is the number of burn in samples
-% \STATE $x_{0} \leftarrow$ initialize
-% %\STATE B $\leftarrow$ \# of burn in samples
-% \FOR{$i = 1 : \text{B} + N$}
-% \STATE Sample $x^{*} \sim q( \cdot | x_{i-1})$
-% \STATE Sample $\alpha \sim \text{Uniform}(0,1)$
-% \STATE $r_{accept} \leftarrow \frac{f(x^{*})q(x_{i-1} | x^{*})} {f(x_{i-1})q(x^{*} | x_{i-1})} $ \hfill $\triangleright$  eq. \eqref{accept_ratio}
-% \IF{$\alpha < \text{min} \{ 1, r_{accept} \} $}
-% \STATE $x_{i} \leftarrow x^{*}$
-% \ELSE
-% \STATE $x_{i} \leftarrow x_{i-1}$
-% \ENDIF
-% \ENDFOR
-% \STATE Discard samples $x_{1}, \ldots, x_{\text{B}}$
-% \STATE \textbf{Output}: Remaining samples $x_{1}, \ldots, x_{N}$
-% \end{algorithmic}
-% %\caption{Say thing at end}
-% \end{algorithm}
-%
 We sample a new value for all $M$ auxiliary variables, denoted $z_{k,t,m}^{*}$, using MH, with the proposal distribution
@@
 \end{equation}
-% and choose $z_{k,t,m}^{*}$ as the next sample in the constructed sequence if $\alpha < \text{min}(1, r_{accept})$, where $\alpha$ is a sample from a $\text{Uniform}(0,1)$ distribution.
 
@@
 \begin{split}
-%\pi_{t}(c_{1:N_{1}, 1}, \ldots, c_{1:N_{t}, t}, & \theta_{1:K_{1}, 1}, \ldots, \theta_{1:K_{t}, t}, m_{1:K_{1}, 1}, \ldots, m_{1:K_{t}, t}) = \\
-%\pi_{t-1}(c_{1:N_{1}, 1}, \ldots, c_{1:N_{t-1}, t-1}, & \theta_{1:K_{1}, 1}, \ldots, \theta_{1:K_{t-1}, t-1}, m_{1:K_{1}, 1}, \ldots, m_{1:K_{t-1}, t-1}) \\
 \pi_{t}(\bold{c}_{1:t}, \Theta_{1:t}, \bold{m}_{1:t}) & =  \pi_{t-1}(\bold{c}_{1:t-1}, \Theta_{1:t-1}, \bold{m}_{1:t-1})\\
@@
 where $\bold{c}_{t} = c_{1:N_{t},t}$, $\Theta_{t} = \theta_{1:K_{t},t}$, and $\bold{m}_{t} = m_{1:K_{t},t} $, and D is given by \eqref{D_distro}.
-%The distributions used in the SMC algorithm are described in detail in the following sections.
 
@@
 \ENDFOR
-\STATE $\tilde{w}_{t}^{(l)} \leftarrow w_{t-1}^{(l)} \times \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})} {P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) } $% \times r $  % \hfill $\triangleright$ eq.~\eqref{smc_weight_eqn}
+\STATE $\tilde{w}_{t}^{(l)} \leftarrow w_{t-1}^{(l)} \times \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})} {P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) } $
 \ENDFOR
@@
 
-% \subsection{Calculation of particle weights $\tilde{w}_{t}^{(l)}$}
-% In Algorithm~\ref{alg:SMC}, the weight $\tilde{w}_{t}^{(l)}$ of each particle $l \in \{ 1. \ldots, L \}$ at time $t$ is given in terms of the particle's weight at time $t-1$ by the equation
-% \begin{equation}
-% \tilde{w}_{t}^{(l)} \leftarrow w_{t-1}^{(l)} \times \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})} {P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) } \times r
-% \end{equation}
-% where $r$ denotes the computed ratio for the particle (defined in Algorithm~\ref{alg:SMC}), and we define
-% \begin{equation}
-% \label{smc_weight_eqn}
-% \begin{split}
-% & \frac{ P(\bold{x}_{1:N_{t},t}, c_{1:N_{t}, t}^{(l)} | \theta_{1:K_{t}^{(l)}}, m_{1:K_{t},t}^{(l)})} {P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t} ) }  \\
-% & = \frac{P(\bold{x}_{1:N_{t},t} | \theta_{1:K_{t}^{(l)}}) P(c_{1:N_{t}, t}^{(l)} | m_{1:K_{t},t}^{(l)}) )}{P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t})}
-% \end{split}
-% \end{equation}
-% where $P(\bold{x}_{1:N_{t},t} | \theta_{1:K_{t}^{(l)}})$ is defined by \eqref{likelihood}, $P(c_{1:N_{t}, t}^{(l)} | m_{1:K_{t},t}^{(l)}))$ is defined by \eqref{C_distro}, and $P(c_{1:N_{t}, t}^{(l)} | m_{t}^{(l)}, \theta_{t-1}^{(l)}, \bold{x}_{1:N_{t},t})$ is defined by \eqref{mcmc_assig_vars}.
-
-%%%%%
-% need to complete this equation
-%%%%%
-
 \subsection{Resampling Particles and Particle Weights}
@@
 In this step, $L$ new particles are sampled from current set of $L$ particles.
-% We will describe a specific, simple method known as Multinomial Resampling. In this method of resampling, we draw
-% \begin{equation}
-% \Upsilon \sim \mathcal{M}n((w_{t}^{(1)}, \ldots, w_{t}^{(L)}))
-% \end{equation}
-% where $\mathcal{M}n$ denotes the multinomial distribution, and $\Upsilon$ represents a vector of counts showing the multiplicity of each of the old particles in the new particle set (i.e.
-$\Upsilon_{i}$ denotes how many times particle $i$ is duplicated in the new particle set).
 Resampling strategies such as those described in \cite{gasthaus_thesis} and \cite{douc2005comparison} can be used.
@@
 
-%The nature of object detection and tracking results may vary depending on the application; in this study, due to the performance metrics we wish to compute (see Section~\ref{sec:performanceevaluationmetrics}), we desire the approximate spatial region for each moving object, for each frame that this object can be seen within the video.
-We choose to make our detection and tracking results depend completely on the inferred parameters of the multivariate normal factor of each cluster.
 Each inferred cluster is taken to be a distinct object, and the sequence of means and covariance matrices for a given cluster are used to determine the position and spatial region, respectively, of a given object over a sequence of time steps.
@@
 We report the maximum a posteriori (MAP) sample as our result.
-%Furthermore, as both inference procedures described in this paper involve sampling, and each sample represents a result, we must determine a way to narrow these samples down to get a single result.
-We have found that calculating a value proportional to joint probability of the model for each sample and choosing the sample that yields the highest value gives reasonable results.
 
@@
 
-%%%%%
-% I still need to include performance metric results for these synthetic datasets...maybe
-%%%%%
 \begin{figure*}[!]
   \centering               
-  \subfloat[]{\includegraphics[width=0.32\textwidth]{../img/synth1_cross.pdf}} \hspace{0.5mm}
-  \subfloat[]{\includegraphics[width=0.32\textwidth]{../img/synth1_bounce.pdf}} \hspace{0.5mm}
-  \subfloat[]{\includegraphics[width=0.32\textwidth]{../img/synth4.pdf}}
+  \subfloat[]{\includegraphics[width=0.32\textwidth]{synth1_cross.pdf}} \hspace{0.5mm}
+  \subfloat[]{\includegraphics[width=0.32\textwidth]{synth1_bounce.pdf}} \hspace{0.5mm}
+  \subfloat[]{\includegraphics[width=0.32\textwidth]{synth4.pdf}}
   \caption{Each plot shows a sample from the posterior distribution of the model for synthetic experiments one (a and b) and two (c), where the vertical axis represents frame number, the horizontal axes represent spatial position, objects are denoted by marker colors and marker types, and the mean and standard deviation are shown.
@@
   \centering
-  \subfloat[]{\label{fig:pets2000_results} \includegraphics[width=0.48\textwidth]{../img/pets2000_front.pdf}} \hspace{1mm}           
-  \subfloat[]{\label{fig:pets2001_results} \includegraphics[width=0.48\textwidth]{../img/pets2001_back.pdf}} \\
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f1_w}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f4_w}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f9_w}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2000_f11_w}} \\           
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f1_w.pdf}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f3_w.pdf}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f5_w.pdf}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2001_f8_w.pdf}} \\
+  \subfloat[]{\label{fig:pets2000_results} \includegraphics[width=0.48\textwidth]{pets2000_front.pdf}} \hspace{1mm}           
+  \subfloat[]{\label{fig:pets2001_results} \includegraphics[width=0.48\textwidth]{pets2001_back.pdf}} \\
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{pets2000_f1_w}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{pets2000_f4_w}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{pets2000_f9_w}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{pets2000_f11_w}} \\           
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{pets2001_f1_w.pdf}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{pets2001_f3_w.pdf}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{pets2001_f5_w.pdf}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{pets2001_f8_w.pdf}} \\
   \caption{Results from the PETS2000 (a) and PETS2001 (b) dataset.
@@
   \centering             
-  \subfloat[]{\includegraphics[width=0.49\textwidth]{../img/pets2000_pm_v2.pdf}} \hspace{1mm}
-  \subfloat[]{\includegraphics[width=0.49\textwidth]{../img/pets2001_pm_v2.pdf}}
+  \subfloat[]{\includegraphics[width=0.49\textwidth]{pets2000_pm_v2.pdf}} \hspace{1mm}
+  \subfloat[]{\includegraphics[width=0.49\textwidth]{pets2001_pm_v2.pdf}}
   \caption{The SFDA (blue solid line) and ATA (red dashed line) vs  confidence values from which an object's oval region is computed for (a) PETS2000 and (b) PETS2001 video datasets.}
@@
   \centering
-  \subfloat[]{\includegraphics[width=0.43\textwidth]{../img/pets2009_1-50_front.pdf}}
+  \subfloat[]{\includegraphics[width=0.43\textwidth]{pets2009_1-50_front.pdf}}
   \hspace{4mm}
-  \subfloat[]{\includegraphics[width=0.53\textwidth]{../img/pets2009full_pm_b_v2.pdf}} \\
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2009_f3_w.pdf}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2009_f6_w.pdf}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2009_f13_w.pdf}} \hspace{1pt}
-  \subfloat[]{\includegraphics[width=0.24\textwidth]{../img/pets2009_f14_w.pdf}}
+  \subfloat[]{\includegraphics[width=0.53\textwidth]{pets2009full_pm_b_v2.pdf}} \\
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{pets2009_f3_w.pdf}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{pets2009_f6_w.pdf}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{pets2009_f13_w.pdf}} \hspace{1pt}
+  \subfloat[]{\includegraphics[width=0.24\textwidth]{pets2009_f14_w.pdf}}
   \caption{ (a) Results for the PETS2009/2010 dataset, showing a sample from the posterior distribution of the state for frames 1-50, where the vertical axis denotes time, the horizontal axes represent spatial position, color represents assignment, and the mean and standard deviation are shown.
@@
   Arsic \cite{arsic2009multi} & $ 0.18 $ & $0.02$ \\
-  % Anonymous & $ 0.52311  $ & $0.246275$ \\ \hline
   \hline
@@
 \end{table}
-%
-% table with more sig figs
-%
-% \begin{table}
-% \begin{tabular}[!] {| l | l | l |}
-%   \hline
-%   \textsc{Method Name} & \textsc{SFDA}  & \textsc{ATA} \\ \hline \hline
-%   Breitenstein \cite{breitenstein2009markovian} & $ 0.567265 $ & $0.30044$ \\ \hline
-%   Yang \cite{yang2009probabilistic} & $ 0.552439 $ & $0.448237$ \\ \hline
-%   Conte \cite{conte2010performance} & $ 0.52966  $ & $0.058003$ \\ \hline
-%   \textbf{GPUDDPM} & $ \bold{0.5079} $ & $\bold{0.2996}$ \\ \hline
-%   Berclaz \cite{berclaz2009multiple} & $ 0.482081 $ & $0.145195$ \\ \hline
-%   Alahi 1 \cite{alahi2009sparsity} & $ 0.430043 $ & $0.041053$ \\ \hline
-%   Alahi 2 \cite{alahi2009sparsity} & $ 0.419141 $ & $0.045609$ \\ \hline
-%   Bolme 1 \cite{bolme2009simple} & $ 0.413812 $ & NA \\ \hline
-%   Ge \cite{ge2009evaluation} & $ 0.383235 $ & $0.040118$ \\ \hline
-%   Bolme 2 \cite{bolme2009simple} & $ 0.338979 $ & NA \\ \hline
-%   Arsic \cite{arsic2009multi} & $ 0.177768 $ & $0.016111$ \\
-%   % Anonymous & $ 0.52311  $ & $0.246275$ \\ \hline
-%   \hline
-% \end{tabular}
-% \caption{SFDA and ATA performance metric results are shown for our method (in bold) and for ten other algorithms on the PETS2009/2010 benchmark dataset.
-Results are listed in descending order of the SFDA value.
-Results are provided by the authors of \cite{ellis_2010}.}
-% \label{benchmark_results_table}
-% \end{table}
 
@@
 Detection and tracking performance was also shown to be fairly robust to minor variations in these parameter values.
-%Figure~\ref{fig:sens_surf} shows a heatmap and contour plot of the sensitivity analysis results.
-From this figure, we can see that the SFDA and ATA achieve their maximal values at different $\alpha$ and $\rho$ parameters, though both achieve a reasonably optimal value at the intermediate parameter values $\alpha = 10$ and $\rho = 0.85$.
-We can also see that detection and tracking performance is fairly robust to minor variations in these parameter values.
-
-% \begin{figure*}[!]
-%   \centering             
-%   \subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2009full_sens_sfda_contour5.pdf}}
-%   \hspace{6mm}
-%   \subfloat[]{\includegraphics[width=0.48\textwidth]{../img/pets2009full_sens_ata_contour5.pdf}}
-%   \caption{Heatmap and contour plot showing the sensitivity of performance metrics SFDA and ATA to variation of parameters $\alpha$ and $\rho$.}
-%   \label{fig:sens_surf}
-% \end{figure*}
 
@@
 
-% More to include: more discussion of parameter choices, thoughts about deletion params vs.
-dataset size (and the tradeoff between tracking through occlusion vs clusters rapidly gaining size), mvn cluster size vs.
-tracking in occlusion vs.
-sampling correct number of objects.
-discuss how we detect movement, which isn’t exactly same as authors in the perf.
-metric paper (ie our tracking is designed to stop if people stop moving).
-objects hard to track through occlusion becuase they are small, similar appearance, and our algorithm doesn’t assume dynamics to keep robust (we could, probably for better accuracy but less generality).
-background (appearance) is like a natural prior
-
 \begin{small}
@@
 \label{sec:modelbackground}
-
-% Dirichlet process mixture (DPM) models (Section~\ref{sec:dpmixture}) fall under the heading of Bayesian nonparametric (BNP) models.
-These models have been widely used in the past decade to perform nonparametric density estimation and cluster analysis.
-Since we have time series data and wish to find spatiotemporal clusters, we are interested in using a closely related class of models known as dependent Dirichlet process mixture (DDPM) models (Section~\ref{sec:ddpmixture}).
-These models are particularly useful for estimating the number of latent classes (clusters) in time dependent data.
-In this work, estimating the number of clusters in video extraction data is equivalent to estimating the number of distinct objects in a video.
-
-% This section provides background information that allows us to introduce DDPM models and describes the motivation behind our use of these models.
-This background is helpful for the definition of the particular DDPM chosen, the Generalized Polya Urn dependent Dirichlet process mixture (GPUDDPM) model, given in Section~\ref{sec:gpuddpm}.
 
@@
 Hence, a time-dependent DDPM is a model which remains a Dirichlet process, marginally, at each time step, yet allows cluster parameters at a given time step to vary from (and remain dependent upon) the parameters in neighboring time steps.
-
-% \subsection{Generalized Polya Urn Dependent Dirichlet Process Mixture Model}
-% \label{sec:gpuddpmixture}
-
-% The time dependent video extraction data contains a collection of clusters, each representing a video object, which might change in number as time progresses (which is to say, clusters may be created, or be ``born'', and may disappear, or ``die'', at some intermediate time step), since objects can enter and exit a scene.
-To handle these challenges, the specific DDPM chosen to model extraction data in this work is known as the Generalized Polya Urn dependent Dirichlet process mixture (GPUDDPM), introduced by \cite{caron_2007} in 2006.
-
-% The GPUDDPM, when applied to data extracted from a video with T frames, can be viewed as a sequence of DPMs (one for each ``time-step'' $t \in \{1, \ldots, T \}$), which are linked together by dependencies between cluster parameters in neighboring time steps.
-More specifically, the parameters at a given time step are distributed as a function of the parameters in the previous time step, and the distribution and number of distinct cluster assignments at a given time step are distributed according to all previous cluster assignments and a deletion procedure, described below.
-
-% A transition kernel $P(\theta_{k,t} | \theta_{k,t-1})$ specifies how mixture component parameters in time step $t$ are dependent upon associated mixture component parameters in time step $t-1$.
-One caveat is that each mixture component must be drawn independently from $\mathbb{G}_{0}$ (the base distribution of the DP, which acts as a prior distribution for the cluster parameters) which we achieve by making $\mathbb{G}_{0}$ the invariant distribution of the transition kernel $P(\theta_{k,t} | \theta_{k,t-1})$ (which, one should note, is a markov chain).
-
-% Recall that each observation $i$ is given a cluster assignments $c_{i}$ (Section~\ref{sec:dpmixture}).
-Cluster size $m_{k}$ denotes the number of observations assigned to a given cluster $k$; in a DPM, cluster size can be found by counting the number of these assigned observations.
-In contrast, to account for varying numbers of clusters at different points in time---and in particular, to allow clusters to diminish in size (i.e.
-to diminish in number of assigned observations) and even die off---the GPUDDPM has a deletion procedure that allows observations to be considered unassigned (or ``removed'') from their assigned clusters at a later time step.
-This procedure is introduced and described in detail in \cite{caron_2007}.
-
-% The deletion procedure operates in the following way: at each time step, all previous assignments (that have not yet been deleted) are independently considered for deletion.
-Specifically, each remaining assignment is removed from its cluster with probability $\rho$.
-It can be shown that performing deletion in this way is equivalent to, for each cluster $k$, drawing $r \sim \text{Binomial}(m_{k,t-1}, \rho)$, and reducing the previous-time cluster size $m_{k,t-1}$ by $r$ to find the next cluster size $m_{k,t}$.
-At each time step, this deletion procedure is carried out on all existing assignments, including those at the current time.
-Hence, the size $m_{k,t}$ of a given cluster $k$ at time $t$ is dependent upon the cluster size at the previous time step, $m_{k,t-1}$, and on the assignments $c_{i}$ for all observations $i \in \{ 1, \ldots, N_{t} \}$ at time $t$.
-We refer to the conditional distribution over $m_{k,t} | m_{k,t-1}, c_{i}, \rho$ as Del, which we define to be
-% \begin{equation}
-% \begin{split}
-% \label{del_step}
-% \text{Del} &(\cdot | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) := m_{k,t-1} \\
-% &- \text{Binomial}( \cdot | {m_{k,t-1}, \hspace{1mm} \rho}) + \sum_{i=1}^{N_{t}} \mathbb{I}(c_{i,t} = k)
-% \end{split}
-% \end{equation}
-% $\forall k \in \{1, \ldots, K_{t} \}$, where $\rho$ is a deletion parameter, $K_{t}$ is the number of clusters at time $t$, and $\mathbb{I}(c_{i,t} = k)$ is an indicator function whose value is 1 if $c_{i,t} = k$ and 0 otherwise.
-This process adheres to what \cite{caron_2007} calls a ``uniform deletion strategy'' over all the observations' assignments (since assignments to each cluster have equal probability of being deleted), though a more complex deletion strategy, dependent upon cluster size, can also be implemented and is described in \cite{caron_2007}.
-
-% We also define a distribution, which we refer to as C, that is useful during inference procedures (Section~\ref{sec:inference}) for inferring assigments of observations. We define C to be
-% \begin{equation}
-% \label{C}
-% \begin{split}
-% \text{C} &(\cdot | m_{1:K_{t},t}, \hspace{1mm} \alpha) := \text{Discrete}(\cdot | P(c_{i,t}=1 | m_{1:K_{t},t}), \\
-% &  \ldots, P(c_{i,t}=K_{t+1}  | m_{1:K_{t},t}))
-% \end{split}
-% \end{equation}
-% where
-% \begin{equation}
-% \begin{split}
-% P (c_{i,t}=k  | m_{1:K_{t},t})
-% = \begin{cases}
-% \frac{m_{k,t}}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{1mm} \text{if} \hspace{1mm} k \in \{ 1, \ldots, K_{t} \} \\
-% \frac{\alpha}{\sum_{k=1}^{K_{t}} m_{k,t} + \alpha}, \hspace{1mm} \text{if} \hspace{1mm} k = K_{t} + 1
-% \end{cases}
-% \end{split}
-% \end{equation}
-% $\forall i \in \{1, \ldots, N_{t} \}$, where there exists $K_{t}$ clusters at time $t$, and we give a newly created cluster the index $K_{t+1}$.
-Distributions \eqref{del_step} and \eqref{C} together comprise what \cite{caron_2007} refers to as the ``Generalized Polya Urn''.
-Using the Del and C distributions given above, we can define the GPUDDPM generatively as, for each time step $t \in \{1, \ldots, T\}$, and each cluster $k \in \{ 1, \ldots, K_{t} \}$ at time $t$,
-% \begin{align}
-% \label{gpuddpm_def}
-% \begin{split}
-% m_{k,t} | m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho  &\sim \text{Del}(m_{k,t-1}, \hspace{1mm} c_{1:N_{t}, t}, \hspace{1mm} \rho) \\
-% \theta_{k, t} | \theta_{k, t-1}   &\sim
-% \begin{cases}
-%   P(\theta_{k, t} | \theta_{k, t-1}) \hspace{1mm} \text{if} \hspace{1mm} k \in \{ 1, \ldots, K_{t} \} \\
-%   \mathbb{G}_{0}   \hspace{2mm} \text{if} \hspace{2mm} k = K_{t+1}
-% \end{cases} \\
-% c_{i,t} | m_{1:K_{t},t}, \hspace{1mm} \alpha  &\sim  \text{C} (m_{1:K_{t},t}, \hspace{1mm} \alpha) \\
-% \bold{x}_{i,t} | c_{i,t}, \theta_{1:K_{t}, t} &\sim \text{F}(\theta_{c_{i,t}, t})
-% \end{split}
-% \end{align}
-% where we specify distributions for F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ in Section~\ref{sec:modelspecification}.
-The graphical model corresponding with this formulation is shown in Figure \ref{fig:gpuddpm_gm_1}.
-% \begin{figure}[h]
-%         \center{\includegraphics[width=80mm]{../img/gpuddp_gm_1.pdf}}
-%         \caption{\label{fig:gpuddpm_gm_1} Graphical Model of the Generalized Polya Urn dependent Dirichlet process mixture.
-Note that all observations at time $t$, $\bold{x}_{1:N,t}$, and their associated assignments $c_{1:N,t}$ are denoted respectively as $\bold{x}_{t}$ and $c_{t}$ (this also holds for time step $t-1$).}
-% \end{figure}
-
-% \subsection{Deletion Variable Formulation of the GPUDDPM}
-
-% To allow for modeling time-dependent data, the GPUDDPM allows clusters to ``lose'' assigned observations in a deletion step, which modifies the values of the cluster sizes at each time.
-Instead of incorporating the cluster size variable $m$ directly in the model (as it was in the definition of the GPUDDPM given by \eqref{gpuddpm_def}), we can formulate an equivalent model which makes use of new set of variables called the deletion variables (denoted as $d$), which represent the times at which assignments are deleted.
-More precisely, we introduce variables $d_{1}, \ldots, d_{N}$ to the model, which denote the times at which the assignments of observations $1, \ldots, N$ are considered removed from their assigned cluster.
-At each time step, cluster sizes can be recontructed from all previous assignment and deletion variables.
-With these deletion variables, we can compute the size of cluster $k$ at time $t$, $m_{k,t}$, by
-% \begin{equation}
-% \label{compute_clust_size}
-% m_{k,t} = \sum_{t' = 1}^{t} \mathbb{I}[(c_{t'}=k) \wedge (t < d_{t'})]
-% \end{equation}
-% where $\mathbb{I}[\cdot]$ is an indicator function that evaluates to 1 if its argument is true, and 0 otherwise.
-Additionally, for an observation $\bold{x}_{i}$ at a given time-step $t$, the deletion time $d_{i}$ can be defined to be $d_{i} = t + l_{i}$, where $l_{i}$ is considered the lifespan of an assignment, is distributed geometrically, and can be expressed as
-% \begin{equation}
-% \label{del_rho_form}
-% l_{i} | \rho  \sim  \rho(1 - \rho)^{l_{i}}
-% \end{equation}
-% This alternate, deletion variable formulation of the GPUDDPM is useful for the MCMC inference algorithm described in Section~\ref{sec:MCMC}.
-We can write the new formulation of the GPUDDPM as, for each time step $t \in \{1, \ldots, T\}$ and clusters $k \in \{ 1, \ldots, K_{t} \} $ at time $t$,
-% \begin{align}
-% \begin{split}
-% d_{i,t} | \rho  &\sim \text{Geo}(\rho) + t + 1 \\
-% \theta_{k, t} | \theta_{k, t-1}  &\sim
-% \begin{cases}
-% P(\theta_{k,t} | \theta_{k,t-1}) \hspace{2mm} \text{if} \hspace{2mm} k \in \{ 1, \ldots, K_{t} \} \\
-% \mathbb{G}_{0} \hspace{2mm} \text{if} \hspace{2mm} k = K_{t} + 1
-% \end{cases} \\
-% c_{i,t} | \bold{c}_{1:t-1}, \bold{d}_{1:t-1}, \alpha  &\sim  \text{C}(\bold{c}_{1:t-1}, \bold{d}_{1:t-1}, \alpha) \\
-% \bold{x}_{i,t} | c_{i,t}, \theta_{c_{i,t},t}  &\sim  \text{F}(\theta_{c_{i,t}, t})
-% \end{split}
-% \end{align}
-% where $\bold{c}_{t} = c_{1,t}, \ldots, c_{N_{t}, t}$ (for $N_{t}$ observations at time $t$), $\bold{d}_{t} = d_{1,t}, \ldots, d_{N_{t}, t}$, and we specify distributions for F, $\mathbb{G}_{0}$, and $P(\theta_{k, t} | \theta_{k, t-1})$ in Section~\ref{sec:modelspecification}.
-This formulation of the GPUDDPM is also used by \cite{gasthaus_thesis} and \cite{caron_2007}.
-The associated graphical model for this formulation is given in Figure~\ref{fig:gpuddpm_gm_2}.
-% \begin{figure}[h]
-%         \center{\includegraphics[width=82mm]{../img/gpuddp_gm_3.pdf}}
-%         \caption{Graphical model of the deletion variable and auxiliary variable transition kernel representation of the GPUDDPM}
-%         \label{fig:gpuddpm_gm_2}
-% \end{figure}
 

### 2012-06-14 newStuff_ltx/bigPictureIdeas.tex
files for new work in newStuff_ltx directory, most-up-to-date paper in submitCorrections_20120422 directory

% header stuff
% ------------

\documentclass{article}
\usepackage[]{natbib}
\usepackage{amsmath, epsfig}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{easybmat}
\usepackage{footmisc}
\usepackage[usenames,dvipsnames]{color}
\usepackage{subfig}
%\usepackage{fullpage} % Include this if you want to cram lots of things on a page 
\renewcommand\algorithmiccomment[1]{// \textit{#1}}
\newcommand{\ignore}[1]{}
\newcommand{\comment}[1]{}
\newcommand{\frank}[1]{\textcolor{red}{\textsf{\emph{\textbf{\textcolor{blue}{#1}}}}}}
\newcommand{\willie}[1]{\textcolor{green}{\textsf{\emph{\textbf{\textcolor{green}{#1}}}}}}
\DeclareMathOperator*{\argmax}{arg\,max}

\begin{document}

% title and author related
% ------------------------

\title{Big Picture Ideas}
\author{Willie}
\maketitle

% introduction
% ------------

\begin{abstract}
This document contains big picture ideas and directions for research. Currently there are three related but distinct directions for projects that I am looking into: (1) A nonparametric Bayesian treatment of learning object types within a Dependent Dirichlet Process (DDP) tracking scheme, (2) a sophisticated appearance model for tracking a specific object type with a DDP tracker (based on the work of Sigal), and (3) ways to represent and infer the structure of abstract objects within a DDP tracking scheme.
\end{abstract}

\section{NPBayes Treatment of Object Types}

Here, I am trying to extend the current ``simple'' Multivariate-Normal-Multinomial object appearance model (MNM model) of the GPUDDPM such that each object has an assignment, or ``type''. Generatively speaking, we start with a mixture model (where each mixture component is an object type), we draw object appearance parameters given a component of this mixture, and we draw image observations (foreground pixels) given these appearance parameters. I.e., in order to incorporate object types, we make the prior distribution over the object appearance parameters a mixture model, where we might seek to infer the parameters and/or number of mixture components. To start off, I am fixing the number of type-clusters and inferring component parameters only (and I'll do inference on a video with, for example, two types of objects to gauge performance). Next, I want to incorporate a NPB prior to estimate the number of types.

We are left with an important modeling consideration: as each object-cluster is actually a sequence of object appearance parameters in the GPUDDPM model, should the parameters of a single object be allowed to be drawn from different ``types'' at different points in time? In the framework of the GPUDDPM, I believe this is necessary, since each $\theta_{k,t}$ is drawn from the base distribution $\mathbb{G}_{0}$; thus, if the base distribution is a mixture, the cluster parameters are able to switch between modes of the mixture at different points in time. This may prove to be ok, as the transition kernel should keep parameters similar between adjacent time steps, and therefore keep an object within a mode of a mixture. An alternate solution could be to have a separate GPUDDPM for each ``type'' (i.e. a mixture of GPUDDPMs)---however, I really don't think inference would be feasible in a model like this.

\willie{NOTE: I think I might've come up with a way to solve this problem in the framework of the GPUDDPM. I have graphical models in notebook to show to Frank. Basically, consider a DPM with a mixture model base distribution (i.e. each cluster parameter now has an assignment variable); then consider a GPUDDPM, where we extend each cluster parameter to a sequence of parameters, but still keep a single assignment variable for each sequence.}

The idea in this line of research is to figure out a way to add object type clusters to the GPUDDPM, using (nearly) the same representation as before. Once this is developed, I'll be able to assign types to objects in the current model (i.e. this will yield object clusters with similar oval / color parameters) and it may be possible to apply this technique to the other developments below.

\section{Sophisticated Appearance Model for Specific Tracker}

Sigal has developed a sophisticated model for estimating human pose (i.e. estimating the position and orientation, in $\mathbb{R}^{3}$, of a collection of nodes, each representing a part of the body). His ``Loose-limbed Human Body Model'' is structured as an undirected graphical model, with kinematic and penetration constraints between the nodes, and specific likelihood distributions for each node defined on foreground pixel and edge image features.

Sigal's model uses a lot of information to make really good estimates about the pose of a single, upright human in a video. The model is able to incorporate multiple camera views, externally trained body-part detectors (e.g. head, hands, feet), separate likelihoods for image features common to each body part, and constraints that are justified under assumptions about the orientation, body structure, and behavior of the typical human walking in a video. Even then, the model sometimes makes mistakes.

I believe that estimating the 3D pose of a deformable figure to a high precision is probably very hard or impossible without many of the pieces of information and dependencies that Sigal is making use of. An initial plan was to infer parameters of a model similar to Sigal's (instead of fully specifying certain parts of the model as Sigal does) for different animals by setting up a general model with more latent parameters (e.g. something like a ``general vertebrate model''); the strategy would've been, given foreground-pixel data from different animals, to infer distributions over and/or contraints between the parameters of the model. I still think this project might give results, but I am not optimistic about our ability to accurately estimate pose parameters that match up with the true pose (mostly because these Sigal-esque models are very high dimensional, and we use only a single camera, no edge image data, noisier foreground-pixel data, often much smaller and lower resolution figures, there is high variability between the poses of different vertebrates, and we don't expect to incorporate the many object-specific assumptions as Sigal is able to do with humans). Relating to this topic, I've been working with ideas on inferring more-general representations of arbitrary objects (instead of the 3D pose of vertebrates), which I write about in the next section.

However, while I am more interested in developing methods to automatically learn object structure from a video, I still think the GPUDDPM could contribute to Sigal's work. He developed a very good appearance distribution for humans, and we have a model for localizing/segmenting and tracking multiple objects simultaneously, which makes use of (i.e. needs) an appearance distribution. I don't think Sigal has developed a way to track multiple humans at once (in fact, all the tracking that he does on single humans is very basic---through, it has also shown to be very effective). To apply Sigal's appearance distribution to the GPUDDPM tracking scheme we need a prior over the model parameters (for a single human) and a transition kernel between these model parameters that has the prior as its stationary distribution).

I define the chief aspects of Sigal's Loose-limbed Human Body Model here:\\ \\
The model is a 10-node undirected graphical model, where each node represents a body part. The $i_{th}$ node has 6 fixed parameters $\Phi_{i} = [l_{i},w_{i}^{p}, w_{i}^d, o_{i}^{p}, o_{i}^{d}, \epsilon_{i}^{d}]$ (length, width at both ends, offset at both ends, and eccentricity) and six parameters to be inferred, $\bold{X}_{i}$ $\in$ $\mathbb{R}^6$.\\\\
Sigal defines $\bold{X}_{i} = [\bold{x}_{i},\bold{q}_{i}]$, where $\bold{x}_{i} = [\bold{x}_{x,i}, \bold{x}_{y,i}, \bold{x}_{z,i}]$, and $\bold{q}$ $=$ SO(3) $=$ $[q_{x,i},q_{y,i},q_{z,i},q_{w,i}]$ $\in \mathbb{R}^{4}$ (the rotation group). All together, $\bold{X}_{i}$ is in 7 continuous dimensions, but lies on a 6 dimensional manifold.\\\\
Each (undirected) edge between parts $i$ and $j$ has a potential function $\psi_{ij}(\bold{X}_{i}, \bold{X}_{j})$, which is composed of kinematic contraints $\psi_{ij}^{K}(\bold{X}_{i}, \bold{X}_{j})$, and interpenetration constraints $\psi_{ij}^{P}(\bold{X}_{i}, \bold{X}_{j})$.\\\\
The foreground pixel likelihood is defined for each part configuration $\bold{X}_{i}$ to be a function of pixels in an image partitioned into three subsets: those representing the given body part configuration, those correlated with the body part (e.g. pixels representing the feet if the body part is a leg), and pixels that have nothing to do with the given body part. \\\\
Overall, this is a 60 parameter model. I need to define reasonable priors over these parameters and a transition kernel whose stationary distribution is equal to this prior. Then we could try and see if we get any accuracy in pose estimation with our data.

\section{Representing and Inferring Object Structure}

Depending on the success we have inferring the three-dimensional pose of objects, we could try to infer other structural representations that might provide a less-realistic 3D depiction, but be more applicable to wider (and more arbitrary) classes of objects. For example, suppose we chose to model (the foreground pixels of) a single object in a video with a DDP mixture of bivariate Gaussians. This model would incorporate dependencies between adjacent frames for each body part, and would automatically infer a wide scale of structures and complexities---it would, in fact, be a nonparametric representation of object structure.

Didn't finish writing. We can talk more about this in person.

\begin{small}
\bibliographystyle{plainnat}
\bibliography{../ltx/paper_refs} 
\end{small}

\end{document}

### 2012-10-09 newStuff/litReview.tex
reorganized repo, removed superfluous files, cleaned things up a bit.

--- 2012-03-23 newStuff_ltx/litReview.tex
@@
 \begin{abstract}
-This is a review of graphical models for representing humans and other objects.
+This is a review of graphical models for representing humans and other articulated bodies.
 Models of both deformable and non-deformable objects allow us, given image and video data, to infer the pose and/or orientation of objects.
@@
 \section{Data}
-We take the data to be the same as in the previous paper: frame difference (i.e. foreground, i.e. silhouette) pixel positions, and additional local appearance features
+We take our data to be the same as in the previous paper: frame difference (i.e. foreground, i.e. silhouette) pixel positions, and additional local appearance features
 
@@
 
-Sigal states, ``we combine techniques from discriminative learning, graphical models, belief propagation, and particle filtering to achieve reliable multi-component object detection and tracking." [p98]
+Sigal breaks down his framework into five components [p103]
+\begin{enumerate} 
+\item a graphical model
+\item an inference algorithm that provides the ability to infer a state of each node in the graph
+\item a local evidence distribution (or image likelihood)
+\item a proposal process for some or all nodes in a graphical model
+\item a set of spatial and/or temporal constraints corresponding to the edges in a graph
+\end{enumerate}
+Each object is represented by a undirected graphical model.
+The model for each object contains a set of four loosely connected regions; for example, a human is represented by head, left arm, right arm, and leg regions.
+Spatial constraints govern the relative positions of the regions for a given object type, and are modeled with a mixture of Gaussians.
+Sigal defines a likelihood, which models the probability of image data given one of the regions.
+Features for the likelihood for each region for a given object are found in a supervised manner, by using a boosted classifier.
 
